### 2021-07-08 latex_files/ch1/1_1_motivation.tex
added all content from thesis to the paper

\subsubsection*{\textit{Explanation of simulation-based models}}

A simulator-based model is a parameterised stochastic data generating
mechanism \cite{Gutmann2016}. The key characteristic of these
models is that although we can sample (simulate) data points, we
cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
given a set of parameters $\thetab$, produces samples $\yb$ in a
stochastic manner i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models are particularly captivating due to the
modelling freedom they provide; any physical process that can be
conceptualised as a computer program of finite (deterministic or
stochastic) steps can be modelled as a simulator-based model without
any compromise. The modelling freedom includes any amount of hidden
(unobserved) internal variables or logic-based decisions. As always,
this degree of freedom comes at a cost; performing the inference is
particularly demanding from both computational and mathematical
perspective. Unfortunately, the algorithms deployed so far permit the
inference only at low-dimensionality parametric spaces, i.e.\
$\thetab \in \mathbb{R}^D$ where $D$ is small.

\subsubsection*{\textit{Example}}

For underlying the importance of simulator-based models, let us use
the tuberculosis disease spread example as described in
\cite{Tanaka2006}. An overview of the disease spread model is
presented at figure~\ref{fig:tuberculosis_model}. At each stage one of
the following \textit{unobserved} events may happen; (a) the
transmission of a specific haplotype to a new host, (b) the mutation
of an existent haplotype or (c) the exclusion of an infectious host
(recovers/dies) from the population. The random process, which stops
when $m$ infectious hosts are reached\footnote{We suppose that the
  unaffected population is infinite, so a new host can always be added
  until we reach $m$ simultaneous hosts.}, can be parameterised by the
transmission rate $\alpha$, the mutation rate $\tau$ and the exclusion
rate $\delta$, creating a $3D$-parametric space
$\thetab = (\alpha, \tau, \delta)$. The outcome of the process is a
variable-size tuple $\yb_{\thetab}$, containing the population
contaminated by each different haplotype, as described in
figure~\ref{fig:tuberculosis_model}. Let's say that the disease has
been spread in a real population and when $m$ hosts were contaminated
simultaneously, the vector with the infectious populations has been
measured to be $\data$. We would like to discover the parameters
$\thetab = (\alpha, \tau, \delta)$ that generated the spreading
process and led to the specific outcome $\data$. Computing
$p(\yb=\data|\thetab)$ requires tracking all tree-paths that could
generate the specific tuple; such exhaustive enumeration becomes
intractable when $m$ grows larger, as in real-case scenarios. In
figure~\ref{fig:tuberculosis_model} we can observe that a transmission
followed by a recovery/death creates a loop, reinstating the process
to the previous step, which also complicates the exhaustive
enumeration. Hence, representing the process with a simulator-based
model\footnote{which is simple and efficient} and performing
likelihood-free inference is the recommended solution.

\begin{figure}[!ht]
    \begin{center}
      \includegraphics[width=0.75\textwidth]{./latex_files/images/chapter1/tuber_model_1.png}
    \end{center}
    \caption[The tuberculosis spreading process.]{Depiction of a random example from the tuberculosis
      spreading process. The image has been taken from
      \cite{Lintusaari2017}.}
    \label{fig:tuberculosis_model}
\end{figure}

\subsubsection*{\textit{Goal of Simulation-Based Models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one(many) parameter configuration(s) $\thetab^*$ that
\textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. In our case, following the approach of Bayesian ML, we treat
the parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

\subsubsection*{\textit{Robust Optimisation Monte Carlo (ROMC) method}}

The ROMC method~\cite{Ikonomov2019} is very a recent
likelihood-free approach. Its fundamental idea is the transformation
of the stochastic data generation process $M_r(\thetab)$ to a
deterministic mapping $g_i(\thetab)$, by sampling the variables that
produce the randomness $\vb_i \sim p(\V)$. Formally, in every
stochastic process the randomness is influenced by a vector of random
variables $\V$, whose state is unknown prior to the execution of the
simulation; sampling the state makes the procedure deterministic,
namely $g_i(\thetab) = M_d(\thetab, \V=\vb_i)$. This approach
initially introduced at \cite{Meeds2015} with the title
\textit{Optimisation Monte Carlo (OMC)}. The ROMC extended this
approach by resolving a fundamental failure-mode of OMC\@. The ROMC
describes a methodology for approximating the posterior through a
series of steps, without explicitly enforcing which algorithms must be
utilised for each step\footnote{The implementation chooses a specific
  algorithm for each task, but this choice has just a demonstrative
  value; any appropriate algorithm can be used instead.}; in this
sense, it can be perceived as a meta-algorithm.

\subsubsection*{\textit{Implementation}}

The most important contribution of this work is the implementation of
the ROMC method in the Python package Engine for Likelihood-Free
Inference (ELFI) \cite{1708.00707}. Since the method has been
published quite recently, it has not been implemented by now in any ML
software. This work attempts to provide the research community with a
robust and extensible implementation for further experimentation.

### 2021-07-08 latex_files/ch1/1_4_online_notebooks.tex
added all content from thesis to the paper

Since the dissertation is mainly focused on the implementation of the
ROMC, we have created some interactive notebooks\footnote{We have used
  the jupyter notebooks \autocite{Kluyver:2016aa}} supporting the main
document. The reader is advised to exploit the notebooks in order to
(a) review the code used for performing the inference (b) understand
the functionalities of our implementation in a practical sense (c)
check the validity of our claims and (c) interactively execute
experiments.

The notebooks are provided in two formats:\footnote{The
  notebooks in the two repositories are identical.}

\begin{itemize}
\item The first format is located in
  \href{https://github.com/givasile/edinburgh-thesis/tree/master/notebook_examples}{this}
  Github repository. In order for the reader to experiment with the
  examples, they should clone the repository into their pc, execute
  the installation instructions and run the notebooks using the
  Jupyter package.
\item The second format is the google colab notebook
  \autocite{Bisong2019}. Following the links below, the reader can
  view the notebook and make an online clone of it. Hence, without any
  installation overhead, they may interactivelly explore all the
  provided functionalities.
\end{itemize}

The notebooks provide a practical overview of the implemented
functionalities and it are easy to use; in particular, the google
colab version is entirely plug-and-play. Therefore, we encourage the
reader of the dissertation to use them as supporting material.  The
following list contains the implemented examples along with their
links:

\begin{itemize}
\item Simple 1D example
  \begin{itemize}
  \item \href{https://colab.research.google.com/drive/1lGRp0XrNfZ64NN0ASB_tYEKowXwlveDC}{Google colab}
  \item \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/example_1D.ipynb}{Github repository}
  \end{itemize}

\item Simple 2D example
  \begin{itemize}
  \item
    \href{https://colab.research.google.com/drive/1Fof_WmCi1YizzSI_63aEsbLXsno5gSZ3}{Google colab}
  \item \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/example_2D.ipynb}{Github repository}
  \end{itemize}

\item Moving Average example  
  \begin{itemize}
  \item
    \href{https://colab.research.google.com/drive/1nkdACQ370SSc0KB1bHv4sBRaxMlMqoNH}{Google colab}
  \item
    \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/example_MA2.ipynb}{Github repository}
  \end{itemize}

\item Extensibility example
  \begin{itemize}
  \item
    \href{https://colab.research.google.com/drive/1RzB-V1QueP1y1nyzv_VOqR1nVz3DUH3v}{Google colab}
  \item
    \href{https://github.com/givasile/edinburgh-thesis/blob/master/notebook_examples/extending_romc.ipynb}{Github repository}
  \end{itemize}

\end{itemize}

### 2021-07-08 latex_files/ch2/2_1_simulator_based_models.tex
added all content from thesis to the paper

As already stated at Chapter~\ref{sec:introduction}, in
simulator-based models we cannot evaluate the posterior
$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood and forms the basis of all likelihood-free inference
approaches,

\begin{equation} \label{eq:likelihood}
  L(\thetab) =
  \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in B_{d,\epsilon}(\data))
\end{equation}
%
where $c_\epsilon$ is a proportionality factor dependent on
$\epsilon$, needed when
$\Pr(M_r(\thetab) \in B_{d, \epsilon}(\data)) \rightarrow 0$, as
$\epsilon \rightarrow 0$. Equation~\ref{eq:likelihood} describes that
the likelihood of a specific parameter configuration $\thetab$ is
proportional to the probability that the simulator will produce
outputs equal to the observations, using this configuration.

\subsubsection{Approximate Bayesian Computation (ABC) Rejection
  Sampling}

ABC rejection sampling is a modified version of the traditional
rejection sampling method, for cases when the evaluation of the
likelihood is intractable. In the typical rejection sampling, a sample
obtained from the prior $\thetab \sim p(\thetab)$ gets accepted
with probability $L(\thetab)/ \text{max}_{\thetab}
L(\thetab)$. Though we cannot use this approach out-of-the-box
(evaluating $L(\thetab)$ is impossible in our case), we can
modify the method incorporating the simulator.

In the discrete case scenario where $\Y_{\thetab}$ can take a finite
set of values, the likelihood becomes
$L(\thetab) = \Pr(\Y_{\thetab} = \data)$ and the posterior
$p(\thetab|\data) \propto \Pr(\Y_{\thetab}=\data)p(\thetab)$; hence, we can
sample from the prior $\thetab_i \sim p(\thetab)$, run the simulator
$\yb_i = M_r(\thetab_i)$ and accept $\thetab_i$ only if
$\yb_i = \data$.

The method above becomes less useful as the finite set of
$\Y_{\thetab}$ values grows larger, since the probability of
accepting a sample becomes smaller. In the limit
where the set becomes infinite (i.e.\ continuous case) the probability
becomes zero. In order for the method to work in this set-up, a
relaxation is introduced; we relax the acceptance criterion by letting
$\yb_{i}$ lie in a larger set of points i.e.\
$\yb_{i} \in \region(\data), \epsilon > 0$. The region can be
defined as $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$
where $d(\cdot, \cdot)$ can represent any valid distance. With this
modification, the maintained samples follow the approximate posterior,

\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto Pr(\yb \in
  \region(\data)) p(\thetab)
\end{equation}

\noindent
This method is called Rejection ABC.

\subsubsection{Summary Statistics}

When $\yb \in \mathbb{R}^D$ lies in a high-dimensional space, generating
samples inside $\region (\data)$ becomes rare even when $\epsilon$ is
relatively large; this is the curse of dimensionality. As a
representative example lets make the following hypothesis;

\begin{itemize}
\item $d$ is set to be the Euclidean distance, hence
  $\region(\data) := \{ \yb: ||\yb - \data||_2^2 < \epsilon^2 \}$ is a
  hyper-sphere with radius $\epsilon$ and volume $V_{hypersphere} = \frac{\pi^{D/2}}{\Gamma(D/2 + 1)} \epsilon^D$
\item the prior $p(\thetab)$ is a uniform distribution in a hyper-cube with side of
  length $2\epsilon$ and volume $V_{hypercube} = (2\epsilon)^D$
\item the generative model is the identity function $\yb=f(\thetab)= \thetab $
\end{itemize}

\noindent
The probability of drawing a sample inside the hypersphere equals the
fraction of the volume of the hypersphere inscribed in the hypercube:

\begin{equation}
  Pr(\yb \in \region (\data))
  = Pr(\thetab \in \region (\data))
  = \frac{V_{hypersphere}}{V_{hypercube}}
  = \frac{\pi^{D/2}}{2^D\Gamma(D/2 + 1)} \rightarrow 0, \quad \text{as} \quad D \rightarrow \infty
\end{equation}

\noindent
We observe that the probability tends to $0$, independently of
$\epsilon$; enlarging $\epsilon$ will not increase the acceptance
rate. Intuitively, we can think that in high-dimensional spaces the
volume of the hypercube concentrates at its corners. This generates
the need for a mapping
$T: \mathbb{R}^{D_1} \rightarrow \mathbb{R}^{D_2}$ where $D_1 > D_2$,
for squeezing the dimensionality of the output. This
dimensionality-reduction step that redefines the area as
$\region(\data) := \{\yb: d(T(\yb), T(\data)) \leq \epsilon \}$ is
called \textit{summary statistic} extraction, since the distance is
not measured on the actual outputs, but on a summarisation (i.e.\
lower-dimension representation) of them.

\subsubsection{Approximations introduced so far}

So far, we have introduced some approximations for inferring the
posterior as
$p_{d,\epsilon}(\thetab|\data) \propto Pr(\Y_{\thetab} \in
\region(\data))p(\thetab)$ where
$\region(\data) := \{\yb: d(T(\yb), T(\data)) < \epsilon \}$. These
approximations introduce two different types of errors:

\begin{itemize}
\item $\epsilon$ is chosen to be \textit{big enough}, so that enough
  samples are accepted. This modification leads to the approximate
  posterior introduced in \eqref{eq:approx_posterior}
\item $T$ introduces some loss of information, making possible a $\yb$
  far away from $\data$ i.e.\ $\yb: d(\yb,\data)>\epsilon$, to enter
  the acceptance region after the dimensionality reduction
  $d(T(\yb), T(\data)) \leq \epsilon$
\end{itemize}

\noindent
In the following sections, we will not use the summary statistics in
our expressions for the notation not to clutter. One could understand
it as absorbing the mapping $T(\cdot)$ inside the simulator. In any
case, all the propositions that will be expressed in the following
sections are valid with the use of summary statistics.

\subsubsection{Optimisation Monte Carlo (OMC)}

Before we define the likelihood approximation as introduced in the
OMC, approach lets define the indicator function based on
$\region(\yb)$. The indicator function $\indicator{\region(\yb)}(\xb)$
returns 1 if $\xb \in \region(\yb)$ and 0 otherwise. If
$d(\cdot,\cdot)$ is a formal distance, due to symmetry
$\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so
the expressions can be used interchangeably.

\begin{gather} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \xb \in \region(\yb) \\
      0 & \mbox{otherwise} 
    \end{array} \right. \end{gather}

\noindent
Based on equation~\eqref{eq:approx_posterior} and the indicator
function as defined above~\eqref{eq:indicator}, we can approximate the
likelihood as:

\begin{gather} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) =
  \int_{\yb \in B_\epsilon(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb\\
  \approx \frac{1}{N} \sum_i^N \indicator{\region(\data)}(\yb_i),\text{ where }
  \yb_i \sim M_r(\thetab) \label{eq:init_view}\\
  \approx \frac{1}{N} \sum_i^N \indicator{\region (\data)} (\yb_i)
  \text{ where } \yb_i = M_d(\thetab, \vb_i), \vb_i \sim p(\vb) \label{eq:alt_view}
\end{gather}
%
This approach is quite intuitive; approximating the likelihood of a
specific $\thetab$ requires sampling from the data generator and count
the fraction of samples that lie inside the area around the
observations. Nevertheless, by using the approximation of equation
\eqref{eq:init_view} we need to draw $N$ new samples for each distinct
evaluation of $L_{d,\epsilon}(\thetab)$; this makes this approach
quite inconvenient from a computational point-of-view. For this
reason, we choose to approximate the integral as in equation
\eqref{eq:alt_view}; the nuisance variables are sampled once
$\vb_i \sim p(\vb)$ and we count the fraction of samples that lie
inside the area using the deterministic simulators
$M_d(\thetab, \vb_i) \: \forall i$. Hence, the evaluation
$L_{d,\epsilon}(\thetab)$ for each different $\thetab$ does not imply
drawing new samples all over again. Based on this approach, the
unnormalised approximate posterior can be defined as:

\begin{equation} \label{eq:aprox_posterior}
  p_{d,\epsilon}(\thetab|\data)
  \propto p(\thetab) \sum_i^N \indicator{ \region(\data)} (\yb_i)
\end{equation}

\subsubsection*{Further approximations for sampling and computing expectations}

The posterior approximation in \eqref{eq:approx_posterior} does not
provide any obvious way for drawing samples. In fact, the set
$\mathcal{S}_i = \{ \thetab: M_d(\thetab, \vb_i) \in \region(\data) \}$ can
represent any arbitrary shape in the D-dimensional Euclidean space; it
can be non-convex, can contain disjoint sets of $\thetab$ etc. We need
some further simplification of the posterior for being able to draw
samples from it.

As a side-note, weighted sampling could be performed in a
straightforward fashion with importance sampling. Using the prior as
the proposal distribution $\thetab_i \sim p(\thetab)$ and we can
compute the weight as
$w_i = \frac{L_{d,\epsilon}(\thetab_i)}{p(\thetab_i)}$, where
$L_{d,\epsilon}(\thetab_i)$ is computed with the expression
\eqref{eq:approx_likelihood}. This approach has the same drawbacks as
ABC rejection sampling; when the prior is wide or the dimensionality
$D$ is high, drawing a sample with non-zero weight is rare, leading to
either poor Effective Sample Size (ESS) or huge execution time.

The OMC proposes a quite drastic simplification of the posterior; it
squeezes all regions $\mathcal{S}_i$ into a single point
$\thetab_i^* \in \mathcal{S}_i$ attaching a weight $w_i$ proportional
to the volume of $\mathcal{S}_i$. For obtaining a
$\thetab_i^* \in \mathcal{S}_i$, a gradient based optimiser is used
for minimising $g_i(\thetab) = d(\data, f_i(\thetab))$ and the
estimation of the volume of $\mathcal{S}_i$ is done using the Hessian
approximation $\hess_i \approx \jac_i^{*T}\jac_i^*$, where $\jac_i^*$
is the Jacobian matrix of $g_i(\thetab)$ at $\thetab_i^*$. Hence,

\begin{gather} \label{eq:OMC_posterior}
    p(\thetab|\data) \propto p(\thetab) \sum_i^N w_i \delta(\thetab - \thetab_i^*)\\
  \thetab_i^* = \text{argmin}_{\thetab} \:g_i(\thetab) \\
  w_i \propto \frac{1}{\sqrt{det( \jac_i^{*T}\jac_i^*)}}
\end{gather}

The distribution \eqref{eq:OMC_posterior} provides weighted samples
automatically and an expectation can be computed easily with the
following equation,

\begin{equation}
  \label{eq:OMC_expectation}
  E_{p(\thetab|\data)}[h(\thetab)] = \frac{\sum_i^N w_i p(\thetab_i^*)h(\thetab_i^*)}{\sum_i^N w_i p(\thetab_i^*)}
\end{equation}

### 2021-07-08 latex_files/ch2/2_2_ROMC.tex
added all content from thesis to the paper

The simplifications introduced by OMC, although quite useful from a
computational point-of-view, they suffer from some significant failure modes:

\begin{itemize}
\item The whole acceptable region $\mathcal{S}_i$, for each nuisance variable,
  shrinks to a single point $\thetab_i^*$; this simplification may add
  significant error when then the area $\mathcal{S}_i$ is relatively big.
\item The weight $w_i$ is computed based only at the curvature at the
  point $\thetab_i^*$. This approach is error prone at many cases
  e.g.\ when $g_i$ is almost flat at $\thetab_i^*$, leading to a
  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0 \Rightarrow w_i
  \rightarrow \infty$, thus dominating the posterior.
\item There is no way to solve the optimisation problem
  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
  is not differentiable.
\end{itemize}

\subsubsection{Sampling and computing expectation in ROMC}

The ROMC approach resolves the aforementioned issues. Instead of
collapsing the acceptance regions $\mathcal{S}_i$ into single points,
it tries to approximate them with a bounding box.\footnote{The
  description on how to estimate the bounding box is provided in the
  following chapters.}. A uniform distribution is then defined on the
bounding box area, used as the proposal distribution for importance
sampling. If we define as $q_i$, the uniform distribution defined on
the $i-th$ bounding box, weighted sampling is performed as:

\begin{gather}
  \label{eq:sampling}
  \thetab_{ij} \sim q_i \\
  w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
\end{gather}

\noindent
Having defined the procedure for obtaining weighted samples, any
expectation $E_{p(\thetab|\data)}[h(\thetab)]$, can be approximated
as,

\begin{equation} \label{eq:expectation}
  E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}

\subsubsection{Construction of the proposal region}

In this section we will describe mathematically the steps needed for
computing the proposal distributions $q_i$. There will be also
presented a Bayesian optimisation alternative when gradients are not
available.

\subsubsection*{Define and solve deterministic optimisation problems}

For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$ a
deterministic function is defined as
$f_i(\thetab) = M_d(\thetab,\vb_i)$. For constructing the proposal
region, we search for a point
$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$; this point can be
obtained by solving the the following optimisation problem:

\begin{subequations}
\begin{alignat}{2}
&\!\min_{\thetab}        &\qquad& g_i(\thetab) = d(\data,  f_i(\thetab))\label{eq:optProb}\\
&\text{subject to} &      & g_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
We maintain a list of the solutions $\thetab_i^*$ of the optimisation
problems. If for a specific set of nuisance variables $\vb_i$, there
is no feasible solution we add nothing to the list. The optimisation
problem can be treated as unconstrained, accepting the optimal point
$\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$ only if
$g_i(\thetab_i^*) < \epsilon$.

\subsubsection*{Gradient-based approach}
\label{subsubsec:GB_approach}

The nature of the generative model $M_r(\thetab)$, specifies the
properties of the objective function $g_i$. If $g_i$ is continuous
with smooth gradients $\nabla_{\thetab} g_i$ any gradient-based
iterative algorithm can be used for solving~\ref{eq:optProb}. The
gradients $\nabla_{\thetab} g_i$ can be either provided in closed form
or approximated by finite differences.

\subsubsection*{Bayesian optimisation approach}
\label{subsubsec:GP_approach}

In cases where the gradients are not available, the Bayesian
optimisation scheme provides an alternative
choice~\cite{Shahriari2016}. With this approach, apart from
obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
the distance $g_i$ is fitted; this approximate model can be used in
the following steps for making the method more
efficient. Specifically, in the construction of the proposal region
and in
equations~\eqref{eq:approx_posterior},~\eqref{eq:sampling},~\eqref{eq:expectation}
it could replace $g_i$ in the evaluation of the indicator
function, providing a major speed-up.

\subsubsection*{Construction of the proposal area $q_i$}

After obtaining a $\thetab_i^*$ such that
$g_i(\thetab_i^*) < \epsilon$, we need to construct a bounding box
around it. The bounding box $\mathcal{\hat{S}}_i$ must contain the
acceptance region around $\thetab_i^*$, i.e.\
$\{ \thetab : g_i(\thetab) < \epsilon$,
$d(\thetab, \thetab_i^*) < M \}$. The second condition
$d(\thetab, \thetab_i^*) < M$ is meant to describe that if
$\mathcal{S}_i := \{ \thetab : g_i(\thetab) < \epsilon \} $ contains a
number of disjoint sets of $\thetab$ that respect
$g_i(\thetab) < \epsilon$, we want our bounding box to fit only the
one that contains $\thetab_i^*$. We seek for a bounding box that is as
tight as possible to the local acceptance region (enlarging the
bounding box without a reason decreases the acceptance rate) but large
enough for not discarding accepted areas.

In contrast with the OMC approach, we construct the bounding box by
obtaining search directions and querying the indicator function as we move on them. The
search directions $\mathbf{v}_d$ are computed as the eigenvectors of
the curvature at $\thetab_i^*$ and a line-search method is used to
obtain the limit point where
$g_i(\thetab_i^* + \kappa \vb_d) \geq
\epsilon$\footnote{$-\kappa$ is used as well for the opposite direction along the search line}. The Algorithm~\ref{alg:region_construction} describes the method in-depth. After the limits are obtained along all
search directions, we define bounding box and the uniform distribution $q_i$. This is the proposal distribution used for the importance
sampling as explained in \eqref{eq:sampling}.

\subsubsection*{Fitting a local surrogate model $\hat{g}_i$}

After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
are no longer interested in the surface outside the box. In the future
steps (e.g. sampling, evaluating the posterior) we will only evaluate
$g_i$ inside the corresponding bounding box. Hence, we could fit a
local surrogate model $\hat{g}_i$ for representing the local area
around $\theta_i^*$. Doing so, in the future steps we can exploit
$\hat{g}_i$ for evaluating the indicator function instead of running
the whole deterministic simulator.

Any ML regression model may be chosen as local surrogates. The choice
should consider the properties of the local region (i.e. size,
smoothness). The ROMC proposes fitting a simple quadratic model. The
training set $X: \mathbb{R}^{N \times D}$ is created by sampling $N$
points from the bounding box and the labels $Y: \mathbb{R}^{N}$ are
the computed by evaluating $g_i$. The quadratic model is fitted on the
data points, for minimising the square error.

This additional step places an
additional step in the training part, increasing the computational
demands, but promises a major speed at the inference phase (sampling,
posterior evaluation). It is frequent in ML problems, to be quite
generous with the execution time at the training phase, but quite
eager at the inference phase. Fitting a local surrogate model aligns
with this requirement.

### 2021-07-08 latex_files/ch2/2_4_ELFI.tex
added all content from thesis to the paper

The Engine for Likelihood-Free Inference (ELFI) \cite{1708.00707}
is a Python package dedicated to Likelihood-Free Inference (LFI). ELFI
models in a convenient manner all the fundamental components of a
probabilistic model such as priors, simulators, summaries and
distances. Furthermore, ELFI already supports some recently proposed
likelihood-free inference methods.

\subsubsection{Modelling}
\label{sec:modelling}

ELFI models the probabilistic model as a Directed Acyclic Graph (DAG);
it implements this functionality based on the package
\pinline{NetworkX}, which is designed for creating general purpose
graphs. Although not restricted to that, in most cases the structure
of a likelihood-free model follows the pattern presented in
figure~\ref{fig:elfi}; there are edges that connect the prior
distributions to the simulator, the simulator is connected to the
summary statistics that consequently are connected to the
distance. The distance is the output node. Samples can be obtained
from all nodes through sequential sampling. The nodes that are defined
as \pinline{elfi.Prior}\footnote{The \pinline{elfi.Prior} functionality
  is a wrapper around the \pinline{scipy.stats} package.} are automatically
considered as the parameters of interest and are the only nodes that,
apart from sampling, should also provide PDF evaluation. The function
passed as argument in the \pinline{elfi.Summary} node can be any valid
Python function. Finally, the observations should be passed in the
appropriate node through the argument \pinline{observed}.

\begin{figure}[!ht]
    \begin{center}
      \includegraphics[width=0.8\textwidth]{./latex_files/images/chapter2/elfi.png}
    \end{center}
    \caption[Baseline example for creating an \pinline{ELFI} model]{Baseline example for creating an \pinline{ELFI} model. Image taken from \cite{1708.00707}}
    \label{fig:elfi}
\end{figure}

\subsubsection{Inference Methods}
\label{sec:inference-methods}

The inference Methods implemented at the ELFI follow some common
guidelines;

\begin{itemize}
\item the initial argument should is the output node of the model. It
  is followed by the rest hyper-parameters of the method.
\item each inference method provides a central sampling
  functionality. In most cases it is named
  \pinline{<method_name>.sample()}.
\end{itemize}

The collection of likelihood-free inference methods implemented so far
contain the \textit{ABC Rejection Sampler}, the \textit{Sequential
  Monte Carlo ABC Sampler} and the \textit{Bayesian Optimisation for
  Likelihood-Free Inference (BOLFI)}. The latter has methodological
similarities to the ROMC method that we implement in the current work.

### 2021-07-08 latex_files/ch3/3_1_1_general_design.tex
added all content from thesis to the paper

In figure \ref{fig:romc_overview} we present an overview of our
implementation; one may interpret figure~\ref{fig:romc_overview} as a
depiction of the main class of our implementation, called
\pinline{ROMC}, while the entities inside the green and blue
ellipses are the main functions of the class. Following Python's
naming principles, the methods starting with an underscore (green
ellipses) represent internal (private) functions and are not meant to
be used by a user, whereas the rest of the methods (blue ellipses) are
the functionalities the user interacts with. As mentioned before, the
implementation favours extensibility; the building blocks that compose
the method have been designed in a modular fashion so that a
practitioner may replace them without the method to collapse.

Figure \ref{fig:romc_overview} groups the ROMC implementation into the
training, the inference and the evaluation part. The training part includes all the
steps until the computation of the proposal regions; sampling the
nuisance variables, defining the optimisation problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalised posterior (and
the normalised one, in low-dimensional cases), sampling and computing
an expectation. Moreover, the ROMC implementation provides some
utilities for inspecting the training process, such as plotting the
histogram of the distances
$d^*_i = g_i(\theta_i^*), \: \forall i \in \{1, \ldots, n_1 \}$ after
solving the optimisation problems and visualising the constructed
bounding box\footnote{if the parametric space is up to $2D$}. Finally,
two functionalities for evaluating the inference are implemented; (a)
computing the Effective Sample Size (ESS) of the weighted samples and
(b) measuring the divergence between the approximate posterior the
ground-truth, if the latter is available.\footnote{Normally, the
  ground-truth posterior is not available; However, this functionality
  is useful in cases where the posterior can be computed numerically
  or with an alternative method (i.e.\ ABC Rejection Sampling) and we
  would like to measure the discrepancy between the two
  approximations.}

\subsubsection*{Parallelising the processes}

As stated above, the most critical advantage of the ROMC method is that it can be fully parallelised. In our implementation, we have exploited this fact by implementing a parallel version in the following tasks; (a) solving the optimisation problems, (b) constructing bounding box regions, (c) sampling and (d) evaluating the posterior. Parallelism has been achieved using the package \pinline{multiprocessing}. The specific package enables concurrency, using subprocesses instead of threads, hence side-stepping the Global Interpreter (GIL). In our implementation we use the \pinline{Pool} object, which offers a convenient means of parallelising the execution of a function across multiple input values (data parallelism). For activating the parallel version of the algorithms, the user has to just pass the argument \pinline{parallelize=True} at the initialisation of the \pinline{ROMC} method.

\subsubsection*{Simple one-dimensional example}

For illustrating the functionalities we choose as running example the
following model, introduced by~\autocite{Ikonomov2019},

\begin{gather} \label{eq:1D_example} p(\theta) =
\mathcal{U}(\theta;-2.5,2.5)\\ p(y|\theta) = \left\{
    \begin{array}{ll} \theta^4 + u & \mbox{if } \theta \in [-0.5, 0.5]
\\ |\theta| - c + u & \mbox{otherwise}
    \end{array} \right.\\ u \sim \mathcal{N}(0,1)
\end{gather}

\noindent

In the model \eqref{eq:1D_example}, the prior is the uniform
distribution in the range $[-2.5, 2.5]$ and the likelihood a Gaussian
distribution. There is only one observation $y_0 = 0$. The inference
in this particular example can be performed quite easily, without
incorporating a likelihood-free inference approach. We can exploit this
fact for validating the accuracy of our implementation. The
ground-truth posterior, approximated computationally, is shown in
figure \ref{fig:example_gt}.

\begin{figure}[h]
    \begin{center}
\includegraphics[width=0.7\textwidth]{./Thesis/images/chapter3/example_gt.png}
    \end{center}
  \caption[Ground-truth posterior distribution of the simple 1D
example.]{Ground-truth posterior distribution of the simple 1D
example.}
  \label{fig:example_gt}
\end{figure}

\subsubsection*{ELFI code for modelling the example}

In the following code snippet, we code the model at \textit{ELFI} and
we initialise the ROMC inference method. We observe that the
initialisation of the ROMC inference method is quite intuitive; we
just pass the final (distance) node of the simulator as argument, as
in all $\textit{ELFI}$ inference methods. The argument
\pinline{bounds}, although optional, is important for many
functionalities (e.g. approximating the partition function, setting
the bounds of the Bayesian optimisation etc.) so it is recommended to
be passed.

\begin{pythoncode}
  import elfi import scipy.stats as ss
  import numpy as np

  def simulator(t1, batch_size=1,random_state=None):
      if t1 < -0.5:
          y = ss.norm(loc=-t1-c, scale=1).rvs(random_state=random_state)
      elif t1 <= 0.5:
          y = ss.norm(loc=t1**4, scale=1).rvs(random_state=random_state)
      else:
          y = ss.norm(loc=t1-c, scale=1).rvs(random_state=random_state)
      return y

  # observation
  y = 0

  # Elfi graph
  t1 = elfi.Prior('uniform', -2.5, 5)
  sim = elfi.Simulator(simulator, t1, observed=y)
  d = elfi.Distance('euclidean', sim)

  # Initialise the ROMC inference method
  bounds = [(-2.5, 2.5)] # limits of the prior
  parallelize = True # activate parallel execution
  romc = elfi.ROMC(d, bounds=bounds, parallelize=parallelize)
\end{pythoncode}

\begin{figure}[!ht]
    \begin{center}
\includegraphics[width=0.8\textwidth]{./Thesis/graphs/ROMC.png}
    \end{center}
    \caption[Overview of the ROMC implementation.]{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
    \label{fig:romc_overview}
\end{figure}

### 2021-07-08 latex_files/ch3/3_1_2_training.tex
added all content from thesis to the paper

The training part contains the following 6 functionalities:

\begin{enumerate}[label=(\roman*)]
\item \mintinline{python}{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
\item \mintinline{python}{romc.estimate_regions(eps_filter,}

      \mintinline{python}{                      use_surrogate=None, region_args=None,}

      \mintinline{python}{                      fit_models=False, fit_models_args=None,}

      \mintinline{python}{                      eps_region=None, eps_cutoff=None)}

    \item \mintinline{python}{romc.fit_posterior(n1, eps_filter, use_bo=False, optimizer_args=None,}

          \mintinline{python}{                   seed=None, use_surrogate=None, region_args=None,}

          \mintinline{python}{                   fit_models=False, fit_models_args=None,}

          \mintinline{python}{                   eps_region=None, eps_cutoff=None)}

\item \mintinline{python}{romc.distance_hist(savefig=False, **kwargs)}
\item \mintinline{python}{romc.visualize_region(i, savefig=False)}
\item \mintinline{python}{romc.compute_eps(quantile)}
\end{enumerate}

\subsubsection*{Function (i): Define and solve the optimisation problems}

\pinline{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
\vspace{5mm}

\noindent
This routine is responsible for (a) drawing the nuisance variables,
(b) defining the optimisation problems and (c) solving them using either a
gradient-based optimiser or Bayesian optimisation. The aforementioned
tasks are done in a sequential fashion, as show in
figure~\ref{fig:romc_overview}. The definition of the optimisation
problems is performed by drawing $n_1$ integer numbers from a discrete
uniform distribution $u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each
integer $u_i$ is the seed used in ELFI's random simulator. Hence from
an algorithmic point-of-view drawing the state of all random
variables $\vb_i$ as described in the previous chapter, traces back to
just setting the seed that initialises the state of the pseudo-random
generator, before asking a sample from the simulator.

Finally, passing an integer number as the argument \pinline{seed}
absorbs all the randomness of the optimisation part (e.g.\ drawing
initial points for the optimisation procedure), making the whole
process reproducible and deterministic.

Setting the argument \pinline{use_bo=True}, chooses the Bayesian
Optimisation scheme for obtaining $\thetab_i^*$. In this case, apart
from obtaining the optimal points $\thetab_i^*$, we also fit a Gaussian
Process (GP) as surrogate model $\hat{d}_i$. In the following steps,
$\hat{d}_i(\thetab)$ will replace $g_i(\thetab)$ when calling the
indicator function.

\subsubsection*{Function (ii): Construct bounding boxes and fit local surrogate models}

\mintinline{python}{romc.estimate_regions(eps_filter,}

      \mintinline{python}{                      use_surrogate=None, region_args=None,}

      \mintinline{python}{                      fit_models=False, fit_models_args=None,}

      \mintinline{python}{                      eps_region=None, eps_cutoff=None)}
\vspace{5mm}

This routine constructs the bounding boxes around the optimal points
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian $\hess_i = \jac_i^T \jac_i$. The
eigenvectores are computed using the function
\pinline{numpy.linalg.eig()} that calls, under the hood, the
\pinline{_geev LAPACK}. A check is performed so that the matrix
$\hess_i$ is not singular; if this is the case, the eigenvectors are
set to be the vectors of the standard Euclidean basis i.e.\
$\{ \mathbf{e_1} = (1, 0, \ldots), \mathbf{e_2} = (0,1,0,\ldots),
\text{etc} \}$. Afterwards, the limits are obtained by repeteadly querying
the distance function ($g_i(\thetab)$ or $\hat{d}(\thetab)$) along the
search directions. In section \ref{subsec:developers}, we provide some
details regarding the way the bounding box is defined as a class and
sampling is performed on it.

\subsubsection*{Function (iii): Perform all training steps in a single call}

\mintinline{python}{romc.fit_posterior(n1, eps_filter, use_bo=False, optimizer_args=None,}

          \mintinline{python}{                seed=None, use_surrogate=None, region_args=None,}

          \mintinline{python}{                fit_models=False, fit_models_args=None,}

          \mintinline{python}{                eps_region=None, eps_cutoff=None)}

\vspace{5mm}
\noindent

This function merges all steps for constructing the bounding box into
a single command. If the user doesn't want to manually inspect the
histogram of the distances before deciding where to set the threshold
$\epsilon$, he may call \pinline{romc.fit_posterior()} and the whole
training part will be done end-to-end. There are two alternatives for
setting the threshold $\epsilon$; the first is to set to a specific
value blindly and the second is to set at as a specific quantile of
the histogram of distances. In the second scenario the
\pinline{quantile} argument must be set to a floating number in the
range $[0,1]$ and \pinline{eps='auto'}.

\subsubsection*{Function (iv): Plot the histogramm of the optimal points}

\pinline{romc.distance_hist(**kwargs)}
\vspace{5mm}
\noindent

This function can serve as an intermediate step of manual inspection,
for helping the user choose which threshold $\epsilon$ to use. It
plots a histogram of the distances at the optimal point
$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or
$d_i^*$ in case \pinline{use_bo=True}. The function accepts all
keyword arguments and forwards them to the underlying
\pinline{matplotlib.hist()} function; in this way the user may
customise some properties of the histogram, such as the number of bins
or the range of values.

\subsubsection*{Function (v): Plot the acceptance region of the objective functions}

\pinline{romc.visualize_region(i)}
\vspace{5mm}
\noindent

It can be used as an inspection utility for cases where the parametric
space is up to two dimensional. The argument $\pinline{i}$ is the
index of the corresponding optimization problem i.e.\ $i<n_1$.

\subsubsection*{Function (vi): Compute $\epsilon$ automatically based on the distribution of $d^*$}

\pinline{romc.compute_eps(quantile)}
\vspace{5mm}
\noindent

\noindent
This function return the appropriate distance value $d_{i=\kappa}^*$
where $\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the
collection $\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the
number of accepted solutions. It can be used to automate the selection
of the threshold $\epsilon$, e.g.\
\pinline{eps=romc.compute_eps(quantile=0.9)}.

\subsubsection*{Example}

Here we will illustrate the aforementioned functionalities using the
simple 1D example we set up in the previous section. The following
code snippet performs the training part in ELFI.

\begin{pythoncode}
  n1 = 500 # number of optimisation problems
  seed = 21 # seed for solving the optimisation problems
  eps = .75 # threshold for bounding box
  use_bo = False # set to True for switching to Bayesian optimisation

  # Training step-by-step
  romc.solve_problems(n1=n1, seed=seed, use_bo=use_bo)
  romc.theta_hist(bins=100)
  romc.estimate_regions(eps=eps)
  romc.visualize_region(i=1)

  # Equivalent one-line command
  # romc.fit_posterior(n1=n1, eps=eps, use_bo=use_bo, seed=seed)
\end{pythoncode}

As stated before, switching to the Bayesian optimisation scheme needs
nothing more the setting the argument \pinline{use_bo=True}; all the
following command remain unchanged. In figure
\ref{fig:example_training_hist} we illustrate the distribution of the
distances obtained and the acceptance area of the first optimisation
problem. We observe that most optimal points produce almost zero
distance.

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_theta_dist.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_region.png}
    \end{center}
  \caption[Histogram of distances at the 1D example.]{Histogram of distances and visualisation of a specific region.}
  \label{fig:example_training_hist}
\end{figure}

### 2021-07-08 latex_files/ch3/3_1_3_inference.tex
added all content from thesis to the paper

The inference part contains the 4 following functionalities:

\begin{enumerate}[label=(\roman*)]
\item \mintinline{python}{romc.sample(n2, seed=None)}
\item \mintinline{python}{romc.compute_expectation(h)}  
\item \mintinline{python}{romc.eval_unnorm_posterior(theta)}
\item \mintinline{python}{romc.eval_posterior(theta)}
\end{enumerate}

\subsubsection*{Function (i): Perform weighted sampling}

\mintinline{python}{romc.sample(n2)}
\vspace{5mm}

\noindent
This is the basic inference utility of the ROMC implementation; we
draw $n_2$ samples for each bounding box region. This gives a total of
$k \times n_2$, where $k < n_1$ is the number of the optimal points
remained after filtering\footnote{From the $n_1$ optimisation
  problems, only the ones with $g_i(\thetab_*) < \epsilon$ are kept
  for building a bounding box}. The samples are drawn from a uniform
distribution $q_i$ defined over the corresponding bounding box and the
weight $w_i$ is computed as in equation~\eqref{eq:sampling}. The
function stores an \pinline{elfi.Result} object as
\pinline{romc.result} attribute. The \pinline{elfi.Result} provides
some usefull functionalities for inspecting the obtained samples e.g.\
\pinline{romc.result.summary()} prints the number of the obtained
samples and their mean. A complete overview of these functionalities is
provided in ELFI's
\href{https://elfi.readthedocs.io/en/latest/api.html#elfi.methods.results.Sample}{official
  documentation}.

\subsubsection*{Function (ii): Compute an expectation}

\mintinline{python}{romc.compute_expectation(h)}
\vspace{5mm}

\noindent
This function computes the expectation
$E_{p(\thetab|\data)}[h(\thetab)]$ using
expression~\eqref{eq:expectation}. The argument \pinline{h} can be
any python \pinline{Callable}.

\subsubsection*{Function (iii): Evaluate the unnormalised posterior}
\mintinline{python}{romc.eval_unorm_posterior(theta, eps_cutoff=False)}
\vspace{5mm}

\noindent
This function computes the unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}.

\subsubsection*{Function (iv): Evaluate the normalised posterior}
\mintinline{python}{romc.eval_posterior(theta, eps_cutoff=False)}
\vspace{5mm}

\noindent
This function evaluates the normalised posterior. For doing so it
needs to approximate the partition function
$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; this is done using
the Riemann integral approximation. Unfortunately, the Riemann
approximation does not scale well in high-dimensional spaces, hence
the approximation is tractable only at low-dimensional parametric
spaces. Given that this functionality is particularly useful for
plotting the posterior, we could say that it is meaningful to be used
for up to $3D$ parametric spaces, even though it is not restricted to
that. Finally, for this functionality to work, the \pinline{bounds}
arguments must have been set at the initialisation of the
\pinline{elfi.ROMC} object.\footnote{The argument \pinline{bounds}
  should define a bounding box \emph{containing} all the mass of the
  prior; it may also contain redundant areas. For example, if the
  prior is the uniform defined over a unit circle i.e.\
  $c=(0,0), r=1$, the best bounds arguments is
  \pinline{bounds=[(-1,1),(-1,1)]}. However, any argument
  \pinline{bounds=[(-t,t),(-t,t)]} where $t\geq1$ is technically
  correct.}

\subsubsection*{Example - Sampling and compute expectation}

With the following code snippet, we perform weighted sampling from the
ROMC approximate posterior. Afterwards, we used some ELFI's built-in
tools to get a summary of the obtained samples. In figure
\ref{fig:example_sampling}, we observe the histogram of the weighted
samples and the acceptance region of the first deterministic function
(as before) alongside with the obtained samples obtained from
it. Finally, in the code snippet we demonstrate how to use the
\pinline{compute_expectation} function; in the current example we
define \pinline{h} in order to compute firstly the empirical mean and
afterwards the empirical variance. In both cases, the empirical result
is close to the ground truth $\mu = 0$ and $\sigma^2 = 1$.

\begin{pythoncode}
  seed = 21
  n2 = 50
  romc.sample(n2=n2, seed=seed)

  # visualize region, adding the samples now
  romc.visualize_region(i=1)

  # Visualise marginal (built-in ELFI tool)
  romc.result.plot_marginals(weights=romc.result.weights, bins=100, range=(-3,3))

  # Summarize the samples (built-in ELFI tool)
  romc.result.summary()
  # Number of samples: 1720
  # Sample means: theta: -0.0792

  # compute expectation
  print("Expected value   : %.3f" % romc.compute_expectation(h=lambda x: np.squeeze(x)))
  # Expected value   : -0.079

  print("Expected variance: %.3f" % romc.compute_expectation(h=lambda x: np.squeeze(x)**2))
  # Expected variance: 1.061
\end{pythoncode}

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_marginal.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_region_samples.png}
    \end{center}
  \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples. (b) Right: Acceptance region around $\theta_1^*$ with the obtained samples plotted inside.}
  \label{fig:example_sampling}
\end{figure}

\subsubsection*{Example - Evaluate Posterior}

The \pinline{romc.eval_unnorm_posterior(theta)} evaluates the
posterior at point $\theta$ using expression
\eqref{eq:aprox_posterior}. The \pinline{romc.eval_posterior(theta)}
approximates the partition function
$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
Riemann approximation as explained above. In our simple example, this
utility can provide a nice plot of the approximate posterior as
illustrated in figure~\ref{fig:approx_posterior}. We observe that the
approximation is quite close to the ground-truth posterior.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.75\textwidth]{./Thesis/images/chapter3/example_posterior.png}
    \end{center}
  \caption[Approximate posterior evaluation, at the 1D example.]{Approximate posterior evaluation.}
  \label{fig:approx_posterior}
\end{figure}

### 2021-07-08 latex_files/ch3/3_1_4_evaluation.tex
added all content from thesis to the paper

The ROMC implementation provides two functions for evaluating the inference results,

\begin{enumerate}[label=(\roman*)]
\item \mintinline{python}{romc.compute_divergence(gt_posterior, bounds=None, step=0.1,}

      \mintinline{python}{                        distance="Jensen-Shannon")}

\item \mintinline{python}{romc.compute_ess()}
\end{enumerate}

\subsubsection*{Function (i): Compute the divergence between ROMC approximation and a ground-truth posterior}

\pinline{romc.compute_divergence(gt_posterior, bounds=None, step=0.1,}

\pinline{                     distance="Jensen-Shannon")}
\vspace{5mm}

\noindent
This function computes the divergence between the ROMC approximation
and the ground truth posterior. Since the computation is performed
using the Riemann's approximation, this method can only work in low
dimensional parameter spaces; it is suggested to be used for up to the
three-dimensional parameter space. As mentioned in the beginning of
this chapter, in a real-case scenario it is not expected the
ground-truth posterior to be available. However, in cases where the
posterior can be approximated decently well with a numerical approach
(as in the current example) or with some other inference approach,
this function can provide a numerical measure of the agreement
between the two approximations. The argument \pinline{step} defines
the step used in the Riemann's approximation and the argument
\pinline{distance} can take either the \pinline{Jensen-Shannon} or the
\pinline{KL-divergence} value, for computing the appropriate distance.

\subsubsection*{Function (ii): Compute the effective sample size of the weighted samples}

\mintinline{python}{romc.compute_ess()}
\vspace{5mm}

\noindent
This function computes the Effective Sample Size (ESS) using the
following expression~\autocite{Sudman1967},

\begin{equation} \label{eq:ESS}
  ESS = \frac{(\sum_i w_i)^2}{\sum_i w_i^2}
\end{equation}

The ESS is a useful measure of the \textbf{actual} sample size, when
the samples are weighted. For example if in a population of $100$
samples one has a very large weight (e.g.\ $\approx 100$) whereas the
rest have small (i.e.\ $\approx 1$), the real sample size is close to
1; one sample dominates over the rest. Hence, the ESS provides a
measure of the equivalent uniformly weighted sample population.

\newpage

\begin{pythoncode}
res = romc.compute_divergence(wrapper, distance="Jensen-Shannon")                                 
print("Jensen-Shannon divergence: %.3f" % res)
# Jensen-Shannon divergence: 0.025

print("Nof Samples: %d, ESS: %.3f" % (len(romc.result.weights), romc.compute_ess()))
# Nof Samples: 19950, ESS: 16694.816
\end{pythoncode}

### 2021-07-08 latex_files/ch3/3_1_user.tex
added all content from thesis to the paper

In section \ref{subsec:general_design} we present an overview of the implementation. Afterwards, we divide the functionalities in 3 parts; in section \ref{subsec:training} we present the methods used for training (fitting) the model, in section \ref{subsec:inference} those used for performing the inference and in section \ref{subsec:evaluation} those used for evaluating the approximate posterior and the obtained samples.

\subsubsection{General design} 
\label{subsec:general_design}
\input{Thesis/ch3/3_1_1_general_design}

\subsubsection{Training part} 
\label{subsec:training}
\input{Thesis/ch3/3_1_2_training}

\subsubsection{Inference part} 
\label{subsec:inference}
\input{Thesis/ch3/3_1_3_inference}

\subsubsection{Evaluation part} 
\label{subsec:evaluation}
\input{Thesis/ch3/3_1_4_evaluation}

### 2021-07-08 latex_files/ch3/3_2_developer.tex
added all content from thesis to the paper

In the previous sections we described ROMC from the user's
point-of-view; the demonstration focused on performing the inference
with the ready-to-use tools. Apart from this use-case, a practitioner
can also use ROMC as a meta-algorithm, adding custom algorithms as
part of the method. Our implementation allows such intervention. In
the rest of the chapter, we will initially present the main internal
classes and then then we will exhibit how to add custom methods.

\subsubsection{Entities presentation}

In figure \ref{fig:example_training} we provide an overview of the
classes used in the algorithm. The class $\pinline{ROMC}$ can be
thought as the interface between the user and the method. The rest of
the classes are responsible for the back-end functionality. It is
important for the reader to remember the sequence of the events for
performing the inference, as demonstrated in figure
\ref{fig:romc_overview}.

The class \pinline{ROMC} is the main class of the method; it is
initialised when the user calls the method for the first time and its
state is updated throughout the inference.The initialisation of the
\pinline{ROMC} object sets the attributes \pinline{model},
\pinline{model_prior}, \pinline{bounds} and \pinline{inference_state} to
the appropriate values; the rest of the attributes are set to
\pinline{None}.\footnote{in all cases, we use the value None for
  indicating that an attribute is not yet initialised."}.

The \pinline{_sample_nuisance()} routine samples $n_1$ integers from
the discrete integer distribution. The \pinline{_define_objectives()}
routine initialises $n_1$ \pinline{OptimisationProblem} objects,
appends them to a \pinline{List} and stores this \pinline{List} as the
\pinline{romc.optim_problems} attribute. The
\pinline{_define_objectives()} function, initialises only the
attributes \pinline{objective} and \pinline{bounds} of the
\pinline{OptimisationProblem}; the rest are set to \pinline{None}. The
attribute \pinline{objective}, which is a \pinline{Callable}, is
created by setting the seed of the \pinline{model.generate()} to a
specific integer, turning the random generator to a deterministic.

Afterwards, depending on the boolean argument
\pinline{use_bo=True/False}, the function \linebreak \pinline{_solve_gradients}
or \pinline{_solve_bo} is called. Both of them follow the same
sequence of steps; for each optimisation problem they (a) solve it,
(b) create a \pinline{RomcOptimisationResult} object with the solution
and (c) store it as the \pinline{OptimisationProblem.result}
attribute. Although they both apply the same sequence of steps, each
method uses a different optimisation scheme;
\pinline{_solve_gradients} uses the \pinline{minimize} class of the
\pinline{scipy.optimize} library \autocite{2020SciPy-NMeth}, whereas
\pinline{_solve_bo} uses the \pinline{BoDeterministic} class.
\pinline{BoDeterministic} is a class we implemented for performing
Bayesian Optimisation and fitting a Gaussian Process surrogate model
to the objective function. It relies on the Gpy framework
\autocite{gpy2014} for fitting the Gaussian Process. The only
difference between \pinline{_solve_gradients} and \pinline{_solve_bo},
as effect in the state of the \pinline{OptimisationProblem} class, is
the initialisation of the attribute 
\pinline{OptimisationProblem.surrogate}, which is done only by
\pinline{_solve_bo}. The attribute is set to a \pinline{Callable} that
wraps the \pinline{GPyRegression.predict_mean} function. If
\pinline{_solve_gradients} is called, the attribute remains
\pinline{None}.

At this level, all optimisation problems have been solved. The method
\pinline{_filter_solutions} is used for discarding the optimisation
results that are over threshold. Afterwards, \pinline{_build_boxes}
estimates the bounding boxes around the accepted objective
functions. For each accepted objective function, a
\pinline{RegionConstructor} object is initialised in order to
construct the region. In the current implementation, for each
objective function we construct a single bounding box, but this may
change in a future approach; for being able to support multiple
bounding boxes per objective function, we decided to return a
\pinline{List} of \pinline{NDimBoundingBox} objects\footnote{Each
  \pinline{NDimBoundingBox} object represents a region.}. The
\pinline{List} is stored as the \pinline{OptimisationProblem.regions}
attribute.

By now, we have estimated the bounding boxes around the optimal
points. The last step, before defining the posterior, is the fitting
local surrogate models. This step is optional and is performed only
if the argument \pinline{fit_models} is set to \pinline{True}. The
routine \pinline{_fit_models}, fits a quadratic model on the area
around the optimal point for each objective function. For achieving
so, it asks for samples from the \pinline{NDimBoundingBox} object and
evaluates them using the \pinline{OptimisationProblem.objective}
function. Afterwards, based on these points, it fits a quadratic model
using the \pinline{linear_model.LinearRegression} and
\pinline{preprocessing.PolynomialFeatures} functions of the
\pinline{scikit-learn} package \autocite{scikit-learn}. The trained
model is stored as the attribute
\pinline{OptimisationProblem.local_surrogate}.

Finally, the \pinline{_define_posterior} method is used for creating
the \pinline{RomcPosterior} object and storing it as the
\pinline{ROMC.posterior} attribute. The method collects (a) all the
bounding boxes created so far (accessing the
\pinline{OptimisationProblem.regions} attributes and (b) all the objective functions. As the objective function it collects either \pinline{OptimisationProblem.local_surrogate} or \linebreak \pinline{OptimisationProblem.surrogate} or \pinline{OptimisationProblem.objective}, depending on which one is available based on the previous steps.

The description above summarises the sequence of steps needed for the
training part of the method. The conclusion is that a \pinline{ROMC}
object is initialised when the user calls the method. Throughout the
inference process, the \pinline{ROMC} object is always as a specific state,
which gets updated whenever an algorithmic step is executed. The rest
of the classes provide objects which are stored as attributes of
\pinline{ROMC}.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=\textwidth]{./Thesis/graphs/RomcEntityDiagram.png}
    \end{center}
  \caption[Overview of the entities (classes) of the ROMC implementation.]{Overview of the entities (classes) used at the deployment of the ROMC inference method, at the \pinline{ELFI} package.}
  \label{fig:example_training}
\end{figure}

\subsubsection{Extensibility of the ROMC method}

In this section we will explain how a practitioner may replace some
parts of the ROMC method with their custom alogorithms. We can locate
at least four such points; (a) the gradient-based solver (b) the
Bayesian Optimisation solver (c) the bounding box region construction
and (d) the fitting of the local surrogate models. Each of the
aforementioned tasks may be approached with completely different
algorithms than the ones we propose, without the rest of the method to
change.

The four replacable parts described above, are solved using the four
methods of the \linebreak \pinline{OptimisatioProblem} class;
\pinline{solve_gradients(**kwargs)}, \pinline{solve_bo(**kwargs)}, \linebreak
\pinline{build_region(**kwargs)},
\pinline{fit_local_surrogate(**kwargs)}. Therefore, the practitioner
should not alter at all the basic \pinline{ROMC} class. Instead, they
should deploy a custom optimisation problem class which inherits the
basic \pinline{OptimisatioProblem} class and overwrites the above four
functiones with custom ones. The only rule that should be followed
carefully, is that the cutom methods must have the same effect on the
attributes of the \pinline{ROMC} class, i.e. update them with the
appropriate objects as presented in table \ref{tab:extensibility}. For
example, a function that replaces \pinline{solve_gradients()} must
have the effect of storing a \pinline{RomcOptimisationResult} object
to the \pinline{OptimisationObject} attribute.

\begin{center} \label{tab:extensibility} \captionof{table}[Demonstration of the appropriate effect of each replaceable routine]{Table
    explaining the appropriate effect of each replaceable routine,
    i.e. which object they should attach to the appropriate
    attribute. The functions of the first column (\pinline{ROMC}
    class) call the correpsonding functions of the second column
    (\pinline{OptimisationProblem} class). The functions of the second
    column should execute their main functionality and update the
    appropriate attribute with a certain object, as described in the
    third column. }

\begin{tabular}{ c|c }
\hline
\pinline{OptimisationProblem} & \pinline{Effect} \\
\hline \hline
\pinline{solve_gradients()} & \pinline{result <- RomcOptimisationResult} \\
\hline
\multirow{ 2}{*}{\pinline{solve_bo()}} & \pinline{result <- RomcOptimisationResult} \\
& \pinline{surrogate <- Callable} \\
\hline
\pinline{build_region()} & \pinline{regions <- List[NDimBoundingBox]}\\
\hline
\pinline{fit_local_surrogate()} & \pinline{local_surrogate <- Callable}\\
\hline
\end{tabular}
\end{center}

\subsubsection*{Example: use a Neural Network as a local surrogate model}

Let's say we have observed that the local area around $\theta_i^*$ is
too complex to be represented by a simple quadratic model\footnote{as
  in the current implementation}. Hence, the user selects a neural
network as a good alternative. In the following snippet, we
demonstrate how they could implement this enhancement, without much
effort; (a) they have to develop the neural network using the package
of their choice (b) they must create a custom optimisation class which
inherites the basic \pinline{OptimisationClass} and (c) they have to
overwrite the \pinline{fit_local_surrogate} routine, with one that
sets the neural network's prediction function as the
\pinline{local_surrogate} attribute. The argument \pinline{**kwargs}
may be used for passing all the important arguments e.g.\ training
epochs, gradient step etc. If, for example, they would like to set the
size of the training set dynamically, we may replace \pinline{x =
  self.regions[0].sample(30)} with \pinline{x =
  self.regions[0].sample(kwargs["nof_examples"])}. Finally, they must
pass the custom optimisation class, when calling the \pinline{ROMC}
method.

\newpage

\begin{pythoncode}
  class NeuralNetwork:
      def __init__(self, **kwargs):
          # set the input arguments

      def train(x, y):
          # training code

      def predict(x):
          # prediction code

  # Inherit the base optimisation class
  class customOptim(elfi.methods.parameter_inference.OptimisationProblem):
      def __init__(self, **kwargs):
          super(customOptim, self).__init__(**kwargs)

      # overwrite the function you want to replace
      def fit_local_surrogate(**kwargs):
          # init and train the NN
          x = self.regions[0].sample(30) # 30 training points
          y = [np.array([self.objective(ii) for ii in x])]
          nn = NeuralNet()
          nn.train(x,y)

          # set the appropriate attribute
          self.local_surrogate = nn.predict

          # update the state
          self.state["local_surrogate"] = True

  # pass the custom inference method as argument
  romc = elfi.ROMC(dist, bounds, custom_optim_class=customOptim)
\end{pythoncode}

### 2021-07-08 latex_files/ch4/4_1_ex1.tex
added all content from thesis to the paper

This examples is implemented for validating that the ROMC
implementation works accurately in a multidimensional parameter
space.

\subsubsection*{Problem Definition}

The equations describing this inference problem are presented below.

\begin{gather} \label{eq:ex1_equations}
  p(\thetab) = p(\theta_1)p(\theta_2)
  = \mathcal{U}(\theta_1; -2.5, 2.5) \mathcal{U}(\theta_2; -2.5, 2.5)\\
  p(\yb|\thetab) = \mathcal{N}(\yb;\thetab, \mathcal{I})\\
  p(\thetab|\yb) = \frac{1}{Z}p(\thetab)p(\yb|\thetab)\\
  Z = \int_{\thetab} p(\thetab)p(\yb|\thetab) d\thetab
\end{gather}

\noindent
In this simple example, it is feasible to evaluate the likelihood and
the unnormalised posterior. Additionally, the partition function $Z$
can be estimated by numerical integration (i.e. using Riemann's
approximation). Hence, it is feasible to compute the ground-truth
posterior with numerical approximation. Setting the observation
$\data$ to $(-0.5,0.5)$, the ground-truth posterior is illustrated in
figure \ref{fig:ex2_4}. In table \ref{tab:ex2}, we present the
ground-truth statistics i.e.\ $\mu, \sigma$ of the marginal posterior
distributions.

\subsubsection*{Performing the inference}

We perform the inference using the following hyperparameters
$n_1=500, n_2=30, \epsilon=0.4$. This set-up leads to a total of
$15000$ samples. As observed in the histogram of distances (figure
\ref{fig:ex2_1}), in the gradient-based approach, all optimisation
problems reach an almost zero-distance end point; hence all optimal
points are accepted. In the Bayesian optimisation scheme, the vast
majority of the optimisation procedures has the same behaviour; there
are only 4 optimal distances above the limit. In figure
\ref{fig:ex2_2}, the acceptance area of a specific optimisation
problem is demonstrated. We observe that both optimisation schemes
lead to a similar bounding box construction. This specific example is
representative of the rest of the optimisation problems; due to the
simplicity of the objective function, in most cases the optimal points
are similar and the surrogate model represents accurately the local
region. Hence, similar proposal regions are obtained by the two
optimisation alternatives.

The histograms of the marginal distributions, based on the weighted
samples, are presented in figure \ref{fig:ex2_3}. In the same figure,
we also plot the ground-truth distribution with the red dotted
line. We observe that the weighted samples follow quite accurately the
ground-truth distribution. This is also confirmed by the statistics
provided in table \ref{tab:ex2}; the sample mean $\mu$ and standard
deviation $\sigma$ are similar to the ground-truth for both parameters
$\theta_1$ and $\theta_2$. We also observe that both optimisation
schemes produce accurate samples.

Finally, the ground-truth and the approximate posteriors are presented
in figure \ref{fig:ex2_4}. We also confirm that the approximations are
close to the ground truth posteriors. As a note, we can observe that
the approximate posteriors present a diamond-shape in the mode of the
posterior; this happens due to the approximation of the circular
Gaussian-shape posterior with a sum of square boxes. The divergence
between the ground-truth distribution and the approximate ones is
$0.077$, using the Jensen-Shannon distance, which confirms the
satisfying matching between the two posteriors.

In this experiment we observed that the implementation fulfilled the
theoretical expectations for the ROMC inference method.

\begin{center} \label{tab:ex2}
\begin{tabular}{ c|c|c|c|c|c }
\hline
& $\mu_{\theta_1}$ & $\sigma_{\theta_1}$ & $\mu_{\theta_2}$ & $\sigma_{\theta_2}$ & Divergence\\
\hline \hline
Ground-truth & -0.45 & 0.935 & 0.45 & 0.935 & \\
\hline
ROMC (gradient-based) & -0.474 & 0.994 & 0.502 & 0.966 & 0.068\\
\hline
ROMC (Bayesian optimisation) & -0.485 & 0.987 & 0.511 & 0.939 & 0.069\\
\hline
\end{tabular}
\end{center}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_distance_hist.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_distance_hist_bo.png}
    \end{center}
    \caption[2D example, histogram of distances]{Histogram of distances $d_i^*, i \in \{1, \ldots,
      n_1\}$. The left graph corresponds to the gradient-based
      approach and the right one to the Bayesian optimisation
      approach.}
  \label{fig:ex2_1}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_region_1.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_region_1_bo.png}
    \end{center}
    \caption[2D example, the bounding box of the $1^{st}$ optimisation problem.]{The acceptance region and the bounding box of the
      $1^{st}$ optimisation problem; In the left plot, it is
      constructed with a gradient-based approach and in the right
      figure with Bayesian optimisation. We observe that both
      approaches construct a similar bounding box.}
  \label{fig:ex2_2}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_hist_t1_romc.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_hist_t2_romc.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_hist_t1_romc_bo.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_hist_t2_romc_bo.png}\\
    \end{center}
    \caption[2D example, histogram of the weighted samples.]{Histogram of the weighted samples. The first row of plots
      refers to the gradient-based optimisation scheme, while the
      second row to the Bayesian optimisation. We represent the
      ground-truth marginal distribution $p(\theta_i|\data$ with the
      red dotted line. We observe that the samples follow the ground
      truth distribution quite accurately.}
  \label{fig:ex2_3}
\end{figure}

\begin{figure}[ht]
  \begin{center}
      \includegraphics[width=0.5\textwidth]{./Thesis/images/chapter4/ex2D_gt_posterior.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_romc_posterior.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/ex2D_romc_posterior_bo.png}
    \end{center}
    \caption[2D example, evaluation of the approximate posterior.]{(a) First row: Ground-truth posterior approximated
      computationally. (b) Second row (left): ROMC approximate
      posteriors using gradient-based approach. The divergence from
      the ground-truth using the Jensen-Shannon distance is
      $0.068$. (c) Second row (right): ROMC approximate posterior
      using Bayesian optimisation. The divergence from the
      ground-truth using the Jensen-Shannon distance is $0.069$}
  \label{fig:ex2_4}
\end{figure}

In this simple artificial 2D example, with the ground-truth
information available, we confirmed that our implementation produces
accurate approximations. In the following section, we will question it
in a more involved case.

### 2021-07-08 latex_files/ch4/4_2_ma2.tex
added all content from thesis to the paper

The second example is the second-order moving average (MA2) which is
used by the ELFI package as a fundamental model for testing all
inference implementations. This example is chosen to confirm that our
implementation of the ROMC approach produces sensible results in a
general model, since the previous ones where artificially created by
us.

\subsubsection*{Problem definition}

The second-order moving average (MA2) is a common model used for
univariate time series analysis. The observation at time $t$ is given by:

\begin{gather} \label{eq:ma2}
y_t = w_t + \theta_1 w_{t-1} + \theta_2 w_{t-2}\\
\theta_1, \theta_2 \in \R, \quad  w_k \sim \mathcal{N}(0,1), k \in \mathbb{Z}
\end{gather}

\noindent
The random variables $w_{k} \sim \mathcal{N}(0,1), k \in \mathbb{Z}$
represent an independent and identically distributed white noise and
$\theta_1, \theta_2$ the dependence from the previous
observations. The number of consecutive observations $T$ is a
hyper-parameter of the model; in our case we will set
$T=100$. Generating an MA2 time-series is pretty easy and efficient
using a simulator, therefore using a likelihood-free inference method
quite convenient. At the specific example, we use the prior proposed
by \cite{Marin2012} for guaranteeing that the inference problem is
identifiable, i.e.\ loosely speaking the likelihood will have just one
mode. The multivariate prior, which is given in the equation
\eqref{eq:ma2_prior}, follows a triangular shape as plotted in figure
\ref{fig:ma2_1}.

\begin{equation} \label{eq:ma2_prior}
p(\thetab) = p(\theta_1)p(\theta_2|\theta_1)
= \mathcal{U}(\theta_1;-2,2)\mathcal{U}(\theta_2;\theta_1-1, \theta_1+1)
\end{equation}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_prior_samples.png}
    \end{center}
    \caption[MA2 example, prior distribution.]{Prior distribution proposed by \cite{Marin2012}. The
      samples follow a triangular shape.}
  \label{fig:ma2_1}
\end{figure}

\noindent
The vector $\yb_0 = (y_1, \ldots, y_{100})$ used as the observation,
has been generated with $\thetab=(0.6, 0.2)$. The dimensionality of
the output $\yb$ is quite large, therefore we use summary
statistics. Considering that the ouptput vector represents a
time-series signal, we prefer the autocovariance as the summary
statistic; we incorporate the autocovariances with $lag=1$ and
$lag=2$, as shown in equation \eqref{eq:ma2_summary}. Hence, the
distance is defined as the squared euclidean distance between the
summary statistics; this is the same choice as in \cite{Marin2012}.

\begin{gather} \label{eq:ma2_summary}
  s_1(\yb) = \frac{1}{T-1} \sum_{t=2}^T y_ty_{t-1}\\
  s_2(\yb) = \frac{1}{T-2} \sum_{t=3}^T y_ty_{t-2} \\
  s(\yb) = (s_1(\yb), s_2(\yb))\\
  d = ||s(\yb) - s(\yb_0)||_2^2
\end{gather}

\subsubsection*{Perform the inference}

As in the previous example, we perform the inference using the two
optimisation alternatives, the gradient-based and the Bayesian
optimisation. In this way, we compare the results obtained in each
step. For comparing the samples drawn with our implementation, we will
use the samples obtained with Rejection ABC. In figure \ref{fig:ma2_2}
we observe that in most cases the optimal distance $d_i^*$ is close to
zero in both optimisation approaches.

In figure \ref{fig:ma2_5}, we have chosen three different
deterministic optimisation cases (i.e. three different seeds) for
illustrating three different cases. In the first case, both
optimisation schemes lead to the creation of almost the same bounding
box. In the second case, the bounding box has a similar shape, though
different size and it is shifted along the $\theta_1$ axis. Finally,
in the third case, the bounding boxes are completely different. We can
thus conclude that although fitting a surrogate model has important
computational advantages, there is no guarantee that it will reproduce
accurately the local region around the optimal point. This
approximation error may lead to the construction of a considerably
different proposal region, which in turn, explains the differences in
the histogram of the marginal distributions presented in figure
\ref{fig:ma2_3} and in the approximate posteriors in figure
\ref{fig:ma2_4}.

In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
marginal posteriors, using the three different methods; (a) Rejection
ABC (first line), (b) ROMC with gradient-based optimisation (second
line) and (c) ROMC with Bayesian optimisation (third
line). Undoubtedly, we observe that there is a significant similarity
between the three approaches. The Rejection ABC inference has been set
to infer 10000 accepted samples, with threshold $\epsilon=0.1$. The
large number of samples and the small distance from the observations
let us treat them as ground-truth information. In the table \ref{tab:ma2}
we present the empirical mean $\mu$ and standard deviation $\sigma$
for each inference approach. We observe that there is a significant
agreement between the approaches, which verifies that the ROMC
implementation provides sensible results. Finally, in figure
\ref{fig:ma2_4} we provide the ROMC approximate posteriors using
gradients and Bayesian optimisation; as confirmed by the statistics in
\ref{tab:ma2}, both posteriors have a single mode, located at the same
point, with a larger variance observed in the Bayesian Optimisation
case. The posteriors mode is quite close to the parameter
configuration that generated the data i.e.\ $\thetab = (0.6, 0.2)$.

\begin{center} \label{tab:ma2}
\begin{tabular}{ c|c|c|c|c }
\hline
& $\mu_{\theta_1}$ & $\sigma_{\theta_1}$ & $\mu_{\theta_2}$ & $\sigma_{\theta_2}$ \\
\hline \hline
Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
\hline
ROMC (gradient-based) & 0.495 & 0.136 & 0.048 & 0.178 \\
\hline
ROMC (Bayesian optimisation) & 0.510 & 0.156 & 0.108 & 0.183 \\
\hline
\end{tabular}
\end{center}

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist_bo.png}
    \end{center}
    \caption[MA2 example, histogram of distances]{Histogram of distances
      $d_i^*, i \in \{ 1, \ldots, n_1 \}$. The left graph corresponds
      to the gradient-based approach and the right one to the Bayesian
      optimisation approach.}
  \label{fig:ma2_2}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.png}\\
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_2.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_2_bo.png}\\
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
    \end{center}
  \caption[MA2 example, the acceptance region in three distinct optimisation problems.]{Visualisation of the acceptance region in 3 different optimisation problems. Each row illustrates a different optimisation problem, the left column corresponds to the gradient-based approach and the right column to the Bayesian optimisation approach. The examples have been chosen to illustrate three different cases; in the first case, both optimisation schemes lead to similar optimal point and bounding box, in the second case the bounding box is similar in shape but a little bit shifted to the right relatively to the gradient-based approach and in the third case, both the optimal point and the bounding box is completely different.}
  \label{fig:ma2_5}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
    \end{center}
    \caption[MA2 example, evaluation of the marginal distributions.]{Histogram of the marginal distributions using
      different inference approaches; (a) in the first row, the
      approximate posterior samples are obtained using Rejection ABC
      sampling (b) in the second row, using ROMC sampling with
      gradient-based approach and (c) in the third row, using ROMC
      sampling with Bayesian optimisation approach. The vertical (red)
      line represents the samples mean $\mu$ and the horizontal
      (black) the standard deviation $\sigma$.}
  \label{fig:ma2_3}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.png}
      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.png}
    \end{center}
    \caption[MA2 example, approximate posterior.]{ROMC approximate posteriors using gradient-based approach
      (left) and Bayesian optimisation approach (right).}
  \label{fig:ma2_4}
\end{figure}

### 2021-07-08 latex_files/ch4/4_3_exec.tex
added all content from thesis to the paper

In this section, we will present the execution time of the basic ROMC
functionalities. Apart from performing the inference accurately, one
of the notable advantages of ROMC is its efficiency. In terms of
performance, ROMC holds two key advantages.

Firstly, all its subparts are parallelisable; optimising the objective
functions, constructing the bounding boxes, fitting the local
surrogate models, sampling and evaluating the posterior can be
executed in a fully-parallel fashion. Therefore, the speed-up can be
extended as much as the computational resources allow. Specifically,
at the CPU level, the parallelisation can be incorporate all available
cores. A similar design can be utilised in the case of having a
cluster of PCs available. Parallelising the process at the GPU level
is not that trivial, since the parallel tasks are more complicated
than simple floating-point operations. Our current implementation
supports parallelisation only at the CPU level, exploiting all the cores of the CPU.\footnote{At a future update, it is possible to support parallelising the processes using a cluster of PCs.}. In subsection
\ref{subsubsec:parallel}, we will demonstrate the speed-up achieved
through parallelisation.

The second important advantage concerns the execution of the
training and the inference phase in distinct time-slots. Therefore, one
can consume a lot of training time and resources but ask for
accelerated inference. The use of a lightweight surrogate model around
the optimal point exploits this essential characteristic; it trades
some additional computational burden at the training phase, for
exercising faster inference later. In figures \ref{fig:exec_solve},
\ref{fig:exec_regions}, \ref{fig:exec_posterior},
\ref{fig:exec_sample} we can observe this advantage. The
example measured in these figures is the simple 1D example, used in
the previous chapter. We observe that fitting local surrogate models
slows down the training phase (estimating the regions) by a factor of
$1.5$, but provides a speed-up of $15$ at the inference phase i.e.\
approximating unnormalised posterior. This outcome would be even more
potent in larger models, where running the simulator is even more
expensive.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_solve_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_solve_bo.png}
    \end{center}
    \caption[Execution time for solving the optimisation problems.]{Execution time for defining and solving the optimisation
      problems. We observe that the Bayesian optimisation scheme is much more expensive, performing the task slower by a factor of $75$.}
  \label{fig:exec_solve}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_grad_fit.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_bo.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_bo_fit.png}
    \end{center}
    \caption[Execution time for constructing the n-dimensional bounding box regions.]{Execution time for constructing the n-dimensional
      bounding box region and, optionally, fitting the local surrogate
      models. We observe that fitting the surrogate models incurs a
      small increase by a factor of $1.5$.}
  \label{fig:exec_regions}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_bo.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_grad_fit.png}
    \end{center}
    \caption[Execution time for evaluation the unnormalised posterior.]{Execution time for evaluating the unnormalised posterior
      approximation. We observe that using a Gaussian Process surrogate model is almost 5 times faster than the simulator and the quadratic local surrogate model 15 times faster.}
  \label{fig:exec_posterior}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_grad.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_bo.png}\\
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_grad_fit.png}
    \end{center}
    \caption[Execution time for sampling from the approximate posterior.]{Execution time for sampling from the approximate
      posterior. We observe a small speed-up using the local surrogate model.}
  \label{fig:exec_sample}
\end{figure}

\subsubsection{The effect of parallelisation} \label{subsubsec:parallel}

In this section, we measure the execution times for (a) solving the optimisation problems, (b) construct the bounding boxes, (c) sample from the posterior and (d) evaluate the approximate posterior using parallelisation. The model used for this experiment is the simple two-dimensional example of section \ref{subsec:ex1}. All experiments
have been executed in a laptop Dell XPS 15 with an 8-core CPU. Therefore the maximum expected speed-up can reach a factor of 8. Normally the observed speed-up is lower due to the overhead of setting-up the parallel processes. The experiments confirm our general expectation; the parallel version performs all tasks between 2.5 and 6 times faster compared to the sequential, as shown in figures \ref{fig:exec_parallel} and \ref{fig:exec_parallel2}. The most benefited task is solving the optimisation problems which is performed almost 6 times faster. Sampling is executed 3.5 times faster, whereas evaluating the posterior and constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/solve_problems_parallel.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/estimate_regions_parallel.png}
    \end{center}
    \caption[Execution time of the training part exploiting parallelisation]{Execution
      time of the training part exploiting parallelisation. At the left figure, we measure
      the execution time for solving the optimisation problems. At the
      right figure, we measure the execution time for constructing the
      bounding boxes.}
  \label{fig:exec_parallel}
\end{figure}

\begin{figure}[ht]
    \begin{center}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/sample_parallel.png}
      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/eval_post_parallel.png}
    \end{center}
    \caption[Execution time of the inference part exploiting parallelisation.]{Execution
      time of the inference part exploiting parallelisation. At the left figure, we measure
      the execution time for sampling $n_2=50$ points per region. At the right figure we measure the execution time for evaluating the posterior at a batch of $50$ points.}
  \label{fig:exec_parallel2}
\end{figure}

### 2021-07-08 latex_files/ch5/5_1_outcomes.tex
added all content from thesis to the paper

In this dissertation, we have studied the likelihood-free inference
approaches, focusing on the novel ROMC method that we implemented in
an open-source software package. The main contribution of the
dissertation is this implementation which can be used mainly by the
research community for further experimentation.

In chapter 2, we presented the simulator-based models explaining the
particularities of the inference when the likelihood is not
tractable. We then presented an overview of the Optimisation
Monte-Carlo methods (OMC and ROMC) examining their strategies on
approximating the posterior. We discussed the point-of-view of the
ROMC approach, demonstrating the mathematic modelling it
introduces. We also presented the aspect of ROMC as a
meta-algorithm. Finally, at the end of chapter 2, we transformed the
mathematical modelling of ROMC into algorithmic form. Up to this
point, the dissertation mainly restates the ideas presented in the
original paper \autocite{Ikonomov2019} and, hence, it can be used
together with the paper by a reader who wants to understand the ROMC
approach.

The notable contribution of the dissertation is the implementation of
the method in the ELFI package. Due to the novelty of the method, it
has not been implemented in any other package so far. We tried to
implement the method focusing on four principles: simplicity,
accuracy, efficiency and extensibility. We tried to provide the user
with a simple-to-use method. Therefore, we followed the guidelines of
the ELFI package, aligning ROMC with the other provided inference methods. We
also kept the function calls simple, asking only for the necessary
arguments. Secondly, we tested the implementation on a range of
examples for ensuring accurate inference results. We designed some
artificial examples for evaluating our implementation using
ground-truth information. We also tested that it works smoothly under
general models, not artificially created by us. Thirdly, we tried to
solve the tasks efficiently. We applied parallel processing which offers a major speedup in the the execution time of all tasks. Furthermore, we avoided using redundant calls, we
exploited vectorisation when possible, and we avoided expensive
for-loops.  We measured the execution time of the method for providing
an overview of the time needed for performing each task. Finally, we
preserved extensibility; this was a significant priority in the
implementation design. Apart from offering a ready-to-use inference
method, we want our implementation to serve as the initial point for
researchers who would like to further experiment with the method. This
requirement aligns with the nature of ROMC as a meta-algorithm; one
can replace the method involved in a specific task without the rest of
the algorithm to collapse. Finally, we provide extensive documentation
of all implemented methods and a collection of examples for
illustrating the main use-cases. We wanted the reader to be able to
observe and interact in-practice with the functionalities we offer.

### 2021-07-08 latex_files/ch5/5_2_future.tex
added all content from thesis to the paper

Future research directions involve both the theoretical and
implementation aspect of the ROMC approach.

In the theoretical part, an essential drawback of the ROMC approach is
the creation of a single bounding box for each objective
function. This simplification can be error-prone in situations where
we have ignorance about the properties of the objective
function. Consider, for example, the scenario where an objective
function defines multiple tiny disjoint acceptance areas; the
optimiser will (hopefully) reach only one, and a bounding box will be
constructed around it, compressing all the mass of the distribution
there. Therefore, this small region will erroneously dominate the
specific objective function. In general, the existence of multiple
disjoint acceptance regions complicates the inference. A naive
approach for improving this failure would be solving each optimisation
problem more than once, using different starting points. With this
approach, we hope that the optimiser will reach a different local
minimum in each execution. However, apart from the added computational
complexity, this approach does not guarantee an improvement. In
general, identifying all the acceptance regions in an unconstrained
problem with many local optimums is quite challenging.

Further research on the optimisation part, may also focus on modelling
the case where the parameter space is constrained. Consider, for
example, the case where a parameter represents a probability (as in
the tuberculosis example). In this case, the parameter is constrained
in the region $[0,1]$. Defining a prior with zero mass outside of this
area ensures that all non-acceptable values will not contribute to the
posterior. However, the optimiser does not consider the prior. Hence,
we can observe the phenomenon where the majority of the (or all)
acceptance regions is cancelled-out by the prior. In this case, the
posterior will be dominated by a few acceptance region parts that
overlap with the prior, which is not robust. Dealing with such
constrained non-linear optimisation problems, which is particularly
challenging, can serve as a second research line.

The previous two research directions overlap with the field of
optimisation. Another research avenue, more concentrated on the ROMC
method, is the approximation of the acceptance regions with a more
complex geometrical shape than the bounding box. In the current
implementation, the bounding box only serves as the proposal region
for sampling; running the indicator function afterwards, we manage to
discard the samples that are above the distance threshold. This
operation is quite expensive from a computational point of view. The
workaround proposed from the current implementation is fitting a local
surrogate model for avoiding running the expensive simulator. However,
fitting a surrogate model for every objective function is still
expensive. Approximating the acceptance region accurately with a more
complicated geometrical structure could remove the burden of running
the indicator function later. For example, a simple idea would be
approximating the acceptance region with a set of non-overlapping
bounding boxes. We insist on using the square as the basic shape
because we can efficiently sample from a uniform distribution defined
on it. A future research direction could investigate this capability.

On the implementation side, the improvements are more concrete. A critical
improvement would be performing the training and the inference phase
in a batched fashion. For example, the practitioner may indicate that
he would like to approximate the posterior in batches of N
optimisation problems; this means that the approximation should be
updated after each N new bounding boxes are constructed. Utilising
this feature, the user may interactively decide when he would like to
terminate the process. Finally, the implementation could also support
saving the state of the training or the inference process for
continuing later. In the current implementation, although the user can
save the inference result, he cannot interrupt the training phase and
continue later.

### 2021-07-08 latex_files/python_scripts/example_1D.py

@@ -1,0 +1,300 @@
+""" This script generates the figures used in the implementation example.
+"""
+
+import os
+import scipy.stats as ss
+import scipy.integrate as integrate
+import matplotlib.pyplot as plt
+import timeit
+import numpy as np
+import elfi
+
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+
+prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images'
+np.random.seed(21)
+
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    r"""Implements the distribution
+    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
+                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
+                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
+    """
+
+    def rvs(self, theta, seed=None):
+        """Vectorized sampling from likelihood.
+
+        """
+        assert isinstance(theta, np.ndarray)
+        theta = theta.astype(np.float)
+        samples = np.empty_like(theta)
+
+        c = 0.5 - 0.5 ** 4
+
+        tmp_theta = theta[theta <= -0.5]
+        samples[theta <= -
+                0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta <= -0.5] = np.inf
+
+        tmp_theta = theta[theta <= 0.5]
+        samples[theta <= 0.5] = ss.norm(
+            loc=tmp_theta**4, scale=1).rvs(random_state=seed)
+        theta[theta <= 0.5] = np.inf
+
+        tmp_theta = theta[theta < np.inf]
+        samples[theta < np.inf] = ss.norm(
+            loc=tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta < np.inf] = np.inf
+
[…]

### 2021-07-08 latex_files/python_scripts/example_1D_gp.py

@@ -1,0 +1,290 @@
+""" This script generates the figures used in the implementation example.
+"""
+
+import timeit
+import numpy as np
+import elfi
+
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+import matplotlib.pyplot as plt
+
+import scipy.integrate as integrate
+import scipy.stats as ss
+
+import os
+# from matplotlib import rc
+# rc('font', **{'family':'sans-serif','sans-serif':['Helvetica']})
+# ## for Palatino and other serif fonts use:
+# #rc('font',**{'family':'serif','serif':['Palatino']})
+# rc('text', usetex=True)
+
+
+prepath = '/home/givasile/ORwDS/edinburgh-thesis/'
+np.random.seed(21)
+
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    r"""Implements the distribution
+    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
+                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
+                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
+    """
+
+    def rvs(self, theta, seed=None):
+        """Vectorized sampling from likelihood.
+
+        """
+        assert isinstance(theta, np.ndarray)
+        theta = theta.astype(np.float)
+        samples = np.empty_like(theta)
+
+        c = 0.5 - 0.5 ** 4
+
+        tmp_theta = theta[theta <= -0.5]
+        samples[theta <= -0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta <= -0.5] = np.inf
+
+        tmp_theta = theta[theta <= 0.5]
+        samples[theta <= 0.5] = ss.norm(loc=tmp_theta**4, scale=1).rvs(random_state=seed)
+        theta[theta <= 0.5] = np.inf
[…]

### 2021-07-08 latex_files/python_scripts/example_2D.py

@@ -1,0 +1,338 @@
+import logging
+import matplotlib.pyplot as plt
+import scipy.integrate as integrate
+from elfi.methods.parameter_inference import ROMC
+from elfi.examples import ma2
+import elfi
+import numpy as np
+import scipy.stats as ss
+import matplotlib
+import os
+matplotlib.rcParams['text.usetex'] = True
+
+logging.basicConfig(level=logging.INFO)
+prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images/chapter4/'
+
+
+def plot_marginal(samples, weights, mean, std, marg, title, xlabel, ylabel, bins,
+                  range, ylim, savepath):
+    plt.figure()
+    plt.title(title)
+    plt.hist(x=samples,
+             weights=weights,
+             bins=bins, density=True, range=range)
+    # plt.axvline(mean, 0, 1,
+    #             color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
+    # plt.axhline(.2,
+    #             (mean-std-range[0])/(range[1] - range[0]),
+    #             (mean+std-range[0])/(range[1] - range[0]),
+    #             color="k",
+    #             linestyle="--", label=r"$\sigma = %.3f$" % (std))
+    x = np.linspace(-3, 3, 40)
+    y = [marg(tmp_x) for tmp_x in x]
+    plt.plot(x, y, "r--", label=r"$p(\theta_1|y_0)$")
+    plt.plot
+    plt.xlabel(xlabel)
+    plt.ylabel(ylabel)
+    plt.ylim(ylim)
+    plt.legend()
+    plt.savefig(savepath, bbox_inches='tight')
+    plt.show(block=False)
+
+
+# Set seed for reproducibility
+seed = 21
+np.random.seed(seed)
+
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
[…]

### 2021-07-08 latex_files/python_scripts/example_MA2.py

@@ -1,0 +1,205 @@
+import logging
+import matplotlib.pyplot as plt
+from elfi.methods.parameter_inference import ROMC
+from elfi.examples import ma2
+import elfi
+import numpy as np
+import scipy.stats
+import matplotlib
+import os
+matplotlib.rcParams['text.usetex'] = True
+
+logging.basicConfig(level=logging.INFO)
+prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images/chapter4/'
+
+
+def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
+    plt.figure()
+    plt.title(title)
+    plt.hist(x=samples,
+             weights=weights,
+             bins=bins, density=True, range=range)
+    plt.axvline(mean, 0, 1,
+                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
+    plt.axhline(1,
+                (mean-std-range[0])/(range[1] - range[0]),
+                (mean+std-range[0])/(range[1] - range[0]),
+                color="k",
+                linestyle="--", label=r"$\sigma = %.3f$" % (std))
+    plt.xlabel(xlabel)
+    plt.ylabel(ylabel)
+    plt.ylim(ylim)
+    plt.legend()
+    plt.savefig(savepath, bbox_inches='tight')
+    plt.show(block=False)
+
+
+# Set seed for reproducibility
+seed = 1
+np.random.seed(seed)
+model = ma2.get_model(seed_obs=seed)
+
+# plot prior samples
+
+x = model.generate(1000)
+plt.figure()
+plt.title("Samples from the prior")
+plt.xlabel(r"$\theta_1$")
+plt.ylabel(r"$\theta_2$")
+plt.plot(x["t1"], x["t2"], "bo")
+plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
+            bbox_inches="tight")
+plt.show(block=False)
+
+####### ROMC with gradients ################
+n1 = 1000
+n2 = 20
+bounds = [(-2, 2), (-1.25, 1.25)]
+eps = .01
+vis_ind_1 = 12
+vis_ind_2 = 1
+vis_ind_3 = 2
+
+romc = ROMC(model, bounds=bounds, discrepancy_name="d")
+romc.solve_problems(n1=n1, seed=seed)
+romc.distance_hist(savefig=os.path.join(
[…]

### 2021-07-08 latex_files/python_scripts/execution_times.py

@@ -1,0 +1,347 @@
+""" This script generates the figures used in the implementation example.
+"""
+
+import os
+import scipy.stats as ss
+import scipy.integrate as integrate
+import matplotlib.pyplot as plt
+import timeit
+import numpy as np
+import elfi
+
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+
+
+prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/images/chapter4'
+np.random.seed(21)
+
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    r"""Implements the distribution
+    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
+                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
+                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
+    """
+
+    def rvs(self, theta, seed=None):
+        """Vectorized sampling from likelihood.
+
+        """
+        assert isinstance(theta, np.ndarray)
+        theta = theta.astype(np.float)
+        samples = np.empty_like(theta)
+
+        c = 0.5 - 0.5 ** 4
+
+        tmp_theta = theta[theta <= -0.5]
+        samples[theta <= -
+                0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta <= -0.5] = np.inf
+
+        tmp_theta = theta[theta <= 0.5]
+        samples[theta <= 0.5] = ss.norm(
+            loc=tmp_theta**4, scale=1).rvs(random_state=seed)
+        theta[theta <= 0.5] = np.inf
+
+        tmp_theta = theta[theta < np.inf]
+        samples[theta < np.inf] = ss.norm(
+            loc=tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta < np.inf] = np.inf
+
[…]

### 2021-07-08 latex_files/python_scripts/execution_times_parallel.py

@@ -1,0 +1,317 @@
+"""Testing the execution times using parallelisation."""
+
+import timeit
+import time
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy.integrate as integrate
+import scipy.ndimage as ndimage
+import scipy.stats as ss
+import os
+import elfi
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+
+np.random.seed(21)
+
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    def rvs(self, th1, th2, seed=None):
+        assert isinstance(th1, np.ndarray)
+        assert isinstance(th2, np.ndarray)
+        assert th1.ndim == 2
+        assert th2.ndim == 2
+        assert np.allclose(th1.shape, th2.shape)
+
+        x = []
+        for i in range(th1.shape[0]):
+            cur_th = np.concatenate((th1[i], th2[i]))
+            x.append(ss.multivariate_normal(
+                mean=cur_th, cov=1).rvs(random_state=seed))
+        return np.array(x)
+
+    def pdf(self, x, th1, th2):
+        assert isinstance(th1, float)
+        assert isinstance(th2, float)
+        assert isinstance(x, np.ndarray)
+
+        th = np.stack((th1, th2))
+        rv = ss.multivariate_normal(mean=th, cov=1)
+        nof_points = x.shape[0]
+        prod = 1
+        for i in range(nof_points):
+            prod *= rv.pdf(x[i])
+        return prod
+
+
+def create_factor(x):
+    lik = Likelihood()
+    pr = Prior()
+
+    def tmp_func(th1, th2):
+        return lik.pdf(x, th1, th2) * pr.pdf(th1) * pr.pdf(th2)
[…]

### 2021-07-09 latex_files/ch2/2_1_LFI_intro.tex
added content in chapter 2

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism \cite{Gutmann2016}. The key characteristic
of these models is that although we can sample (simulate) data points,
we cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
given a set of parameters $\thetab$, produces samples $\yb$ in a
stochastic manner i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models are particularly captivating due to the
modelling freedom they provide; any physical process that can be
conceptualised as a computer program of finite steps can be modelled
as a simulator-based model without any compromise. The modelling
freedom includes any amount of hidden (unobserved) internal variables
or rule-based decisions. Hence, implicit models are often used to
model physical phenomena in many fiels such as biology, neuroscience
etc. This degree of freedom comes at a cost; performing the inference
is particularly demanding from both the computational and the
mathematical perspective. The class of methods that try to approximate
the posterior distribution in these cases are called likelihood-free
inference methods. Unfortunately, the algorithms deployed so far
permit the inference only at low-dimensional parametric spaces,
i.e.\ $\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms, small refers to dimensionality $D$ less than a hundred.}.

\subsubsection*{\textit{Goal of implicit models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data in the best manner i.e.\ generate
samples $M_r(\thetab^*)$ that are as close as possible to the observed
data $\data$. In our case, following the Bayesian approach, we treat
the parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

### 2021-07-09 latex_files/ch2/2_2_LFI_background.tex
added content in chapter 2

As already stated, in simulator-based models we cannot evaluate the
posterior $p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
following equation allows incorporating the simulator in the place of
the likelihood and forms the basis of all likelihood-free inference
approaches,

\begin{equation} \label{eq:likelihood}
  L(\thetab) =
  \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in B_{d,\epsilon}(\data))
\end{equation}
%
where $c_\epsilon$ is a proportionality factor dependend on
$\epsilon$ and $B_{d,\epsilon}(\data)$ is a region around $\data$.

Intuitively, equation~\ref{eq:likelihood} describes that the
likelihood of a specific parameter configuration $\thetab$ is equal to
the probability that the simulator will produce outputs infinitively
close to the observations $\data$, using this configuration.

For continuous cases, the probability of replicating the data
infinitely close becomes zero. In this scenario a relaxation must be
introduced by accepting simulated data that fall close to the data
i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
$\epsilon > 0$. The region can be
defined as $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$
where $d(\cdot, \cdot)$ can represent any valid distance.

Such relaxation, introduces the approximate likelihood,

\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = Pr(\yb \in \region(\data)), \text{where  } \epsilon > 0
\end{equation}

and consequently the approximate posterior,

\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab)
\end{equation}

### 2021-07-09 latex_files/ch2/2_3_OMC.tex
added content in chapter 2

--- 2021-07-08 latex_files/ch2/2_1_simulator_based_models.tex
@@
-As already stated at Chapter~\ref{sec:introduction}, in
-simulator-based models we cannot evaluate the posterior
-$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
-intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
-following equation allows incorporating the simulator in the place of
-the likelihood and forms the basis of all likelihood-free inference
-approaches,
-
-\begin{equation} \label{eq:likelihood}
-  L(\thetab) =
-  \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
-  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in B_{d,\epsilon}(\data))
-\end{equation}
-%
-where $c_\epsilon$ is a proportionality factor dependent on
-$\epsilon$, needed when
-$\Pr(M_r(\thetab) \in B_{d, \epsilon}(\data)) \rightarrow 0$, as
-$\epsilon \rightarrow 0$. Equation~\ref{eq:likelihood} describes that
-the likelihood of a specific parameter configuration $\thetab$ is
-proportional to the probability that the simulator will produce
-outputs equal to the observations, using this configuration.
-
-\subsubsection{Approximate Bayesian Computation (ABC) Rejection
-  Sampling}
-
-ABC rejection sampling is a modified version of the traditional
-rejection sampling method, for cases when the evaluation of the
-likelihood is intractable. In the typical rejection sampling, a sample
-obtained from the prior $\thetab \sim p(\thetab)$ gets accepted
-with probability $L(\thetab)/ \text{max}_{\thetab}
-L(\thetab)$. Though we cannot use this approach out-of-the-box
-(evaluating $L(\thetab)$ is impossible in our case), we can
-modify the method incorporating the simulator.
-
-In the discrete case scenario where $\Y_{\thetab}$ can take a finite
-set of values, the likelihood becomes
-$L(\thetab) = \Pr(\Y_{\thetab} = \data)$ and the posterior
-$p(\thetab|\data) \propto \Pr(\Y_{\thetab}=\data)p(\thetab)$; hence, we can
-sample from the prior $\thetab_i \sim p(\thetab)$, run the simulator
-$\yb_i = M_r(\thetab_i)$ and accept $\thetab_i$ only if
-$\yb_i = \data$.
-
-The method above becomes less useful as the finite set of
-$\Y_{\thetab}$ values grows larger, since the probability of
-accepting a sample becomes smaller. In the limit
-where the set becomes infinite (i.e.\ continuous case) the probability
-becomes zero. In order for the method to work in this set-up, a
-relaxation is introduced; we relax the acceptance criterion by letting
-$\yb_{i}$ lie in a larger set of points i.e.\
-$\yb_{i} \in \region(\data), \epsilon > 0$. The region can be
-defined as $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$
-where $d(\cdot, \cdot)$ can represent any valid distance. With this
-modification, the maintained samples follow the approximate posterior,
-
-\begin{equation} \label{eq:approx_posterior}
-  p_{d,\epsilon}(\thetab|\data) \propto Pr(\yb \in
-  \region(\data)) p(\thetab)
-\end{equation}
-
-\noindent
-This method is called Rejection ABC.
-
-\subsubsection{Summary Statistics}
-
-When $\yb \in \mathbb{R}^D$ lies in a high-dimensional space, generating
-samples inside $\region (\data)$ becomes rare even when $\epsilon$ is
-relatively large; this is the curse of dimensionality. As a
-representative example lets make the following hypothesis;
-
-\begin{itemize}
-\item $d$ is set to be the Euclidean distance, hence
-  $\region(\data) := \{ \yb: ||\yb - \data||_2^2 < \epsilon^2 \}$ is a
-  hyper-sphere with radius $\epsilon$ and volume $V_{hypersphere} = \frac{\pi^{D/2}}{\Gamma(D/2 + 1)} \epsilon^D$
-\item the prior $p(\thetab)$ is a uniform distribution in a hyper-cube with side of
-  length $2\epsilon$ and volume $V_{hypercube} = (2\epsilon)^D$
-\item the generative model is the identity function $\yb=f(\thetab)= \thetab $
-\end{itemize}
-
-\noindent
-The probability of drawing a sample inside the hypersphere equals the
-fraction of the volume of the hypersphere inscribed in the hypercube:
-
-\begin{equation}
-  Pr(\yb \in \region (\data))
-  = Pr(\thetab \in \region (\data))
-  = \frac{V_{hypersphere}}{V_{hypercube}}
-  = \frac{\pi^{D/2}}{2^D\Gamma(D/2 + 1)} \rightarrow 0, \quad \text{as} \quad D \rightarrow \infty
-\end{equation}
-
-\noindent
-We observe that the probability tends to $0$, independently of
-$\epsilon$; enlarging $\epsilon$ will not increase the acceptance
-rate. Intuitively, we can think that in high-dimensional spaces the
-volume of the hypercube concentrates at its corners. This generates
-the need for a mapping
-$T: \mathbb{R}^{D_1} \rightarrow \mathbb{R}^{D_2}$ where $D_1 > D_2$,
-for squeezing the dimensionality of the output. This
-dimensionality-reduction step that redefines the area as
-$\region(\data) := \{\yb: d(T(\yb), T(\data)) \leq \epsilon \}$ is
-called \textit{summary statistic} extraction, since the distance is
-not measured on the actual outputs, but on a summarisation (i.e.\
-lower-dimension representation) of them.
-
-\subsubsection{Approximations introduced so far}
-
-So far, we have introduced some approximations for inferring the
-posterior as
-$p_{d,\epsilon}(\thetab|\data) \propto Pr(\Y_{\thetab} \in
-\region(\data))p(\thetab)$ where
-$\region(\data) := \{\yb: d(T(\yb), T(\data)) < \epsilon \}$. These
-approximations introduce two different types of errors:
-
-\begin{itemize}
-\item $\epsilon$ is chosen to be \textit{big enough}, so that enough
-  samples are accepted. This modification leads to the approximate
-  posterior introduced in \eqref{eq:approx_posterior}
-\item $T$ introduces some loss of information, making possible a $\yb$
-  far away from $\data$ i.e.\ $\yb: d(\yb,\data)>\epsilon$, to enter
-  the acceptance region after the dimensionality reduction
-  $d(T(\yb), T(\data)) \leq \epsilon$
-\end{itemize}
-
-\noindent
-In the following sections, we will not use the summary statistics in
-our expressions for the notation not to clutter. One could understand
-it as absorbing the mapping $T(\cdot)$ inside the simulator. In any
-case, all the propositions that will be expressed in the following
-sections are valid with the use of summary statistics.
-
-\subsubsection{Optimisation Monte Carlo (OMC)}
-
 Before we define the likelihood approximation as introduced in the
@@
 
-\begin{gather} \label{eq:approx_likelihood}
+\begin{gather} \label{eq:approx_likelihood_omc}
   L_{d, \epsilon}(\thetab) =
@@
 
-\begin{equation} \label{eq:aprox_posterior}
+\begin{equation} \label{eq:approx_posterior_omc}
   p_{d,\epsilon}(\thetab|\data)
@@
 
-The posterior approximation in \eqref{eq:approx_posterior} does not
+The posterior approximation in \eqref{eq:approx_posterior_omc} does not
 provide any obvious way for drawing samples. In fact, the set
@@
 $L_{d,\epsilon}(\thetab_i)$ is computed with the expression
-\eqref{eq:approx_likelihood}. This approach has the same drawbacks as
+\eqref{eq:approx_likelihood_omc}. This approach has the same drawbacks as
 ABC rejection sampling; when the prior is wide or the dimensionality

### 2021-07-09 latex_files/ch2/2_4_ROMC.tex
added content in chapter 2

--- 2021-07-08 latex_files/ch2/2_2_ROMC.tex
@@
+\subsubsection*{\textit{Robust Optimisation Monte Carlo (ROMC) method}}
+
+The ROMC method~\cite{Ikonomov2019} is very a recent likelihood-free
+approach. Its fundamental idea is the transformation of the stochastic
+data generation process $M_r(\thetab)$ to a deterministic function
+$g_i(\thetab)$, by sampling the variables that produce the randomness
+$\vb_i \sim p(\V)$. Formally, in every stochastic process the
+randomness is influenced by a vector of random variables $\V$, whose
+state is unknown prior to the execution of the simulation; sampling in
+advance this state makes the procedure deterministic, i.e.
+$g_i(\thetab) = M_d(\thetab, \V=\vb_i)$. This approach was initially
+introduced by \cite{Meeds2015} as \textit{Optimisation Monte Carlo
+  (OMC)}. The ROMC extended this approach by resolving a fundamental
+failure-mode of OMC. The ROMC describes a methodology for
+approximating the posterior through a series of steps, without
+explicitly enforcing which algorithms must be utilised for each
+step\footnote{The implementation chooses a specific algorithm for each
+  task, but this choice has just a demonstrative value; any
+  appropriate algorithm can be used instead.}; in this sense, it can
+be perceived as a meta-algorithm.
+
 The simplifications introduced by OMC, although quite useful from a

### 2021-07-09 latex_files/ch2/2_5_ROMC_algorithmic.tex
added content in chapter 2

In this section, we will provide the algorithmic description of the
ROMC method; how to solve the optimisation problems using either the
gradient-based approach or the Bayesian optimisation alternative and
the construction of the bounding box. Afterwards, we will discuss the
advantages and disadvantages of each choice in terms of accuracy and
efficiency.

At a high-level, the ROMC method can be split into the training and
the inference part.

\subsubsection*{Training part}
\noindent
At the training (fitting) part, the goal is the estimation of the
proposal regions $q_i$. The tasks are (a) sampling the nuisance
variables $\vb_i \sim p(\vb)$ (b) defining the optimisation problems
$\min_{\thetab} \: g_i(\thetab)$ (c) obtaining $\thetab_i^*$ (d)
checking whether $d_i^* \leq \epsilon$ and (e) building the bounding
box for obtaining the proposal region $q_i$. If gradients are
available, using a gradient-based method is advised for obtaining
$\thetab_i^*$ much faster. Providing $\nabla_{\thetab} g_i$ in
closed-form provides an upgrade in both accuracy and efficiency; If
closed-form description is not available, approximate gradients with
finite-differences requires two evaluations of $g_i$ for
\textbf{every} parameter $\thetab_d$, which works adequately well for
low-dimensional problems. When gradients are not available or $g_i$ is
not differentiable, the Bayesian optimisation paradigm exists as an
alternative solution. In this scenario, the training part becomes
slower due to fitting of the surrogate model and the blind
optimisation steps. Nevertheless, the subsequent task of computing the
proposal region $q_i$ becomes faster since $\hat{d}_i$ can be used
instead of $g_i$; hence we avoid to run the simulator
$M_d(\thetab, \vb_i)$ for each query. The
algorithms~\ref{alg:training_GB} and~\ref{alg:training_GP} present the
above procedure.

\subsubsection*{Inference Part}
Performing the inference includes one or more of the following three
tasks; (a) evaluating the unnormalised posterior
$p_{d, \epsilon}(\thetab|\data)$ (b) sampling from the posterior
$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$ (c) computing an
expectation $E_{\thetab|\data}[h(\thetab)]$. Computing an expectation
can be done easily after weighted samples are obtained using the
equation~\ref{eq:expectation}, so we will not discuss it separately.

\noindent
Evaluating the unnormalised posterior requires solely the
deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
there is no need for solving the optimisation problems and building
the proposal regions. The evaluation requires iterating over all $g_i$
and evaluating the distance from the observed data. In contrast, using
the GP approach, the optimisation part should be performed first for
fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
indicator function on them. This provides an important speed-up,
especially when running the simulator is computationally
expensive. % The evaluation of the posterior is presented analytically
% in~\ref{alg:posterior_GB} and~\ref{alg:posterior_GP}.

\noindent
Sampling is performed by getting $n_2$ samples from each proposal
distribution $q_i$. For each sample $\thetab_{ij}$, the indicator
function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
checking if it lies inside the acceptance region. If so the
corresponding weight is computed as in \eqref{eq:sampling}. As before,
if a surrogate model $\hat{d}$ is available, it can be utilised for
evaluating the indicator function. At the sampling task, the
computational benefit of using the surrogate model is more valuable
compared to the evaluation of the posterior, because the indicator
function must be evaluated for a total of $n_1 \times n_2$ points.

\noindent
In summary, we can state that the choice of using a Bayesian
optimisation approach provides a significant speed-up in the inference
part with the cost of making the training part slower and a possible
approximation error. It is typical in many Machine-Learning use cases,
being able to provide enough time and computational resources for the
training phase, but asking for efficiency in the inference
part.

\begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
    \begin{algorithmic}[1]
      \For{$i \gets 1 \textrm{ to } n$}
        \State Obtain $\thetab_i^*$ using a Gradient Optimiser
        \If{$g_i(\thetab_i^*) > \epsilon$}
        \State{go to} 1
        \Else
        \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
        \EndIf      
      \EndFor
      \Return{$q_i, p(\theta), g_i(\theta)$}
    \end{algorithmic}
\end{algorithm}
\end{minipage}
\hfill
\begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
    \begin{algorithmic}[1]
      \For{$i \gets 1 \textrm{ to } n$}
        \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
        \If{$g_i(\thetab_i^*) > \epsilon$}
        \State{go to} 1
        \Else
        \State Approximate $H_i \approx \jac^T_i \jac_i$
        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
        \EndIf      
      \EndFor
      \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
    \end{algorithmic}
\end{algorithm}
\end{minipage}

\begin{algorithm}[!ht]
	\caption{Computation of the proposal distribution $q_i$; Needs, a model of distance $d$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
	\begin{algorithmic}[1]
	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
		\State $k \gets 0$
		\Repeat
        	\Repeat
                \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$}
        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
        	\State $k \gets k + 1$
    	\Until $k = K$
    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
	\EndFor
	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
	\end{algorithmic}
\end{algorithm}

% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Evaluate unnormalised posterior - Gradient approach. Requires $g_i(\theta), p(\theta)$}\label{alg:posterior_GB}
%     \begin{algorithmic}[1]
%       \State $k \leftarrow 0$
%         \For {$i \gets 1 \textrm{ to } n_1$}
%           \If {$g_i(\theta) > \epsilon$}
%             \State $k \leftarrow k + 1$
%           \EndIf
%           \EndFor
%       \Return{$kp(\theta)$}
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}
% \hfill
% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Evaluate unnormalised posterior - GP approach. Requires $\hat{d}_i(\theta), p(\theta)$}\label{alg:posterior_GP}
%     \begin{algorithmic}[1]
%       \State $k \leftarrow 0$
%         \For {$i \gets 1 \textrm{ to } n_1$}
%           \If {$d_i(\theta) > \epsilon$}
%             \State $k \leftarrow k + 1$
%           \EndIf
%           \EndFor
%       \Return{$kp(\theta)$}
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}

% \begin{minipage}{0.46\textwidth}
\begin{algorithm}[H]
    \centering
    \caption{Sampling. Requires a function of distance $(g_i(\theta)$ or $\hat{d}_i$ or $\hat{g}_i), p(\theta), q_i$}\label{alg:sampling_GB}
    \begin{algorithmic}[1]
      \For {$i \gets 1 \textrm{ to } n_1$}
      \For {$j \gets 1 \textrm{ to } n_2$}
          \State $\thetab_{ij} \sim q_i$
          \If {$g_i(\thetab_{ij}) > \epsilon$}
            \State Reject $\theta_{ij}$
          \Else {}
            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$
            \State Accept $\thetab_{ij}$, with weight $w_{ij}$
          \EndIf
      \EndFor
      \EndFor
    \end{algorithmic}
\end{algorithm}
% \end{minipage}
% \hfill
% \begin{minipage}{0.46\textwidth}
% \begin{algorithm}[H]
%     \centering
%     \caption{Sampling - GP approach. Requires $\hat{d}_i(\theta), p(\theta), q_i$}\label{alg:sampling_GP}
%     \begin{algorithmic}[1]
%       \For {$i \gets 1 \textrm{ to } n_1$}
%       \For {$j \gets 1 \textrm{ to } n_2$}
%           \State $\theta_{ij} \sim q_i$
%           \If {$\hat{d}_i(\theta_{ij}) > \epsilon$}
%             \State Reject $\theta_{ij}$
%           \Else {}
%             \State $w_{ij} = \frac{p(\theta_{ij})}{q(\theta_{ij})}$
%             \State Accept $\theta_{ij}$, with weight $w_{ij}$
%           \EndIf
%       \EndFor
%       \EndFor
%     \end{algorithmic}
% \end{algorithm}
% \end{minipage}

### 2021-07-09 latex_files/ch3/3_2_1_general_design.tex
added content in chapter 2

--- 2021-07-08 latex_files/ch3/3_1_1_general_design.tex
@@
 For illustrating the functionalities we choose as running example the
-following model, introduced by~\autocite{Ikonomov2019},
+following model, introduced by~\cite{Ikonomov2019},
 
@@
     \begin{center}
-\includegraphics[width=0.7\textwidth]{./Thesis/images/chapter3/example_gt.png}
+\includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
     \end{center}
@@
     \begin{center}
-\includegraphics[width=0.8\textwidth]{./Thesis/graphs/ROMC.png}
+\includegraphics[width=0.8\textwidth]{./latex_files/graphs/ROMC.png}
     \end{center}

### 2021-07-09 latex_files/ch3/3_2_2_training.tex
added content in chapter 2

--- 2021-07-08 latex_files/ch3/3_1_2_training.tex
@@
 
-\begin{enumerate}[label=(\roman*)]
+\begin{enumerate}
 \item \mintinline{python}{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
@@
     \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_theta_dist.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_region.png}
+      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
+      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region.png}
     \end{center}

### 2021-07-09 latex_files/ch3/3_2_3_inference.tex
added content in chapter 2

--- 2021-07-08 latex_files/ch3/3_1_3_inference.tex
@@
 
-\begin{enumerate}[label=(\roman*)]
+\begin{enumerate}
 \item \mintinline{python}{romc.sample(n2, seed=None)}
@@
     \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_marginal.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter3/example_region_samples.png}
+      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_marginal.png}
+      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
     \end{center}
@@
     \begin{center}
-      \includegraphics[width=0.75\textwidth]{./Thesis/images/chapter3/example_posterior.png}
+      \includegraphics[width=0.75\textwidth]{./latex_files/images/chapter3/example_posterior.png}
     \end{center}

### 2021-07-09 paper.tex
added content in chapter 2

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% Vasilis packages
\usepackage{amsmath,amssymb,bbold, xfrac}
\usepackage{xcolor}

% must be removed
\usepackage{multirow}
\usepackage{caption}

%% just to show content -> they should be replaced by the journals packages in the end
\usepackage{algorithm} 
\usepackage[noend]{algpseudocode}

%% just to show content -> they should be replaced by the journals packages in the end
% Python style code
\usepackage[finalizecache, cachedir=./minted_files]{minted}
\usemintedstyle{tango}
\usepackage{listings}
\usepackage{xcolor}
\newcommand{\pinline}[1]{\mintinline{python}{#1}}
\newminted{python}{framesep=2mm,
baselinestretch=1.2,
fontsize=\small,
frame=single
}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% Vasilis' custom commands
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\X}{\mathbf{x}}
\newcommand{\Z}{\mathbf{z}}
\newcommand{\V}{\mathbf{V}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\hess}{\mathbf{H}}
\newcommand{\Thetab}{\mathbf{\Theta}}

\newcommand{\vb}{\mathbf{v}}
\newcommand{\yb}{\mathbf{y}}
\newcommand{\xb}{\mathbf{x}}

\newcommand{\jac}{\mathbf{J}}
\newcommand{\hessian}{\mathbf{H}}
\newcommand{\thetab}{\boldsymbol{\theta}}
\newcommand{\thetabij}{\thetab_{ij}}
\newcommand{\thetabi}{\thetab_i}

\newcommand{\region}{B_{d,\epsilon}}
\newcommand{\indicator}[1]{\mathbb{1}_{#1}}
\newcommand{\regioni}{B_{d,\epsilon}^i}
\newcommand{\data}{\mathbf{y_0}}

\newcommand{\R}{\mathbb{R}}

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Achim Zeileis\\Universit\"at Innsbruck
   \And Second Author\\Plus Affiliation}
\Plainauthor{Achim Zeileis, Second Author}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
\Plaintitle{An Extendable Python Implementation of ROMC}
\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}

%% - \Abstract{} almost as usual
\Abstract{
  This short article illustrates how to write a manuscript for the
  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
  it is recommended to keep the {\LaTeX} code as simple as possible,
  i.e., avoid inclusion of packages/commands that are not necessary.
  For outlining the typical structure of a JSS article some brief text snippets
  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
  discussing count data regression in \proglang{R}. Editorial comments and
  instructions are marked by vertical bars.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Achim Zeileis\\
  Journal of Statistical Software\\
  \emph{and}\\
  Department of Statistics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  Universit\"atsstr.~15\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@R-project.org}\\
  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

%% See jss_template_article.tex for formatting instructions and macros to e.g. typset code
%% I only kept the informative sidebars as comments

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

% \section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
\section{Introduction} \label{sec:intro}

% @Vasilis: we can work on the intro when we have the basic structure of the article in place.

% \begin{leftbar}
% The introduction is in principle ``as usual''. However, it should usually embed
% both the implemented \emph{methods} and the \emph{software} into the respective
% relevant literature. For the latter both competing and complementary software
% should be discussed (within the same software environment and beyond), bringing
% out relative (dis)advantages. All software mentioned should be properly
% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
% \textsc{Bib}{\TeX}.)

% For writing about software JSS requires authors to use the markup
% \verb|\proglang{}| (programming languages and large programmable systems),
% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
% arguments, etc.). If there is such markup in (sub)section titles (as above), a
% plain text version has to be provided in the {\LaTeX} command as well. Below we
% also illustrate how abbrevations should be introduced and citation commands can
% be employed. See the {\LaTeX} code for more details.
% \end{leftbar}

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

% \begin{leftbar}
% As the synopsis above is a code listing that is not meant to be executed,
% one can use either the dedicated \verb|{Code}| environment or a simple
% \verb|{verbatim}| environment for this. Again, spaces before and after should be
% avoided.

% Finally, there might be a reference to a \verb|{table}| such as
% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
% column headers and captions in sentence style, and if possible avoiding vertical
% lines.
% \end{leftbar}

%% -- Illustrations ------------------------------------------------------------

%% - Virtually all JSS manuscripts list source code along with the generated
%%   output. The style files provide dedicated environments for this.
%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%%   or knitr using the render_sweave() hook - are used (without the need to
%%   load Sweave.sty).
%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
%% - The code input should use "the usual" command prompt in the respective
%%   software system.
%% - For R code, the prompt "R> " should be used with "+  " as the
%%   continuation prompt.
%% - Comments within the code chunks should be avoided - these should be made
%%   within the regular LaTeX text.

% \begin{leftbar}
% For code input and output, the style files provide dedicated environments.
% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
% hook. Please make sure that all code is properly spaced, e.g., using
% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
% use ``the usual'' command prompt in the respective software system. For
% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
% the continuation prompt. Generally, comments within the code chunks should be
% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
% before and after code input/output should be avoided (see above).
% \end{leftbar}

\section{Background}

\subsection{Likelihood-free inference and implicit models}
\input{latex_files/ch2/2_1_LFI_intro.tex}

\subsection{Background of Likehood-free inference methods}
\input{latex_files/ch2/2_2_LFI_background.tex}

\subsection{Optimisation Monte Carlo (OMC)}
\input{latex_files/ch2/2_3_OMC.tex}

\subsection{Robust optimisation Monte Carlo (ROMC)}
\input{latex_files/ch2/2_4_ROMC.tex}

\subsection{Algorithmic view of ROMC} % better titl
\input{latex_files/ch2/2_5_ROMC_algorithmic.tex}

\section{ROMC module in ELFI}  % better title needed

\subsection{Engine for likelihood-free inference (ELFI)}
\input{latex_files/ch3/3_1_ELFI.tex}

\subsection{Design Principles}
\input{latex_files/ch3/3_2_design_principles.tex}

\subsection{Extensibility} % better title
\input{latex_files/ch3/3_3_extensibility.tex}

\section{Example}
\input{latex_files/ch4/4_2_ma2.tex}

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}
lalala

%% -- Optional special unnumbered sections -------------------------------------

% \section*{Computational details}

% \begin{leftbar}
% If necessary or useful, information about certain computational details
% such as version numbers, operating systems, or compilers could be included
% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
% maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}

% The results in this paper were obtained using
% \proglang{R}~3.4.1 with the
% \pkg{MASS}~7.3.47 package. \proglang{R} itself
% and all packages used are available from the Comprehensive
% \proglang{R} Archive Network (CRAN) at
% \url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

% \newpage

% \begin{appendix}

% \section{More technical details} \label{app:technical}

% \begin{leftbar}
% Appendices can be included after the bibliography (with a page break). Each
% section within the appendix should have a proper section title (rather than
% just \emph{Appendix}).

% For more technical style details, please check out JSS's style FAQ at
% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
% which includes the following topics:
% \begin{itemize}
%   \item Title vs.\ sentence case.
%   \item Graphics formatting.
%   \item Naming conventions.
%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
%   \item Trouble shooting.
%   \item Many other potentially helpful details\dots
% \end{itemize}
% \end{leftbar}

% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

% \begin{leftbar}
% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
% formats of author-year citations and allow to include additional details (e.g.,
% pages, chapters, \dots) in brackets. In case you are not familiar with these
% commands see the JSS style FAQ for details.

% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
% when acquiring the entries automatically from mixed online sources. However,
% it is important that informations are complete and presented in a consistent
% style to avoid confusions. JSS requires the following format.
% \begin{itemize}
%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
%     be used in the references.
%   \item Titles should be in title case.
%   \item Journal titles should not be abbreviated and in title case.
%   \item DOIs should be included where available.
%   \item Software should be properly cited as well. For \proglang{R} packages
%     \code{citation("pkgname")} typically provides a good starting point.
% \end{itemize}
% \end{leftbar}

% \end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2021-07-17 latex_files/ch1/1_1_motivation.tex
updated most content in chapter 2,3

@@
-\subsubsection*{\textit{Explanation of simulation-based models}}
-
-A simulator-based model is a parameterised stochastic data generating
-mechanism \cite{Gutmann2016}. The key characteristic of these
-models is that although we can sample (simulate) data points, we
-cannot evaluate the likelihood of a specific set of observations
+An implicit or simulator-based model is a parameterised stochastic
+data generating mechanism \cite{Gutmann2016}. The key characteristic
+of these models is that although we can sample (simulate) data points,
+we cannot evaluate the likelihood of a specific set of observations
 $\data$. Formally, a simulator-based model is described as a
@@
   \textit{random} simulator. In the next chapters we will introduce
-  $M_d$ witch stands for the \textit{deterministic} simulator.} that
+  $M_d$ indicating the \textit{deterministic} simulator.} that
 given a set of parameters $\thetab$, produces samples $\yb$ in a
@@
 modelling freedom they provide; any physical process that can be
-conceptualised as a computer program of finite (deterministic or
-stochastic) steps can be modelled as a simulator-based model without
-any compromise. The modelling freedom includes any amount of hidden
-(unobserved) internal variables or logic-based decisions. As always,
-this degree of freedom comes at a cost; performing the inference is
-particularly demanding from both computational and mathematical
-perspective. Unfortunately, the algorithms deployed so far permit the
-inference only at low-dimensionality parametric spaces, i.e.\
-$\thetab \in \mathbb{R}^D$ where $D$ is small.
+conceptualised as a computer program of finite steps can be modelled
+as a simulator-based model without any compromise. The modelling
+freedom includes any amount of hidden (unobserved) internal variables
+or rule-based decisions. Hence, implicit models are often used to
+model physical phenomena in many fiels such as biology, neuroscience
+etc. This degree of freedom comes at a cost; performing the inference
+is particularly demanding from both the computational and the
+mathematical perspective. The class of methods that try to approximate
+the posterior distribution in these cases are called likelihood-free
+inference methods. Unfortunately, the algorithms deployed so far
+permit the inference only at low-dimensional parametric spaces,
+i.e.\ $\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms, small refers to dimensionality $D$ less than a hundred.}.
 
-\subsubsection*{\textit{Example}}
-
-For underlying the importance of simulator-based models, let us use
-the tuberculosis disease spread example as described in
-\cite{Tanaka2006}. An overview of the disease spread model is
-presented at figure~\ref{fig:tuberculosis_model}. At each stage one of
-the following \textit{unobserved} events may happen; (a) the
-transmission of a specific haplotype to a new host, (b) the mutation
-of an existent haplotype or (c) the exclusion of an infectious host
-(recovers/dies) from the population. The random process, which stops
-when $m$ infectious hosts are reached\footnote{We suppose that the
-  unaffected population is infinite, so a new host can always be added
-  until we reach $m$ simultaneous hosts.}, can be parameterised by the
-transmission rate $\alpha$, the mutation rate $\tau$ and the exclusion
-rate $\delta$, creating a $3D$-parametric space
-$\thetab = (\alpha, \tau, \delta)$. The outcome of the process is a
-variable-size tuple $\yb_{\thetab}$, containing the population
-contaminated by each different haplotype, as described in
-figure~\ref{fig:tuberculosis_model}. Let's say that the disease has
-been spread in a real population and when $m$ hosts were contaminated
-simultaneously, the vector with the infectious populations has been
-measured to be $\data$. We would like to discover the parameters
-$\thetab = (\alpha, \tau, \delta)$ that generated the spreading
-process and led to the specific outcome $\data$. Computing
-$p(\yb=\data|\thetab)$ requires tracking all tree-paths that could
-generate the specific tuple; such exhaustive enumeration becomes
-intractable when $m$ grows larger, as in real-case scenarios. In
-figure~\ref{fig:tuberculosis_model} we can observe that a transmission
-followed by a recovery/death creates a loop, reinstating the process
-to the previous step, which also complicates the exhaustive
-enumeration. Hence, representing the process with a simulator-based
-model\footnote{which is simple and efficient} and performing
-likelihood-free inference is the recommended solution.
-
-\begin{figure}[!ht]
-    \begin{center}
-      \includegraphics[width=0.75\textwidth]{./latex_files/images/chapter1/tuber_model_1.png}
-    \end{center}
-    \caption[The tuberculosis spreading process.]{Depiction of a random example from the tuberculosis
-      spreading process. The image has been taken from
-      \cite{Lintusaari2017}.}
-    \label{fig:tuberculosis_model}
-\end{figure}
-
-\subsubsection*{\textit{Goal of Simulation-Based Models}}
+\subsubsection*{\textit{Goal of implicit models}}
 
 As in most Machine Learning (ML) concepts, the fundamental goal is the
-derivation of one(many) parameter configuration(s) $\thetab^*$ that
-\textit{describe} the data best i.e.\ generate samples
-$M_r(\thetab^*)$ that are as close as possible to the observed data
-$\data$. In our case, following the approach of Bayesian ML, we treat
+derivation of one (or many) parameter configuration(s) $\thetab^*$
+that \textit{describe} the data best i.e.\ generate
+samples $M_r(\thetab^*)$ that are as close as possible to the observed
+data $\data$. In our case, following the Bayesian approach, we treat
 the parameters of interest $\thetab$ as random variables and we try to
 \textit{infer} a posterior distribution $p(\thetab|\data)$ on them.
-
-\subsubsection*{\textit{Robust Optimisation Monte Carlo (ROMC) method}}
-
-The ROMC method~\cite{Ikonomov2019} is very a recent
-likelihood-free approach. Its fundamental idea is the transformation
-of the stochastic data generation process $M_r(\thetab)$ to a
-deterministic mapping $g_i(\thetab)$, by sampling the variables that
-produce the randomness $\vb_i \sim p(\V)$. Formally, in every
-stochastic process the randomness is influenced by a vector of random
-variables $\V$, whose state is unknown prior to the execution of the
-simulation; sampling the state makes the procedure deterministic,
-namely $g_i(\thetab) = M_d(\thetab, \V=\vb_i)$. This approach
-initially introduced at \cite{Meeds2015} with the title
-\textit{Optimisation Monte Carlo (OMC)}. The ROMC extended this
-approach by resolving a fundamental failure-mode of OMC\@. The ROMC
-describes a methodology for approximating the posterior through a
-series of steps, without explicitly enforcing which algorithms must be
-utilised for each step\footnote{The implementation chooses a specific
-  algorithm for each task, but this choice has just a demonstrative
-  value; any appropriate algorithm can be used instead.}; in this
-sense, it can be perceived as a meta-algorithm.
-
-\subsubsection*{\textit{Implementation}}
-
-The most important contribution of this work is the implementation of
-the ROMC method in the Python package Engine for Likelihood-Free
-Inference (ELFI) \cite{1708.00707}. Since the method has been
-published quite recently, it has not been implemented by now in any ML
-software. This work attempts to provide the research community with a
-robust and extensible implementation for further experimentation.

### 2021-07-17 latex_files/ch3/3_1_ELFI.tex
updated most content in chapter 2,3

\pkg{Engine for Likelihood-Free Inference (ELFI)}\footnote{Extended
  documentation can be found
  \href{https://elfi.readthedocs.io/en/latest/}{here}}~\cite{1708.00707}
is a \proglang{Python} package dedicated to Likelihood-Free Inference
(LFI). We implemented ROMC at \pkg{ELFI} since it provides convenient
modules for all the fundamental components of a probabilistic model
(e.g. prior, simulator, summaries etc.). Furthermore, \pkg{ELFI}
already supports some recently proposed likelihood-free inference
methods, making it straightforward to perform comparisons between
methods. The implementation was officially added to \pkg{ELFI} at
version \pkg{v0.8.0}.

% \subsubsection{Modelling at ELFI}
% \label{sec:modelling}

\pkg{ELFI} models the probabilistic model as a Directed Acyclic Graph
(DAG). This functionality is based on the package \pkg{NetworkX},
which supports general-purpose graphs. In most cases, the structure of
a likelihood-free model follows the pattern of figure~\ref{fig:elfi};
some edges connect the prior distributions to the simulator, the
simulator is connected to the summary statistics that, in turn, lead
to the output node. Samples can be obtained from all nodes through
sequential sampling. \pkg{ELFI} automatically considers as parameters
of interest, i.e. those we try to infer a posterior distribution,
as the ones  defined through the \code{elfi.Prior} class. Finally, the
observations are passed to the appropriate node through the argument
\code{observed}.

\begin{figure}[h]
    \begin{center}
      \includegraphics[width=0.8\textwidth]{./latex_files/images/chapter2/elfi.png}
    \end{center}
    \caption[Baseline example for creating an \pkg{ELFI} model]{Baseline example for creating an \pkg{ELFI} model. Image taken from \cite{1708.00707}}
    \label{fig:elfi}
\end{figure}

% \subsubsection{Inference at ELFI}
% \label{sec:inference-methods}

All inference methods at \pkg{ELFI} share the following rules;

\begin{itemize}
\item they follow the signature \code{elfi.<Class name>(<output node>,
    *arg)}; the initial argument is the output node model and the rest
  of the arguments are hyper-parameters of the method.
\item they provide a central sampling functionality, which in most
  cases is named \\ \code{<method_name>.sample()}.
\end{itemize}

### 2021-07-17 latex_files/ch3/3_2_1_general_design.tex
updated most content in chapter 2,3

@@
-In figure \ref{fig:romc_overview} we present an overview of our
-implementation; one may interpret figure~\ref{fig:romc_overview} as a
-depiction of the main class of our implementation, called
-\pinline{ROMC}, while the entities inside the green and blue
-ellipses are the main functions of the class. Following Python's
-naming principles, the methods starting with an underscore (green
-ellipses) represent internal (private) functions and are not meant to
-be used by a user, whereas the rest of the methods (blue ellipses) are
-the functionalities the user interacts with. As mentioned before, the
-implementation favours extensibility; the building blocks that compose
-the method have been designed in a modular fashion so that a
-practitioner may replace them without the method to collapse.
+The overview of our implementation is illustrated in
+figure~\ref{fig:romc_overview}; one may interpret
+figure~\ref{fig:romc_overview} as a depiction of the main class and
+the ellipses are the main routines of the class. Following
+\proglang{Python}'s naming principles, the methods starting with an
+underscore (green ellipses) represent internal (private) functions,
+whereas the rest (blue ellipses) are the public methods that the user
+interacts with.
 
-Figure \ref{fig:romc_overview} groups the ROMC implementation into the
-training, the inference and the evaluation part. The training part includes all the
-steps until the computation of the proposal regions; sampling the
-nuisance variables, defining the optimisation problems, solving them,
+Figure~\ref{fig:romc_overview} splits ROMC into the training, the
+inference and the evaluation part. The training part includes all the
+steps until the computation of the proposal regions i.e. sampling the
+nuisance variables, defining the optimization problems, solving them,
 constructing the regions and fitting local surrogate models. The
-inference part comprises of evaluating the unnormalised posterior (and
-the normalised one, in low-dimensional cases), sampling and computing
-an expectation. Moreover, the ROMC implementation provides some
-utilities for inspecting the training process, such as plotting the
-histogram of the distances
-$d^*_i = g_i(\theta_i^*), \: \forall i \in \{1, \ldots, n_1 \}$ after
-solving the optimisation problems and visualising the constructed
-bounding box\footnote{if the parametric space is up to $2D$}. Finally,
-two functionalities for evaluating the inference are implemented; (a)
-computing the Effective Sample Size (ESS) of the weighted samples and
-(b) measuring the divergence between the approximate posterior the
-ground-truth, if the latter is available.\footnote{Normally, the
-  ground-truth posterior is not available; However, this functionality
-  is useful in cases where the posterior can be computed numerically
-  or with an alternative method (i.e.\ ABC Rejection Sampling) and we
-  would like to measure the discrepancy between the two
-  approximations.}
+inference part comprises of evaluating the unnormalized posterior (and
+the normalized when is possible), sampling and computing an
+expectation. From an algorithmic point of view, the training and the
+inference parts are \emph{isolated}. We also provide some utilities
+for inspecting the training process, such as plotting the histogram of
+the final distances or visualizing the constructed bounding
+boxes. Finally, in the evaluation part, we provide two methods for
+evaluating the inference; (a) computing the Effective Sample Size
+(ESS) of the samples and (b) measuring the divergence between the
+approximate posterior the ground-truth, if the latter is
+available.\footnote{Normally, the ground-truth posterior is not
+  available; However, this functionality is useful in cases where the
+  posterior can be computed numerically or with an alternative method
+  (i.e.\ ABC Rejection Sampling), and we would like to measure the
+  discrepancy between the two approximations.}
 
-\subsubsection*{Parallelising the processes}
+\begin{figure}[H]
+    \begin{center}
+\includegraphics[width=0.8\textwidth]{./latex_files/graphs/ROMC.png}
+    \end{center}
+    \caption[Overview of the ROMC implementation.]{Overview of the
+ROMC implementation. The training part follows a sequential pattern;
+the functions in the green ellipses must be called in a sequential
+fashion for completing the training part and define the posterior
+distribution. The functions in blue ellipses are the functionalities
+provided to the user.}
+    \label{fig:romc_overview}
+\end{figure}
 
-As stated above, the most critical advantage of the ROMC method is that it can be fully parallelised.
-In our implementation, we have exploited this fact by implementing a parallel version in the following tasks; (a) solving the optimisation problems, (b) constructing bounding box regions, (c) sampling and (d) evaluating the posterior.
-Parallelism has been achieved using the package \pinline{multiprocessing}.
-The specific package enables concurrency, using subprocesses instead of threads, hence side-stepping the Global Interpreter (GIL).
-In our implementation we use the \pinline{Pool} object, which offers a convenient means of parallelising the execution of a function across multiple input values (data parallelism).
-For activating the parallel version of the algorithms, the user has to just pass the argument \pinline{parallelize=True} at the initialisation of the \pinline{ROMC} method.
+\subsubsection*{Parallelizing ROMC method}
+
+ROMC has the significant advantage of being fully parallelizable. We
+exploit this fact by implementing a parallel version of all the major
+components of the method; (a) solving the optimization problems, (b)
+constructing bounding box regions, (c) sampling and (d) evaluating the
+posterior. We parallelize these processes using the built-in
+\proglang{Python} package \pkg{multiprocessing}. The specific package
+enables concurrency, using subprocesses instead of threads, for
+side-stepping the Global Interpreter (GIL). For activating the
+parallel version of the algorithm, the user simpy has to pass the
+argument \code{parallelize=True} when instantiating \code{ROMC}.
+
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> elfi.ROMC(<output_node>, parallelize=True)
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 
-For illustrating the functionalities we choose as running example the
-following model, introduced by~\cite{Ikonomov2019},
+For illustrating the functionalities we will use the following running
+example introduced by~\cite{Ikonomov2019},
 
-\begin{gather} \label{eq:1D_example} p(\theta) =
-\mathcal{U}(\theta;-2.5,2.5)\\ p(y|\theta) = \left\{
+\begin{gather} \label{eq:1D_example}
+  p(\theta) = \mathcal{U}(\theta;-2.5,2.5)\\ \label{eq:1D_example_eq_2}
+  p(y|\theta) = \left\{
     \begin{array}{ll} \theta^4 + u & \mbox{if } \theta \in [-0.5, 0.5]
 \\ |\theta| - c + u & \mbox{otherwise}
-    \end{array} \right.\\ u \sim \mathcal{N}(0,1)
+    \end{array} \right.\\ 
+  u \sim \mathcal{N}(0,1)
 \end{gather}
@@
 
-In the model \eqref{eq:1D_example}, the prior is the uniform
-distribution in the range $[-2.5, 2.5]$ and the likelihood a Gaussian
-distribution. There is only one observation $y_0 = 0$. The inference
-in this particular example can be performed quite easily, without
-incorporating a likelihood-free inference approach. We can exploit this
-fact for validating the accuracy of our implementation. The
-ground-truth posterior, approximated computationally, is shown in
-figure \ref{fig:example_gt}.
+The prior is a uniform distribution in the range $[-2.5, 2.5]$ and the
+likelihood is defined at~\ref{eq:1D_example_eq_2}. There is only one
+observation $y_0 = 0$. The inference in this particular example can be
+performed quite easily without using a likelihood-free inference
+approach. We can exploit this fact for validating the accuracy of our
+implementation. The ground-truth posterior approximated
+computationally is shown in figure~\ref{fig:example_gt}.
 
-\begin{figure}[h]
+\begin{figure}[!h]
     \begin{center}
@@
 
-\subsubsection*{ELFI code for modelling the example}
+In the following code snippet, we code the model at \pkg{ELFI}.
 
-In the following code snippet, we code the model at \textit{ELFI} and
-we initialise the ROMC inference method. We observe that the
-initialisation of the ROMC inference method is quite intuitive; we
-just pass the final (distance) node of the simulator as argument, as
-in all $\textit{ELFI}$ inference methods. The argument
-\pinline{bounds}, although optional, is important for many
-functionalities (e.g. approximating the partition function, setting
-the bounds of the Bayesian optimisation etc.) so it is recommended to
-be passed.
-
-\begin{pythoncode}
+\begin{Code}
+------------------------------ python snippet ------------------------------
   import elfi import scipy.stats as ss
@@
   romc = elfi.ROMC(d, bounds=bounds, parallelize=parallelize)
-\end{pythoncode}
-
-\begin{figure}[!ht]
-    \begin{center}
-\includegraphics[width=0.8\textwidth]{./latex_files/graphs/ROMC.png}
-    \end{center}
-    \caption[Overview of the ROMC implementation.]{Overview of the
-ROMC implementation. The training part follows a sequential pattern;
-the functions in the green ellipses must be called in a sequential
-fashion for completing the training part and define the posterior
-distribution. The functions in blue ellipses are the functionalities
-provided to the user.}
-    \label{fig:romc_overview}
-\end{figure}
+----------------------------------------------------------------------------    
+\end{Code}

### 2021-07-17 latex_files/ch3/3_2_2_training.tex
updated most content in chapter 2,3

@@
-The training part contains the following 6 functionalities:
+The training part contains six methods.
 
-\begin{enumerate}
-\item \mintinline{python}{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
-\item \mintinline{python}{romc.estimate_regions(eps_filter,}
-
-      \mintinline{python}{                      use_surrogate=None, region_args=None,}
-
-      \mintinline{python}{                      fit_models=False, fit_models_args=None,}
-
-      \mintinline{python}{                      eps_region=None, eps_cutoff=None)}
-
-    \item \mintinline{python}{romc.fit_posterior(n1, eps_filter, use_bo=False, optimizer_args=None,}
-
-          \mintinline{python}{                   seed=None, use_surrogate=None, region_args=None,}
-
-          \mintinline{python}{                   fit_models=False, fit_models_args=None,}
-
-          \mintinline{python}{                   eps_region=None, eps_cutoff=None)}
-
-\item \mintinline{python}{romc.distance_hist(savefig=False, **kwargs)}
-\item \mintinline{python}{romc.visualize_region(i, savefig=False)}
-\item \mintinline{python}{romc.compute_eps(quantile)}
-\end{enumerate}
-
-\subsubsection*{Function (i): Define and solve the optimisation problems}
-
-\pinline{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.solve_problems(n1,
+                        use_bo=False,
+                        optimizer_args=None,
+                        seed=None)
+----------------------------------------------------------------------------                    
+\end{Code}
 
 \noindent
-This routine is responsible for (a) drawing the nuisance variables,
-(b) defining the optimisation problems and (c) solving them using either a
-gradient-based optimiser or Bayesian optimisation. The aforementioned
-tasks are done in a sequential fashion, as show in
-figure~\ref{fig:romc_overview}. The definition of the optimisation
-problems is performed by drawing $n_1$ integer numbers from a discrete
-uniform distribution $u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each
-integer $u_i$ is the seed used in ELFI's random simulator. Hence from
-an algorithmic point-of-view drawing the state of all random
-variables $\vb_i$ as described in the previous chapter, traces back to
-just setting the seed that initialises the state of the pseudo-random
-generator, before asking a sample from the simulator.
+This method (a) draws the nuisance variables, (b) defines the
+optimisation problems and (c) solves them using either a
+gradient-based optimizer or Bayesian optimization. These tasks are
+compleed sequentially, as shown in figure~\ref{fig:romc_overview}. The
+definition of the optimisation problems is performed by drawing
+\code{n_1} integer numbers from a discrete uniform distribution
+$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
+used in \pkg{ELFI}'s random simulator. The seed $u_i$ is the
+computational analogous of the random variables $\vb_i$, that were
+defined in the previous chapter; freezing the \code{seed} absorbs all
+the randomness of the simulator transforming it to a deterministic
+function. Setting the argument \code{use_bo=True}, chooses the
+Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
+case, apart from obtaining the optimal points $\thetab_i^*$, we also
+fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
+scenario, in the following steps, $\hat{d}_i(\thetab)$ will replace
+$g_i(\thetab)$ as the distance function.
 
-Finally, passing an integer number as the argument \pinline{seed}
-absorbs all the randomness of the optimisation part (e.g.\ drawing
-initial points for the optimisation procedure), making the whole
-process reproducible and deterministic.
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.distance_hist(**kwargs)
+----------------------------------------------------------------------------  
+\end{Code}
 
-Setting the argument \pinline{use_bo=True}, chooses the Bayesian
-Optimisation scheme for obtaining $\thetab_i^*$. In this case, apart
-from obtaining the optimal points $\thetab_i^*$, we also fit a Gaussian
-Process (GP) as surrogate model $\hat{d}_i$. In the following steps,
-$\hat{d}_i(\thetab)$ will replace $g_i(\thetab)$ when calling the
-indicator function.
+\noindent
+This function helps the user decide which threshold $\epsilon$ to
+use. It plots a histogram of the distances at the optimal point
+$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $d_i^*$ in case
+\code{use_bo=True}. The function accepts all keyword arguments and
+forwards them to the underlying \code{matplotlib.hist()} function; in
+this way the user may customise some properties of the histogram, such
+as the number of bins or the range of values.
 
-\subsubsection*{Function (ii): Construct bounding boxes and fit local surrogate models}
+\begin{Code}
+---------------------------------- python ----------------------------------  
+>>> romc.estimate_regions(eps_filter,
+                          use_surrogate=None,
+                          region_args=None,
+                          fit_models=False,
+                          fit_models_args=None,
+                          eps_region=None,
+                          eps_cutoff=None)
+----------------------------------------------------------------------------
+\end{Code}
 
-\mintinline{python}{romc.estimate_regions(eps_filter,}
-
-      \mintinline{python}{                      use_surrogate=None, region_args=None,}
-
-      \mintinline{python}{                      fit_models=False, fit_models_args=None,}
-
-      \mintinline{python}{                      eps_region=None, eps_cutoff=None)}
-\vspace{5mm}
-
-This routine constructs the bounding boxes around the optimal points
+\noindent
+This method constructs a bounding box around the optimal point
 $\thetab_i^* : i = 1, 2, \ldots, n_1$ following
@@
 eigenvectores are computed using the function
-\pinline{numpy.linalg.eig()} that calls, under the hood, the
-\pinline{_geev LAPACK}. A check is performed so that the matrix
+\code{numpy.linalg.eig()}. A check is performed so that the matrix
 $\hess_i$ is not singular; if this is the case, the eigenvectors are
@@
 $\{ \mathbf{e_1} = (1, 0, \ldots), \mathbf{e_2} = (0,1,0,\ldots),
-\text{etc} \}$. Afterwards, the limits are obtained by repeteadly querying
-the distance function ($g_i(\thetab)$ or $\hat{d}(\thetab)$) along the
-search directions. In section \ref{subsec:developers}, we provide some
-details regarding the way the bounding box is defined as a class and
-sampling is performed on it.
+\text{etc} \}$. The limits of the bounding box are obtained by
+repeteadly querying the distance function ($g_i(\thetab)$ or
+$\hat{d}(\thetab)$) along the search directions given by $\hess_i$.
 
-\subsubsection*{Function (iii): Perform all training steps in a single call}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.fit_posterior(n1,
+                       eps_filter,
+                       use_bo=False,
+                       optimizer_args=None,
+                       seed=None,
+                       use_surrogate=None,
+                       region_args=None,
+                       fit_models=False,
+                       fit_models_args=None,
+                       eps_region=None,
+                       eps_cutoff=None)
+----------------------------------------------------------------------------
+\end{Code}
 
-\mintinline{python}{romc.fit_posterior(n1, eps_filter, use_bo=False, optimizer_args=None,}
+\noindent
+The function \code{fit_posterior} is a syntactic sugar for applying
+\code{solve_problems} and \code{estimate_regions} into a single
+step.
 
-          \mintinline{python}{                seed=None, use_surrogate=None, region_args=None,}
+% merges all steps for constructing the bounding box into
+% a single command. If the user doesn't want to manually inspect the
+% histogram of the distances before deciding where to set the threshold
+% $\epsilon$, he may call \code{romc.fit_posterior()} and the whole
+% training part will be done end-to-end. There are two alternatives for
+% setting the threshold $\epsilon$; the first is to set to a specific
+% value blindly, and the second is to set at as a specific quantile of
+% the histogram of distances. In the second scenario the
+% \code{quantile} argument must be set to a floating number in the
+% range $[0,1]$ and \code{eps='auto'}.
 
-          \mintinline{python}{                fit_models=False, fit_models_args=None,}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.visualize_region(i)
+----------------------------------------------------------------------------
+\end{Code}
 
-          \mintinline{python}{                eps_region=None, eps_cutoff=None)}
-
-\vspace{5mm}
 \noindent
 
-This function merges all steps for constructing the bounding box into
-a single command. If the user doesn't want to manually inspect the
-histogram of the distances before deciding where to set the threshold
-$\epsilon$, he may call \pinline{romc.fit_posterior()} and the whole
-training part will be done end-to-end. There are two alternatives for
-setting the threshold $\epsilon$; the first is to set to a specific
-value blindly and the second is to set at as a specific quantile of
-the histogram of distances. In the second scenario the
-\pinline{quantile} argument must be set to a floating number in the
-range $[0,1]$ and \pinline{eps='auto'}.
-
-\subsubsection*{Function (iv): Plot the histogramm of the optimal points}
-
-\pinline{romc.distance_hist(**kwargs)}
-\vspace{5mm}
-\noindent
-
-This function can serve as an intermediate step of manual inspection,
-for helping the user choose which threshold $\epsilon$ to use. It
-plots a histogram of the distances at the optimal point
-$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or
-$d_i^*$ in case \pinline{use_bo=True}. The function accepts all
-keyword arguments and forwards them to the underlying
-\pinline{matplotlib.hist()} function; in this way the user may
-customise some properties of the histogram, such as the number of bins
-or the range of values.
-
-\subsubsection*{Function (v): Plot the acceptance region of the objective functions}
-
-\pinline{romc.visualize_region(i)}
-\vspace{5mm}
-\noindent
-
-It can be used as an inspection utility for cases where the parametric
-space is up to two dimensional. The argument $\pinline{i}$ is the
+It can be used for plotting the bounding box around the optimal point,
+when the parameter space is $1D$ or $2D$. The argument \code{i} is the
 index of the corresponding optimization problem i.e.\ $i<n_1$.
 
-\subsubsection*{Function (vi): Compute $\epsilon$ automatically based on the distribution of $d^*$}
-
-\pinline{romc.compute_eps(quantile)}
-\vspace{5mm}
-\noindent
+\begin{Code}
+---------------------------------- python ----------------------------------  
+>>> romc.compute_eps(quantile)
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 of the threshold $\epsilon$, e.g.\
-\pinline{eps=romc.compute_eps(quantile=0.9)}.
+\code{eps=romc.compute_eps(quantile=0.9)}.
 
@@
 
-Here we will illustrate the aforementioned functionalities using the
-simple 1D example we set up in the previous section. The following
-code snippet performs the training part in ELFI.
+In the following snippet, we put together the routines described above
+to perform the training part at our running example.
 
-\begin{pythoncode}
+\begin{Code}
+------------------------------ python snippet ------------------------------  
   n1 = 500 # number of optimisation problems
   seed = 21 # seed for solving the optimisation problems
-  eps = .75 # threshold for bounding box
+  eps = .75 
   use_bo = False # set to True for switching to Bayesian optimisation
@@
   romc.solve_problems(n1=n1, seed=seed, use_bo=use_bo)
-  romc.theta_hist(bins=100)
-  romc.estimate_regions(eps=eps)
-  romc.visualize_region(i=1)
+  romc.theta_hist(bins=100) # plot hist to decide which eps to use
+
+  eps = .75 # threshold for the bounding box based on histogram inspection
+  romc.estimate_regions(eps=eps) # build the bounding boxes
+
+  romc.visualize_region(i=1) # for inspecting visually the bounding box
 
@@
   # romc.fit_posterior(n1=n1, eps=eps, use_bo=use_bo, seed=seed)
-\end{pythoncode}
+----------------------------------------------------------------------------  
+\end{Code}
 
-As stated before, switching to the Bayesian optimisation scheme needs
-nothing more the setting the argument \pinline{use_bo=True}; all the
-following command remain unchanged. In figure
-\ref{fig:example_training_hist} we illustrate the distribution of the
-distances obtained and the acceptance area of the first optimisation
-problem. We observe that most optimal points produce almost zero
-distance.
+In figure \ref{fig:example_training_hist} we illustrate the
+distribution of the distances obtained and the acceptance area of the
+first optimisation problem. We observe that most optimal points
+produce almost zero distance.
 
@@
     \end{center}
-  \caption[Histogram of distances at the 1D example.]{Histogram of distances and visualisation of a specific region.}
+    \caption[Histogram of distances at the 1D example.]{Histogram of
+      distances and visualisation of a specific region.}
     \label{fig:example_training_hist}

### 2021-07-17 latex_files/ch3/3_2_3_inference.tex
updated most content in chapter 2,3

@@
-The inference part contains the 4 following functionalities:
+The inference part contains in four methods.
 
-\begin{enumerate}
-\item \mintinline{python}{romc.sample(n2, seed=None)}
-\item \mintinline{python}{romc.compute_expectation(h)}  
-\item \mintinline{python}{romc.eval_unnorm_posterior(theta)}
-\item \mintinline{python}{romc.eval_posterior(theta)}
-\end{enumerate}
-
-\subsubsection*{Function (i): Perform weighted sampling}
-
-\mintinline{python}{romc.sample(n2)}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.sample(n2)
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 remained after filtering\footnote{From the $n_1$ optimisation
-  problems, only the ones with $g_i(\thetab_*) < \epsilon$ are kept
-  for building a bounding box}. The samples are drawn from a uniform
-distribution $q_i$ defined over the corresponding bounding box and the
-weight $w_i$ is computed as in equation~\eqref{eq:sampling}. The
-function stores an \pinline{elfi.Result} object as
-\pinline{romc.result} attribute. The \pinline{elfi.Result} provides
-some usefull functionalities for inspecting the obtained samples e.g.\
-\pinline{romc.result.summary()} prints the number of the obtained
-samples and their mean. A complete overview of these functionalities is
-provided in ELFI's
-\href{https://elfi.readthedocs.io/en/latest/api.html#elfi.methods.results.Sample}{official
-  documentation}.
+  problems, only the ones with $g_i(\thetab_*) < \epsilon$ are
+  maintained for building a bounding box}. The samples are drawn from
+a uniform distribution $q_i$ defined over the corresponding bounding
+box and the weight $w_i$ is computed as in
+equation~\eqref{eq:sampling}.
 
-\subsubsection*{Function (ii): Compute an expectation}
-
-\mintinline{python}{romc.compute_expectation(h)}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.compute_expectation(h)
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 $E_{p(\thetab|\data)}[h(\thetab)]$ using
-expression~\eqref{eq:expectation}. The argument \pinline{h} can be
-any python \pinline{Callable}.
+expression~\eqref{eq:expectation}. The argument \code{h} can be
+any python \code{Callable}.
 
-\subsubsection*{Function (iii): Evaluate the unnormalised posterior}
-\mintinline{python}{romc.eval_unorm_posterior(theta, eps_cutoff=False)}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.eval_unorm_posterior(theta, eps_cutoff=False)
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 
-\subsubsection*{Function (iv): Evaluate the normalised posterior}
-\mintinline{python}{romc.eval_posterior(theta, eps_cutoff=False)}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------
+>>> romc.eval_posterior(theta, eps_cutoff=False)
+----------------------------------------------------------------------------
+\end{Code}
 
 \noindent
-This function evaluates the normalised posterior. For doing so it
-needs to approximate the partition function
-$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; this is done using
-the Riemann integral approximation. Unfortunately, the Riemann
-approximation does not scale well in high-dimensional spaces, hence
-the approximation is tractable only at low-dimensional parametric
-spaces. Given that this functionality is particularly useful for
-plotting the posterior, we could say that it is meaningful to be used
-for up to $3D$ parametric spaces, even though it is not restricted to
-that. Finally, for this functionality to work, the \pinline{bounds}
-arguments must have been set at the initialisation of the
-\pinline{elfi.ROMC} object.\footnote{The argument \pinline{bounds}
-  should define a bounding box \emph{containing} all the mass of the
-  prior; it may also contain redundant areas. For example, if the
-  prior is the uniform defined over a unit circle i.e.\
-  $c=(0,0), r=1$, the best bounds arguments is
-  \pinline{bounds=[(-1,1),(-1,1)]}. However, any argument
-  \pinline{bounds=[(-t,t),(-t,t)]} where $t\geq1$ is technically
-  correct.}
+This function evaluates the normalised posterior estimating the
+partition function $Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$;
+with Riemann's integral approximation. The approximation is
+computationally feasible only in low-dimensional parametric
+spaces.
 
@@
 
-With the following code snippet, we perform weighted sampling from the
-ROMC approximate posterior. Afterwards, we used some ELFI's built-in
-tools to get a summary of the obtained samples. In figure
+In the following code snippet, we perform weighted sampling from the
+ROMC approximate posterior. Afterwards, we used some of ELFI's
+built-in tools to get a summary of the obtained samples. In figure
 \ref{fig:example_sampling}, we observe the histogram of the weighted
 samples and the acceptance region of the first deterministic function
-(as before) alongside with the obtained samples obtained from
-it. Finally, in the code snippet we demonstrate how to use the
-\pinline{compute_expectation} function; in the current example we
-define \pinline{h} in order to compute firstly the empirical mean and
-afterwards the empirical variance. In both cases, the empirical result
-is close to the ground truth $\mu = 0$ and $\sigma^2 = 1$.
+(as before) alongside the samples obtained from it. Finally, in the
+code snippet, we demonstrate how to use the \code{compute_expectation}
+function; in the current example, we define \code{h} to compute
+firstly the empirical mean and afterwards the empirical variance. In
+both cases, the empirical result is close to the ground truth
+$\mu = 0$ and $\sigma^2 = 1$. The
+\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
+point $\theta$ using expression \eqref{eq:aprox_posterior}. The
+\code{romc.eval_posterior(theta)} approximates the partition function
+$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
+Riemann approximation as explained above. In our simple example, this
+utility can provide a nice plot of the approximate posterior as
+illustrated in figure~\ref{fig:approx_posterior}. We observe that the
+approximation is quite close to the ground-truth posterior.
 
-\begin{pythoncode}
+\begin{Code}
+------------------------------ python snippet ------------------------------  
   seed = 21
@@
   # Expected variance: 1.061
-\end{pythoncode}
+
+  # eval unnorm posterior
+  romc.eval_unnorm_posterior(theta=0)
+
+  # check eval posterior
+  romc.eval_posterior(theta=0)
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 
-\subsubsection*{Example - Evaluate Posterior}
-
-The \pinline{romc.eval_unnorm_posterior(theta)} evaluates the
-posterior at point $\theta$ using expression
-\eqref{eq:aprox_posterior}. The \pinline{romc.eval_posterior(theta)}
-approximates the partition function
-$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
-Riemann approximation as explained above. In our simple example, this
-utility can provide a nice plot of the approximate posterior as
-illustrated in figure~\ref{fig:approx_posterior}. We observe that the
-approximation is quite close to the ground-truth posterior.
-
-\begin{figure}[ht]
+\begin{figure}[h]
     \begin{center}

### 2021-07-17 latex_files/ch3/3_2_4_evaluation.tex
added content in chapter 2
updated most content in chapter 2,3

--- 2021-07-08 latex_files/ch3/3_1_4_evaluation.tex
@@
-The ROMC implementation provides two functions for evaluating the inference results,
+The ROMC implementation provides two functions for evaluating the inference results.
 
-\begin{enumerate}[label=(\roman*)]
-\item \mintinline{python}{romc.compute_divergence(gt_posterior, bounds=None, step=0.1,}
-
-      \mintinline{python}{                        distance="Jensen-Shannon")}
-
-\item \mintinline{python}{romc.compute_ess()}
-\end{enumerate}
-
-\subsubsection*{Function (i): Compute the divergence between ROMC approximation and a ground-truth posterior}
-
-\pinline{romc.compute_divergence(gt_posterior, bounds=None, step=0.1,}
-
-\pinline{                     distance="Jensen-Shannon")}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------  
+>>> romc.compute_divergence(gt_posterior,
+                            bounds=None,
+                            step=0.1,
+                            distance="Jensen-Shannon")
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 dimensional parameter spaces; it is suggested to be used for up to the
-three-dimensional parameter space. As mentioned in the beginning of
+three-dimensional parameter space. As mentioned at the beginning of
 this chapter, in a real-case scenario it is not expected the
@@
 (as in the current example) or with some other inference approach,
-this function can provide a numerical measure of the agreement
-between the two approximations. The argument \pinline{step} defines
-the step used in the Riemann's approximation and the argument
-\pinline{distance} can take either the \pinline{Jensen-Shannon} or the
-\pinline{KL-divergence} value, for computing the appropriate distance.
+this function can provide a numerical measure of the agreement between
+the two approximations. The argument \code{step} defines the step used
+in the Riemann's approximation and the argument \code{distance} can
+take either the \code{Jensen-Shannon} or the \code{KL-divergence}
+value, for computing the appropriate distance.
 
-\subsubsection*{Function (ii): Compute the effective sample size of the weighted samples}
-
-\mintinline{python}{romc.compute_ess()}
-\vspace{5mm}
+\begin{Code}
+---------------------------------- python ----------------------------------  
+>>> romc.compute_ess()
+----------------------------------------------------------------------------
+\end{Code}
 
@@
 This function computes the Effective Sample Size (ESS) using the
-following expression~\autocite{Sudman1967},
+following expression~\cite{Sudman1967},
 
@@
 
-The ESS is a useful measure of the \textbf{actual} sample size, when
-the samples are weighted. For example if in a population of $100$
-samples one has a very large weight (e.g.\ $\approx 100$) whereas the
-rest have small (i.e.\ $\approx 1$), the real sample size is close to
-1; one sample dominates over the rest. Hence, the ESS provides a
+The ESS is a valuable measure of the \textbf{actual} sample size when
+the samples are weighted. For example, if in a population of $100$
+samples one has an enormous weight (e.g.\ $\approx 100$) whereas the
+rest have small (i.e.\ $\approx 1$), the actual sample size is close
+to 1; one sample dominates over the rest. Hence, the ESS provides a
 measure of the equivalent uniformly weighted sample population.
 
-\newpage
-
-\begin{pythoncode}
+\begin{Code}
+------------------------------ python snippet ------------------------------  
   res = romc.compute_divergence(wrapper, distance="Jensen-Shannon")                                 
@@
 
-print("Nof Samples: %d, ESS: %.3f" % (len(romc.result.weights), romc.compute_ess()))
+  nof_samples = len(romc.result.weights)
+  ess = romc.compute_ess()
+  print("Nof Samples: %d, ESS: %.3f" % (nof_samples, ess))
   # Nof Samples: 19950, ESS: 16694.816
-\end{pythoncode}
+----------------------------------------------------------------------------  
+\end{Code}

### 2021-07-17 paper.tex
updated most content in chapter 2,3

@@
 \usepackage{xcolor}
-
-% must be removed
 \usepackage{multirow}
 \usepackage{caption}
+\usepackage{float}
 
-%% just to show content -> they should be replaced by the journals packages in the end
+%% maybe they should be replaced by the journals packages in the end
 \usepackage{algorithm} 
 \usepackage[noend]{algpseudocode}
-
-%% just to show content -> they should be replaced by the journals packages in the end
-% Python style code
-\usepackage[finalizecache, cachedir=./minted_files]{minted}
-\usemintedstyle{tango}
-\usepackage{listings}
-\usepackage{xcolor}
-\newcommand{\pinline}[1]{\mintinline{python}{#1}}
-\newminted{python}{framesep=2mm,
-baselinestretch=1.2,
-fontsize=\small,
-frame=single
-}
 
@@
 % should be discussed (within the same software environment and beyond), bringing
-% out relative (dis)advantages. All software mentioned should be properly
+p% out relative (dis)advantages. All software mentioned should be properly
 % \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
@@
 
-%% - Virtually all JSS manuscripts list source code along with the generated
-%%   output. The style files provide dedicated environments for this.
-%% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%%   or knitr using the render_sweave() hook - are used (without the need to
-%%   load Sweave.sty).
-%% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-%% - The code input should use "the usual" command prompt in the respective
-%%   software system.
-%% - For R code, the prompt "R> " should be used with "+  " as the
-%%   continuation prompt.
-%% - Comments within the code chunks should be avoided - these should be made
-%%   within the regular LaTeX text.
+% - Virtually all JSS manuscripts list source code along with the generated
+%   output. The style files provide dedicated environments for this.
+% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
+%   or knitr using the render_sweave() hook - are used (without the need to
+%   load Sweave.sty).
+% - Equivalently, {CodeInput} and {CodeOutput} can be used.
+% - The code input should use "the usual" command prompt in the respective
+%   software system.
+% - For R code, the prompt "R> " should be used with "+  " as the
+%   continuation prompt.
+% - Comments within the code chunks should be avoided - these should be made
+%   within the regular LaTeX text.
 
-% \begin{leftbar}
-% For code input and output, the style files provide dedicated environments.
-% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-% hook. Please make sure that all code is properly spaced, e.g., using
-% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-% use ``the usual'' command prompt in the respective software system. For
-% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-% the continuation prompt. Generally, comments within the code chunks should be
-% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-% before and after code input/output should be avoided (see above).
-% \end{leftbar}
+\begin{leftbar}
+For code input and output, the style files provide dedicated environments.
+Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
+or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
+produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
+hook. Please make sure that all code is properly spaced, e.g., using
+\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
+use ``the usual'' command prompt in the respective software system. For
+\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
+the continuation prompt. Generally, comments within the code chunks should be
+avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
+before and after code input/output should be avoided (see above).
+\end{leftbar}
 
@@
 \section{ROMC module in ELFI}  % better title needed
+In this section, we present our implementation. At~\ref{subsec:ELFI}
+we introduce the reader to the \proglang{Python} package \pkg{Engine
+  for likelihood-free inference (ELFI)}
+package. At~\ref{subsec:general_design} we explain the general design
+that we follow. Then, we divide the implementation into three parts;
+at~\ref{subsec:training} we present the training (fitting) part,
+at~\ref{subsec:inference} the inference part and
+at~\ref{subsec:evaluation} the evaluation part. Finally,
+at~\ref{subsec:extensibility} we describe how a user may extend ROMC
+with its custom modules.
 
 \subsection{Engine for likelihood-free inference (ELFI)}
+\label{subsec:ELFI}
 \input{latex_files/ch3/3_1_ELFI.tex}
 
-\subsection{Design Principles}
-\input{latex_files/ch3/3_2_design_principles.tex}
+\subsection{Design principles} 
+\label{subsec:general_design}
+\input{latex_files/ch3/3_2_1_general_design}
+
+\subsection{Training part} 
+\label{subsec:training}
+\input{latex_files/ch3/3_2_2_training}
+
+\subsection{Inference part} 
+\label{subsec:inference}
+\input{latex_files/ch3/3_2_3_inference}
+
+\subsection{Evaluation part} 
+\label{subsec:evaluation}
+\input{latex_files/ch3/3_2_4_evaluation}
 
 \subsection{Extensibility} % better title
+\label{subsec:extensibility}
 \input{latex_files/ch3/3_3_extensibility.tex}
@@
 %% -- Summary/conclusions/discussion -------------------------------------------
-
 \section{Summary and discussion} \label{sec:summary}
-lalala
 
@@
 
-% The results in this paper were obtained using
-% \proglang{R}~3.4.1 with the
-% \pkg{MASS}~7.3.47 package. \proglang{R} itself
-% and all packages used are available from the Comprehensive
-% \proglang{R} Archive Network (CRAN) at
-% \url{https://CRAN.R-project.org/}.
+The results in this paper were obtained using
+\proglang{R}~3.4.1 with the
+\pkg{MASS}~7.3.47 package. \proglang{R} itself
+and all packages used are available from the Comprehensive
+\proglang{R} Archive Network (CRAN) at
+\url{https://CRAN.R-project.org/}.
 

### 2021-07-18 latex_files/ch4/4_2_ma2.tex
updated most content in chapter 2,3
standarized ma2 example

@@
-The second example is the second-order moving average (MA2) which is
-used by the ELFI package as a fundamental model for testing all
-inference implementations. This example is chosen to confirm that our
-implementation of the ROMC approach produces sensible results in a
-general model, since the previous ones where artificially created by
-us.
-
-\subsubsection*{Problem definition}
-
-The second-order moving average (MA2) is a common model used for
-univariate time series analysis. The observation at time $t$ is given by:
+Second-order moving average (MA2) is a probabilistic model commonly
+used for univariate time series analysis. The observation at time $t$
+is given by:
 
@@
 \noindent
-The random variables $w_{k} \sim \mathcal{N}(0,1), k \in \mathbb{Z}$
-represent an independent and identically distributed white noise and
-$\theta_1, \theta_2$ the dependence from the previous
-observations. The number of consecutive observations $T$ is a
-hyper-parameter of the model; in our case we will set
-$T=100$. Generating an MA2 time-series is pretty easy and efficient
-using a simulator, therefore using a likelihood-free inference method
-quite convenient. At the specific example, we use the prior proposed
-by \cite{Marin2012} for guaranteeing that the inference problem is
-identifiable, i.e.\ loosely speaking the likelihood will have just one
-mode. The multivariate prior, which is given in the equation
-\eqref{eq:ma2_prior}, follows a triangular shape as plotted in figure
-\ref{fig:ma2_1}.
+The random variable $w_{k} \sim \mathcal{N}(0,1), k \in \mathbb{Z}$
+represent white noise and $\theta_1, \theta_2$, the two parameters of
+interest, the dependence from the previous observations. The number of
+consecutive observations $T$ is a hyper-parameter of the model that in
+our case is set at $T=100$. For guaranteeing that the inference
+problem is identifiable, i.e.\ the likelihood has only one mode, we
+use the prior proposed by \cite{Marin2012} given in the equation
+\eqref{eq:ma2_prior}:
 
@@
 
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_prior_samples.png}
-    \end{center}
-    \caption[MA2 example, prior distribution.]{Prior distribution proposed by \cite{Marin2012}. The
-      samples follow a triangular shape.}
-  \label{fig:ma2_1}
-\end{figure}
+% \begin{figure}[ht]
+%     \begin{center}
+%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_prior_samples.png}
+%     \end{center}
+%     \caption[MA2 example, prior distribution.]{Prior distribution proposed by \cite{Marin2012}. The
+%       samples follow a triangular shape.}
+%   \label{fig:ma2_1}
+% \end{figure}
 
 \noindent
-The vector $\yb_0 = (y_1, \ldots, y_{100})$ used as the observation,
-has been generated with $\thetab=(0.6, 0.2)$. The dimensionality of
-the output $\yb$ is quite large, therefore we use summary
-statistics. Considering that the ouptput vector represents a
-time-series signal, we prefer the autocovariance as the summary
-statistic; we incorporate the autocovariances with $lag=1$ and
-$lag=2$, as shown in equation \eqref{eq:ma2_summary}. Hence, the
-distance is defined as the squared euclidean distance between the
-summary statistics; this is the same choice as in \cite{Marin2012}.
+The observation vector $\yb_0 = (y_1, \ldots, y_{100})$ is generated
+with $\thetab^*=(0.6, 0.2)$. The dimensionality of the output $\yb$ is
+high, therefore we use summary statistics. Considering that the
+ouptput vector represents a time-series signal, we choose as summary
+statistics the autocovariances with $lag=1$ and $lag=2$, as shown in
+equations \eqref{eq:ma2_summary_1} and \eqref{eq:ma2_summary_2}. The
+final distance node is the squared euclidean
+distance \eqref{eq:ma2_summary_4}.
 
-\begin{gather} \label{eq:ma2_summary}
-  s_1(\yb) = \frac{1}{T-1} \sum_{t=2}^T y_ty_{t-1}\\
+\begin{gather} \label{eq:ma2_summary_1}
+  s_1(\yb) = \frac{1}{T-1} \sum_{t=2}^T y_ty_{t-1}\\ \label{eq:ma2_summary_2}
   s_2(\yb) = \frac{1}{T-2} \sum_{t=3}^T y_ty_{t-2} \\
-  s(\yb) = (s_1(\yb), s_2(\yb))\\
+  s(\yb) = (s_1(\yb), s_2(\yb))\\ \label{eq:ma2_summary_4}
   d = ||s(\yb) - s(\yb_0)||_2^2 
@@
 
-As in the previous example, we perform the inference using the two
-optimisation alternatives, the gradient-based and the Bayesian
-optimisation. In this way, we compare the results obtained in each
-step. For comparing the samples drawn with our implementation, we will
-use the samples obtained with Rejection ABC. In figure \ref{fig:ma2_2}
-we observe that in most cases the optimal distance $d_i^*$ is close to
-zero in both optimisation approaches.
+As in the previous example, we perform the inference using both
+optimisation alternatives (gradient-based and Bayesian optimisation)
+and we compare the results at each step. In the end, we use the
+Rejection ABC algorithm for evaluating both approximations. Rejection
+ABC is a robust method in terms of accuracy and can be used for
+baseline comparison.
 
-In figure \ref{fig:ma2_5}, we have chosen three different
-deterministic optimisation cases (i.e. three different seeds) for
-illustrating three different cases. In the first case, both
-optimisation schemes lead to the creation of almost the same bounding
-box. In the second case, the bounding box has a similar shape, though
-different size and it is shifted along the $\theta_1$ axis. Finally,
-in the third case, the bounding boxes are completely different. We can
-thus conclude that although fitting a surrogate model has important
-computational advantages, there is no guarantee that it will reproduce
-accurately the local region around the optimal point. This
-approximation error may lead to the construction of a considerably
-different proposal region, which in turn, explains the differences in
-the histogram of the marginal distributions presented in figure
-\ref{fig:ma2_3} and in the approximate posteriors in figure
-\ref{fig:ma2_4}.
-
-In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
-marginal posteriors, using the three different methods; (a) Rejection
-ABC (first line), (b) ROMC with gradient-based optimisation (second
-line) and (c) ROMC with Bayesian optimisation (third
-line). Undoubtedly, we observe that there is a significant similarity
-between the three approaches. The Rejection ABC inference has been set
-to infer 10000 accepted samples, with threshold $\epsilon=0.1$. The
-large number of samples and the small distance from the observations
-let us treat them as ground-truth information. In the table \ref{tab:ma2}
-we present the empirical mean $\mu$ and standard deviation $\sigma$
-for each inference approach. We observe that there is a significant
-agreement between the approaches, which verifies that the ROMC
-implementation provides sensible results. Finally, in figure
-\ref{fig:ma2_4} we provide the ROMC approximate posteriors using
-gradients and Bayesian optimisation; as confirmed by the statistics in
-\ref{tab:ma2}, both posteriors have a single mode, located at the same
-point, with a larger variance observed in the Bayesian Optimisation
-case. The posteriors mode is quite close to the parameter
-configuration that generated the data i.e.\ $\thetab = (0.6, 0.2)$.
+In figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
+same deterministic simulator, i.e. with \code{seed=799981517}, in the
+gradient-based and the Bayesian optimisation case. The acceptance
+regions are similar in both cases. The optimal points are close and
+the search directions are similar. This indicates that both
+optimization scemes work robustly. Therefore, we observe that as we
+move away from the optimal pint, the surrogate model fitted by the
+Bayesian Optimisation tends to deviate from the real distance. This
+divergence affects the bounding box approximation and explains some
+small differences between the posterior approximations, as shown in
+figure \ref{fig:ma2_3}. In figure \ref{fig:ma2_3}, we demonstrate the
+histograms of the marginal posteriors, using the three different
+methods; (a) Rejection ABC (first line), (b) ROMC with gradient-based
+optimisation (second line) and (c) ROMC with Bayesian optimisation
+(third line). Undoubtedly, we observe that there is a significant
+similarity between the three approaches. The Rejection ABC inference
+has been set to infer 10000 accepted samples, with threshold
+$\epsilon=0.1$. At table \ref{tab:ma2} we present the empirical mean
+$\mu$ and standard deviation $\sigma$ for each inference approach, for
+verifying their agreement.
 
@@
 \hline
-ROMC (gradient-based) & 0.495 & 0.136 & 0.048 & 0.178 \\
+ROMC (gradient-based) & 0.504 & 0.141 & 0.031 & 0.172 \\
 \hline
-ROMC (Bayesian optimisation) & 0.510 & 0.156 & 0.108 & 0.183 \\
+ROMC (Bayesian optimisation) & 0.5 & 0.165 & 0.07 & 0.167 \\
 \hline
@@
 
-\begin{figure}[h]
+% \begin{figure}[H]
+%     \begin{center}
+%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist.png}
+%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist_bo.png} \\
+%     \end{center}
+%     \caption[MA2 example, histogram of distances]{Histogram of distances
+%       $d_i^*, i \in \{ 1, \ldots, n_1 \}$. The left graph corresponds
+%       to the gradient-based approach and the right one to the Bayesian
+%       optimisation approach.}
+%   \label{fig:ma2_2}
+% \end{figure}
+
+\begin{figure}[H]
     \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist_bo.png}
-    \end{center}
-    \caption[MA2 example, histogram of distances]{Histogram of distances
-      $d_i^*, i \in \{ 1, \ldots, n_1 \}$. The left graph corresponds
-      to the gradient-based approach and the right one to the Bayesian
-      optimisation approach.}
-  \label{fig:ma2_2}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.png}\\
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_2.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_2_bo.png}\\
       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
@@
     \end{center}
-  \caption[MA2 example, the acceptance region in three distinct optimisation problems.]{Visualisation of the acceptance region in 3 different optimisation problems.
-Each row illustrates a different optimisation problem, the left column corresponds to the gradient-based approach and the right column to the Bayesian optimisation approach.
-The examples have been chosen to illustrate three different cases; in the first case, both optimisation schemes lead to similar optimal point and bounding box, in the second case the bounding box is similar in shape but a little bit shifted to the right relatively to the gradient-based approach and in the third case, both the optimal point and the bounding box is completely different.}
+  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimisation problem.
+In the left figure the region obtained with gradient-based optimiser and in the right one with Bayesia Optimisation.}
   \label{fig:ma2_5}
@@
 
-\begin{figure}[ht]
+\begin{figure}[H]
     \begin{center}
@@
     \end{center}
-    \caption[MA2 example, evaluation of the marginal distributions.]{Histogram of the marginal distributions using
-      different inference approaches; (a) in the first row, the
-      approximate posterior samples are obtained using Rejection ABC
-      sampling (b) in the second row, using ROMC sampling with
-      gradient-based approach and (c) in the third row, using ROMC
-      sampling with Bayesian optimisation approach. The vertical (red)
-      line represents the samples mean $\mu$ and the horizontal
-      (black) the standard deviation $\sigma$.}
+    \caption[MA2 example, evaluation of the marginal
+    distributions.]{Histogram of the marginal posterior distributions
+      using three different inference approaches; (a) in the first
+      row, the samples are obtained using Rejection ABC sampling (b)
+      in the second row, using ROMC with a gradient-based optimiser
+      and (c) in the third row, using ROMC with Bayesian optimization
+      approach. The vertical (red) line represents the samples mean
+      $\mu$ and the horizontal (black) the standard deviation
+      $\sigma$.}
   \label{fig:ma2_3}
@@
 
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.png}
-    \end{center}
-    \caption[MA2 example, approximate posterior.]{ROMC approximate posteriors using gradient-based approach
-      (left) and Bayesian optimisation approach (right).}
-  \label{fig:ma2_4}
-\end{figure}
+% \begin{figure}[H]
+%     \begin{center}
+%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.png}
+%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.png}
+%     \end{center}
+%     \caption[MA2 example, approximate posterior.]{ROMC approximate posteriors using gradient-based approach
+%       (left) and Bayesian optimisation approach (right).}
+%   \label{fig:ma2_4}
+% \end{figure}

### 2021-07-18 latex_files/python_scripts/example_MA2.py

@@ -1,17 +1,16 @@
 import logging
 import matplotlib.pyplot as plt
-from elfi.methods.parameter_inference import ROMC
 from elfi.examples import ma2
 import elfi
 import numpy as np
 import scipy.stats
 import matplotlib
 import os
-matplotlib.rcParams['text.usetex'] = True
+matplotlib.rcParams['text.usetex'] = False
 
 logging.basicConfig(level=logging.INFO)
-prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images/chapter4/'
-
+prepath = './../images/chapter4/'
+savefig = True
 
 def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
     plt.figure()
@@ -30,7 +29,8 @@ def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, rang
     plt.ylabel(ylabel)
     plt.ylim(ylim)
     plt.legend()
-    plt.savefig(savepath, bbox_inches='tight')
+    if savepath:
+        plt.savefig(savepath, bbox_inches='tight')
     plt.show(block=False)
 
 
@@ -41,136 +41,100 @@ model = ma2.get_model(seed_obs=seed)
 
 # plot prior samples
 
-x = model.generate(1000)
+x = model.generate(1000, outputs=["t1", "t2"])
 plt.figure()
 plt.title("Samples from the prior")
 plt.xlabel(r"$\theta_1$")
 plt.ylabel(r"$\theta_2$")
 plt.plot(x["t1"], x["t2"], "bo")
-plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
-            bbox_inches="tight")
+if savefig:
+    plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
+                bbox_inches="tight")
 plt.show(block=False)
 
 ####### ROMC with gradients ################
-n1 = 1000
-n2 = 20
+n1 = 100
+n2 = 200
 bounds = [(-2, 2), (-1.25, 1.25)]
 eps = .01
-vis_ind_1 = 12
-vis_ind_2 = 1
-vis_ind_3 = 2
+vis_ind_1 = 1
+vis_ind_2 = 3
+vis_ind_3 = 12
 
-romc = ROMC(model, bounds=bounds, discrepancy_name="d")
+romc = elfi.ROMC(model, bounds=bounds, discrepancy_name="d")
[…]

### 2021-07-18 latex_files/python_scripts/example_MA2_bo.py

@@ -1,0 +1,220 @@
+import logging
+import matplotlib.pyplot as plt
+from elfi.examples import ma2
+import elfi
+import numpy as np
+import scipy.stats
+import matplotlib
+import os
+matplotlib.rcParams['text.usetex'] = False
+
+logging.basicConfig(level=logging.INFO)
+prepath = './../images/chapter4/'
+savefig = True
+
+def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
+    plt.figure()
+    plt.title(title)
+    plt.hist(x=samples,
+             weights=weights,
+             bins=bins, density=True, range=range)
+    plt.axvline(mean, 0, 1,
+                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
+    plt.axhline(1,
+                (mean-std-range[0])/(range[1] - range[0]),
+                (mean+std-range[0])/(range[1] - range[0]),
+                color="k",
+                linestyle="--", label=r"$\sigma = %.3f$" % (std))
+    plt.xlabel(xlabel)
+    plt.ylabel(ylabel)
+    plt.ylim(ylim)
+    plt.legend()
+    if savepath:
+        plt.savefig(savepath, bbox_inches='tight')
+    plt.show(block=False)
+
+
+####### ROMC with BO ###########
+# Set seed for reproducibility
+seed = 1
+np.random.seed(seed)
+model = ma2.get_model(seed_obs=seed)
+
+# plot prior samples
+
+x = model.generate(1000, outputs=["t1", "t2"])
+plt.figure()
+plt.title("Samples from the prior")
+plt.xlabel(r"$\theta_1$")
+plt.ylabel(r"$\theta_2$")
+plt.plot(x["t1"], x["t2"], "bo")
+if savefig:
+    plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
+                bbox_inches="tight")
+plt.show(block=False)
+
+n1 = 100
+n2 = 200
+bounds = [(-2, 2), (-1.25, 1.25)]
+eps = .03
+vis_ind_1 = 1
+vis_ind_2 = 3
+vis_ind_3 = 12
+
+use_bo = True
+romc1 = elfi.ROMC(model, bounds=bounds, discrepancy_name="d")
+romc1.solve_problems(n1=n1, seed=seed, use_bo=use_bo)
+if savefig:
+    romc1.distance_hist(savefig=os.path.join(
[…]

### 2021-07-19 latex_files/ch3/3_2_1_general_design.tex
reordered document

@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \begin{center}
-\includegraphics[width=0.8\textwidth]{./latex_files/graphs/ROMC.png}
+\includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
     \end{center}
@@
 approach. We can exploit this fact for validating the accuracy of our
-implementation. The ground-truth posterior approximated
-computationally is shown in figure~\ref{fig:example_gt}.
+implementation.
 
-\begin{figure}[!h]
-    \begin{center}
-\includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
-    \end{center}
-  \caption[Ground-truth posterior distribution of the simple 1D
-example.]{Ground-truth posterior distribution of the simple 1D
-example.}
-  \label{fig:example_gt}
-\end{figure}
+% \begin{figure}[!h]
+%     \begin{center}
+% \includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
+%     \end{center}
+%   \caption[Ground-truth posterior distribution of the simple 1D
+% example.]{Ground-truth posterior distribution of the simple 1D
+% example.}
+%   \label{fig:example_gt}
+% \end{figure}
 

### 2021-07-19 latex_files/ch3/3_2_2_training.tex
reordered document

@@
 ---------------------------------- python ----------------------------------
->>> romc.fit_posterior(n1,
-                       eps_filter,
-                       use_bo=False,
-                       optimizer_args=None,
-                       seed=None,
-                       use_surrogate=None,
-                       region_args=None,
-                       fit_models=False,
-                       fit_models_args=None,
-                       eps_region=None,
-                       eps_cutoff=None)
+>>> romc.fit_posterior(args*)  # training in a single call
+>>> romc.visualize_region(i)   # acceptance region
+>>> romc.compute_eps(quantile) # estimates ε
 ----------------------------------------------------------------------------
@@
 The function \code{fit_posterior} is a syntactic sugar for applying
-\code{solve_problems} and \code{estimate_regions} into a single
-step.
-
-% merges all steps for constructing the bounding box into
-% a single command. If the user doesn't want to manually inspect the
-% histogram of the distances before deciding where to set the threshold
-% $\epsilon$, he may call \code{romc.fit_posterior()} and the whole
-% training part will be done end-to-end. There are two alternatives for
-% setting the threshold $\epsilon$; the first is to set to a specific
-% value blindly, and the second is to set at as a specific quantile of
-% the histogram of distances. In the second scenario the
-% \code{quantile} argument must be set to a floating number in the
-% range $[0,1]$ and \code{eps='auto'}.
-
-\begin{Code}
----------------------------------- python ----------------------------------
->>> romc.visualize_region(i)
-----------------------------------------------------------------------------
-\end{Code}
-
-\noindent
-
-It can be used for plotting the bounding box around the optimal point,
-when the parameter space is $1D$ or $2D$. The argument \code{i} is the
-index of the corresponding optimization problem i.e.\ $i<n_1$.
-
-\begin{Code}
----------------------------------- python ----------------------------------  
->>> romc.compute_eps(quantile)
-----------------------------------------------------------------------------
-\end{Code}
-
-\noindent
-This function return the appropriate distance value $d_{i=\kappa}^*$
-where $\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the
-collection $\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the
-number of accepted solutions. It can be used to automate the selection
-of the threshold $\epsilon$, e.g.\
-\code{eps=romc.compute_eps(quantile=0.9)}.
+\code{solve_problems} and \\ \code{estimate_regions} into a single
+step. The function \code{visualize_region} can be used for plotting
+the bounding box around the optimal point, when the parameter space is
+$1D$ or $2D$. The argument \code{i} is the index of the corresponding
+optimization problem i.e.\ $i<n_1$. \code{compute_eps} returns the
+appropriate distance value $d_{i=\kappa}^*$ where
+$\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the collection
+$\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the number of
+accepted solutions. It can be used to automate the selection of the
+threshold $\epsilon$, e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.
 
@@
 ------------------------------ python snippet ------------------------------
+  # Trainining (fitting) part
   n1 = 500 # number of optimisation problems
@@
 
+% \begin{figure}[h]
+%   \begin{center}
+%     \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
+%     \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
+%     \end{center}
+%     \caption[Histogram of distances at the 1D example.]{Histogram of
+%       distances and visualisation of a specific region.}
+%     \label{fig:example_training_hist}
+% \end{figure}
+
 \begin{figure}[h]
   \begin{center}
-    \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
-    \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region.png}
+    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
+    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
+    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}
+
     \end{center}

### 2021-07-19 latex_files/ch3/3_2_3_inference.tex
reordered document

@@
 ---------------------------------- python ----------------------------------
->>> romc.compute_expectation(h)
+>>> romc.compute_expectation(h) # E[h(x)|theta]
+>>> romc.eval_unorm_posterior(theta, eps_cutoff=False) # unnorm p(theta)
+>>> romc.eval_posterior(theta, eps_cutoff=False) # normalised p(theta)
 ----------------------------------------------------------------------------
@@
 $E_{p(\thetab|\data)}[h(\thetab)]$ using
-expression~\eqref{eq:expectation}. The argument \code{h} can be
-any python \code{Callable}.
-
-\begin{Code}
----------------------------------- python ----------------------------------
->>> romc.eval_unorm_posterior(theta, eps_cutoff=False)
-----------------------------------------------------------------------------
-\end{Code}
-
-\noindent
-This function computes the unnormalised posterior approximation using
-expression~\eqref{eq:approx_posterior}.
-
-\begin{Code}
----------------------------------- python ----------------------------------
->>> romc.eval_posterior(theta, eps_cutoff=False)
-----------------------------------------------------------------------------
-\end{Code}
-
-\noindent
-This function evaluates the normalised posterior estimating the
-partition function $Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$;
-with Riemann's integral approximation. The approximation is
-computationally feasible only in low-dimensional parametric
-spaces.
+expression~\eqref{eq:expectation}. The argument \code{h} can be any
+python \code{Callable}. \code{eval_unnorm_posterior} computes the
+unnormalised posterior approximation using
+expression~\eqref{eq:approx_posterior}. \code{eval_posterior}
+evaluates the normalised posterior estimating the partition function
+$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; with Riemann's
+integral approximation. The approximation is computationally feasible
+only in low-dimensional parametric spaces.
 
@@
 ------------------------------ python snippet ------------------------------  
+  # Inference part
   seed = 21
@@
   # Visualise marginal (built-in ELFI tool)
-  romc.result.plot_marginals(weights=romc.result.weights, bins=100, range=(-3,3))
+  romc.result.plot_marginals(weights=romc.result.weights,
+                             bins=100, range=(-3,3))
 
@@
   # compute expectation
-  print("Expected value   : %.3f" % romc.compute_expectation(h=lambda x: np.squeeze(x)))
+  exp_val = romc.compute_expectation(h=lambda x: np.squeeze(x))
+  print("Expected value   : %.3f" % exp_val)
   # Expected value   : -0.079
 
-  print("Expected variance: %.3f" % romc.compute_expectation(h=lambda x: np.squeeze(x)**2))
+  exp_var = romc.compute_expectation(h=lambda x: np.squeeze(x)**2)
+  print("Expected variance: %.3f" % exp_var)
   # Expected variance: 1.061
@@
 
-\begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_marginal.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
-    \end{center}
-  \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples.
+% \begin{figure}[h]
+%     \begin{center}
+%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_marginal.png}
+%     \end{center}
+%   \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples.
 (b) Right: Acceptance region around $\theta_1^*$ with the obtained samples plotted inside.}
-  \label{fig:example_sampling}
-\end{figure}
+%   \label{fig:example_sampling}
+% \end{figure}
 
-\begin{figure}[h]
-    \begin{center}
-      \includegraphics[width=0.75\textwidth]{./latex_files/images/chapter3/example_posterior.png}
-    \end{center}
-  \caption[Approximate posterior evaluation, at the 1D example.]{Approximate posterior evaluation.}
-  \label{fig:approx_posterior}
-\end{figure}
+% \begin{figure}[h]
+%     \begin{center}
+%       \includegraphics[width=0.6\textwidth]{./latex_files/images/chapter3/example_posterior.png}
+%     \end{center}
+%   \caption[Approximate posterior evaluation, at the 1D example.]{Approximate posterior evaluation.}
+%   \label{fig:approx_posterior}
+% \end{figure}

### 2021-07-19 latex_files/ch3/3_2_4_evaluation.tex
reordered document

@@
-The ROMC implementation provides two functions for evaluating the inference results.
+The ROMC implementation provides two functions for evaluating the
+inference results.
 
@@
                             distance="Jensen-Shannon")
+>>> romc.compute_ess()
 ----------------------------------------------------------------------------
@@
 take either the \code{Jensen-Shannon} or the \code{KL-divergence}
-value, for computing the appropriate distance.
-
-\begin{Code}
----------------------------------- python ----------------------------------  
->>> romc.compute_ess()
-----------------------------------------------------------------------------
-\end{Code}
-
-\noindent
-This function computes the Effective Sample Size (ESS) using the
-following expression~\cite{Sudman1967},
+value, for computing the appropriate distance. \code{compute_ess}
+computes the Effective Sample Size (ESS) using the following
+expression~\cite{Sudman1967},
 
@@
 ------------------------------ python snippet ------------------------------  
+  # Evaluation part
   res = romc.compute_divergence(wrapper, distance="Jensen-Shannon")                                 

### 2021-07-19 latex_files/ch4/4_2_ma2.tex
reordered document

@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
     \end{center}
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
     \begin{center}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
-      \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
     \end{center}

### 2021-07-19 latex_files/python_scripts/example_MA2.py

@@ -6,7 +6,7 @@ import numpy as np
 import scipy.stats
 import matplotlib
 import os
-matplotlib.rcParams['text.usetex'] = False
+matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
 prepath = './../images/chapter4/'

### 2021-07-19 latex_files/ch2/2_4_ROMC.tex
updated most content in chapter 2,3
initial draft (without intro and conclusion)

--- 2021-07-08 latex_files/ch2/2_2_ROMC.tex
@@
-The simplifications introduced by OMC, although quite useful from a
-computational point-of-view, they suffer from some significant failure modes:
+The simplifications introduced by OMC, although beneficial from a
+computational point-of-view, they suffer from some drawbacks:
 
 \begin{itemize}
-\item The whole acceptable region $\mathcal{S}_i$, for each nuisance variable,
-  shrinks to a single point $\thetab_i^*$; this simplification may add
-  significant error when then the area $\mathcal{S}_i$ is relatively big.
+\item The whole acceptable region $\mathcal{S}_i$
+  shrinks to a single point $\thetab_i^*$/. This simplification adds
+  significant apprimation error when then the area $\mathcal{S}_i$ is relatively big.
 \item The weight $w_i$ is computed based only at the curvature at the
-  point $\thetab_i^*$. This approach is error prone at many cases
-  e.g.\ when $g_i$ is almost flat at $\thetab_i^*$, leading to a
-  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0 \Rightarrow w_i
-  \rightarrow \infty$, thus dominating the posterior.
+  point $\thetab_i^*$.
+Using only local information for estimating the weight can also be misleading.
+For example, if $g_i$ is almost flat at $\thetab_i^*$, i.e.
+$\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then \Rightarrow w_i
+  \rightarrow \infty$ dominating the posterior.
 \item There is no way to solve the optimisation problem
@@
 
-The ROMC approach resolves the aforementioned issues. Instead of
-collapsing the acceptance regions $\mathcal{S}_i$ into single points,
-it tries to approximate them with a bounding box.\footnote{The
-  description on how to estimate the bounding box is provided in the
-  following chapters.}. A uniform distribution is then defined on the
-bounding box area, used as the proposal distribution for importance
-sampling. If we define as $q_i$, the uniform distribution defined on
-the $i-th$ bounding box, weighted sampling is performed as:
+ROMC resolves the failure-modes of OMC mainly by approaching the
+acceptance regions $\mathcal{S}_i$ in a more sophisticated
+way. Instead of collapsing the acceptance regions into single points,
+we approximate them with a bounding box region. A uniform distribution
+defined on the bounding box area, is then used as the \emph{proposal
+  distribution} for importance sampling. If $q_i$ is the uniform
+distribution defined on the $i-th$ bounding box, weighted sampling is
+performed by drawing samples from $q_i$ and assigning the weights as
+in~\eqref{eq:sampling}.
 
 \begin{gather}
-  \label{eq:sampling}
-  \thetab_{ij} \sim q_i \\
+  \thetab_{ij} \sim q_i \\  \label{eq:sampling}
   w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
@@
 \noindent
+
 Having defined the procedure for obtaining weighted samples, any
@@
 
-\subsubsection{Construction of the proposal region}
+\subsubsection{Proposal region estimation}
 
-In this section we will describe mathematically the steps needed for
-computing the proposal distributions $q_i$. There will be also
-presented a Bayesian optimisation alternative when gradients are not
-available.
+The constrution of the proposal regions $q_i$ is performed in two
+steps; (a) solving the deterministic optimisation problems as in OMC
+and (b) constructing the bounding box regions.
 
-\subsubsection*{Define and solve deterministic optimisation problems}
+\subsubsection*{Solving the deterministic optimisation problems}
 
@@
 deterministic function is defined as
-$f_i(\thetab) = M_d(\thetab,\vb_i)$. For constructing the proposal
-region, we search for a point
-$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$; this point can be
+$f_i(\thetab) = M_d(\thetab,\vb_i)$. We look for a point
+$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$. This point can be
 obtained by solving the the following optimisation problem:
@@
 %
-We maintain a list of the solutions $\thetab_i^*$ of the optimisation
-problems. If for a specific set of nuisance variables $\vb_i$, there
-is no feasible solution we add nothing to the list. The optimisation
-problem can be treated as unconstrained, accepting the optimal point
-$\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$ only if
-$g_i(\thetab_i^*) < \epsilon$.
+The optimisation problem can be treated as unconstrained, accepting
+the optimal point $\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$
+only if $g_i(\thetab_i^*) < \epsilon$. The nature of the generative
+model $M_r(\thetab)$, specifies the properties of the objective
+function $g_i$. If $g_i$ is differentiable any gradient-based
+iterative algorithm can be used for~\ref{eq:optProb}. The gradients
+$\nabla_{\thetab} g_i$ can be either provided in closed form or
+approximated by finite differences. Alternatively, if $g_i$ is not
+differentiable, the Bayesian Optimisation approach provides an
+alternative choice~\cite{Shahriari2016}. In this case, apart from
+obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
+the distance $g_i$ is also fitted; the approximate model can then
+substitute the actual model in the following steps, e.g. construction
+of bounding box, sampling, for accelerating the method.
 
-\subsubsection*{Gradient-based approach}
-\label{subsubsec:GB_approach}
-
-The nature of the generative model $M_r(\thetab)$, specifies the
-properties of the objective function $g_i$. If $g_i$ is continuous
-with smooth gradients $\nabla_{\thetab} g_i$ any gradient-based
-iterative algorithm can be used for solving~\ref{eq:optProb}. The
-gradients $\nabla_{\thetab} g_i$ can be either provided in closed form
-or approximated by finite differences.
-
-\subsubsection*{Bayesian optimisation approach}
-\label{subsubsec:GP_approach}
-
-In cases where the gradients are not available, the Bayesian
-optimisation scheme provides an alternative
-choice~\cite{Shahriari2016}. With this approach, apart from
-obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
-the distance $g_i$ is fitted; this approximate model can be used in
-the following steps for making the method more
-efficient. Specifically, in the construction of the proposal region
-and in
-equations~\eqref{eq:approx_posterior},~\eqref{eq:sampling},~\eqref{eq:expectation}
-it could replace $g_i$ in the evaluation of the indicator
-function, providing a major speed-up.
-
-\subsubsection*{Construction of the proposal area $q_i$}
+\subsubsection*{Bounding boxes construction}
 
@@
 $g_i(\thetab_i^*) < \epsilon$, we need to construct a bounding box
-around it. The bounding box $\mathcal{\hat{S}}_i$ must contain the
-acceptance region around $\thetab_i^*$, i.e.\
+area $\mathcal{\hat{S}}_i$ around it. The bounding box must contain
+the acceptance region around $\thetab_i^*$, formally
 $\{ \thetab : g_i(\thetab) < \epsilon$,
 $d(\thetab, \thetab_i^*) < M \}$. The second condition
-$d(\thetab, \thetab_i^*) < M$ is meant to describe that if
+$d(\thetab, \thetab_i^*) < M$ describes that if
 $\mathcal{S}_i := \{ \thetab : g_i(\thetab) < \epsilon \} $ contains a
-number of disjoint sets of $\thetab$ that respect
-$g_i(\thetab) < \epsilon$, we want our bounding box to fit only the
-one that contains $\thetab_i^*$. We seek for a bounding box that is as
-tight as possible to the local acceptance region (enlarging the
-bounding box without a reason decreases the acceptance rate) but large
-enough for not discarding accepted areas.
+number of disjoint sets of $\thetab$ that they fulfill
+$g_i(\thetab) < \epsilon$, we want the bounding box to match only the
+set that contains $\thetab_i^*$. We seek for a bounding box that is as
+tight as possible to the local acceptance region for ensuring high
+acceptance rate in the importance sampling.
 
-In contrast with the OMC approach, we construct the bounding box by
-obtaining search directions and querying the indicator function as we move on them. The
-search directions $\mathbf{v}_d$ are computed as the eigenvectors of
-the curvature at $\thetab_i^*$ and a line-search method is used to
-obtain the limit point where
-$g_i(\thetab_i^* + \kappa \vb_d) \geq
-\epsilon$\footnote{$-\kappa$ is used as well for the opposite direction along the search line}.
-The Algorithm~\ref{alg:region_construction} describes the method in-depth.
-After the limits are obtained along all
-search directions, we define bounding box and the uniform distribution $q_i$. This is the proposal distribution used for the importance
-sampling as explained in \eqref{eq:sampling}.
+Bounding boxes are built in two steps. Firstly, the search directions
+$\mathbf{v}_d$ are computed as the eigenvectors of the curvature at
+$\thetab_i^*$ and then the limits are obtained with a line-search
+method looking for the smallest
+$\kappa: g_i(\thetab_i^* + \kappa \vb_d) \geq
+\epsilon$\footnote{$-\kappa$ is used as well for the opposite
+  direction along the search line}. The
+Algorithm~\ref{alg:region_construction} describes the method
+in-depth. After limits are found along all directions the uniform
+distribution $q_i$ can be defined and is used as the proposal region
+for importance sampling.
 
@@
 After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
-are no longer interested in the surface outside the box. In the future
-steps (e.g. sampling, evaluating the posterior) we will only evaluate
-$g_i$ inside the corresponding bounding box. Hence, we could fit a
-local surrogate model $\hat{g}_i$ for representing the local area
-around $\theta_i^*$. Doing so, in the future steps we can exploit
-$\hat{g}_i$ for evaluating the indicator function instead of running
-the whole deterministic simulator.
+are no longer interested in $g_i(\thetab)$ outside
+$\mathcal{\hat{S}}_i$. In the inference steps (e.g. sampling,
+evaluating the posterior), we evaluate $g_i$ for checking if a sample
+from the bounding box belongs to the acceptance region. Instead of
+running the deterministic simulator $M_d(\thetab_{ij}, \vb_i)$ for
+evaluating $g_i$, ROMC proposes fitting a \emph{light} surrogate model
+$\hat{g}_i$ for approximating $g_i$ around $\theta_i^*$.
 
-Any ML regression model may be chosen as local surrogates. The choice
-should consider the properties of the local region (i.e. size,
-smoothness). The ROMC proposes fitting a simple quadratic model. The
-training set $X: \mathbb{R}^{N \times D}$ is created by sampling $N$
-points from the bounding box and the labels $Y: \mathbb{R}^{N}$ are
-the computed by evaluating $g_i$. The quadratic model is fitted on the
-data points, for minimising the square error.
-
-This additional step places an
-additional step in the training part, increasing the computational
-demands, but promises a major speed at the inference phase (sampling,
-posterior evaluation). It is frequent in ML problems, to be quite
-generous with the execution time at the training phase, but quite
-eager at the inference phase. Fitting a local surrogate model aligns
-with this requirement.
+ROMC proposes fitting a simple quadratic model\footnote{Any ML
+  regression model may be chosen as local surrogate. The choice should
+  consider the properties of the local region (i.e. size,
+  smoothness). } by sampling $N$ points from the bounding box and the
+labels $Y: \mathbb{R}^{N}$ are computed by evaluating $g_i$. This
+additional step adds an extra effort in the training part, but
+promises a significant speed-up at the inference phase. Such imbalance
+is usual in the context of ML where is usual to be quite generous with
+the execution time at the training phase but quite eager at the
+inference phase. Fitting a local surrogate model aligns with this
+requirement.

### 2021-07-19 latex_files/ch2/2_5_ROMC_algorithmic.tex
initial draft (without intro and conclusion)

@@
-In this section, we will provide the algorithmic description of the
-ROMC method; how to solve the optimisation problems using either the
-gradient-based approach or the Bayesian optimisation alternative and
-the construction of the bounding box. Afterwards, we will discuss the
-advantages and disadvantages of each choice in terms of accuracy and
-efficiency.
-
-At a high-level, the ROMC method can be split into the training and
-the inference part.
+Here we present ROMC as a sequence of algorithmic steps. The reader
+may think of the algorithmic description as the backbone of the
+implementation that will follow. At a high-level, ROMC can be split
+into the training and the inference part.
 
@@
 closed-form description is not available, approximate gradients with
-finite-differences requires two evaluations of $g_i$ for
-\textbf{every} parameter $\thetab_d$, which works adequately well for
-low-dimensional problems. When gradients are not available or $g_i$ is
-not differentiable, the Bayesian optimisation paradigm exists as an
-alternative solution. In this scenario, the training part becomes
-slower due to fitting of the surrogate model and the blind
-optimisation steps. Nevertheless, the subsequent task of computing the
-proposal region $q_i$ becomes faster since $\hat{d}_i$ can be used
-instead of $g_i$; hence we avoid to run the simulator
-$M_d(\thetab, \vb_i)$ for each query. The
-algorithms~\ref{alg:training_GB} and~\ref{alg:training_GP} present the
-above procedure.
+finite-differences requires two evaluations of $g_i$ for \emph{every}
+parameter $\theta_d$, which can only work in low-dimensional
+problems. When gradients are not available, the Bayesian optimisation
+paradigm exists as an alternative solution. In this scenario, the
+training part becomes slower due to fitting of the surrogate model and
+the blind optimisation steps. Nevertheless, the subsequent task of
+computing the proposal region $q_i$ becomes faster since $\hat{d}_i$
+can be used instead of $g_i$. The algorithms~\ref{alg:training_GB}
+and~\ref{alg:training_GP} present the above procedure.
 
@@
 Performing the inference includes one or more of the following three
-tasks; (a) evaluating the unnormalised posterior
-$p_{d, \epsilon}(\thetab|\data)$ (b) sampling from the posterior
-$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$ (c) computing an
-expectation $E_{\thetab|\data}[h(\thetab)]$. Computing an expectation
-can be done easily after weighted samples are obtained using the
-equation~\ref{eq:expectation}, so we will not discuss it separately.
-
-\noindent
-Evaluating the unnormalised posterior requires solely the
-deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
-there is no need for solving the optimisation problems and building
-the proposal regions. The evaluation requires iterating over all $g_i$
-and evaluating the distance from the observed data. In contrast, using
-the GP approach, the optimisation part should be performed first for
-fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
-indicator function on them. This provides an important speed-up,
-especially when running the simulator is computationally
-expensive. % The evaluation of the posterior is presented analytically
-% in~\ref{alg:posterior_GB} and~\ref{alg:posterior_GP}.
+tasks; (a) sampling from the posterior
+$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
+expectation $E_{\thetab|\data}[h(\thetab)]$ and (c) evaluating the
+unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
+expectation can be done easily after weighted samples are obtained
+using the equation~\ref{eq:expectation}, so we do not discuss it
+separately.
 
@@
 function must be evaluated for a total of $n_1 \times n_2$ points.
+
+\noindent
+Evaluating the unnormalised posterior requires only the
+deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
+there is no need for solving the optimisation problems and building
+the proposal regions. The evaluation requires iterating over all $g_i$
+and evaluating the distance from the observed data. In contrast, using
+the GP approach, the optimisation part should be performed first for
+fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
+indicator function on them. This provides an important speed-up,
+especially when running the simulator is computationally
+expensive.
 

### 2021-07-19 latex_files/ch3/3_2_2_training.tex
initial draft (without intro and conclusion)

@@
 >>> romc.visualize_region(i)   # acceptance region
->>> romc.compute_eps(quantile) # estimates ε
+>>> romc.compute_eps(quantile) # estimates eps
 ----------------------------------------------------------------------------

### 2021-07-19 latex_files/ch3/3_2_3_inference.tex
initial draft (without intro and conclusion)

@@
   romc.result.summary()
-  # Number of samples: 1720
-  # Sample means: theta: -0.0792
+  # Number of samples: 19300
+  # Sample means: theta: -0.0132
 
@@
   print("Expected value   : %.3f" % exp_val)
-  # Expected value   : -0.079
+  # Expected value   : -0.013
 
@@
   print("Expected variance: %.3f" % exp_var)
-  # Expected variance: 1.061
+  # Expected variance: 1.116
 

### 2021-07-19 latex_files/ch4/4_3_exec.tex
initial draft (without intro and conclusion)

@@
-In this section, we will present the execution time of the basic ROMC
-functionalities. Apart from performing the inference accurately, one
-of the notable advantages of ROMC is its efficiency. In terms of
-performance, ROMC holds two key advantages.
-
-Firstly, all its subparts are parallelisable; optimising the objective
-functions, constructing the bounding boxes, fitting the local
-surrogate models, sampling and evaluating the posterior can be
-executed in a fully-parallel fashion. Therefore, the speed-up can be
-extended as much as the computational resources allow. Specifically,
-at the CPU level, the parallelisation can be incorporate all available
-cores. A similar design can be utilised in the case of having a
-cluster of PCs available. Parallelising the process at the GPU level
-is not that trivial, since the parallel tasks are more complicated
-than simple floating-point operations. Our current implementation
-supports parallelisation only at the CPU level, exploiting all the cores of the CPU.\footnote{At a future update, it is possible to support parallelising the processes using a cluster of PCs.}.
-In subsection
-\ref{subsubsec:parallel}, we will demonstrate the speed-up achieved
-through parallelisation.
-
-The second important advantage concerns the execution of the
-training and the inference phase in distinct time-slots. Therefore, one
-can consume a lot of training time and resources but ask for
-accelerated inference. The use of a lightweight surrogate model around
-the optimal point exploits this essential characteristic; it trades
-some additional computational burden at the training phase, for
-exercising faster inference later. In figures \ref{fig:exec_solve},
-\ref{fig:exec_regions}, \ref{fig:exec_posterior},
-\ref{fig:exec_sample} we can observe this advantage. The
-example measured in these figures is the simple 1D example, used in
-the previous chapter. We observe that fitting local surrogate models
-slows down the training phase (estimating the regions) by a factor of
-$1.5$, but provides a speed-up of $15$ at the inference phase i.e.\
-approximating unnormalised posterior. This outcome would be even more
-potent in larger models, where running the simulator is even more
-expensive.
+As already stated, it is straightforward for the two basic parts of
+the training part to be parallelised. There is a batch of $n_1$
+independent objectives functions to be optimised and $n_1$ bounding
+boxes to be built. Both tasks can be applied in a parallel utilizing
+all the available CPU cores. Our implementation supports
+parallelization using the built-in \proglang{Python}
+\pkg{multiprocess}. In figure \label{fig:exec_parallel} we observe the
+execution times for performing the inference. The parallel version
+performs all tasks between 2.5 and 6 times faster compared to the
+sequential. Optimisation problems are solved almost 6 times
+faster. Sampling is executed 3.5 times faster, whereas evaluating the
+posterior and constructing the bounding boxes almost 2.5 times faster.
 
@@
     \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_solve_grad.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_solve_bo.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/solve_problems_parallel.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/estimate_regions_parallel.png}\\
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/sample_parallel.png}
+      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/eval_post_parallel.png}
+
     \end{center}
-    \caption[Execution time for solving the optimisation problems.]{Execution time for defining and solving the optimisation
-      problems. We observe that the Bayesian optimisation scheme is much more expensive, performing the task slower by a factor of $75$.}
-  \label{fig:exec_solve}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_grad.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_grad_fit.png}\\
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_bo.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_regions_bo_fit.png}
-    \end{center}
-    \caption[Execution time for constructing the n-dimensional bounding box regions.]{Execution time for constructing the n-dimensional
-      bounding box region and, optionally, fitting the local surrogate
-      models. We observe that fitting the surrogate models incurs a
-      small increase by a factor of $1.5$.}
-  \label{fig:exec_regions}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_grad.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_bo.png}\\
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_posterior_grad_fit.png}
-    \end{center}
-    \caption[Execution time for evaluation the unnormalised posterior.]{Execution time for evaluating the unnormalised posterior
-      approximation. We observe that using a Gaussian Process surrogate model is almost 5 times faster than the simulator and the quadratic local surrogate model 15 times faster.}
-  \label{fig:exec_posterior}
-\end{figure}
-
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_grad.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_bo.png}\\
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/exec_sample_grad_fit.png}
-    \end{center}
-    \caption[Execution time for sampling from the approximate posterior.]{Execution time for sampling from the approximate
-      posterior. We observe a small speed-up using the local surrogate model.}
-  \label{fig:exec_sample}
-\end{figure}
-
-\subsubsection{The effect of parallelisation} \label{subsubsec:parallel}
-
-In this section, we measure the execution times for (a) solving the optimisation problems, (b) construct the bounding boxes, (c) sample from the posterior and (d) evaluate the approximate posterior using parallelisation.
-The model used for this experiment is the simple two-dimensional example of section \ref{subsec:ex1}.
-All experiments
-have been executed in a laptop Dell XPS 15 with an 8-core CPU.
-Therefore the maximum expected speed-up can reach a factor of 8.
-Normally the observed speed-up is lower due to the overhead of setting-up the parallel processes.
-The experiments confirm our general expectation; the parallel version performs all tasks between 2.5 and 6 times faster compared to the sequential, as shown in figures \ref{fig:exec_parallel} and \ref{fig:exec_parallel2}.
-The most benefited task is solving the optimisation problems which is performed almost 6 times faster.
-Sampling is executed 3.5 times faster, whereas evaluating the posterior and constructing the bounding boxes almost 2.5 times faster.
-
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/solve_problems_parallel.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/estimate_regions_parallel.png}
-    \end{center}
-    \caption[Execution time of the training part exploiting parallelisation]{Execution
-      time of the training part exploiting parallelisation. At the left figure, we measure
-      the execution time for solving the optimisation problems. At the
-      right figure, we measure the execution time for constructing the
-      bounding boxes.}
+    \caption[Execution time exploiting parallelisation]{In the first
+      line, we compare the parallel and sequential execution of the
+      training part and at the second line the inference part. At the
+      left figure, we measure the execution time for sampling $n_2=50$
+      points per region. At the right figure we measure the execution
+      time for evaluating the posterior at a batch of $50$ points.}
   \label{fig:exec_parallel}
 \end{figure}
-
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/sample_parallel.png}
-      \includegraphics[width=0.48\textwidth]{./Thesis/images/chapter4/eval_post_parallel.png}
-    \end{center}
-    \caption[Execution time of the inference part exploiting parallelisation.]{Execution
-      time of the inference part exploiting parallelisation. At the left figure, we measure
-      the execution time for sampling $n_2=50$ points per region. At the right figure we measure the execution time for evaluating the posterior at a batch of $50$ points.}
-  \label{fig:exec_parallel2}
-\end{figure}

### 2021-07-19 paper.tex
standarized ma2 example
Updates from Overleaf
reordered document
initial draft (without intro and conclusion)

@@
 \documentclass[article]{jss}
-
 %% -- LaTeX packages and custom commands ---------------------------------------
@@
 \usepackage{amsmath,amssymb,bbold, xfrac}
-\usepackage{xcolor}
+% \usepackage{xcolor}
 \usepackage{multirow}
@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
-\Plaintitle{An Extendable Python Implementation of ROMC}
-\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}
+\title{A Parallelized and Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
+\Plaintitle{A Parallelized and Extendable Python Implementation of ROMC}
+\Shorttitle{A Parallelized and Extendable \proglang{Python} Implementation of ROMC}
 
@@
 % should be discussed (within the same software environment and beyond), bringing
-p% out relative (dis)advantages. All software mentioned should be properly
+% out relative (dis)advantages. All software mentioned should be properly
 % \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
@@
 
-\begin{leftbar}
-For code input and output, the style files provide dedicated environments.
-Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-hook. Please make sure that all code is properly spaced, e.g., using
-\code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-use ``the usual'' command prompt in the respective software system. For
-\proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-the continuation prompt. Generally, comments within the code chunks should be
-avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-before and after code input/output should be avoided (see above).
-\end{leftbar}
+% \begin{leftbar}
+% For code input and output, the style files provide dedicated environments.
+% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
+% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
+% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
+% hook. Please make sure that all code is properly spaced, e.g., using
+% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
+% use ``the usual'' command prompt in the respective software system. For
+% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
+% the continuation prompt. Generally, comments within the code chunks should be
+% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
+% before and after code input/output should be avoided (see above).
+% \end{leftbar}
 
+\clearpage
 \section{Background}
-
-\subsection{Likelihood-free inference and implicit models}
 \input{latex_files/ch2/2_1_LFI_intro.tex}
@@
 
-\section{Example}
+\section{Use-case illustration}
+\subsection{Moving average example}
 \input{latex_files/ch4/4_2_ma2.tex}
+
+\subsection{Execution time}
+\input{latex_files/ch4/4_3_exec.tex}
 
@@
 
-The results in this paper were obtained using
-\proglang{R}~3.4.1 with the
-\pkg{MASS}~7.3.47 package. \proglang{R} itself
-and all packages used are available from the Comprehensive
-\proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}.
+% The results in this paper were obtained using
+% \proglang{R}~3.4.1 with the
+% \pkg{MASS}~7.3.47 package. \proglang{R} itself
+% and all packages used are available from the Comprehensive
+% \proglang{R} Archive Network (CRAN) at
+% \url{https://CRAN.R-project.org/}.
 
@@
 
-\begin{leftbar}
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-\end{leftbar}
+% \begin{leftbar}
+% All acknowledgments (note the AE spelling) should be collected in this
+% unnumbered section before the references. It may contain the usual information
+% about funding and feedback from colleagues/reviewers/etc. Furthermore,
+% information such as relative contributions of the authors may be added here
+% (if any).
+% \end{leftbar}
 

### 2021-07-19 latex_files/python_scripts/example_1D.py

@@ -12,10 +12,10 @@ import elfi
 import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 
-prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images'
+prepath = './../images/'
+savefig = True
+
 np.random.seed(21)
-
-
 class Prior:
     def rvs(self, size=None, random_state=None):
         # size from (BS,) -> (BS,1)
@@ -172,8 +172,9 @@ y = np.squeeze(np.array([gt_posterior_pdf(th) for th in theta]))
 plt.plot(theta, y, 'r-.', label=r'Posterior: $p(\theta|y_0)$')
 
 plt.legend()
-plt.savefig(os.path.join(
-    prepath, "chapter3/example_gt.png"), bbox_inches='tight')
+if savefig:
+    plt.savefig(os.path.join(
+        prepath, "chapter3/example_gt.png"), bbox_inches='tight')
 plt.show(block=False)
 
 
@@ -199,22 +200,33 @@ dim = data.shape[-1]
 romc = elfi.ROMC(dist, bounds)
 
 ############# TRAINING ###################
-n1 = 1000
+n1 = 500
 seed = 21
 romc.solve_problems(n1=n1, seed=seed, use_bo=False, optimizer_args=None)
-romc.distance_hist(bins=100,
-                   savefig=os.path.join(prepath, "chapter3/example_theta_dist.png"))
-eps = 1.
+if savefig:
+    romc.distance_hist(bins=100,
+                       savefig=os.path.join(prepath, "chapter3/example_theta_dist.png"))
+else:
+    romc.distance_hist(bins=100)
+    
+eps = .75
 romc.estimate_regions(eps_filter=eps, use_surrogate=False,
-                      region_args=None, fit_models=True)
-romc.visualize_region(0,
-                      savefig=os.path.join(prepath, "chapter3/example_region.png"))
+                      region_args=None, fit_models=False)
+if savefig:
+    romc.visualize_region(0,
+                         savefig=os.path.join(prepath, "chapter3/example_region.png"))
+else:
+    romc.visualize_region(0)
 
+    
 ############# INFERENCE ##################
 n2 = 50
 tmp = romc.sample(n2=n2)
-romc.visualize_region(i=0, savefig=os.path.join(
-    prepath, "chapter3/example_region_samples.png"))
+if savefig:
[…]

### 2021-07-19 latex_files/python_scripts/example_1D_gp.py

@@ -1,290 +1,0 @@
-""" This script generates the figures used in the implementation example.
-"""
-
-import timeit
-import numpy as np
-import elfi
-
-import matplotlib
-matplotlib.rcParams['text.usetex'] = True
-import matplotlib.pyplot as plt
-
-import scipy.integrate as integrate
-import scipy.stats as ss
-
-import os
-# from matplotlib import rc
-# rc('font', **{'family':'sans-serif','sans-serif':['Helvetica']})
-# ## for Palatino and other serif fonts use:
-# #rc('font',**{'family':'serif','serif':['Palatino']})
-# rc('text', usetex=True)
-
-
-prepath = '/home/givasile/ORwDS/edinburgh-thesis/'
-np.random.seed(21)
-
-
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
-
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
-
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
-
-
-class Likelihood:
-    r"""Implements the distribution
-    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
-                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
-                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
-    """
-
-    def rvs(self, theta, seed=None):
-        """Vectorized sampling from likelihood.
-
-        """
-        assert isinstance(theta, np.ndarray)
-        theta = theta.astype(np.float)
-        samples = np.empty_like(theta)
-
-        c = 0.5 - 0.5 ** 4
-
-        tmp_theta = theta[theta <= -0.5]
-        samples[theta <= -0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
-        theta[theta <= -0.5] = np.inf
-
-        tmp_theta = theta[theta <= 0.5]
-        samples[theta <= 0.5] = ss.norm(loc=tmp_theta**4, scale=1).rvs(random_state=seed)
-        theta[theta <= 0.5] = np.inf
[…]

### 2021-07-19 latex_files/python_scripts/example_2D.py

@@ -1,338 +1,0 @@
-import logging
-import matplotlib.pyplot as plt
-import scipy.integrate as integrate
-from elfi.methods.parameter_inference import ROMC
-from elfi.examples import ma2
-import elfi
-import numpy as np
-import scipy.stats as ss
-import matplotlib
-import os
-matplotlib.rcParams['text.usetex'] = True
-
-logging.basicConfig(level=logging.INFO)
-prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images/chapter4/'
-
-
-def plot_marginal(samples, weights, mean, std, marg, title, xlabel, ylabel, bins,
-                  range, ylim, savepath):
-    plt.figure()
-    plt.title(title)
-    plt.hist(x=samples,
-             weights=weights,
-             bins=bins, density=True, range=range)
-    # plt.axvline(mean, 0, 1,
-    #             color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
-    # plt.axhline(.2,
-    #             (mean-std-range[0])/(range[1] - range[0]),
-    #             (mean+std-range[0])/(range[1] - range[0]),
-    #             color="k",
-    #             linestyle="--", label=r"$\sigma = %.3f$" % (std))
-    x = np.linspace(-3, 3, 40)
-    y = [marg(tmp_x) for tmp_x in x]
-    plt.plot(x, y, "r--", label=r"$p(\theta_1|y_0)$")
-    plt.plot
-    plt.xlabel(xlabel)
-    plt.ylabel(ylabel)
-    plt.ylim(ylim)
-    plt.legend()
-    plt.savefig(savepath, bbox_inches='tight')
-    plt.show(block=False)
-
-
-# Set seed for reproducibility
-seed = 21
-np.random.seed(seed)
-
-
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
-
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
-
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
-
-
-class Likelihood:
[…]

### 2021-07-19 latex_files/python_scripts/execution_times.py

@@ -1,347 +1,0 @@
-""" This script generates the figures used in the implementation example.
-"""
-
-import os
-import scipy.stats as ss
-import scipy.integrate as integrate
-import matplotlib.pyplot as plt
-import timeit
-import numpy as np
-import elfi
-
-import matplotlib
-matplotlib.rcParams['text.usetex'] = True
-
-
-prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/images/chapter4'
-np.random.seed(21)
-
-
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
-
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
-
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
-
-
-class Likelihood:
-    r"""Implements the distribution
-    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
-                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
-                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
-    """
-
-    def rvs(self, theta, seed=None):
-        """Vectorized sampling from likelihood.
-
-        """
-        assert isinstance(theta, np.ndarray)
-        theta = theta.astype(np.float)
-        samples = np.empty_like(theta)
-
-        c = 0.5 - 0.5 ** 4
-
-        tmp_theta = theta[theta <= -0.5]
-        samples[theta <= -
-                0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
-        theta[theta <= -0.5] = np.inf
-
-        tmp_theta = theta[theta <= 0.5]
-        samples[theta <= 0.5] = ss.norm(
-            loc=tmp_theta**4, scale=1).rvs(random_state=seed)
-        theta[theta <= 0.5] = np.inf
-
-        tmp_theta = theta[theta < np.inf]
-        samples[theta < np.inf] = ss.norm(
-            loc=tmp_theta - c, scale=1).rvs(random_state=seed)
-        theta[theta < np.inf] = np.inf
-
[…]

### 2021-07-19 latex_files/python_scripts/execution_times_parallel.py

@@ -13,7 +13,8 @@ import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 
 np.random.seed(21)
-
+prepath = './../images/chapter4' 
+savefig = True
 
 class Prior:
     def rvs(self, size=None, random_state=None):
@@ -206,7 +207,7 @@ bounds = [(-2.5, 2.5), (-2.5, 2.5)]
 # LINEAR
 seed = 21
 eps = 1
-n1 = np.linspace(1, 501, 10)
+n1 = np.linspace(1, 501, 5)
 solve_grad = []
 estimate_regions = []
 sample = []
@@ -266,52 +267,53 @@ for i, n in enumerate(n1):
     eval_post_parallel.append(toc-tic)
 
 
-prepath = '/home/givasile/ORwDS/edinburgh-thesis/Thesis/tmp_images/chapter4'
-
-
 plt.figure()
-plt.title("Solve optimisation problems: sequential vs parallel")
+plt.title("Optimization problems: Sequential vs Parallel Execution")
 plt.plot(n1, solve_grad, "bo--", label="linear")
-plt.plot(n1, solve_grad_parallel, "ro--", label="parallel")
+plt.plot(n1, solve_grad_parallel, "go--", label="parallel")
 plt.xlabel(r"$n_1$")
 plt.ylabel("time (sec)")
 plt.legend()
-plt.savefig(os.path.join(prepath, "solve_problems_parallel"),
-            bbox_inches="tight")
+if savefig:
+    plt.savefig(os.path.join(prepath, "solve_problems_parallel"),
+                bbox_inches="tight")
 plt.show(block=False)
 
 
 plt.figure()
-plt.title("Construct bounding boxes: sequential vs parallel")
+plt.title("Acceptance Regions: Sequential vs Parallel Execution")
 plt.plot(n1, estimate_regions, "bo--", label="linear")
-plt.plot(n1, estimate_regions_parallel, "ro--", label="parallel")
+plt.plot(n1, estimate_regions_parallel, "go--", label="parallel")
 plt.xlabel(r"$n_1$")
 plt.ylabel("time (sec)")
 plt.legend()
-plt.savefig(os.path.join(prepath, "estimate_regions_parallel"),
-            bbox_inches="tight")
+if savefig:
+    plt.savefig(os.path.join(prepath, "estimate_regions_parallel"),
+                bbox_inches="tight")
 plt.show(block=False)
 
 
 plt.figure()
[…]

### 2021-07-23 latex_files/ch2/2_2_LFI_background.tex
create tikz figures and plots

@@
-As already stated, in simulator-based models we cannot evaluate the
-posterior $p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
+In simulator-based models we cannot evaluate the posterior
+$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
 intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
 following equation allows incorporating the simulator in the place of
-the likelihood and forms the basis of all likelihood-free inference
-approaches,
+the likelihood,
 

### 2021-07-23 latex_files/ch3/3_2_1_general_design.tex
create tikz figures and plots

@@
 
-\begin{figure}[h]
-    \begin{center}
-\includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
-    \end{center}
-    \caption[Overview of the ROMC implementation.]{Overview of the
-ROMC implementation. The training part follows a sequential pattern;
-the functions in the green ellipses must be called in a sequential
-fashion for completing the training part and define the posterior
-distribution. The functions in blue ellipses are the functionalities
-provided to the user.}
-    \label{fig:romc_overview}
-\end{figure}
+% \begin{figure}[h]
+%     \begin{center}
+% \includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
+%     \end{center}
+%     \caption[Overview of the ROMC implementation.]{Overview of the
+%       ROMC implementation. The training part follows a sequential
+%       pattern; the functions in the green ellipses must be called in a
+%       sequential fashion for completing the training part and define
+%       the posterior distribution. The functions in blue ellipses are
+%       the functionalities provided to the user.}
+%     \label{fig:romc_overview}
+% \end{figure}
 

### 2021-07-23 latex_files/ch3/3_2_2_training.tex
create tikz figures and plots

@@
 
+% \begin{figure}[h]
+%   \begin{center}
+%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
+%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
+%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}
+
+%     \end{center}
+%     \caption[Histogram of distances at the 1D example.]{Histogram of
+%       distances and visualisation of a specific region.}
+%     \label{fig:example_training_hist}
+%   \end{figure}
+
 \begin{figure}[h]
   \begin{center}    
-    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
-    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
-    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}
-
+    \input{latex_files/images/chapter3/example_gt.tex}
+    \input{latex_files/images/chapter3/example_posterior.tex}
   \end{center}

### 2021-07-23 latex_files/ch4/4_2_ma2.tex
create tikz figures and plots

@@
     \begin{center}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
+      % \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
+      \input{./latex_files/images/chapter4/ma2_region_3.tex}
+      \input{./latex_files/images/chapter4/ma2_region_3_bo.tex}
+      % \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
     \end{center}
@@
 
+% \begin{figure}[h]
+%     \begin{center}
+%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
+%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
+%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
+%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
+%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
+%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
+%     \end{center}
+%     \caption[MA2 example, evaluation of the marginal
+%     distributions.]{Histogram of the marginal posterior distributions
+%       using three different inference approaches; (a) in the first
+%       row, the samples are obtained using Rejection ABC sampling (b)
+%       in the second row, using ROMC with a gradient-based optimiser
+%       and (c) in the third row, using ROMC with Bayesian optimization
+%       approach. The vertical (red) line represents the samples mean
+%       $\mu$ and the horizontal (black) the standard deviation
+%       $\sigma$.}
+%   \label{fig:ma2_3}
+% \end{figure}
+
 \begin{figure}[h]
   \begin{center}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
+    \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
+    \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}\\
+    \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
+    \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}\\
+    \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
+    \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}\\
     \end{center}

### 2021-07-23 latex_files/ch4/4_3_exec.tex
create tikz figures and plots

@@
 
-\begin{figure}[ht]
-    \begin{center}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/solve_problems_parallel.png}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/estimate_regions_parallel.png}\\
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/sample_parallel.png}
-      \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/eval_post_parallel.png}
-
-    \end{center}
+\begin{figure}[ht] \label{fig:exec_parallel}
+  \centering
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Solve optimization problems},
+          xlabel={$n_1$},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Proposal regions},
+          xlabel={$n_1$},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Sample},
+          xlabel={$n_1$},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Evaluate posterior},
+          xlabel={$n_1$},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
     \caption[Execution time exploiting parallelisation]{In the first
@@
       time for evaluating the posterior at a batch of $50$ points.}
-  \label{fig:exec_parallel}
 \end{figure}

### 2021-07-23 latex_files/graphs/romc_graph.tex
create tikz figures and plots

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
\tikzstyle{train_process} = [rectangle, minimum width=3cm, minimum height=.7cm, text centered, draw=black, fill=green!40]
\tikzstyle{infer_process} = [rectangle, minimum width=3cm, minimum height=.7cm, text centered, draw=black, fill=blue!30]

\tikzstyle{decision} = [diamond, minimum width=.1cm, minimum height=.1cm, text centered, draw=black, fill=green!30]

\tikzstyle{public_func_rev} = [draw=black, rotate=90, anchor=north, fill=blue!20, rounded corners]
\tikzstyle{public_func} = [draw=black, fill=blue!20, rounded corners]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{figure}[h]
  \begin{center}
  %   \resizebox{.32\textwidth}{!}{
  %     \begin{tikzpicture}[node distance=1.4cm, scale=.1]
  %     \end{tikzpicture}
  %   }
    \resizebox{.8\textwidth}{!}{    
      \begin{tikzpicture}[node distance=1.2cm, scale=.1]
        % first graph
        % private functions
        \node (n1) [train_process, xshift=-.8cm] { $\_sample\_nuisance()$  };
        \node (n2) [train_process, below of=n1] { $\_define\_objectives()$  };
        \node (n3) [decision, below of=n2, yshift=-.5cm] { $grads$  };
        \node (n4) [train_process, left of=n3, yshift=-1cm, xshift=-1.5cm] { $\_solve\_gradients()$  };
        \node (n5) [train_process, right of=n3, yshift=-1cm, xshift=1.5cm] { $\_solve\_bo()$  };
        \node (n6) [train_process, below of=n3, yshift=-1.5cm] { $\_filter\_solutions()$  };
        \node (n7) [train_process, below of=n6] { $\_build\_boxes()$  };
        \node (n8) [decision, below of=n7] { $h?$  };
        \node (n9) [train_process, right of=n8, yshift=-.5cm, xshift=1cm] { $\_fit\_models()$  };

        \node (n10) [train_process, below of=n8, yshift=-1cm] { $\_define\_posterior()$  };

        % public functions
        \node (n11) [public_func_rev, right of=n3, yshift=-5.6cm, xshift=-0.3cm, minimum width=4.7cm] {$solve\_problems()$};
        \node (n12) [public_func_rev, right of=n9, yshift=-3.4cm, xshift=-.45cm, minimum width=4.7cm] {$estimate\_regions()$};
        \node (n13) [public_func_rev, right of=n5, yshift=-4.5cm, xshift=-2.3cm, minimum width=10.5cm] {$fit\_posterior()$};
        \node (n14) [public_func, left of=n3, yshift=0.3cm, xshift=-1.5cm] {$distance\_hist()$};
        \node (n15) [public_func, left of=n6, xshift=-2.2cm, yshift=-.3cm] {$compute\_eps()$};        
        \node (n16) [public_func, left of=n7, xshift=-1.8cm, yshift=-1.5cm] {$visualize\_region()$};

        % public functions inference
        \node (n17) [public_func, below of=n10, xshift=-2cm, yshift=-.5cm] {$sample()$};
        \node (n18) [public_func, below of=n17] {$compute\_expectation()$};        
        \node (n19) [public_func, below of=n10, xshift=4cm, yshift=-.5cm] {$eval\_unnorm\_posterior()$};
        \node (n20) [public_func, below of=n19] {$eval\_posterior()$};

        % public functions evaluation
        \node (n21) [public_func, below of=n18, yshift=-.7cm] {$compute\_ess()$};
        \node (n22) [public_func, below of=n20, yshift=-.7cm] {$compute\_divergence()$};

        % add headers
        \node (algorithm) [below of=n1, yshift=2.5cm, xshift=1cm, minimum width=4cm, minimum height=1cm] {\huge{Implementation Design} };

        % backgrounds
        \draw [ultra thick, draw=black, fill=yellow, opacity=0.15, rounded corners=20pt] (-115,8) rectangle (70, -108);

        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115,8) rectangle (57, -48);
        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115,-50) rectangle (57, -108);

        \draw [ultra thick, draw=black, fill=green, opacity=0.15, rounded corners=20pt] (-115,-110) rectangle (70, -140);

        \draw [ultra thick, draw=black, fill=cyan, opacity=0.15, rounded corners=20pt] (-115,-142) rectangle (70, -160);

        \draw [ultra thick, draw=black, fill=magenta, opacity=0.05, rounded corners=20pt] (-115,18) rectangle (-64, -160);

        \draw [ultra thick, draw=black, fill=magenta, opacity=0.05, rounded corners=20pt] (-60,18) rectangle (70, -160);

        % arrows
        \draw [arrow] (n1) -- (n2);
        \draw [arrow] (n2) -- (n3);
        \draw [arrow] (n3) -- (n4);
        \draw [arrow] (n3) -- (n5);
        \draw [arrow] (n4) -- (n6);
        \draw [arrow] (n5) -- (n6);
        \draw [arrow] (n6) -- (n7);
        \draw [arrow] (n7) -- (n8);
        \draw [arrow] (n8) -- (n9);
        \draw [arrow] (n9) -- (n10);
        \draw [arrow] (n8) -- (n10);

        % second graph
        \node (pro1) [train_process, left of=n1, xshift=-7cm, yshift=-1cm, minimum width=4cm, minimum height=1cm] { define $d_i(\thetab) \forall i$  };
        \node (pro2) [train_process, below of=pro1, yshift=-1cm, minimum width=4cm, minimum height=1cm] { Solve $\thetab_i^*, d_i^*, \forall i$  };
        \node (filter) [train_process, below of=pro2, yshift=-1.5cm, minimum width=4cm, minimum height=1cm] { Filter solutions  };
        \node (proposal_region) [train_process, below of=filter, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $q_i \forall i$};
        \node (surrogate) [train_process, below of=proposal_region, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $\Tilde{d}_i \forall i$};
        \node (posterior) [train_process, below of=surrogate, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Define $p_{d,\epsilon}(\thetab)$};

        \node (sample) [infer_process, below of=posterior, yshift=-1.2cm, minimum width=4cm, minimum height=1cm] {Draw $\{w_{ij}, \thetab_{ij} \}$ };

        % add headers
        \node (algorithm) [below of=pro1, yshift=3.5cm, minimum width=4cm, minimum height=1cm] {\huge{Algorithm} };

        \draw [arrow] (pro1) -- (pro2);
        \draw [arrow] (pro2) -- (filter);
        \draw [arrow] (filter) -- (proposal_region);
        \draw [arrow] (proposal_region) -- (surrogate);
        \draw [arrow] (surrogate) -- (posterior);
        \draw [arrow] (posterior) -- (sample);

        % % backgrounds
        % \draw [ultra thick, draw=black, fill=yellow, opacity=0.15, rounded corners=5pt] (-72,8) rectangle (-62,-108);

        % % \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-60,8) rectangle (57, -48);
        % % \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-60,-50) rectangle (57, -108);

        % % \draw [ultra thick, draw=black, fill=green, opacity=0.15, rounded corners=20pt] (-60,-110) rectangle (70, -140);

        % % \draw [ultra thick, draw=black, fill=cyan, opacity=0.15, rounded corners=20pt] (-60,-142) rectangle (70, -160);

        % \draw [arrow] (pro1) -- (pro2);
        % \draw [arrow] (pro2) -- (filter);
        % \draw [arrow] (filter) -- (proposal_region);
        % \draw [arrow] (proposal_region) -- (surrogate);

      \end{tikzpicture}
    }
  \end{center}
  \caption{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
\end{figure}

### 2021-07-23 latex_files/images/chapter3/example_marginal.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
tick align=outside,
tick pos=left,
title={Histogram of the samples drawn},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta\)},
xmin=-3.3, xmax=3.3,
xtick style={color=black},
xtick={-4,-3,-2,-1,0,1,2,3,4},
xticklabels={
  \(\displaystyle {\ensuremath{-}4}\),
  \(\displaystyle {\ensuremath{-}3}\),
  \(\displaystyle {\ensuremath{-}2}\),
  \(\displaystyle {\ensuremath{-}1}\),
  \(\displaystyle {0}\),
  \(\displaystyle {1}\),
  \(\displaystyle {2}\),
  \(\displaystyle {3}\),
  \(\displaystyle {4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=0.6,
ytick style={color=black},
ytick={0,0.1,0.2,0.3,0.4,0.5,0.6},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.1}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.3}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {0.6}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-3,0) rectangle (axis cs:-2.91428571428571,0);
\draw[draw=none,fill=color0] (axis cs:-2.91428571428571,0) rectangle (axis cs:-2.82857142857143,0);
\draw[draw=none,fill=color0] (axis cs:-2.82857142857143,0) rectangle (axis cs:-2.74285714285714,0);
\draw[draw=none,fill=color0] (axis cs:-2.74285714285714,0) rectangle (axis cs:-2.65714285714286,0);
\draw[draw=none,fill=color0] (axis cs:-2.65714285714286,0) rectangle (axis cs:-2.57142857142857,0);
\draw[draw=none,fill=color0] (axis cs:-2.57142857142857,0) rectangle (axis cs:-2.48571428571429,0.00722007830101246);
\draw[draw=none,fill=color0] (axis cs:-2.48571428571429,0) rectangle (axis cs:-2.4,0.0252702740535435);
\draw[draw=none,fill=color0] (axis cs:-2.4,0) rectangle (axis cs:-2.31428571428571,0.0347466268236223);
\draw[draw=none,fill=color0] (axis cs:-2.31428571428571,0) rectangle (axis cs:-2.22857142857143,0.0374541561865022);
\draw[draw=none,fill=color0] (axis cs:-2.22857142857143,0) rectangle (axis cs:-2.14285714285714,0.0510691609826969);
\draw[draw=none,fill=color0] (axis cs:-2.14285714285714,0) rectangle (axis cs:-2.05714285714286,0.0647744167576544);
\draw[draw=none,fill=color0] (axis cs:-2.05714285714286,0) rectangle (axis cs:-1.97142857142857,0.080503873056289);
\draw[draw=none,fill=color0] (axis cs:-1.97142857142857,0) rectangle (axis cs:-1.88571428571429,0.089941546835469);
\draw[draw=none,fill=color0] (axis cs:-1.88571428571429,0) rectangle (axis cs:-1.8,0.113625982262183);
\draw[draw=none,fill=color0] (axis cs:-1.8,0) rectangle (axis cs:-1.71428571428571,0.12614508231626);
\draw[draw=none,fill=color0] (axis cs:-1.71428571428571,0) rectangle (axis cs:-1.62857142857143,0.145794009692587);
\draw[draw=none,fill=color0] (axis cs:-1.62857142857143,0) rectangle (axis cs:-1.54285714285714,0.171554217630842);
\draw[draw=none,fill=color0] (axis cs:-1.54285714285714,0) rectangle (axis cs:-1.45714285714286,0.157720031886223);
\draw[draw=none,fill=color0] (axis cs:-1.45714285714286,0) rectangle (axis cs:-1.37142857142857,0.174635643905738);
\draw[draw=none,fill=color0] (axis cs:-1.37142857142857,0) rectangle (axis cs:-1.28571428571429,0.199983275941078);
\draw[draw=none,fill=color0] (axis cs:-1.28571428571429,0) rectangle (axis cs:-1.2,0.216744171997);
\draw[draw=none,fill=color0] (axis cs:-1.2,0) rectangle (axis cs:-1.11428571428571,0.240854076323595);
\draw[draw=none,fill=color0] (axis cs:-1.11428571428571,0) rectangle (axis cs:-1.02857142857143,0.277844084619317);
\draw[draw=none,fill=color0] (axis cs:-1.02857142857143,0) rectangle (axis cs:-0.942857142857143,0.253682608304859);
\draw[draw=none,fill=color0] (axis cs:-0.942857142857143,0) rectangle (axis cs:-0.857142857142857,0.2729834247631);
\draw[draw=none,fill=color0] (axis cs:-0.857142857142857,0) rectangle (axis cs:-0.771428571428571,0.303178823657691);
\draw[draw=none,fill=color0] (axis cs:-0.771428571428571,0) rectangle (axis cs:-0.685714285714285,0.297596155971373);
\draw[draw=none,fill=color0] (axis cs:-0.685714285714285,0) rectangle (axis cs:-0.6,0.292864426084818);
\draw[draw=none,fill=color0] (axis cs:-0.6,0) rectangle (axis cs:-0.514285714285714,0.327185584008558);
\draw[draw=none,fill=color0] (axis cs:-0.514285714285714,0) rectangle (axis cs:-0.428571428571428,0.303294860630386);
\draw[draw=none,fill=color0] (axis cs:-0.428571428571428,0) rectangle (axis cs:-0.342857142857143,0.313867118142585);
\draw[draw=none,fill=color0] (axis cs:-0.342857142857143,0) rectangle (axis cs:-0.257142857142857,0.316600433499394);
\draw[draw=none,fill=color0] (axis cs:-0.257142857142857,0) rectangle (axis cs:-0.171428571428571,0.325419243424203);
\draw[draw=none,fill=color0] (axis cs:-0.171428571428571,0) rectangle (axis cs:-0.0857142857142859,0.322737500055257);
\draw[draw=none,fill=color0] (axis cs:-0.0857142857142859,0) rectangle (axis cs:0,0.331556309980064);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.0857142857142859,0.327533694926642);
\draw[draw=none,fill=color0] (axis cs:0.0857142857142859,0) rectangle (axis cs:0.171428571428571,0.351940138183459);
\draw[draw=none,fill=color0] (axis cs:0.171428571428571,0) rectangle (axis cs:0.257142857142857,0.308245771465366);
\draw[draw=none,fill=color0] (axis cs:0.257142857142857,0) rectangle (axis cs:0.342857142857143,0.329080854562574);
\draw[draw=none,fill=color0] (axis cs:0.342857142857143,0) rectangle (axis cs:0.428571428571429,0.311752666640144);
\draw[draw=none,fill=color0] (axis cs:0.428571428571429,0) rectangle (axis cs:0.514285714285714,0.312139456549128);
\draw[draw=none,fill=color0] (axis cs:0.514285714285714,0) rectangle (axis cs:0.6,0.322673035070425);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.685714285714286,0.302650210782081);
\draw[draw=none,fill=color0] (axis cs:0.685714285714286,0) rectangle (axis cs:0.771428571428571,0.313763974166856);
\draw[draw=none,fill=color0] (axis cs:0.771428571428571,0) rectangle (axis cs:0.857142857142857,0.278656343428181);
\draw[draw=none,fill=color0] (axis cs:0.857142857142857,0) rectangle (axis cs:0.942857142857143,0.269785961515508);
\draw[draw=none,fill=color0] (axis cs:0.942857142857143,0) rectangle (axis cs:1.02857142857143,0.245830773152508);
\draw[draw=none,fill=color0] (axis cs:1.02857142857143,0) rectangle (axis cs:1.11428571428571,0.227754791406045);
\draw[draw=none,fill=color0] (axis cs:1.11428571428571,0) rectangle (axis cs:1.2,0.218858623499437);
\draw[draw=none,fill=color0] (axis cs:1.2,0) rectangle (axis cs:1.28571428571429,0.218433154599559);
\draw[draw=none,fill=color0] (axis cs:1.28571428571429,0) rectangle (axis cs:1.37142857142857,0.21010427855946);
\draw[draw=none,fill=color0] (axis cs:1.37142857142857,0) rectangle (axis cs:1.45714285714286,0.193691493421625);
\draw[draw=none,fill=color0] (axis cs:1.45714285714286,0) rectangle (axis cs:1.54285714285714,0.194078283330608);
\draw[draw=none,fill=color0] (axis cs:1.54285714285714,0) rectangle (axis cs:1.62857142857143,0.130644738257426);
\draw[draw=none,fill=color0] (axis cs:1.62857142857143,0) rectangle (axis cs:1.71428571428571,0.150177628661059);
\draw[draw=none,fill=color0] (axis cs:1.71428571428571,0) rectangle (axis cs:1.8,0.0986572127845487);
\draw[draw=none,fill=color0] (axis cs:1.8,0) rectangle (axis cs:1.88571428571429,0.0960012554095325);
\draw[draw=none,fill=color0] (axis cs:1.88571428571429,0) rectangle (axis cs:1.97142857142857,0.0732709017583105);
\draw[draw=none,fill=color0] (axis cs:1.97142857142857,0) rectangle (axis cs:2.05714285714286,0.0897223658870459);
\draw[draw=none,fill=color0] (axis cs:2.05714285714286,0) rectangle (axis cs:2.14285714285714,0.0617058834797238);
\draw[draw=none,fill=color0] (axis cs:2.14285714285714,0) rectangle (axis cs:2.22857142857143,0.0434622927727019);
\draw[draw=none,fill=color0] (axis cs:2.22857142857143,0) rectangle (axis cs:2.31428571428571,0.0347466268236225);
\draw[draw=none,fill=color0] (axis cs:2.31428571428571,0) rectangle (axis cs:2.4,0.0374541561865018);
\draw[draw=none,fill=color0] (axis cs:2.4,0) rectangle (axis cs:2.48571428571429,0.0320390974607428);
\draw[draw=none,fill=color0] (axis cs:2.48571428571429,0) rectangle (axis cs:2.57142857142857,0.00496380383194607);
\draw[draw=none,fill=color0] (axis cs:2.57142857142857,0) rectangle (axis cs:2.65714285714286,0);
\draw[draw=none,fill=color0] (axis cs:2.65714285714286,0) rectangle (axis cs:2.74285714285714,0);
\draw[draw=none,fill=color0] (axis cs:2.74285714285714,0) rectangle (axis cs:2.82857142857143,0);
\draw[draw=none,fill=color0] (axis cs:2.82857142857143,0) rectangle (axis cs:2.91428571428571,0);
\draw[draw=none,fill=color0] (axis cs:2.91428571428571,0) rectangle (axis cs:3,0);
\addplot [semithick, red, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0
-2.59322023391724 0
-2.49152541160583 0.0369433164596558
-2.38983058929443 0.0452905893325806
-2.28813552856445 0.0549525022506714
-2.18644070625305 0.0659897327423096
-2.08474564552307 0.0784285068511963
-1.98305082321167 0.0922528505325317
-1.88135588169098 0.107397437095642
-1.77966105937958 0.123741984367371
-1.67796611785889 0.141106963157654
-1.5762711763382 0.159253358840942
-1.27118647098541 0.21515953540802
-1.16949152946472 0.232988834381104
-1.06779658794403 0.249699831008911
-0.96610164642334 0.264856100082397
-0.864406824111938 0.278041839599609
-0.762711882591248 0.288881063461304
-0.661016941070557 0.297054767608643
-0.559322118759155 0.302316904067993
-0.457627058029175 0.304275751113892
-0.355932235717773 0.304529428482056
0.254237294197083 0.304565906524658
0.355932235717773 0.304529428482056
0.457627058029175 0.304275751113892
0.559322118759155 0.302316904067993
0.661016941070557 0.297054767608643
0.762711882591248 0.288881063461304
0.864406824111938 0.278041839599609
0.96610164642334 0.264856100082397
1.06779658794403 0.249699831008911
1.16949152946472 0.232988834381104
1.27118647098541 0.21515953540802
1.3728814125061 0.196650266647339
1.5762711763382 0.159253358840942
1.67796611785889 0.141106963157654
1.77966105937958 0.123741984367371
1.88135588169098 0.107397437095642
1.98305082321167 0.0922528505325317
2.08474564552307 0.0784285068511963
2.18644070625305 0.0659897327423096
2.28813552856445 0.0549525022506714
2.38983058929443 0.0452905893325806
2.49152541160583 0.0369433164596558
2.59322023391724 0
3 0
};
\end{axis}

\end{tikzpicture}

### 2021-07-23 latex_files/images/chapter4/mae2_hist_t1_rejection.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={Rejection ABC - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.02965548855742);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.17070418562009);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.6643746253394);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.50922105857047);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,1.85479036637399);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.25677915300258);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.3484608060933);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.53887654712789);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.74339715786876);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.79981663669382);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.02549455199407);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.87739342007828);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.67287280933742);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.11012377023167);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.97612750802216);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.77865933213442);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.67992524419055);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.68697767904369);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.24972671814944);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,2.45424732889032);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,2.35551324094642);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,2.12278289079306);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.9464720194647);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.960576889171);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.5162734934236);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.25533340385768);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.42459184033288);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.10723227194188);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.853344617229104);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.768715398991495);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.66292887619451);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.550089918544377);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.500722874572446);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.430198526041119);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.352621742656649);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.373779047216054);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.282097394125319);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.148101131915795);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.112838957650128);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.0916816530907302);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.112838957650129);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.0634719136781969);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0493670439719316);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0423146091187979);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0564194788250647);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.00705243485313299);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0.0141048697062662);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0.0282097394125324);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0.014104869706266);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0.00705243485313299);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0.00705243485313309);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.51592730438119 0
0.51592730438119 3.5
};
\addlegendentry{$\mu = 0.516$}
\addplot [semithick, black, dashed]
table {%
0.336596813861936 1
0.648443255776683 1
};
\addlegendentry{$\sigma = 0.142$}
\end{axis}

\end{tikzpicture}

### 2021-07-23 latex_files/images/chapter4/mae2_hist_t1_romc.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.02018167274615);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.39039538321726);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.64837235793467);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.83766715107148);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.03114988210955);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.20118016090059);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.42565363240795);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.83858430947187);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.02313277298727);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.49050664276752);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.58012851385441);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.63066296452957);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.34895434170502);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.09544450074894);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.92569341781802);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.63923846537203);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.52616414203809);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.21569834562492);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.07749639488345);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.95995493778819);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.85302292337607);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.59560434037885);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.66428652195943);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.3535415296862);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.45042249313527);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.1846280343355);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.19830863147962);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.979977468894081);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.941727636062617);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.705807134291806);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.614789283904501);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.496689435089055);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.456206035376906);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.425773686627776);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.449226138874808);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.309907404693004);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.224752667367446);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.192924339317897);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.0974393551692405);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.176730979433038);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.170309474651111);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.145461043103651);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0304323487491329);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0206604936462);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0217772770865351);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.503688298846839 0
0.503688298846839 3.5
};
\addlegendentry{$\mu = 0.504$}
\addplot [semithick, black, dashed]
table {%
0.323676768279387 1
0.634437489183658 1
};
\addlegendentry{$\sigma = 0.141$}
\end{axis}

\end{tikzpicture}

### 2021-07-23 latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (Bayesian Optimisation) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.67339010906917);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,2.12285439456296);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,2.41994534810638);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.13421448089961);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.60269456308737);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.61479552461991);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.51971654115006);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.94571977877467);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.78840727885186);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.70814579929938);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.89534026545558);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.97807132899427);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.47304140381032);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,2.18508791101595);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.50193553644923);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.27646251850645);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,1.84996536408459);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.10951864103732);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.94010517958197);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.70154336651221);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.88058820551382);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.72549833117864);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.29579071757466);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.50076618843174);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.46940247180661);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.16860714228384);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.37012519556019);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.767793661318775);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.885098900664694);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.65715629873569);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.82731063538692);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.813727923462651);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.801873920328749);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.459095663040085);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.672961636247559);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.561583398468614);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.382538559466961);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.340555631701063);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.294127452759942);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.239055729867026);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.283755200017783);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.176575255015415);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.283261283220537);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.258071526560991);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0775449371676099);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0155583791132463);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0.0172870879036073);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.500240536632723 0
0.500240536632723 3.5
};
\addlegendentry{$\mu = 0.500$}
\addplot [semithick, black, dashed]
table {%
0.293929386789429 1
0.656599793802562 1
};
\addlegendentry{$\sigma = 0.165$}
\end{axis}

\end{tikzpicture}

### 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_rejection.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={Rejection ABC - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0.016001600160016);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.02000200020002);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0600060006000599);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.0800080008000799);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.132013201320132);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.216021602160216);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.344034403440344);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.412041204120412);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.508050805080508);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.832083208320833);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.884088408840883);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.21612161216122);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.39213921392139);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.71217121712171);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.95219521952195);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.07620762076207);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.15221522152216);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.41624162416241);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.49224922492249);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.48824882488249);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.30823082308231);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.30423042304231);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.2042204220422);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.90819081908191);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.64816481648165);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.50815081508151);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.3001300130013);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.964096409640963);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.920092009200919);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.744074407440743);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.556055605560556);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.472047204720474);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.38003800380038);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.228022802280228);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.28002800280028);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.192019201920192);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.140014001400141);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0960096009600955);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0880088008800883);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0560056005600562);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0760076007600756);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0280028002800281);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.0360036003600358);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.0240024002400241);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0.0280028002800279);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0.00800080008000803);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0.0200020002000201);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0.00400040004000398);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0.0280028002800281);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0.00400040004000398);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0.0160016001600161);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0.0120012001200119);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0.00800080008000796);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0.00800080008000803);
\addplot [semithick, red, dashed]
table {%
0.0698328247281109 0
0.0698328247281109 3
};
\addlegendentry{$\mu = 0.070$}
\addplot [semithick, black, dashed]
table {%
-0.136927686891309 1
0.240559901293153 1
};
\addlegendentry{$\sigma = 0.172$}
\end{axis}

\end{tikzpicture}

### 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_romc.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0132189786765283);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0698717444330784);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.203477850342276);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.355023998741048);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.514595955621997);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.712880635769923);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.916830592493502);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.952710677472651);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.1953733574632);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.3804390589346);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.52584782437642);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.59477535604688);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.7604846958848);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.91108663152097);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.09048705641672);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.27366433236291);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.32654024706902);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.32811393500671);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.26296325438667);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.10968604925645);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15437878668663);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.78361790856873);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.58785112912109);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.44590447714217);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.22086710205366);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.01125186875443);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.805413486505622);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.800220316311271);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.62758674954756);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.600834054606967);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.396254622708316);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.300574396097251);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.197497836179084);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.168384609331968);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.156581949799354);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.068455425289165);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0550790778188679);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.066094893382642);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0275395389094342);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0275395389094339);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0);
\addplot [semithick, red, dashed]
table {%
0.0308218751896741 0
0.0308218751896741 3
};
\addlegendentry{$\mu = 0.031$}
\addplot [semithick, black, dashed]
table {%
-0.179871006878183 1
0.197679132295466 1
};
\addlegendentry{$\sigma = 0.172$}
\end{axis}

\end{tikzpicture}

### 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex
create tikz figures and plots

% This file was created by tikzplotlib v0.9.9.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (Bayesian Optimisation) - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0710656818158832);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.0516473550957937);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.194048417709784);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.2505503544856);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.348585934523831);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.648086654282988);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.583763447022691);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.929922090706508);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.79776958941701);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,0.987907371884556);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.05978215064711);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.2930717702704);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.55481463251829);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.01101046094927);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.35069632905971);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.34206596162857);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.80716185647295);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.71681269742808);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.65275918915001);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.45763197551135);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.24848041479702);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.8837125413398);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.80455589005721);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.47147764701123);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.35887832193294);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.990739211197902);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.834853199472739);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.568255755544842);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.531441844471339);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.415606131606362);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.366790615823915);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.350473827399395);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.220883466441019);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.136737383987297);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.185283200787522);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.154672366305157);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0558276893202577);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.179619522160829);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0788869523003634);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0133500996200616);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.0134849491111732);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.00943946437782135);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0.0140243470756202);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0.00337123727779334);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0);
\addplot [semithick, red, dashed]
table {%
0.070389896781954 0
0.070389896781954 3
};
\addlegendentry{$\mu = 0.070$}
\addplot [semithick, black, dashed]
table {%
-0.131187651225067 1
0.236045424145365 1
};
\addlegendentry{$\sigma = 0.167$}
\end{axis}

\end{tikzpicture}

### 2021-07-23 paper.tex
create tikz figures and plots

@@
 \usepackage{amsmath,amssymb,bbold, xfrac}
-% \usepackage{xcolor}
+
 \usepackage{multirow}
 \usepackage{caption}
+\usepackage{subcaption} % for subfigure
 \usepackage{float}
 
-%% maybe they should be replaced by the journals packages in the end
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
+\usepackage{pgfplots}
+\pgfplotsset{width=7cm} % ,compat=1.9}
+
+\usepackage[utf8]{inputenc}
+\DeclareUnicodeCharacter{2212}{−}
+\usepgfplotslibrary{groupplots,dateplot}
+\usetikzlibrary{patterns,shapes.arrows}
+\pgfplotsset{compat=newest}
+
+%% for algorithms
 \usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
+% \usepackage{algorithmic}
+\usepackage{algpseudocode} % algorithmicx package
 
@@
 
-\subsection{Background of Likehood-free inference methods}
+\subsection{Description of Likehood-free inference methods}
 \input{latex_files/ch2/2_2_LFI_background.tex}
@@
 
+\subsection{Engine for likelihood-free inference (ELFI)}
+\label{subsec:ELFI}
+\input{latex_files/ch3/3_1_ELFI.tex}
+
+%% -- Design Principles -----------------------------------------------------
+\section{Design principles}  % better title needed
+
+For developing an extenbable implementation of ROMC, we initially
+translate ROMC's mathematic description to an algorithmic one. The
+pure algorithmic view serves as the driver for designing the actual
+implementation at \pkg{ELFI}.
+
+\input{latex_files/graphs/romc_graph}
+
 \subsection{Algorithmic view of ROMC} % better titl
@@
 
-\section{ROMC module in ELFI}  % better title needed
+\subsection{General implementation principles}
+\label{subsec:general_design}
+\input{latex_files/ch3/3_2_1_general_design}
+
+%% -- ROMC implementation ----------------------------------------------------
+\section{Implemented functionalities}
+
 In this section, we present our implementation. At~\ref{subsec:ELFI}
@@
 with its custom modules.
-
-\subsection{Engine for likelihood-free inference (ELFI)}
-\label{subsec:ELFI}
-\input{latex_files/ch3/3_1_ELFI.tex}
-
-\subsection{Design principles} 
-\label{subsec:general_design}
-\input{latex_files/ch3/3_2_1_general_design}
 
@@
 
+%% -- Use-case illustration ----------------------------------------------------
 \section{Use-case illustration}

### 2021-07-23 latex_files/python_scripts/example_1D.py

@@ -8,7 +8,7 @@ import matplotlib.pyplot as plt
 import timeit
 import numpy as np
 import elfi
-
+import tikzplotlib as tplt
 import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 
@@ -175,6 +175,8 @@ plt.legend()
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_gt.png"), bbox_inches='tight')
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "chapter3/example_gt.tex"))
 plt.show(block=False)
 
 
@@ -252,6 +254,8 @@ plt.ylim([0, .6])
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_marginal.png"), bbox_inches='tight')
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "chapter3/example_marginal.tex"))
 plt.show(block=False)
 
 
@@ -292,6 +296,8 @@ plt.legend()
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_posterior.png"), bbox_inches='tight')
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "chapter3/example_post.tex"))
 plt.show(block=False)
 
 

### 2021-07-23 latex_files/python_scripts/example_MA2.py

@@ -6,6 +6,7 @@ import numpy as np
 import scipy.stats
 import matplotlib
 import os
+import tikzplotlib as tplt
 matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
@@ -31,8 +32,9 @@ def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, rang
     plt.legend()
     if savepath:
         plt.savefig(savepath, bbox_inches='tight')
+        # tplt.clean_figure()
+        tplt.save(savepath[:-3] + "tex")
     plt.show(block=False)
-
 
 # Set seed for reproducibility
 seed = 1
@@ -50,6 +52,8 @@ plt.plot(x["t1"], x["t2"], "bo")
 if savefig:
     plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
                 bbox_inches="tight")
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "ma2_prior_samples.tex"))
 plt.show(block=False)
 
 ####### ROMC with gradients ################
@@ -155,6 +159,7 @@ def plot_romc_posterior(title, posterior, nof_points, savefig):
     plt.colorbar()
     if savefig:
         plt.savefig(savefig, bbox_inches='tight')
+        tplt.save(savefig[:-3] + "tex")
     plt.show(block=False)
 
     

### 2021-07-23 latex_files/python_scripts/example_MA2_bo.py

@@ -6,6 +6,7 @@ import numpy as np
 import scipy.stats
 import matplotlib
 import os
+import tikzplotlib as tplt
 matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
@@ -31,6 +32,7 @@ def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, rang
     plt.legend()
     if savepath:
         plt.savefig(savepath, bbox_inches='tight')
+        tplt.save(savepath[:-3] + "tex")        
     plt.show(block=False)
 
 
@@ -51,6 +53,8 @@ plt.plot(x["t1"], x["t2"], "bo")
 if savefig:
     plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
                 bbox_inches="tight")
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "ma2_prior_samples.tex"))
 plt.show(block=False)
 
 n1 = 100
@@ -204,6 +208,7 @@ def plot_romc_posterior(title, posterior, nof_points, savefig):
     plt.colorbar()
     if savefig:
         plt.savefig(savefig, bbox_inches='tight')
+        tplt.save(savefig[:-3] + "tex")        
     plt.show(block=False)
 
 

### 2021-07-23 latex_files/ch2/2_1_LFI_intro.tex
create tikz figures and plots
applied fixes

@@
 inference methods. Unfortunately, the algorithms deployed so far
-permit the inference only at low-dimensional parametric spaces,
-i.e.\ $\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms, small refers to dimensionality $D$ less than a hundred.}.
+permit the inference only at low-dimensional parametric spaces, i.e.\
+$\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms,
+  small refers to dimensionality $D$ less than a hundred.}.
 
@@
 derivation of one (or many) parameter configuration(s) $\thetab^*$
-that \textit{describe} the data in the best manner i.e.\ generate
-samples $M_r(\thetab^*)$ that are as close as possible to the observed
-data $\data$. In our case, following the Bayesian approach, we treat
-the parameters of interest $\thetab$ as random variables and we try to
+that \textit{describe} the data best i.e.\ generate samples
+$M_r(\thetab^*)$ that are as close as possible to the observed data
+$\data$. Following the Bayesian approach, we treat the
+parameters of interest $\thetab$ as random variables and we try to
 \textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

### 2021-07-23 latex_files/ch2/2_2_LFI_background.tex
applied fixes

@@
-In simulator-based models we cannot evaluate the posterior
+In simulator-based models it is infeasible to evaluate the posterior
 $p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
@@
 the probability that the simulator will produce outputs infinitively
-close to the observations $\data$, using this configuration.
-
-For continuous cases, the probability of replicating the data
-infinitely close becomes zero. In this scenario a relaxation must be
+close to the observations $\data$, using this configuration. For
+continuous cases, the probability of perfectly replicating the
+observed data becomes zero. In this scenario a relaxation must be
 introduced by accepting simulated data that fall close to the data
 i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
-$\epsilon > 0$. The region can be
-defined as $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$
-where $d(\cdot, \cdot)$ can represent any valid distance.
+$\epsilon > 0$. The region can be defined as
+$\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$ where
+$d(\cdot, \cdot)$ can represent any valid distance.
 
-Such relaxation, introduces the approximate likelihood,
+This relaxation introduces the approximate likelihood,
 
@@
 
-and consequently the approximate posterior,
+and the approximate posterior,
 

### 2021-07-23 latex_files/ch2/2_3_OMC.tex
applied fixes

@@
-Before we define the likelihood approximation as introduced in the
-OMC, approach lets define the indicator function based on
+For describing the likelihood approximation introduced in OMC, it is
+convenient to firstly define the indicator function based on
 $\region(\yb)$. The indicator function $\indicator{\region(\yb)}(\xb)$
@@
 \noindent
-Based on equation~\eqref{eq:approx_posterior} and the indicator
-function as defined above~\eqref{eq:indicator}, we can approximate the
-likelihood as:
+Based on equation~\eqref{eq:approx_likelihood} and the indicator
+function~\eqref{eq:indicator}, we can approximate the likelihood as:
 
@@
 %
-This approach is quite intuitive; approximating the likelihood of a
+The procedure is quite intuitive; approximating the likelihood of a
 specific $\thetab$ requires sampling from the data generator and count
@@
 quite inconvenient from a computational point-of-view. For this
-reason, we choose to approximate the integral as in equation
-\eqref{eq:alt_view}; the nuisance variables are sampled once
-$\vb_i \sim p(\vb)$ and we count the fraction of samples that lie
-inside the area using the deterministic simulators
+reason, OMC approximates the integral as in equation
+\eqref{eq:alt_view}; the nuisance variables, $\vb_i \sim p(\vb)$, are
+sampled once and we count the fraction of samples that lie inside the
+area using the deterministic simulators
 $M_d(\thetab, \vb_i) \: \forall i$. Hence, the evaluation
@@
 
-The posterior approximation in \eqref{eq:approx_posterior_omc} does not
-provide any obvious way for drawing samples. In fact, the set
-$\mathcal{S}_i = \{ \thetab: M_d(\thetab, \vb_i) \in \region(\data) \}$ can
-represent any arbitrary shape in the D-dimensional Euclidean space; it
-can be non-convex, can contain disjoint sets of $\thetab$ etc. We need
-some further simplification of the posterior for being able to draw
-samples from it.
+The posterior approximation in \eqref{eq:approx_posterior_omc} does
+not provide any convenient way for drawing samples.\footnote{Weighted
+  sampling could be performed in a straightforward fashion with
+  importance sampling, using the prior as the proposal distribution
+  $\thetab_i \sim p(\thetab)$. However, this approach has the same
+  drawbacks as ABC rejection sampling; when the prior is wide or the
+  dimensionality $D$ is high, drawing a sample with non-zero weight is
+  rare, leading to either poor Effective Sample Size (ESS) or huge
+  execution time.} In fact, the set
+$\mathcal{S}_i = \{ \thetab: M_d(\thetab, \vb_i) \in \region(\data)
+\}$ can represent any arbitrary shape in the D-dimensional Euclidean
+space; it can be non-convex, can contain disjoint sets of $\thetab$
+etc. We need some further simplification of the posterior for being
+able to draw samples from it.
 
-As a side-note, weighted sampling could be performed in a
-straightforward fashion with importance sampling. Using the prior as
-the proposal distribution $\thetab_i \sim p(\thetab)$ and we can
-compute the weight as
-$w_i = \frac{L_{d,\epsilon}(\thetab_i)}{p(\thetab_i)}$, where
-$L_{d,\epsilon}(\thetab_i)$ is computed with the expression
-\eqref{eq:approx_likelihood_omc}. This approach has the same drawbacks as
-ABC rejection sampling; when the prior is wide or the dimensionality
-$D$ is high, drawing a sample with non-zero weight is rare, leading to
-either poor Effective Sample Size (ESS) or huge execution time.
-
-The OMC proposes a quite drastic simplification of the posterior; it
+OMC proposes a quite drastic simplification of the posterior; it
 squeezes all regions $\mathcal{S}_i$ into a single point

### 2021-07-23 latex_files/ch2/2_5_ROMC_algorithmic.tex
create tikz figures and plots
applied fixes

@@
-Here we present ROMC as a sequence of algorithmic steps. The reader
-may think of the algorithmic description as the backbone of the
-implementation that will follow. At a high-level, ROMC can be split
-into the training and the inference part.
+For designing an extendable implementation of ROMC, we firstly define
+ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
+serves as the driver for the implementation at \pkg{ELFI} that
+follows. At a high level, ROMC can be split into the training and the
+inference part. In general terms, the training part covers all the
+appropriate steps for estimating the proposal regions and the
+inference part for getting weighted
+samples. Algorithm~\ref{alg:romc_algorithm} defines formally ROMC;
+Steps 2-10 (before the vertical line) form the training part, while
+steps 12 and 13 the inference part.
+
+\begin{algorithm}[!ht]
+	\caption{ROMC}\label{alg:romc_algorithm}
+	\begin{algorithmic}[1]
+    \Procedure{ROMC}{$M_r(\thetab)$}
+    \For{$i \gets 1 \textrm{ to } n_1$}
+      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define deterministic distance function}
+      \State Obtain $\thetab_i^*, d_i^*$ \Comment{Solve optimization problem}
+    \EndFor
+    \State $\text{L} \gets \{d_{i, i = 1, ..., n_1}: d_i^* < \epsilon \} $ \Comment{List with filtered solutions}
+    \For{$i \gets 1 \textrm{ to } len(L)$}
+      \State Construct $q_i$ \Comment{Compute proposal area}
+      \State Compute $\Tilde{d}_i$ \Comment{Fit surrogate model in $q_i$}
+      \EndFor
+    \\\hrulefill
+    \State Draw $\{w_{ij}, \thetab_{ij}\}$ \Comment{Draw $n_2$ samples per $q_i$}
+    \State Draw $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}    
+    \EndProcedure
+	\end{algorithmic}
+\end{algorithm}
 
@@
 At the training (fitting) part, the goal is the estimation of the
-proposal regions $q_i$. The tasks are (a) sampling the nuisance
-variables $\vb_i \sim p(\vb)$ (b) defining the optimisation problems
-$\min_{\thetab} \: g_i(\thetab)$ (c) obtaining $\thetab_i^*$ (d)
-checking whether $d_i^* \leq \epsilon$ and (e) building the bounding
-box for obtaining the proposal region $q_i$. If gradients are
-available, using a gradient-based method is advised for obtaining
-$\thetab_i^*$ much faster. Providing $\nabla_{\thetab} g_i$ in
-closed-form provides an upgrade in both accuracy and efficiency; If
-closed-form description is not available, approximate gradients with
-finite-differences requires two evaluations of $g_i$ for \emph{every}
-parameter $\theta_d$, which can only work in low-dimensional
-problems. When gradients are not available, the Bayesian optimisation
-paradigm exists as an alternative solution. In this scenario, the
-training part becomes slower due to fitting of the surrogate model and
-the blind optimisation steps. Nevertheless, the subsequent task of
-computing the proposal region $q_i$ becomes faster since $\hat{d}_i$
-can be used instead of $g_i$. The algorithms~\ref{alg:training_GB}
-and~\ref{alg:training_GP} present the above procedure.
+proposal regions $q_i$. The tasks are; (a) sampling the nuisance
+variables $\vb_i \sim p(\vb)$ for obtaining the deterministic
+functions $g_i(\thetab)$, (b) defining the optimisation problems
+$\min_{\thetab} \: g_i(\thetab)$, (c) obtaining $\thetab_i^*$ and
+$d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and (e) building
+the bounding box for obtaining the proposal region $q_i$. If gradients
+are available, using a gradient-based method is advised for obtaining
+$\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
+in closed-form, the procedure is upgraded in terms of both accuracy
+and efficiency; If closed-form description is not available, gradients
+approximation with finite-differences requires two evaluations of
+$g_i$ for \emph{each} parameter $\theta_d$, which can only work in
+low-dimensional problems. If $g_i(\thetab)$ is not differentiable, the
+Bayesian Optimisation can be used as an alternative solution. In this
+scenario, the training part becomes slower due to fitting of the
+surrogate model and the blind optimisation
+steps. Algorithm~\ref{alg:training_GB} defines formally the steps of a
+gradient-based optimiser and algorithm~\ref{alg:training_GP} the steps
+of the Bayesian Optimisation. After obtaning the optimal points
+$\thetab^*$, the training part gets completed by estimating the
+proposal regions. Algorithm~\ref{alg:region_construction} describes
+the line search approach for finding the region's boundaries.
 
@@
 $ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
-expectation $E_{\thetab|\data}[h(\thetab)]$ and (c) evaluating the
+expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
 unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
@@
 function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
-checking if it lies inside the acceptance region. If so the
-corresponding weight is computed as in \eqref{eq:sampling}. As before,
-if a surrogate model $\hat{d}$ is available, it can be utilised for
-evaluating the indicator function. At the sampling task, the
-computational benefit of using the surrogate model is more valuable
-compared to the evaluation of the posterior, because the indicator
-function must be evaluated for a total of $n_1 \times n_2$ points.
+checking if it lies inside the acceptance region. If this is the case,
+the corresponding weight is computed as in \eqref{eq:sampling}. As
+before, if a surrogate model $\hat{d}$ is available, it can be
+utilised for evaluating the indicator function.
 
 \noindent
-Evaluating the unnormalised posterior requires only the
-deterministic functions $g_i$ and the prior distribution $p(\thetab)$;
-there is no need for solving the optimisation problems and building
-the proposal regions. The evaluation requires iterating over all $g_i$
-and evaluating the distance from the observed data. In contrast, using
-the GP approach, the optimisation part should be performed first for
-fitting the surrogate models $\hat{d}_i(\thetab)$ and evaluate the
-indicator function on them. This provides an important speed-up,
-especially when running the simulator is computationally
-expensive.
-
-\noindent
-In summary, we can state that the choice of using a Bayesian
-optimisation approach provides a significant speed-up in the inference
-part with the cost of making the training part slower and a possible
-approximation error. It is typical in many Machine-Learning use cases,
-being able to provide enough time and computational resources for the
-training phase, but asking for efficiency in the inference
-part.
+Evaluating the unnormalised posterior requires the deterministic
+functions $g_i$ and the prior distribution $p(\thetab)$\footnote{There
+  is no need for solving the optimisation problems and building the
+  proposal regions.}. The evaluation requires iterating over all $g_i$
+and evaluating the distance from the observed data.
 
@@
 
-% \begin{minipage}{0.46\textwidth}
-% \begin{algorithm}[H]
-%     \centering
-%     \caption{Evaluate unnormalised posterior - Gradient approach. Requires $g_i(\theta), p(\theta)$}\label{alg:posterior_GB}
-%     \begin{algorithmic}[1]
-%       \State $k \leftarrow 0$
-%         \For {$i \gets 1 \textrm{ to } n_1$}
-%           \If {$g_i(\theta) > \epsilon$}
-%             \State $k \leftarrow k + 1$
-%           \EndIf
-%           \EndFor
-%       \Return{$kp(\theta)$}
-%     \end{algorithmic}
-% \end{algorithm}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.46\textwidth}
-% \begin{algorithm}[H]
-%     \centering
-%     \caption{Evaluate unnormalised posterior - GP approach. Requires $\hat{d}_i(\theta), p(\theta)$}\label{alg:posterior_GP}
-%     \begin{algorithmic}[1]
-%       \State $k \leftarrow 0$
-%         \For {$i \gets 1 \textrm{ to } n_1$}
-%           \If {$d_i(\theta) > \epsilon$}
-%             \State $k \leftarrow k + 1$
-%           \EndIf
-%           \EndFor
-%       \Return{$kp(\theta)$}
-%     \end{algorithmic}
-% \end{algorithm}
-% \end{minipage}
-
-% \begin{minipage}{0.46\textwidth}
 \begin{algorithm}[H]
@@
 \end{algorithm}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.46\textwidth}
-% \begin{algorithm}[H]
-%     \centering
-%     \caption{Sampling - GP approach. Requires $\hat{d}_i(\theta), p(\theta), q_i$}\label{alg:sampling_GP}
-%     \begin{algorithmic}[1]
-%       \For {$i \gets 1 \textrm{ to } n_1$}
-%       \For {$j \gets 1 \textrm{ to } n_2$}
-%           \State $\theta_{ij} \sim q_i$
-%           \If {$\hat{d}_i(\theta_{ij}) > \epsilon$}
-%             \State Reject $\theta_{ij}$
-%           \Else {}
-%             \State $w_{ij} = \frac{p(\theta_{ij})}{q(\theta_{ij})}$
-%             \State Accept $\theta_{ij}$, with weight $w_{ij}$
-%           \EndIf
-%       \EndFor
-%       \EndFor
-%     \end{algorithmic}
-% \end{algorithm}
-% \end{minipage}

### 2021-07-23 latex_files/ch3/3_2_1_general_design.tex
applied fixes

@@
 The overview of our implementation is illustrated in
-figure~\ref{fig:romc_overview}; one may interpret
-figure~\ref{fig:romc_overview} as a depiction of the main class and
+figure~\ref{graph:romc_overview}; one may interpret
+figure~\ref{graph:romc_overview} as a depiction of the main class and
 the ellipses are the main routines of the class. Following
@@
 
-Figure~\ref{fig:romc_overview} splits ROMC into the training, the
-inference and the evaluation part. The training part includes all the
-steps until the computation of the proposal regions i.e. sampling the
-nuisance variables, defining the optimization problems, solving them,
+\input{latex_files/graphs/romc_graph}
+
+In figure~\ref{graph:romc_overview}, it can be easily observed that the
+implementation follows consistently the algorithmic description
+described above, which follows the partition into training, inference
+and evaluation. The training part includes all the steps until the
+computation of the proposal regions i.e. sampling the nuisance
+variables, defining the optimization problems, solving them,
 constructing the regions and fitting local surrogate models. The
@@
   discrepancy between the two approximations.}
-
-% \begin{figure}[h]
-%     \begin{center}
-% \includegraphics[width=0.5\textwidth]{./latex_files/graphs/ROMC.png}
-%     \end{center}
-%     \caption[Overview of the ROMC implementation.]{Overview of the
-%       ROMC implementation. The training part follows a sequential
-%       pattern; the functions in the green ellipses must be called in a
-%       sequential fashion for completing the training part and define
-%       the posterior distribution. The functions in blue ellipses are
-%       the functionalities provided to the user.}
-%     \label{fig:romc_overview}
-% \end{figure}
 
@@
 
-% \begin{figure}[!h]
-%     \begin{center}
-% \includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
-%     \end{center}
-%   \caption[Ground-truth posterior distribution of the simple 1D
-% example.]{Ground-truth posterior distribution of the simple 1D
-% example.}
-%   \label{fig:example_gt}
-% \end{figure}
-
 In the following code snippet, we code the model at \pkg{ELFI}.

### 2021-07-23 paper.tex
applied fixes

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{A Parallelized and Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
-\Plaintitle{A Parallelized and Extendable Python Implementation of ROMC}
-\Shorttitle{A Parallelized and Extendable \proglang{Python} Implementation of ROMC}
+\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
+\Plaintitle{An Extendable Python Implementation of ROMC}
+\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}
 
@@
 
-For developing an extenbable implementation of ROMC, we initially
-translate ROMC's mathematic description to an algorithmic one. The
-pure algorithmic view serves as the driver for designing the actual
-implementation at \pkg{ELFI}.
-
-\input{latex_files/graphs/romc_graph}
-
-\subsection{Algorithmic view of ROMC} % better titl
 \input{latex_files/ch2/2_5_ROMC_algorithmic.tex}

### 2021-07-25 latex_files/ch1/1_1_motivation.tex
fixed warnings

@@
 data generating mechanism \cite{Gutmann2016}. The key characteristic
-of these models is that although we can sample (simulate) data points,
-we cannot evaluate the likelihood of a specific set of observations
-$\data$. Formally, a simulator-based model is described as a
+of these models is that although it is feasible to sample (simulate) data points, it is impossible to evaluate the likelihood. Formally, a simulator-based model is a
 parameterised family of probability density functions
 $\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
-unknown or intractable to evaluate. Whereas evaluating
-$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
-feasible. Practically, a simulator can be understood as a black-box
+unknown or intractable to evaluate. Practically, a simulator can be understood as a black-box
 machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the

### 2021-07-25 latex_files/ch3/3_2_1_general_design.tex
fixed warnings

@@
+\input{latex_files/graphs/romc_graph}
+
 The overview of our implementation is illustrated in
-figure~\ref{graph:romc_overview}; one may interpret
-figure~\ref{graph:romc_overview} as a depiction of the main class and
+figure~\ref{fig:romc_overview}; one may interpret
+figure~\ref{fig:romc_overview} as a depiction of the main class and
 the ellipses are the main routines of the class. Following
@@
 
-\input{latex_files/graphs/romc_graph}
-
-In figure~\ref{graph:romc_overview}, it can be easily observed that the
+In figure~\ref{fig:romc_overview}, it can be easily observed that the
 implementation follows consistently the algorithmic description

### 2021-07-25 latex_files/ch3/3_2_3_inference.tex
fixed warnings

@@
 In the following code snippet, we perform weighted sampling from the
-ROMC approximate posterior. Afterwards, we used some of ELFI's
+ROMC approximate posterior. Afterwards, we use some of ELFI's
 built-in tools to get a summary of the obtained samples. In figure
-\ref{fig:example_sampling}, we observe the histogram of the weighted
+\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
 samples and the acceptance region of the first deterministic function
@@
 \code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
-point $\theta$ using expression \eqref{eq:aprox_posterior}. The
+point $\theta$ using expression \eqref{eq:approx_posterior_omc}. The
 \code{romc.eval_posterior(theta)} approximates the partition function
@@
 utility can provide a nice plot of the approximate posterior as
-illustrated in figure~\ref{fig:approx_posterior}. We observe that the
+illustrated in figure~\ref{fig:running_example_romc_inference}. We observe that the
 approximation is quite close to the ground-truth posterior.

### 2021-07-25 latex_files/ch4/4_2_ma2.tex
fixed warnings

@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \begin{center}
-      % \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
-      \input{./latex_files/images/chapter4/ma2_region_3.tex}
-      \input{./latex_files/images/chapter4/ma2_region_3_bo.tex}
-      % \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
+        % \input{./latex_files/images/chapter4/ma2_region_3.tex}
+        % \input{./latex_files/images/chapter4/ma2_region_3_bo.tex}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
     \end{center}
   \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimisation problem.
-In the left figure the region obtained with gradient-based optimiser and in the right one with Bayesia Optimisation.}
+In the left figure the region obtained with gradient-based optimiser and in the right one with Bayesian Optimisation.}
   \label{fig:ma2_5}
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \begin{center}

### 2021-07-27 latex_files/ch2/2_1_LFI_intro.tex
minor improvements

@@
-An implicit or simulator-based model is a parameterised stochastic
-data generating mechanism \cite{Gutmann2016}. The key characteristic
-of these models is that although we can sample (simulate) data points,
-we cannot evaluate the likelihood of a specific set of observations
-$\data$. Formally, a simulator-based model is described as a
-parameterised family of probability density functions
-$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
-unknown or intractable to evaluate. Whereas evaluating
-$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
-feasible. Practically, a simulator can be understood as a black-box
-machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
-  \textit{random} simulator. In the next chapters we will introduce
-  $M_d$ witch stands for the \textit{deterministic} simulator.} that
-given a set of parameters $\thetab$, produces samples $\yb$ in a
-stochastic manner i.e.\ $M_r(\thetab) \rightarrow \yb$.
-
-Simulator-based models are particularly captivating due to the
-modelling freedom they provide; any physical process that can be
-conceptualised as a computer program of finite steps can be modelled
-as a simulator-based model without any compromise. The modelling
-freedom includes any amount of hidden (unobserved) internal variables
-or rule-based decisions. Hence, implicit models are often used to
-model physical phenomena in many fiels such as biology, neuroscience
-etc. This degree of freedom comes at a cost; performing the inference
-is particularly demanding from both the computational and the
-mathematical perspective. The class of methods that try to approximate
-the posterior distribution in these cases are called likelihood-free
-inference methods. Unfortunately, the algorithms deployed so far
-permit the inference only at low-dimensional parametric spaces, i.e.\
-$\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms,
-  small refers to dimensionality $D$ less than a hundred.}.
+An implicit or simulator-based\footnote{The terms can be used
+  interchangeably} model is a parameterised stochastic data generating
+mechanism \cite{Gutmann2016}. The key characteristic of these models
+is that although it is feasible to sample (simulate) data points, it
+is impossible to evaluate the likelihood. Formally, a simulator-based
+model is a parameterised family of probability density functions
+$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$ whose closed-form is either
+unknown or intractable to evaluate. Practically, a simulator can be
+understood as a black-box machine $M_r$\footnote{The subscript $r$ in
+  $M_r$ indicates the \textit{random} simulator. In the next chapters
+  we will introduce $M_d$ witch stands for the \textit{deterministic}
+  simulator.} that given a set of parameters $\thetab$, produces
+samples $\yb$ in a stochastic manner i.e.\
+$M_r(\thetab) \rightarrow \yb$. Simulator-based models are
+particularly captivating due to the modelling freedom they provide;
+any physical process, that can be conceptualised as a computer program
+of finite steps, can be represented as a simulator-based model. For
+this reason, they are quite popular in many fiels such as biology,
+neuroscience etc. This degree of freedom comes at a cost; performing
+the inference is particularly demanding from both computational and
+mathematical perspective.
 
@@
 $M_r(\thetab^*)$ that are as close as possible to the observed data
-$\data$. Following the Bayesian approach, we treat the
-parameters of interest $\thetab$ as random variables and we try to
-\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.
+$\data$. Following the Bayesian approach, we treat the parameters of
+interest $\thetab$ as random variables and we try to \textit{infer} a
+posterior distribution $p(\thetab|\data)$ on them. The class of
+methods for approximate the posterior distribution in implicit models
+is called likelihood-free inference methods. Unfortunately, the
+algorithms deployed so far permit the inference only at
+low-dimensional parametric spaces, i.e.\ $\thetab \in \mathbb{R}^D$
+where $D$ is small\footnote{In broad terms, small refers to
+  dimensionality $D$ less than a hundred.}.

### 2021-07-27 latex_files/ch2/2_3_OMC.tex
minor improvements

@@
 For describing the likelihood approximation introduced in OMC, it is
-convenient to firstly define the indicator function based on
-$\region(\yb)$. The indicator function $\indicator{\region(\yb)}(\xb)$
-returns 1 if $\xb \in \region(\yb)$ and 0 otherwise. If
-$d(\cdot,\cdot)$ is a formal distance, due to symmetry
+convenient to firstly define the indicator function
+$\indicator{\region(\yb)}(\xb)$, that returns 1 if
+$\xb \in \region(\yb)$ and 0 otherwise. If $d(\cdot,\cdot)$ is a
+formal distance, due to symmetry
 $\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so

### 2021-07-27 latex_files/ch3/3_2_2_training.tex
fixed warnings
minor improvements

@@
 ------------------------------ python snippet ------------------------------
-  # Trainining (fitting) part
+  # Training (fitting) part
   n1 = 500 # number of optimisation problems
@@
 \end{Code}
-
-In figure \ref{fig:example_training_hist} we illustrate the
-distribution of the distances obtained and the acceptance area of the
-first optimisation problem. We observe that most optimal points
-produce almost zero distance.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \begin{center}    
@@
     distances and visualisation of a specific region.}
-  \label{fig:example_training_hist}
+  \label{fig:running_example_romc_inference}
 \end{figure}

### 2021-07-27 latex_files/ch2/2_1_implicit_models.tex
michael updates

%% Not used; Michael combined the text with the LFI intro

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic
of these models is that we can sample (simulate) data points,
but that we cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
data generating machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
produces samples $\yb$ in a stochastic manner for any given a set of parameters $\thetab$.% i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models provide considerable modelling freedom; any physical process that can be
conceptualised as a computer program of finite steps can be modelled
as a simulator-based model without any compromise. The modelling
freedom allows for any amount of hidden (unobserved) internal variables
or rule-based decisions. Hence, implicit models are often used to
model physical phenomena in the natural sciences such as genetics, epidemiology, or neuroscience, to name a few examples.
The modelling freedom, however, makes parameter inference difficult.

% ;performing the inference
% is particularly demanding from both the computational and the
% mathematical perspective. The class of methods that try to approximate
% the posterior distribution in these cases are called likelihood-free
% inference methods. Unfortunately, the algorithms deployed so far
% permit the inference only at low-dimensional parametric spaces, i.e.\
% $\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms,
%   small refers to dimensionality $D$ less than a hundred.}.

\subsubsection*{\textit{Goal of implicit models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. Following the Bayesian approach, we treat the
parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

### 2021-07-27 latex_files/ch2/2_2_LFI_background.tex
minor improvements
michael updates

@@
-In simulator-based models it is infeasible to evaluate the posterior
-$p(\thetab|\data) \propto L(\thetab)p(\thetab)$, due to the
-intractability of the likelihood $L(\thetab) = p(\data|\thetab)$. The
-following equation allows incorporating the simulator in the place of
-the likelihood,
+An implicit or simulator-based model is a parameterised stochastic
+data generating mechanism. The key characteristic
+of these models is that we can sample (simulate) data points,
+but that we cannot evaluate the likelihood of a specific set of observations
+$\data$. Formally, a simulator-based model is described as a
+parameterised family of probability density functions
+$\{ p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either
+unknown or computationally intractable to evaluate. Whereas evaluating
+$p(\yb|\thetab)$ is intractable, sampling is
+feasible. Practically, a simulator can be understood as a black-box
+data generating machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
+  \textit{random} simulator. In the next chapters we will introduce
+  $M_d$ witch stands for the \textit{deterministic} simulator.} that
+produces samples $\yb$ in a stochastic manner for any given a set of parameters $\thetab$.% i.e.\ $M_r(\thetab) \rightarrow \yb$.
 
+Simulator-based models provide considerable modelling freedom; any physical process that can be
+conceptualised as a computer program of finite steps can be modelled
+as a simulator-based model without any compromise. The modelling
+freedom allows for any amount of hidden (unobserved) internal variables
+or rule-based decisions. Hence, implicit models are often used to
+model physical phenomena in the natural sciences such as genetics, epidemiology, or neuroscience, to name a few examples.
+The modelling freedom, however, makes parameter inference difficult.
+
+The main difficulty in parameter inference with implicit models is that the likelihood function
+$L(\thetab) = p(\data|\thetab)$ is intractable.
+The methods that we consider below resort to the basic definition of the likelihood as the (rescaled) probability for a parameter of the model to generate data $\yb$ that is similar to the observed data $\data$.
+More formally, the likelihood $L(\thetab)$ equals
 \begin{equation} \label{eq:likelihood}
-  L(\thetab) =
-  \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
-  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in B_{d,\epsilon}(\data))
+  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
+  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in \region(\data))
 \end{equation}
 %
-where $c_\epsilon$ is a proportionality factor dependend on
-$\epsilon$ and $B_{d,\epsilon}(\data)$ is a region around $\data$.
+where $c_\epsilon$ is a proportionality factor that depends on
+$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$ that is defined via a distance function $d$, i.e.\ $\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$.
 
-Intuitively, equation~\ref{eq:likelihood} describes that the
-likelihood of a specific parameter configuration $\thetab$ is equal to
-the probability that the simulator will produce outputs infinitively
-close to the observations $\data$, using this configuration. For
-continuous cases, the probability of perfectly replicating the
-observed data becomes zero. In this scenario a relaxation must be
-introduced by accepting simulated data that fall close to the data
-i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
-$\epsilon > 0$. The region can be defined as
-$\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$ where
-$d(\cdot, \cdot)$ can represent any valid distance.
-
-This relaxation introduces the approximate likelihood,
-
+The basic characterisation of the likelihood in \eqref{eq:likelihood} highlights two sources of intractability: The first is the computation of the probability $\Pr(M_r(\thetab) \in \region(\data))$, the second is the limit of $\epsilon \to 0$.
+Whilst the probability can be approximated with samples, this becomes computationally infeasible if $\epsilon$ is too small.
+Hence, a large class of inference methods for implicit models work with $\epsilon >0$, which leads to the approximate likelihood function  $L_{d, \epsilon}(\thetab)$,
 \begin{equation} \label{eq:approx_likelihood}
-  L_{d, \epsilon}(\thetab) = Pr(\yb \in \region(\data)), \text{where  } \epsilon > 0
+  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)), \quad \text{where  } \epsilon > 0.
+\end{equation}
+Most often in likelihood-free inference, we are interested in the posterior distribution of the parameters $\thetab$ given some observed data.
+The approximate likelihood function in \eqref{eq:approx_likelihood} implies the approximate posterior  $p_{d,\epsilon}(\thetab|\data)$,
+\begin{equation} \label{eq:approx_posterior}
+  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
 \end{equation}
 
-and the approximate posterior,
+% likelihood of a specific parameter configuration $\thetab$ is equal to
+% the probability that the simulator will produce outputs infinitively
+% close to the observations $\data$, using this configuration. For
+% continuous cases, the probability of perfectly replicating the
+% observed data becomes zero. In this scenario a relaxation must be
+% introduced by accepting simulated data that fall close to the data
+% i.e. in a region $B_{d,\epsilon}(\data)$ around $\data$, where
+% $\epsilon > 0$. The region can be defined as
+% $\region (\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$ where
+% $d(\cdot, \cdot)$ can represent any valid distance.
 
-\begin{equation} \label{eq:approx_posterior}
-  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab)
-\end{equation}
+% This relaxation introduces the approximate likelihood,
+
+% and the approximate posterior,

### 2021-07-27 paper.tex
fixed warnings
minor improvements
michael updates

@@
 %% -- LaTeX packages and custom commands ---------------------------------------
+
+\usepackage[usenames,dvipsnames]{xcolor}
 
@@
 
+%% TikZ packages
 \usepackage{tikz}
@@
 \usepackage{pgfplots}
-\pgfplotsset{width=7cm} % ,compat=1.9}
-
+\pgfplotsset{width=6.75cm, compat=newest}
 \usepackage[utf8]{inputenc}
@@
 \usetikzlibrary{patterns,shapes.arrows}
-\pgfplotsset{compat=newest}
 
@@
 \usepackage{algorithm}
-% \usepackage{algorithmic}
 \usepackage{algpseudocode} % algorithmicx package
@@
 \section{Background}
-\input{latex_files/ch2/2_1_LFI_intro.tex}
+We first give a short introduction to implicit parametric models and likelihood-free inference.
+We then provide the necessary background on two of these methods: Optimisation Monte Carlo (OMC) and its recent robust generalisation, robust OMC (ROMC).
+%\input{latex_files/ch2/2_1_implicit_models}
 
-\subsection{Description of Likehood-free inference methods}
+\subsection{Implicit models and likelihood-free inference}
 \input{latex_files/ch2/2_2_LFI_background.tex}

### 2021-07-29 latex_files/ch2/2_2_LFI_background.tex
added references for LFI background

@@
 An implicit or simulator-based model is a parameterised stochastic
-data generating mechanism. The key characteristic
-of these models is that we can sample (simulate) data points,
-but that we cannot evaluate the likelihood of a specific set of observations
-$\data$. Formally, a simulator-based model is described as a
-parameterised family of probability density functions
-$\{ p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either
-unknown or computationally intractable to evaluate. Whereas evaluating
-$p(\yb|\thetab)$ is intractable, sampling is
-feasible. Practically, a simulator can be understood as a black-box
-data generating machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
-  \textit{random} simulator. In the next chapters we will introduce
-  $M_d$ witch stands for the \textit{deterministic} simulator.} that
-produces samples $\yb$ in a stochastic manner for any given a set of parameters $\thetab$.% i.e.\ $M_r(\thetab) \rightarrow \yb$.
+data generating mechanism. The key characteristic of these models is
+that we can sample, or simulate, data points, but that we cannot
+evaluate the likelihood of a specific set of observations
+$\data$. Formally, a simulator-based model is a parameterised family
+of probability density functions $\{ p(\yb|\thetab) \}_{\thetab}$,
+whose closed-form is either unknown or computationally intractable to
+evaluate. Whereas evaluating $p(\yb|\thetab)$ is intractable, sampling
+is feasible. Practically, a simulator can be understood as a black-box
+data generating machine $M_r$\footnote{The subscript $r$ in $M_r$
+indicates the \textit{random} simulator. In the next chapters we will
+introduce $M_d$ which stands for the \textit{deterministic}
+simulator.} that generates samples $\yb$ in a stochastic manner for
+any given a set of parameters $\thetab$.% i.e.\
+$M_r(\thetab) \rightarrow \yb$.
 
-Simulator-based models provide considerable modelling freedom; any physical process that can be
-conceptualised as a computer program of finite steps can be modelled
-as a simulator-based model without any compromise. The modelling
-freedom allows for any amount of hidden (unobserved) internal variables
-or rule-based decisions. Hence, implicit models are often used to
-model physical phenomena in the natural sciences such as genetics, epidemiology, or neuroscience, to name a few examples.
-The modelling freedom, however, makes parameter inference difficult.
+Simulator-based models provide considerable modelling freedom; any
+physical process that can be conceptualised as a computer program of
+finite steps can be modelled as a simulator-based model without any
+compromise. The modelling freedom allows for any amount of hidden
+(unobserved) internal variables or rule-based decisions. Hence,
+implicit models are often used to model physical phenomena in the
+natural sciences such as genetics, epidemiology, or neuroscience, to
+name a few examples. Further background on simulator-based models and
+example applications can be found in the articles
+by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.
 
-The main difficulty in parameter inference with implicit models is that the likelihood function
-$L(\thetab) = p(\data|\thetab)$ is intractable.
-The methods that we consider below resort to the basic definition of the likelihood as the (rescaled) probability for a parameter of the model to generate data $\yb$ that is similar to the observed data $\data$.
-More formally, the likelihood $L(\thetab)$ equals
+The modelling freedom of simulator-based models, however, comes at the
+price of difficulties in inferring their parameters. The main
+difficulty is that the likelihood function $L(\thetab) =
+p(\data|\thetab)$ is intractable. To better see the sources of the
+intractability, and to address them, we go back to the basic
+characterisation of the likelihood as the (rescaled) probability for a
+parameter of the model to generate data $\yb$ that is similar to the
+observed data $\data$. More formally, the likelihood $L(\thetab)$
+equals
 \begin{equation} \label{eq:likelihood}
@@
 \end{equation}
-%
 where $c_\epsilon$ is a proportionality factor that depends on
-$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$ that is defined via a distance function $d$, i.e.\ $\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon \}$.
+$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$
+that is defined via a distance function $d$, i.e.\ $\region(\data)
+:= \{\yb: d(\yb, \data) \leq \epsilon \}$.
 
-The basic characterisation of the likelihood in \eqref{eq:likelihood} highlights two sources of intractability: The first is the computation of the probability $\Pr(M_r(\thetab) \in \region(\data))$, the second is the limit of $\epsilon \to 0$.
-Whilst the probability can be approximated with samples, this becomes computationally infeasible if $\epsilon$ is too small.
-Hence, a large class of inference methods for implicit models work with $\epsilon >0$, which leads to the approximate likelihood function  $L_{d, \epsilon}(\thetab)$,
+The basic characterisation of the likelihood in \eqref{eq:likelihood}
+highlights two sources of intractability: The first is the computation
+of the probability $\Pr(M_r(\thetab) \in \region(\data))$, the second
+is the limit of $\epsilon \to 0$.  Whilst the probability can be
+approximated with samples, this becomes computationally infeasible if
+$\epsilon$ is too small. Hence, a large class of inference methods
+work with $\epsilon >0$, which leads to the approximate likelihood
+function $L_{d, \epsilon}(\thetab)$,
 \begin{equation} \label{eq:approx_likelihood}
@@
 \end{equation}
+
+The approximation in \eqref{eq:approx_likelihood} is by no means the
+only strategy to deal with the intractabilities of the likelihood
+function in \eqref{eq:likelihood}. Other strategies include modelling
+the (stochastic) relationship between $\thetab$ and $\yb$, and its
+reverse, or framing likelihood-free inference as a ratio estimation
+problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
+Thomas2016, Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both OMC and robust OMC that we introduce next
+are based on the approximation in \eqref{eq:approx_likelihood}.
 

### 2021-07-29 paper.tex
extended section intro

@@
 \section{Background}
-We first give a short introduction to implicit parametric models and likelihood-free inference.
-We then provide the necessary background on two of these methods: Optimisation Monte Carlo (OMC) and its recent robust generalisation, robust OMC (ROMC).
+We first give a short introduction to implicit parametric models and
+likelihood-free inference. We then focus on the methods relevant for
+this paper: first the earlier Optmisation Monte Carlo (OMC) and then
+its recent robust generalisation, robust OMC, for which we provide the
+software package as part of the Engine for likelihood-free inference
+(ELFI), a larger inference platform that we briefly introduce at the
+end of this background section.
+
 %\input{latex_files/ch2/2_1_implicit_models}

### 2021-08-18 latex_files/ch2/2_2_LFI_background.tex
Michael's edits on overleaf, LFI background and OMC

@@
 evaluate. Whereas evaluating $p(\yb|\thetab)$ is intractable, sampling
-is feasible. Practically, a simulator can be understood as a black-box
-data generating machine $M_r$\footnote{The subscript $r$ in $M_r$
-indicates the \textit{random} simulator. In the next chapters we will
-introduce $M_d$ which stands for the \textit{deterministic}
-simulator.} that generates samples $\yb$ in a stochastic manner for
-any given a set of parameters $\thetab$.% i.e.\
-$M_r(\thetab) \rightarrow \yb$.
+is feasible. Practically, a simulator can be understood as a random black-box
+machine (computer code) $M_r$ that generates samples $\yb$ in a stochastic manner for
+any given a set of parameters $\thetab$.
+We can isolate the randomness by introducing the stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator becomes a deterministic mapping $M_d$ that maps $(\thetab, \vb)$ to the data $\yb$.
+The distribution $p(\vb)$ is determined by the random draws performed when running $M_r$.
+Importantly, we can fix the value of $\vb$ while changing $\thetab$ by fixing the random seed used with $M_r$.
 
@@
 problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
-Thomas2016, Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both OMC and robust OMC that we introduce next
+Thomas2016, Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both OMC and robust OMC, which we introduce next,
 are based on the approximation in \eqref{eq:approx_likelihood}.

### 2021-08-18 latex_files/ch2/2_3_OMC.tex
Michael's edits on overleaf, LFI background and OMC

@@
-For describing the likelihood approximation introduced in OMC, it is
-convenient to firstly define the indicator function
-$\indicator{\region(\yb)}(\xb)$, that returns 1 if
-$\xb \in \region(\yb)$ and 0 otherwise. If $d(\cdot,\cdot)$ is a
-formal distance, due to symmetry
-$\indicator{\region(\yb)}(\xb) = \indicator{\region(\xb)}(\yb)$, so
-the expressions can be used interchangeably.
-
-\begin{gather} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
+Our description of OMC follows \citet{Ikonomov2019} who base their explanation of OMC on the approximate likelihood function in \eqref{eq:approx_likelihood}. With the indicator function
+%
+\begin{equation} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
   \left\{
@@
       0 & \mbox{otherwise} 
-    \end{array} \right. \end{gather}
-
-\noindent
-Based on equation~\eqref{eq:approx_likelihood} and the indicator
-function~\eqref{eq:indicator}, we can approximate the likelihood as:
-
+    \end{array} \right. \end{equation}
+%
+we can write the approximate likelihood function $L_{d, \epsilon}(\thetab)$ in \eqref{eq:approx_likelihood} as
 \begin{gather} \label{eq:approx_likelihood_omc}
-  L_{d, \epsilon}(\thetab) =
+  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)) =
   \int_{\yb \in B_\epsilon(\data)}p(\yb|\thetab)d\yb =
-  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb\\
-  \approx \frac{1}{N} \sum_i^N \indicator{\region(\data)}(\yb_i),\text{ where }
-  \yb_i \sim M_r(\thetab) \label{eq:init_view}\\
-  \approx \frac{1}{N} \sum_i^N \indicator{\region (\data)} (\yb_i)
-  \text{ where } \yb_i = M_d(\thetab, \vb_i), \vb_i \sim p(\vb) \label{eq:alt_view}
+  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb
 \end{gather}
-%
-The procedure is quite intuitive; approximating the likelihood of a
-specific $\thetab$ requires sampling from the data generator and count
-the fraction of samples that lie inside the area around the
-observations. Nevertheless, by using the approximation of equation
-\eqref{eq:init_view} we need to draw $N$ new samples for each distinct
-evaluation of $L_{d,\epsilon}(\thetab)$; this makes this approach
-quite inconvenient from a computational point-of-view. For this
-reason, OMC approximates the integral as in equation
-\eqref{eq:alt_view}; the nuisance variables, $\vb_i \sim p(\vb)$, are
-sampled once and we count the fraction of samples that lie inside the
-area using the deterministic simulators
-$M_d(\thetab, \vb_i) \: \forall i$. Hence, the evaluation
-$L_{d,\epsilon}(\thetab)$ for each different $\thetab$ does not imply
-drawing new samples all over again. Based on this approach, the
-unnormalised approximate posterior can be defined as:
-
+which we can approximate as a sample average
+\begin{equation}
+L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
+  \text{ where } \yb_i = M_d(\thetab, \vb_i), \quad \vb_i \sim p(\vb). \label{eq:alt_view}
+\end{equation}
+We thus approximate the likelihood of a
+specific $\thetab$ by counting the fraction of samples that lie inside the area around the
+observations.
+The crucial point in \eqref{eq:alt_view} is that we may only sample the nuisance variables $\vb_i \sim p(\vb)$ once and use the sample set of samples to approximate $L_{d,\epsilon}(\thetab)$ for different values of $\thetab$.
+Based on this approach, we introduce the approximate posterior
 \begin{equation} \label{eq:approx_posterior_omc}
   p_{d,\epsilon}(\thetab|\data)
-  \propto p(\thetab) \sum_i^N \indicator{ \region(\data)} (\yb_i)
+  \propto p(\thetab) \sum_i^N \indicator{ \region(\data)} (\yb_i) \text{ where } \yb_i = M_d(\thetab, \vb_i), \quad \vb_i \sim p(\vb).
 \end{equation}

### 2021-08-20 latex_files/ch2/2_3_OMC.tex
edited description of OMC

@@
-Our description of OMC follows \citet{Ikonomov2019} who base their explanation of OMC on the approximate likelihood function in \eqref{eq:approx_likelihood}. With the indicator function
+Our description of OMC \citep{Meeds2015}
+follows \citet{Ikonomov2019} who base their explanation of OMC on the
+approximate likelihood function in \eqref{eq:approx_likelihood}. With
+the indicator function
 %
@@
 \end{equation}
-We thus approximate the likelihood of a
-specific $\thetab$ by counting the fraction of samples that lie inside the area around the
-observations.
-The crucial point in \eqref{eq:alt_view} is that we may only sample the nuisance variables $\vb_i \sim p(\vb)$ once and use the sample set of samples to approximate $L_{d,\epsilon}(\thetab)$ for different values of $\thetab$.
-Based on this approach, we introduce the approximate posterior
+We thus approximate the likelihood of a specific $\thetab$ by counting
+the fraction of samples that lie inside the area around the
+observations. The fact that we isolated the randomness of the
+simulator via the $\vb_i$ has two crucial consequences: First, we can
+sample the nuisance variables $\vb_i \sim p(\vb)$ only once and use
+the same set of samples to approximate $L_{d,\epsilon}(\thetab)$ for
+different values of $\thetab$. Secondly, since
+$\yb_i=M_d(\thetab, \vb_i)$ is a function of $\thetab$, checking
+whether $\yb_i$ is contained in region $\region (\data)$ for a
+particular value of $\thetab$ is the same as checking whether
+$\thetab$ is in the acceptance region $\mathcal{S}_i = \{ \thetab:
+M_d(\thetab, \vb_i) \in \region(\data) \}$. This leads to the
+likelihood approximation
+\begin{equation}
+L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\mathcal{S}_i}(\thetab) \label{eq:alt_view_theta}
+\end{equation}
+and the corresponding approximate posterior
 \begin{equation} \label{eq:approx_posterior_omc}
-  p_{d,\epsilon}(\thetab|\data)
-  \propto p(\thetab) \sum_i^N \indicator{ \region(\data)} (\yb_i) \text{ where } \yb_i = M_d(\thetab, \vb_i), \quad \vb_i \sim p(\vb).
+  p_{d,\epsilon}(\thetab|\data) \propto
+  p(\thetab) \sum_i^N  \indicator{\mathcal{S}_i}(\thetab).
 \end{equation}
+As argued by \citet{Ikonomov2019}, this derivations provide a unique
+perspective for likelihood-free inference by shifting the focus onto
+the geometry of the acceptance regions $\mathcal{S}_i = \{ \thetab:
+M_d(\thetab, \vb_i) \in \region(\data)\}$. Indeed, the task of
+approximating the likelihood and/or posterior becomes a task of
+characterising the sets $\mathcal{S}_i$.
 
-\subsubsection*{Further approximations for sampling and computing expectations}
-
-The posterior approximation in \eqref{eq:approx_posterior_omc} does
-not provide any convenient way for drawing samples.\footnote{Weighted
-  sampling could be performed in a straightforward fashion with
-  importance sampling, using the prior as the proposal distribution
-  $\thetab_i \sim p(\thetab)$. However, this approach has the same
-  drawbacks as ABC rejection sampling; when the prior is wide or the
-  dimensionality $D$ is high, drawing a sample with non-zero weight is
-  rare, leading to either poor Effective Sample Size (ESS) or huge
-  execution time.} In fact, the set
-$\mathcal{S}_i = \{ \thetab: M_d(\thetab, \vb_i) \in \region(\data)
-\}$ can represent any arbitrary shape in the D-dimensional Euclidean
-space; it can be non-convex, can contain disjoint sets of $\thetab$
-etc. We need some further simplification of the posterior for being
-able to draw samples from it.
-
-OMC proposes a quite drastic simplification of the posterior; it
-squeezes all regions $\mathcal{S}_i$ into a single point
-$\thetab_i^* \in \mathcal{S}_i$ attaching a weight $w_i$ proportional
-to the volume of $\mathcal{S}_i$. For obtaining a
-$\thetab_i^* \in \mathcal{S}_i$, a gradient based optimiser is used
-for minimising $g_i(\thetab) = d(\data, f_i(\thetab))$ and the
-estimation of the volume of $\mathcal{S}_i$ is done using the Hessian
-approximation $\hess_i \approx \jac_i^{*T}\jac_i^*$, where $\jac_i^*$
-is the Jacobian matrix of $g_i(\thetab)$ at $\thetab_i^*$. Hence,
-
-\begin{gather} \label{eq:OMC_posterior}
-    p(\thetab|\data) \propto p(\thetab) \sum_i^N w_i \delta(\thetab - \thetab_i^*)\\
-  \thetab_i^* = \text{argmin}_{\thetab} \:g_i(\thetab) \\
-  w_i \propto \frac{1}{\sqrt{det( \jac_i^{*T}\jac_i^*)}}
-\end{gather}
-
-The distribution \eqref{eq:OMC_posterior} provides weighted samples
-automatically and an expectation can be computed easily with the
-following equation,
-
+OMC by \citet{Meeds2015} assumes that the distance $d$ is the
+Euclidean distance between summary statistics $\Phi$ of the observed
+and generated data, and that the $\mathcal{S}_i$ can be well
+approximated by infinitesimally small ellipses. These assumptions
+leads to an approximation of the posterior in terms of weighted
+samples $\thetab_i^*$. In more detail, OMC considers the limit of
+$\epsilon \to 0$ and represents the sets $\mathcal{S}_i$ by points
+$\thetab_i^*$ which are those parameter configurations that achieve
+the smallest distance between observed and simulated data for each
+realisation $\vb_i \sim p(\vb)$, i.e.\ they are solutions to the
+deterministic optimisation problems
 \begin{equation}
-  \label{eq:OMC_expectation}
-  E_{p(\thetab|\data)}[h(\thetab)] = \frac{\sum_i^N w_i p(\thetab_i^*)h(\thetab_i^*)}{\sum_i^N w_i p(\thetab_i^*)}
+\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(M_d(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
 \end{equation}
+The weighting for each $\thetab_i^*$ is proportional to the prior
+density at $\thetab_i^*$ and inversely proportional to the determinant
+of the Jacobian matrix of the summary statistics at $\thetab_i^*$. For
+further details on OMC we refer the reader to \citep{Meeds2015,
+Ikonomov2019}.

### 2021-08-24 latex_files/ch2/2_2_LFI_background.tex
plots after changes

@@
 of probability density functions $\{ p(\yb|\thetab) \}_{\thetab}$,
-whose closed-form is either unknown or computationally intractable to
-evaluate. Whereas evaluating $p(\yb|\thetab)$ is intractable, sampling
-is feasible. Practically, a simulator can be understood as a random black-box
-machine (computer code) $M_r$ that generates samples $\yb$ in a stochastic manner for
-any given a set of parameters $\thetab$.
-We can isolate the randomness by introducing the stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator becomes a deterministic mapping $M_d$ that maps $(\thetab, \vb)$ to the data $\yb$.
-The distribution $p(\vb)$ is determined by the random draws performed when running $M_r$.
-Importantly, we can fix the value of $\vb$ while changing $\thetab$ by fixing the random seed used with $M_r$.
+whose closed-form is either unknown or computationally
+intractable. Whereas evaluating $p(\yb|\thetab)$ is intractable,
+sampling is feasible. Practically, a simulator can be understood as a
+random black-box machine (computer code) $M_r$ that generates samples
+$\yb$ in a stochastic manner for any given a set of parameters
+$\thetab$. We can isolate the randomness by introducing the stochastic
+nuisance variables $\vb \sim p(\vb)$ so that the simulator becomes a
+deterministic mapping $M_d$ that maps $(\thetab, \vb)$ to the data
+$\yb$. The distribution $p(\vb)$ is determined by the random draws
+performed when running $M_r$. Importantly, we can fix the value of
+$\vb$ while changing $\thetab$ by fixing the random seed used with
+$M_r$.
 

### 2021-08-24 latex_files/ch2/2_4_ROMC.tex
applied fixes
minor improvements
plots after changes

@@
 way. Instead of collapsing the acceptance regions into single points,
-we approximate them with a bounding box region. A uniform distribution
+it approximates them with a bounding box region. A uniform distribution
 defined on the bounding box area, is then used as the \emph{proposal
@@
 
-\subsubsection{Proposal region estimation}
-
 The constrution of the proposal regions $q_i$ is performed in two
@@
 
-For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$ a
-deterministic function is defined as
-$f_i(\thetab) = M_d(\thetab,\vb_i)$. We look for a point
-$\thetab_* : d(f_i(\thetab_*), \data) < \epsilon$. This point can be
-obtained by solving the the following optimisation problem:
+For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$,
+we search for a point $\thetab^*$ such that
+$d(M_d(\thetab^*,\vb_i), \data) < \epsilon$. For notation convenience,
+we name the function $d(M_d(\thetab^*,\vb_i)$ as $d_i(\thetab)$.  Obtaining
+this point involves solving the following optimisation problem:
 
@@
 \begin{alignat}{2}
-&\!\min_{\thetab}        &\qquad& g_i(\thetab) = d(\data,  f_i(\thetab))\label{eq:optProb}\\
-&\text{subject to} &      & g_i(\thetab) \leq \epsilon
+  &\!\min_{\thetab}        &\qquad& d_i(\thetab) = d(M_d(\thetab^*,\vb_i), \data)\label{eq:optProb}\\
+  &\text{subject to} & & d_i(\thetab) \leq \epsilon
 \end{alignat}
@@
 The optimisation problem can be treated as unconstrained, accepting
-the optimal point $\thetab_i^* = \text{argmin}_{\thetab} g_i(\thetab)$
-only if $g_i(\thetab_i^*) < \epsilon$. The nature of the generative
+the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
+only if $d_i(\thetab_i^*) < \epsilon$. The nature of the generative
 model $M_r(\thetab)$, specifies the properties of the objective
-function $g_i$. If $g_i$ is differentiable any gradient-based
+function $d_i$. If $d_i$ is differentiable any gradient-based
 iterative algorithm can be used for~\ref{eq:optProb}. The gradients
 $\nabla_{\thetab} g_i$ can be either provided in closed form or
-approximated by finite differences. Alternatively, if $g_i$ is not
+approximated by finite differences. Alternatively, if $d_i$ is not
 differentiable, the Bayesian Optimisation approach provides an
@@
 obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
-the distance $g_i$ is also fitted; the approximate model can then
-substitute the actual model in the following steps, e.g. construction
-of bounding box, sampling, for accelerating the method.
+the distance $d_i$ is also fitted; the approximate model can then
+substitute the actual model in the following steps.
 
-\subsubsection*{Bounding boxes construction}
+\subsubsection*{Proposal region computation}
 
-After obtaining a $\thetab_i^*$ such that
-$g_i(\thetab_i^*) < \epsilon$, we need to construct a bounding box
-area $\mathcal{\hat{S}}_i$ around it. The bounding box must contain
-the acceptance region around $\thetab_i^*$, formally
-$\{ \thetab : g_i(\thetab) < \epsilon$,
+After obtaining $\thetab_i^*$, we approximate the acceptance region
+$\mathcal{S}_i$ by constructing a bounding box $\mathcal{\hat{S}}_i$
+around it. The bounding must the part of the acceptance region
+$\mathcal{S}_i$ around $\thetab_i^*$, formally
+$\{ \thetab : d_i(\thetab) < \epsilon$,
 $d(\thetab, \thetab_i^*) < M \}$. The second condition
 $d(\thetab, \thetab_i^*) < M$ describes that if
-$\mathcal{S}_i := \{ \thetab : g_i(\thetab) < \epsilon \} $ contains a
+$\mathcal{S}_i := \{ \thetab : d_i(\thetab) < \epsilon \} $ contains a
 number of disjoint sets of $\thetab$ that they fulfill
-$g_i(\thetab) < \epsilon$, we want the bounding box to match only the
+$d_i(\thetab) < \epsilon$, we want the bounding box to match only the
 set that contains $\thetab_i^*$. We seek for a bounding box that is as
@@
 method looking for the smallest
-$\kappa: g_i(\thetab_i^* + \kappa \vb_d) \geq
+$\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
 \epsilon$\footnote{$-\kappa$ is used as well for the opposite

### 2021-08-24 latex_files/images/chapter3/example_marginal.tex
plots after changes

@@
-% This file was created by tikzplotlib v0.9.9.
+% This file was created with tikzplotlib v0.9.12.
 \begin{tikzpicture}
@@
 \draw[draw=none,fill=color0] (axis cs:-2.65714285714286,0) rectangle (axis cs:-2.57142857142857,0);
-\draw[draw=none,fill=color0] (axis cs:-2.57142857142857,0) rectangle (axis cs:-2.48571428571429,0.00722007830101246);
-\draw[draw=none,fill=color0] (axis cs:-2.48571428571429,0) rectangle (axis cs:-2.4,0.0252702740535435);
-\draw[draw=none,fill=color0] (axis cs:-2.4,0) rectangle (axis cs:-2.31428571428571,0.0347466268236223);
-\draw[draw=none,fill=color0] (axis cs:-2.31428571428571,0) rectangle (axis cs:-2.22857142857143,0.0374541561865022);
-\draw[draw=none,fill=color0] (axis cs:-2.22857142857143,0) rectangle (axis cs:-2.14285714285714,0.0510691609826969);
-\draw[draw=none,fill=color0] (axis cs:-2.14285714285714,0) rectangle (axis cs:-2.05714285714286,0.0647744167576544);
-\draw[draw=none,fill=color0] (axis cs:-2.05714285714286,0) rectangle (axis cs:-1.97142857142857,0.080503873056289);
-\draw[draw=none,fill=color0] (axis cs:-1.97142857142857,0) rectangle (axis cs:-1.88571428571429,0.089941546835469);
-\draw[draw=none,fill=color0] (axis cs:-1.88571428571429,0) rectangle (axis cs:-1.8,0.113625982262183);
-\draw[draw=none,fill=color0] (axis cs:-1.8,0) rectangle (axis cs:-1.71428571428571,0.12614508231626);
-\draw[draw=none,fill=color0] (axis cs:-1.71428571428571,0) rectangle (axis cs:-1.62857142857143,0.145794009692587);
-\draw[draw=none,fill=color0] (axis cs:-1.62857142857143,0) rectangle (axis cs:-1.54285714285714,0.171554217630842);
-\draw[draw=none,fill=color0] (axis cs:-1.54285714285714,0) rectangle (axis cs:-1.45714285714286,0.157720031886223);
-\draw[draw=none,fill=color0] (axis cs:-1.45714285714286,0) rectangle (axis cs:-1.37142857142857,0.174635643905738);
-\draw[draw=none,fill=color0] (axis cs:-1.37142857142857,0) rectangle (axis cs:-1.28571428571429,0.199983275941078);
-\draw[draw=none,fill=color0] (axis cs:-1.28571428571429,0) rectangle (axis cs:-1.2,0.216744171997);
-\draw[draw=none,fill=color0] (axis cs:-1.2,0) rectangle (axis cs:-1.11428571428571,0.240854076323595);
-\draw[draw=none,fill=color0] (axis cs:-1.11428571428571,0) rectangle (axis cs:-1.02857142857143,0.277844084619317);
-\draw[draw=none,fill=color0] (axis cs:-1.02857142857143,0) rectangle (axis cs:-0.942857142857143,0.253682608304859);
-\draw[draw=none,fill=color0] (axis cs:-0.942857142857143,0) rectangle (axis cs:-0.857142857142857,0.2729834247631);
-\draw[draw=none,fill=color0] (axis cs:-0.857142857142857,0) rectangle (axis cs:-0.771428571428571,0.303178823657691);
-\draw[draw=none,fill=color0] (axis cs:-0.771428571428571,0) rectangle (axis cs:-0.685714285714285,0.297596155971373);
-\draw[draw=none,fill=color0] (axis cs:-0.685714285714285,0) rectangle (axis cs:-0.6,0.292864426084818);
-\draw[draw=none,fill=color0] (axis cs:-0.6,0) rectangle (axis cs:-0.514285714285714,0.327185584008558);
-\draw[draw=none,fill=color0] (axis cs:-0.514285714285714,0) rectangle (axis cs:-0.428571428571428,0.303294860630386);
-\draw[draw=none,fill=color0] (axis cs:-0.428571428571428,0) rectangle (axis cs:-0.342857142857143,0.313867118142585);
-\draw[draw=none,fill=color0] (axis cs:-0.342857142857143,0) rectangle (axis cs:-0.257142857142857,0.316600433499394);
-\draw[draw=none,fill=color0] (axis cs:-0.257142857142857,0) rectangle (axis cs:-0.171428571428571,0.325419243424203);
-\draw[draw=none,fill=color0] (axis cs:-0.171428571428571,0) rectangle (axis cs:-0.0857142857142859,0.322737500055257);
-\draw[draw=none,fill=color0] (axis cs:-0.0857142857142859,0) rectangle (axis cs:0,0.331556309980064);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.0857142857142859,0.327533694926642);
-\draw[draw=none,fill=color0] (axis cs:0.0857142857142859,0) rectangle (axis cs:0.171428571428571,0.351940138183459);
-\draw[draw=none,fill=color0] (axis cs:0.171428571428571,0) rectangle (axis cs:0.257142857142857,0.308245771465366);
-\draw[draw=none,fill=color0] (axis cs:0.257142857142857,0) rectangle (axis cs:0.342857142857143,0.329080854562574);
-\draw[draw=none,fill=color0] (axis cs:0.342857142857143,0) rectangle (axis cs:0.428571428571429,0.311752666640144);
-\draw[draw=none,fill=color0] (axis cs:0.428571428571429,0) rectangle (axis cs:0.514285714285714,0.312139456549128);
-\draw[draw=none,fill=color0] (axis cs:0.514285714285714,0) rectangle (axis cs:0.6,0.322673035070425);
-\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.685714285714286,0.302650210782081);
-\draw[draw=none,fill=color0] (axis cs:0.685714285714286,0) rectangle (axis cs:0.771428571428571,0.313763974166856);
-\draw[draw=none,fill=color0] (axis cs:0.771428571428571,0) rectangle (axis cs:0.857142857142857,0.278656343428181);
-\draw[draw=none,fill=color0] (axis cs:0.857142857142857,0) rectangle (axis cs:0.942857142857143,0.269785961515508);
-\draw[draw=none,fill=color0] (axis cs:0.942857142857143,0) rectangle (axis cs:1.02857142857143,0.245830773152508);
-\draw[draw=none,fill=color0] (axis cs:1.02857142857143,0) rectangle (axis cs:1.11428571428571,0.227754791406045);
-\draw[draw=none,fill=color0] (axis cs:1.11428571428571,0) rectangle (axis cs:1.2,0.218858623499437);
-\draw[draw=none,fill=color0] (axis cs:1.2,0) rectangle (axis cs:1.28571428571429,0.218433154599559);
-\draw[draw=none,fill=color0] (axis cs:1.28571428571429,0) rectangle (axis cs:1.37142857142857,0.21010427855946);
-\draw[draw=none,fill=color0] (axis cs:1.37142857142857,0) rectangle (axis cs:1.45714285714286,0.193691493421625);
-\draw[draw=none,fill=color0] (axis cs:1.45714285714286,0) rectangle (axis cs:1.54285714285714,0.194078283330608);
-\draw[draw=none,fill=color0] (axis cs:1.54285714285714,0) rectangle (axis cs:1.62857142857143,0.130644738257426);
-\draw[draw=none,fill=color0] (axis cs:1.62857142857143,0) rectangle (axis cs:1.71428571428571,0.150177628661059);
-\draw[draw=none,fill=color0] (axis cs:1.71428571428571,0) rectangle (axis cs:1.8,0.0986572127845487);
-\draw[draw=none,fill=color0] (axis cs:1.8,0) rectangle (axis cs:1.88571428571429,0.0960012554095325);
-\draw[draw=none,fill=color0] (axis cs:1.88571428571429,0) rectangle (axis cs:1.97142857142857,0.0732709017583105);
-\draw[draw=none,fill=color0] (axis cs:1.97142857142857,0) rectangle (axis cs:2.05714285714286,0.0897223658870459);
-\draw[draw=none,fill=color0] (axis cs:2.05714285714286,0) rectangle (axis cs:2.14285714285714,0.0617058834797238);
-\draw[draw=none,fill=color0] (axis cs:2.14285714285714,0) rectangle (axis cs:2.22857142857143,0.0434622927727019);
-\draw[draw=none,fill=color0] (axis cs:2.22857142857143,0) rectangle (axis cs:2.31428571428571,0.0347466268236225);
-\draw[draw=none,fill=color0] (axis cs:2.31428571428571,0) rectangle (axis cs:2.4,0.0374541561865018);
-\draw[draw=none,fill=color0] (axis cs:2.4,0) rectangle (axis cs:2.48571428571429,0.0320390974607428);
-\draw[draw=none,fill=color0] (axis cs:2.48571428571429,0) rectangle (axis cs:2.57142857142857,0.00496380383194607);
+\draw[draw=none,fill=color0] (axis cs:-2.57142857142857,0) rectangle (axis cs:-2.48571428571429,0.006208033765389);
+\draw[draw=none,fill=color0] (axis cs:-2.48571428571429,0) rectangle (axis cs:-2.4,0.0261624280112821);
+\draw[draw=none,fill=color0] (axis cs:-2.4,0) rectangle (axis cs:-2.31428571428571,0.0337007547263973);
+\draw[draw=none,fill=color0] (axis cs:-2.31428571428571,0) rectangle (axis cs:-2.22857142857143,0.0363613406258499);
+\draw[draw=none,fill=color0] (axis cs:-2.22857142857143,0) rectangle (axis cs:-2.14285714285714,0.0541896684870789);
+\draw[draw=none,fill=color0] (axis cs:-2.14285714285714,0) rectangle (axis cs:-2.05714285714286,0.0670526704061584);
+\draw[draw=none,fill=color0] (axis cs:-2.05714285714286,0) rectangle (axis cs:-1.97142857142857,0.0841007399119992);
+\draw[draw=none,fill=color0] (axis cs:-1.97142857142857,0) rectangle (axis cs:-1.88571428571429,0.0882637860955287);
+\draw[draw=none,fill=color0] (axis cs:-1.88571428571429,0) rectangle (axis cs:-1.8,0.107105098505383);
+\draw[draw=none,fill=color0] (axis cs:-1.8,0) rectangle (axis cs:-1.71428571428571,0.134226458530087);
+\draw[draw=none,fill=color0] (axis cs:-1.71428571428571,0) rectangle (axis cs:-1.62857142857143,0.13742807026801);
+\draw[draw=none,fill=color0] (axis cs:-1.62857142857143,0) rectangle (axis cs:-1.54285714285714,0.172328991470496);
+\draw[draw=none,fill=color0] (axis cs:-1.54285714285714,0) rectangle (axis cs:-1.45714285714286,0.159523545491678);
+\draw[draw=none,fill=color0] (axis cs:-1.45714285714286,0) rectangle (axis cs:-1.37142857142857,0.181085502173434);
+\draw[draw=none,fill=color0] (axis cs:-1.37142857142857,0) rectangle (axis cs:-1.28571428571429,0.198590015309879);
+\draw[draw=none,fill=color0] (axis cs:-1.28571428571429,0) rectangle (axis cs:-1.2,0.221267556259859);
+\draw[draw=none,fill=color0] (axis cs:-1.2,0) rectangle (axis cs:-1.11428571428571,0.234145572772051);
+\draw[draw=none,fill=color0] (axis cs:-1.11428571428571,0) rectangle (axis cs:-1.02857142857143,0.277788991057236);
+\draw[draw=none,fill=color0] (axis cs:-1.02857142857143,0) rectangle (axis cs:-0.942857142857143,0.250781241453748);
+\draw[draw=none,fill=color0] (axis cs:-0.942857142857143,0) rectangle (axis cs:-0.857142857142857,0.265410460009239);
+\draw[draw=none,fill=color0] (axis cs:-0.857142857142857,0) rectangle (axis cs:-0.771428571428571,0.303461442846629);
+\draw[draw=none,fill=color0] (axis cs:-0.771428571428571,0) rectangle (axis cs:-0.685714285714285,0.302778779346469);
+\draw[draw=none,fill=color0] (axis cs:-0.685714285714285,0) rectangle (axis cs:-0.6,0.294470204004708);
+\draw[draw=none,fill=color0] (axis cs:-0.6,0) rectangle (axis cs:-0.514285714285714,0.32589124301026);
+\draw[draw=none,fill=color0] (axis cs:-0.514285714285714,0) rectangle (axis cs:-0.428571428571428,0.307263137822596);
+\draw[draw=none,fill=color0] (axis cs:-0.428571428571428,0) rectangle (axis cs:-0.342857142857143,0.310432718428548);
+\draw[draw=none,fill=color0] (axis cs:-0.342857142857143,0) rectangle (axis cs:-0.257142857142857,0.313571769361836);
+\draw[draw=none,fill=color0] (axis cs:-0.257142857142857,0) rectangle (axis cs:-0.171428571428571,0.322468416267153);
+\draw[draw=none,fill=color0] (axis cs:-0.171428571428571,0) rectangle (axis cs:-0.0857142857142859,0.329442694767694);
+\draw[draw=none,fill=color0] (axis cs:-0.0857142857142859,0) rectangle (axis cs:0,0.328704477273017);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.0857142857142859,0.315996626149429);
+\draw[draw=none,fill=color0] (axis cs:0.0857142857142859,0) rectangle (axis cs:0.171428571428571,0.35869562652787);
+\draw[draw=none,fill=color0] (axis cs:0.171428571428571,0) rectangle (axis cs:0.257142857142857,0.299321419039204);
+\draw[draw=none,fill=color0] (axis cs:0.257142857142857,0) rectangle (axis cs:0.342857142857143,0.327344655623505);
+\draw[draw=none,fill=color0] (axis cs:0.342857142857143,0) rectangle (axis cs:0.428571428571429,0.324602990921247);
+\draw[draw=none,fill=color0] (axis cs:0.428571428571429,0) rectangle (axis cs:0.514285714285714,0.300242814569848);
+\draw[draw=none,fill=color0] (axis cs:0.514285714285714,0) rectangle (axis cs:0.6,0.325196567835609);
+\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.685714285714286,0.3008173729996);
+\draw[draw=none,fill=color0] (axis cs:0.685714285714286,0) rectangle (axis cs:0.771428571428571,0.306931815801259);
+\draw[draw=none,fill=color0] (axis cs:0.771428571428571,0) rectangle (axis cs:0.857142857142857,0.292671455749884);
+\draw[draw=none,fill=color0] (axis cs:0.857142857142857,0) rectangle (axis cs:0.942857142857143,0.266064095296049);
+\draw[draw=none,fill=color0] (axis cs:0.942857142857143,0) rectangle (axis cs:1.02857142857143,0.241445667943122);
+\draw[draw=none,fill=color0] (axis cs:1.02857142857143,0) rectangle (axis cs:1.11428571428571,0.23182281521762);
+\draw[draw=none,fill=color0] (axis cs:1.11428571428571,0) rectangle (axis cs:1.2,0.217147051423478);
+\draw[draw=none,fill=color0] (axis cs:1.2,0) rectangle (axis cs:1.28571428571429,0.224310513797232);
+\draw[draw=none,fill=color0] (axis cs:1.28571428571429,0) rectangle (axis cs:1.37142857142857,0.212377916164659);
+\draw[draw=none,fill=color0] (axis cs:1.37142857142857,0) rectangle (axis cs:1.45714285714286,0.190746892354591);
+\draw[draw=none,fill=color0] (axis cs:1.45714285714286,0) rectangle (axis cs:1.54285714285714,0.196218210084616);
+\draw[draw=none,fill=color0] (axis cs:1.54285714285714,0) rectangle (axis cs:1.62857142857143,0.133681428800119);
+\draw[draw=none,fill=color0] (axis cs:1.62857142857143,0) rectangle (axis cs:1.71428571428571,0.142913401618273);
+\draw[draw=none,fill=color0] (axis cs:1.71428571428571,0) rectangle (axis cs:1.8,0.108955396863223);
+\draw[draw=none,fill=color0] (axis cs:1.8,0) rectangle (axis cs:1.88571428571429,0.0924363415213623);
+\draw[draw=none,fill=color0] (axis cs:1.88571428571429,0) rectangle (axis cs:1.97142857142857,0.0739859090188589);
+\draw[draw=none,fill=color0] (axis cs:1.97142857142857,0) rectangle (axis cs:2.05714285714286,0.0907797314146689);
+\draw[draw=none,fill=color0] (axis cs:2.05714285714286,0) rectangle (axis cs:2.14285714285714,0.0652904576612423);
+\draw[draw=none,fill=color0] (axis cs:2.14285714285714,0) rectangle (axis cs:2.22857142857143,0.0433796619328517);
+\draw[draw=none,fill=color0] (axis cs:2.22857142857143,0) rectangle (axis cs:2.31428571428571,0.0363613406258499);
+\draw[draw=none,fill=color0] (axis cs:2.31428571428571,0) rectangle (axis cs:2.4,0.0359179096426075);
+\draw[draw=none,fill=color0] (axis cs:2.4,0) rectangle (axis cs:2.48571428571429,0.0301533068604609);
+\draw[draw=none,fill=color0] (axis cs:2.48571428571429,0) rectangle (axis cs:2.57142857142857,0.00665146474863107);
 \draw[draw=none,fill=color0] (axis cs:2.57142857142857,0) rectangle (axis cs:2.65714285714286,0);

### 2021-08-24 latex_files/images/chapter3/example_post.tex
plots after changes

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.75,0.75,0}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={Approximate Posterior},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta\)},
xmin=-3, xmax=3,
xtick style={color=black},
xtick={-3,-2,-1,0,1,2,3},
xticklabels={
  \(\displaystyle {\ensuremath{-}3}\),
  \(\displaystyle {\ensuremath{-}2}\),
  \(\displaystyle {\ensuremath{-}1}\),
  \(\displaystyle {0}\),
  \(\displaystyle {1}\),
  \(\displaystyle {2}\),
  \(\displaystyle {3}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=0.6,
ytick style={color=black},
ytick={0,0.1,0.2,0.3,0.4,0.5,0.6},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.1}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.3}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {0.6}\)
}
]
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-3,0) rectangle (axis cs:-2.88,0);
\addlegendimage{ybar,ybar legend,draw=none,fill=color0,fill opacity=0.5}
\addlegendentry{samples histogram}

\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.88,0) rectangle (axis cs:-2.76,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.76,0) rectangle (axis cs:-2.64,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.64,0) rectangle (axis cs:-2.52,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.52,0) rectangle (axis cs:-2.4,0.0231217584119079);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.4,0) rectangle (axis cs:-2.28,0.0357912150759672);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.28,0) rectangle (axis cs:-2.16,0.0432881444129306);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.16,0) rectangle (axis cs:-2.04,0.0678262794417402);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.04,0) rectangle (axis cs:-1.92,0.0853043382950843);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.92,0) rectangle (axis cs:-1.8,0.104057207601703);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.8,0) rectangle (axis cs:-1.68,0.133234851902131);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.68,0) rectangle (axis cs:-1.56,0.165264196422909);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.56,0) rectangle (axis cs:-1.44,0.16070619494247);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.44,0) rectangle (axis cs:-1.32,0.180066797779813);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.32,0) rectangle (axis cs:-1.2,0.221049487169422);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.2,0) rectangle (axis cs:-1.08,0.246171403878041);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.08,0) rectangle (axis cs:-0.96,0.263219759376131);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.96,0) rectangle (axis cs:-0.84,0.275654702395393);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.84,0) rectangle (axis cs:-0.72,0.30275211056918);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.72,0) rectangle (axis cs:-0.6,0.289942517702741);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.6,0) rectangle (axis cs:-0.48,0.327857582723436);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.48,0) rectangle (axis cs:-0.36,0.302668457836128);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.36,0) rectangle (axis cs:-0.24,0.314413087062547);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.24,0) rectangle (axis cs:-0.12,0.319501604166278);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.12,0) rectangle (axis cs:0,0.333969594590972);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0,0) rectangle (axis cs:0.12,0.336688308415186);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.12,0) rectangle (axis cs:0.24,0.311239645561226);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.24,0) rectangle (axis cs:0.36,0.328195053578145);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.36,0) rectangle (axis cs:0.48,0.316839373811389);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.48,0) rectangle (axis cs:0.6,0.315180976253135);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.6,0) rectangle (axis cs:0.72,0.305533027707716);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.72,0) rectangle (axis cs:0.84,0.294804743438855);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.84,0) rectangle (axis cs:0.96,0.270682012152812);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.96,0) rectangle (axis cs:1.08,0.238103562512508);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.08,0) rectangle (axis cs:1.2,0.217233993067405);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.2,0) rectangle (axis cs:1.32,0.213737380323862);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.32,0) rectangle (axis cs:1.44,0.205697780692993);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.44,0) rectangle (axis cs:1.56,0.187837922186227);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.56,0) rectangle (axis cs:1.68,0.140154434384905);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.68,0) rectangle (axis cs:1.8,0.116289453613951);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.8,0) rectangle (axis cs:1.92,0.0886000414831757);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.92,0) rectangle (axis cs:2.04,0.0841267652067261);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.04,0) rectangle (axis cs:2.16,0.0663927432898532);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.16,0) rectangle (axis cs:2.28,0.0377377498258489);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.28,0) rectangle (axis cs:2.4,0.0361079514925686);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.4,0) rectangle (axis cs:2.52,0.0262891225779227);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.52,0) rectangle (axis cs:2.64,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.64,0) rectangle (axis cs:2.76,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.76,0) rectangle (axis cs:2.88,0);
\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.88,0) rectangle (axis cs:3,0);
\addplot [semithick, blue, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0
-2.51677846908569 0
-2.4765100479126 0.200000047683716
2.4765100479126 0.200000047683716
2.51677846908569 0
3 0
};
\addlegendentry{Prior}
\addplot [semithick, green!50!black, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0.0149635076522827
-2.91946315765381 0.0183337926864624
-2.83892607688904 0.0223180055618286
-2.75838923454285 0.0269923210144043
-2.67785239219666 0.0324347019195557
-2.63758397102356 0.0354681015014648
-2.59731554985046 0.0387222766876221
-2.55704689025879 0.0422066450119019
-2.51677846908569 0.0459299087524414
-2.4765100479126 0.0499007701873779
-2.4362416267395 0.0541269779205322
-2.39597320556641 0.058616042137146
-2.35570478439331 0.0633745193481445
-2.31543612480164 0.0684083700180054
-2.27516770362854 0.073722243309021
-2.23489928245544 0.0793203115463257
-2.19463086128235 0.0852051973342896
-2.15436244010925 0.0913783311843872
-2.11409401893616 0.0978399515151978
-2.07382559776306 0.104588747024536
-2.03355693817139 0.111621975898743
-1.99328863620758 0.118935108184814
-1.9530200958252 0.126521944999695
-1.9127516746521 0.134374856948853
-1.872483253479 0.142483830451965
-1.83221471309662 0.150837302207947
-1.79194629192352 0.159421920776367
-1.71140944957733 0.177220344543457
-1.63087248802185 0.195732235908508
-1.55033552646637 0.214780211448669
-1.3892617225647 0.253633737564087
-1.30872488021851 0.272953271865845
-1.22818791866302 0.291845083236694
-1.14765095710754 0.310027003288269
-1.10738253593445 0.318762183189392
-1.06711411476135 0.327212452888489
-1.02684569358826 0.335342526435852
-0.986577153205872 0.343117833137512
-0.946308732032776 0.350504517555237
-0.906040191650391 0.357470154762268
-0.865771770477295 0.363983392715454
-0.825503349304199 0.370014905929565
-0.785234928131104 0.375536918640137
-0.744966506958008 0.380523920059204
-0.704697966575623 0.384952306747437
-0.664429545402527 0.388801217079163
-0.624161005020142 0.392052412033081
-0.583892583847046 0.394690275192261
-0.54362416267395 0.39670205116272
-0.503355741500854 0.398078083992004
-0.463087320327759 0.398520588874817
-0.382550358772278 0.398850798606873
-0.18120801448822 0.398941993713379
0.422818779945374 0.398738622665405
0.463087320327759 0.398520588874817
0.503355741500854 0.398078083992004
0.54362416267395 0.39670205116272
0.583892583847046 0.394690275192261
0.624161005020142 0.392052412033081
0.664429545402527 0.388801217079163
0.704697966575623 0.384952306747437
0.744966506958008 0.380523920059204
0.785234928131104 0.375536918640137
0.825503349304199 0.370014905929565
0.865771770477295 0.363983392715454
0.906040191650391 0.357470154762268
0.946308732032776 0.350504517555237
0.986577153205872 0.343117833137512
1.02684569358826 0.335342526435852
1.06711411476135 0.327212452888489
1.10738253593445 0.318762183189392
1.14765095710754 0.310027003288269
1.22818791866302 0.291845083236694
1.30872488021851 0.272953271865845
1.42953014373779 0.24389910697937
1.55033552646637 0.214780211448669
1.63087248802185 0.195732235908508
1.71140944957733 0.177220344543457
1.79194629192352 0.159421920776367
1.83221471309662 0.150837302207947
1.872483253479 0.142483830451965
1.9127516746521 0.134374856948853
1.9530200958252 0.126521944999695
1.99328863620758 0.118935108184814
2.03355693817139 0.111621975898743
2.07382559776306 0.104588747024536
2.11409401893616 0.0978399515151978
2.15436244010925 0.0913783311843872
2.19463086128235 0.0852051973342896
2.23489928245544 0.0793203115463257
2.27516770362854 0.073722243309021
2.31543612480164 0.0684083700180054
2.35570478439331 0.0633745193481445
2.39597320556641 0.058616042137146
2.4362416267395 0.0541269779205322
2.4765100479126 0.0499007701873779
2.51677846908569 0.0459299087524414
2.55704689025879 0.0422066450119019
2.59731554985046 0.0387222766876221
2.63758397102356 0.0354681015014648
2.67785239219666 0.0324347019195557
2.75838923454285 0.0269923210144043
2.83892607688904 0.0223180055618286
2.91946315765381 0.0183337926864624
3 0.0149635076522827
};
\addlegendentry{Likelihood}
\addplot [semithick, blue!50.1960784313725!black, dash pattern=on 1pt off 3pt on 3pt off 3pt, mark=*, mark size=3, mark options={solid}]
table {%
-3 0
-2.9597315788269 0
-2.91946315765381 0
-2.87919473648071 0
-2.83892607688904 0
-2.79865765571594 0
-2.75838923454285 0
-2.71812081336975 0
-2.67785239219666 0
-2.63758397102356 0
-2.59731554985046 0
-2.55704689025879 0
-2.51677846908569 0
-2.4765100479126 0.0550098419189453
-2.4362416267395 0.0559921264648438
-2.39597320556641 0.060903787612915
-2.35570478439331 0.0648330450057983
-2.31543612480164 0.0658153295516968
-2.27516770362854 0.0697445869445801
-2.23489928245544 0.0746562480926514
-2.19463086128235 0.0834970474243164
-2.15436244010925 0.0903732776641846
-2.11409401893616 0.0982317924499512
-2.07382559776306 0.105108022689819
-2.03355693817139 0.111984252929688
-1.99328863620758 0.116895914077759
-1.9530200958252 0.123772144317627
-1.9127516746521 0.126718997955322
-1.872483253479 0.131630659103394
-1.83221471309662 0.13948917388916
-1.79194629192352 0.14538311958313
-1.75167787075043 0.149312376976013
-1.71140944957733 0.154223918914795
-1.67114090919495 0.160117864608765
-1.63087248802185 0.168958783149719
-1.59060406684875 0.176817297935486
-1.55033552646637 0.181728839874268
-1.51006710529327 0.190569758415222
-1.46979868412018 0.196463704109192
-1.42953014373779 0.199410557746887
-1.3892617225647 0.20923376083374
-1.3489933013916 0.214145421981812
-1.30872488021851 0.217092275619507
-1.26845633983612 0.224950909614563
-1.22818791866302 0.230844736099243
-1.18791949748993 0.235756397247314
-1.14765095710754 0.236738681793213
-1.10738253593445 0.245579600334167
-1.06711411476135 0.246561884880066
-1.02684569358826 0.250491142272949
-0.986577153205872 0.255402803421021
-0.946308732032776 0.259332060813904
-0.906040191650391 0.26719057559967
-0.865771770477295 0.27308452129364
-0.825503349304199 0.275049090385437
-0.785234928131104 0.281925320625305
-0.744966506958008 0.28487229347229
-0.704697966575623 0.288801550865173
-0.664429545402527 0.288801550865173
-0.624161005020142 0.287819266319275
-0.583892583847046 0.29076623916626
-0.54362416267395 0.292730808258057
-0.503355741500854 0.29076623916626
-0.463087320327759 0.29076623916626
-0.422818779945374 0.289783954620361
-0.382550358772278 0.289783954620361
-0.342281818389893 0.288801550865173
-0.302013397216797 0.288801550865173
-0.261744976043701 0.289783954620361
-0.221476554870605 0.287819266319275
-0.18120801448822 0.288801550865173
-0.140939593315125 0.288801550865173
-0.100671172142029 0.288801550865173
-0.0604026317596436 0.288801550865173
-0.0201342105865479 0.288801550865173
0.0201342105865479 0.288801550865173
0.0604026317596436 0.288801550865173
0.100671172142029 0.288801550865173
0.140939593315125 0.288801550865173
0.18120801448822 0.288801550865173
0.221476554870605 0.287819266319275
0.261744976043701 0.289783954620361
0.302013397216797 0.288801550865173
0.342281818389893 0.288801550865173
0.382550358772278 0.289783954620361
0.422818779945374 0.289783954620361
0.463087320327759 0.29076623916626
0.503355741500854 0.29076623916626
0.54362416267395 0.292730808258057
0.583892583847046 0.29076623916626
0.624161005020142 0.287819266319275
0.664429545402527 0.288801550865173
0.704697966575623 0.288801550865173
0.744966506958008 0.28487229347229
0.785234928131104 0.281925320625305
0.825503349304199 0.275049090385437
0.865771770477295 0.27308452129364
0.906040191650391 0.26719057559967
0.946308732032776 0.259332060813904
0.986577153205872 0.255402803421021
1.02684569358826 0.250491142272949
1.06711411476135 0.246561884880066
1.10738253593445 0.245579600334167
1.14765095710754 0.236738681793213
1.18791949748993 0.235756397247314
1.22818791866302 0.230844736099243
1.26845633983612 0.224950909614563
1.30872488021851 0.217092275619507
1.3489933013916 0.214145421981812
1.3892617225647 0.20923376083374
1.42953014373779 0.199410557746887
1.46979868412018 0.196463704109192
1.51006710529327 0.190569758415222
1.55033552646637 0.181728839874268
1.59060406684875 0.176817297935486
1.63087248802185 0.168958783149719
1.67114090919495 0.160117864608765
1.71140944957733 0.154223918914795
1.75167787075043 0.149312376976013
1.79194629192352 0.14538311958313
1.83221471309662 0.13948917388916
1.872483253479 0.131630659103394
1.9127516746521 0.126718997955322
1.9530200958252 0.123772144317627
1.99328863620758 0.116895914077759
2.03355693817139 0.111984252929688
2.07382559776306 0.105108022689819
2.11409401893616 0.0982317924499512
2.15436244010925 0.0903732776641846
2.19463086128235 0.0834970474243164
2.23489928245544 0.0746562480926514
2.27516770362854 0.0697445869445801
2.31543612480164 0.0658153295516968
2.35570478439331 0.0648330450057983
2.39597320556641 0.060903787612915
2.4362416267395 0.0559921264648438
2.4765100479126 0.0550098419189453
2.51677846908569 0
2.55704689025879 0
2.59731554985046 0
2.63758397102356 0
2.67785239219666 0
2.71812081336975 0
2.75838923454285 0
2.79865765571594 0
2.83892607688904 0
2.87919473648071 0
2.91946315765381 0
2.9597315788269 0
3 0
};
\addlegendentry{ROMC Posterior}
\addplot [semithick, red, dash pattern=on 1pt off 3pt on 3pt off 3pt]
table {%
-3 0
-2.51677846908569 0
-2.4765100479126 0.0380961894989014
-2.39597320556641 0.0447498559951782
-2.35570478439331 0.0483826398849487
-2.31543612480164 0.0522257089614868
-2.27516770362854 0.0562825202941895
-2.23489928245544 0.0605562925338745
-2.19463086128235 0.0650490522384644
-2.15436244010925 0.0697618722915649
-2.11409401893616 0.0746949911117554
-2.07382559776306 0.0798472166061401
-2.03355693817139 0.0852166414260864
-1.9530200958252 0.0965919494628906
-1.872483253479 0.10877788066864
-1.79194629192352 0.121709108352661
-1.71140944957733 0.135297179222107
-1.63087248802185 0.14942991733551
-1.55033552646637 0.163971900939941
-1.22818791866302 0.222806215286255
-1.14765095710754 0.236687064170837
-1.10738253593445 0.243355870246887
-1.06711411476135 0.249807119369507
-1.02684569358826 0.256013989448547
-0.986577153205872 0.261949896812439
-0.946308732032776 0.267589211463928
-0.906040191650391 0.272907018661499
-0.865771770477295 0.277879595756531
-0.825503349304199 0.282484292984009
-0.785234928131104 0.286700010299683
-0.744966506958008 0.290507197380066
-0.704697966575623 0.293888092041016
-0.664429545402527 0.296826481819153
-0.624161005020142 0.29930853843689
-0.583892583847046 0.301322460174561
-0.54362416267395 0.302858352661133
-0.503355741500854 0.303908824920654
-0.422818779945374 0.304413080215454
-0.261744976043701 0.304565191268921
0.422818779945374 0.304413080215454
0.503355741500854 0.303908824920654
0.54362416267395 0.302858352661133
0.583892583847046 0.301322460174561
0.624161005020142 0.29930853843689
0.664429545402527 0.296826481819153
0.704697966575623 0.293888092041016
0.744966506958008 0.290507197380066
0.785234928131104 0.286700010299683
0.825503349304199 0.282484292984009
0.865771770477295 0.277879595756531
0.906040191650391 0.272907018661499
0.946308732032776 0.267589211463928
0.986577153205872 0.261949896812439
1.02684569358826 0.256013989448547
1.06711411476135 0.249807119369507
1.10738253593445 0.243355870246887
1.18791949748993 0.229828119277954
1.26845633983612 0.215649008750916
1.3489933013916 0.20103645324707
1.63087248802185 0.14942991733551
1.71140944957733 0.135297179222107
1.79194629192352 0.121709108352661
1.872483253479 0.10877788066864
1.9530200958252 0.0965919494628906
2.03355693817139 0.0852166414260864
2.07382559776306 0.0798472166061401
2.11409401893616 0.0746949911117554
2.15436244010925 0.0697618722915649
2.19463086128235 0.0650490522384644
2.23489928245544 0.0605562925338745
2.27516770362854 0.0562825202941895
2.31543612480164 0.0522257089614868
2.35570478439331 0.0483826398849487
2.4362416267395 0.0413227081298828
2.4765100479126 0.0380961894989014
2.51677846908569 0
3 0
};
\addlegendentry{True Posterior}
\end{axis}

\end{tikzpicture}

### 2021-08-24 latex_files/images/chapter4/mae2_hist_t1_romc.tex
plots after changes

@@
 tick pos=left,
-title={ROMC (gradient-based) - \(\displaystyle \theta_1\)},
+title={ROMC (gradient-based) - Histogram of the parameter \(\displaystyle \theta_1\)},
 x grid style={white!69.0196078431373!black},
@@
 ]
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.02018167274615);
-\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.39039538321726);
-\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.64837235793467);
-\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.83766715107148);
-\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.03114988210955);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.20118016090059);
-\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.42565363240795);
-\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.83858430947187);
-\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.02313277298727);
-\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.49050664276752);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.58012851385441);
-\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.63066296452957);
-\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.34895434170502);
-\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.09544450074894);
-\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.92569341781802);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.63923846537203);
-\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.52616414203809);
-\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.21569834562492);
-\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.07749639488345);
-\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.95995493778819);
-\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.85302292337607);
-\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.59560434037885);
-\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.66428652195943);
-\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.3535415296862);
-\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.45042249313527);
-\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.1846280343355);
-\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.19830863147962);
-\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.979977468894081);
-\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.941727636062617);
-\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.705807134291806);
-\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.614789283904501);
-\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.496689435089055);
-\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.456206035376906);
-\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.425773686627776);
-\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.449226138874808);
-\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.309907404693004);
-\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.224752667367446);
-\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.192924339317897);
-\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.0974393551692405);
-\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.176730979433038);
-\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.170309474651111);
-\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.145461043103651);
-\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0304323487491329);
-\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0206604936462);
-\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0);
-\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0217772770865351);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.23969307221967);
+\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.22260605418476);
+\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.58099417689925);
+\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.86697039022021);
+\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,1.9327905830084);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.28560027755721);
+\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.65528249842742);
+\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.76542029517717);
+\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.08006607906211);
+\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.37496095644731);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.55021219168023);
+\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.48979291667414);
+\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.29497485978374);
+\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.21369852560985);
+\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.95263129250428);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.68186180204391);
+\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.34177450679284);
+\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.16665229531096);
+\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.90550699050888);
+\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.80681861350051);
+\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.82068989944754);
+\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.66308656575326);
+\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.55350554347052);
+\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.55706027108261);
+\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.34606411506585);
+\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.25058859381965);
+\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.02541420681756);
+\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.08883061316133);
+\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.901026681844632);
+\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.90959073604536);
+\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.663388353950532);
+\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.635304321110202);
+\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.584894248497155);
+\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.541909616027169);
+\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.364723435432413);
+\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.337243430980688);
+\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.304888876342628);
+\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.11082071864216);
+\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.14353193766205);
+\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.12562311229572);
+\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.107714286929387);
+\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.0820282878821482);
+\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0827149079075519);
+\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0269285717323464);
+\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0307869571533697);
+\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0);
 \draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
@@
 table {%
-0.503688298846839 0
-0.503688298846839 3.5
+0.50403516578119 0
+0.50403516578119 3.5
 };
@@
 table {%
-0.323676768279387 1
-0.634437489183658 1
+0.321847690178847 1
+0.637029674539771 1
 };
-\addlegendentry{$\sigma = 0.141$}
+\addlegendentry{$\sigma = 0.143$}
 \end{axis}

### 2021-08-24 latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex
plots after changes

@@
 tick pos=left,
-title={ROMC (Bayesian Optimisation) - \(\displaystyle \theta_1\)},
+title={ROMC (Bayesian Optimisation) - Histogram of the parameter \(\displaystyle \theta_1\)},
 x grid style={white!69.0196078431373!black},
@@
 ]
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.67339010906917);
-\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,2.12285439456296);
-\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,2.41994534810638);
-\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.13421448089961);
-\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.60269456308737);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.61479552461991);
-\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.51971654115006);
-\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.94571977877467);
-\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.78840727885186);
-\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.70814579929938);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.89534026545558);
-\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.97807132899427);
-\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.47304140381032);
-\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,2.18508791101595);
-\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.50193553644923);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.27646251850645);
-\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,1.84996536408459);
-\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.10951864103732);
-\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.94010517958197);
-\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.70154336651221);
-\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.88058820551382);
-\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.72549833117864);
-\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.29579071757466);
-\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.50076618843174);
-\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.46940247180661);
-\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.16860714228384);
-\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.37012519556019);
-\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.767793661318775);
-\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.885098900664694);
-\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.65715629873569);
-\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.82731063538692);
-\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.813727923462651);
-\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.801873920328749);
-\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.459095663040085);
-\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.672961636247559);
-\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.561583398468614);
-\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.382538559466961);
-\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.340555631701063);
-\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.294127452759942);
-\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.239055729867026);
-\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.283755200017783);
-\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.176575255015415);
-\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.283261283220537);
-\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.258071526560991);
-\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0775449371676099);
-\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0155583791132463);
-\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
-\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0.0172870879036073);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.59230812208571);
+\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,2.04310565145056);
+\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,2.19382815490065);
+\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.20021699441589);
+\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.51506116421806);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.44880980882892);
+\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.65904829491329);
+\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.73250343027527);
+\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.82852250487442);
+\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.82331312648089);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.90135211837164);
+\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.88136478592358);
+\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.53560317010369);
+\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,2.3564598017404);
+\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.22420707945385);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.33534256536061);
+\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.00495652659749);
+\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,1.74992091039568);
+\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.83870977452066);
+\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.6578788920657);
+\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.55026416918985);
+\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.62152961093496);
+\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.6437730803435);
+\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.66640493128548);
+\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.32365272171391);
+\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.14155317519524);
+\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.17480604967746);
+\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.20896637137188);
+\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,1.26617555858479);
+\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,1.10815016078187);
+\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,1.01575133068398);
+\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.778653581806476);
+\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.658517776017212);
+\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.522857721601219);
+\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.649384569447495);
+\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.571069158563703);
+\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.362376122626364);
+\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.275087556542002);
+\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.255847252846728);
+\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.406941618766892);
+\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.192257669195406);
+\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.204927790769102);
+\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.203525138747285);
+\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.113584712363094);
+\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0662010640185658);
+\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.116253091726246);
+\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0.0456418048890116);
+\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
 \draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
@@
 table {%
-0.500240536632723 0
-0.500240536632723 3.5
+0.502312841390367 0
+0.502312841390367 3.5
 };
-\addlegendentry{$\mu = 0.500$}
+\addlegendentry{$\mu = 0.502$}
 \addplot [semithick, black, dashed]
 table {%
-0.293929386789429 1
-0.656599793802562 1
+0.294820390999603 1
+0.660267860059204 1
 };
-\addlegendentry{$\sigma = 0.165$}
+\addlegendentry{$\sigma = 0.166$}
 \end{axis}

### 2021-08-24 latex_files/images/chapter4/mae2_hist_t2_romc.tex
plots after changes

@@
 tick pos=left,
-title={ROMC (gradient-based) - \(\displaystyle \theta_2\)},
+title={ROMC (gradient-based) - Histogram of the parameter \(\displaystyle \theta_2\)},
 x grid style={white!69.0196078431373!black},
@@
 \draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
-\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0132189786765283);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0698717444330784);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.203477850342276);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.355023998741048);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.514595955621997);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.712880635769923);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.916830592493502);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.952710677472651);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.1953733574632);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.3804390589346);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.52584782437642);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.59477535604688);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.7604846958848);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.91108663152097);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.09048705641672);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.27366433236291);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.32654024706902);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.32811393500671);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.26296325438667);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.10968604925645);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15437878668663);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.78361790856873);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.58785112912109);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.44590447714217);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.22086710205366);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.01125186875443);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.805413486505622);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.800220316311271);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.62758674954756);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.600834054606967);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.396254622708316);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.300574396097251);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.197497836179084);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.168384609331968);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.156581949799354);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.068455425289165);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0550790778188679);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.066094893382642);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0275395389094342);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0275395389094339);
+\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00554428884192465);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0609051679887177);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.172932144280507);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.33874934889258);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.484235646321314);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.665208164482786);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.92463814505837);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.12139409849263);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.18629739177248);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.38911741112759);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.3586665597235);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.4168730451184);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.86440804043856);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.77193230571172);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.14800442101592);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.34373098482187);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.24893542723874);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.36527262608105);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.32011000423491);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.35112752816106);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.09237672296862);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.72225524667604);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65885907591305);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.4089028988961);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.18892677075576);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.11057927007035);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.808010795099994);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.746145310081448);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.68232985247479);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.586634734027065);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.400470451389202);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.259104715630191);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.230773630659991);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.211235556781049);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.129869421833243);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0903839207491964);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0434117816322698);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0482353129247447);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0289411877548468);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0144705938774233);
 \draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
@@
 table {%
-0.0308218751896741 0
-0.0308218751896741 3
+0.0325607299281219 0
+0.0325607299281219 3
 };
-\addlegendentry{$\mu = 0.031$}
+\addlegendentry{$\mu = 0.033$}
 \addplot [semithick, black, dashed]
 table {%
--0.179871006878183 1
-0.197679132295466 1
+-0.17677281465975 1
+0.198406420501618 1
 };
-\addlegendentry{$\sigma = 0.172$}
+\addlegendentry{$\sigma = 0.171$}
 \end{axis}

### 2021-08-24 latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex
plots after changes

@@
 tick pos=left,
-title={ROMC (Bayesian Optimisation) - \(\displaystyle \theta_2\)},
+title={ROMC (Bayesian Optimisation) - Histogram of the parameter \(\displaystyle \theta_2\)},
 x grid style={white!69.0196078431373!black},
@@
 \draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
-\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0710656818158832);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.0516473550957937);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.194048417709784);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.2505503544856);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.348585934523831);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.648086654282988);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.583763447022691);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.929922090706508);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.79776958941701);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,0.987907371884556);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.05978215064711);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.2930717702704);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.55481463251829);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.01101046094927);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.35069632905971);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.34206596162857);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.80716185647295);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.71681269742808);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.65275918915001);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.45763197551135);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.24848041479702);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.8837125413398);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.80455589005721);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.47147764701123);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.35887832193294);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.990739211197902);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.834853199472739);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.568255755544842);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.531441844471339);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.415606131606362);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.366790615823915);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.350473827399395);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.220883466441019);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.136737383987297);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.185283200787522);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.154672366305157);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0558276893202577);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.179619522160829);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0788869523003634);
-\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0133500996200616);
-\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.0134849491111732);
-\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.00943946437782135);
-\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0.0140243470756202);
-\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0.00337123727779334);
+\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00425643655908195);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0584436505421199);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.032329412084104);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.170488346927624);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.241312187040701);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.353203225767864);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.454783676389647);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.645870310110015);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.736855945766172);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.7054880882772);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,0.853720954547648);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.16488627133908);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.37944888929298);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.75863459008645);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.04186645728356);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,1.95487930439736);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.2931926651875);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.65546746491156);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.64247045528465);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.53538619434107);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.47437216167486);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.25894133814364);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,2.12306099510411);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.79209215192117);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.53877177811189);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.34176813031492);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.869256102075021);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.984211536278622);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.744495637346589);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.487332826886714);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.631223471040121);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.5510631368371);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.369415701030125);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.376101401475276);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.135239439335051);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.17253048109158);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.187926500765093);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.106413103399642);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0572991796412681);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0435129826888125);
+\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0491139237583741);
+\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.011416843630065);
+\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.0114168436300651);
+\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,3.98076835078974e-05);
+\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0);
 \draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
@@
 table {%
-0.070389896781954 0
-0.070389896781954 3
+0.0793529734576715 0
+0.0793529734576715 3
 };
-\addlegendentry{$\mu = 0.070$}
+\addlegendentry{$\mu = 0.079$}
 \addplot [semithick, black, dashed]
 table {%
--0.131187651225067 1
-0.236045424145365 1
+-0.122023559466631 1
+0.246600101073509 1
 };
-\addlegendentry{$\sigma = 0.167$}
+\addlegendentry{$\sigma = 0.168$}
 \end{axis}

### 2021-08-24 latex_files/python_scripts/example_1D.py

@@ -4,6 +4,7 @@
 import os
 import scipy.stats as ss
 import scipy.integrate as integrate
+
 import matplotlib.pyplot as plt
 import timeit
 import numpy as np

### 2021-08-24 latex_files/python_scripts/example_MA2.py

@@ -98,7 +98,6 @@ else:
 print(romc.result.summary())
 print(romc.result.samples_cov())
 
-
 name = "t1"
 if savefig:
     plot_marginal(romc.result.samples["t1"], romc.result.weights,

### 2021-08-25 latex_files/ch2/2_4_ROMC.tex
changed notation to be inline with ROMC paper

@@
 \item The whole acceptable region $\mathcal{S}_i$
-  shrinks to a single point $\thetab_i^*$/. This simplification adds
+  shrinks to a single point $\thetab_i^*$. This simplification adds
   significant apprimation error when then the area $\mathcal{S}_i$ is relatively big.
@@
 ROMC resolves the failure-modes of OMC mainly by approaching the
-acceptance regions $\mathcal{S}_i$ in a more sophisticated
-way. Instead of collapsing the acceptance regions into single points,
-it approximates them with a bounding box region. A uniform distribution
-defined on the bounding box area, is then used as the \emph{proposal
-  distribution} for importance sampling. If $q_i$ is the uniform
-distribution defined on the $i-th$ bounding box, weighted sampling is
-performed by drawing samples from $q_i$ and assigning the weights as
+acceptance regions $\mathcal{S}_i$ in a more sophisticated way.
+Instead of collapsing the regions into single points, it approximates
+them with a bounding box. Afterwards, a uniform distribution defined
+on the bounding box area, is used as the \emph{proposal distribution}
+for importance sampling. Defining as $q_i$ the uniform distribution
+defined on the $i-th$ bounding box, weighted sampling is performed by
+drawing samples from $q_i$ and assigning the weights as
 in~\eqref{eq:sampling}.
@@
 For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$,
-we search for a point $\thetab^*$ such that
-$d(M_d(\thetab^*,\vb_i), \data) < \epsilon$. For notation convenience,
-we name the function $d(M_d(\thetab^*,\vb_i)$ as $d_i(\thetab)$.  Obtaining
-this point involves solving the following optimisation problem:
+we search for a point $\thetab^*_i$ such that
+$d(M_d(\thetab^*_i,\vb_i), \data) < \epsilon$. For notation
+convenience, we name the distance function $d(M_d(\thetab,\vb_i), \data)$ as
+$d_i(\thetab)$.  Obtaining this point involves solving the following
+optimisation problem:
 
@@
 iterative algorithm can be used for~\ref{eq:optProb}. The gradients
-$\nabla_{\thetab} g_i$ can be either provided in closed form or
+$\nabla_{\thetab} d_i$ can be either provided in closed form or
 approximated by finite differences. Alternatively, if $d_i$ is not
-differentiable, the Bayesian Optimisation approach provides an
-alternative choice~\cite{Shahriari2016}. In this case, apart from
-obtaining an optimal $\thetab_i^* $, a surrogate model $\hat{d}_i$ of
-the distance $d_i$ is also fitted; the approximate model can then
-substitute the actual model in the following steps.
+differentiable, the Bayesian Optimisation
+approach~\cite{Shahriari2016} provides an alternative choice. In this
+case, apart from obtaining an optimal $\thetab_i^* $, a surrogate
+model $\hat{d}_i$ of the distance $d_i$ is also fitted; the
+approximate model can then substitute the actual model in the
+following steps.
 
@@
 
-After obtaining $\thetab_i^*$, we approximate the acceptance region
-$\mathcal{S}_i$ by constructing a bounding box $\mathcal{\hat{S}}_i$
-around it. The bounding must the part of the acceptance region
-$\mathcal{S}_i$ around $\thetab_i^*$, formally
-$\{ \thetab : d_i(\thetab) < \epsilon$,
-$d(\thetab, \thetab_i^*) < M \}$. The second condition
+After obtaining $\thetab_i^*$, we initially approximate the acceptance
+region $\mathcal{S}_i$ by constructing a bounding box
+$\mathcal{\hat{S}}_i$ around it. The bounding must the part of the
+acceptance region $\mathcal{S}_i$ \emph{around} $\thetab_i^*$;
+formally $\{ \thetab : d_i(\thetab) < \epsilon$,
+$d(\thetab, \thetab_i^*) < M \}$. The condition
 $d(\thetab, \thetab_i^*) < M$ describes that if
-$\mathcal{S}_i := \{ \thetab : d_i(\thetab) < \epsilon \} $ contains a
-number of disjoint sets of $\thetab$ that they fulfill
+$\mathcal{S}_i := \{ \thetab : d_i(\thetab) < \epsilon \} $ contains
+many \emph{disjoint} sets of $\thetab$ that they fulfill
 $d_i(\thetab) < \epsilon$, we want the bounding box to match only the
-set that contains $\thetab_i^*$. We seek for a bounding box that is as
-tight as possible to the local acceptance region for ensuring high
-acceptance rate in the importance sampling.
+set that contains $\thetab_i^*$. Ideally, we want the bounding box to
+be as tight as possible to $\mathcal{S}_i$, to ensure high acceptance
+rate in the importance sampling, and big enough for not discarding
+valid parts of $\mathcal{S}_i$.
 
-Bounding boxes are built in two steps. Firstly, the search directions
-$\mathbf{v}_d$ are computed as the eigenvectors of the curvature at
-$\thetab_i^*$ and then the limits are obtained with a line-search
-method looking for the smallest
+The bounding boxes are built in two steps; firstly, we define their
+axes $\mathbf{v}_d$ in the $d$-dimensional space and, then, the limits
+are obtained with a line-search method looking for the smallest
 $\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
 \epsilon$\footnote{$-\kappa$ is used as well for the opposite
-  direction along the search line}. The
-Algorithm~\ref{alg:region_construction} describes the method
-in-depth. After limits are found along all directions the uniform
-distribution $q_i$ can be defined and is used as the proposal region
-for importance sampling.
+  direction along the search
+  line}. Algorithm~\ref{alg:region_construction} describes the method
+in-depth. After limits are found along all directions, the uniform
+distribution $q_i$ is defined on the bounding box, and is used as the
+proposal region for importance sampling.
 
@@
 After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
-are no longer interested in $g_i(\thetab)$ outside
-$\mathcal{\hat{S}}_i$. In the inference steps (e.g. sampling,
-evaluating the posterior), we evaluate $g_i$ for checking if a sample
-from the bounding box belongs to the acceptance region. Instead of
-running the deterministic simulator $M_d(\thetab_{ij}, \vb_i)$ for
-evaluating $g_i$, ROMC proposes fitting a \emph{light} surrogate model
-$\hat{g}_i$ for approximating $g_i$ around $\theta_i^*$.
+are no longer interested in $d_i(\thetab)$ outside
+$\mathcal{\hat{S}}_i$. In the inference steps
+(i.e. sampling~\eqref{eq:sampling}, computing
+expectation~\eqref{eq:expectation}, evaluating the
+posterior~\eqref{eq:approx_posterior_omc}), we need to evaluate the
+distance function $d_i$ for checking if the samples obtained from the
+uniform distribution $\thetab_{ij} \sim q_i$ lie inside the acceptance
+region $\mathcal{S}_i$. Evaluating $d_i$ involves running the
+deterministic simulator $M_d(\thetab_{ij}, \vb_i)$, which can be
+inneficient when the model is big.\footnote{If Bayesian Optimization
+  has been used for obtaining $\thetab_i^*$, then $\hat{d}_i$ can be
+  used as an alternative.}. Hence, a \emph{light} surrogate model
+$\Tilde{d}_i$ can be fit for approximating $d_i$ around $\theta_i^*$.
 
@@
   consider the properties of the local region (i.e. size,
-  smoothness). } by sampling $N$ points from the bounding box and the
-labels $Y: \mathbb{R}^{N}$ are computed by evaluating $g_i$. This
-additional step adds an extra effort in the training part, but
-promises a significant speed-up at the inference phase. Such imbalance
-is usual in the context of ML where is usual to be quite generous with
-the execution time at the training phase but quite eager at the
-inference phase. Fitting a local surrogate model aligns with this
-requirement.
+  smoothness). } by sampling training points from the bounding box and
+labeling them by evaluating $d_i$. This additional step adds an extra
+effort in the training part, but promises a significant speed-up at
+the inference phase. Such time-badget imbalance between the training
+and the inference part is demanded in many ML contexts; it is usual to
+be quite generous with the execution time at the training phase but
+quite eager at the inference phase. Fitting a local surrogate model
+aligns with this requirement.

### 2021-08-25 latex_files/ch2/2_5_ROMC_algorithmic.tex
minor improvements
changed notation to be inline with ROMC paper

@@
 ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
-serves as the driver for the implementation at \pkg{ELFI} that
-follows. At a high level, ROMC can be split into the training and the
-inference part. In general terms, the training part covers all the
-appropriate steps for estimating the proposal regions and the
-inference part for getting weighted
-samples. Algorithm~\ref{alg:romc_algorithm} defines formally ROMC;
+serves as the driver for the implementation at \pkg{ELFI}. At a high
+level, ROMC can be split into the training and the inference part. In
+general terms, the training part covers all steps for estimating the
+proposal regions and the inference part for getting weighted
+samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
 Steps 2-10 (before the vertical line) form the training part, while
-steps 12 and 13 the inference part.
+steps 12-17 the inference part.
 
@@
 	\begin{algorithmic}[1]
-    \Procedure{ROMC}{$M_r(\thetab)$}
+    \Procedure{ROMC}{$M_r(\thetab), n_1, n_2, \epsilon$}
     \For{$i \gets 1 \textrm{ to } n_1$}
-      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define deterministic distance function}
-      \State Obtain $\thetab_i^*, d_i^*$ \Comment{Solve optimization problem}
+      \State $\vb_i \sim p(\vb)$ \Comment{Draw nuisance variables}
+      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define distance function}
+      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimization problem}
+      \If{$d^*_i > \epsilon$}
+        \State Go to 2 \Comment{Filter solution}
+      \EndIf
+      \State Define $q_i$ over $\mathcal{\hat{S}}$ \Comment{Estimate proposal area}
+      \State (Optional) Fit $\Tilde{d}_i$ on $\mathcal{\hat{S}}$ \Comment{Fit surrogate model}
+      \\\hrulefill
+      \For{$j \gets 1 \textrm{ to } n_2$}      
+        \State $\thetab_{ij} \sim q_i$, compute $w_{ij}$ \Comment{Sample}
       \EndFor
-    \State $\text{L} \gets \{d_{i, i = 1, ..., n_1}: d_i^* < \epsilon \} $ \Comment{List with filtered solutions}
-    \For{$i \gets 1 \textrm{ to } len(L)$}
-      \State Construct $q_i$ \Comment{Compute proposal area}
-      \State Compute $\Tilde{d}_i$ \Comment{Fit surrogate model in $q_i$}
     \EndFor
-    \\\hrulefill
-    \State Draw $\{w_{ij}, \thetab_{ij}\}$ \Comment{Draw $n_2$ samples per $q_i$}
-    \State Draw $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}    
+    \State $E_{p(\thetab|\data)}[h(\thetab)]$ \Comment{Estimate an expectation}
+    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}
     \EndProcedure
@@
 At the training (fitting) part, the goal is the estimation of the
-proposal regions $q_i$. The tasks are; (a) sampling the nuisance
-variables $\vb_i \sim p(\vb)$ for obtaining the deterministic
-functions $g_i(\thetab)$, (b) defining the optimisation problems
-$\min_{\thetab} \: g_i(\thetab)$, (c) obtaining $\thetab_i^*$ and
-$d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and (e) building
-the bounding box for obtaining the proposal region $q_i$. If gradients
-are available, using a gradient-based method is advised for obtaining
+proposal regions $\mathcal{\hat{S}}_i$. The tasks are; (a) sampling
+the nuisance variables $\vb_i \sim p(\vb)$ for obtaining the
+deterministic functions $d_i(\thetab)$, (b) defining the optimisation
+problems $\min_{\thetab} \: g_i(\thetab)$, (c) obtaining
+$\thetab_i^*$, $d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and
+(e) computing the proposal distribution $q_i$. If gradients are
+available, using a gradient-based method is advised for obtaining
 $\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
@@
 scenario, the training part becomes slower due to fitting of the
-surrogate model and the blind optimisation
-steps. Algorithm~\ref{alg:training_GB} defines formally the steps of a
-gradient-based optimiser and algorithm~\ref{alg:training_GP} the steps
-of the Bayesian Optimisation. After obtaning the optimal points
-$\thetab^*$, the training part gets completed by estimating the
-proposal regions. Algorithm~\ref{alg:region_construction} describes
-the line search approach for finding the region's boundaries.
+surrogate model and the blind optimisation steps. After obtaning the
+optimal points $\thetab^*$, the training part gets completed by
+estimating the proposal regions;
+Algorithm~\ref{alg:region_construction} describes the line search
+approach for finding the region's boundaries.
 
@@
 expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
-unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Computing an
-expectation can be done easily after weighted samples are obtained
-using the equation~\ref{eq:expectation}, so we do not discuss it
-separately.
+unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Sampling is
+performed by getting $n_2$ samples from each proposal distribution
+$q_i$. For each sample $\thetab_{ij}$, the distance
+function\footnote{As before, if a surrogate model $\hat{d}$ is
+  available, it can be utilised as the distance function.} is
+evaluated for checking if it lies inside the acceptance
+region. Algorithm~\ref{alg:sampling_GB} defines the steps for
+computing a weighted sample. Computing an expectation can be done
+easily after weighted samples are obtained using the
+equation~\ref{eq:expectation}, so we do not discuss it
+separately. Evaluating the unnormalised posterior requires the
+deterministic functions $g_i$ and the prior distribution
+$p(\thetab)$\footnote{There is no need for solving the optimisation
+  problems and building the proposal regions.}. The evaluation
+requires iterating over all $g_i$ and evaluating the distance from the
+observed data.
 
-\noindent
-Sampling is performed by getting $n_2$ samples from each proposal
-distribution $q_i$. For each sample $\thetab_{ij}$, the indicator
-function is evaluated $\indicator{\regioni(\data)}(\thetab_{ij})$ for
-checking if it lies inside the acceptance region. If this is the case,
-the corresponding weight is computed as in \eqref{eq:sampling}. As
-before, if a surrogate model $\hat{d}$ is available, it can be
-utilised for evaluating the indicator function.
-
-\noindent
-Evaluating the unnormalised posterior requires the deterministic
-functions $g_i$ and the prior distribution $p(\thetab)$\footnote{There
-  is no need for solving the optimisation problems and building the
-  proposal regions.}. The evaluation requires iterating over all $g_i$
-and evaluating the distance from the observed data.
-
-\begin{minipage}{0.46\textwidth}
-\begin{algorithm}[H]
-    \centering
-    \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
-    \begin{algorithmic}[1]
-      \For{$i \gets 1 \textrm{ to } n$}
-        \State Obtain $\thetab_i^*$ using a Gradient Optimiser
-        \If{$g_i(\thetab_i^*) > \epsilon$}
-        \State{go to} 1
-        \Else
-        \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
-        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
-        \EndIf      
-      \EndFor
-      \Return{$q_i, p(\theta), g_i(\theta)$}
-    \end{algorithmic}
-\end{algorithm}
-\end{minipage}
-\hfill
-\begin{minipage}{0.46\textwidth}
-\begin{algorithm}[H]
-    \centering
-    \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
-    \begin{algorithmic}[1]
-      \For{$i \gets 1 \textrm{ to } n$}
-        \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
-        \If{$g_i(\thetab_i^*) > \epsilon$}
-        \State{go to} 1
-        \Else
-        \State Approximate $H_i \approx \jac^T_i \jac_i$
-        \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
-        \EndIf      
-      \EndFor
-      \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
-    \end{algorithmic}
-\end{algorithm}
-\end{minipage}
+% \begin{minipage}{0.46\textwidth}
+% \begin{algorithm}[H]
+%     \centering
+%     \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
+%     \begin{algorithmic}[1]
+%       \For{$i \gets 1 \textrm{ to } n$}
+%         \State Obtain $\thetab_i^*$ using a Gradient Optimiser
+%         \If{$g_i(\thetab_i^*) > \epsilon$}
+%         \State{go to} 1
+%         \Else
+%         \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
+%         \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
+%         \EndIf      
+%       \EndFor
+%       \Return{$q_i, p(\theta), g_i(\theta)$}
+%     \end{algorithmic}
+% \end{algorithm}
+% \end{minipage}
+% \hfill
+% \begin{minipage}{0.46\textwidth}
+% \begin{algorithm}[H]
+%     \centering
+%     \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
+%     \begin{algorithmic}[1]
+%       \For{$i \gets 1 \textrm{ to } n$}
+%         \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
+%         \If{$g_i(\thetab_i^*) > \epsilon$}
+%         \State{go to} 1
+%         \Else
+%         \State Approximate $H_i \approx \jac^T_i \jac_i$
+%         \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
+%         \EndIf      
+%       \EndFor
+%       \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
+%     \end{algorithmic}
+% \end{algorithm}
+% \end{minipage}
 
 \begin{algorithm}[!ht]
-	\caption{Computation of the proposal distribution $q_i$; Needs, a model of distance $d$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
+	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; Requires, a model of distance $d_i(\thetab)$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta\_\text{start}$, maximum iterations $M$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
 	\begin{algorithmic}[1]
@@
 		\State $k \gets 0$
+		\State $\eta \gets \eta\_\text{start}$ \Comment{Initialize $\eta$}
 		\Repeat
+          \State $j \gets 0$
         	\Repeat
             \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
-        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$}
+            \State $j \gets j + 1$
+        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$ or $j \geq M$} \Comment{Check distance or maximum iterations}
         	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
@@
       \Until $k = K$
+      \If{$\Tilde{\thetab} = \thetab_i^*$} \Comment{Check if no step has been done}
+        \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d$ \Comment{Then, make the minimum step}
+      \EndIf
     	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
@@
     \centering
-    \caption{Sampling. Requires a function of distance $(g_i(\theta)$ or $\hat{d}_i$ or $\hat{g}_i), p(\theta), q_i$}\label{alg:sampling_GB}
+    \caption{Sampling. Requires a function of distance $d_i$, the prior distribution $p(\thetab)$, the proposal distribution $q_i$}\label{alg:sampling_GB}
     \begin{algorithmic}[1]
-      \For {$i \gets 1 \textrm{ to } n_1$}
-      \For {$j \gets 1 \textrm{ to } n_2$}
       \State $\thetab_{ij} \sim q_i$
-          \If {$g_i(\thetab_{ij}) > \epsilon$}
-            \State Reject $\theta_{ij}$
+          \If {$d_i(\thetab_{ij}) > \epsilon$}
+            \State Go to 2 \Comment{Reject sample}
           \Else {}
-            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$
-            \State Accept $\thetab_{ij}$, with weight $w_{ij}$
+            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$ \Comment{Compute weight}
+            \State Store $(w_{ij}, \thetab_{ij})$ \Comment{Store weighted sample}
           \EndIf
-      \EndFor
-      \EndFor
     \end{algorithmic}

### 2021-08-25 latex_files/ch3/3_2_1_general_design.tex
minor improvements
changed notation to be inline with ROMC paper

@@
 whereas the rest (blue ellipses) are the public methods that the user
-interacts with.
-
-In figure~\ref{fig:romc_overview}, it can be easily observed that the
-implementation follows consistently the algorithmic description
-described above, which follows the partition into training, inference
-and evaluation. The training part includes all the steps until the
-computation of the proposal regions i.e. sampling the nuisance
-variables, defining the optimization problems, solving them,
-constructing the regions and fitting local surrogate models. The
-inference part comprises of evaluating the unnormalized posterior (and
-the normalized when is possible), sampling and computing an
-expectation. From an algorithmic point of view, the training and the
-inference parts are \emph{isolated}. We also provide some utilities
-for inspecting the training process, such as plotting the histogram of
-the final distances or visualizing the constructed bounding
+interacts with. In figure~\ref{fig:romc_overview}, it can be easily
+observed that the implementation follows consistently the algorithmic
+view presented in~\ref{alg:romc_algorithm}. The training part includes
+all the steps until the computation of the proposal regions
+i.e. sampling the nuisance variables, defining the optimization
+problems, solving them, constructing the regions and fitting local
+surrogate models. The inference part comprises of evaluating the
+unnormalized posterior (and the normalized when is possible), sampling
+and computing an expectation. We also provide some utilities for
+inspecting the training process, such as plotting the histogram of the
+final distances or visualizing the constructed bounding
 boxes. Finally, in the evaluation part, we provide two methods for

### 2021-08-25 latex_files/ch3/3_2_2_training.tex
changed notation to be inline with ROMC paper

@@
 definition of the optimisation problems is performed by drawing
-\code{n_1} integer numbers from a discrete uniform distribution
+\code{n1} integer numbers from a discrete uniform distribution
 $u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
@@
 fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
-scenario, in the following steps, $\hat{d}_i(\thetab)$ will replace
-$g_i(\thetab)$ as the distance function.
+scenario, in the following steps, $\hat{d}_i(\thetab)$ replaces
+$d_i(\thetab)$ as the distance function.
 
@@
 use. It plots a histogram of the distances at the optimal point
-$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $d_i^*$ in case
+$d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $\hat{d}_i^*$ in case
 \code{use_bo=True}. The function accepts all keyword arguments and
@@
 Algorithm~\ref{alg:region_construction}. The Hessian matrix is
-approximated based on the Jacobian $\hess_i = \jac_i^T \jac_i$. The
-eigenvectores are computed using the function
-\code{numpy.linalg.eig()}. A check is performed so that the matrix
-$\hess_i$ is not singular; if this is the case, the eigenvectors are
-set to be the vectors of the standard Euclidean basis i.e.\
-$\{ \mathbf{e_1} = (1, 0, \ldots), \mathbf{e_2} = (0,1,0,\ldots),
-\text{etc} \}$. The limits of the bounding box are obtained by
-repeteadly querying the distance function ($g_i(\thetab)$ or
-$\hat{d}(\thetab)$) along the search directions given by $\hess_i$.
+approximated based on the Jacobian
+$\hess_i = \jac_i^T \jac_i$\footnote{The Jacobian matrix is computed
+  on the output of the summary statistics
+  $\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))$ at $\thetab_i^*$} if
+gradients have been used, or as the Hessian of the GP, in case of
+Bayesian Optimisation. The eigenvectors, which are the axes of the
+bounding box, are computed using the function
+\code{numpy.linalg.eig()}. Afterwards, the limits along the
+eigenvectors are computed with
+Algorithm~\ref{alg:region_construction}.
 

### 2021-08-25 latex_files/graphs/romc_graph.tex
applied fixes
fixed warnings
changed notation to be inline with ROMC paper

@@
         \node (n2) [train_process, below of=n1] { $\_define\_objectives()$  };
-        \node (n3) [decision, below of=n2, yshift=-.5cm] { $grads$  };
+        \node (n3) [decision, below of=n2, yshift=-.5cm] { $grads?$  };
         \node (n4) [train_process, left of=n3, yshift=-1cm, xshift=-1.5cm] { $\_solve\_gradients()$  };
@@
         \node (n7) [train_process, below of=n6] { $\_build\_boxes()$  };
-        \node (n8) [decision, below of=n7] { $h?$  };
-        \node (n9) [train_process, right of=n8, yshift=-.5cm, xshift=1cm] { $\_fit\_models()$  };
+        \node (n8) [decision, below of=n7] { $fit?$  };
+        \node (n9) [train_process, right of=n8, yshift=-1cm, xshift=1cm] { $\_fit\_models()$  };
 
@@
         \node (proposal_region) [train_process, below of=filter, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $q_i \forall i$};
-        \node (surrogate) [train_process, below of=proposal_region, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Construct $\Tilde{d}_i \forall i$};
-        \node (posterior) [train_process, below of=surrogate, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Define $p_{d,\epsilon}(\thetab)$};
+        \node (surrogate) [train_process, below of=proposal_region, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Fit $\Tilde{d}_i \forall i$};
+        \node (posterior) [train_process, below of=surrogate, yshift=-0.15cm, minimum width=4cm, minimum height=1cm] {Define $p_{d,\epsilon|\data}(\thetab)$};
 
@@
 provided to the user.}
+\label{fig:romc_overview}
 \end{figure}

### 2021-08-25 paper.tex
added argmin and argmax macros
plots after changes
changed notation to be inline with ROMC paper

@@
 \newcommand{\R}{\mathbb{R}}
+
+\DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
 
@@
 
-In this section, we present our implementation. At~\ref{subsec:ELFI}
+In this section, we present the implementation. At~\ref{subsec:ELFI}
 we introduce the reader to the \proglang{Python} package \pkg{Engine
-  for likelihood-free inference (ELFI)}
-package. At~\ref{subsec:general_design} we explain the general design
-that we follow. Then, we divide the implementation into three parts;
-at~\ref{subsec:training} we present the training (fitting) part,
-at~\ref{subsec:inference} the inference part and
+  for likelihood-free inference
+  (ELFI)}. At~\ref{subsec:general_design} we explain the general
+design that we follow. Then, we divide the implementation into three
+parts; at~\ref{subsec:training} we present the training (fitting)
+part, at~\ref{subsec:inference} the inference part and
 at~\ref{subsec:evaluation} the evaluation part. Finally,
@@
 \section{Use-case illustration}
+
+In this section, we illustrate the implementation performing two
+experiments; In chapter~\ref{subsec:ma2} we perform the inference at
+the second-order moving average (MA2) model, which is a standard model
+of \pkg{ELFI}, and in chapter~\ref{subsec:exec_time} we measure the
+execution speed-up achieved by the parallelization of ROMC.
+
 \subsection{Moving average example}
+\label{subsec:ma2}
 \input{latex_files/ch4/4_2_ma2.tex}
@@
 \subsection{Execution time}
+\label{subsec:exec_time}
 \input{latex_files/ch4/4_3_exec.tex}

### 2021-08-25 latex_files/python_scripts/example_MA2.py

@@ -103,7 +103,7 @@ if savefig:
     plot_marginal(romc.result.samples["t1"], romc.result.weights,
                   romc.result.sample_means_array[0],
                   np.sqrt(romc.result.samples_cov()[0, 0]),
-                  r"ROMC (gradient-based) - Histogram of the parameter $\theta_1$",
+                  r"ROMC (gradient-based) - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -112,7 +112,7 @@ else:
     plot_marginal(romc.result.samples["t1"], romc.result.weights,
                   romc.result.sample_means_array[0],
                   np.sqrt(romc.result.samples_cov()[0, 0]),
-                  r"ROMC (gradient-based) - Histogram of the parameter $\theta_1$",
+                  r"ROMC (gradient-based) - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -123,7 +123,7 @@ if savefig:
     plot_marginal(romc.result.samples["t2"], romc.result.weights,
                   romc.result.sample_means_array[1],
                   np.sqrt(romc.result.samples_cov()[1, 1]),
-                  r"ROMC (gradient-based) - Histogram of the parameter $\theta_2$",
+                  r"ROMC (gradient-based) - $\theta_2$",
                   r"$\theta_2$",
                   r"density",
                   60,
@@ -132,7 +132,7 @@ else:
     plot_marginal(romc.result.samples["t2"], romc.result.weights,
                   romc.result.sample_means_array[1],
                   np.sqrt(romc.result.samples_cov()[1, 1]),
-                  r"ROMC (gradient-based) - Histogram of the parameter $\theta_2$",
+                  r"ROMC (gradient-based) - $\theta_2$",
                   r"$\theta_2$",
                   r"density",
                   60,

### 2021-08-25 latex_files/python_scripts/example_MA2_bo.py

@@ -103,7 +103,7 @@ if savefig:
     plot_marginal(romc1.result.samples["t1"], romc1.result.weights,
                   romc1.result.sample_means_array[0],
                   np.sqrt(romc1.result.samples_cov()[0, 0]),
-                  r"ROMC (Bayesian Optimisation) - Histogram of the parameter $\theta_1$",
+                  r"ROMC (Bayesian Optimisation) - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -112,7 +112,7 @@ else:
     plot_marginal(romc1.result.samples["t1"], romc1.result.weights,
                   romc1.result.sample_means_array[0],
                   np.sqrt(romc1.result.samples_cov()[0, 0]),
-                  r"ROMC (Bayesian Optimisation) - Histogram of the parameter $\theta_1$",
+                  r"ROMC (Bayesian Optimisation) - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -123,7 +123,7 @@ if savefig:
     plot_marginal(romc1.result.samples["t2"], romc1.result.weights,
                   romc1.result.sample_means_array[1],
                   np.sqrt(romc1.result.samples_cov()[1, 1]),
-                  r"ROMC (Bayesian Optimisation) - Histogram of the parameter $\theta_2$",
+                  r"ROMC (Bayesian Optimisation) - $\theta_2$",
                   r"$\theta_2$",
                   r"density",
                   60,
@@ -132,7 +132,7 @@ else:
     plot_marginal(romc1.result.samples["t2"], romc1.result.weights,
                   romc1.result.sample_means_array[1],
                   np.sqrt(romc1.result.samples_cov()[1, 1]),
-                  r"ROMC (Bayesian Optimisation) - Histogram of the parameter $\theta_2$",
+                  r"ROMC (Bayesian Optimisation) - $\theta_2$",
                   r"$\theta_2$",
                   r"density",
                   60,
@@ -150,7 +150,7 @@ if savefig:
     plot_marginal(result.samples["t1"], None,
                   result.sample_means_array[0],
                   np.std(result.samples_array[:, 0]),
-                  r"Rejection ABC - Histogram of the parameter $\theta_1$",
+                  r"Rejection ABC - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -159,7 +159,7 @@ else:
     plot_marginal(result.samples["t1"], None,
                   result.sample_means_array[0],
[…]

### 2021-08-26 latex_files/ch2/2_4_ROMC.tex
small changes

@@
   shrinks to a single point $\thetab_i^*$. This simplification adds
-  significant apprimation error when then the area $\mathcal{S}_i$ is relatively big.
+  significant approximation error when then the area $\mathcal{S}_i$ is relatively big.
 \item The weight $w_i$ is computed based only at the curvature at the
-  point $\thetab_i^*$.
-Using only local information for estimating the weight can also be misleading.
-For example, if $g_i$ is almost flat at $\thetab_i^*$, i.e.
-$\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then \Rightarrow w_i
-  \rightarrow \infty$ dominating the posterior.
+  point $\thetab_i^*$. Using only local information for estimating the
+  weight can also be misleading. For example, if $g_i$ is almost flat
+  at $\thetab_i^*$, i.e.
+  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then w_i \rightarrow
+  \infty$, dominating the posterior.
 \item There is no way to solve the optimisation problem
-  $\thetab_i^* = \text{argmin}_{\thetab} \: g_i(\thetab)$ when $g_i$
-  is not differentiable.
+  of~\eqref{eq:omc_opt_prob} when
+  $||\Phi(\data)-\Phi(M_d(\thetab, \vb_i))||_2$ is not differentiable.
 \end{itemize}
@@
 
-ROMC resolves the failure-modes of OMC mainly by approaching the
-acceptance regions $\mathcal{S}_i$ in a more sophisticated way.
-Instead of collapsing the regions into single points, it approximates
+ROMC resolves the failure-modes of OMC, mainly, by approaching the
+acceptance regions $\mathcal{S}_i$ in a more sophisticated way;
+instead of collapsing the regions into single points, it approximates
 them with a bounding box. Afterwards, a uniform distribution defined
@@
 
-The constrution of the proposal regions $q_i$ is performed in two
-steps; (a) solving the deterministic optimisation problems as in OMC
-and (b) constructing the bounding box regions.
+The constrution of the proposal regions is performed in two steps; (a)
+solving the optimisation problems as in OMC and (b) constructing the
+bounding box regions.
 
@@
 $d(M_d(\thetab^*_i,\vb_i), \data) < \epsilon$. For notation
-convenience, we name the distance function $d(M_d(\thetab,\vb_i), \data)$ as
-$d_i(\thetab)$.  Obtaining this point involves solving the following
+convenience, we name the distance $d(M_d(\thetab,\vb_i), \data)$ as
+$d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
 optimisation problem:
@@
 function $d_i$. If $d_i$ is differentiable any gradient-based
-iterative algorithm can be used for~\ref{eq:optProb}. The gradients
+optimizer can be used for~\ref{eq:optProb}. The gradients
 $\nabla_{\thetab} d_i$ can be either provided in closed form or
-approximated by finite differences. Alternatively, if $d_i$ is not
-differentiable, the Bayesian Optimisation
-approach~\cite{Shahriari2016} provides an alternative choice. In this
-case, apart from obtaining an optimal $\thetab_i^* $, a surrogate
-model $\hat{d}_i$ of the distance $d_i$ is also fitted; the
-approximate model can then substitute the actual model in the
-following steps.
+approximated by finite differences. In case $d_i$ is not
+differentiable, Bayesian Optimisation~\cite{Shahriari2016} provides an
+alternative choice. In this scenario, apart from obtaining an optimal
+$\thetab_i^* $, a surrogate model $\hat{d}_i$ of the distance $d_i$ is
+also fitted; $\hat{d}_i$ can then substitute the actual model in the
+following steps\footnote{Such as proposal region construction, sampling etc}.
 
-\subsubsection*{Proposal region computation}
+\subsubsection*{Proposal region estimation}
 
-After obtaining $\thetab_i^*$, we initially approximate the acceptance
-region $\mathcal{S}_i$ by constructing a bounding box
-$\mathcal{\hat{S}}_i$ around it. The bounding must the part of the
-acceptance region $\mathcal{S}_i$ \emph{around} $\thetab_i^*$;
-formally $\{ \thetab : d_i(\thetab) < \epsilon$,
-$d(\thetab, \thetab_i^*) < M \}$. The condition
-$d(\thetab, \thetab_i^*) < M$ describes that if
-$\mathcal{S}_i := \{ \thetab : d_i(\thetab) < \epsilon \} $ contains
-many \emph{disjoint} sets of $\thetab$ that they fulfill
+The acceptance region $\mathcal{S}_i$ is approximated with a bounding
+box $\mathcal{\hat{S}}_i$ around it. The bounding box must contain the
+part of the acceptance region \emph{around} $\thetab_i^*$; formally
+$\{ \thetab : \thetab \in \mathcal{S}_i \quad \text{and} \quad
+||\thetab - \thetab_i^*||_2 < M \}$. The condition
+$||\thetab - \thetab_i^*||_2 < M$ restricts that if $\mathcal{S}_i$
+contains many \emph{disjoint} sets of $\thetab$ that they fulfill
 $d_i(\thetab) < \epsilon$, we want the bounding box to match only the
@@
 axes $\mathbf{v}_d$ in the $d$-dimensional space and, then, the limits
-are obtained with a line-search method looking for the smallest
+along the axes are obtained with a line-search method looking for the
+smallest
 $\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
@@
   line}. Algorithm~\ref{alg:region_construction} describes the method
-in-depth. After limits are found along all directions, the uniform
-distribution $q_i$ is defined on the bounding box, and is used as the
+in-depth. After the bounding boxes construction, a uniform
+distribution $q_i$ is defined on each bounding box, and is used as the
 proposal region for importance sampling.

### 2021-08-26 latex_files/ch2/2_5_ROMC_algorithmic.tex
small changes

@@
 deterministic functions $d_i(\thetab)$, (b) defining the optimisation
-problems $\min_{\thetab} \: g_i(\thetab)$, (c) obtaining
+problems $\min_{\thetab} \: d_i(\thetab)$, (c) obtaining
 $\thetab_i^*$, $d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and
-(e) computing the proposal distribution $q_i$. If gradients are
-available, using a gradient-based method is advised for obtaining
-$\thetab_i^*$ much faster. In case $\nabla_{\thetab} g_i$ is available
-in closed-form, the procedure is upgraded in terms of both accuracy
-and efficiency; If closed-form description is not available, gradients
+(e) computing the proposal distribution $q_i$. If $d_i(\thetab)$ is
+differentiable, using a gradient-based method is advised for obtaining
+$\thetab_i^*$ faster. In case $\nabla_{\thetab} d_i$ is available in
+closed-form, the procedure is upgraded in terms of both accuracy and
+efficiency. If closed-form description is not available, gradients
 approximation with finite-differences requires two evaluations of
-$g_i$ for \emph{each} parameter $\theta_d$, which can only work in
-low-dimensional problems. If $g_i(\thetab)$ is not differentiable, the
+$d_i$ for \emph{each} parameter $\theta_d$, which can only work in
+low-dimensional problems. If $d_i(\thetab)$ is not differentiable, the
 Bayesian Optimisation can be used as an alternative solution. In this

### 2021-08-26 latex_files/ch4/4_2_ma2.tex
changed notation to be inline with ROMC paper
small changes

@@
 Second-order moving average (MA2) is a probabilistic model commonly
-used for univariate time series analysis. The observation at time $t$
-is given by:
+used for time series analysis. The observation at time $t$ is given
+by:
 
@@
 \hline
-ROMC (gradient-based) & 0.504 & 0.141 & 0.031 & 0.172 \\
+ROMC (gradient-based) & 0.504 & 0.143 & 0.033 & 0.171 \\
 \hline
-ROMC (Bayesian optimisation) & 0.5 & 0.165 & 0.07 & 0.167 \\
+ROMC (Bayesian optimisation) & 0.502 & 0.166 & 0.079 & 0.168 \\
 \hline
@@
         % \input{./latex_files/images/chapter4/ma2_region_3_bo.tex}
-        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_3.png}
-        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_3_bo.png}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.png}
     \end{center}

### 2021-08-26 latex_files/python_scripts/example_1D.py

@@ -14,7 +14,7 @@ import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 
 prepath = './../images/'
-savefig = True
+savefig = False
 
 np.random.seed(21)
 class Prior:

### 2021-08-26 latex_files/python_scripts/example_MA2.py

@@ -11,7 +11,7 @@ matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
 prepath = './../images/chapter4/'
-savefig = True
+savefig = False
 
 def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
     plt.figure()

### 2021-08-26 latex_files/python_scripts/example_MA2_bo.py

@@ -11,7 +11,7 @@ matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
 prepath = './../images/chapter4/'
-savefig = True
+savefig = False
 
 def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
     plt.figure()
@@ -103,7 +103,7 @@ if savefig:
     plot_marginal(romc1.result.samples["t1"], romc1.result.weights,
                   romc1.result.sample_means_array[0],
                   np.sqrt(romc1.result.samples_cov()[0, 0]),
-                  r"ROMC (Bayesian Optimisation) - $\theta_1$",
+                  r"ROMC (BO) - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -112,7 +112,7 @@ else:
     plot_marginal(romc1.result.samples["t1"], romc1.result.weights,
                   romc1.result.sample_means_array[0],
                   np.sqrt(romc1.result.samples_cov()[0, 0]),
-                  r"ROMC (Bayesian Optimisation) - $\theta_1$",
+                  r"ROMC (BO) - $\theta_1$",
                   r"$\theta_1$",
                   r"density",
                   60,
@@ -123,7 +123,7 @@ if savefig:
     plot_marginal(romc1.result.samples["t2"], romc1.result.weights,
                   romc1.result.sample_means_array[1],
                   np.sqrt(romc1.result.samples_cov()[1, 1]),
-                  r"ROMC (Bayesian Optimisation) - $\theta_2$",
+                  r"ROMC (BO) - $\theta_2$",
                   r"$\theta_2$",
                   r"density",
                   60,
@@ -132,7 +132,7 @@ else:
     plot_marginal(romc1.result.samples["t2"], romc1.result.weights,
                   romc1.result.sample_means_array[1],
                   np.sqrt(romc1.result.samples_cov()[1, 1]),
-                  r"ROMC (Bayesian Optimisation) - $\theta_2$",
+                  r"ROMC (BO) - $\theta_2$",
                   r"$\theta_2$",
                   r"density",
                   60,
@@ -213,12 +213,12 @@ def plot_romc_posterior(title, posterior, nof_points, savefig):
 
 
 if savefig:
-    plot_romc_posterior('MAE2, ROMC Unnorm Posterior (Bayesian optimisation)',
+    plot_romc_posterior('MAE2, ROMC Unnorm Posterior (BO)',
[…]

### 2021-08-27 latex_files/images/chapter3/example_post.tex
added Neural net

@@
-% This file was created with tikzplotlib v0.9.12.
+% This file was created by tikzplotlib v0.9.9.
 \begin{tikzpicture}
@@
 \draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.64,0) rectangle (axis cs:-2.52,0);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.52,0) rectangle (axis cs:-2.4,0.0231217584119079);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.4,0) rectangle (axis cs:-2.28,0.0357912150759672);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.28,0) rectangle (axis cs:-2.16,0.0432881444129306);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.16,0) rectangle (axis cs:-2.04,0.0678262794417402);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.04,0) rectangle (axis cs:-1.92,0.0853043382950843);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.92,0) rectangle (axis cs:-1.8,0.104057207601703);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.8,0) rectangle (axis cs:-1.68,0.133234851902131);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.68,0) rectangle (axis cs:-1.56,0.165264196422909);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.56,0) rectangle (axis cs:-1.44,0.16070619494247);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.44,0) rectangle (axis cs:-1.32,0.180066797779813);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.32,0) rectangle (axis cs:-1.2,0.221049487169422);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.2,0) rectangle (axis cs:-1.08,0.246171403878041);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.08,0) rectangle (axis cs:-0.96,0.263219759376131);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.96,0) rectangle (axis cs:-0.84,0.275654702395393);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.84,0) rectangle (axis cs:-0.72,0.30275211056918);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.72,0) rectangle (axis cs:-0.6,0.289942517702741);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.6,0) rectangle (axis cs:-0.48,0.327857582723436);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.48,0) rectangle (axis cs:-0.36,0.302668457836128);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.36,0) rectangle (axis cs:-0.24,0.314413087062547);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.24,0) rectangle (axis cs:-0.12,0.319501604166278);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.12,0) rectangle (axis cs:0,0.333969594590972);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0,0) rectangle (axis cs:0.12,0.336688308415186);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.12,0) rectangle (axis cs:0.24,0.311239645561226);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.24,0) rectangle (axis cs:0.36,0.328195053578145);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.36,0) rectangle (axis cs:0.48,0.316839373811389);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.48,0) rectangle (axis cs:0.6,0.315180976253135);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.6,0) rectangle (axis cs:0.72,0.305533027707716);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.72,0) rectangle (axis cs:0.84,0.294804743438855);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.84,0) rectangle (axis cs:0.96,0.270682012152812);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.96,0) rectangle (axis cs:1.08,0.238103562512508);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.08,0) rectangle (axis cs:1.2,0.217233993067405);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.2,0) rectangle (axis cs:1.32,0.213737380323862);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.32,0) rectangle (axis cs:1.44,0.205697780692993);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.44,0) rectangle (axis cs:1.56,0.187837922186227);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.56,0) rectangle (axis cs:1.68,0.140154434384905);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.68,0) rectangle (axis cs:1.8,0.116289453613951);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.8,0) rectangle (axis cs:1.92,0.0886000414831757);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.92,0) rectangle (axis cs:2.04,0.0841267652067261);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.04,0) rectangle (axis cs:2.16,0.0663927432898532);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.16,0) rectangle (axis cs:2.28,0.0377377498258489);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.28,0) rectangle (axis cs:2.4,0.0361079514925686);
-\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.4,0) rectangle (axis cs:2.52,0.0262891225779227);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.52,0) rectangle (axis cs:-2.4,0.0232073945389685);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.4,0) rectangle (axis cs:-2.28,0.0370673662775194);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.28,0) rectangle (axis cs:-2.16,0.0432283769706026);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.16,0) rectangle (axis cs:-2.04,0.0612233170218504);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-2.04,0) rectangle (axis cs:-1.92,0.0863738703892799);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.92,0) rectangle (axis cs:-1.8,0.109332614272474);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.8,0) rectangle (axis cs:-1.68,0.1270696943844);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.68,0) rectangle (axis cs:-1.56,0.162608319593082);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.56,0) rectangle (axis cs:-1.44,0.165932870953625);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.44,0) rectangle (axis cs:-1.32,0.186184085473939);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.32,0) rectangle (axis cs:-1.2,0.21004533914476);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.2,0) rectangle (axis cs:-1.08,0.260281980894789);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-1.08,0) rectangle (axis cs:-0.96,0.258872960512065);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.96,0) rectangle (axis cs:-0.84,0.267861221254142);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.84,0) rectangle (axis cs:-0.72,0.301309339097609);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.72,0) rectangle (axis cs:-0.6,0.296677069473362);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.6,0) rectangle (axis cs:-0.48,0.318079444437077);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.48,0) rectangle (axis cs:-0.36,0.310739645449952);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.36,0) rectangle (axis cs:-0.24,0.317600561692622);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.24,0) rectangle (axis cs:-0.12,0.313833964721813);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:-0.12,0) rectangle (axis cs:0,0.340218562084569);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0,0) rectangle (axis cs:0.12,0.343469439176735);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.12,0) rectangle (axis cs:0.24,0.315528472894501);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.24,0) rectangle (axis cs:0.36,0.327491332222326);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.36,0) rectangle (axis cs:0.48,0.320667253113843);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.48,0) rectangle (axis cs:0.6,0.309533229305267);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.6,0) rectangle (axis cs:0.72,0.307884767550316);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.72,0) rectangle (axis cs:0.84,0.289208340516575);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.84,0) rectangle (axis cs:0.96,0.272217212371973);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:0.96,0) rectangle (axis cs:1.08,0.235895798061777);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.08,0) rectangle (axis cs:1.2,0.221437222892658);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.2,0) rectangle (axis cs:1.32,0.207761436825051);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.32,0) rectangle (axis cs:1.44,0.206149812204291);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.44,0) rectangle (axis cs:1.56,0.191175517156526);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.56,0) rectangle (axis cs:1.68,0.130919174907133);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.68,0) rectangle (axis cs:1.8,0.118127480060059);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.8,0) rectangle (axis cs:1.92,0.0924796253810799);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:1.92,0) rectangle (axis cs:2.04,0.0812258808863897);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.04,0) rectangle (axis cs:2.16,0.0618587575866078);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.16,0) rectangle (axis cs:2.28,0.0384119216754122);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.28,0) rectangle (axis cs:2.4,0.0377120161258239);
+\draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.4,0) rectangle (axis cs:2.52,0.026430643780492);
 \draw[draw=none,fill=color0,fill opacity=0.5] (axis cs:2.52,0) rectangle (axis cs:2.64,0);

### 2021-09-23 latex_files/ch2/2_1_implicit_models.tex
multiple edits until Sec 2.4

@@
 An implicit or simulator-based model is a parameterised stochastic
-data generating mechanism. The key characteristic
-of these models is that we can sample (simulate) data points,
-but that we cannot evaluate the likelihood of a specific set of observations
-$\data$. Formally, a simulator-based model is described as a
-parameterised family of probability density functions
-$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
+data generating mechanism. The key characteristic of these models is
+that we can sample (simulate) data points, but that we cannot evaluate
+the likelihood. Formally, a simulator-based model is described as a
+parameterised family of probability density functions $\{
+p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
 unknown or intractable to evaluate. Whereas evaluating
@@
 feasible. Practically, a simulator can be understood as a black-box
-data generating machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
-  \textit{random} simulator. In the next chapters we will introduce
-  $M_d$ witch stands for the \textit{deterministic} simulator.} that
-produces samples $\yb$ in a stochastic manner for any given a set of parameters $\thetab$.% i.e.\ $M_r(\thetab) \rightarrow \yb$.
+data generating machine $\simulator$\footnote{The subscript $r$ in $M_r$
+indicates the \textit{random} simulator. In the next chapters we will
+introduce $M_d$ witch stands for the \textit{deterministic}
+simulator.} that produces samples $\yb$ in a stochastic manner for any
+given a set of parameters $\thetab$.% i.e.\
+$M_r(\thetab) \rightarrow \yb$.
 

### 2021-09-23 latex_files/ch2/2_2_LFI_background.tex
multiple edits until Sec 2.4

@@
 that we can sample, or simulate, data points, but that we cannot
-evaluate the likelihood of a specific set of observations
-$\data$. Formally, a simulator-based model is a parameterised family
-of probability density functions $\{ p(\yb|\thetab) \}_{\thetab}$,
-whose closed-form is either unknown or computationally
-intractable. Whereas evaluating $p(\yb|\thetab)$ is intractable,
-sampling is feasible. Practically, a simulator can be understood as a
-random black-box machine (computer code) $M_r$ that generates samples
-$\yb$ in a stochastic manner for any given a set of parameters
-$\thetab$. We can isolate the randomness by introducing the stochastic
-nuisance variables $\vb \sim p(\vb)$ so that the simulator becomes a
-deterministic mapping $M_d$ that maps $(\thetab, \vb)$ to the data
-$\yb$. The distribution $p(\vb)$ is determined by the random draws
-performed when running $M_r$. Importantly, we can fix the value of
-$\vb$ while changing $\thetab$ by fixing the random seed used with
-$M_r$.
+evaluate the likelihood. Formally, a simulator-based model is a
+parameterised family of probability density functions $\{
+p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either unknown or
+computationally intractable. Whereas evaluating $p(\yb|\thetab)$ is
+intractable, sampling is feasible. Practically, a simulator can be
+understood as a random black-box machine (computer code) that
+generates samples $\yb$ in a stochastic manner for any given a set of
+parameters $\thetab$. We can isolate the randomness by introducing the
+stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator
+becomes a deterministic mapping $\simulator$ that maps
+$(\thetab, \vb)$ to the data $\yb$, i.e.\
+$\yb=\simulator(\thetab,\vb)$. The distribution $p(\vb)$ is determined
+by the random draws performed when running the computer
+code. Importantly, we can fix the value of $\vb$ while changing
+$\thetab$ by fixing the random seed used.
 
@@
 The modelling freedom of simulator-based models, however, comes at the
-price of difficulties in inferring their parameters. The main
-difficulty is that the likelihood function $L(\thetab) =
-p(\data|\thetab)$ is intractable. To better see the sources of the
-intractability, and to address them, we go back to the basic
-characterisation of the likelihood as the (rescaled) probability for a
-parameter of the model to generate data $\yb$ that is similar to the
-observed data $\data$. More formally, the likelihood $L(\thetab)$
+price of difficulties in inferring their parameters. Denoting observed
+data by $\data$, the main difficulty is that the likelihood function
+$L(\thetab) = p(\data|\thetab)$ is intractable. To better see the
+sources of the intractability, and to address them, we go back to the
+basic characterisation of the likelihood as the (rescaled) probability
+for a parameter of the model to generate data $\yb$ that is similar to
+the observed data $\data$. More formally, the likelihood $L(\thetab)$
 equals
@@
   L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
-  \lim_{\epsilon \to 0} c_\epsilon \Pr(M_r(\thetab) \in \region(\data))
+  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data))
 \end{equation}
@@
 highlights two sources of intractability: The first is the computation
-of the probability $\Pr(M_r(\thetab) \in \region(\data))$, the second
+of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$, the second
 is the limit of $\epsilon \to 0$.  Whilst the probability can be

### 2021-09-23 latex_files/ch2/2_3_OMC.tex
small changes
multiple edits until Sec 2.4

@@
   L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)) =
-  \int_{\yb \in B_\epsilon(\data)}p(\yb|\thetab)d\yb =
+  \int_{\yb \in B_{d,\epsilon}(\data)}p(\yb|\thetab)d\yb =
   \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb
@@
 L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
-  \text{ where } \yb_i = M_d(\thetab, \vb_i), \quad \vb_i \sim p(\vb). \label{eq:alt_view}
+ \quad \text{ where } \yb_i = \simulator(\thetab, \vb_i),\; \vb_i \sim p(\vb). \label{eq:alt_view}
 \end{equation}
@@
 different values of $\thetab$. Secondly, since
-$\yb_i=M_d(\thetab, \vb_i)$ is a function of $\thetab$, checking
-whether $\yb_i$ is contained in region $\region (\data)$ for a
-particular value of $\thetab$ is the same as checking whether
-$\thetab$ is in the acceptance region $\mathcal{S}_i = \{ \thetab:
-M_d(\thetab, \vb_i) \in \region(\data) \}$. This leads to the
-likelihood approximation
+$\yb_i=\simulator(\thetab, \vb_i)$ is a function of $\thetab$,
+checking whether $\yb_i$ is contained in region $\region (\data)$ for
+a particular value of $\thetab$ is the same as checking whether
+$\thetab$ is in the acceptance region $\accregioni
+= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}$. This
+leads to the likelihood approximation
 \begin{equation}
-L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\mathcal{S}_i}(\thetab) \label{eq:alt_view_theta}
+L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\accregioni}(\thetab) \label{eq:alt_view_theta}
 \end{equation}
@@
   p_{d,\epsilon}(\thetab|\data) \propto
-  p(\thetab) \sum_i^N  \indicator{\mathcal{S}_i}(\thetab).
+  p(\thetab) \sum_i^N  \indicator{\accregioni}(\thetab).
 \end{equation}
-As argued by \citet{Ikonomov2019}, this derivations provide a unique
+As argued by \citet{Ikonomov2019}, these derivations provide a unique
 perspective for likelihood-free inference by shifting the focus onto
-the geometry of the acceptance regions $\mathcal{S}_i = \{ \thetab:
-M_d(\thetab, \vb_i) \in \region(\data)\}$. Indeed, the task of
-approximating the likelihood and/or posterior becomes a task of
-characterising the sets $\mathcal{S}_i$.
+the geometry of the acceptance regions $\accregioni$. Indeed, the task
+of approximating the likelihood and/or posterior becomes a task of
+characterising the sets $\accregioni$.
 
@@
 Euclidean distance between summary statistics $\Phi$ of the observed
-and generated data, and that the $\mathcal{S}_i$ can be well
-approximated by infinitesimally small ellipses. These assumptions
-leads to an approximation of the posterior in terms of weighted
-samples $\thetab_i^*$. In more detail, OMC considers the limit of
-$\epsilon \to 0$ and represents the sets $\mathcal{S}_i$ by points
-$\thetab_i^*$ which are those parameter configurations that achieve
-the smallest distance between observed and simulated data for each
-realisation $\vb_i \sim p(\vb)$, i.e.\ they are solutions to the
-deterministic optimisation problems
-\begin{equation}
-\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(M_d(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
+and generated data, and that the $\accregioni$ can be well
+approximated by infinitesimally small ellipses. These assumptions leads to an approximation of the posterior in
+terms of weighted samples $\thetab_i^*$ that achieve the smallest
+distance between observed and simulated data for each realisation
+$\vb_i \sim p(\vb)$, i.e.\
+\begin{equation} \label{eq:omc_opt_prob}
+\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
 \end{equation}

### 2021-09-23 latex_files/ch2/2_4_ROMC.tex
multiple edits until Sec 2.4

@@
-The simplifications introduced by OMC, although beneficial from a
-computational point-of-view, they suffer from some drawbacks:
+\citet{Ikonomov2019} showed that considering infinitesimally small ellipses can lead
+to highly overconfident posteriors. We refer the reader to their paper
+for the technical details and conditions for this issue to
+occur. Intuitively, it happens because the weights in OMC are only
+computed from information at $\thetab_i^*$. And using only local
+information can be misleading. For example if the curvature of
+$||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at $\thetab_i^*$
+is nearly flat, the curvature alone may wrongly indicate that
+$\accregioni$ is much larger than it actually is. We implement in our
+software package the robust generalisation of OMC
+by \citet{Ikonomov2019} that resolves this issue.
 
-\begin{itemize}
-\item The whole acceptable region $\mathcal{S}_i$
-  shrinks to a single point $\thetab_i^*$. This simplification adds
-  significant approximation error when then the area $\mathcal{S}_i$ is relatively big.
-\item The weight $w_i$ is computed based only at the curvature at the
-  point $\thetab_i^*$. Using only local information for estimating the
-  weight can also be misleading. For example, if $g_i$ is almost flat
-  at $\thetab_i^*$, i.e.
-  $\text{det}(\jac_i^{*T}\jac_i^*) \rightarrow 0, then w_i \rightarrow
-  \infty$, dominating the posterior.
-\item There is no way to solve the optimisation problem
-  of~\eqref{eq:omc_opt_prob} when
-  $||\Phi(\data)-\Phi(M_d(\thetab, \vb_i))||_2$ is not differentiable.
-\end{itemize}
-
-\subsubsection{Sampling and computing expectation in ROMC}
-
-ROMC resolves the failure-modes of OMC, mainly, by approaching the
-acceptance regions $\mathcal{S}_i$ in a more sophisticated way;
-instead of collapsing the regions into single points, it approximates
-them with a bounding box. Afterwards, a uniform distribution defined
-on the bounding box area, is used as the \emph{proposal distribution}
-for importance sampling. Defining as $q_i$ the uniform distribution
-defined on the $i-th$ bounding box, weighted sampling is performed by
-drawing samples from $q_i$ and assigning the weights as
-in~\eqref{eq:sampling}.
-
-\begin{gather}
-  \thetab_{ij} \sim q_i \\  \label{eq:sampling}
-  w_{ij} = \frac{\indicator{\region(\data)}(M_d(\thetab_{ij}, \vb_i)) p(\thetab_{ij})}{q(\thetab_{ij})}
-\end{gather}
-
-\noindent
-
-Having defined the procedure for obtaining weighted samples, any
-expectation $E_{p(\thetab|\data)}[h(\thetab)]$, can be approximated
-as,
-
+ROMC approximates the acceptance regions $\accregioni$ and defines
+local proposal distributions $q_i(\thetab)$ on them from which the
+posterior samples $\thetab_{ij} \sim q_i$ are generated. The samples
+are assigned (importance) weights $w_{ij}$ that compensate for using
+the proposal distributions $q_i(\thetab)$ and not the prior
+$p(\thetab)$,
+\begin{equation}
+  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
+  \label{eq:sampling}
+\end{equation}
+Given the weighted samples, any expectation
+$\Ex_{p(\thetab|\data)}[h(\thetab)]$ of some function $h(\thetab)$, can be approximated as
 \begin{equation} \label{eq:expectation}
-  E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
+  \Ex_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
 \end{equation}
-
-The constrution of the proposal regions is performed in two steps; (a)
-solving the optimisation problems as in OMC and (b) constructing the
-bounding box regions.
-
+\citet{Ikonomov2019} considered uniform distributions as proposal
+distributions so that the main work is to approximate the acceptance
+regions $\accregioni$ and to represent them such that uniform sampling is
+easy. The approximation of the acceptance regions contains two compulsory
+and one optional step: (1) solving the optimisation problems as in OMC, (2)
+constructing bounding boxes around $\accregioni$ and optionally, (3)
+refining the approximation via a probabilistic surrogate model of the distance.
 \subsubsection*{Solving the deterministic optimisation problems}
-
-For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$,
-we search for a point $\thetab^*_i$ such that
-$d(M_d(\thetab^*_i,\vb_i), \data) < \epsilon$. For notation
-convenience, we name the distance $d(M_d(\thetab,\vb_i), \data)$ as
+For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$, we
+search for a point $\thetab^*_i$ such that
+$d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon$. For notational
+convenience, we denote the distance $d(\simulator(\thetab,\vb_i), \data)$ by
 $d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
 optimisation problem:
-
 \begin{subequations}
 \begin{alignat}{2}      
-  &\!\min_{\thetab}        &\qquad& d_i(\thetab) = d(M_d(\thetab^*,\vb_i), \data)\label{eq:optProb}\\
-  &\text{subject to} & & d_i(\thetab) \leq \epsilon
+  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
+  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
 \end{alignat}
@@
 the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
-only if $d_i(\thetab_i^*) < \epsilon$. The nature of the generative
-model $M_r(\thetab)$, specifies the properties of the objective
-function $d_i$. If $d_i$ is differentiable any gradient-based
-optimizer can be used for~\ref{eq:optProb}. The gradients
-$\nabla_{\thetab} d_i$ can be either provided in closed form or
-approximated by finite differences. In case $d_i$ is not
-differentiable, Bayesian Optimisation~\cite{Shahriari2016} provides an
-alternative choice. In this scenario, apart from obtaining an optimal
-$\thetab_i^* $, a surrogate model $\hat{d}_i$ of the distance $d_i$ is
-also fitted; $\hat{d}_i$ can then substitute the actual model in the
-following steps\footnote{Such as proposal region construction, sampling etc}.
+only if $d_i(\thetab_i^*) \le \epsilon$. If $d_i(\thetab)$ is
+differentiable any gradient-based optimizer can be used
+for~\ref{eq:optProb}. The gradients $\nabla_{\thetab} d_i(\thetab)$
+can be either provided in closed form or approximated by finite
+differences. In case $d_i$ is not differentiable, Bayesian
+Optimisation~\cite{Shahriari2016} provides an alternative choice. In
+this scenario, apart from obtaining an optimal $\thetab_i^* $, a
+surrogate model $\hat{d}_i(\thetab)$ of the distance function
+$d_i(\thetab)$ is also automatically obtained; $\hat{d}_i$ can then
+substitute the actual distance function in downstream steps of the
+algorithms, which can lead to computational gains if running the
+simulator and hence evaluating the distance function $d_i(\thetab)$ is
+computationally expensive.
 
-\subsubsection*{Proposal region estimation}
+\subsubsection*{Estimating the acceptance regions}
+The acceptance region $\accregioni$ is approximated by a bounding box
+$\accregionihat$ around it. Ideally, we want the bounding box to be as
+tight as possible to $\mathcal{S}_i$, to ensure high acceptance rate
+in the importance sampling, and big enough for not discarding valid
+parts of $\mathcal{S}_i$. \TODO{did remove the issue of disconnected
+sets; which needed to be handled by multiple restarts with the same
+seed. Is this something that the code would support? It's expensive
+because we don't know when we need it! I think it's an open research
+question, not related to this software paper, so I omitted it.} The
+bounding boxes are built in two steps. First, we define their axes
+$\mathbf{v}_d$ based on the (estimated) curvature of the distance
+$\thetab_i^*$, second, we determine the size of the box via a
+one-dimensional line-search method along each axix, see
+Algorithm~\ref{alg:region_construction} for the details. After the
+bounding boxes construction, a uniform distribution $q_i$ is defined
+on each bounding box, and is used as the proposal region for
+importance sampling.
 
-The acceptance region $\mathcal{S}_i$ is approximated with a bounding
-box $\mathcal{\hat{S}}_i$ around it. The bounding box must contain the
-part of the acceptance region \emph{around} $\thetab_i^*$; formally
-$\{ \thetab : \thetab \in \mathcal{S}_i \quad \text{and} \quad
-||\thetab - \thetab_i^*||_2 < M \}$. The condition
-$||\thetab - \thetab_i^*||_2 < M$ restricts that if $\mathcal{S}_i$
-contains many \emph{disjoint} sets of $\thetab$ that they fulfill
-$d_i(\thetab) < \epsilon$, we want the bounding box to match only the
-set that contains $\thetab_i^*$. Ideally, we want the bounding box to
-be as tight as possible to $\mathcal{S}_i$, to ensure high acceptance
-rate in the importance sampling, and big enough for not discarding
-valid parts of $\mathcal{S}_i$.
+\subsubsection*{Refining the estimate via a local surrogate model (optional)}
+When computing the weight $w_{ij}$ in \eqref{eq:sampling}, we need to
+check whether the samples $\thetab_{ij} \sim q_i$ lie inside the
+acceptance region $\accregioni$. This can be considered to be a
+safety-mechanism that corrects for any inaccuracies in the
+construction of $\accregionihat$ above. However, this check involves
+evaluting the distance function $d_i(\thetab_{ij})$ and hence running
+the simulator $\simulator(\thetab_{ij}, \vb_i)$, which can be
+expensive if the model is complex.
 
-The bounding boxes are built in two steps; firstly, we define their
-axes $\mathbf{v}_d$ in the $d$-dimensional space and, then, the limits
-along the axes are obtained with a line-search method looking for the
-smallest
-$\kappa: d_i(\thetab_i^* + \kappa \vb_d) \geq
-\epsilon$\footnote{$-\kappa$ is used as well for the opposite
-  direction along the search
-  line}. Algorithm~\ref{alg:region_construction} describes the method
-in-depth. After the bounding boxes construction, a uniform
-distribution $q_i$ is defined on each bounding box, and is used as the
-proposal region for importance sampling.
+\citet{Ikonomov2019} thus proposed to construct a surrogate model $\hat{d}_i(\thetab)$
+of the distance function $d_i(\thetab)$ valid in $\accregionihat$ and
+that is is cheap to evaluate. They used a simple quadratic model
+whilst other regression models are, in principle, possible too. The
+advantage of using a quadratic model is that it has ellipsoidal
+isocontours, which thus naturally allowed \citet{Ikonomov2019} to
+replace the bounding box approximation of $\accregioni$ with a
+tigher-fitting ellipsoidal approximation.\footnote{The difference to
+the infinitesimal ellipsoidal model in OMC is the estimation procedure: OMC uses
+information at $\thetab_i^*$ whilst, here, information in
+$\accregionihat$ is used, which results in a more stable fit.}
 
-\subsubsection*{Fitting a local surrogate model $\hat{g}_i$}
-
-After the construction of the bounding box $\mathcal{\hat{S}}_i$, we
-are no longer interested in $d_i(\thetab)$ outside
-$\mathcal{\hat{S}}_i$. In the inference steps
-(i.e. sampling~\eqref{eq:sampling}, computing
-expectation~\eqref{eq:expectation}, evaluating the
-posterior~\eqref{eq:approx_posterior_omc}), we need to evaluate the
-distance function $d_i$ for checking if the samples obtained from the
-uniform distribution $\thetab_{ij} \sim q_i$ lie inside the acceptance
-region $\mathcal{S}_i$. Evaluating $d_i$ involves running the
-deterministic simulator $M_d(\thetab_{ij}, \vb_i)$, which can be
-inneficient when the model is big.\footnote{If Bayesian Optimization
-  has been used for obtaining $\thetab_i^*$, then $\hat{d}_i$ can be
-  used as an alternative.}. Hence, a \emph{light} surrogate model
-$\Tilde{d}_i$ can be fit for approximating $d_i$ around $\theta_i^*$.
-
-ROMC proposes fitting a simple quadratic model\footnote{Any ML
-  regression model may be chosen as local surrogate. The choice should
-  consider the properties of the local region (i.e. size,
-  smoothness). } by sampling training points from the bounding box and
-labeling them by evaluating $d_i$. This additional step adds an extra
-effort in the training part, but promises a significant speed-up at
-the inference phase. Such time-badget imbalance between the training
-and the inference part is demanded in many ML contexts; it is usual to
-be quite generous with the execution time at the training phase but
-quite eager at the inference phase. Fitting a local surrogate model
-aligns with this requirement.
+The training data for fitting the quadratic model is obtained by
+sampling $\thetab \sim q_i$ and evaluating the corresponding distances
+$d_i(\thetab)$. The generation of the training data adds an extra
+computational cost, but leads to a significant speed-up when
+evaluating the weights $w_{ij}$. Moreover, the extra cost is largely
+eliminated if Bayesian Optimization with a Gaussian processe (GP)
+surrogate model was used to obtain $\thetab_i^*$ in the first place
+(see above). In this case, we can use the GP model of the distance
+function instead of $d_i(\thetab)$ to generate the training data. This
+essentially replaces the global GP model with a simpler local
+quadratic model which is typically more robust.

### 2021-09-23 latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex
changed notation to be inline with ROMC paper
small changes

@@
 tick pos=left,
-title={ROMC (Bayesian Optimisation) - Histogram of the parameter \(\displaystyle \theta_1\)},
+title={ROMC (BO) - \(\displaystyle \theta_1\)},
 x grid style={white!69.0196078431373!black},
@@
 ]
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.59230812208571);
-\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,2.04310565145056);
-\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,2.19382815490065);
-\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.20021699441589);
-\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.51506116421806);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.44880980882892);
-\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.65904829491329);
-\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.73250343027527);
-\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.82852250487442);
-\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,2.82331312648089);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.90135211837164);
-\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.88136478592358);
-\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.53560317010369);
-\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,2.3564598017404);
-\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.22420707945385);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.33534256536061);
-\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.00495652659749);
-\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,1.74992091039568);
-\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.83870977452066);
-\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.6578788920657);
-\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.55026416918985);
-\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.62152961093496);
-\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.6437730803435);
-\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.66640493128548);
-\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.32365272171391);
-\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.14155317519524);
-\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.17480604967746);
-\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.20896637137188);
-\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,1.26617555858479);
-\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,1.10815016078187);
-\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,1.01575133068398);
-\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.778653581806476);
-\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.658517776017212);
-\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.522857721601219);
-\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.649384569447495);
-\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.571069158563703);
-\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.362376122626364);
-\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.275087556542002);
-\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.255847252846728);
-\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.406941618766892);
-\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.192257669195406);
-\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.204927790769102);
-\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.203525138747285);
-\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.113584712363094);
-\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0662010640185658);
-\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.116253091726246);
-\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0.0456418048890116);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.61050759359378);
+\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.93197020366126);
+\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,2.17257862381295);
+\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.3025919446582);
+\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.42910062383181);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.77208977160826);
+\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.91890254170224);
+\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.79720744902878);
+\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.89385988809184);
+\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.03707063266652);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.96625904171146);
+\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,2.87224832339805);
+\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.65419765058084);
+\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,2.62169851287656);
+\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.32485253886239);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.14150464429012);
+\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.24693369229431);
+\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.33863073514552);
+\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.14663274237078);
+\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.90924828657755);
+\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.71866221248833);
+\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.67767714633082);
+\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.5285521718297);
+\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.07857259816532);
+\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.06467730589615);
+\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.31604561147561);
+\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.16959736883555);
+\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.934947898685309);
+\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.697393683133471);
+\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.826583507333401);
+\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.526592417514156);
+\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.744305041869243);
+\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.507002670950614);
+\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.550468818640814);
+\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.531452195144458);
+\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.449213153956996);
+\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.46212842931813);
+\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.307647048984769);
+\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.39272346125025);
+\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.33527111008569);
+\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.236505059054111);
+\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.27054174354952);
+\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0792063139920926);
+\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0290615761452173);
+\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0742518004133052);
+\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.037125900206652);
+\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0.0323749806537868);
 \draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
@@
 table {%
-0.502312841390367 0
-0.502312841390367 3.5
+0.494409356161276 0
+0.494409356161276 3.5
 };
-\addlegendentry{$\mu = 0.502$}
+\addlegendentry{$\mu = 0.494$}
 \addplot [semithick, black, dashed]
 table {%
-0.294820390999603 1
-0.660267860059204 1
+0.292577624981475 1
+0.645122958573331 1
 };
-\addlegendentry{$\sigma = 0.166$}
+\addlegendentry{$\sigma = 0.160$}
 \end{axis}

### 2021-09-23 latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (Neural Network) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,0.918657455133441);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.20251419656731);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.33742778487049);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.56194622164953);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.10132928493464);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.35262182295654);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.99077324408847);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,3.06195916743878);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.20534143070652);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.54249833812081);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.48746191660371);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.64921604449934);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.65237906359403);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.41841807637564);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.70295622336452);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.57664640133676);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.13473725538676);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.24111391157059);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.17436219440432);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,2.02392362475434);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,2.1307420959631);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.34480742455423);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.48585437983289);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.43000242757426);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.44793856735925);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.04402163798497);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.00795945546565);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.973472415485145);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,1.03878684968805);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,1.07033952922709);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.499611406245489);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.427526350563684);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.301712601546415);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.33618635385698);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.291403030644556);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.251052806310352);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.313679809271219);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.230501210162083);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.16371073729186);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.133740397779522);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.128624644858993);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.097498278063311);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.119954440233943);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0362421197162086);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0137859575455791);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0112280810853149);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.502656907372089 0
0.502656907372089 3.5
};
\addlegendentry{$\mu = 0.503$}
\addplot [semithick, black, dashed]
table {%
0.32355294654317 1
0.632292249675427 1
};
\addlegendentry{$\sigma = 0.140$}
\end{axis}

\end{tikzpicture}

### 2021-09-23 latex_files/images/chapter4/mae2_hist_t2_romc.tex
changed notation to be inline with ROMC paper

--- 2021-07-23 latex_files/images/chapter4/mae2_hist_t2_romc.tex
@@
 \draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
-\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0132189786765283);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0698717444330784);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.203477850342276);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.355023998741048);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.514595955621997);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.712880635769923);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.916830592493502);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.952710677472651);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.1953733574632);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.3804390589346);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.52584782437642);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.59477535604688);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.7604846958848);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.91108663152097);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.09048705641672);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.27366433236291);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.32654024706902);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.32811393500671);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.26296325438667);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.10968604925645);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15437878668663);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.78361790856873);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.58785112912109);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.44590447714217);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.22086710205366);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.01125186875443);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.805413486505622);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.800220316311271);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.62758674954756);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.600834054606967);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.396254622708316);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.300574396097251);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.197497836179084);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.168384609331968);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.156581949799354);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.068455425289165);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0550790778188679);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.066094893382642);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0275395389094342);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0275395389094339);
+\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00551481774368941);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0605814218347914);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.168699423699822);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.340262186728983);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.48172967019454);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.661723210849365);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.926167683238105);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.1131441320768);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.19038558853121);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.39560413532974);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.38890799766854);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.45486223069012);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.89014334843844);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.80399178478282);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.15561058973456);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.33061276076321);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.24024442959031);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.34890403251458);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.31095726097247);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.33900171768764);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.07598627244674);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.71987564844275);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65165022964706);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.40256222683377);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.18758980571923);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.0980208933351);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.803715750920935);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.742179116696122);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.678702875114039);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.583516431505742);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.398341719580722);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.257727424376927);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.229546935490747);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.210112717761985);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.129179090828181);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0899034779939157);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0431810229330879);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0479789143700981);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0287873486220588);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0143936743110293);
 \draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
@@
 table {%
-0.0308218751896741 0
-0.0308218751896741 3
+0.0318388478807404 0
+0.0318388478807404 3
 };
-\addlegendentry{$\mu = 0.031$}
+\addlegendentry{$\mu = 0.032$}
 \addplot [semithick, black, dashed]
 table {%
--0.179871006878183 1
-0.197679132295466 1
+-0.177435597378694 1
+0.197481062716323 1
 };
-\addlegendentry{$\sigma = 0.172$}
+\addlegendentry{$\sigma = 0.170$}
 \end{axis}

### 2021-09-23 latex_files/ch3/3_1_ELFI.tex
create tikz figures and plots
fixed warnings
light edits in ELFI section

@@
   \href{https://elfi.readthedocs.io/en/latest/}{here}}~\cite{1708.00707}
-is a \proglang{Python} package dedicated to Likelihood-Free Inference
-(LFI). We implemented ROMC at \pkg{ELFI} since it provides convenient
-modules for all the fundamental components of a probabilistic model
-(e.g. prior, simulator, summaries etc.). Furthermore, \pkg{ELFI}
-already supports some recently proposed likelihood-free inference
-methods, making it straightforward to perform comparisons between
-methods. The implementation was officially added to \pkg{ELFI} at
-version \pkg{v0.8.0}.
+is a \proglang{Python} package for Likelihood-Free Inference
+(LFI). We selected to implemented ROMC in \pkg{ELFI} since it provides
+convenient modules for all the fundamental components of a
+probabilistic model (e.g.\ prior, simulator, summaries
+etc.). Furthermore, \pkg{ELFI} already supports some recently proposed
+likelihood-free inference methods, making it straightforward to
+perform comparisons between methods. The implementation was officially
+added to \pkg{ELFI} in version \pkg{v0.8.0}.
 
@@
 to the output node. Samples can be obtained from all nodes through
-sequential sampling. \pkg{ELFI} automatically considers as parameters
-of interest, i.e. those we try to infer a posterior distribution,
-as the ones  defined through the \code{elfi.Prior} class. Finally, the
-observations are passed to the appropriate node through the argument
+sequential (ancestral) sampling. \pkg{ELFI} automatically considers as
+parameters of interest, i.e. those we try to infer a posterior
+distribution, as the ones defined through the \code{elfi.Prior}
+class. Finally, the observations are passed to the appropriate node
+through the argument
 \code{observed}.
 
-\begin{figure}[h]
+\begin{figure}[ht]
     \begin{center}
@@
 
-All inference methods at \pkg{ELFI} share the following rules;
+All inference methods in \pkg{ELFI} share the following rules;
 

### 2021-10-05 latex_files/ch2/2-3-ROMC.tex
reordered files and fixed graph colors

--- 2021-09-23 latex_files/ch2/2_4_ROMC.tex
@@
-\citet{Ikonomov2019} showed that considering infinitesimally small ellipses can lead
-to highly overconfident posteriors. We refer the reader to their paper
-for the technical details and conditions for this issue to
-occur. Intuitively, it happens because the weights in OMC are only
-computed from information at $\thetab_i^*$. And using only local
-information can be misleading. For example if the curvature of
-$||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at $\thetab_i^*$
-is nearly flat, the curvature alone may wrongly indicate that
-$\accregioni$ is much larger than it actually is. We implement in our
-software package the robust generalisation of OMC
-by \citet{Ikonomov2019} that resolves this issue.
+\citet{Ikonomov2019} showed that considering infinitesimally small
+ellipses can lead to highly overconfident posteriors. We refer the
+reader to their paper for the technical details and conditions for
+this issue to occur. Intuitively, it happens because the weights in
+OMC are only computed from information at $\thetab_i^*$. And using
+only local information can be misleading. For example if the curvature
+of $||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at
+$\thetab_i^*$ is nearly flat, the curvature alone may wrongly indicate
+that $\accregioni$ is much larger than it actually is. We implement in
+our software package the robust generalisation of OMC by
+\citet{Ikonomov2019} that resolves this issue.
 
@@
 distributions so that the main work is to approximate the acceptance
-regions $\accregioni$ and to represent them such that uniform sampling is
-easy. The approximation of the acceptance regions contains two compulsory
-and one optional step: (1) solving the optimisation problems as in OMC, (2)
-constructing bounding boxes around $\accregioni$ and optionally, (3)
-refining the approximation via a probabilistic surrogate model of the distance.
+regions $\accregioni$ and to represent them such that uniform sampling
+is easy. The approximation of the acceptance regions contains two
+compulsory and one optional step: (1) solving the optimisation
+problems as in OMC, (2) constructing bounding boxes around
+$\accregioni$ and optionally, (3) refining the approximation via a
+probabilistic surrogate model of the distance.
 \subsubsection*{Solving the deterministic optimisation problems}
@@
 in the importance sampling, and big enough for not discarding valid
-parts of $\mathcal{S}_i$. \TODO{did remove the issue of disconnected
-sets; which needed to be handled by multiple restarts with the same
-seed. Is this something that the code would support? It's expensive
-because we don't know when we need it! I think it's an open research
-question, not related to this software paper, so I omitted it.} The
-bounding boxes are built in two steps. First, we define their axes
-$\mathbf{v}_d$ based on the (estimated) curvature of the distance
-$\thetab_i^*$, second, we determine the size of the box via a
-one-dimensional line-search method along each axix, see
-Algorithm~\ref{alg:region_construction} for the details. After the
-bounding boxes construction, a uniform distribution $q_i$ is defined
-on each bounding box, and is used as the proposal region for
-importance sampling.
+parts of $\mathcal{S}_i$. The bounding boxes are built in two
+steps. First, we define their axes $\mathbf{v}_d$ based on the
+(estimated) curvature of the distance $\thetab_i^*$, second, we
+determine the size of the box via a one-dimensional line-search method
+along each axix, see Algorithm~\ref{alg:region_construction} for the
+details. After the bounding boxes construction, a uniform distribution
+$q_i$ is defined on each bounding box, and is used as the proposal
+region for importance sampling.
 

### 2021-10-05 latex_files/ch3/3-0-design-principles.tex
reordered files and fixed graph colors

--- 2021-08-26 latex_files/ch2/2_5_ROMC_algorithmic.tex
@@
 serves as the driver for the implementation at \pkg{ELFI}. At a high
-level, ROMC can be split into the training and the inference part. In
+level, ROMC can be split into the trainind the inference part. In
 general terms, the training part covers all steps for estimating the
@@
 $\thetab_i^*$, $d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and
-(e) computing the proposal distribution $q_i$. If $d_i(\thetab)$ is
-differentiable, using a gradient-based method is advised for obtaining
-$\thetab_i^*$ faster. In case $\nabla_{\thetab} d_i$ is available in
-closed-form, the procedure is upgraded in terms of both accuracy and
-efficiency. If closed-form description is not available, gradients
-approximation with finite-differences requires two evaluations of
-$d_i$ for \emph{each} parameter $\theta_d$, which can only work in
-low-dimensional problems. If $d_i(\thetab)$ is not differentiable, the
-Bayesian Optimisation can be used as an alternative solution. In this
-scenario, the training part becomes slower due to fitting of the
-surrogate model and the blind optimisation steps. After obtaning the
-optimal points $\thetab^*$, the training part gets completed by
-estimating the proposal regions;
-Algorithm~\ref{alg:region_construction} describes the line search
-approach for finding the region's boundaries.
+(e) computing the proposal distribution $q_i$.
+
+If $d_i(\thetab)$ is differentiable, using a gradient-based method is
+advised for obtaining $\thetab_i^*$ faster. In this case, the
+gradients $\nabla_{\thetab} d_i$ gradients are approximated
+automatically with finite-differences. This procedure requires about
+two evaluations of $d_i$ for \emph{each} parameter $\theta_d$, which
+can only work in low-dimensional problems. If $d_i(\thetab)$ is not
+differentiable, the Bayesian Optimisation can be used as an
+alternative solution. In this scenario, the training part becomes
+slower due to fitting of the surrogate model and the blind
+optimisation steps.
+
+After obtaining the optimal points $\thetab^*$, the training part gets
+completed by estimating the proposal
+regions. Algorithm~\ref{alg:region_construction} describes the line
+search approach for finding the region's boundaries. An important step
+for the proposal region estimation is deciding the axes of the
+bounding box i.e. the directions $\vb_d$ we follow for reaching the
+boundaries. We approximate them as the direction of the highest
+curvature of $d_i$ at $\thetab_i^*$. This estimation is given by the
+eigenvalues of the Hessian matrix $\hessian_i$ of
+$d_i$\footnote{Either the real distance $d_i$ or the Gaussian Process
+  approximation $\hat{d}_i$} at $\thetab_i$. The Hessian matrix is
+approximated numerically. In case where the distance function is the
+Euclidean, the Hessian matrix can be also computed as
+$\hessian_i = \jac_i^T\jac_i$, where $\jac_i$ is the Jacobian matrix
+of the summary statistics $\Phi(M_d(\thetab, \vb=\vb_i))$ at
+$\thetab_i^*$. The approximation through the Jacobian matrix has the
+computational advantage of using only first-order derivatives.
 

### 2021-10-05 latex_files/ch4/4-2-inference.tex
reordered files and fixed graph colors

--- 2021-07-25 latex_files/ch3/3_2_3_inference.tex
@@
-The inference part contains in four methods.
+The inference part contains four methods.
 
@@
 remained after filtering\footnote{From the $n_1$ optimisation
-  problems, only the ones with $g_i(\thetab_*) < \epsilon$ are
+  problems, only the ones with $d_i(\thetab_*) < \epsilon$ are
   maintained for building a bounding box}. The samples are drawn from
@@
 box and the weight $w_i$ is computed as in
-equation~\eqref{eq:sampling}.
+algorithm~\ref{alg:sampling_GB}.
 

### 2021-10-05 latex_files/ch4/4-3-evaluation.tex
reordered files and fixed graph colors

--- 2021-07-19 latex_files/ch3/3_2_4_evaluation.tex
@@
   print("Nof Samples: %d, ESS: %.3f" % (nof_samples, ess))
-  # Nof Samples: 19950, ESS: 16694.816
+  # Nof Samples: 19300, ESS: 16138.825
 ----------------------------------------------------------------------------  

### 2021-10-05 latex_files/ch4/4-4-extensibility.tex
reordered files and fixed graph colors

ROMC can be conceived as a meta-algorithm; it describes a sequence of
steps for approximating the posterior distribution, without
\emph{explicitly} enforcing the methods that solve these steps. Even
though for each step a specific algotihm is proposed
by~\cite{Ikonomov2019}, in general ROMC works well if a practitioner
thinks of an alternative way of approaching a specific
step. Considering this particularity, we designed the implementation
to support such \emph{extensibility} capability.

We have specified four specific points where a user may intervene with
their custom modules; (a) the gradient-based optimization, (b) the
Bayesian Optimization, (c) the proposal region construction and (d)
the surrogate model fitting. These are the four critical parts of the
ROMC procedure, whereas the rest of the code is the backbone of the
algorithm; each of them may be approached with a completely different
algorithm than the one we propose, without the rest of the method to
collapse.

The four replaceable parts described above, are solved using the four
methods of the \linebreak \code{OptimisatioProblem} class; (a)
\code{solve_gradients(**kwargs)}, (b) \code{solve_bo(**kwargs)},
\linebreak (c) \code{build_region(**kwargs)}, (d)
\code{fit_local_surrogate(**kwargs)}. Therefore, the user should
create a custom class that inherits the basic
\code{OptimisatioProblem} class and overwrites one or more of the
above four functions with custom ones.

We illustrate that in the
following example; suppose we want to blindly set the proposal region
as a predefined n-dimensional bounding box around the optimal
point. The following snippet illustrates how:

\begin{Code}
---------------------------------- python ----------------------------------
class customOptim(OptimisationProblem):
    def __init__(self, **kwargs):
        super(customOptim, self).__init__(**kwargs)

    def fit_local_surrogate(self, **kwargs):
        nof_samples = 500
        objective = self.objective

        local_surrogates = []
        for i in range(len(self.regions)):
            # prepare dataset
            x = self.regions[i].sample(nof_samples)
            y = np.array([objective(ii) for ii in x])

            # train Neural Network
            mlp = MLPRegressor(hidden_layer_sizes=(10,10), solver='adam')
            model = Pipeline([('linear', mlp)])
            model = model.fit(x, y)
            local_surrogates.append(self.create_local_surrogate(model))

        self.local_surrogates = local_surrogates
        self.state["local_surrogates"] = True

    def create_local_surrogate(self, model):
        def _local_surrogate(th):
            th = np.expand_dims(th, 0)
            return float(model.predict(th))
        return _local_surrogate
----------------------------------------------------------------------------    
\end{Code}

In the same way, the user may replace each of the other three
functionalities. The only restriction that must be respected, concerns
the side-effects each method has at the \code{OptimisatioProblem}
class level. In the following four snippets we present the signature
of each method, pointing out with a left arrow the side-effects that
must be included in the custom routine that will replace them.

\begin{Code}
---------------------------------- python ----------------------------------
def solve_gradients(self, **kwargs):
  self.state["attempted"] = True

  # custom solution procedure

  if success:
    self.result = RomcOpimisationResult(res.x, res.y, jac, hess_inv)
    self.state["solved"] = True
    return True 
  else:
    return False 
----------------------------------------------------------------------------    
def solve_bo(self, **kwargs):
  self.state["attempted"] = True

  # custom procedure

  if success:
    self.result = RomcOpimisationResult(x, y)
    self.surrogate <- Callable
    self.state["solved"] = True
    self.state["has_fit_surrogate"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------    
def build_region(self, **kwargs):
    self.eps_region = eps_region

    if success:
        # construct region
        self.regions <- List[NDimBoudningBox]
        self.state["region"] = True
        return True
    else:
        return False
----------------------------------------------------------------------------    
def fit_local_surrogate(self, **kwargs):
  if success:
    self.local_surrogate = local_surrogates
    self.state["local_surrogates"] = True
    return True
  else:
    return False
----------------------------------------------------------------------------    
\end{Code}

The two classes that may be needed for creating the custom routines
are \\ (a) \code{RomcOpimisationResult} and (b)
\code{NDimBoundingBox}. We present their signatures below.

\begin{Code}
---------------------------------- python ----------------------------------
class RomcOpimisationResult:
    def __init__(self, x_min, f_min, hess_appr):
        Parameters
        ----------
        x_min: np.ndarray (D,) or float, the minimum point
        f_min: float, distance at the minimum point
        hess_appr: np.ndarray (DxD), Hessian approximation at x_min
        """
----------------------------------------------------------------------------    
class NDimBoundingBox:
    def __init__(self, rotation, center, limits, eps_region):
        Parameters
        ----------
        rotation: np.array (D,D),  rotation matrix for the bounding box
        center: np.array (D,) center of the bounding box
        limits: np.ndarray, shape: (D,2), the limits of the bounding box
        eps_region: float, distance threshold 
----------------------------------------------------------------------------    
\end{Code}

% Let's say we have observed that the local area around $\theta_i^*$ is
% too complex to be represented by a simple quadratic model\footnote{as
%   in the current implementation}. Hence, the user selects a neural
% network as a good alternative. In the following snippet, we
% demonstrate how they could implement this enhancement without much
% effort; (a) they have to develop the neural network using the package
% of their choice (b) they must create a custom optimisation class which
% inherits the basic \code{OptimisationClass} and (c) they have to
% overwrite the \code{fit_local_surrogate} routine, with one that
% sets the neural network's prediction function as the
% \code{local_surrogate} attribute. The argument \code{**kwargs}
% may be used for passing all the important arguments, e.g.\ training
% epochs, gradient step etc. If, for example, they would like to set the
% size of the training set dynamically, we may replace \code{x =
%   self.regions[0].sample(30)} with \code{x =
%   self.regions[0].sample(kwargs["nof_examples"])}. Finally, they must
% pass the custom optimisation class, when calling the \code{ROMC}
% method.

% \begin{Code}
% ------------------------------ python snippet ------------------------------  
%   class NeuralNetwork:
%       def __init__(self, **kwargs):
%           # set the input arguments

%       def train(x, y):
%           # training code

%       def predict(x):
%           # prediction code

%   # Inherit the base optimisation class
%   class customOptim(elfi.methods.parameter_inference.OptimisationProblem):
%       def __init__(self, **kwargs):
%           super(customOptim, self).__init__(**kwargs)

%       # overwrite the function you want to replace
%       def fit_local_surrogate(**kwargs):
%           # init and train the NN
%           x = self.regions[0].sample(30) # 30 training points
%           y = [np.array([self.objective(ii) for ii in x])]
%           nn = NeuralNet()
%           nn.train(x,y)

%           # set the appropriate attribute
%           self.local_surrogate = nn.predict

%           # update the state
%           self.state["local_surrogate"] = True

%   # pass the custom inference method as argument
%   romc = elfi.ROMC(dist, bounds, custom_optim_class=customOptim)
% ----------------------------------------------------------------------------
% \end{Code}

### 2021-10-05 latex_files/ch5/5-1-ma2.tex
reordered files and fixed graph colors

--- 2021-08-26 latex_files/ch4/4_2_ma2.tex
@@
 optimisation alternatives (gradient-based and Bayesian optimisation)
-and we compare the results at each step. In the end, we use the
-Rejection ABC algorithm for evaluating both approximations. Rejection
-ABC is a robust method in terms of accuracy and can be used for
-baseline comparison.
+and, in the end, by extending the implementation with a Neural Network
+surrogate model. In the end, we use the Rejection ABC algorithm for
+evaluating our approaches; Rejection ABC is a robust method in terms
+of accuracy and can be used for baseline comparison.
+
+The implementation's extension is performed using a multi-layer
+perceptron as local surrogate function, as illustrated in chapter
+... This intervention can be done easily by coding a custom
+optimisation function with the surrogate model of our own
+preference. The surrogate model substitutes the real distance function
+$d_i$ inside the proposal region $q_i \quad \forall i$ at the
+inference phase i.e. sampling, computing an expectation and evaluating
+the posterior. In our example we use a neural network of two hidden
+layers of 10 neurons each and we train it in 500 examples for each
+proposal region.
 
 In figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
-same deterministic simulator, i.e. with \code{seed=799981517}, in the
-gradient-based and the Bayesian optimisation case. The acceptance
-regions are similar in both cases. The optimal points are close and
-the search directions are similar. This indicates that both
-optimization scemes work robustly. Therefore, we observe that as we
-move away from the optimal pint, the surrogate model fitted by the
-Bayesian Optimisation tends to deviate from the real distance. This
-divergence affects the bounding box approximation and explains some
-small differences between the posterior approximations, as shown in
-figure \ref{fig:ma2_3}. In figure \ref{fig:ma2_3}, we demonstrate the
-histograms of the marginal posteriors, using the three different
-methods; (a) Rejection ABC (first line), (b) ROMC with gradient-based
-optimisation (second line) and (c) ROMC with Bayesian optimisation
-(third line). Undoubtedly, we observe that there is a significant
-similarity between the three approaches. The Rejection ABC inference
-has been set to infer 10000 accepted samples, with threshold
-$\epsilon=0.1$. At table \ref{tab:ma2} we present the empirical mean
-$\mu$ and standard deviation $\sigma$ for each inference approach, for
-verifying their agreement.
+same deterministic simulator, in the gradient-based and the Bayesian
+optimisation case. The acceptance regions are similar even though the
+optimal points differ significantly due to the different optimisation
+schemens.
+
+In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
+marginal posteriors, using all four approaches; (a) Rejection ABC
+(first column), (b) ROMC with gradient-based optimisation (second
+column) (c) ROMC with Bayesian optimisation (third column) and (d)
+ROMC with the Neural Network extendion. We observe that there is a
+significant aggreement between the different approaches. The Rejection
+ABC inference has been set to infer 10000 accepted samples, with
+threshold $\epsilon=0.1$. At table \ref{tab:ma2} we present the
+empirical mean $\mu$ and standard deviation $\sigma$ for each
+inference approach. In figure \ref{fig:ma2_4}, we illustrate the
+unnormalized posterior for the three different variations of the ROMC
+method.
+
+\subsubsection*{Extending the implementation}
 
@@
 \hline
-ROMC (gradient-based) & 0.504 & 0.143 & 0.033 & 0.171 \\
+ROMC (gradient-based) & 0.503 & 0.143 & 0.032 & 0.17 \\
 \hline
-ROMC (Bayesian optimisation) & 0.502 & 0.166 & 0.079 & 0.168 \\
+ROMC (Bayesian optimisation) & 0.494 & 0.16 & 0.086 & 0.167 \\
+\hline
+ROMC (Neural Network) & 0.503 & 0.140 & 0.03 & 0.171 \\
 \hline
@@
 
-% \begin{figure}[h]
-%     \begin{center}
-%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.png}
-%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.png}\\
-%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.png}
-%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.png}\\
-%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.png}
-%       \includegraphics[width=0.35\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.png}\\
-%     \end{center}
-%     \caption[MA2 example, evaluation of the marginal
-%     distributions.]{Histogram of the marginal posterior distributions
-%       using three different inference approaches; (a) in the first
-%       row, the samples are obtained using Rejection ABC sampling (b)
-%       in the second row, using ROMC with a gradient-based optimiser
-%       and (c) in the third row, using ROMC with Bayesian optimization
-%       approach. The vertical (red) line represents the samples mean
-%       $\mu$ and the horizontal (black) the standard deviation
-%       $\sigma$.}
-%   \label{fig:ma2_3}
-% \end{figure}
-
 \begin{figure}[ht]
   \begin{center}
+    \resizebox{.24\columnwidth}{!}{%
       \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
-    \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}\\
+    }
+    \resizebox{.24\columnwidth}{!}{%
       \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
-    \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}\\
+    }
+    \resizebox{.24\columnwidth}{!}{%
       \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
-    \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}\\
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex}
+    }\\
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex}
+    }
     \end{center}
@@
 
-% \begin{figure}[H]
-%     \begin{center}
-%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.png}
-%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.png}
-%     \end{center}
-%     \caption[MA2 example, approximate posterior.]{ROMC approximate posteriors using gradient-based approach
-%       (left) and Bayesian optimisation approach (right).}
-%   \label{fig:ma2_4}
-% \end{figure}
+\begin{figure}[ht]
+  \begin{center}
+    \resizebox{.32\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_romc_posterior.tex}
+    }
+    \resizebox{.32\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_romc_posterior_bo.tex}
+    }
+    \resizebox{.32\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_romc_posterior_nn.tex}
+    }
+    \end{center}
+    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
+  \label{fig:ma2_4}
+\end{figure}

### 2021-10-05 latex_files/ch5/5-2-execution-time.tex
reordered files and fixed graph colors

--- 2021-07-23 latex_files/ch4/4_3_exec.tex
@@
 all the available CPU cores. Our implementation supports
-parallelization using the built-in \proglang{Python}
-\pkg{multiprocess}. In figure \label{fig:exec_parallel} we observe the
+parallelization using the built-in \proglang{Python} package
+\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
 execution times for performing the inference. The parallel version
@@
 
-\begin{figure}[ht] \label{fig:exec_parallel}
+\begin{figure}[ht]
   \centering
@@
     \end{minipage}\hfill
-    \caption[Execution time exploiting parallelisation]{In the first
+    \caption[Execution time exploiting parallelization]{In the first
       line, we compare the parallel and sequential execution of the
@@
       time for evaluating the posterior at a batch of $50$ points.}
+      \label{fig:exec_parallel}
 \end{figure}

### 2021-10-05 latex_files/graphs/romc_graph.tex
reordered files and fixed graph colors

@@
       \begin{tikzpicture}[node distance=1.2cm, scale=.1]
+
+        % backgrounds
+        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115, 8) rectangle (70, -108);
+        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115, -110) rectangle (70, -138);
+        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115, -142) rectangle (70, -158);
+
         % first graph
@@
         % add headers
-        \node (algorithm) [below of=n1, yshift=2.5cm, xshift=1cm, minimum width=4cm, minimum height=1cm] {\huge{Implementation Design} };
-
-        % backgrounds
-        \draw [ultra thick, draw=black, fill=yellow, opacity=0.15, rounded corners=20pt] (-115,8) rectangle (70, -108);
-
-        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115,8) rectangle (57, -48);
-        \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-115,-50) rectangle (57, -108);
-
-        \draw [ultra thick, draw=black, fill=green, opacity=0.15, rounded corners=20pt] (-115,-110) rectangle (70, -140);
-
-        \draw [ultra thick, draw=black, fill=cyan, opacity=0.15, rounded corners=20pt] (-115,-142) rectangle (70, -160);
-
-        \draw [ultra thick, draw=black, fill=magenta, opacity=0.05, rounded corners=20pt] (-115,18) rectangle (-64, -160);
-
-        \draw [ultra thick, draw=black, fill=magenta, opacity=0.05, rounded corners=20pt] (-60,18) rectangle (70, -160);
+        \node (impl_design) [below of=n1, yshift=2.5cm, xshift=1cm, minimum width=4cm, minimum height=1cm] {\huge{Implementation Design} };
 
@@
 
-        % % backgrounds
-        % \draw [ultra thick, draw=black, fill=yellow, opacity=0.15, rounded corners=5pt] (-72,8) rectangle (-62,-108);
-
-        % % \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-60,8) rectangle (57, -48);
-        % % \draw [ultra thick, draw=black, fill=red, opacity=0.15, rounded corners=20pt] (-60,-50) rectangle (57, -108);
-
-        % % \draw [ultra thick, draw=black, fill=green, opacity=0.15, rounded corners=20pt] (-60,-110) rectangle (70, -140);
-
-        % % \draw [ultra thick, draw=black, fill=cyan, opacity=0.15, rounded corners=20pt] (-60,-142) rectangle (70, -160);
-
-        % \draw [arrow] (pro1) -- (pro2);
-        % \draw [arrow] (pro2) -- (filter);
-        % \draw [arrow] (filter) -- (proposal_region);
-        % \draw [arrow] (proposal_region) -- (surrogate);
-
       \end{tikzpicture}

### 2021-10-24 latex_files/ch4/4-3-evaluation.tex
updated some images

@@
   print("Jensen-Shannon divergence: %.3f" % res)
-  # Jensen-Shannon divergence: 0.025
+  # Jensen-Shannon divergence: 0.035
 
@@
   print("Nof Samples: %d, ESS: %.3f" % (nof_samples, ess))
-  # Nof Samples: 19300, ESS: 16138.825
+  # Nof Samples: 19300, ESS: 16196.214
 ----------------------------------------------------------------------------  

### 2021-10-24 latex_files/images/chapter3/example_marginal.tex
updated some images

@@
 \draw[draw=none,fill=color0] (axis cs:-2.65714285714286,0) rectangle (axis cs:-2.57142857142857,0);
-\draw[draw=none,fill=color0] (axis cs:-2.57142857142857,0) rectangle (axis cs:-2.48571428571429,0.006208033765389);
-\draw[draw=none,fill=color0] (axis cs:-2.48571428571429,0) rectangle (axis cs:-2.4,0.0261624280112821);
-\draw[draw=none,fill=color0] (axis cs:-2.4,0) rectangle (axis cs:-2.31428571428571,0.0337007547263973);
-\draw[draw=none,fill=color0] (axis cs:-2.31428571428571,0) rectangle (axis cs:-2.22857142857143,0.0363613406258499);
-\draw[draw=none,fill=color0] (axis cs:-2.22857142857143,0) rectangle (axis cs:-2.14285714285714,0.0541896684870789);
-\draw[draw=none,fill=color0] (axis cs:-2.14285714285714,0) rectangle (axis cs:-2.05714285714286,0.0670526704061584);
-\draw[draw=none,fill=color0] (axis cs:-2.05714285714286,0) rectangle (axis cs:-1.97142857142857,0.0841007399119992);
-\draw[draw=none,fill=color0] (axis cs:-1.97142857142857,0) rectangle (axis cs:-1.88571428571429,0.0882637860955287);
-\draw[draw=none,fill=color0] (axis cs:-1.88571428571429,0) rectangle (axis cs:-1.8,0.107105098505383);
-\draw[draw=none,fill=color0] (axis cs:-1.8,0) rectangle (axis cs:-1.71428571428571,0.134226458530087);
-\draw[draw=none,fill=color0] (axis cs:-1.71428571428571,0) rectangle (axis cs:-1.62857142857143,0.13742807026801);
-\draw[draw=none,fill=color0] (axis cs:-1.62857142857143,0) rectangle (axis cs:-1.54285714285714,0.172328991470496);
-\draw[draw=none,fill=color0] (axis cs:-1.54285714285714,0) rectangle (axis cs:-1.45714285714286,0.159523545491678);
-\draw[draw=none,fill=color0] (axis cs:-1.45714285714286,0) rectangle (axis cs:-1.37142857142857,0.181085502173434);
-\draw[draw=none,fill=color0] (axis cs:-1.37142857142857,0) rectangle (axis cs:-1.28571428571429,0.198590015309879);
-\draw[draw=none,fill=color0] (axis cs:-1.28571428571429,0) rectangle (axis cs:-1.2,0.221267556259859);
-\draw[draw=none,fill=color0] (axis cs:-1.2,0) rectangle (axis cs:-1.11428571428571,0.234145572772051);
-\draw[draw=none,fill=color0] (axis cs:-1.11428571428571,0) rectangle (axis cs:-1.02857142857143,0.277788991057236);
-\draw[draw=none,fill=color0] (axis cs:-1.02857142857143,0) rectangle (axis cs:-0.942857142857143,0.250781241453748);
-\draw[draw=none,fill=color0] (axis cs:-0.942857142857143,0) rectangle (axis cs:-0.857142857142857,0.265410460009239);
-\draw[draw=none,fill=color0] (axis cs:-0.857142857142857,0) rectangle (axis cs:-0.771428571428571,0.303461442846629);
-\draw[draw=none,fill=color0] (axis cs:-0.771428571428571,0) rectangle (axis cs:-0.685714285714285,0.302778779346469);
-\draw[draw=none,fill=color0] (axis cs:-0.685714285714285,0) rectangle (axis cs:-0.6,0.294470204004708);
-\draw[draw=none,fill=color0] (axis cs:-0.6,0) rectangle (axis cs:-0.514285714285714,0.32589124301026);
-\draw[draw=none,fill=color0] (axis cs:-0.514285714285714,0) rectangle (axis cs:-0.428571428571428,0.307263137822596);
-\draw[draw=none,fill=color0] (axis cs:-0.428571428571428,0) rectangle (axis cs:-0.342857142857143,0.310432718428548);
-\draw[draw=none,fill=color0] (axis cs:-0.342857142857143,0) rectangle (axis cs:-0.257142857142857,0.313571769361836);
-\draw[draw=none,fill=color0] (axis cs:-0.257142857142857,0) rectangle (axis cs:-0.171428571428571,0.322468416267153);
-\draw[draw=none,fill=color0] (axis cs:-0.171428571428571,0) rectangle (axis cs:-0.0857142857142859,0.329442694767694);
-\draw[draw=none,fill=color0] (axis cs:-0.0857142857142859,0) rectangle (axis cs:0,0.328704477273017);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.0857142857142859,0.315996626149429);
-\draw[draw=none,fill=color0] (axis cs:0.0857142857142859,0) rectangle (axis cs:0.171428571428571,0.35869562652787);
-\draw[draw=none,fill=color0] (axis cs:0.171428571428571,0) rectangle (axis cs:0.257142857142857,0.299321419039204);
-\draw[draw=none,fill=color0] (axis cs:0.257142857142857,0) rectangle (axis cs:0.342857142857143,0.327344655623505);
-\draw[draw=none,fill=color0] (axis cs:0.342857142857143,0) rectangle (axis cs:0.428571428571429,0.324602990921247);
-\draw[draw=none,fill=color0] (axis cs:0.428571428571429,0) rectangle (axis cs:0.514285714285714,0.300242814569848);
-\draw[draw=none,fill=color0] (axis cs:0.514285714285714,0) rectangle (axis cs:0.6,0.325196567835609);
-\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.685714285714286,0.3008173729996);
-\draw[draw=none,fill=color0] (axis cs:0.685714285714286,0) rectangle (axis cs:0.771428571428571,0.306931815801259);
-\draw[draw=none,fill=color0] (axis cs:0.771428571428571,0) rectangle (axis cs:0.857142857142857,0.292671455749884);
-\draw[draw=none,fill=color0] (axis cs:0.857142857142857,0) rectangle (axis cs:0.942857142857143,0.266064095296049);
-\draw[draw=none,fill=color0] (axis cs:0.942857142857143,0) rectangle (axis cs:1.02857142857143,0.241445667943122);
-\draw[draw=none,fill=color0] (axis cs:1.02857142857143,0) rectangle (axis cs:1.11428571428571,0.23182281521762);
-\draw[draw=none,fill=color0] (axis cs:1.11428571428571,0) rectangle (axis cs:1.2,0.217147051423478);
-\draw[draw=none,fill=color0] (axis cs:1.2,0) rectangle (axis cs:1.28571428571429,0.224310513797232);
-\draw[draw=none,fill=color0] (axis cs:1.28571428571429,0) rectangle (axis cs:1.37142857142857,0.212377916164659);
-\draw[draw=none,fill=color0] (axis cs:1.37142857142857,0) rectangle (axis cs:1.45714285714286,0.190746892354591);
-\draw[draw=none,fill=color0] (axis cs:1.45714285714286,0) rectangle (axis cs:1.54285714285714,0.196218210084616);
-\draw[draw=none,fill=color0] (axis cs:1.54285714285714,0) rectangle (axis cs:1.62857142857143,0.133681428800119);
-\draw[draw=none,fill=color0] (axis cs:1.62857142857143,0) rectangle (axis cs:1.71428571428571,0.142913401618273);
-\draw[draw=none,fill=color0] (axis cs:1.71428571428571,0) rectangle (axis cs:1.8,0.108955396863223);
-\draw[draw=none,fill=color0] (axis cs:1.8,0) rectangle (axis cs:1.88571428571429,0.0924363415213623);
-\draw[draw=none,fill=color0] (axis cs:1.88571428571429,0) rectangle (axis cs:1.97142857142857,0.0739859090188589);
-\draw[draw=none,fill=color0] (axis cs:1.97142857142857,0) rectangle (axis cs:2.05714285714286,0.0907797314146689);
-\draw[draw=none,fill=color0] (axis cs:2.05714285714286,0) rectangle (axis cs:2.14285714285714,0.0652904576612423);
-\draw[draw=none,fill=color0] (axis cs:2.14285714285714,0) rectangle (axis cs:2.22857142857143,0.0433796619328517);
-\draw[draw=none,fill=color0] (axis cs:2.22857142857143,0) rectangle (axis cs:2.31428571428571,0.0363613406258499);
-\draw[draw=none,fill=color0] (axis cs:2.31428571428571,0) rectangle (axis cs:2.4,0.0359179096426075);
-\draw[draw=none,fill=color0] (axis cs:2.4,0) rectangle (axis cs:2.48571428571429,0.0301533068604609);
-\draw[draw=none,fill=color0] (axis cs:2.48571428571429,0) rectangle (axis cs:2.57142857142857,0.00665146474863107);
+\draw[draw=none,fill=color0] (axis cs:-2.57142857142857,0) rectangle (axis cs:-2.48571428571429,0.00620818024343236);
+\draw[draw=none,fill=color0] (axis cs:-2.48571428571429,0) rectangle (axis cs:-2.4,0.0261630453116077);
+\draw[draw=none,fill=color0] (axis cs:-2.4,0) rectangle (axis cs:-2.31428571428571,0.0337015498929184);
+\draw[draw=none,fill=color0] (axis cs:-2.31428571428571,0) rectangle (axis cs:-2.22857142857143,0.0363621985686753);
+\draw[draw=none,fill=color0] (axis cs:-2.22857142857143,0) rectangle (axis cs:-2.14285714285714,0.054190947087826);
+\draw[draw=none,fill=color0] (axis cs:-2.14285714285714,0) rectangle (axis cs:-2.05714285714286,0.0670542525083721);
+\draw[draw=none,fill=color0] (axis cs:-2.05714285714286,0) rectangle (axis cs:-1.97142857142857,0.0841027242620032);
+\draw[draw=none,fill=color0] (axis cs:-1.97142857142857,0) rectangle (axis cs:-1.88571428571429,0.0882658686722632);
+\draw[draw=none,fill=color0] (axis cs:-1.88571428571429,0) rectangle (axis cs:-1.8,0.107107625641326);
+\draw[draw=none,fill=color0] (axis cs:-1.8,0) rectangle (axis cs:-1.71428571428571,0.134229125094031);
+\draw[draw=none,fill=color0] (axis cs:-1.71428571428571,0) rectangle (axis cs:-1.62857142857143,0.137431312871964);
+\draw[draw=none,fill=color0] (axis cs:-1.62857142857143,0) rectangle (axis cs:-1.54285714285714,0.172331556064062);
+\draw[draw=none,fill=color0] (axis cs:-1.54285714285714,0) rectangle (axis cs:-1.45714285714286,0.159526808938298);
+\draw[draw=none,fill=color0] (axis cs:-1.45714285714286,0) rectangle (axis cs:-1.37142857142857,0.181088273376146);
+\draw[draw=none,fill=color0] (axis cs:-1.37142857142857,0) rectangle (axis cs:-1.28571428571429,0.198593700028377);
+\draw[draw=none,fill=color0] (axis cs:-1.28571428571429,0) rectangle (axis cs:-1.2,0.221271776053061);
+\draw[draw=none,fill=color0] (axis cs:-1.2,0) rectangle (axis cs:-1.11428571428571,0.234818762077649);
+\draw[draw=none,fill=color0] (axis cs:-1.11428571428571,0) rectangle (axis cs:-1.02857142857143,0.276972225850834);
+\draw[draw=none,fill=color0] (axis cs:-1.02857142857143,0) rectangle (axis cs:-0.942857142857143,0.250286660370399);
+\draw[draw=none,fill=color0] (axis cs:-0.942857142857143,0) rectangle (axis cs:-0.857142857142857,0.266079382024805);
+\draw[draw=none,fill=color0] (axis cs:-0.857142857142857,0) rectangle (axis cs:-0.771428571428571,0.303023660053582);
+\draw[draw=none,fill=color0] (axis cs:-0.771428571428571,0) rectangle (axis cs:-0.685714285714285,0.303811444292793);
+\draw[draw=none,fill=color0] (axis cs:-0.685714285714285,0) rectangle (axis cs:-0.6,0.294340015485417);
+\draw[draw=none,fill=color0] (axis cs:-0.6,0) rectangle (axis cs:-0.514285714285714,0.325167704449779);
+\draw[draw=none,fill=color0] (axis cs:-0.514285714285714,0) rectangle (axis cs:-0.428571428571428,0.30751913529919);
+\draw[draw=none,fill=color0] (axis cs:-0.428571428571428,0) rectangle (axis cs:-0.342857142857143,0.309628234907852);
+\draw[draw=none,fill=color0] (axis cs:-0.342857142857143,0) rectangle (axis cs:-0.257142857142857,0.314430015080012);
+\draw[draw=none,fill=color0] (axis cs:-0.257142857142857,0) rectangle (axis cs:-0.171428571428571,0.322025576461367);
+\draw[draw=none,fill=color0] (axis cs:-0.171428571428571,0) rectangle (axis cs:-0.0857142857142859,0.330218732748452);
+\draw[draw=none,fill=color0] (axis cs:-0.0857142857142859,0) rectangle (axis cs:0,0.327214742275935);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.0857142857142859,0.316669244232633);
+\draw[draw=none,fill=color0] (axis cs:0.0857142857142859,0) rectangle (axis cs:0.171428571428571,0.358477864714324);
+\draw[draw=none,fill=color0] (axis cs:0.171428571428571,0) rectangle (axis cs:0.257142857142857,0.300973118739705);
+\draw[draw=none,fill=color0] (axis cs:0.257142857142857,0) rectangle (axis cs:0.342857142857143,0.326183215390922);
+\draw[draw=none,fill=color0] (axis cs:0.342857142857143,0) rectangle (axis cs:0.428571428571429,0.324785323789832);
+\draw[draw=none,fill=color0] (axis cs:0.428571428571429,0) rectangle (axis cs:0.514285714285714,0.300400548746278);
+\draw[draw=none,fill=color0] (axis cs:0.514285714285714,0) rectangle (axis cs:0.6,0.324864903010947);
+\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.685714285714286,0.301298442599607);
+\draw[draw=none,fill=color0] (axis cs:0.685714285714286,0) rectangle (axis cs:0.771428571428571,0.30662874891955);
+\draw[draw=none,fill=color0] (axis cs:0.771428571428571,0) rectangle (axis cs:0.857142857142857,0.292636319456018);
+\draw[draw=none,fill=color0] (axis cs:0.857142857142857,0) rectangle (axis cs:0.942857142857143,0.267089387485241);
+\draw[draw=none,fill=color0] (axis cs:0.942857142857143,0) rectangle (axis cs:1.02857142857143,0.240312230825788);
+\draw[draw=none,fill=color0] (axis cs:1.02857142857143,0) rectangle (axis cs:1.11428571428571,0.231829286054268);
+\draw[draw=none,fill=color0] (axis cs:1.11428571428571,0) rectangle (axis cs:1.2,0.217155177979677);
+\draw[draw=none,fill=color0] (axis cs:1.2,0) rectangle (axis cs:1.28571428571429,0.224316807381749);
+\draw[draw=none,fill=color0] (axis cs:1.28571428571429,0) rectangle (axis cs:1.37142857142857,0.212381425709292);
+\draw[draw=none,fill=color0] (axis cs:1.37142857142857,0) rectangle (axis cs:1.45714285714286,0.190750892513498);
+\draw[draw=none,fill=color0] (axis cs:1.45714285714286,0) rectangle (axis cs:1.54285714285714,0.196222839837058);
+\draw[draw=none,fill=color0] (axis cs:1.54285714285714,0) rectangle (axis cs:1.62857142857143,0.133683582005886);
+\draw[draw=none,fill=color0] (axis cs:1.62857142857143,0) rectangle (axis cs:1.71428571428571,0.142916773648167);
+\draw[draw=none,fill=color0] (axis cs:1.71428571428571,0) rectangle (axis cs:1.8,0.108957967656803);
+\draw[draw=none,fill=color0] (axis cs:1.8,0) rectangle (axis cs:1.88571428571429,0.0924385225491971);
+\draw[draw=none,fill=color0] (axis cs:1.88571428571429,0) rectangle (axis cs:1.97142857142857,0.0739876547102644);
+\draw[draw=none,fill=color0] (axis cs:1.97142857142857,0) rectangle (axis cs:2.05714285714286,0.0907818733549252);
+\draw[draw=none,fill=color0] (axis cs:2.05714285714286,0) rectangle (axis cs:2.14285714285714,0.065291998184192);
+\draw[draw=none,fill=color0] (axis cs:2.14285714285714,0) rectangle (axis cs:2.22857142857143,0.043380685472388);
+\draw[draw=none,fill=color0] (axis cs:2.22857142857143,0) rectangle (axis cs:2.31428571428571,0.0363621985686753);
+\draw[draw=none,fill=color0] (axis cs:2.31428571428571,0) rectangle (axis cs:2.4,0.0359187571227155);
+\draw[draw=none,fill=color0] (axis cs:2.4,0) rectangle (axis cs:2.48571428571429,0.0301540183252429);
+\draw[draw=none,fill=color0] (axis cs:2.48571428571429,0) rectangle (axis cs:2.57142857142857,0.00665162168939181);
 \draw[draw=none,fill=color0] (axis cs:2.57142857142857,0) rectangle (axis cs:2.65714285714286,0);
@@
 -3 0
--2.59322023391724 0
--2.49152541160583 0.0369433164596558
--2.38983058929443 0.0452905893325806
--2.28813552856445 0.0549525022506714
--2.18644070625305 0.0659897327423096
--2.08474564552307 0.0784285068511963
--1.98305082321167 0.0922528505325317
--1.88135588169098 0.107397437095642
--1.77966105937958 0.123741984367371
--1.67796611785889 0.141106963157654
--1.5762711763382 0.159253358840942
--1.27118647098541 0.21515953540802
--1.16949152946472 0.232988834381104
--1.06779658794403 0.249699831008911
--0.96610164642334 0.264856100082397
--0.864406824111938 0.278041839599609
--0.762711882591248 0.288881063461304
--0.661016941070557 0.297054767608643
--0.559322118759155 0.302316904067993
--0.457627058029175 0.304275751113892
--0.355932235717773 0.304529428482056
-0.254237294197083 0.304565906524658
-0.355932235717773 0.304529428482056
-0.457627058029175 0.304275751113892
-0.559322118759155 0.302316904067993
-0.661016941070557 0.297054767608643
-0.762711882591248 0.288881063461304
-0.864406824111938 0.278041839599609
-0.96610164642334 0.264856100082397
-1.06779658794403 0.249699831008911
-1.16949152946472 0.232988834381104
-1.27118647098541 0.21515953540802
-1.3728814125061 0.196650266647339
-1.5762711763382 0.159253358840942
-1.67796611785889 0.141106963157654
-1.77966105937958 0.123741984367371
-1.88135588169098 0.107397437095642
-1.98305082321167 0.0922528505325317
-2.08474564552307 0.0784285068511963
-2.18644070625305 0.0659897327423096
-2.28813552856445 0.0549525022506714
-2.38983058929443 0.0452905893325806
-2.49152541160583 0.0369433164596558
-2.59322023391724 0
+-2.89830508474576 0
+-2.79661016949153 0
+-2.69491525423729 0
+-2.59322033898305 0
+-2.49152542372881 0.036943365219351
+-2.38983050847458 0.0452905809004646
+-2.28813559322034 0.0549525585640254
+-2.1864406779661 0.0659897576831711
+-2.08474576271186 0.0784284692399008
+-1.98305084745763 0.0922527949339697
+-1.88135593220339 0.107397436643172
+-1.77966101694915 0.123741930739631
+-1.67796610169492 0.141106964487117
+-1.57627118644068 0.159253358442953
+-1.47457627118644 0.177884185408375
+-1.3728813559322 0.19665032476711
+-1.27118644067797 0.215159529401958
+-1.16949152542373 0.232988825745372
+-1.06779661016949 0.24969979654604
+-0.966101694915254 0.264856035195966
+-0.864406779661017 0.278041836013959
+-0.762711864406779 0.288881021349384
+-0.661016949152542 0.297054723966203
+-0.559322033898305 0.30231695501038
+-0.457627118644067 0.304275803005703
+-0.35593220338983 0.304529357038898
+-0.254237288135593 0.304565924185496
+-0.152542372881356 0.304568537614607
+-0.0508474576271185 0.304568582253443
+0.050847457627119 0.304568582253443
+0.152542372881356 0.304568537614607
+0.254237288135593 0.304565924185496
+0.355932203389831 0.304529357038898
+0.457627118644068 0.304275803005703
+0.559322033898305 0.30231695501038
+0.661016949152542 0.297054723966203
+0.76271186440678 0.288881021349384
+0.864406779661017 0.278041836013959
+0.966101694915254 0.264856035195966
+1.06779661016949 0.24969979654604
+1.16949152542373 0.232988825745372
+1.27118644067797 0.215159529401958
+1.3728813559322 0.19665032476711
+1.47457627118644 0.177884185408375
+1.57627118644068 0.159253358442953
+1.67796610169492 0.141106964487117
+1.77966101694915 0.123741930739631
+1.88135593220339 0.107397436643172
+1.98305084745763 0.0922527949339697
+2.08474576271187 0.0784284692399007
+2.1864406779661 0.065989757683171
+2.28813559322034 0.0549525585640254
+2.38983050847458 0.0452905809004645
+2.49152542372881 0.0369433652193509
+2.59322033898305 0
+2.69491525423729 0
+2.79661016949153 0
+2.89830508474576 0
 3 0

### 2021-10-24 latex_files/python_scripts/example_1D.py

@@ -14,7 +14,7 @@ import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 
 prepath = './../images/'
-savefig = True
+savefig = False
 
 np.random.seed(21)
 class Prior:
@@ -255,7 +255,7 @@ plt.ylim([0, .6])
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_marginal.png"), bbox_inches='tight')
-    tplt.clean_figure()
+    # tplt.clean_figure()
     tplt.save(os.path.join(prepath, "chapter3/example_marginal.tex"))
 plt.show(block=False)
 
@@ -297,7 +297,7 @@ plt.legend()
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_posterior.png"), bbox_inches='tight')
-    tplt.clean_figure()
+    # tplt.clean_figure()
     tplt.save(os.path.join(prepath, "chapter3/example_post.tex"))
 plt.show(block=False)
 
@@ -312,7 +312,6 @@ def wrapper(x):
         tmp = x[i]
         res.append(gt_posterior_pdf(float(tmp)))
     return np.array(res)
-
 
 print(romc.compute_divergence(wrapper, distance="Jensen-Shannon"))
 

### 2021-10-24 latex_files/python_scripts/example_MA2.py

@@ -11,7 +11,7 @@ matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
 prepath = './../images/chapter4/'
-savefig = True
+savefig = False
 
 def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
     plt.figure()
@@ -161,7 +161,7 @@ def plot_romc_posterior(title, posterior, nof_points, savefig):
         tplt.save(savefig[:-3] + "tex")
     plt.show(block=False)
 
-    
+breakpoint()
 # from elfi.methods.parameter_inference import ROMC
 if savefig:
     plot_romc_posterior('ROMC (gradient-based)',

### 2021-10-24 latex_files/python_scripts/example_extensibility.py

@@ -11,7 +11,7 @@ matplotlib.rcParams['text.usetex'] = True
 
 logging.basicConfig(level=logging.INFO)
 prepath = './../images/chapter4/'
-savefig = True
+savefig = False
 
 def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
     plt.figure()
@@ -58,8 +58,8 @@ plt.show(block=False)
 
 
 ####### ROMC with gradients ################
-from elfi.methods.inference.romc import OptimisationProblem
-from elfi.methods.utils import NDimBoundingBox
+from elfi.methods.inference.romc import OptimisationProblem, NDimBoundingBox
+# from elfi.methods.utils import NDimBoundingBox
 from sklearn.pipeline import Pipeline
 from sklearn.neural_network import MLPRegressor
 from functools import partial
@@ -81,6 +81,14 @@ class customOptim(OptimisationProblem):
         nof_samples = 500
         objective = self.objective
 
+        def local_surrogate(theta, model_scikit):
+            assert theta.ndim == 1
+            theta = np.expand_dims(theta, 0)
+            return float(model_scikit.predict(theta))
+
+        def create_local_surrogate(model):
+            return partial(local_surrogate, model_scikit=model)
+
         local_surrogates = []
         for i in range(len(self.regions)):
             # prepare dataset
@@ -91,7 +99,7 @@ class customOptim(OptimisationProblem):
             mlp = MLPRegressor(hidden_layer_sizes=(10,10), solver='adam')
             model = Pipeline([('linear', mlp)])
             model = model.fit(x, y)
-            local_surrogates.append(self.create_local_surrogate(model))
+            local_surrogates.append(create_local_surrogate(model))
 
         self.local_surrogates = local_surrogates
         self.state["local_surrogates"] = True

### 2022-01-27 latex_files/ch4/4-2-inference.tex
added initial version of intro, abstract

@@
   # Number of samples: 19300
-  # Sample means: theta: -0.0132
+  # Sample means: theta: -0.0116
 
@@
   print("Expected value   : %.3f" % exp_val)
-  # Expected value   : -0.013
+  # Expected value   : -0.012
 
@@
   print("Expected variance: %.3f" % exp_var)
-  # Expected variance: 1.116
+  # Expected variance: 1.120
 

### 2022-01-27 paper.tex
small changes
added Neural net
multiple edits until Sec 2.4
reordered files and fixed graph colors
added initial version of intro, abstract

@@
 \usepackage{framed}
+\usepackage{marginnote}
 
@@
 
+\newcommand{\simulator}{g}
 \newcommand{\region}{B_{d,\epsilon}}
@@
 \newcommand{\data}{\mathbf{y_0}}
+\newcommand{\accregioni}{C^i_{\epsilon}}
+\newcommand{\accregionihat}{\hat{C}^i_{\epsilon}}
+
+\newcommand{\danger}{\fontencoding{U}\fontfamily{futs}\selectfont\char 66\relax}
+\newcommand{\TODO}[1]{\textcolor{red}{TODO #1}\marginpar{\textcolor{red}{\Large \danger}}}
 
 \newcommand{\R}{\mathbb{R}}
+
+\newcommand{\Ex}{\mathbb{E}}
 
@@
 %% - \AND starts a new line, \And does not.
-\author{Achim Zeileis\\Universit\"at Innsbruck
-   \And Second Author\\Plus Affiliation}
-\Plainauthor{Achim Zeileis, Second Author}
+\author{Vasilis Gkolemis\\ATHENA RC
+  \And Michael Gutmann\\University of Edinburgh
+   \And Henri Pesonen \\University of Oslo
+}
+\Plainauthor{Vasilis Gkolemis, Michael Gutmann}
 
@@
 \Abstract{
-  This short article illustrates how to write a manuscript for the
-  \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
-  Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
-  it is recommended to keep the {\LaTeX} code as simple as possible,
-  i.e., avoid inclusion of packages/commands that are not necessary.
-  For outlining the typical structure of a JSS article some brief text snippets
-  are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
-  discussing count data regression in \proglang{R}. Editorial comments and
-  instructions are marked by vertical bars.
+  Likelihood-free inference (LFI) or Approximate Bayesian Computation (ABC) is the general framework for performing the inference in statistical models with intractable likelihood.
+In general, the intractability of the likelihood makes the inference problem quite challenging, therefore, many LFI methods either lack accuracy or are pretty inefficient.
+Robust Optimisation Monte Carlo (ROMC) is a highly-parallelizable approximate Bayesian computation (ABC) technique.
+In this paper, we present the implementation of ROMC in the \proglang{Python} package ELFI.
+Our implementation can be exploited in two directions.
+A scientist may use it as an out-of-the-box ABC algorithm for performing the inference in an LFI problem; we provide an easy-to-use API harmonized with the rest of the ELFI framework.
+Therefore, it is directly comparable with the rest of the methods included in ELFI.
+On the other direction, we have carefully designed ROMC in isolated algorithmic steps.
+Therefore, we allow the researcher to experiment with ROMC, altering or replacing parts of the algorithm with custom-created functionalities.
+In both scenarios, the algorithmic steps of ROMC can run in a fully-parallelizable manner, exploiting all CPU cores.
+We also provide some functionalities for visualization and evaluation needs.
+Finally, we test the robustness of our implementation on some typical LFI models that ship.
+
+  % This short article illustrates how to write a manuscript for the
+  % \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
+  % Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
+  % it is recommended to keep the {\LaTeX} code as simple as possible,
+  % i.e., avoid inclusion of packages/commands that are not necessary.
+  % For outlining the typical structure of a JSS article some brief text snippets
+  % are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
+  % discussing count data regression in \proglang{R}. Editorial comments and
+  % instructions are marked by vertical bars.
 }
@@
 %% - Should be comma-separated and in sentence case.
-\Keywords{JSS, style guide, comma-separated, not capitalized, \proglang{R}}
-\Plainkeywords{JSS, style guide, comma-separated, not capitalized, R}
+\Keywords{Bayesian Inference, Simulator-based models, Likelihood-free, \proglang{Python}, \proglang{ELFI}}
+\Plainkeywords{Bayesian Inference, Simulator-based models, Likelihood-free, Python, ELFI}
 
@@
 \Address{
-  Achim Zeileis\\
-  Journal of Statistical Software\\
-  \emph{and}\\
-  Department of Statistics\\
-  Faculty of Economics and Statistics\\
-  Universit\"at Innsbruck\\
-  Universit\"atsstr.~15\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{Achim.Zeileis@R-project.org}\\
-  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+  Vasilis Gkolemis\\
+  Information Management Systems Institute (IMSI)\\
+  ATHENA Research and Innovation Center\\
+  Athens, Greece\\
+  E-mail: \email{vgkolemis@athenarc.gr}\\
+  URL: \url{https://givasile.github.io}
 }
@@
 
-% @Vasilis: we can work on the intro when we have the basic structure of the article in place.
+Simulator-based models are particularly captivating due to the modeling freedom they provide.
+In essence, a simulator-based model can describe any data generating mechanism that can be written as a finite set of algorithmic steps.
+This modeling freedom comes at a cost; performing the inference, i.e., sample or evaluate the posterior distribution, is challenging.
+
+Optimization Monte Carlo (OMC) proposed by [..] was a novel ABC approach approximating the posterior distribution.
+Afterward, Forneron and Ng provided a similar method under "reverse sampler ".
+The key idea is turning the stochastic data-generating mechanism into a set of fully-parallelizable deterministic optimization processes.
+Robust OMC (ROCM) provided an upgrade in OMC by locating and resolving some fundamental failure modes of the latter.
+
+In this paper, we present an extendable and parallelizable implementation of ROMC at the Python package 'Engine for Likelihood-Free inference (ELFI) .' We have followed specific design principles for making the implementation extendible.
+Each step of the method is designed as a fully-isolated and parallelizable step; The user may easily replace parts of the technique with custom components without the backbone method collapse.
+Therefore, our implementation is a friendly tool for the research community to experiment with alterations of ROMC.
+Apart from the main functionalities, we also provide essential visualization and evaluation tools for gaining insights while fitting data.
+Finally, we verify the robustness of the implementation on the LFI examples supplied by the ELFI package.
 
@@
 
-\clearpage
+% \clearpage
 \section{Background}
@@
 \subsection{Implicit models and likelihood-free inference}
-\input{latex_files/ch2/2_2_LFI_background.tex}
+\input{latex_files/ch2/2-1-implicit-models.tex}
 
 \subsection{Optimisation Monte Carlo (OMC)}
-\input{latex_files/ch2/2_3_OMC.tex}
+\input{latex_files/ch2/2-2-OMC.tex}
 
 \subsection{Robust optimisation Monte Carlo (ROMC)}
-\input{latex_files/ch2/2_4_ROMC.tex}
+\input{latex_files/ch2/2-3-ROMC.tex}
 
@@
 \label{subsec:ELFI}
-\input{latex_files/ch3/3_1_ELFI.tex}
+\input{latex_files/ch2/2-4-ELFI.tex}
 
@@
 
-\input{latex_files/ch2/2_5_ROMC_algorithmic.tex}
+\input{latex_files/ch3/3-0-design-principles.tex}
 
@@
 \label{subsec:general_design}
-\input{latex_files/ch3/3_2_1_general_design}
+\input{latex_files/ch3/3-1-implementation-principles.tex}
 
@@
 \label{subsec:training}
-\input{latex_files/ch3/3_2_2_training}
+\input{latex_files/ch4/4-1-training.tex}
 
@@
 \label{subsec:inference}
-\input{latex_files/ch3/3_2_3_inference}
+\input{latex_files/ch4/4-2-inference.tex}
 
@@
 \label{subsec:evaluation}
-\input{latex_files/ch3/3_2_4_evaluation}
+\input{latex_files/ch4/4-3-evaluation.tex}
 
@@
 \label{subsec:extensibility}
-\input{latex_files/ch3/3_3_extensibility.tex}
+\input{latex_files/ch4/4-4-extensibility.tex}
 
@@
 \label{subsec:ma2}
-\input{latex_files/ch4/4_2_ma2.tex}
+\input{latex_files/ch5/5-1-ma2.tex}
 
@@
 \label{subsec:exec_time}
-\input{latex_files/ch4/4_3_exec.tex}
+\input{latex_files/ch5/5-2-execution-time.tex}
 
@@
 
-\section*{Acknowledgments}
+% \section*{Acknowledgments}
 

### 2022-01-28 paper.tex
abstract and intro

@@
 %% - \Abstract{} almost as usual
-\Abstract{
-  Likelihood-free inference (LFI) or Approximate Bayesian Computation (ABC) is the general framework for performing the inference in statistical models with intractable likelihood.
-In general, the intractability of the likelihood makes the inference problem quite challenging, therefore, many LFI methods either lack accuracy or are pretty inefficient.
-Robust Optimisation Monte Carlo (ROMC) is a highly-parallelizable approximate Bayesian computation (ABC) technique.
-In this paper, we present the implementation of ROMC in the \proglang{Python} package ELFI.
-Our implementation can be exploited in two directions.
-A scientist may use it as an out-of-the-box ABC algorithm for performing the inference in an LFI problem; we provide an easy-to-use API harmonized with the rest of the ELFI framework.
-Therefore, it is directly comparable with the rest of the methods included in ELFI.
-On the other direction, we have carefully designed ROMC in isolated algorithmic steps.
-Therefore, we allow the researcher to experiment with ROMC, altering or replacing parts of the algorithm with custom-created functionalities.
-In both scenarios, the algorithmic steps of ROMC can run in a fully-parallelizable manner, exploiting all CPU cores.
-We also provide some functionalities for visualization and evaluation needs.
-Finally, we test the robustness of our implementation on some typical LFI models that ship.
+\Abstract{ Performing the inference in statistical models with
+  intractable likelihood is challenging; most likelihood-free
+  inference (LFI) methods encounter problems in accuracy and
+  efficiency. In this paper, we present the implementation of Robust
+  Optimisation Monte Carlo (ROMC) in the Python package ELFI. ROMC is
+  a novel and efficient (highly-parallelizable) LFI technique that
+  provides accurate weighted samples from the posterior.  Our
+  implementation can be used in two ways. A scientist may use it as an
+  out-of-the-box LFI algorithm. We provide an easy-to-use API
+  harmonized with the rest of the ELFI framework; therefore, it is
+  straightforward to compare it with all the methods included in
+  ELFI. On the other direction, we have carefully split ROMC into
+  isolated algorithmic steps. Therefore, we allow the researcher to
+  experiment, replacing components of the algorithm with
+  custom-designed subparts. In both scenarios, the algorithmic steps
+  of ROMC run in a fully-parallelizable manner, exploiting all CPU
+  cores. We also provide helpful functionalities for (i) inspecting
+  the inference process and (ii) evaluating the approximated
+  posterior. Finally, we test the robustness of our implementation on
+  some typical LFI examples.
 
@@
 
-Simulator-based models are particularly captivating due to the modeling freedom they provide.
-In essence, a simulator-based model can describe any data generating mechanism that can be written as a finite set of algorithmic steps.
-This modeling freedom comes at a cost; performing the inference, i.e., sample or evaluate the posterior distribution, is challenging.
+Simulator-based models are particularly captivating due to the
+modeling freedom they provide. In essence, a simulator-based model can
+describe any data generating mechanism that can be written as a finite
+set of algorithmic steps. This modeling freedom comes at a cost;
+performing the inference, i.e., sample or evaluate the posterior
+distribution, is challenging.
 
-Optimization Monte Carlo (OMC) proposed by [..] was a novel ABC approach approximating the posterior distribution.
-Afterward, Forneron and Ng provided a similar method under "reverse sampler ".
-The key idea is turning the stochastic data-generating mechanism into a set of fully-parallelizable deterministic optimization processes.
-Robust OMC (ROCM) provided an upgrade in OMC by locating and resolving some fundamental failure modes of the latter.
+Optimization Monte Carlo (OMC) proposed by~\citep{Meeds2015} was a
+novel LFI approach for approximating the posterior
+distribution. Afterwards, \citep{Forneron2016} provided a similar
+method under `reverse sampler'. The key idea is turning the stochastic
+data-generating mechanism into a set of fully-parallelizable
+deterministic optimization processes. In their work,
+~\citep{Ikonomov2019}, located some critical failure modes of OMC, and
+proposed Robust OMC (ROMC) an alternative version of OMC with the
+appropriate improvements.
 
-In this paper, we present an extendable and parallelizable implementation of ROMC at the Python package 'Engine for Likelihood-Free inference (ELFI) .' We have followed specific design principles for making the implementation extendible.
-Each step of the method is designed as a fully-isolated and parallelizable step; The user may easily replace parts of the technique with custom components without the backbone method collapse.
-Therefore, our implementation is a friendly tool for the research community to experiment with alterations of ROMC.
-Apart from the main functionalities, we also provide essential visualization and evaluation tools for gaining insights while fitting data.
-Finally, we verify the robustness of the implementation on the LFI examples supplied by the ELFI package.
+In this paper, we present an extendible and parallelizable
+implementation of ROMC at the \proglang{Python} package
+\proglang{Engine for Likelihood-Free inference (ELFI)}. We follow
+specific principles for making the implementation extendible. As we
+will describe analytically in the next chapter, ROMC can be understood
+not only as a specific algorithm, but mainly as a framework for
+obtaining weighted samples from the posterior; it defines some
+algorithmic task that must be fulfilled, without enforcing which
+specific algorithm must be used for the tasks. Therefore, a researcher
+may adopt ROMC's backbone approach, but experiment with alternative
+methods to solve the distinct tasks. We have designed ROMC for
+faciliating such alterations. Finally, we have tested the accuracy and
+the efficienct of our implementation on the LFI examples supported by
+the \proglang{ELFI} package.
 

### 2022-02-01 latex_files/ch5/5-1-ma2.tex
added conclusion and the extensibility part

@@
-Second-order moving average (MA2) is a probabilistic model commonly
-used for time series analysis. The observation at time $t$ is given
-by:
+\subsubsection*{Model Definition}
+
+MA2 is a probabilistic model for time series analysis. The observation
+at time \(t\) is given by,
 
@@
 \noindent
-The random variable $w_{k} \sim \mathcal{N}(0,1), k \in \mathbb{Z}$
-represent white noise and $\theta_1, \theta_2$, the two parameters of
-interest, the dependence from the previous observations. The number of
-consecutive observations $T$ is a hyper-parameter of the model that in
-our case is set at $T=100$. For guaranteeing that the inference
-problem is identifiable, i.e.\ the likelihood has only one mode, we
-use the prior proposed by \cite{Marin2012} given in the equation
-\eqref{eq:ma2_prior}:
+The r.v. \(w_{k} \sim \mathcal{N}(0,1) \) is white noise and the two
+parameters of interest, \(\theta_1, \theta_2\), model the dependence
+from the previous observations. The number of sequential observations
+\(T\) is a hyper-parameter, in our case set to \(T=100\). For securing
+that the inference problem is identifiable, i.e.\ the likelihood has
+only one mode, we use the prior proposed by~\cite{Marin2012},
 
@@
 \noindent
-The observation vector $\yb_0 = (y_1, \ldots, y_{100})$ is generated
-with $\thetab^*=(0.6, 0.2)$. The dimensionality of the output $\yb$ is
-high, therefore we use summary statistics. Considering that the
-ouptput vector represents a time-series signal, we choose as summary
-statistics the autocovariances with $lag=1$ and $lag=2$, as shown in
-equations \eqref{eq:ma2_summary_1} and \eqref{eq:ma2_summary_2}. The
-final distance node is the squared euclidean
-distance \eqref{eq:ma2_summary_4}.
+The observation vector \(\yb_0 = (y_1, \ldots, y_{100})\) is generated
+with \(\thetab^*=(0.6, 0.2)\). The dimensionality of the output
+\(\yb\) is high, therefore we use summary statistics. Considering that
+the ouptput vector represents a time-series signal, we select the
+autocovariances with \(\mathrm{lag}=1\) and \(\mathrm{lag}=2\), as shown in equations
+\eqref{eq:ma2_summary_1} and \eqref{eq:ma2_summary_2}. The final
+distance node is the squared euclidean distance
+\eqref{eq:ma2_summary_4}.
 
@@
 
-\subsubsection*{Perform the inference}
+\subsubsection*{Inference}
 
-As in the previous example, we perform the inference using both
-optimisation alternatives (gradient-based and Bayesian optimisation)
-and, in the end, by extending the implementation with a Neural Network
-surrogate model. In the end, we use the Rejection ABC algorithm for
-evaluating our approaches; Rejection ABC is a robust method in terms
-of accuracy and can be used for baseline comparison.
-
-The implementation's extension is performed using a multi-layer
-perceptron as local surrogate function, as illustrated in chapter
-... This intervention can be done easily by coding a custom
-optimisation function with the surrogate model of our own
-preference. The surrogate model substitutes the real distance function
-$d_i$ inside the proposal region $q_i \quad \forall i$ at the
-inference phase i.e. sampling, computing an expectation and evaluating
-the posterior. In our example we use a neural network of two hidden
-layers of 10 neurons each and we train it in 500 examples for each
-proposal region.
+In order to show all the capabilities of the implementation, we
+perform the inference (i) using the gradient-based optimizer, (ii)
+using the Bayesian Optimization scheme and (iii) fitting a Neural
+Network (NN) as a surrogate model. We use the Rejection ABC algorithm
+for evaluating all approaches. Replacing the typical quadratic
+surrogate model with a NN serves as an illustrator of the
+extensibility of our implementation. The replacement is done by coding
+a custom optimisation function with a surrogate model of our own
+preference, as shown in Chapter~\ref{subsec:extensibility}. For the
+definition of the NN, we use the \pkg{MLPRegressor} class of the
+\pkg{scikit-learn} package. Therefore, the NN substitutes the real
+distance function \(d_i\) inside the proposal region
+\(q_i \quad \forall i\) at the inference phase i.e. sampling,
+computing an expectation and evaluating the posterior. In our example
+we use a neural network of two hidden layers of 10 neurons each and we
+train it sampling 500 examples from each proposal region.
 
@@
 same deterministic simulator, in the gradient-based and the Bayesian
-optimisation case. The acceptance regions are similar even though the
-optimal points differ significantly due to the different optimisation
-schemens.
+optimisation case. The acceptance regions are quite similar even though the
+different optimisation schemes lead to different optimal points.
 
 In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
-marginal posteriors, using all four approaches; (a) Rejection ABC
-(first column), (b) ROMC with gradient-based optimisation (second
-column) (c) ROMC with Bayesian optimisation (third column) and (d)
-ROMC with the Neural Network extendion. We observe that there is a
-significant aggreement between the different approaches. The Rejection
-ABC inference has been set to infer 10000 accepted samples, with
-threshold $\epsilon=0.1$. At table \ref{tab:ma2} we present the
-empirical mean $\mu$ and standard deviation $\sigma$ for each
-inference approach. In figure \ref{fig:ma2_4}, we illustrate the
+marginal posteriors, for each approach; (a) Rejection ABC (first
+column), (b) ROMC with gradient-based optimisation (second column) (c)
+ROMC with Bayesian optimisation (third column) and (d) ROMC with the
+NN extension. We observe a significant aggreement between the
+different approaches. At table \ref{tab:ma2} we present the empirical
+mean \(\mu\) and standard deviation \(\sigma\) for each inference
+approach and finally, in figure \ref{fig:ma2_4}, we illustrate the
 unnormalized posterior for the three different variations of the ROMC
-method.
-
-\subsubsection*{Extending the implementation}
+method. The results show that all ROMC variations provide consistent
+results between them and in comparison with the standard Rejection ABC
+algorithm.
 
@@
 \hline
-& $\mu_{\theta_1}$ & $\sigma_{\theta_1}$ & $\mu_{\theta_2}$ & $\sigma_{\theta_2}$ \\
+& \(\mu_{\theta_1}\) & \(\sigma_{\theta_1}\) & \(\mu_{\theta_2}\) & \(\sigma_{\theta_2}\) \\
 \hline \hline
@@
 %     \caption[MA2 example, histogram of distances]{Histogram of distances
-%       $d_i^*, i \in \{ 1, \ldots, n_1 \}$. The left graph corresponds
+%       \(d_i^*, i \in \{ 1, \ldots, n_1 \}\). The left graph corresponds
 %       to the gradient-based approach and the right one to the Bayesian
@@
       approach. The vertical (red) line represents the samples mean
-      $\mu$ and the horizontal (black) the standard deviation
-      $\sigma$.}
+      \(\mu\) and the horizontal (black) the standard deviation
+      \(\sigma\).}
   \label{fig:ma2_3}
@@
 \end{figure}
+
+\subsubsection*{Parallelization}
+
+As already stated, it is straightforward for the two basic parts of
+the training part to be parallelised. There is a batch of \(n_1\)
+independent objectives functions to be optimised and \(n_1\) bounding
+boxes to be built. Both tasks can be applied in a parallel utilizing
+all the available CPU cores. Our implementation supports
+parallelization using the built-in \proglang{Python} package
+\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
+execution times for performing the inference. The parallel version
+performs all tasks between 2.5 and 6 times faster compared to the
+sequential; the optimization problems are solved almost 6 times
+faster, sampling 3.5 times faster and evaluating the posterior and
+constructing the bounding boxes almost 2.5 times faster.
+
+\begin{figure}[ht]
+  \centering
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Solve optimization problems},
+          xlabel={\(n_1\)},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Proposal regions},
+          xlabel={\(n_1\)},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Sample},
+          xlabel={\(n_1\)},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \begin{minipage}[t]{.49\textwidth}
+      \begin{tikzpicture}
+        \begin{axis}[
+          title={Evaluate posterior},
+          xlabel={\(n_1\)},
+          ylabel={seconds},
+          xmin=1, xmax=500,
+          ymin=0, ymax=7,
+          xtick={1,100,200,300,400,500},
+          ytick={0, 1, 2, 3, 4, 5, 6, 7},
+          legend pos=north west,
+          ymajorgrids=true,
+          grid style=dashed,
+          ]
+          \addplot[color=blue, mark=o] coordinates {
+            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+          };
+          \addlegendentry{sequential}
+          \addplot[color=red, mark=o] coordinates {
+            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+          };
+          \addlegendentry{parallel}          
+        \end{axis}
+      \end{tikzpicture}
+    \end{minipage}\hfill
+    \caption[Execution time exploiting parallelization]{In the first
+      line, we compare the parallel and sequential execution of the
+      training part and at the second line the inference part. At the
+      left figure, we measure the execution time for sampling \(n_2=50\)
+      points per region. At the right figure we measure the execution
+      time for evaluating the posterior at a batch of \(50\) points.}
+      \label{fig:exec_parallel}
+\end{figure}

### 2022-02-01 latex_files/ch5/5-2-execution-time.tex
added conclusion and the extensibility part

@@
 As already stated, it is straightforward for the two basic parts of
-the training part to be parallelised. There is a batch of $n_1$
-independent objectives functions to be optimised and $n_1$ bounding
+the training part to be parallelised. There is a batch of \(n_1\)
+independent objectives functions to be optimised and \(n_1\) bounding
 boxes to be built. Both tasks can be applied in a parallel utilizing
@@
           title={Solve optimization problems},
-          xlabel={$n_1$},
+          xlabel={\(n_1\)},
           ylabel={seconds},
@@
           title={Proposal regions},
-          xlabel={$n_1$},
+          xlabel={\(n_1\)},
           ylabel={seconds},
@@
           title={Sample},
-          xlabel={$n_1$},
+          xlabel={\(n_1\)},
           ylabel={seconds},
@@
           title={Evaluate posterior},
-          xlabel={$n_1$},
+          xlabel={\(n_1\)},
           ylabel={seconds},
@@
       training part and at the second line the inference part. At the
-      left figure, we measure the execution time for sampling $n_2=50$
+      left figure, we measure the execution time for sampling \(n_2=50\)
       points per region. At the right figure we measure the execution
-      time for evaluating the posterior at a batch of $50$ points.}
+      time for evaluating the posterior at a batch of \(50\) points.}
       \label{fig:exec_parallel}

### 2022-02-01 paper.tex
added conclusion and the extensibility part

@@
 
-In this section, we illustrate the implementation performing two
-experiments; In chapter~\ref{subsec:ma2} we perform the inference at
-the second-order moving average (MA2) model, which is a standard model
-of \pkg{ELFI}, and in chapter~\ref{subsec:exec_time} we measure the
-execution speed-up achieved by the parallelization of ROMC.
+In this section, we test the implementation using the second-order
+moving average (MA2) example, which is a standard model of
+\pkg{ELFI}. We perform the inference using three different versions of
+ROMC; (i) using a gradient-based optimizer, (ii) using the Bayesian
+Optimization scheme and (iii) fitting a Neural Network as a surrogate
+model. The later illustrates how to extend the implementation,
+replacing part of ROMC with a user-defined component. Finally, we
+measure the execution speed-up achieved by the parallelized version of
+ROMC.
 
-\subsection{Moving average example}
-\label{subsec:ma2}
+% \subsection{Moving average example}
+% \label{subsec:ma2}
 \input{latex_files/ch5/5-1-ma2.tex}
 
-\subsection{Execution time}
-\label{subsec:exec_time}
-\input{latex_files/ch5/5-2-execution-time.tex}
+% \subsection{Execution time}
+% \label{subsec:exec_time}
+% \input{latex_files/ch5/5-2-execution-time.tex}
 
@@
 \section{Summary and discussion} \label{sec:summary}
+
+In this paper, we presented the implementation details we followed for
+developing the LFI method ROMC at the \proglang{ELFI} package. We paid
+thorough attention to two specific use-case scenarios. Firstly, we
+illustrate how a user may take advantage of our ready-to-use API for
+solving its LFI problem. Secondly, we focus on the scenario where a
+researcher wants to intervene and alter parts of the method. Our
+implementation is designed to support this as well.
+
+By all means, there are still open challenges for the future. Two
+directions may enable ROMC to solve high-dimensional problems
+efficiently. The first one is enabling ROMC's execution into a cluster
+of computers. ROMC can be characterized as an \textit{embarrassingly
+  parallel} workload; each optimization problem is an entirely
+independent task. Therefore, supporting inference into a cluster of
+computers can radically speed up the inference. The second one refers
+to implementing the method in a framework that supports automatic
+differentiation. Automatic differentiation is necessary for
+efficiently solving optimization problems, especially in
+high-dimensional parametric models.
 

### 2022-02-01 latex_files/python_scripts/example_MA2.py

@@ -161,7 +161,6 @@ def plot_romc_posterior(title, posterior, nof_points, savefig):
         tplt.save(savefig[:-3] + "tex")
     plt.show(block=False)
 
-breakpoint()
 # from elfi.methods.parameter_inference import ROMC
 if savefig:
     plot_romc_posterior('ROMC (gradient-based)',

### 2022-02-02 latex_files/ch2/2-1-implicit-models.tex
reordered files and fixed graph colors
Add comments and some changes. Ask HP.

--- 2021-09-23 latex_files/ch2/2_2_LFI_background.tex
@@
 by the random draws performed when running the computer
-code. Importantly, we can fix the value of $\vb$ while changing
+code. \textbf{HP comment: I don't think that the
+distribution $p(\vb)$ is determined by the random draws, but $p(\vb)$ 
+should define the way we choose the `random' draws. How about: The distribution 
+$p(\vb)$ is defined as a random number generating process within the computer 
+code.} Importantly, we can fix the value of $\vb$ while changing
 $\thetab$ by fixing the random seed used.
@@
 implicit models are often used to model physical phenomena in the
-natural sciences such as genetics, epidemiology, or neuroscience, to
-name a few examples. Further background on simulator-based models and
+natural sciences such as e.g.~genetics, epidemiology, or neuroscience. 
+Further background on simulator-based models and
 example applications can be found in the articles
@@
 data by $\data$, the main difficulty is that the likelihood function
-$L(\thetab) = p(\data|\thetab)$ is intractable. To better see the
+$L(\thetab) = p(\data|\thetab)$ is generally intractable. To better see the
 sources of the intractability, and to address them, we go back to the
 basic characterisation of the likelihood as the (rescaled) probability
-for a parameter of the model to generate data $\yb$ that is similar to
+of a parameter of the model to generate data $\yb$ that is similar to
 the observed data $\data$. More formally, the likelihood $L(\thetab)$
@@
   L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
-  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data))
+  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data)  \mid \theta)
 \end{equation}
@@
 highlights two sources of intractability: The first is the computation
-of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$, the second
-is the limit of $\epsilon \to 0$.  Whilst the probability can be
-approximated with samples, this becomes computationally infeasible if
-$\epsilon$ is too small. Hence, a large class of inference methods
-work with $\epsilon >0$, which leads to the approximate likelihood
-function $L_{d, \epsilon}(\thetab)$,
+of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$,
+the second is the limit of $\epsilon \to 0$.  Whilst the probability
+can be approximated with samples, this becomes computationally
+infeasible if $\epsilon$ is too small. Hence, a large class of
+inference methods work with $\epsilon >0$, which leads to the
+approximate likelihood function $L_{d, \epsilon}(\thetab)$,
 \begin{equation} \label{eq:approx_likelihood}
-  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)), \quad \text{where  } \epsilon > 0.
+  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
 \end{equation}
-Most often in likelihood-free inference, we are interested in the posterior distribution of the parameters $\thetab$ given some observed data.
-The approximate likelihood function in \eqref{eq:approx_likelihood} implies the approximate posterior  $p_{d,\epsilon}(\thetab|\data)$,
+Most often in likelihood-free inference, we are interested in the
+posterior distribution of the parameters $\thetab$ given some observed
+data. \textbf{HP comment: thinking about it, I think we're often more 
+interested in a sample from the posterior, but this is maybe just semantics.} The approximate likelihood function in
+\eqref{eq:approx_likelihood} implies the approximate posterior
+$p_{d,\epsilon}(\thetab|\data)$,
 \begin{equation} \label{eq:approx_posterior}
@@
 problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
-Thomas2016, Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both OMC and robust OMC, which we introduce next,
-are based on the approximation in \eqref{eq:approx_likelihood}.
+  Thomas2016, Papamakarios2019, Chen2019, Thomas2020,
+  Hermans2020}. However, both OMC and robust OMC, which we introduce
+next, are based on the approximation in \eqref{eq:approx_likelihood}.
 

### 2022-02-02 latex_files/ch2/2-2-OMC.tex
reordered files and fixed graph colors
Add comments and some changes. Ask HP.

--- 2021-09-23 latex_files/ch2/2_3_OMC.tex
@@
-Our description of OMC \citep{Meeds2015}
-follows \citet{Ikonomov2019} who base their explanation of OMC on the
-approximate likelihood function in \eqref{eq:approx_likelihood}. With
-the indicator function
+Our description of OMC \citep{Meeds2015} follows \citet{Ikonomov2019}
+who base their explanation of OMC on the approximate likelihood
+function in \eqref{eq:approx_likelihood}. With the indicator function
 %
@@
 We thus approximate the likelihood of a specific $\thetab$ by counting
-the fraction of samples that lie inside the area around the
+the fraction of samples that lie inside the volume around the
 observations. The fact that we isolated the randomness of the
@@
 OMC by \citet{Meeds2015} assumes that the distance $d$ is the
-Euclidean distance between summary statistics $\Phi$ of the observed
+Euclidean distance $||\cdot||_2$ between summary statistics $\Phi$ of the observed
 and generated data, and that the $\accregioni$ can be well
-approximated by infinitesimally small ellipses. These assumptions leads to an approximation of the posterior in
-terms of weighted samples $\thetab_i^*$ that achieve the smallest
-distance between observed and simulated data for each realisation
-$\vb_i \sim p(\vb)$, i.e.\
+approximated by infinitesimally small ellipses. These assumptions
+leads to an approximation of the posterior in terms of weighted
+samples $\thetab_i^*$ that achieve the smallest distance between
+observed and simulated data for each realisation $\vb_i \sim p(\vb)$,
+i.e.\
 \begin{equation} \label{eq:omc_opt_prob}

### 2022-02-02 latex_files/ch2/2-3-ROMC.tex
Add comments and some changes. Ask HP.

@@
 this issue to occur. Intuitively, it happens because the weights in
-OMC are only computed from information at $\thetab_i^*$. And using
+OMC are only computed from information at $\thetab_i^*$, and using
 only local information can be misleading. For example if the curvature
@@
 $\thetab_i^*$ is nearly flat, the curvature alone may wrongly indicate
-that $\accregioni$ is much larger than it actually is. We implement in
-our software package the robust generalisation of OMC by
+that $\accregioni$ is much larger than it actually is. In
+our software package we implement the robust generalisation of OMC by
 \citet{Ikonomov2019} that resolves this issue.
 
-ROMC approximates the acceptance regions $\accregioni$ and defines
-local proposal distributions $q_i(\thetab)$ on them from which the
+ROMC approximates the acceptance regions $\accregioni$ and on them defines
+local proposal distributions $q_i(\thetab)$ from which the
 posterior samples $\thetab_{ij} \sim q_i$ are generated. The samples
@@
 \citet{Ikonomov2019} considered uniform distributions as proposal
-distributions so that the main work is to approximate the acceptance
-regions $\accregioni$ and to represent them such that uniform sampling
+distributions so that the main task is to approximate the acceptance
+regions $\accregioni$ and to represent them so that uniform sampling
 is easy. The approximation of the acceptance regions contains two
@@
 differences. In case $d_i$ is not differentiable, Bayesian
-Optimisation~\cite{Shahriari2016} provides an alternative choice. In
+Optimisation~\citep{Shahriari2016} provides an alternative approach. In
 this scenario, apart from obtaining an optimal $\thetab_i^* $, a
@@
 steps. First, we define their axes $\mathbf{v}_d$ based on the
-(estimated) curvature of the distance $\thetab_i^*$, second, we
+(estimated) curvature of the distance $\thetab_i^*$. Second, we
 determine the size of the box via a one-dimensional line-search method
@@
 evaluating the weights $w_{ij}$. Moreover, the extra cost is largely
-eliminated if Bayesian Optimization with a Gaussian processe (GP)
+eliminated if Bayesian Optimisation with a Gaussian process (GP)
 surrogate model was used to obtain $\thetab_i^*$ in the first place
-(see above). In this case, we can use the GP model of the distance
+(see above \textbf{HP comment: what above this refers to?}). In this case, we can use the GP model of the distance
 function instead of $d_i(\thetab)$ to generate the training data. This

### 2022-02-02 latex_files/ch2/2-4-ELFI.tex
reordered files and fixed graph colors
Add comments and some changes. Ask HP.

--- 2021-09-23 latex_files/ch3/3_1_ELFI.tex
@@
   documentation can be found
-  \href{https://elfi.readthedocs.io/en/latest/}{here}}~\cite{1708.00707}
-is a \proglang{Python} package for Likelihood-Free Inference
-(LFI). We selected to implemented ROMC in \pkg{ELFI} since it provides
+  \href{https://elfi.readthedocs.io/en/latest/}{https://elfi.readthedocs.io}}~\cite{1708.00707}
+is a \proglang{Python} package for LFI. We
+selected to implemented ROMC in \pkg{ELFI} since it provides
 convenient modules for all the fundamental components of a
@@
 
-\pkg{ELFI} models the probabilistic model as a Directed Acyclic Graph
+\pkg{ELFI} handles the probabilistic model as a Directed Acyclic Graph
 (DAG). This functionality is based on the package \pkg{NetworkX},
 which supports general-purpose graphs. In most cases, the structure of
-a likelihood-free model follows the pattern of figure~\ref{fig:elfi};
+a likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
 some edges connect the prior distributions to the simulator, the

### 2022-02-02 latex_files/ch3/3-0-design-principles.tex
Add comments and some changes. Ask HP.

@@
 serves as the driver for the implementation at \pkg{ELFI}. At a high
-level, ROMC can be split into the trainind the inference part. In
+level, ROMC can be split into the training and the inference part. In
 general terms, the training part covers all steps for estimating the
-proposal regions and the inference part for getting weighted
+proposal regions and the inference part calculates the weighted
 samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
-Steps 2-10 (before the vertical line) form the training part, while
-steps 12-17 the inference part.
+Steps 2-10 (before the horizontal line) and steps 12-17 form the training 
+part and the inference part, respectively. \textbf{HP comment Algorithm 1: I got a bit 
+lost with $M_r(\thetab)$ (Line 1) and $M_d(\thetab, \vb=\vb_i)$ (Line 4). 
+What is their relationship? Also, are Lines 13, 16 and 17 still under construction?
+Maybe refer to Eq. (X)s in the Algorithm, if wanting to keep the Algorithm compact.}
 
@@
       \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define distance function}
-      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimization problem}
+      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimisation problem}
       \If{$d^*_i > \epsilon$}
@@
     \State $E_{p(\thetab|\data)}[h(\thetab)]$ \Comment{Estimate an expectation}
-    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}
+    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalised posterior}
     \EndProcedure
@@
 $\hessian_i = \jac_i^T\jac_i$, where $\jac_i$ is the Jacobian matrix
-of the summary statistics $\Phi(M_d(\thetab, \vb=\vb_i))$ at
+of the summary statistics $\Phi(M_d(\thetab, \vb=\vb_i))$ \textbf{is 
+$M_d(\cdot, \cdot)$ the deterministic simulator function, or deterministic 
+evaluation of the distance function? If so, what is it's relationship to 
+the summary statistics?} at
 $\thetab_i^*$. The approximation through the Jacobian matrix has the
@@
 region. Algorithm~\ref{alg:sampling_GB} defines the steps for
-computing a weighted sample. Computing an expectation can be done
-easily after weighted samples are obtained using the
-equation~\ref{eq:expectation}, so we do not discuss it
-separately. Evaluating the unnormalised posterior requires the
+computing a weighted sample. Computing the expectation is easy after weighted 
+samples are obtained using the equation~\ref{eq:expectation}, so we do not 
+discuss it separately. Evaluating the unnormalised posterior requires the
 deterministic functions $g_i$ and the prior distribution
@@
 requires iterating over all $g_i$ and evaluating the distance from the
-observed data.
+observed data. \textbf{HP comment: Algorithm 2 - introduce the notation $||\thetab||$, or even 
+better, define the variable separately, instead of using the same notation 
+that was used for the norm - could it be $N$ if $\thetab \in \mathbb{R}^N$. 
+Note that Line 20 would lead to an infinite loop.}
 
@@
 \begin{algorithm}[!ht]
-	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; Requires, a model of distance $d_i(\thetab)$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta\_\text{start}$, maximum iterations $M$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
+	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; 
+  Requires: a model of distance $d_i(\thetab)$, 
+  an optimal point $\thetab_i^*$, 
+  a number of refinements $K$, 
+  a step size $\eta\_\text{start}$, 
+  maximum iterations $M$ and 
+  a curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
 	\begin{algorithmic}[1]

### 2022-02-02 latex_files/ch3/3-1-implementation-principles.tex
Add comments and some changes. Ask HP.

--- 2021-08-25 latex_files/ch3/3_2_1_general_design.tex
@@
 The overview of our implementation is illustrated in
-figure~\ref{fig:romc_overview}; one may interpret
-figure~\ref{fig:romc_overview} as a depiction of the main class and
+Figure~\ref{fig:romc_overview}; one may interpret
+Figure~\ref{fig:romc_overview} as a depiction of the main class and
 the ellipses are the main routines of the class. Following
@@
 whereas the rest (blue ellipses) are the public methods that the user
-interacts with. In figure~\ref{fig:romc_overview}, it can be easily
-observed that the implementation follows consistently the algorithmic
-view presented in~\ref{alg:romc_algorithm}. The training part includes
+interacts with. In Figure~\ref{fig:romc_overview}, it can be easily
+observed that the implementation follows consistently Algorithm~\ref{alg:romc_algorithm}. The training part includes
 all the steps until the computation of the proposal regions
-i.e. sampling the nuisance variables, defining the optimization
+i.e.~sampling the nuisance variables, defining the optimisation
 problems, solving them, constructing the regions and fitting local
 surrogate models. The inference part comprises of evaluating the
-unnormalized posterior (and the normalized when is possible), sampling
+unnormalised posterior (and the normalised when is possible), sampling
 and computing an expectation. We also provide some utilities for
 inspecting the training process, such as plotting the histogram of the
-final distances or visualizing the constructed bounding
+final distances or visualising the constructed bounding
 boxes. Finally, in the evaluation part, we provide two methods for
@@
 
-\subsubsection*{Parallelizing ROMC method}
+\subsubsection*{Parallelising ROMC method}
 
-ROMC has the significant advantage of being fully parallelizable. We
+ROMC has the significant advantage of being fully parallelisable. We
 exploit this fact by implementing a parallel version of all the major
-components of the method; (a) solving the optimization problems, (b)
+components of the method; (a) solving the optimisation problems, (b)
 constructing bounding box regions, (c) sampling and (d) evaluating the
-posterior. We parallelize these processes using the built-in
+posterior. We parallelise these processes using the built-in
 \proglang{Python} package \pkg{multiprocessing}. The specific package
@@
 For illustrating the functionalities we will use the following running
-example introduced by~\cite{Ikonomov2019},
+example introduced by~\cite{Ikonomov2019}, \textbf{HP comment: what is c in Eq. \eqref{eq:1D_example_eq_2}
+and in python snippet? As it is, it will not run. }
 

### 2022-02-02 latex_files/ch4/4-1-training.tex
updated some images
Add comments and some changes. Ask HP.

--- 2021-08-25 latex_files/ch3/3_2_2_training.tex
@@
 optimisation problems and (c) solves them using either a
-gradient-based optimizer or Bayesian optimization. These tasks are
-compleed sequentially, as shown in figure~\ref{fig:romc_overview}. The
+gradient-based optimiser or Bayesian optimisation. These tasks are
+completed sequentially, as shown in Figure~\ref{fig:romc_overview}. The
 definition of the optimisation problems is performed by drawing
@@
 $u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
-used in \pkg{ELFI}'s random simulator. The seed $u_i$ is the
-computational analogous of the random variables $\vb_i$, that were
-defined in the previous chapter; freezing the \code{seed} absorbs all
-the randomness of the simulator transforming it to a deterministic
+used in \pkg{ELFI}'s random simulator. The seeds $u_i$ are
+computationally analogous to the random variables $\vb_i$, that were
+defined in the previous chapter; setting the \code{seed} eliminates all randomness of the simulator transforming it to a deterministic
 function. Setting the argument \code{use_bo=True}, chooses the
@@
 case, apart from obtaining the optimal points $\thetab_i^*$, we also
-fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
+fit a GP as a surrogate model $\hat{d}_i$. In this
 scenario, in the following steps, $\hat{d}_i(\thetab)$ replaces
@@
 \code{use_bo=True}. The function accepts all keyword arguments and
-forwards them to the underlying \code{matplotlib.hist()} function; in
+forwards them to the underlying \code{matplotlib.pyplot.hist()} function; in
 this way the user may customise some properties of the histogram, such
@@
 gradients have been used, or as the Hessian of the GP, in case of
-Bayesian Optimisation. The eigenvectors, which are the axes of the
-bounding box, are computed using the function
-\code{numpy.linalg.eig()}. Afterwards, the limits along the
+Bayesian Optimisation. The eigenvectors are computed using the function
+\code{numpy.linalg.eig()}, and used as the axes of the
+bounding box. Afterwards, the limits along the
 eigenvectors are computed with
@@
 the bounding box around the optimal point, when the parameter space is
-$1D$ or $2D$. The argument \code{i} is the index of the corresponding
-optimization problem i.e.\ $i<n_1$. \code{compute_eps} returns the
+one- or two-dimensional. The argument \code{i} is the index of the corresponding
+optimisation problem i.e.\ $i<n_1$. \code{compute_eps} returns the
 appropriate distance value $d_{i=\kappa}^*$ where
-$\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the collection
+$\kappa = \lfloor quantile \cdot n \rfloor$ from the collection
 $\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the number of
@@
 
-% \begin{figure}[h]
-%   \begin{center}
-%     \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
-%     \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
-%     \end{center}
-%     \caption[Histogram of distances at the 1D example.]{Histogram of
-%       distances and visualisation of a specific region.}
-%     \label{fig:example_training_hist}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \begin{center}
-%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_theta_dist.png}
-%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_region_samples.png}
-%     \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter3/example_posterior.png}
-
-%     \end{center}
-%     \caption[Histogram of distances at the 1D example.]{Histogram of
-%       distances and visualisation of a specific region.}
-%     \label{fig:example_training_hist}
-%   \end{figure}
-
 \begin{figure}[ht]
@@
     \input{latex_files/images/chapter3/example_gt.tex}
-    \input{latex_files/images/chapter3/example_posterior.tex}
+    \input{latex_files/images/chapter3/example_post.tex}
   \end{center}
-  \caption[Histogram of distances at the 1D example.]{Histogram of
+  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
     distances and visualisation of a specific region.}

### 2022-02-02 latex_files/ch4/4-2-inference.tex
Add comments and some changes. Ask HP.

@@
 box and the weight $w_i$ is computed as in
-algorithm~\ref{alg:sampling_GB}.
+Algorithm~\ref{alg:sampling_GB}.
 
@@
 expression~\eqref{eq:expectation}. The argument \code{h} can be any
-python \code{Callable}. \code{eval_unnorm_posterior} computes the
+python \code{Callable}. Method \code{eval_unnorm_posterior} computes the
 unnormalised posterior approximation using
-expression~\eqref{eq:approx_posterior}. \code{eval_posterior}
+expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
 evaluates the normalised posterior estimating the partition function
@@
 In the following code snippet, we perform weighted sampling from the
-ROMC approximate posterior. Afterwards, we use some of ELFI's
-built-in tools to get a summary of the obtained samples. In figure
+ROMC approximate posterior. Afterwards, we use some of \pkg{ELFI}'s
+built-in tools to get a summary of the obtained samples. In Figure
 \ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
@@
 utility can provide a nice plot of the approximate posterior as
-illustrated in figure~\ref{fig:running_example_romc_inference}. We observe that the
+illustrated in Figure~\ref{fig:running_example_romc_inference}. We observe that the
 approximation is quite close to the ground-truth posterior.

### 2022-02-02 latex_files/ch4/4-3-evaluation.tex
Add comments and some changes. Ask HP.

@@
 This function computes the divergence between the ROMC approximation
-and the ground truth posterior. Since the computation is performed
+and the ground truth posterior. Because the computation is performed
 using the Riemann's approximation, this method can only work in low
@@
 ground-truth posterior to be available. However, in cases where the
-posterior can be approximated decently well with a numerical approach
+posterior can be approximated acceptably well with a numerical approach
 (as in the current example) or with some other inference approach,
@@
 in the Riemann's approximation and the argument \code{distance} can
-take either the \code{Jensen-Shannon} or the \code{KL-divergence}
-value, for computing the appropriate distance. \code{compute_ess}
+be either \code{"Jensen-Shannon"} or \code{"KL-divergence"}, for computing 
+the appropriate distance. Method \code{compute_ess}
 computes the Effective Sample Size (ESS) using the following

### 2022-02-02 latex_files/ch5/5-1-ma2.tex
Add comments and some changes. Ask HP.

@@
 \begin{gather} \label{eq:ma2}
-y_t = w_t + \theta_1 w_{t-1} + \theta_2 w_{t-2}\\
+y_t = w_t + \theta_1 w_{t-1} + \theta_2 w_{t-2}, \quad t=1, \ldots, T\\
 \theta_1, \theta_2 \in \R, \quad  w_k \sim \mathcal{N}(0,1), k \in \mathbb{Z}
@@
 from the previous observations. The number of sequential observations
-\(T\) is a hyper-parameter, in our case set to \(T=100\). For securing
+\(T\) is a constant and set to \(T=100\). For securing
 that the inference problem is identifiable, i.e.\ the likelihood has
@@
 \eqref{eq:ma2_summary_1} and \eqref{eq:ma2_summary_2}. The final
-distance node is the squared euclidean distance
+distance node is the squared Euclidean distance
 \eqref{eq:ma2_summary_4}.
@@
 perform the inference (i) using the gradient-based optimizer, (ii)
-using the Bayesian Optimization scheme and (iii) fitting a Neural
+using the Bayesian Optimisation scheme and (iii) fitting a Neural
 Network (NN) as a surrogate model. We use the Rejection ABC algorithm
@@
 distance function \(d_i\) inside the proposal region
-\(q_i \quad \forall i\) at the inference phase i.e. sampling,
+\(q_i~ \forall i\) at the inference phase i.e.~sampling,
 computing an expectation and evaluating the posterior. In our example
@@
 
-In figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
+In Figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
 same deterministic simulator, in the gradient-based and the Bayesian
@@
 
-In figure \ref{fig:ma2_3}, we demonstrate the histograms of the
+In Figure \ref{fig:ma2_3}, we demonstrate the histograms of the
 marginal posteriors, for each approach; (a) Rejection ABC (first
@@
 ROMC with Bayesian optimisation (third column) and (d) ROMC with the
-NN extension. We observe a significant aggreement between the
-different approaches. At table \ref{tab:ma2} we present the empirical
+NN extension. We observe a significant agreement between the
+different approaches. At Table \ref{tab:ma2} we present the empirical
 mean \(\mu\) and standard deviation \(\sigma\) for each inference
-approach and finally, in figure \ref{fig:ma2_4}, we illustrate the
-unnormalized posterior for the three different variations of the ROMC
+approach and finally, in Figure \ref{fig:ma2_4}, we illustrate the
+unnormalised posterior for the three different variations of the ROMC
 method. The results show that all ROMC variations provide consistent
@@
 
-\begin{center} \label{tab:ma2}
+\begin{table}
+\begin{center}
 \begin{tabular}{ c|c|c|c|c }
@@
 \end{center}
-
+\caption{\textbf{HP comment: Needs a fine caption to get Table number.}\label{tab:ma2}}
+\end{table}
 % \begin{figure}[H]
@@
       in the second row, using ROMC with a gradient-based optimiser
-      and (c) in the third row, using ROMC with Bayesian optimization
+      and (c) in the third row, using ROMC with Bayesian optimisation
       approach. The vertical (red) line represents the samples mean
@@
     \end{center}
-    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
+    \caption[MA2 example, posterior distribution.]{The unnormalised posterior distribution using the ROMC method with (a) a gradient-based optimisation (b) Bayesian Optimisation (c) gradient-based with a Neural Network as a surrogate model.}
   \label{fig:ma2_4}
@@
 
-\subsubsection*{Parallelization}
+\subsubsection*{Parallelisation}
 
@@
 independent objectives functions to be optimised and \(n_1\) bounding
-boxes to be built. Both tasks can be applied in a parallel utilizing
+boxes to be built. Both tasks can be applied in a parallel utilising
 all the available CPU cores. Our implementation supports
-parallelization using the built-in \proglang{Python} package
-\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
+parallelisation using the built-in \proglang{Python} package
+\pkg{multiprocess}. In Figure~\ref{fig:exec_parallel} we observe the
 execution times for performing the inference. The parallel version
 performs all tasks between 2.5 and 6 times faster compared to the
-sequential; the optimization problems are solved almost 6 times
+sequential; the optimisation problems are solved almost 6 times
 faster, sampling 3.5 times faster and evaluating the posterior and
@@
         \begin{axis}[
-          title={Solve optimization problems},
+          title={Solve optimisation problems},
           xlabel={\(n_1\)},
@@
     \end{minipage}\hfill
-    \caption[Execution time exploiting parallelization]{In the first
+    \caption[Execution time exploiting parallelisation]{In the first
       line, we compare the parallel and sequential execution of the

### 2022-02-02 latex_files/ch5/5-2-execution-time.tex
Add comments and some changes. Ask HP.

@@
 independent objectives functions to be optimised and \(n_1\) bounding
-boxes to be built. Both tasks can be applied in a parallel utilizing
+boxes to be built. Both tasks can be applied in a parallel utilising
 all the available CPU cores. Our implementation supports
-parallelization using the built-in \proglang{Python} package
-\pkg{multiprocess}. In figure~\ref{fig:exec_parallel} we observe the
+parallelisation using the built-in \proglang{Python} package
+\pkg{multiprocess}. In Figure~\ref{fig:exec_parallel} we observe the
 execution times for performing the inference. The parallel version

### 2022-02-02 latex_files/graphs/romc_graph.tex
Add comments and some changes. Ask HP.

@@
 ROMC implementation. The training part follows a sequential pattern;
-the functions in the green ellipses must be called in a sequential
+the functions in the green ellipses \textbf{There are no ellipses.} must be called in a sequential
 fashion for completing the training part and define the posterior

### 2022-02-02 paper.tex
Add comments and some changes. Ask HP.

@@
   out-of-the-box LFI algorithm. We provide an easy-to-use API
-  harmonized with the rest of the ELFI framework; therefore, it is
+  harmonised with the rest of the ELFI framework; therefore, it is
   straightforward to compare it with all the methods included in
@@
 Simulator-based models are particularly captivating due to the
-modeling freedom they provide. In essence, a simulator-based model can
+modelling freedom they provide. In essence, a simulator-based model can
 describe any data generating mechanism that can be written as a finite
-set of algorithmic steps. This modeling freedom comes at a cost;
+set of algorithmic steps. This modelling freedom comes at a cost;
 performing the inference, i.e., sample or evaluate the posterior
@@
 data-generating mechanism into a set of fully-parallelizable
-deterministic optimization processes. In their work,
-~\citep{Ikonomov2019}, located some critical failure modes of OMC, and
+deterministic optimisation processes. In their work,
+~\citet{Ikonomov2019}, located some critical failure modes of OMC, and
 proposed Robust OMC (ROMC) an alternative version of OMC with the
@@
 
-In this paper, we present an extendible and parallelizable
+In this paper, we present an extendible and parallelisable
 implementation of ROMC at the \proglang{Python} package
@@
 methods to solve the distinct tasks. We have designed ROMC for
-faciliating such alterations. Finally, we have tested the accuracy and
-the efficienct of our implementation on the LFI examples supported by
+facilitating such alterations. Finally, we have tested the accuracy and
+the efficiency of our implementation on the LFI examples supported by
 the \proglang{ELFI} package.
@@
 We first give a short introduction to implicit parametric models and
-likelihood-free inference. We then focus on the methods relevant for
-this paper: first the earlier Optmisation Monte Carlo (OMC) and then
+LFI. We then focus on the methods relevant for
+this paper: first the earlier OMC and then
 its recent robust generalisation, robust OMC, for which we provide the
-software package as part of the Engine for likelihood-free inference
-(ELFI), a larger inference platform that we briefly introduce at the
+software package as part of ELFI, a larger inference platform that we briefly introduce at the
 end of this background section.
@@
 In this section, we present the implementation. At~\ref{subsec:ELFI}
-we introduce the reader to the \proglang{Python} package \pkg{Engine
-  for likelihood-free inference
-  (ELFI)}. At~\ref{subsec:general_design} we explain the general
+we introduce the reader to the \proglang{Python} package \pkg{ELFI}. At~\ref{subsec:general_design} we explain the general
 design that we follow. Then, we divide the implementation into three
@@
 In this section, we test the implementation using the second-order
-moving average (MA2) example, which is a standard model of
+moving average (MA2) example, which is one of the standard models of
 \pkg{ELFI}. We perform the inference using three different versions of
-ROMC; (i) using a gradient-based optimizer, (ii) using the Bayesian
-Optimization scheme and (iii) fitting a Neural Network as a surrogate
+ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
+Optimisation scheme and (iii) fitting a Neural Network as a surrogate
 model. The later illustrates how to extend the implementation,
 replacing part of ROMC with a user-defined component. Finally, we
-measure the execution speed-up achieved by the parallelized version of
+measure the execution speed-up achieved by the parallelised version of
 ROMC.
@@
 In this paper, we presented the implementation details we followed for
-developing the LFI method ROMC at the \proglang{ELFI} package. We paid
+developing the LFI method ROMC at the \pkg{ELFI} package. We paid
 thorough attention to two specific use-case scenarios. Firstly, we
@@
 
-By all means, there are still open challenges for the future. Two
+There are still open challenges for the left for future research. Two
 directions may enable ROMC to solve high-dimensional problems
@@
 differentiation. Automatic differentiation is necessary for
-efficiently solving optimization problems, especially in
+efficiently solving optimisation problems, especially in
 high-dimensional parametric models.

### 2022-02-02 latex_files/python_scripts/example_1D.py

@@ -39,7 +39,7 @@ class Likelihood:
     """
 
     def rvs(self, theta, seed=None):
-        """Vectorized sampling from likelihood.
+        """Vectorised sampling from likelihood.
 
         """
         assert isinstance(theta, np.ndarray)

### 2022-02-02 latex_files/python_scripts/execution_times_parallel.py

@@ -268,7 +268,7 @@ for i, n in enumerate(n1):
 
 
 plt.figure()
-plt.title("Optimization problems: Sequential vs Parallel Execution")
+plt.title("Optimisation problems: Sequential vs Parallel Execution")
 plt.plot(n1, solve_grad, "bo--", label="linear")
 plt.plot(n1, solve_grad_parallel, "go--", label="parallel")
 plt.xlabel(r"$n_1$")

### 2022-02-07 latex_files/ch2/2-2-OMC.tex
addressed henri's comments

--- 2021-09-23 latex_files/ch2/2_3_OMC.tex
@@
-Our description of OMC \citep{Meeds2015}
-follows \citet{Ikonomov2019} who base their explanation of OMC on the
-approximate likelihood function in \eqref{eq:approx_likelihood}. With
-the indicator function
+Our description of OMC \citet{Meeds2015} follows \citet{Ikonomov2019}
+who base their explanation of OMC on the approximate likelihood
+function in \eqref{eq:approx_likelihood}. With the indicator function
 %
@@
 %
-we can write the approximate likelihood function $L_{d, \epsilon}(\thetab)$ in \eqref{eq:approx_likelihood} as
+we can write the approximate likelihood function \(L_{d, \epsilon}(\thetab)\) in \eqref{eq:approx_likelihood} as
 \begin{gather} \label{eq:approx_likelihood_omc}
@@
 \end{equation}
-We thus approximate the likelihood of a specific $\thetab$ by counting
-the fraction of samples that lie inside the area around the
+We thus approximate the likelihood of a specific \(\thetab\) by counting
+the fraction of samples that lie inside the volume around the
 observations. The fact that we isolated the randomness of the
-simulator via the $\vb_i$ has two crucial consequences: First, we can
-sample the nuisance variables $\vb_i \sim p(\vb)$ only once and use
-the same set of samples to approximate $L_{d,\epsilon}(\thetab)$ for
-different values of $\thetab$. Secondly, since
-$\yb_i=\simulator(\thetab, \vb_i)$ is a function of $\thetab$,
-checking whether $\yb_i$ is contained in region $\region (\data)$ for
-a particular value of $\thetab$ is the same as checking whether
-$\thetab$ is in the acceptance region $\accregioni
-= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}$. This
+simulator via the \(\vb_i\) has two crucial consequences: First, we can
+sample the nuisance variables \(\vb_i \sim p(\vb)\) only once and use
+the same set of samples to approximate \(L_{d,\epsilon}(\thetab)\) for
+different values of \(\thetab\). Secondly, since
+\(\yb_i=\simulator(\thetab, \vb_i)\) is a function of \(\thetab\),
+checking whether \(\yb_i\) is contained in region \(\region (\data)\) for
+a particular value of \(\thetab\) is the same as checking whether
+\(\thetab\) is in the acceptance region \(\accregioni
+= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}\). This
 leads to the likelihood approximation
@@
 perspective for likelihood-free inference by shifting the focus onto
-the geometry of the acceptance regions $\accregioni$. Indeed, the task
-of approximating the likelihood and/or posterior becomes a task of
-characterising the sets $\accregioni$.
-
-OMC by \citet{Meeds2015} assumes that the distance $d$ is the
-Euclidean distance between summary statistics $\Phi$ of the observed
-and generated data, and that the $\accregioni$ can be well
-approximated by infinitesimally small ellipses. These assumptions leads to an approximation of the posterior in
-terms of weighted samples $\thetab_i^*$ that achieve the smallest
-distance between observed and simulated data for each realisation
-$\vb_i \sim p(\vb)$, i.e.\
+the geometry of the acceptance regions \(\accregioni\). Indeed, the
+task of approximating the likelihood and/or posterior becomes a task
+of characterising the sets \(\accregioni\). OMC by \citet{Meeds2015}
+assumes that the distance \(d\) is the Euclidean distance
+\(||\cdot||_2\) between summary statistics \(\Phi\) of the observed
+and generated data, and that the \(\accregioni\) can be well
+approximated by infinitesimally small ellipses. These assumptions
+leads to an approximation of the posterior in terms of weighted
+samples \(\thetab_i^*\) that achieve the smallest distance between
+observed and simulated data for each realisation
+\(\vb_i \sim p(\vb)\), i.e.\
 \begin{equation} \label{eq:omc_opt_prob}
@@
 \end{equation}
-The weighting for each $\thetab_i^*$ is proportional to the prior
-density at $\thetab_i^*$ and inversely proportional to the determinant
-of the Jacobian matrix of the summary statistics at $\thetab_i^*$. For
+The weighting for each \(\thetab_i^*\) is proportional to the prior
+density at \(\thetab_i^*\) and inversely proportional to the determinant
+of the Jacobian matrix of the summary statistics at \(\thetab_i^*\). For
 further details on OMC we refer the reader to \citep{Meeds2015,

### 2022-02-07 latex_files/ch2/2-3-ROMC.tex
addressed henri's comments

@@
 this issue to occur. Intuitively, it happens because the weights in
-OMC are only computed from information at $\thetab_i^*$, and using
+OMC are only computed from information at \(\thetab_i^*\), and using
 only local information can be misleading. For example if the curvature
-of $||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at
-$\thetab_i^*$ is nearly flat, the curvature alone may wrongly indicate
-that $\accregioni$ is much larger than it actually is. In
+of \(||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2\) at
+\(\thetab_i^*\) is nearly flat, the curvature alone may wrongly
+indicate that \(\accregioni\) is much larger than it actually is. In
 our software package we implement the robust generalisation of OMC by
@@
 
-ROMC approximates the acceptance regions $\accregioni$ and on them defines
-local proposal distributions $q_i(\thetab)$ from which the
-posterior samples $\thetab_{ij} \sim q_i$ are generated. The samples
-are assigned (importance) weights $w_{ij}$ that compensate for using
-the proposal distributions $q_i(\thetab)$ and not the prior
-$p(\thetab)$,
+ROMC approximates the acceptance regions \(\accregioni\) and on them defines
+local proposal distributions \(q_i(\thetab)\) from which the
+posterior samples \(\thetab_{ij} \sim q_i\) are generated. The samples
+are assigned (importance) weights \(w_{ij}\) that compensate for using
+the proposal distributions \(q_i(\thetab)\) and not the prior
+\(p(\thetab)\),
 \begin{equation}
@@
 Given the weighted samples, any expectation
-$\Ex_{p(\thetab|\data)}[h(\thetab)]$ of some function $h(\thetab)$, can be approximated as
+\(\Ex_{p(\thetab|\data)}[h(\thetab)]\) of some function \(h(\thetab)\), can be approximated as
 \begin{equation} \label{eq:expectation}
@@
 distributions so that the main task is to approximate the acceptance
-regions $\accregioni$ and to represent them so that uniform sampling
+regions \(\accregioni\) and to represent them so that uniform sampling
 is easy. The approximation of the acceptance regions contains two
@@
 problems as in OMC, (2) constructing bounding boxes around
-$\accregioni$ and optionally, (3) refining the approximation via a
-probabilistic surrogate model of the distance.
+\(\accregioni\) and optionally, (3) refining the approximation via a
+surrogate model of the distance.
+
 \subsubsection*{Solving the deterministic optimisation problems}
-For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$, we
-search for a point $\thetab^*_i$ such that
-$d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon$. For notational
-convenience, we denote the distance $d(\simulator(\thetab,\vb_i), \data)$ by
-$d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
+For each set of nuisance variables \(\vb_i, i = \{1,2,\ldots,n_1 \}\), we
+search for a point \(\thetab^*_i\) such that
+\(d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon\). For notational
+convenience, we denote the distance \(d(\simulator(\thetab,\vb_i), \data)\) by
+\(d_i(\thetab)\).  Obtaining \(\theta_i^*\) involves solving the following
 optimisation problem:
@@
 The optimisation problem can be treated as unconstrained, accepting
-the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
-only if $d_i(\thetab_i^*) \le \epsilon$. If $d_i(\thetab)$ is
+the optimal point
+\(\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)\) only if
+\(d_i(\thetab_i^*) \le \epsilon\). If \(d_i(\thetab)\) is
 differentiable any gradient-based optimizer can be used
-for~\ref{eq:optProb}. The gradients $\nabla_{\thetab} d_i(\thetab)$
+for~\ref{eq:optProb}. The gradients \(\nabla_{\thetab} d_i(\thetab)\)
 can be either provided in closed form or approximated by finite
-differences. In case $d_i$ is not differentiable, Bayesian
-Optimisation~\citep{Shahriari2016} provides an alternative approach. In
-this scenario, apart from obtaining an optimal $\thetab_i^* $, a
-surrogate model $\hat{d}_i(\thetab)$ of the distance function
-$d_i(\thetab)$ is also automatically obtained; $\hat{d}_i$ can then
-substitute the actual distance function in downstream steps of the
-algorithms, which can lead to computational gains if running the
-simulator and hence evaluating the distance function $d_i(\thetab)$ is
-computationally expensive.
+differences. In case \(d_i\) is not differentiable, Bayesian
+Optimisation~\citep{Shahriari2016} provides an alternative
+approach. In this scenario, apart from obtaining an optimal
+\(\thetab_i^* \), a surrogate model \(\hat{d}_i(\thetab)\) of the
+distance function \(d_i(\thetab)\) is also automatically obtained;
+\(\hat{d}_i\) can then substitute the actual distance function in
+downstream steps of the algorithms, which can lead to computational
+gains if running the simulator for evaluating the actual distance
+\(d_i(\thetab)\) is expensive.
 
 \subsubsection*{Estimating the acceptance regions}
-The acceptance region $\accregioni$ is approximated by a bounding box
-$\accregionihat$ around it. Ideally, we want the bounding box to be as
-tight as possible to $\mathcal{S}_i$, to ensure high acceptance rate
-in the importance sampling, and big enough for not discarding valid
-parts of $\mathcal{S}_i$. The bounding boxes are built in two
-steps. First, we define their axes $\mathbf{v}_d$ based on the
-(estimated) curvature of the distance $\thetab_i^*$. Second, we
+The acceptance region \(\accregioni\) is approximated by a bounding
+box \(\accregionihat\). Ideally, we want the bounding box to be as
+tight as possible to \(\mathcal{S}_i\), to ensure high acceptance rate
+in the importance sampling, but big enough for not discarding valid
+parts of \(\mathcal{S}_i\). The bounding boxes are built in two
+steps. First, we define their axes \(\mathbf{v}_d\) based on the
+(estimated) curvature of the distance at \(\thetab_i^*\). Second, we
 determine the size of the box via a one-dimensional line-search method
@@
 details. After the bounding boxes construction, a uniform distribution
-$q_i$ is defined on each bounding box, and is used as the proposal
+\(q_i\) is defined on each bounding box, and is used as the proposal
 region for importance sampling.
@@
 \subsubsection*{Refining the estimate via a local surrogate model (optional)}
-When computing the weight $w_{ij}$ in \eqref{eq:sampling}, we need to
-check whether the samples $\thetab_{ij} \sim q_i$ lie inside the
-acceptance region $\accregioni$. This can be considered to be a
+When computing the weight \(w_{ij}\) in \eqref{eq:sampling}, we need
+to check whether the samples \(\thetab_{ij} \sim q_i\) lie inside the
+acceptance region \(\accregioni\). This can be considered to be a
 safety-mechanism that corrects for any inaccuracies in the
-construction of $\accregionihat$ above. However, this check involves
-evaluting the distance function $d_i(\thetab_{ij})$ and hence running
-the simulator $\simulator(\thetab_{ij}, \vb_i)$, which can be
-expensive if the model is complex.
+construction of \(\accregionihat\) above. However, this check involves
+evaluting the distance function \(d_i(\thetab_{ij})\), which can be
+expensive if the model is complex. \citet{Ikonomov2019} thus proposed
+to fit a surrogate model \(\tilde{d}_i(\thetab)\) of the distance
+function \(d_i(\thetab)\), on data points that lie inside
+\(\accregionihat\). They used a simple quadratic model whilst other
+regression models are, in principle, possible too. The advantage of
+using a quadratic model is that it has ellipsoidal isocontours, which
+thus naturally allowed \citet{Ikonomov2019} to replace the bounding
+box approximation of \(\accregioni\) with a tigher-fitting ellipsoidal
+approximation.\footnote{The difference to the infinitesimal
+  ellipsoidal model in OMC is the estimation procedure: OMC uses
+  information at \(\thetab_i^*\) whilst, here, information in
+  \(\accregionihat\) is used, which results in a more stable fit.}
 
-\citet{Ikonomov2019} thus proposed to construct a surrogate model $\hat{d}_i(\thetab)$
-of the distance function $d_i(\thetab)$ valid in $\accregionihat$ and
-that is is cheap to evaluate. They used a simple quadratic model
-whilst other regression models are, in principle, possible too. The
-advantage of using a quadratic model is that it has ellipsoidal
-isocontours, which thus naturally allowed \citet{Ikonomov2019} to
-replace the bounding box approximation of $\accregioni$ with a
-tigher-fitting ellipsoidal approximation.\footnote{The difference to
-the infinitesimal ellipsoidal model in OMC is the estimation procedure: OMC uses
-information at $\thetab_i^*$ whilst, here, information in
-$\accregionihat$ is used, which results in a more stable fit.}
-
-The training data for fitting the quadratic model is obtained by
-sampling $\thetab \sim q_i$ and evaluating the corresponding distances
-$d_i(\thetab)$. The generation of the training data adds an extra
-computational cost, but leads to a significant speed-up when
-evaluating the weights $w_{ij}$. Moreover, the extra cost is largely
+The training data for the quadratic model is obtained by sampling
+\(\thetab_{ij} \sim q_i\) and accessing the distances
+\(d_i(\thetab_{ij})\). The generation of the training data adds an
+extra computational cost, but leads to a significant speed-up when
+evaluating the weights \(w_{ij}\). Moreover, the extra cost is largely
 eliminated if Bayesian Optimisation with a Gaussian process (GP)
-surrogate model was used to obtain $\thetab_i^*$ in the first place
-(see above \textbf{HP comment: what above this refers to?}). In this case, we can use the GP model of the distance
-function instead of $d_i(\thetab)$ to generate the training data. This
-essentially replaces the global GP model with a simpler local
-quadratic model which is typically more robust.
+surrogate model \(\hat{d}_i(\thetab)\) was used to obtain
+\(\thetab_i^*\) in the first step. In this case, we can use
+\(\hat{d}_i(\thetab)\) instead of \(d_i(\thetab)\) to generate the
+training data. This essentially replaces the global GP model with a
+simpler local quadratic model which is typically more robust.

### 2022-02-07 latex_files/ch3/3-0-design-principles.tex
addressed henri's comments

--- 2021-08-26 latex_files/ch2/2_5_ROMC_algorithmic.tex
@@
 general terms, the training part covers all steps for estimating the
-proposal regions and the inference part for getting weighted
+proposal regions and the inference part calculates the weighted
 samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
-Steps 2-10 (before the vertical line) form the training part, while
-steps 12-17 the inference part.
+Steps 2-10 (before the horizontal line) form the training part and
+steps 12-17 the inference part, respectively.
 
 \begin{algorithm}[!ht]
-	\caption{ROMC}\label{alg:romc_algorithm}
+	\caption{ROMC. Requires the prior \( p(\thetab) \), the simulator
+    \(M_r(\thetab)\), number of optimisation problems \(n_1\), number
+    of samples per region \(n_2\), acceptance limit
+    \(\epsilon\)}\label{alg:romc_algorithm}
 	\begin{algorithmic}[1]
-    \Procedure{ROMC}{$M_r(\thetab), n_1, n_2, \epsilon$}
-    \For{$i \gets 1 \textrm{ to } n_1$}
-      \State $\vb_i \sim p(\vb)$ \Comment{Draw nuisance variables}
-      \State $d_i(\thetab) = M_d(\thetab, \vb=\vb_i)$ \Comment{Define distance function}
-      \State $\thetab_i^* = \text{argmin}_{\thetab} d_i$, $d_i^*=d_i(\thetab_i^*)$ \Comment{Solve optimization problem}
-      \If{$d^*_i > \epsilon$}
+    \Procedure{ROMC}{}
+    \For{\(i \gets 1 \textrm{ to } n_1\)}
+    \State \(\vb_i \sim p(\vb)\) \Comment{Draw nuisance variables}
+    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \vb=\vb_i) \) \Comment{Define deterministic simulator}
+      \State \( d_i(\thetab) = ||g(\thetab, \vb=\vb_i) - \yb_0||_2 \) \Comment{Define distance function}
+      \State \(\thetab_i^* = \text{argmin}_{\thetab} d_i\), \(d_i^*=d_i(\thetab_i^*)\) \Comment{Solve optimisation problem}
+      \If{\(d^*_i > \epsilon\)}
         \State Go to 2 \Comment{Filter solution}
       \EndIf
-      \State Define $q_i$ over $\mathcal{\hat{S}}$ \Comment{Estimate proposal area}
-      \State (Optional) Fit $\Tilde{d}_i$ on $\mathcal{\hat{S}}$ \Comment{Fit surrogate model}
+      \State Estimate \(\mathcal{\hat{S}}_i\) and define \(q_i\) \Comment{Estimate proposal area}
+      \State (Optional) Fit \(\Tilde{d}_i\) on \(\mathcal{\hat{S}}_i\) \Comment{Fit surrogate model}
       \\\hrulefill
-      \For{$j \gets 1 \textrm{ to } n_2$}      
-        \State $\thetab_{ij} \sim q_i$, compute $w_{ij}$ \Comment{Sample}
+      \For{\(j \gets 1 \textrm{ to } n_2\)}
+      \State \(\thetab_{ij} \sim q_i\), compute \(w_{ij}\) as in Algorithm \ref{alg:sampling_GB} \Comment{Sample}
       \EndFor
     \EndFor
-    \State $E_{p(\thetab|\data)}[h(\thetab)]$ \Comment{Estimate an expectation}
-    \State $p_{d,\epsilon}(\thetab)$ \Comment{Evaluate the unnormalized posterior}
+    \State \(E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
+    \State \(p_{d,\epsilon}(\thetab) \) as in eq.~\eqref{eq:approx_posterior_omc} \Comment{Evaluate the unnormalised posterior}
     \EndProcedure
@@
 At the training (fitting) part, the goal is the estimation of the
-proposal regions $\mathcal{\hat{S}}_i$. The tasks are; (a) sampling
-the nuisance variables $\vb_i \sim p(\vb)$ for obtaining the
-deterministic functions $d_i(\thetab)$, (b) defining the optimisation
-problems $\min_{\thetab} \: d_i(\thetab)$, (c) obtaining
-$\thetab_i^*$, $d_i^*$ (d) checking whether $d_i^* \leq \epsilon$ and
-(e) computing the proposal distribution $q_i$. If $d_i(\thetab)$ is
-differentiable, using a gradient-based method is advised for obtaining
-$\thetab_i^*$ faster. In case $\nabla_{\thetab} d_i$ is available in
-closed-form, the procedure is upgraded in terms of both accuracy and
-efficiency. If closed-form description is not available, gradients
-approximation with finite-differences requires two evaluations of
-$d_i$ for \emph{each} parameter $\theta_d$, which can only work in
-low-dimensional problems. If $d_i(\thetab)$ is not differentiable, the
-Bayesian Optimisation can be used as an alternative solution. In this
-scenario, the training part becomes slower due to fitting of the
-surrogate model and the blind optimisation steps. After obtaning the
-optimal points $\thetab^*$, the training part gets completed by
-estimating the proposal regions;
-Algorithm~\ref{alg:region_construction} describes the line search
-approach for finding the region's boundaries.
+proposal regions \(\mathcal{\hat{S}}_i\). The tasks are; (a) sampling
+the nuisance variables \(\vb_i \sim p(\vb)\) for obtaining the
+deterministic functions \(d_i(\thetab)\), (b) defining the optimisation
+problems \(\min_{\thetab} \: d_i(\thetab)\), (c) obtaining
+\(\thetab_i^*\), \(d_i^*\) (d) checking whether \(d_i^* \leq \epsilon\) and
+(e) computing the proposal distribution \(q_i\).
+
+If \(d_i(\thetab)\) is differentiable, using a gradient-based method
+is advised for obtaining \(\thetab_i^*\) faster. In this case, the
+gradients \(\nabla_{\thetab} d_i\) gradients are approximated
+automatically with finite-differences. This approximation requires two
+evaluations of \(d_i\) for \emph{each} parameter
+\(\theta_d, d \in \{1, \ldots, D\}\)\footnote{\(D\) is the
+  dimensionality of \(\thetab\), i.e.  \(\thetab \in \mathbb{R}^D\)},
+which works in low-dimensional problems. If \(d_i(\thetab)\) is not
+differentiable, Bayesian Optimisation can be used as an alternative
+solution. In this scenario, the training part becomes slower due to
+fitting of the surrogate model and the blind optimisation steps.
+
+After obtaining the optimal points \(\thetab^*\), we estimate the
+proposal regions. Algorithm~\ref{alg:region_construction} describes
+the line search approach for finding the region's boundaries. An
+important step for the proposal region estimation is deciding the axes
+of the bounding box; the directions \(\vb_d, d = \{1, \ldots, D\}\) we
+follow for reaching the boundaries. We approximate them as the
+direction of the highest curvature of \(d_i\) at \(\thetab_i^*\). This
+estimation is given by the eigenvalues of the Hessian matrix
+\(\hessian_i\) of \(d_i\)\footnote{Either the real distance \(d_i\) or
+  the Gaussian Process approximation \(\hat{d}_i\)} at
+\(\thetab_i\). The Hessian matrix is approximated numerically. In case
+where the distance function is the Euclidean, the Hessian matrix can
+be also computed as \(\hessian_i = \jac_i^T\jac_i\), where \(\jac_i\)
+is the Jacobian matrix of the summary statistics
+\(\Phi(g(\thetab, \vb=\vb_i))\) at \(\thetab_i^*\). The approximation
+through the Jacobian matrix has the computational advantage of using
+only first-order derivatives.
 
@@
 tasks; (a) sampling from the posterior
-$ \thetab_i \sim p_{d, \epsilon}(\thetab|\data)$, (b) computing an
-expectation $E_{\thetab|\data}[h(\thetab)]$ or (c) evaluating the
-unnormalised posterior $p_{d, \epsilon}(\thetab|\data)$. Sampling is
-performed by getting $n_2$ samples from each proposal distribution
-$q_i$. For each sample $\thetab_{ij}$, the distance
-function\footnote{As before, if a surrogate model $\hat{d}$ is
+\( \thetab_i \sim p_{d, \epsilon}(\thetab|\data)\), (b) computing an
+expectation \(E_{\thetab|\data}[h(\thetab)]\) and (c) evaluating the
+unnormalised posterior \(p_{d, \epsilon}(\thetab|\data)\). Sampling is
+performed by getting \(n_2\) samples from each proposal distribution
+\(q_i\). For each sample \(\thetab_{ij}\), the distance
+function\footnote{As before, if a surrogate model \(\hat{d}\) is
   available, it can be utilised as the distance function.} is
@@
 region. Algorithm~\ref{alg:sampling_GB} defines the steps for
-computing a weighted sample. Computing an expectation can be done
-easily after weighted samples are obtained using the
-equation~\ref{eq:expectation}, so we do not discuss it
-separately. Evaluating the unnormalised posterior requires the
-deterministic functions $g_i$ and the prior distribution
-$p(\thetab)$\footnote{There is no need for solving the optimisation
-  problems and building the proposal regions.}. The evaluation
-requires iterating over all $g_i$ and evaluating the distance from the
-observed data.
-
-% \begin{minipage}{0.46\textwidth}
-% \begin{algorithm}[H]
-%     \centering
-%     \caption{Training Part - Gradient-based. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GB}
-%     \begin{algorithmic}[1]
-%       \For{$i \gets 1 \textrm{ to } n$}
-%         \State Obtain $\thetab_i^*$ using a Gradient Optimiser
-%         \If{$g_i(\thetab_i^*) > \epsilon$}
-%         \State{go to} 1
-%         \Else
-%         \State Approximate $\jac_i^* = \nabla g_i(\theta)$ and $H_i \approx \jac^T_i\jac_i$
-%         \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
-%         \EndIf      
-%       \EndFor
-%       \Return{$q_i, p(\theta), g_i(\theta)$}
-%     \end{algorithmic}
-% \end{algorithm}
-% \end{minipage}
-% \hfill
-% \begin{minipage}{0.46\textwidth}
-% \begin{algorithm}[H]
-%     \centering
-%     \caption{Training Part - Bayesian optimisation. Requires $g_i(\thetab), p(\thetab)$}\label{alg:training_GP}
-%     \begin{algorithmic}[1]
-%       \For{$i \gets 1 \textrm{ to } n$}
-%         \State Obtain $\thetab_i^*, \hat{d}_i(\thetab)$ using a GP approach
-%         \If{$g_i(\thetab_i^*) > \epsilon$}
-%         \State{go to} 1
-%         \Else
-%         \State Approximate $H_i \approx \jac^T_i \jac_i$
-%         \State Use Algorithm~\ref{alg:region_construction} to obtain $q_i$
-%         \EndIf      
-%       \EndFor
-%       \Return{$q_i, p(\theta), \hat{d}_i(\theta)$}
-%     \end{algorithmic}
-% \end{algorithm}
-% \end{minipage}
+computing a weighted sample. Computing the expectation is easy after
+weighted samples are obtained using the equation~\ref{eq:expectation},
+so we do not discuss it separately. Evaluating the unnormalised
+posterior requires the deterministic functions \(g_i\) and the prior
+distribution \(p(\thetab)\)\footnote{There is no need for solving the
+  optimisation problems and building the proposal regions.}. The
+evaluation requires iterating over all \(g_i\) and evaluating the
+distance from the observed data.
 
 \begin{algorithm}[!ht]
-	\caption{Approximation of $\mathcal{\hat{S}}_i$ with a bounding box; Requires, a model of distance $d_i(\thetab)$, optimal point $\thetab_i^*$, number of refinements $K$, step size $\eta\_\text{start}$, maximum iterations $M$ and curvature matrix $\hessian_i$ ($\jac_i^T\jac_i $ or GP Hessian)}\label{alg:region_construction}
+	\caption{Approximation of \(\mathcal{\hat{S}}_i\) with a bounding box; 
+  Requires: a model of distance \(d_i(\thetab)\), 
+  an optimal point \(\thetab_i^*\), 
+  a number of refinements \(K\), 
+  a step size \(\eta\_\text{start}\), 
+  maximum iterations \(M\) and 
+  a curvature matrix \(\hessian_i\) (\(\jac_i^T\jac_i \) or GP Hessian)}\label{alg:region_construction}
 	\begin{algorithmic}[1]
-	\State Compute eigenvectors $\vb_{d}$ of $\hess_i$ {\scriptsize ($d = 1,\ldots,||\thetab ||)$}
-	\For{$d \gets 1 \textrm{ to } ||\thetab||$}
-		\State $\Tilde{\thetab} \gets \thetab_i^*$ \label{algstep:box_constr_start}
-		\State $k \gets 0$
-		\State $\eta \gets \eta\_\text{start}$ \Comment{Initialize $\eta$}
+	\State Compute eigenvectors \(\vb_{d}\) of \(\hess_i\) {\scriptsize (\(d = 1,\ldots,||\thetab ||)\)}
+	\For{\(d \gets 1 \textrm{ to } D\)}
+		\State \(\Tilde{\thetab} \gets \thetab_i^*\) \label{algstep:box_constr_start}
+		\State \(k \gets 0\)
+		\State \(\eta \gets \eta\_\text{start}\) \Comment{Initialize \(\eta\)}
 		\Repeat
-          \State $j \gets 0$
+          \State \(j \gets 0\)
         	\Repeat
-            \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}$ \Comment{Large step size $\eta$.}
-            \State $j \gets j + 1$
-        	\Until{$d(f_i(\Tilde{\thetab}), \data) > \epsilon$ or $j \geq M$} \Comment{Check distance or maximum iterations}
-        	\State $\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}$
-        	\State $\eta \gets \eta/2$ \Comment{More accurate region boundary}
-        	\State $k \gets k + 1$
-      \Until $k = K$
-      \If{$\Tilde{\thetab} = \thetab_i^*$} \Comment{Check if no step has been done}
-        \State $\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d$ \Comment{Then, make the minimum step}
-      \EndIf
-    	\State Set final $\Tilde{\thetab}$ as region end point. \label{algstep:box_constr_end}
-    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for $\mathbf{v}_{d} = - \mathbf{v}_{d}$
+            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}\) \Comment{Large step size \(\eta\).}
+            \State \(j \gets j + 1\)
+        	\Until{\(d(f_i(\Tilde{\thetab}), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
+        	\State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}\)
+        	\State \(\eta \gets \eta/2\) \Comment{More accurate region boundary}
+        	\State \(k \gets k + 1\)
+      \Until \(k = K\)
+      \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
+        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d\) \Comment{Then, make the minimum step}
+      \EndIf \label{algstep:box_constr_end}
+    	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\mathbf{v}_{d}\)
+    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
 	\EndFor
-	\State Fit a rectangular box around the region end points and define $q_i$ as uniform distribution
+	\State Fit a rectangular box around the region end points and define \(q_i\) as uniform distribution
 	\end{algorithmic}
@@
     \centering
-    \caption{Sampling. Requires a function of distance $d_i$, the prior distribution $p(\thetab)$, the proposal distribution $q_i$}\label{alg:sampling_GB}
+    \caption{Sampling. Requires a function of distance \(d_i\), the prior distribution \(p(\thetab)\), the proposal distribution \(q_i\)}\label{alg:sampling_GB}
     \begin{algorithmic}[1]
-      \State $\thetab_{ij} \sim q_i$
-          \If {$d_i(\thetab_{ij}) > \epsilon$}
+      \State \(\thetab_{ij} \sim q_i\)
+          \If {\(d_i(\thetab_{ij}) > \epsilon\)}
             \State Go to 2 \Comment{Reject sample}
           \Else {}
-            \State $w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}$ \Comment{Compute weight}
-            \State Store $(w_{ij}, \thetab_{ij})$ \Comment{Store weighted sample}
+            \State \(w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}\) \Comment{Compute weight}
+            \State Store \((w_{ij}, \thetab_{ij})\) \Comment{Store weighted sample}
           \EndIf

### 2022-02-07 latex_files/ch3/3-1-implementation-principles.tex
addressed henri's comments

@@
 The overview of our implementation is illustrated in
-Figure~\ref{fig:romc_overview}; one may interpret
-Figure~\ref{fig:romc_overview} as a depiction of the main class and
-the ellipses are the main routines of the class. Following
-\proglang{Python}'s naming principles, the methods starting with an
-underscore (green ellipses) represent internal (private) functions,
-whereas the rest (blue ellipses) are the public methods that the user
-interacts with. In Figure~\ref{fig:romc_overview}, it can be easily
-observed that the implementation follows consistently Algorithm~\ref{alg:romc_algorithm}. The training part includes
-all the steps until the computation of the proposal regions
-i.e.~sampling the nuisance variables, defining the optimisation
-problems, solving them, constructing the regions and fitting local
-surrogate models. The inference part comprises of evaluating the
-unnormalised posterior (and the normalised when is possible), sampling
-and computing an expectation. We also provide some utilities for
-inspecting the training process, such as plotting the histogram of the
-final distances or visualising the constructed bounding
-boxes. Finally, in the evaluation part, we provide two methods for
-evaluating the inference; (a) computing the Effective Sample Size
-(ESS) of the samples and (b) measuring the divergence between the
-approximate posterior the ground-truth, if the latter is
+Figure~\ref{fig:romc_overview}. Following \proglang{Python}'s naming
+principles, the methods starting with an underscore (green rectangles)
+represent internal (private) functions, whereas the rest (blue
+ellipses) are the methods exposed as the public API. In
+Figure~\ref{fig:romc_overview}, it can be easily observed that the
+implementation follows Algorithm~\ref{alg:romc_algorithm}. The
+training part includes all the steps until the computation of the
+proposal regions i.e.~sampling the nuisance variables, defining the
+optimisation problems, solving them, constructing the regions and
+fitting local surrogate models. The inference part comprises of
+evaluating the unnormalised posterior (and the normalised when is
+possible), sampling and computing an expectation. We also provide some
+utilities for inspecting the training process, such as plotting the
+histogram of the final distances or visualising the constructed
+bounding boxes. Finally, in the evaluation part, we provide two
+methods for evaluating the inference; (a) computing the Effective
+Sample Size (ESS) of the samples and (b) measuring the divergence
+between the approximate posterior the ground-truth, if the latter is
 available.\footnote{Normally, the ground-truth posterior is not
@@
 For illustrating the functionalities we will use the following running
-example introduced by~\cite{Ikonomov2019}, \textbf{HP comment: what is c in Eq. \eqref{eq:1D_example_eq_2}
-and in python snippet? As it is, it will not run. }
+example introduced by \citet{Ikonomov2019},
 
@@
 
-The prior is a uniform distribution in the range $[-2.5, 2.5]$ and the
-likelihood is defined at~\ref{eq:1D_example_eq_2}. There is only one
-observation $y_0 = 0$. The inference in this particular example can be
-performed quite easily without using a likelihood-free inference
-approach. We can exploit this fact for validating the accuracy of our
-implementation.
+The prior is a uniform distribution in the range \([-2.5, 2.5]\) and
+the likelihood is defined at~\ref{eq:1D_example_eq_2}. The constant
+\(c\) is \(0.5 - 0.5^4\) ensures the continuity of the pdf. There is
+only one observation \(y_0 = 0\). The inference in this particular
+example can be performed quite easily without using a likelihood-free
+inference approach. We can exploit this fact for validating the
+accuracy of our implementation.
 
@@
 ------------------------------ python snippet ------------------------------
-  import elfi import scipy.stats as ss
+  import elfi
+  import scipy.stats as ss
   import numpy as np
@@
   def simulator(t1, batch_size=1,random_state=None):
+    c = 0.5 - 0.5**4
     if t1 < -0.5:

### 2022-02-07 latex_files/ch4/4-1-training.tex
addressed henri's comments

@@
 gradient-based optimiser or Bayesian optimisation. These tasks are
-completed sequentially, as shown in Figure~\ref{fig:romc_overview}. The
-definition of the optimisation problems is performed by drawing
-\code{n1} integer numbers from a discrete uniform distribution
-$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
-used in \pkg{ELFI}'s random simulator. The seeds $u_i$ are
-computationally analogous to the random variables $\vb_i$, that were
-defined in the previous chapter; setting the \code{seed} eliminates all randomness of the simulator transforming it to a deterministic
-function. Setting the argument \code{use_bo=True}, chooses the
-Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
-case, apart from obtaining the optimal points $\thetab_i^*$, we also
-fit a GP as a surrogate model $\hat{d}_i$. In this
-scenario, in the following steps, $\hat{d}_i(\thetab)$ replaces
-$d_i(\thetab)$ as the distance function.
+completed sequentially, as shown in
+Figure~\ref{fig:romc_overview}. The definition of the optimisation
+problems is performed by drawing \code{n1} integer numbers from a
+discrete uniform distribution \(u_i \sim \mathcal{U}\{1,
+2^{32}-1\}\). Each integer \(u_i\) is the seed used in \pkg{ELFI}'s
+random simulator. The seeds \(u_i\) are computationally analogous to
+the random variables \(\vb_i\), that were defined in the previous
+chapter; setting the \code{seed} eliminates all randomness of the
+simulator transforming it to a deterministic function. Setting the
+argument \code{use_bo=True}, chooses the Bayesian Optimisation scheme
+for obtaining \(\thetab_i^*\). In this case, apart from obtaining the
+optimal points \(\thetab_i^*\), we also fit a GP as a surrogate model
+\(\hat{d}_i\). In this scenario, in the following steps,
+\(\hat{d}_i(\thetab)\) replaces \(d_i(\thetab)\) as the distance
+function.
 
@@
 \noindent
-This function helps the user decide which threshold $\epsilon$ to
+This function helps the user decide which threshold \(\epsilon\) to
 use. It plots a histogram of the distances at the optimal point
-$d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $\hat{d}_i^*$ in case
+\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in case
 \code{use_bo=True}. The function accepts all keyword arguments and
@@
 This method constructs a bounding box around the optimal point
-$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
+\(\thetab_i^* : i = 1, 2, \ldots, n_1\) following
 Algorithm~\ref{alg:region_construction}. The Hessian matrix is
 approximated based on the Jacobian
-$\hess_i = \jac_i^T \jac_i$\footnote{The Jacobian matrix is computed
+\(\hess_i = \jac_i^T \jac_i\)\footnote{The Jacobian matrix is computed
   on the output of the summary statistics
-  $\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))$ at $\thetab_i^*$} if
+  \(\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))\) at \(\thetab_i^*\)} if
 gradients have been used, or as the Hessian of the GP, in case of
@@
 one- or two-dimensional. The argument \code{i} is the index of the corresponding
-optimisation problem i.e.\ $i<n_1$. \code{compute_eps} returns the
-appropriate distance value $d_{i=\kappa}^*$ where
-$\kappa = \lfloor quantile \cdot n \rfloor$ from the collection
-$\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the number of
+optimisation problem i.e.\ \(i<n_1\). \code{compute_eps} returns the
+appropriate distance value \(d_{i=\kappa}^*\) where
+\(\kappa = \lfloor quantile \cdot n \rfloor\) from the collection
+\(\{ d_i^* \} \forall i = \{1, \ldots, n\}\) where \(n\) is the number of
 accepted solutions. It can be used to automate the selection of the
-threshold $\epsilon$, e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.
+threshold \(\epsilon\), e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.
 

### 2022-02-07 latex_files/ch5/5-1-ma2.tex
addressed henri's comments

@@
 \end{center}
-\caption{\textbf{HP comment: Needs a fine caption to get Table number.}\label{tab:ma2}}
+\caption{Comparison of the samples obtained from the estimated
+  posterior with (a) Rejection sampling and (b) the different versions
+  of ROMC. We observe that the obtained samples share similar
+  statistics along all methods. \label{tab:ma2}}
 \end{table}
-% \begin{figure}[H]
-%     \begin{center}
-%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist.png}
-%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/ma2_distance_hist_bo.png} \\
-%     \end{center}
-%     \caption[MA2 example, histogram of distances]{Histogram of distances
-%       \(d_i^*, i \in \{ 1, \ldots, n_1 \}\). The left graph corresponds
-%       to the gradient-based approach and the right one to the Bayesian
-%       optimisation approach.}
-%   \label{fig:ma2_2}
-% \end{figure}
 
@@
   \centering
-    \begin{minipage}[t]{.49\textwidth}
+    \begin{minipage}[t]{.4\textwidth}
       \begin{tikzpicture}
@@
     \end{minipage}\hfill
-    \begin{minipage}[t]{.49\textwidth}
+    \begin{minipage}[t]{.4\textwidth}
       \begin{tikzpicture}

### 2022-02-07 paper.tex
addressed henri's comments

@@
   efficiency. In this paper, we present the implementation of Robust
-  Optimisation Monte Carlo (ROMC) in the Python package ELFI. ROMC is
-  a novel and efficient (highly-parallelizable) LFI technique that
-  provides accurate weighted samples from the posterior.  Our
-  implementation can be used in two ways. A scientist may use it as an
-  out-of-the-box LFI algorithm. We provide an easy-to-use API
-  harmonised with the rest of the ELFI framework; therefore, it is
-  straightforward to compare it with all the methods included in
-  ELFI. On the other direction, we have carefully split ROMC into
-  isolated algorithmic steps. Therefore, we allow the researcher to
-  experiment, replacing components of the algorithm with
-  custom-designed subparts. In both scenarios, the algorithmic steps
+  Optimisation Monte Carlo (ROMC) in the \proglang{Python} package
+  \pkg{ELFI}. ROMC is a novel and efficient (highly-parallelizable)
+  LFI framework that provides accurate weighted samples from the
+  posterior. Our implementation can be used in two ways. Firstly, a
+  scientist may use it as an out-of-the-box LFI algorithm; we provide
+  an easy-to-use API harmonised with the rest of the \pkg{ELFI}
+  framework.  Therefore, it is straightforward to compare it with the
+  rest of the methods included in \pkg{ELFI}. Afterwards, we have
+  carefully split ROMC into isolated algorithmic steps. Hence, we
+  allow the researcher to experiment by replacing part(s) of the
+  algorithm with novel subcomponent(s). In both scenarios, all steps
   of ROMC run in a fully-parallelizable manner, exploiting all CPU
   cores. We also provide helpful functionalities for (i) inspecting
-  the inference process and (ii) evaluating the approximated
-  posterior. Finally, we test the robustness of our implementation on
-  some typical LFI examples.
-
-  % This short article illustrates how to write a manuscript for the
-  % \emph{Journal of Statistical Software} (JSS) using its {\LaTeX} style files.
-  % Generally, we ask to follow JSS's style guide and FAQs precisely. Also,
-  % it is recommended to keep the {\LaTeX} code as simple as possible,
-  % i.e., avoid inclusion of packages/commands that are not necessary.
-  % For outlining the typical structure of a JSS article some brief text snippets
-  % are employed that have been inspired by \cite{Zeileis+Kleiber+Jackman:2008},
-  % discussing count data regression in \proglang{R}. Editorial comments and
-  % instructions are marked by vertical bars.
-}
+  the inference process and (ii) evaluating the obtained
+  samples. Finally, we test the robustness of our implementation on
+  some typical LFI examples.  }
 
@@
 
-Optimization Monte Carlo (OMC) proposed by~\citep{Meeds2015} was a
+Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} was a
 novel LFI approach for approximating the posterior
-distribution. Afterwards, \citep{Forneron2016} provided a similar
-method under `reverse sampler'. The key idea is turning the stochastic
-data-generating mechanism into a set of fully-parallelizable
-deterministic optimisation processes. In their work,
-~\citet{Ikonomov2019}, located some critical failure modes of OMC, and
-proposed Robust OMC (ROMC) an alternative version of OMC with the
-appropriate improvements.
+distribution. Afterwards, \citet{Forneron2016} provided a similar
+method naming it `reverse sampler'. The central idea is turning the
+stochastic data-generating mechanism into a set of
+fully-parallelizable deterministic optimisation processes. In their
+work, ~\citet{Ikonomov2019}, located some critical failure modes of
+OMC, and proposed Robust OMC (ROMC) an alternative version of OMC with
+the appropriate improvements.
 
 In this paper, we present an extendible and parallelisable
-implementation of ROMC at the \proglang{Python} package
-\proglang{Engine for Likelihood-Free inference (ELFI)}. We follow
-specific principles for making the implementation extendible. As we
-will describe analytically in the next chapter, ROMC can be understood
-not only as a specific algorithm, but mainly as a framework for
-obtaining weighted samples from the posterior; it defines some
-algorithmic task that must be fulfilled, without enforcing which
-specific algorithm must be used for the tasks. Therefore, a researcher
-may adopt ROMC's backbone approach, but experiment with alternative
-methods to solve the distinct tasks. We have designed ROMC for
-facilitating such alterations. Finally, we have tested the accuracy and
-the efficiency of our implementation on the LFI examples supported by
-the \proglang{ELFI} package.
+implementation of ROMC at the \proglang{Python} package \pkg{Engine
+  for Likelihood-Free inference (ELFI)}. We choose carefully the
+designing principles for making the implementation extendible. As we
+describe analytically in the next chapter, ROMC can be understood not
+only as a specific algorithm, but mainly as a framework for obtaining
+weighted samples from the posterior; it defines a sequence of
+algorithmic steps, without enforcing a specific algorithm for solving
+them. Therefore, a researcher may adopt ROMC's backbone approach, but
+experiment with alternative methods to solve the tasks. We have
+designed ROMC for facilitating such alterations. Finally, we have
+tested the accuracy and the efficiency of our implementation on the
+LFI examples supported by the \pkg{ELFI} package.
 
@@
 \section{Background}
-We first give a short introduction to implicit parametric models and
-LFI. We then focus on the methods relevant for
-this paper: first the earlier OMC and then
-its recent robust generalisation, robust OMC, for which we provide the
-software package as part of ELFI, a larger inference platform that we briefly introduce at the
-end of this background section.
+We first give a short introduction to simulator-based models and
+LFI. We then focus on the methods forming the basis of this paper; OMC
+and its robust improvement, ROMC. Finally, we introduce ELFI, the
+software package used for our implementation.
 
@@
 
-\subsection{Implicit models and likelihood-free inference}
+\subsection{Simulator-based models and likelihood-free inference}
 \input{latex_files/ch2/2-1-implicit-models.tex}

### 2022-02-14 Appendix/1_3_notation.tex
reordered repo

In this section, we provide an overview of the symbols utilised in the
rest of the document. At this level, the quantities are introduced
quite informally; most of them will be defined formally in the next
chapters. We try to keep the notation as consistent as possible
throughout the document. The symbol $\R^N$, when used, describes that
a variable belongs to the $\text{N-dimensional}$ Euclidean space; $N$
does not represent a specific number. The bold formation ($\xb$)
indicates a vector, while ($x$) a scalar. Random variables are
represented with capital letters ($\Thetab$) while the samples with
lowercase letters ($\thetab$), i.e.\ $\thetab \sim \Thetab$.

\subsubsection*{Random Generator}
\label{sec:random-generator}
\begin{itemize}
\item $M_r(\thetab): \R^D \rightarrow \R$: The black-box data simulator
\end{itemize}

\subsubsection*{Parameters/Random Variables/Symbols}
\label{sec:variables}

\begin{itemize}
\item $D \in \mathbb{N}$, the dimensionality of the parameter-space
\item $\Thetab \in \R^D$, random variable representing the parameters of interest
\item $\data \in \R^N$, the vector with the observations
\item $\epsilon \in R$, the threshold setting the limit on the region
  around $\data$. When further notation is introduced
  regarding $\epsilon$\footnote{If
    $\epsilon_*, \: *:\{filter, region, cutoff\}$ values are not specified
    explicitly, they all share the common value of $\epsilon$}
  \begin{itemize}
  \item $\epsilon_{filter}$, threshold for discarding solutions
  \item $\epsilon_{region}$, threshold for building bounding box regions
  \item $\epsilon_{cutoff}$, threshold for the indicator function
  \end{itemize}

\item $\V \in \R^N$, random variable representing the randomness of
  the generator. It is also called nuisance variable, because we are
  not interested in inferring a posterior distribution on it.
\item $\vb_i \sim \V$, a specific sample drawn from $\V$
\item $\Y_{\thetab}$, random variable describing the simulator $M_r(\thetab)$. 
\item $\yb_i \sim \Y_\theta$, a sample drawn from $\Y_\theta$. It can
  be obtained by executing the simulator $\yb_i \sim M_r(\thetab)$
\end{itemize}

\subsubsection*{Sets}
\label{sec:sets}

\begin{itemize}
\item $\region(\data)$, the set of $\yb$ points close to the
  observations, i.e.\ $\yb := \{\yb: d(\yb, \data) \leq \epsilon \}$
\item $\regioni$, the set of points defined around $\yb_i$ i.e.\
  $\regioni = \region(\yb_i)$
\item $\mathcal{S}_i$, the set of parameters $\thetab$ that generate data close
  to the observations using the i-th deterministic generator, i.e.\
  $\{ \thetab: M_d(\thetab, \vb_i) \in \region(\data) \}$
\end{itemize}

\subsubsection*{Generic Functions}
\label{sec:generic-functions}

\begin{itemize}
\item $p(\cdot)$, any valid pdf
\item $p(\cdot | \cdot)$, any valid conditional distribution
\item $p(\thetab)$, the prior distribution on the parameters
\item $p(\vb)$, the prior distribution on the nuisance variables
\item $p(\thetab|\data)$, the posterior distribution
\item $p_{d,\epsilon}(\thetab|\data)$, the approximate posterior
  distribution  
\item $d(\mathbf{x}, \mathbf{y}): \R^{2N} \rightarrow \R$: any valid
  distance, the $L_2$ norm: $||\mathbf{x}-\mathbf{y}||_2$
\end{itemize}

\subsubsection*{Functions (Mappings)}
\label{sec:functions-mappings}

\begin{itemize}
\item $M_d(\thetab, \vb): \R^D \rightarrow \R$, the deterministic
  generator; all stochastic variables that are part of the data generation process are represented by the parameter $\vb$
\item $f_i(\thetab) = M_d(\thetab, \vb_i)$, deterministic generator associated with sample $\vb_i \sim p(\vb)$
\item $g_i(\thetab) = d(f_i(\thetab), \data)$, distance of the generated data $f_i(\thetab)$ from the observations
\item $T(\mathbf{x}): \mathbb{R}^{D_1} \rightarrow \mathbb{R}^{D_2}$
  where $D_1 > D_2$, the mapping that computes the summary statistic
\item $\indicator{\region(\data)}(\yb)$, the indicator function; returns 1 if $d(\yb, \data) \leq \epsilon$, else 0
\item $L(\thetab)$, the likelihood
\item $L_{d,\epsilon}(\thetab)$, the approximate likelihood
\end{itemize}

### 2022-02-14 article/latex_files/ch1/1_2_outline.tex
reordered repo

The remainder of the dissertation is organised as follows; in Chapter
2, we establish the mathematical formulation. Namely, we initially
describe the simulator-based models and provide some background
information on the fundamental algorithms proposed so far. Afterwards,
we provide the mathematical description of the ROMC approach
\autocite{Ikonomov2019}. Finally, we transform the mathematical
description to algorithms. In Chapter 3, we illustrate the
implementation part; we initially provide some information regarding
the Python package Engine for Likelihood-Free Inference (ELFI)
\autocite{1708.00707} and subsequently, we present the implementation
details of ROMC in this package. In general, the conceptual scheme
followed by the dissertation is

$$\text{Mathematical modelling} \rightarrow
\text{Algorithm} \rightarrow \text{Software}$$

In Chapter 4, we demonstrate the functionalities of the ROMC
implementation at some real-world examples; this chapter demonstrates
the accuracy of the ROMC method and our implementation's at
likelihood-free tasks. Finally, in Chapter 5, we conclude with some
thoughts on the work we have done and some future research ideas.

### 2022-02-14 article/latex_files/ch2/2-4-ELFI.tex
reordered repo

--- 2022-02-02 latex_files/ch2/2-4-ELFI.tex
@@
 likelihood-free inference methods, making it straightforward to
-perform comparisons between methods. The implementation was officially
-added to \pkg{ELFI} in version \pkg{v0.8.0}.
+perform comparisons between methods.
 
@@
 \pkg{ELFI} handles the probabilistic model as a Directed Acyclic Graph
-(DAG). This functionality is based on the package \pkg{NetworkX},
-which supports general-purpose graphs. In most cases, the structure of
-a likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
+(DAG). This functionality is based on the package
+\pkg{NetworkX}~\cite{hagberg2008exploring}, which supports
+general-purpose graphs. In most cases, the structure of a
+likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
 some edges connect the prior distributions to the simulator, the
@@
 class. Finally, the observations are passed to the appropriate node
-through the argument
-\code{observed}.
+through the argument \code{observed}.
 

### 2022-02-14 article/latex_files/ch3/3-0-design-principles.tex
reordered repo

--- 2022-02-07 latex_files/ch3/3-0-design-principles.tex
@@
     \EndFor
-    \State \(E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
+    \State \(\E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
     \State \(p_{d,\epsilon}(\thetab) \) as in eq.~\eqref{eq:approx_posterior_omc} \Comment{Evaluate the unnormalised posterior}
@@
 \( \thetab_i \sim p_{d, \epsilon}(\thetab|\data)\), (b) computing an
-expectation \(E_{\thetab|\data}[h(\thetab)]\) and (c) evaluating the
+expectation \(\E_{\thetab|\data}[h(\thetab)]\) and (c) evaluating the
 unnormalised posterior \(p_{d, \epsilon}(\thetab|\data)\). Sampling is
@@
     	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\mathbf{v}_{d}\)
-    	\State Repeat steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
+    	\State Runc steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
 	\EndFor

### 2022-02-14 article/latex_files/ch4/4-2-inference.tex
reordered repo

--- 2022-02-02 latex_files/ch4/4-2-inference.tex
@@
 This is the basic inference utility of the ROMC implementation; we
-draw $n_2$ samples for each bounding box region. This gives a total of
-$k \times n_2$, where $k < n_1$ is the number of the optimal points
-remained after filtering\footnote{From the $n_1$ optimisation
-  problems, only the ones with $d_i(\thetab_*) < \epsilon$ are
+draw \(n_2\) samples for each bounding box region. This gives a total of
+\(k \times n_2\), where \(k < n_1\) is the number of the optimal points
+remained after filtering\footnote{From the \(n_1\) optimisation
+  problems, only the ones with \(d_i(\thetab_*) < \epsilon\) are
   maintained for building a bounding box}. The samples are drawn from
-a uniform distribution $q_i$ defined over the corresponding bounding
-box and the weight $w_i$ is computed as in
+a uniform distribution \(q_i\) defined over the corresponding bounding
+box and the weight \(w_i\) is computed as in
 Algorithm~\ref{alg:sampling_GB}.
@@
 This function computes the expectation
-$E_{p(\thetab|\data)}[h(\thetab)]$ using
+\(\E_{p(\thetab|\data)}[h(\thetab)]\) using
 expression~\eqref{eq:expectation}. The argument \code{h} can be any
@@
 evaluates the normalised posterior estimating the partition function
-$Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab$; with Riemann's
+\(Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab\); with Riemann's
 integral approximation. The approximation is computationally feasible
@@
 both cases, the empirical result is close to the ground truth
-$\mu = 0$ and $\sigma^2 = 1$. The
+\(\mu = 0\) and \(\sigma^2 = 1\). The
 \code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
-point $\theta$ using expression \eqref{eq:approx_posterior_omc}. The
+point \(\theta\) using expression \eqref{eq:approx_posterior_omc}. The
 \code{romc.eval_posterior(theta)} approximates the partition function
-$Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab$ using the
+\(Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab\) using the
 Riemann approximation as explained above. In our simple example, this
@@
 %   \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples.
-(b) Right: Acceptance region around $\theta_1^*$ with the obtained samples plotted inside.}
+(b) Right: Acceptance region around \(\theta_1^*\) with the obtained samples plotted inside.}
 %   \label{fig:example_sampling}

### 2022-02-14 article/latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex
reordered repo

--- 2021-08-24 latex_files/images/chapter4/mae2_hist_t2_romc.tex
@@
 tick pos=left,
-title={ROMC (gradient-based) - Histogram of the parameter \(\displaystyle \theta_2\)},
+title={ROMC (Neural Network) - \(\displaystyle \theta_2\)},
 x grid style={white!69.0196078431373!black},
@@
 \draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
-\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00554428884192465);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0609051679887177);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.172932144280507);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.33874934889258);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.484235646321314);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.665208164482786);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.92463814505837);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.12139409849263);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.18629739177248);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.38911741112759);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.3586665597235);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.4168730451184);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.86440804043856);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.77193230571172);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.14800442101592);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.34373098482187);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.24893542723874);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.36527262608105);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.32011000423491);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.35112752816106);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.09237672296862);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.72225524667604);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65885907591305);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.4089028988961);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.18892677075576);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.11057927007035);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.808010795099994);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.746145310081448);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.68232985247479);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.586634734027065);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.400470451389202);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.259104715630191);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.230773630659991);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.211235556781049);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.129869421833243);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0903839207491964);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0434117816322698);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0482353129247447);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0289411877548468);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0144705938774233);
+\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0687105672242381);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.184372208227465);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.251601377874948);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.39271865402813);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.516086691042045);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.829805712987072);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.847041859829425);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.888673668222991);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.16327691119626);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.37486835291373);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.19991232691807);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.42636185782617);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.39151225358654);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.82926351678107);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.94594508515498);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.2129470534662);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.34655785188779);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.63158719290653);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.55685248411093);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.63080691718184);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.52614922878044);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.93355077757299);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.72788273613444);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.36820910319437);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.0258751741395);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,0.915348210439271);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.784575688430834);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.649930571025344);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.528011081166735);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.423788895495386);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.255909163140911);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.28970067094874);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.19412252617457);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.189105171805318);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.127165545733375);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0974061925124961);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0718579504601718);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0457277866564734);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0587928685583229);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0979881142638707);
 \draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
@@
 table {%
-0.0325607299281219 0
-0.0325607299281219 3
+0.0301562808023854 0
+0.0301562808023854 3
 };
-\addlegendentry{$\mu = 0.033$}
+\addlegendentry{$\mu = 0.030$}
 \addplot [semithick, black, dashed]
 table {%
--0.17677281465975 1
-0.198406420501618 1
+-0.180267449762181 1
+0.196611267527429 1
 };

### 2022-02-14 latex/article.tex
reordered repo

--- 2022-02-07 paper.tex
@@
 \usepackage{thumbpdf,lmodern}
-
-%% Vasilis packages
+\usepackage{orcidlink}
 \usepackage{amsmath,amssymb,bbold, xfrac}
@@
 
-%% - \author{} with primary affiliation
-%% - \Plainauthor{} without affiliations
-%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
-%% - \AND starts a new line, \And does not.
-\author{Vasilis Gkolemis\\ATHENA RC
-  \And Michael Gutmann\\University of Edinburgh
-   \And Henri Pesonen \\University of Oslo
+\author{Vasilis Gkolemis~\orcidlink{0000-0002-2636-0245}\\ATHENA RC \And
+  Michael Gutmann~\orcidlink{0000-0002-5329-9910}\\University of Edinburgh \And
+  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo
 }
-\Plainauthor{Vasilis Gkolemis, Michael Gutmann}
+\Plainauthor{Vasilis Gkolemis, Michael Gutmann, Henri Pesonen}
 
-%% - \title{} in title case
-%% - \Plaintitle{} without LaTeX markup (if any)
-%% - \Shorttitle{} with LaTeX markup (if any), used as running title
 \title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
 \Plaintitle{An Extendable Python Implementation of ROMC}
-\Shorttitle{An Extendable \proglang{Python} Implementation of ROMC}
+\Shorttitle{An Extendable ROMC implementation}
 
@@
   samples. Finally, we test the robustness of our implementation on
-  some typical LFI examples.  }
+  some typical LFI examples.}
 
-%% - \Keywords{} with LaTeX markup, at least one required
-%% - \Plainkeywords{} without LaTeX markup (if necessary)
-%% - Should be comma-separated and in sentence case.
-\Keywords{Bayesian Inference, Simulator-based models, Likelihood-free, \proglang{Python}, \proglang{ELFI}}
-\Plainkeywords{Bayesian Inference, Simulator-based models, Likelihood-free, Python, ELFI}
+%% - \Keywords{}
+\Keywords{Bayesian inference, implicit models, likelihood-free, \proglang{Python}, \pkg{ELFI}}
+\Plainkeywords{Bayesian inference, implicit models, likelihood-free, Python, ELFI}
 
 %% - \Address{} of at least one author
-%% - May contain multiple affiliations for each author
-%%   (in extra lines, separated by \emph{and}\\).
-%% - May contain multiple authors for the same affiliation
-%%   (in the same first line, separated by comma).
 \Address{
@@
 \begin{document}
-
-%% See jss_template_article.tex for formatting instructions and macros to e.g. typset code
-%% I only kept the informative sidebars as comments
 
@@
 LFI examples supported by the \pkg{ELFI} package.
+
+To the best of our knowledge, this is the first attempt to implement
+the ROMC inference method to a generic LFI framework. Therefore, there
+are not other implementation directly comparable to ours. For this
+reason, we test the implementation against (i) an artificial example
+with tractable likelihood and (ii) the second-order moving average
+(MA2) example from the \pkg{ELFI} package. In the latter, we generate
+data using known parameters and we compare our results with the
+typical Rejection ABC approach.
 
@@
 
-\subsection{Engine for likelihood-free inference (ELFI)}
+\subsection[Engine for likelihood-free inference (ELFI)]{\pkg{Engine for likelihood-free inference (ELFI)}}
 \label{subsec:ELFI}

### 2022-02-14 latex/latex_files/ch2/2-1-implicit-models.tex
reordered repo

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample data points, but we cannot evaluate the
likelihood. Formally, a simulator-based model is a parameterised
family of probability density functions
\(\{ p(\yb|\thetab) \}_{\thetab}\) whose closed-form is either unknown
or computationally intractable. Practically, a simulator
\( M_r(\thetab) \) can be understood as a random black-box machine
(computer code) that generates samples \(\yb\) in a stochastic manner
for any given a set of parameters \(\thetab\), i.e.
\( M_r(\thetab) \rightarrow \yb \). We can isolate the randomness by
introducing the stochastic nuisance variables \(\vb \sim p(\vb)\) so
that the simulator becomes a deterministic mapping \(\simulator\) that
maps \((\thetab, \vb)\) to the data \(\yb\), i.e.\
\(\yb=\simulator(\thetab,\vb)\).  In terms of computer code,
\(p(\vb)\) is defined as a random number generating process within the
computer code.  % Importantly, we can fix the value of \(\vb\) while
% changing \(\thetab\) by fixing the random seed used.

Simulator-based models provide considerable modelling freedom; any
physical process that can be conceptualised as a computer program of
finite steps can be modelled as a simulator-based model without any
compromise. The modelling freedom allows for any amount of hidden
(unobserved) internal variables or rule-based decisions. Hence,
implicit models are often used to model physical phenomena in the
natural sciences such as e.g.~genetics, epidemiology, or neuroscience. 
Further background on simulator-based models and
example applications can be found in the articles
by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.

The modelling freedom of simulator-based models, however, comes at the
price of difficulties in inferring their parameters. Denoting observed
data by \(\data\), the main difficulty is that the likelihood function
\(L(\thetab) = p(\data|\thetab)\) is generally intractable. To better
see the sources of the intractability, and to address them, we go back
to the basic characterisation of the likelihood as the (rescaled)
probability of a parameter of the model to generate data \(\yb\) that
is similar to the observed data \(\data\). More formally, the
likelihood \(L(\thetab)\) equals
\begin{equation} \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data)  \mid \theta)
\end{equation}
where \(c_\epsilon\) is a proportionality factor that depends on
\(\epsilon\) and \(\region(\data)\) is an \(\epsilon\) region around \(\data\)
that is defined via a distance function \(d\), i.e.\ \(\region(\data)
:= \{\yb: d(\yb, \data) \leq \epsilon \}\).

The basic characterisation of the likelihood in \eqref{eq:likelihood}
highlights two sources of intractability: The first is the computation
of the probability
\(\Pr(\simulator(\thetab,\vb) \in \region(\data))\), the second is the
limit of \(\epsilon \to 0\).  Approximating the probability with
samples becomes computationally infeasible if \(\epsilon\) is too
small. Hence, a large class of inference methods work with
\(\epsilon > 0\), which leads to the approximate likelihood function
\(L_{d, \epsilon}(\thetab)\),
\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
\end{equation}

and, in turn, to the approximate posterior,

\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
\end{equation}

The approximation in \eqref{eq:approx_likelihood} is by no means the
only strategy to deal with the intractabilities of the likelihood
function in \eqref{eq:likelihood}. Other strategies include modelling
the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
reverse, or framing likelihood-free inference as a ratio estimation
problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
  Thomas2016, Papamakarios2019, Chen2019, Thomas2020,
  Hermans2020}. However, both OMC and robust OMC, which we introduce
next, are based on the approximation in \eqref{eq:approx_likelihood}.

### 2022-02-14 latex/latex_files/ch4/4-4-extensibility.tex
reordered repo

--- 2021-10-05 latex_files/ch4/4-4-extensibility.tex
@@
 \emph{explicitly} enforcing the methods that solve these steps. Even
-though for each step a specific algotihm is proposed
-by~\cite{Ikonomov2019}, in general ROMC works well if a practitioner
+though for each step a specific algorithm is proposed
+by~\cite{Ikonomov2019}, in general ROMC is functional if a practitioner
 thinks of an alternative way of approaching a specific
@@
 We have specified four specific points where a user may intervene with
-their custom modules; (a) the gradient-based optimization, (b) the
-Bayesian Optimization, (c) the proposal region construction and (d)
+their custom modules; (a) the gradient-based optimisation, (b) the
+Bayesian Optimisation, (c) the proposal region construction and (d)
 the surrogate model fitting. These are the four critical parts of the
 ROMC procedure, whereas the rest of the code is the backbone of the
-algorithm; each of them may be approached with a completely different
-algorithm than the one we propose, without the rest of the method to
+algorithm; each of them can be swapped to a different
+algorithm with the same functionality than the one we propose, without the rest of the method to
 collapse.

### 2022-02-14 latex/latex_files/images/chapter3/example_posterior.tex
reordered repo

--- 2021-08-27 latex_files/images/chapter3/example_post.tex
@@
 legend cell align={left},
-legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
+legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black, font=\tiny},
 tick align=outside,

### 2022-02-14 replication_script/images/chapter4/mae2_hist_t1_romc.tex
reordered repo

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.2891017928294);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.25501979350038);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.61356484847951);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.8947593445762);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,1.9604837129404);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.31923292279405);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.65081439660326);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.75410489838711);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.07726897262135);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.3599490125042);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.53754575168527);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.48128983274092);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.28421573662553);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.20415347732589);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.93220527414237);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.67182230140431);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.33879169620136);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.15217777578996);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.90617769806791);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.79028400808985);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.81288260257434);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.65005170222198);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.55300367553848);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.54492012441559);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.3456563752174);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.24429041156909);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.01082437940841);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.08307433945245);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.909291082406055);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.915784352973493);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.63897857298652);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.662358192832588);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.574938584618925);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.539445093672632);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.362963228305201);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.335571290114039);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.303417920200607);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.110219084548637);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.15297623409581);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.116660395859453);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.105235905447663);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.0815817607102305);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0822347290062214);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0267815192076974);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0305618619739539);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.503211246652872 0
0.503211246652872 3.5
};
\addlegendentry{$\mu = 0.503$}
\addplot [semithick, black, dashed]
table {%
0.320723275794794 1
0.636341466841524 1
};
\addlegendentry{$\sigma = 0.143$}
\end{axis}

\end{tikzpicture}

### 2022-02-14 article/latex_files/python_scripts/example_MA2.py

@@ -1,0 +1,174 @@
+import logging
+import matplotlib.pyplot as plt
+from elfi.examples import ma2
+import elfi
+import numpy as np
+import scipy.stats
+import matplotlib
+import os
+import tikzplotlib as tplt
+matplotlib.rcParams['text.usetex'] = True
+
+logging.basicConfig(level=logging.INFO)
+prepath = './../images/chapter4/'
+savefig = False
+
+def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
+    plt.figure()
+    plt.title(title)
+    plt.hist(x=samples,
+             weights=weights,
+             bins=bins, density=True, range=range)
+    plt.axvline(mean, 0, 1,
+                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
+    plt.axhline(1,
+                (mean-std-range[0])/(range[1] - range[0]),
+                (mean+std-range[0])/(range[1] - range[0]),
+                color="k",
+                linestyle="--", label=r"$\sigma = %.3f$" % (std))
+    plt.xlabel(xlabel)
+    plt.ylabel(ylabel)
+    plt.ylim(ylim)
+    plt.legend()
+    if savepath:
+        plt.savefig(savepath, bbox_inches='tight')
+        # tplt.clean_figure()
+        tplt.save(savepath[:-3] + "tex")
+    plt.show(block=False)
+
+# Set seed for reproducibility
+seed = 1
+np.random.seed(seed)
+model = ma2.get_model(seed_obs=seed)
+
+# plot prior samples
+
+x = model.generate(1000, outputs=["t1", "t2"])
+plt.figure()
+plt.title("Samples from the prior")
+plt.xlabel(r"$\theta_1$")
+plt.ylabel(r"$\theta_2$")
+plt.plot(x["t1"], x["t2"], "bo")
+if savefig:
+    plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
+                bbox_inches="tight")
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "ma2_prior_samples.tex"))
+plt.show(block=False)
+
+####### ROMC with gradients ################
+n1 = 100
+n2 = 200
+bounds = [(-2, 2), (-1.25, 1.25)]
+eps = .01
+vis_ind_1 = 1
+vis_ind_2 = 3
+vis_ind_3 = 12
+
[…]

### 2022-02-14 article/latex_files/python_scripts/example_MA2_bo.py

@@ -1,0 +1,225 @@
+import logging
+import matplotlib.pyplot as plt
+from elfi.examples import ma2
+import elfi
+import numpy as np
+import scipy.stats
+import matplotlib
+import os
+import tikzplotlib as tplt
+matplotlib.rcParams['text.usetex'] = True
+
+logging.basicConfig(level=logging.INFO)
+prepath = './../images/chapter4/'
+savefig = False
+
+def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
+    plt.figure()
+    plt.title(title)
+    plt.hist(x=samples,
+             weights=weights,
+             bins=bins, density=True, range=range)
+    plt.axvline(mean, 0, 1,
+                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
+    plt.axhline(1,
+                (mean-std-range[0])/(range[1] - range[0]),
+                (mean+std-range[0])/(range[1] - range[0]),
+                color="k",
+                linestyle="--", label=r"$\sigma = %.3f$" % (std))
+    plt.xlabel(xlabel)
+    plt.ylabel(ylabel)
+    plt.ylim(ylim)
+    plt.legend()
+    if savepath:
+        plt.savefig(savepath, bbox_inches='tight')
+        tplt.save(savepath[:-3] + "tex")        
+    plt.show(block=False)
+
+
+####### ROMC with BO ###########
+# Set seed for reproducibility
+seed = 1
+np.random.seed(seed)
+model = ma2.get_model(seed_obs=seed)
+
+# plot prior samples
+
+x = model.generate(1000, outputs=["t1", "t2"])
+plt.figure()
+plt.title("Samples from the prior")
+plt.xlabel(r"$\theta_1$")
+plt.ylabel(r"$\theta_2$")
+plt.plot(x["t1"], x["t2"], "bo")
+if savefig:
+    plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
+                bbox_inches="tight")
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "ma2_prior_samples.tex"))
+plt.show(block=False)
+
+n1 = 100
+n2 = 200
+bounds = [(-2, 2), (-1.25, 1.25)]
+eps = .03
+vis_ind_1 = 1
+vis_ind_2 = 3
+vis_ind_3 = 12
+
+use_bo = True
[…]

### 2022-02-14 latex/latex_files/python_scripts/example_extensibility.py

@@ -1,0 +1,222 @@
+import logging
+import matplotlib.pyplot as plt
+from elfi.examples import ma2
+import elfi
+import numpy as np
+import scipy.stats
+import matplotlib
+import os
+import tikzplotlib as tplt
+matplotlib.rcParams['text.usetex'] = True
+
+logging.basicConfig(level=logging.INFO)
+prepath = './../images/chapter4/'
+savefig = False
+
+def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
+    plt.figure()
+    plt.title(title)
+    plt.hist(x=samples,
+             weights=weights,
+             bins=bins, density=True, range=range)
+    plt.axvline(mean, 0, 1,
+                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
+    plt.axhline(1,
+                (mean-std-range[0])/(range[1] - range[0]),
+                (mean+std-range[0])/(range[1] - range[0]),
+                color="k",
+                linestyle="--", label=r"$\sigma = %.3f$" % (std))
+    plt.xlabel(xlabel)
+    plt.ylabel(ylabel)
+    plt.ylim(ylim)
+    plt.legend()
+    if savepath:
+        plt.savefig(savepath, bbox_inches='tight')
+        # tplt.clean_figure()
+        tplt.save(savepath[:-3] + "tex")
+    plt.show(block=False)
+
+# Set seed for reproducibility
+seed = 1
+np.random.seed(seed)
+model = ma2.get_model(seed_obs=seed)
+
+# plot prior samples
+
+x = model.generate(1000, outputs=["t1", "t2"])
+plt.figure()
+plt.title("Samples from the prior")
+plt.xlabel(r"$\theta_1$")
+plt.ylabel(r"$\theta_2$")
+plt.plot(x["t1"], x["t2"], "bo")
+if savefig:
+    plt.savefig(os.path.join(prepath, "mae2_prior_samples_nn.png"),
+                bbox_inches="tight")
+    tplt.clean_figure()
+    tplt.save(os.path.join(prepath, "ma2_prior_samples_nn.tex"))
+plt.show(block=False)
+
+
+####### ROMC with gradients ################
+from elfi.methods.inference.romc import OptimisationProblem, NDimBoundingBox
+# from elfi.methods.utils import NDimBoundingBox
+from sklearn.pipeline import Pipeline
[…]

### 2022-02-14 replication_script/replication_script.py

@@ -1,0 +1,1132 @@
+# Importing all
+from elfi.methods.inference.romc import OptimisationProblem
+from sklearn.pipeline import Pipeline
+from sklearn.neural_network import MLPRegressor
+from functools import partial
+
+import os
+import timeit
+import scipy.stats as ss
+import scipy.integrate as integrate
+import logging
+import matplotlib.pyplot as plt
+from elfi.examples import ma2
+import numpy as np
+import elfi
+import tikzplotlib as tplt
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+logging.basicConfig(level=logging.INFO)
+prepath = './images/'
+savefig = True
+
+np.random.seed(21)
+
+
+######################################################################
+# This part generates the figures used in the implementation example.#
+######################################################################
+
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    r"""Implements the distribution
+    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
+                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
+                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
+    """
+
+    def rvs(self, theta, seed=None):
+        """Vectorised sampling from likelihood.
+
+        """
+        assert isinstance(theta, np.ndarray)
+        theta = theta.astype(np.float)
+        samples = np.empty_like(theta)
+
+        c = 0.5 - 0.5 ** 4
+
+        tmp_theta = theta[theta <= -0.5]
+        samples[theta <= -
+                0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta <= -0.5] = np.inf
+
[…]

### 2022-02-14 replication_script/separate_scripts/example_1D.py

@@ -1,0 +1,320 @@
+""" This script generates the figures used in the implementation example.
+"""
+
+import os
+import scipy.stats as ss
+import scipy.integrate as integrate
+
+import matplotlib.pyplot as plt
+import timeit
+import numpy as np
+import elfi
+import tikzplotlib as tplt
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+
+prepath = './../images/'
+savefig = False
+
+np.random.seed(21)
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    r"""Implements the distribution
+    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
+                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
+                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
+    """
+
+    def rvs(self, theta, seed=None):
+        """Vectorised sampling from likelihood.
+
+        """
+        assert isinstance(theta, np.ndarray)
+        theta = theta.astype(np.float)
+        samples = np.empty_like(theta)
+
+        c = 0.5 - 0.5 ** 4
+
+        tmp_theta = theta[theta <= -0.5]
+        samples[theta <= -
+                0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta <= -0.5] = np.inf
+
+        tmp_theta = theta[theta <= 0.5]
+        samples[theta <= 0.5] = ss.norm(
+            loc=tmp_theta**4, scale=1).rvs(random_state=seed)
+        theta[theta <= 0.5] = np.inf
+
+        tmp_theta = theta[theta < np.inf]
+        samples[theta < np.inf] = ss.norm(
+            loc=tmp_theta - c, scale=1).rvs(random_state=seed)
+        theta[theta < np.inf] = np.inf
+
[…]

### 2022-02-14 replication_script/separate_scripts/execution_times_parallel.py

@@ -1,0 +1,319 @@
+"""Testing the execution times using parallelisation."""
+
+import timeit
+import time
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy.integrate as integrate
+import scipy.ndimage as ndimage
+import scipy.stats as ss
+import os
+import elfi
+import matplotlib
+matplotlib.rcParams['text.usetex'] = True
+
+np.random.seed(21)
+prepath = './../images/chapter4' 
+savefig = True
+
+class Prior:
+    def rvs(self, size=None, random_state=None):
+        # size from (BS,) -> (BS,1)
+        if size is not None:
+            size = np.concatenate((size, [1]))
+        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+    def pdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+    def logpdf(self, theta):
+        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+class Likelihood:
+    def rvs(self, th1, th2, seed=None):
+        assert isinstance(th1, np.ndarray)
+        assert isinstance(th2, np.ndarray)
+        assert th1.ndim == 2
+        assert th2.ndim == 2
+        assert np.allclose(th1.shape, th2.shape)
+
+        x = []
+        for i in range(th1.shape[0]):
+            cur_th = np.concatenate((th1[i], th2[i]))
+            x.append(ss.multivariate_normal(
+                mean=cur_th, cov=1).rvs(random_state=seed))
+        return np.array(x)
+
+    def pdf(self, x, th1, th2):
+        assert isinstance(th1, float)
+        assert isinstance(th2, float)
+        assert isinstance(x, np.ndarray)
+
+        th = np.stack((th1, th2))
+        rv = ss.multivariate_normal(mean=th, cov=1)
+        nof_points = x.shape[0]
+        prod = 1
+        for i in range(nof_points):
+            prod *= rv.pdf(x[i])
+        return prod
+
+
+def create_factor(x):
+    lik = Likelihood()
+    pr = Prior()
+
+    def tmp_func(th1, th2):
[…]

### 2022-02-15 article/latex_files/ch5/5-1-ma2.tex
reordered files

--- 2022-02-07 latex_files/ch5/5-1-ma2.tex
@@
 \begin{figure}[ht]
-  \centering
-    \begin{minipage}[t]{.4\textwidth}
-      \begin{tikzpicture}
-        \begin{axis}[
-          title={Solve optimisation problems},
-          xlabel={\(n_1\)},
-          ylabel={seconds},
-          xmin=1, xmax=500,
-          ymin=0, ymax=7,
-          xtick={1,100,200,300,400,500},
-          ytick={0, 1, 2, 3, 4, 5, 6, 7},
-          legend pos=north west,
-          ymajorgrids=true,
-          grid style=dashed,
-          ]
-          \addplot[color=blue, mark=o] coordinates {
-            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-          };
-          \addlegendentry{sequential}
-          \addplot[color=red, mark=o] coordinates {
-            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-          };
-          \addlegendentry{parallel}          
-        \end{axis}
-      \end{tikzpicture}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{.4\textwidth}
-      \begin{tikzpicture}
-        \begin{axis}[
-          title={Proposal regions},
-          xlabel={\(n_1\)},
-          ylabel={seconds},
-          xmin=1, xmax=500,
-          ymin=0, ymax=7,
-          xtick={1,100,200,300,400,500},
-          ytick={0, 1, 2, 3, 4, 5, 6, 7},
-          legend pos=north west,
-          ymajorgrids=true,
-          grid style=dashed,
-          ]
-          \addplot[color=blue, mark=o] coordinates {
-            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-          };
-          \addlegendentry{sequential}
-          \addplot[color=red, mark=o] coordinates {
-            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-          };
-          \addlegendentry{parallel}          
-        \end{axis}
-      \end{tikzpicture}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{.4\textwidth}
-      \begin{tikzpicture}
-        \begin{axis}[
-          title={Sample},
-          xlabel={\(n_1\)},
-          ylabel={seconds},
-          xmin=1, xmax=500,
-          ymin=0, ymax=7,
-          xtick={1,100,200,300,400,500},
-          ytick={0, 1, 2, 3, 4, 5, 6, 7},
-          legend pos=north west,
-          ymajorgrids=true,
-          grid style=dashed,
-          ]
-          \addplot[color=blue, mark=o] coordinates {
-            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-          };
-          \addlegendentry{sequential}
-          \addplot[color=red, mark=o] coordinates {
-            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-          };
-          \addlegendentry{parallel}          
-        \end{axis}
-      \end{tikzpicture}
-    \end{minipage}\hfill
-    \begin{minipage}[t]{.4\textwidth}
-      \begin{tikzpicture}
-        \begin{axis}[
-          title={Evaluate posterior},
-          xlabel={\(n_1\)},
-          ylabel={seconds},
-          xmin=1, xmax=500,
-          ymin=0, ymax=7,
-          xtick={1,100,200,300,400,500},
-          ytick={0, 1, 2, 3, 4, 5, 6, 7},
-          legend pos=north west,
-          ymajorgrids=true,
-          grid style=dashed,
-          ]
-          \addplot[color=blue, mark=o] coordinates {
-            (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-          };
-          \addlegendentry{sequential}
-          \addplot[color=red, mark=o] coordinates {
-            (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-          };
-          \addlegendentry{parallel}          
-        \end{axis}
-      \end{tikzpicture}
-    \end{minipage}\hfill
+  \begin{center}
+    \resizebox{.47\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/exec_time_solve.tex}
+    }
+    \resizebox{.49\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/exec_time_regions.tex}
+    }
+    \end{center}
     \caption[Execution time exploiting parallelisation]{In the first
@@
 \end{figure}
+
+% \begin{figure}[ht]
+%   \centering
+%     \begin{minipage}[t]{.4\textwidth}
+%       \begin{tikzpicture}
+%         \begin{axis}[
+%           title={Solve optimisation problems},
+%           xlabel={\(n_1\)},
+%           ylabel={seconds},
+%           xmin=1, xmax=500,
+%           ymin=0, ymax=7,
+%           xtick={1,100,200,300,400,500},
+%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
+%           legend pos=north west,
+%           ymajorgrids=true,
+%           grid style=dashed,
+%           ]
+%           \addplot[color=blue, mark=o] coordinates {
+%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+%           };
+%           \addlegendentry{sequential}
+%           \addplot[color=red, mark=o] coordinates {
+%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+%           };
+%           \addlegendentry{parallel}          
+%         \end{axis}
+%       \end{tikzpicture}
+%     \end{minipage}\hfill
+%     \begin{minipage}[t]{.4\textwidth}
+%       \begin{tikzpicture}
+%         \begin{axis}[
+%           title={Proposal regions},
+%           xlabel={\(n_1\)},
+%           ylabel={seconds},
+%           xmin=1, xmax=500,
+%           ymin=0, ymax=7,
+%           xtick={1,100,200,300,400,500},
+%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
+%           legend pos=north west,
+%           ymajorgrids=true,
+%           grid style=dashed,
+%           ]
+%           \addplot[color=blue, mark=o] coordinates {
+%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+%           };
+%           \addlegendentry{sequential}
+%           \addplot[color=red, mark=o] coordinates {
+%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+%           };
+%           \addlegendentry{parallel}          
+%         \end{axis}
+%       \end{tikzpicture}
+%     \end{minipage}\hfill
+%     \begin{minipage}[t]{.4\textwidth}
+%       \begin{tikzpicture}
+%         \begin{axis}[
+%           title={Sample},
+%           xlabel={\(n_1\)},
+%           ylabel={seconds},
+%           xmin=1, xmax=500,
+%           ymin=0, ymax=7,
+%           xtick={1,100,200,300,400,500},
+%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
+%           legend pos=north west,
+%           ymajorgrids=true,
+%           grid style=dashed,
+%           ]
+%           \addplot[color=blue, mark=o] coordinates {
+%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+%           };
+%           \addlegendentry{sequential}
+%           \addplot[color=red, mark=o] coordinates {
+%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+%           };
+%           \addlegendentry{parallel}          
+%         \end{axis}
+%       \end{tikzpicture}
+%     \end{minipage}\hfill
+%     \begin{minipage}[t]{.4\textwidth}
+%       \begin{tikzpicture}
+%         \begin{axis}[
+%           title={Evaluate posterior},
+%           xlabel={\(n_1\)},
+%           ylabel={seconds},
+%           xmin=1, xmax=500,
+%           ymin=0, ymax=7,
+%           xtick={1,100,200,300,400,500},
+%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
+%           legend pos=north west,
+%           ymajorgrids=true,
+%           grid style=dashed,
+%           ]
+%           \addplot[color=blue, mark=o] coordinates {
+%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
+%           };
+%           \addlegendentry{sequential}
+%           \addplot[color=red, mark=o] coordinates {
+%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
+%           };
+%           \addlegendentry{parallel}          
+%         \end{axis}
+%       \end{tikzpicture}
+%     \end{minipage}\hfill
+%     \caption[Execution time exploiting parallelisation]{In the first
+%       line, we compare the parallel and sequential execution of the
+%       training part and at the second line the inference part. At the
+%       left figure, we measure the execution time for sampling \(n_2=50\)
+%       points per region. At the right figure we measure the execution
+%       time for evaluating the posterior at a batch of \(50\) points.}
+%       \label{fig:exec_parallel}
+% \end{figure}

### 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex
reordered files

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (BO) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,0.310209700650407);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,0.366870797321359);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.4774021490772);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,0.847063787984617);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,0.950467021534753);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,1.63746647929829);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,0.63033836109258);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.01425623641141);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,1.15727348863503);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,1.43066001219801);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,2.26780484039086);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,1.57088642283557);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,2.26780484039086);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,1.90093404306951);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,1.86235774750728);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.72575321812577);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.01425623641141);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,1.84427294639855);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,1.78345203790796);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,3.1720295175941);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,2.35296949050992);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,3.68663899199996);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.71687198144523);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,2.49494901962233);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,2.27822359273028);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.42715377524836);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.27700840481905);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.4878209014166);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,1.99251141603071);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,1.75945430650476);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,1.79803060206703);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,1.78570494893043);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,1.72087801094253);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,1.75945430650479);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,1.51407154384225);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,1.00938102922818);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,1.1776112007662);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.672920686152122);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,1.00938102922817);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.841150857690152);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.504690514614091);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.168230171538028);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.607177584483422 0
0.607177584483422 3.5
};
\addlegendentry{$\mu = 0.607$}
\addplot [semithick, black, dashed]
table {%
0.427777567120938 1
0.758013118742592 1
};
\addlegendentry{$\sigma = 0.150$}
\end{axis}

\end{tikzpicture}

### 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex
reordered files

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (Neural Network) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.4847013386292);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.63008300744943);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.78269414852667);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,2.17914288457748);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,2.26619550399379);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.63578624756494);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.81808795446661);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,3.02972033457779);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,2.97861443343929);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.5112192843295);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.38441750293694);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.60452518165752);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.40496357685519);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.39267433638072);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.60955375534895);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.29229691711042);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.33810216344099);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.33307056573176);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.16876578921053);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.94786358717711);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.6348432484266);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.41781765326629);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.42854946707319);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.34821580131638);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.16914162978256);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.02253080695157);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.06697582017199);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,1.05900414126017);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.834884752208823);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.771804085452666);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.651009521206343);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.552051125172065);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.283364267425877);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.321951539666027);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.31247072638644);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.18976114249095);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.255316789610236);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.110079997689329);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.0946414093799529);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.104932268497323);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.0466365637765878);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.0582957047207339);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0233182818882939);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0582957047207339);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0.0349774228324409);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0.0233182818882936);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.491475759372051 0
0.491475759372051 3.5
};
\addlegendentry{$\mu = 0.491$}
\addplot [semithick, black, dashed]
table {%
0.313768529672519 1
0.617478140945993 1
};
\addlegendentry{$\sigma = 0.138$}
\end{axis}

\end{tikzpicture}

### 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t2_romc.tex
reordered files

--- 2021-09-23 latex_files/images/chapter4/mae2_hist_t2_romc.tex
@@
 \draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
-\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.00551481774368941);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0605814218347914);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.168699423699822);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.340262186728983);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.48172967019454);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.661723210849365);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.926167683238105);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.1131441320768);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.19038558853121);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.39560413532974);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.38890799766854);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.45486223069012);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.89014334843844);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.80399178478282);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.15561058973456);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.33061276076321);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.24024442959031);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.34890403251458);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.31095726097247);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.33900171768764);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.07598627244674);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.71987564844275);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65165022964706);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.40256222683377);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.18758980571923);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.0980208933351);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.803715750920935);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.742179116696122);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.678702875114039);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.583516431505742);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.398341719580722);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.257727424376927);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.229546935490747);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.210112717761985);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.129179090828181);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0899034779939157);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0431810229330879);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0479789143700981);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0287873486220588);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0143936743110293);
+\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0143830031521635);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0640495269335328);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.167967992034369);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.322416836713075);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.450460589062093);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.627682829295711);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.78449673556291);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.13618558747044);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.21557183831947);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.29143077400903);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.46122694206085);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.58740403775278);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.71596854606341);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.97590595810952);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.26932406129648);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.32343959220477);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.1892115171771);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.38168476397323);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.31304797611767);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.21128903529658);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.14821419761865);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.81725742398109);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.53374220971548);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.44383714928785);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.24606025648798);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.06408208570229);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.796495552018255);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.705157030158244);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.731994856866652);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.533344891350605);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.478183009636268);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.277899805953886);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.198597889919886);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.151665334936985);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.131771535511397);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0942116661455654);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0673572489817432);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0288673924207473);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0336786244908719);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0144336962103735);
 \draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
@@
 table {%
-0.0318388478807404 0
-0.0318388478807404 3
+0.0325240508501527 0
+0.0325240508501527 3
 };
-\addlegendentry{$\mu = 0.032$}
+\addlegendentry{$\mu = 0.033$}
 \addplot [semithick, black, dashed]
 table {%
--0.177435597378694 1
-0.197481062716323 1
+-0.175528279454964 1
+0.1970811913253 1
 };
-\addlegendentry{$\sigma = 0.170$}
+\addlegendentry{$\sigma = 0.169$}
 \end{axis}

### 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex
reordered files

--- 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t2_romc.tex
@@
 tick pos=left,
-title={ROMC (gradient-based) - \(\displaystyle \theta_2\)},
+title={ROMC (BO) - \(\displaystyle \theta_2\)},
 x grid style={white!69.0196078431373!black},
@@
 \draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
-\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0.0143830031521635);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0640495269335328);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.167967992034369);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.322416836713075);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.450460589062093);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.627682829295711);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.78449673556291);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.13618558747044);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.21557183831947);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.29143077400903);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.46122694206085);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.58740403775278);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.71596854606341);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.97590595810952);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,2.26932406129648);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.32343959220477);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.1892115171771);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.38168476397323);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.31304797611767);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.21128903529658);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.14821419761865);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.81725742398109);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.53374220971548);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.44383714928785);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.24606025648798);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.06408208570229);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.796495552018255);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.705157030158244);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.731994856866652);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.533344891350605);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.478183009636268);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.277899805953886);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.198597889919886);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.151665334936985);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.131771535511397);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0942116661455654);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.0673572489817432);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0288673924207473);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0336786244908719);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0144336962103735);
+\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.0617545870506399);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.494036696405118);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.308772935253199);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.864564218708957);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.11158256691152);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.802809631658317);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.49395771225155);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.71306067268783);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.17926096548025);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.06759941441516);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.5396445345721);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,1.91609586839403);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,1.79258669429275);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,1.59403798219366);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,1.58667684276459);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,1.52979906798699);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15183142210051);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.06111444803853);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.47292129320938);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.85529643854941);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.86505006309548);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,2.67369166423286);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,1.96552066170352);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,1.29148734508544);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,1.68849292724457);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,1.00470598608042);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.703294190256297);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.703294190256294);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,1.2056471832965);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.200941197216084);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.100470598608042);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0);
 \draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
@@
 table {%
-0.0325240508501527 0
-0.0325240508501527 3
+0.100012161271839 0
+0.100012161271839 3
 };
-\addlegendentry{$\mu = 0.033$}
+\addlegendentry{$\mu = 0.100$}
 \addplot [semithick, black, dashed]
 table {%
--0.175528279454964 1
-0.1970811913253 1
+-0.108452084584719 1
+0.278478839382765 1
 };
-\addlegendentry{$\sigma = 0.169$}
+\addlegendentry{$\sigma = 0.176$}
 \end{axis}

### 2022-02-15 latex/latex_files/python_scripts/example_extensibility.py

@@ -1,222 +1,0 @@
-import logging
-import matplotlib.pyplot as plt
-from elfi.examples import ma2
-import elfi
-import numpy as np
-import scipy.stats
-import matplotlib
-import os
-import tikzplotlib as tplt
-matplotlib.rcParams['text.usetex'] = True
-
-logging.basicConfig(level=logging.INFO)
-prepath = './../images/chapter4/'
-savefig = False
-
-def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
-    plt.figure()
-    plt.title(title)
-    plt.hist(x=samples,
-             weights=weights,
-             bins=bins, density=True, range=range)
-    plt.axvline(mean, 0, 1,
-                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
-    plt.axhline(1,
-                (mean-std-range[0])/(range[1] - range[0]),
-                (mean+std-range[0])/(range[1] - range[0]),
-                color="k",
-                linestyle="--", label=r"$\sigma = %.3f$" % (std))
-    plt.xlabel(xlabel)
-    plt.ylabel(ylabel)
-    plt.ylim(ylim)
-    plt.legend()
-    if savepath:
-        plt.savefig(savepath, bbox_inches='tight')
-        # tplt.clean_figure()
-        tplt.save(savepath[:-3] + "tex")
-    plt.show(block=False)
-
-# Set seed for reproducibility
-seed = 1
-np.random.seed(seed)
-model = ma2.get_model(seed_obs=seed)
-
-# plot prior samples
-
-x = model.generate(1000, outputs=["t1", "t2"])
-plt.figure()
-plt.title("Samples from the prior")
-plt.xlabel(r"$\theta_1$")
-plt.ylabel(r"$\theta_2$")
-plt.plot(x["t1"], x["t2"], "bo")
-if savefig:
-    plt.savefig(os.path.join(prepath, "mae2_prior_samples_nn.png"),
-                bbox_inches="tight")
-    tplt.clean_figure()
-    tplt.save(os.path.join(prepath, "ma2_prior_samples_nn.tex"))
-plt.show(block=False)
-
-
-####### ROMC with gradients ################
-from elfi.methods.inference.romc import OptimisationProblem, NDimBoundingBox
-# from elfi.methods.utils import NDimBoundingBox
-from sklearn.pipeline import Pipeline
[…]

### 2022-02-15 latex/latex_files/python_scripts/execution_times_parallel.py

@@ -1,319 +1,0 @@
-"""Testing the execution times using parallelisation."""
-
-import timeit
-import time
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy.integrate as integrate
-import scipy.ndimage as ndimage
-import scipy.stats as ss
-import os
-import elfi
-import matplotlib
-matplotlib.rcParams['text.usetex'] = True
-
-np.random.seed(21)
-prepath = './../images/chapter4' 
-savefig = True
-
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
-
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
-
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
-
-
-class Likelihood:
-    def rvs(self, th1, th2, seed=None):
-        assert isinstance(th1, np.ndarray)
-        assert isinstance(th2, np.ndarray)
-        assert th1.ndim == 2
-        assert th2.ndim == 2
-        assert np.allclose(th1.shape, th2.shape)
-
-        x = []
-        for i in range(th1.shape[0]):
-            cur_th = np.concatenate((th1[i], th2[i]))
-            x.append(ss.multivariate_normal(
-                mean=cur_th, cov=1).rvs(random_state=seed))
-        return np.array(x)
-
-    def pdf(self, x, th1, th2):
-        assert isinstance(th1, float)
-        assert isinstance(th2, float)
-        assert isinstance(x, np.ndarray)
-
-        th = np.stack((th1, th2))
-        rv = ss.multivariate_normal(mean=th, cov=1)
-        nof_points = x.shape[0]
-        prod = 1
-        for i in range(nof_points):
-            prod *= rv.pdf(x[i])
-        return prod
-
-
-def create_factor(x):
-    lik = Likelihood()
-    pr = Prior()
-
-    def tmp_func(th1, th2):
[…]

### 2022-02-15 latex_files/python_scripts/example_1D.py

@@ -1,320 +1,0 @@
-""" This script generates the figures used in the implementation example.
-"""
-
-import os
-import scipy.stats as ss
-import scipy.integrate as integrate
-
-import matplotlib.pyplot as plt
-import timeit
-import numpy as np
-import elfi
-import tikzplotlib as tplt
-import matplotlib
-matplotlib.rcParams['text.usetex'] = True
-
-prepath = './../images/'
-savefig = False
-
-np.random.seed(21)
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
-
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
-
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
-
-
-class Likelihood:
-    r"""Implements the distribution
-    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
-                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
-                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
-    """
-
-    def rvs(self, theta, seed=None):
-        """Vectorised sampling from likelihood.
-
-        """
-        assert isinstance(theta, np.ndarray)
-        theta = theta.astype(np.float)
-        samples = np.empty_like(theta)
-
-        c = 0.5 - 0.5 ** 4
-
-        tmp_theta = theta[theta <= -0.5]
-        samples[theta <= -
-                0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
-        theta[theta <= -0.5] = np.inf
-
-        tmp_theta = theta[theta <= 0.5]
-        samples[theta <= 0.5] = ss.norm(
-            loc=tmp_theta**4, scale=1).rvs(random_state=seed)
-        theta[theta <= 0.5] = np.inf
-
-        tmp_theta = theta[theta < np.inf]
-        samples[theta < np.inf] = ss.norm(
-            loc=tmp_theta - c, scale=1).rvs(random_state=seed)
-        theta[theta < np.inf] = np.inf
-
[…]

### 2022-02-15 latex_files/python_scripts/example_MA2_bo.py

@@ -1,225 +1,0 @@
-import logging
-import matplotlib.pyplot as plt
-from elfi.examples import ma2
-import elfi
-import numpy as np
-import scipy.stats
-import matplotlib
-import os
-import tikzplotlib as tplt
-matplotlib.rcParams['text.usetex'] = True
-
-logging.basicConfig(level=logging.INFO)
-prepath = './../images/chapter4/'
-savefig = False
-
-def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
-    plt.figure()
-    plt.title(title)
-    plt.hist(x=samples,
-             weights=weights,
-             bins=bins, density=True, range=range)
-    plt.axvline(mean, 0, 1,
-                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
-    plt.axhline(1,
-                (mean-std-range[0])/(range[1] - range[0]),
-                (mean+std-range[0])/(range[1] - range[0]),
-                color="k",
-                linestyle="--", label=r"$\sigma = %.3f$" % (std))
-    plt.xlabel(xlabel)
-    plt.ylabel(ylabel)
-    plt.ylim(ylim)
-    plt.legend()
-    if savepath:
-        plt.savefig(savepath, bbox_inches='tight')
-        tplt.save(savepath[:-3] + "tex")        
-    plt.show(block=False)
-
-
-####### ROMC with BO ###########
-# Set seed for reproducibility
-seed = 1
-np.random.seed(seed)
-model = ma2.get_model(seed_obs=seed)
-
-# plot prior samples
-
-x = model.generate(1000, outputs=["t1", "t2"])
-plt.figure()
-plt.title("Samples from the prior")
-plt.xlabel(r"$\theta_1$")
-plt.ylabel(r"$\theta_2$")
-plt.plot(x["t1"], x["t2"], "bo")
-if savefig:
-    plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
-                bbox_inches="tight")
-    tplt.clean_figure()
-    tplt.save(os.path.join(prepath, "ma2_prior_samples.tex"))
-plt.show(block=False)
-
-n1 = 100
-n2 = 200
-bounds = [(-2, 2), (-1.25, 1.25)]
-eps = .03
-vis_ind_1 = 1
-vis_ind_2 = 3
-vis_ind_3 = 12
-
-use_bo = True
[…]

### 2022-02-15 replication_script/replication_script.py

@@ -1,4 +1,3 @@
-# Importing all
 from elfi.methods.inference.romc import OptimisationProblem
 from sklearn.pipeline import Pipeline
 from sklearn.neural_network import MLPRegressor
@@ -6,6 +5,7 @@ from functools import partial
 
 import os
 import timeit
+import time
 import scipy.stats as ss
 import scipy.integrate as integrate
 import logging
@@ -17,16 +17,12 @@ import tikzplotlib as tplt
 import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 logging.basicConfig(level=logging.INFO)
-prepath = './images/'
 savefig = True
 
-np.random.seed(21)
-
-
 ######################################################################
-# This part generates the figures used in the implementation example.#
+# Part 1. Generates the subfigures used in the running example in the#
+# presentation of our implementation i.e. figure 3                   #
 ######################################################################
-
 
 class Prior:
     def rvs(self, size=None, random_state=None):
@@ -148,7 +144,8 @@ def create_gt_posterior(likelihood, prior, data, Z):
         return likelihood.pdf(data, np.array([theta])) * prior.pdf(np.array([theta])) / Z
     return tmp_func
 
-
+prepath = './images/'
+np.random.seed(21)
 data = np.array([0.])
 dim = data.shape[0]
 a = -2.5  # integration left limit
@@ -162,7 +159,7 @@ Z = approximate_Z(factor, a, b)
 gt_posterior_pdf = create_gt_posterior(likelihood, prior, data, Z)
 
 
-############# PLOT GROUND TRUTH ##################
+# plot ground-truth prior, likelihood, posterior
 plt.figure()
 plt.title("Ground-truth PDFs")
 plt.xlim(-3, 3)
@@ -192,56 +189,51 @@ if savefig:
 plt.show(block=False)
 
 
-############# DEFINE ELFI MODEL ##################
+# define elfi model
 def simulator(theta, dim, batch_size=10000, random_state=None):
     likelihood = Likelihood()
     theta = np.repeat(theta, dim, -1)
     return likelihood.rvs(theta, seed=random_state)
-
 
 elfi.new_model("1D_example")
 elfi_prior = elfi.Prior(Prior(), name="theta")
[…]

### 2022-02-15 replication_script/separate_scripts/execution_times_parallel.py

@@ -252,7 +252,7 @@ for i, n in enumerate(n1):
     solve_grad_parallel.append(toc-tic)
 
     tic = timeit.default_timer()
-    romc1.estimate_regions(eps_filter=eps, use_surrogate=False)
+    romc1.estimate_regions(eps_filter=eps, use_surrogate=False, fit_models=False)
     toc = timeit.default_timer()
     estimate_regions_parallel.append(toc-tic)
 

### 2022-02-15 article/article.tex
final version

--- 2022-02-14 latex/article.tex
@@
 %% -- Introduction -------------------------------------------------------------
-
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
-% \section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
 \section{Introduction} \label{sec:intro}
-
-Simulator-based models are particularly captivating due to the
-modelling freedom they provide. In essence, a simulator-based model can
-describe any data generating mechanism that can be written as a finite
-set of algorithmic steps. This modelling freedom comes at a cost;
-performing the inference, i.e., sample or evaluate the posterior
-distribution, is challenging.
-
-Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} was a
-novel LFI approach for approximating the posterior
-distribution. Afterwards, \citet{Forneron2016} provided a similar
-method naming it `reverse sampler'. The central idea is turning the
-stochastic data-generating mechanism into a set of
-fully-parallelizable deterministic optimisation processes. In their
-work, ~\citet{Ikonomov2019}, located some critical failure modes of
-OMC, and proposed Robust OMC (ROMC) an alternative version of OMC with
-the appropriate improvements.
-
-In this paper, we present an extendible and parallelisable
-implementation of ROMC at the \proglang{Python} package \pkg{Engine
-  for Likelihood-Free inference (ELFI)}. We choose carefully the
-designing principles for making the implementation extendible. As we
-describe analytically in the next chapter, ROMC can be understood not
-only as a specific algorithm, but mainly as a framework for obtaining
-weighted samples from the posterior; it defines a sequence of
-algorithmic steps, without enforcing a specific algorithm for solving
-them. Therefore, a researcher may adopt ROMC's backbone approach, but
-experiment with alternative methods to solve the tasks. We have
-designed ROMC for facilitating such alterations. Finally, we have
-tested the accuracy and the efficiency of our implementation on the
-LFI examples supported by the \pkg{ELFI} package.
-
-To the best of our knowledge, this is the first attempt to implement
-the ROMC inference method to a generic LFI framework. Therefore, there
-are not other implementation directly comparable to ours. For this
-reason, we test the implementation against (i) an artificial example
-with tractable likelihood and (ii) the second-order moving average
-(MA2) example from the \pkg{ELFI} package. In the latter, we generate
-data using known parameters and we compare our results with the
-typical Rejection ABC approach.
-
-% \begin{leftbar}
-% The introduction is in principle ``as usual''. However, it should usually embed
-% both the implemented \emph{methods} and the \emph{software} into the respective
-% relevant literature. For the latter both competing and complementary software
-% should be discussed (within the same software environment and beyond), bringing
-% out relative (dis)advantages. All software mentioned should be properly
-% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-% \textsc{Bib}{\TeX}.)
-
-% For writing about software JSS requires authors to use the markup
-% \verb|\proglang{}| (programming languages and large programmable systems),
-% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-% arguments, etc.). If there is such markup in (sub)section titles (as above), a
-% plain text version has to be provided in the {\LaTeX} command as well. Below we
-% also illustrate how abbrevations should be introduced and citation commands can
-% be employed. See the {\LaTeX} code for more details.
-% \end{leftbar}
-
-%% -- Manuscript ---------------------------------------------------------------
-
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
-
-% \begin{leftbar}
-% As the synopsis above is a code listing that is not meant to be executed,
-% one can use either the dedicated \verb|{Code}| environment or a simple
-% \verb|{verbatim}| environment for this. Again, spaces before and after should be
-% avoided.
-
-% Finally, there might be a reference to a \verb|{table}| such as
-% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-% column headers and captions in sentence style, and if possible avoiding vertical
-% lines.
-% \end{leftbar}
-
-%% -- Illustrations ------------------------------------------------------------
-
-% - Virtually all JSS manuscripts list source code along with the generated
-%   output. The style files provide dedicated environments for this.
-% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%   or knitr using the render_sweave() hook - are used (without the need to
-%   load Sweave.sty).
-% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-% - The code input should use "the usual" command prompt in the respective
-%   software system.
-% - For R code, the prompt "R> " should be used with "+  " as the
-%   continuation prompt.
-% - Comments within the code chunks should be avoided - these should be made
-%   within the regular LaTeX text.
-
-% \begin{leftbar}
-% For code input and output, the style files provide dedicated environments.
-% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-% hook. Please make sure that all code is properly spaced, e.g., using
-% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-% use ``the usual'' command prompt in the respective software system. For
-% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-% the continuation prompt. Generally, comments within the code chunks should be
-% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-% before and after code input/output should be avoided (see above).
-% \end{leftbar}
+\input{latex_files/ch1/1-introduction.tex}
 
@@
 \section{Background}
-We first give a short introduction to simulator-based models and
-LFI. We then focus on the methods forming the basis of this paper; OMC
-and its robust improvement, ROMC. Finally, we introduce ELFI, the
-software package used for our implementation.
-
-%\input{latex_files/ch2/2_1_implicit_models}
+\input{latex_files/ch2/2-0-intro.tex}
 
@@
 \section{Design principles}  % better title needed
-
 \input{latex_files/ch3/3-0-design-principles.tex}
@@
 
-%% -- ROMC implementation ----------------------------------------------------
+%% -- ROMC implementation --
 \section{Implemented functionalities}
-
-In this section, we present the implementation. At~\ref{subsec:ELFI}
-we introduce the reader to the \proglang{Python} package \pkg{ELFI}. At~\ref{subsec:general_design} we explain the general
-design that we follow. Then, we divide the implementation into three
-parts; at~\ref{subsec:training} we present the training (fitting)
-part, at~\ref{subsec:inference} the inference part and
-at~\ref{subsec:evaluation} the evaluation part. Finally,
-at~\ref{subsec:extensibility} we describe how a user may extend ROMC
-with its custom modules.
+\input{latex_files/ch4/4-0-intro.tex}
 
@@
 \section{Use-case illustration}
-
-In this section, we test the implementation using the second-order
-moving average (MA2) example, which is one of the standard models of
-\pkg{ELFI}. We perform the inference using three different versions of
-ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
-Optimisation scheme and (iii) fitting a Neural Network as a surrogate
-model. The later illustrates how to extend the implementation,
-replacing part of ROMC with a user-defined component. Finally, we
-measure the execution speed-up achieved by the parallelised version of
-ROMC.
-
-% \subsection{Moving average example}
-% \label{subsec:ma2}
-\input{latex_files/ch5/5-1-ma2.tex}
-
-% \subsection{Execution time}
-% \label{subsec:exec_time}
-% \input{latex_files/ch5/5-2-execution-time.tex}
+\input{latex_files/ch5/5-ma2.tex}
 
@@
 \section{Summary and discussion} \label{sec:summary}
-
-In this paper, we presented the implementation details we followed for
-developing the LFI method ROMC at the \pkg{ELFI} package. We paid
-thorough attention to two specific use-case scenarios. Firstly, we
-illustrate how a user may take advantage of our ready-to-use API for
-solving its LFI problem. Secondly, we focus on the scenario where a
-researcher wants to intervene and alter parts of the method. Our
-implementation is designed to support this as well.
-
-There are still open challenges for the left for future research. Two
-directions may enable ROMC to solve high-dimensional problems
-efficiently. The first one is enabling ROMC's execution into a cluster
-of computers. ROMC can be characterized as an \textit{embarrassingly
-  parallel} workload; each optimization problem is an entirely
-independent task. Therefore, supporting inference into a cluster of
-computers can radically speed up the inference. The second one refers
-to implementing the method in a framework that supports automatic
-differentiation. Automatic differentiation is necessary for
-efficiently solving optimisation problems, especially in
-high-dimensional parametric models.
+\input{latex_files/ch6/6-summary.tex}
 
@@
 
-% \section*{Computational details}
+\section*{Computational details}
 
-% \begin{leftbar}
-% If necessary or useful, information about certain computational details
-% such as version numbers, operating systems, or compilers could be included
-% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-% maps, tables, \dots) that are not cited in the main text can be credited here.
-% \end{leftbar}
-
-% The results in this paper were obtained using
-% \proglang{R}~3.4.1 with the
-% \pkg{MASS}~7.3.47 package. \proglang{R} itself
-% and all packages used are available from the Comprehensive
-% \proglang{R} Archive Network (CRAN) at
-% \url{https://CRAN.R-project.org/}.
+The results in this paper were obtained using \proglang{Python}~3.7.1
+with the \pkg{ELFI}~0.8.2 package. The experiments have been executed
+in a single machine with an Intel® Core™ i7-8750H Processor an with
+Ubuntu 20.04 lts operating system.
 
@@
 
-% \begin{leftbar}
-% All acknowledgments (note the AE spelling) should be collected in this
-% unnumbered section before the references. It may contain the usual information
-% about funding and feedback from colleagues/reviewers/etc. Furthermore,
-% information such as relative contributions of the authors may be added here
-% (if any).
-% \end{leftbar}
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
+\clearpage
 \bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-% \newpage
-
-% \begin{appendix}
-
-% \section{More technical details} \label{app:technical}
-
-% \begin{leftbar}
-% Appendices can be included after the bibliography (with a page break). Each
-% section within the appendix should have a proper section title (rather than
-% just \emph{Appendix}).
-
-% For more technical style details, please check out JSS's style FAQ at
-% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-% which includes the following topics:
-% \begin{itemize}
-%   \item Title vs.\ sentence case.
-%   \item Graphics formatting.
-%   \item Naming conventions.
-%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
-%   \item Trouble shooting.
-%   \item Many other potentially helpful details\dots
-% \end{itemize}
-% \end{leftbar}
-
-% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-% \begin{leftbar}
-% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-% formats of author-year citations and allow to include additional details (e.g.,
-% pages, chapters, \dots) in brackets. In case you are not familiar with these
-% commands see the JSS style FAQ for details.
-
-% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-% when acquiring the entries automatically from mixed online sources. However,
-% it is important that informations are complete and presented in a consistent
-% style to avoid confusions. JSS requires the following format.
-% \begin{itemize}
-%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-%     be used in the references.
-%   \item Titles should be in title case.
-%   \item Journal titles should not be abbreviated and in title case.
-%   \item DOIs should be included where available.
-%   \item Software should be properly cited as well. For \proglang{R} packages
-%     \code{citation("pkgname")} typically provides a good starting point.
-% \end{itemize}
-% \end{leftbar}
-
-% \end{appendix}
 

### 2022-02-15 article/latex_files/ch1/1-introduction.tex
final version

--- 2022-02-14 latex/article.tex
@@
-\documentclass[article]{jss}
-%% -- LaTeX packages and custom commands ---------------------------------------
-
-\usepackage[usenames,dvipsnames]{xcolor}
-
-%% recommended packages
-\usepackage{thumbpdf,lmodern}
-\usepackage{orcidlink}
-\usepackage{amsmath,amssymb,bbold, xfrac}
-
-\usepackage{multirow}
-\usepackage{caption}
-\usepackage{subcaption} % for subfigure
-\usepackage{float}
-
-%% TikZ packages
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
-\usepackage{pgfplots}
-\pgfplotsset{width=6.75cm, compat=newest}
-\usepackage[utf8]{inputenc}
-\DeclareUnicodeCharacter{2212}{−}
-\usepgfplotslibrary{groupplots,dateplot}
-\usetikzlibrary{patterns,shapes.arrows}
-
-%% for algorithms
-\usepackage{algorithm}
-\usepackage{algpseudocode} % algorithmicx package
-
-%% another package (only for this demo article)
-\usepackage{framed}
-\usepackage{marginnote}
-
-%% new custom commands
-\newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\code{#1()}}
-
-%% Vasilis' custom commands
-\newcommand{\bigO}{\mathcal{O}}
-\newcommand{\X}{\mathbf{x}}
-\newcommand{\Z}{\mathbf{z}}
-\newcommand{\V}{\mathbf{V}}
-\newcommand{\Y}{\mathbf{Y}}
-\newcommand{\hess}{\mathbf{H}}
-\newcommand{\Thetab}{\mathbf{\Theta}}
-
-\newcommand{\vb}{\mathbf{v}}
-\newcommand{\yb}{\mathbf{y}}
-\newcommand{\xb}{\mathbf{x}}
-
-\newcommand{\jac}{\mathbf{J}}
-\newcommand{\hessian}{\mathbf{H}}
-\newcommand{\thetab}{\boldsymbol{\theta}}
-\newcommand{\thetabij}{\thetab_{ij}}
-\newcommand{\thetabi}{\thetab_i}
-
-\newcommand{\simulator}{g}
-\newcommand{\region}{B_{d,\epsilon}}
-\newcommand{\indicator}[1]{\mathbb{1}_{#1}}
-\newcommand{\regioni}{B_{d,\epsilon}^i}
-\newcommand{\data}{\mathbf{y_0}}
-\newcommand{\accregioni}{C^i_{\epsilon}}
-\newcommand{\accregionihat}{\hat{C}^i_{\epsilon}}
-
-\newcommand{\danger}{\fontencoding{U}\fontfamily{futs}\selectfont\char 66\relax}
-\newcommand{\TODO}[1]{\textcolor{red}{TODO #1}\marginpar{\textcolor{red}{\Large \danger}}}
-
-\newcommand{\R}{\mathbb{R}}
-
-\newcommand{\Ex}{\mathbb{E}}
-
-\DeclareMathOperator*{\argmax}{argmax}
-\DeclareMathOperator*{\argmin}{argmin}
-
-%% -- Article metainformation (author, title, ...) -----------------------------
-
-\author{Vasilis Gkolemis~\orcidlink{0000-0002-2636-0245}\\ATHENA RC \And
-  Michael Gutmann~\orcidlink{0000-0002-5329-9910}\\University of Edinburgh \And
-  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo
-}
-\Plainauthor{Vasilis Gkolemis, Michael Gutmann, Henri Pesonen}
-
-\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
-\Plaintitle{An Extendable Python Implementation of ROMC}
-\Shorttitle{An Extendable ROMC implementation}
-
-%% - \Abstract{} almost as usual
-\Abstract{ Performing the inference in statistical models with
-  intractable likelihood is challenging; most likelihood-free
-  inference (LFI) methods encounter problems in accuracy and
-  efficiency. In this paper, we present the implementation of Robust
-  Optimisation Monte Carlo (ROMC) in the \proglang{Python} package
-  \pkg{ELFI}. ROMC is a novel and efficient (highly-parallelizable)
-  LFI framework that provides accurate weighted samples from the
-  posterior. Our implementation can be used in two ways. Firstly, a
-  scientist may use it as an out-of-the-box LFI algorithm; we provide
-  an easy-to-use API harmonised with the rest of the \pkg{ELFI}
-  framework.  Therefore, it is straightforward to compare it with the
-  rest of the methods included in \pkg{ELFI}. Afterwards, we have
-  carefully split ROMC into isolated algorithmic steps. Hence, we
-  allow the researcher to experiment by replacing part(s) of the
-  algorithm with novel subcomponent(s). In both scenarios, all steps
-  of ROMC run in a fully-parallelizable manner, exploiting all CPU
-  cores. We also provide helpful functionalities for (i) inspecting
-  the inference process and (ii) evaluating the obtained
-  samples. Finally, we test the robustness of our implementation on
-  some typical LFI examples.}
-
-%% - \Keywords{}
-\Keywords{Bayesian inference, implicit models, likelihood-free, \proglang{Python}, \pkg{ELFI}}
-\Plainkeywords{Bayesian inference, implicit models, likelihood-free, Python, ELFI}
-
-%% - \Address{} of at least one author
-\Address{
-  Vasilis Gkolemis\\
-  Information Management Systems Institute (IMSI)\\
-  ATHENA Research and Innovation Center\\
-  Athens, Greece\\
-  E-mail: \email{vgkolemis@athenarc.gr}\\
-  URL: \url{https://givasile.github.io}
-}
-
-\begin{document}
-
-%% -- Introduction -------------------------------------------------------------
-
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
-% \section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
-\section{Introduction} \label{sec:intro}
-
 Simulator-based models are particularly captivating due to the
@@
 typical Rejection ABC approach.
-
-% \begin{leftbar}
-% The introduction is in principle ``as usual''. However, it should usually embed
-% both the implemented \emph{methods} and the \emph{software} into the respective
-% relevant literature. For the latter both competing and complementary software
-% should be discussed (within the same software environment and beyond), bringing
-% out relative (dis)advantages. All software mentioned should be properly
-% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-% \textsc{Bib}{\TeX}.)
-
-% For writing about software JSS requires authors to use the markup
-% \verb|\proglang{}| (programming languages and large programmable systems),
-% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-% arguments, etc.). If there is such markup in (sub)section titles (as above), a
-% plain text version has to be provided in the {\LaTeX} command as well. Below we
-% also illustrate how abbrevations should be introduced and citation commands can
-% be employed. See the {\LaTeX} code for more details.
-% \end{leftbar}
-
-%% -- Manuscript ---------------------------------------------------------------
-
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
-
-% \begin{leftbar}
-% As the synopsis above is a code listing that is not meant to be executed,
-% one can use either the dedicated \verb|{Code}| environment or a simple
-% \verb|{verbatim}| environment for this. Again, spaces before and after should be
-% avoided.
-
-% Finally, there might be a reference to a \verb|{table}| such as
-% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-% column headers and captions in sentence style, and if possible avoiding vertical
-% lines.
-% \end{leftbar}
-
-%% -- Illustrations ------------------------------------------------------------
-
-% - Virtually all JSS manuscripts list source code along with the generated
-%   output. The style files provide dedicated environments for this.
-% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%   or knitr using the render_sweave() hook - are used (without the need to
-%   load Sweave.sty).
-% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-% - The code input should use "the usual" command prompt in the respective
-%   software system.
-% - For R code, the prompt "R> " should be used with "+  " as the
-%   continuation prompt.
-% - Comments within the code chunks should be avoided - these should be made
-%   within the regular LaTeX text.
-
-% \begin{leftbar}
-% For code input and output, the style files provide dedicated environments.
-% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-% hook. Please make sure that all code is properly spaced, e.g., using
-% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-% use ``the usual'' command prompt in the respective software system. For
-% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-% the continuation prompt. Generally, comments within the code chunks should be
-% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-% before and after code input/output should be avoided (see above).
-% \end{leftbar}
-
-% \clearpage
-\section{Background}
-We first give a short introduction to simulator-based models and
-LFI. We then focus on the methods forming the basis of this paper; OMC
-and its robust improvement, ROMC. Finally, we introduce ELFI, the
-software package used for our implementation.
-
-%\input{latex_files/ch2/2_1_implicit_models}
-
-\subsection{Simulator-based models and likelihood-free inference}
-\input{latex_files/ch2/2-1-implicit-models.tex}
-
-\subsection{Optimisation Monte Carlo (OMC)}
-\input{latex_files/ch2/2-2-OMC.tex}
-
-\subsection{Robust optimisation Monte Carlo (ROMC)}
-\input{latex_files/ch2/2-3-ROMC.tex}
-
-\subsection[Engine for likelihood-free inference (ELFI)]{\pkg{Engine for likelihood-free inference (ELFI)}}
-\label{subsec:ELFI}
-\input{latex_files/ch2/2-4-ELFI.tex}
-
-%% -- Design Principles -----------------------------------------------------
-\section{Design principles}  % better title needed
-
-\input{latex_files/ch3/3-0-design-principles.tex}
-
-\subsection{General implementation principles}
-\label{subsec:general_design}
-\input{latex_files/ch3/3-1-implementation-principles.tex}
-
-%% -- ROMC implementation ----------------------------------------------------
-\section{Implemented functionalities}
-
-In this section, we present the implementation. At~\ref{subsec:ELFI}
-we introduce the reader to the \proglang{Python} package \pkg{ELFI}. At~\ref{subsec:general_design} we explain the general
-design that we follow. Then, we divide the implementation into three
-parts; at~\ref{subsec:training} we present the training (fitting)
-part, at~\ref{subsec:inference} the inference part and
-at~\ref{subsec:evaluation} the evaluation part. Finally,
-at~\ref{subsec:extensibility} we describe how a user may extend ROMC
-with its custom modules.
-
-\subsection{Training part} 
-\label{subsec:training}
-\input{latex_files/ch4/4-1-training.tex}
-
-\subsection{Inference part} 
-\label{subsec:inference}
-\input{latex_files/ch4/4-2-inference.tex}
-
-\subsection{Evaluation part} 
-\label{subsec:evaluation}
-\input{latex_files/ch4/4-3-evaluation.tex}
-
-\subsection{Extensibility} % better title
-\label{subsec:extensibility}
-\input{latex_files/ch4/4-4-extensibility.tex}
-
-%% -- Use-case illustration ----------------------------------------------------
-\section{Use-case illustration}
-
-In this section, we test the implementation using the second-order
-moving average (MA2) example, which is one of the standard models of
-\pkg{ELFI}. We perform the inference using three different versions of
-ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
-Optimisation scheme and (iii) fitting a Neural Network as a surrogate
-model. The later illustrates how to extend the implementation,
-replacing part of ROMC with a user-defined component. Finally, we
-measure the execution speed-up achieved by the parallelised version of
-ROMC.
-
-% \subsection{Moving average example}
-% \label{subsec:ma2}
-\input{latex_files/ch5/5-1-ma2.tex}
-
-% \subsection{Execution time}
-% \label{subsec:exec_time}
-% \input{latex_files/ch5/5-2-execution-time.tex}
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-\section{Summary and discussion} \label{sec:summary}
-
-In this paper, we presented the implementation details we followed for
-developing the LFI method ROMC at the \pkg{ELFI} package. We paid
-thorough attention to two specific use-case scenarios. Firstly, we
-illustrate how a user may take advantage of our ready-to-use API for
-solving its LFI problem. Secondly, we focus on the scenario where a
-researcher wants to intervene and alter parts of the method. Our
-implementation is designed to support this as well.
-
-There are still open challenges for the left for future research. Two
-directions may enable ROMC to solve high-dimensional problems
-efficiently. The first one is enabling ROMC's execution into a cluster
-of computers. ROMC can be characterized as an \textit{embarrassingly
-  parallel} workload; each optimization problem is an entirely
-independent task. Therefore, supporting inference into a cluster of
-computers can radically speed up the inference. The second one refers
-to implementing the method in a framework that supports automatic
-differentiation. Automatic differentiation is necessary for
-efficiently solving optimisation problems, especially in
-high-dimensional parametric models.
-
-%% -- Optional special unnumbered sections -------------------------------------
-
-% \section*{Computational details}
-
-% \begin{leftbar}
-% If necessary or useful, information about certain computational details
-% such as version numbers, operating systems, or compilers could be included
-% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-% maps, tables, \dots) that are not cited in the main text can be credited here.
-% \end{leftbar}
-
-% The results in this paper were obtained using
-% \proglang{R}~3.4.1 with the
-% \pkg{MASS}~7.3.47 package. \proglang{R} itself
-% and all packages used are available from the Comprehensive
-% \proglang{R} Archive Network (CRAN) at
-% \url{https://CRAN.R-project.org/}.
-
-% \section*{Acknowledgments}
-
-% \begin{leftbar}
-% All acknowledgments (note the AE spelling) should be collected in this
-% unnumbered section before the references. It may contain the usual information
-% about funding and feedback from colleagues/reviewers/etc. Furthermore,
-% information such as relative contributions of the authors may be added here
-% (if any).
-% \end{leftbar}
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
-\bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-% \newpage
-
-% \begin{appendix}
-
-% \section{More technical details} \label{app:technical}
-
-% \begin{leftbar}
-% Appendices can be included after the bibliography (with a page break). Each
-% section within the appendix should have a proper section title (rather than
-% just \emph{Appendix}).
-
-% For more technical style details, please check out JSS's style FAQ at
-% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-% which includes the following topics:
-% \begin{itemize}
-%   \item Title vs.\ sentence case.
-%   \item Graphics formatting.
-%   \item Naming conventions.
-%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
-%   \item Trouble shooting.
-%   \item Many other potentially helpful details\dots
-% \end{itemize}
-% \end{leftbar}
-
-% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-% \begin{leftbar}
-% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-% formats of author-year citations and allow to include additional details (e.g.,
-% pages, chapters, \dots) in brackets. In case you are not familiar with these
-% commands see the JSS style FAQ for details.
-
-% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-% when acquiring the entries automatically from mixed online sources. However,
-% it is important that informations are complete and presented in a consistent
-% style to avoid confusions. JSS requires the following format.
-% \begin{itemize}
-%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-%     be used in the references.
-%   \item Titles should be in title case.
-%   \item Journal titles should not be abbreviated and in title case.
-%   \item DOIs should be included where available.
-%   \item Software should be properly cited as well. For \proglang{R} packages
-%     \code{citation("pkgname")} typically provides a good starting point.
-% \end{itemize}
-% \end{leftbar}
-
-% \end{appendix}
-
-%% -----------------------------------------------------------------------------
-
-\end{document}

### 2022-02-15 article/latex_files/ch2/2-1-implicit-models.tex
final version

--- 2022-02-14 latex/latex_files/ch2/2-1-implicit-models.tex
@@
 reverse, or framing likelihood-free inference as a ratio estimation
-problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
-  Thomas2016, Papamakarios2019, Chen2019, Thomas2020,
-  Hermans2020}. However, both OMC and robust OMC, which we introduce
-next, are based on the approximation in \eqref{eq:approx_likelihood}.
+problem, see for example \citet{Blum2010, Wood2010, Papamakarios2016,
+  Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both
+OMC and robust OMC, which we introduce next, are based on the
+approximation in \eqref{eq:approx_likelihood}.

### 2022-02-15 article/latex_files/ch4/4-1-training.tex
final version

--- 2022-02-07 latex_files/ch4/4-1-training.tex
@@
-The training part contains six methods.
-
 \begin{Code}
 ---------------------------------- python ----------------------------------
->>> romc.solve_problems(n1,
-                        use_bo=False,
-                        optimizer_args=None,
-                        seed=None)
+>>> romc.solve_problems(n1, use_bo=False, optimizer_args=None)
 ----------------------------------------------------------------------------
@@
 optimisation problems and (c) solves them using either a
-gradient-based optimiser or Bayesian optimisation. These tasks are
-completed sequentially, as shown in
+gradient-based optimiser or the Bayesian optimisation (B0) scheme. The
+three tasks are completed sequentially, as shown in
 Figure~\ref{fig:romc_overview}. The definition of the optimisation
@@
 2^{32}-1\}\). Each integer \(u_i\) is the seed used in \pkg{ELFI}'s
-random simulator. The seeds \(u_i\) are computationally analogous to
-the random variables \(\vb_i\), that were defined in the previous
-chapter; setting the \code{seed} eliminates all randomness of the
-simulator transforming it to a deterministic function. Setting the
-argument \code{use_bo=True}, chooses the Bayesian Optimisation scheme
-for obtaining \(\thetab_i^*\). In this case, apart from obtaining the
-optimal points \(\thetab_i^*\), we also fit a GP as a surrogate model
-\(\hat{d}_i\). In this scenario, in the following steps,
-\(\hat{d}_i(\thetab)\) replaces \(d_i(\thetab)\) as the distance
-function.
+random simulator. The user may select the Bayesian Optimisation scheme
+by setting the argument \code{use_bo=True} and pass custom arguments
+to the optimizer through \code{optimize_args}.
 
@@
 This function helps the user decide which threshold \(\epsilon\) to
-use. It plots a histogram of the distances at the optimal point
-\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in case
-\code{use_bo=True}. The function accepts all keyword arguments and
-forwards them to the underlying \code{matplotlib.pyplot.hist()} function; in
-this way the user may customise some properties of the histogram, such
-as the number of bins or the range of values.
+use, by plotting a histogram of the distances at the optimal point
+\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in
+case \code{use_bo=True}. The function accepts all keyword arguments
+and forwards them to the underlying \code{pyplot.hist()} function of
+the package \pkg{matplotlib}. In this way the user may customise some
+properties of the histogram, such as the number of bins or the range
+of values.
 
@@
 ---------------------------------- python ----------------------------------
->>> romc.estimate_regions(eps_filter,
-                          use_surrogate=None,
-                          region_args=None,
-                          fit_models=False,
-                          fit_models_args=None,
-                          eps_region=None,
-                          eps_cutoff=None)
+>>> romc.estimate_regions(eps_filter, use_surrogate=None, fit_models=False)
 ----------------------------------------------------------------------------
@@
 \noindent
-This method constructs a bounding box around the optimal point
-\(\thetab_i^* : i = 1, 2, \ldots, n_1\) following
-Algorithm~\ref{alg:region_construction}. The Hessian matrix is
-approximated based on the Jacobian
-\(\hess_i = \jac_i^T \jac_i\)\footnote{The Jacobian matrix is computed
-  on the output of the summary statistics
-  \(\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))\) at \(\thetab_i^*\)} if
-gradients have been used, or as the Hessian of the GP, in case of
-Bayesian Optimisation. The eigenvectors are computed using the function
-\code{numpy.linalg.eig()}, and used as the axes of the
-bounding box. Afterwards, the limits along the
-eigenvectors are computed with
-Algorithm~\ref{alg:region_construction}.
+This method estimates the proposal region around the optimal points,
+following Algorithm~\ref{alg:region_construction}. By default, the
+distance model that will be used follows the decision from the
+previous step; if a gradient-based optimizer has been used, then the
+real distance function \(d\) will be chosen. If BO, then then the
+surrogate model \(\hat{d}\). In case, the user wants to enforce using
+the original distance function \(d\) after BO, they may set
+\code{use_surrogate=False}. Finally, the option \code{fit_models}
+defines whether to fit local surrogate models \(\tilde{d}\) after
+estimating the proposal regions.
 
@@
 the bounding box around the optimal point, when the parameter space is
-one- or two-dimensional. The argument \code{i} is the index of the corresponding
-optimisation problem i.e.\ \(i<n_1\). \code{compute_eps} returns the
+up to 2D. The argument \code{i} is the index of the corresponding
+optimisation problem. Finally, \code{compute_eps} returns the
 appropriate distance value \(d_{i=\kappa}^*\) where
 \(\kappa = \lfloor quantile \cdot n \rfloor\) from the collection
-\(\{ d_i^* \} \forall i = \{1, \ldots, n\}\) where \(n\) is the number of
-accepted solutions. It can be used to automate the selection of the
-threshold \(\epsilon\), e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.
+\(\{ d_i^* \} \forall i = \{1, \ldots, n\}\) where \(n\) is the number
+of accepted solutions. It can be used to automate the selection of the
+threshold \(\epsilon\), e.g.\
+\code{eps=romc.compute_eps(quantile=0.9)}.
 

### 2022-02-15 article/latex_files/ch4/4-2-inference.tex
final version

@@
-The inference part contains four methods.
-
 \begin{Code}
@@
 \noindent
-This function computes the expectation
+The function \code{compute_expectation} computes the expectation
 \(\E_{p(\thetab|\data)}[h(\thetab)]\) using
 expression~\eqref{eq:expectation}. The argument \code{h} can be any
-python \code{Callable}. Method \code{eval_unnorm_posterior} computes the
-unnormalised posterior approximation using
+python \code{Callable}. Method \code{eval_unnorm_posterior} computes
+the unnormalised posterior approximation using
 expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
@@
 
-In the following code snippet, we perform weighted sampling from the
-ROMC approximate posterior. Afterwards, we use some of \pkg{ELFI}'s
-built-in tools to get a summary of the obtained samples. In Figure
-\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
-samples and the acceptance region of the first deterministic function
-(as before) alongside the samples obtained from it. Finally, in the
-code snippet, we demonstrate how to use the \code{compute_expectation}
-function; in the current example, we define \code{h} to compute
-firstly the empirical mean and afterwards the empirical variance. In
-both cases, the empirical result is close to the ground truth
-\(\mu = 0\) and \(\sigma^2 = 1\). The
-\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
-point \(\theta\) using expression \eqref{eq:approx_posterior_omc}. The
-\code{romc.eval_posterior(theta)} approximates the partition function
-\(Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab\) using the
-Riemann approximation as explained above. In our simple example, this
-utility can provide a nice plot of the approximate posterior as
-illustrated in Figure~\ref{fig:running_example_romc_inference}. We observe that the
-approximation is quite close to the ground-truth posterior.
+In the following code snippet, we use the inference utilities to (a)
+get weighted samples from the approximate posterior, (b) compute an
+expectation and (c) evaluate the approximate posterior. We also use
+some of \pkg{ELFI}'s built-in tools to get a summary of the obtained
+samples. For the utility \code{compute_expectation}, we demonstrate
+how to use it in order to compute the samples mean and
+variance. Finally, we evaluate \code{eval_posterior} at multiple
+points to plot the approximate posterior of
+Figure~\ref{fig:running_example_romc_inference}. We observe that the
+approximation is quite close to the ground-truth.
 

### 2022-02-15 article/latex_files/ch4/4-3-evaluation.tex
final version

\begin{Code}
---------------------------------- python ----------------------------------  
>>> romc.compute_divergence(gt_posterior, bounds, step, distance)
>>> romc.compute_ess()
----------------------------------------------------------------------------
\end{Code}

\noindent
The utility \code{compute_divergence} estimate the divergence between
the ROMC approximation and the ground truth posterior. Since the
estimation is performed using the Riemann's approximation, the method
can only work in low dimensional spaces. As mentioned at the beginning
of this chapter, in a real-case scenario it is not expected the
ground-truth posterior to be available. In cases the ground truth
posterior is not availabe (as in real-scenarios), the user may select
the approximation obtained with any other inference approach for
comparing the two methods. The argument \code{step} defines the step
used in the Riemann's approximation and the argument \code{distance}
can be either \code{"Jensen-Shannon"} or \code{"KL-divergence"}.

The method \code{compute_ess} computes the Effective Sample Size (ESS)
using the following expression~\cite{Sudman1967},

\begin{equation} \label{eq:ESS}
  ESS = \frac{(\sum_i w_i)^2}{\sum_i w_i^2}
\end{equation}

The ESS is a valuable measure of the \textbf{actual} sample size in
cases of weighted samples. For example, there are cases where in a big
population, a single sample with huge weight dominates. The ESS
provides a nice metric in these cases.

\begin{Code}
------------------------------ python snippet ------------------------------  
  # Evaluation part
  res = romc.compute_divergence(wrapper, distance="Jensen-Shannon")                                 
  print("Jensen-Shannon divergence: %.3f" % res)
  # Jensen-Shannon divergence: 0.035

  nof_samples = len(romc.result.weights)
  ess = romc.compute_ess()
  print("Nof Samples: %d, ESS: %.3f" % (nof_samples, ess))
  # Nof Samples: 19300, ESS: 16196.214
----------------------------------------------------------------------------  
\end{Code}

### 2022-02-15 article/latex_files/ch5/5-ma2.tex
final version

--- 2022-02-15 article/latex_files/ch5/5-1-ma2.tex
@@
+In this section, we test the implementation using the second-order
+moving average (MA2) example, which is one of the standard models of
+\pkg{ELFI}. We perform the inference using three different versions of
+ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
+Optimisation scheme and (iii) fitting a Neural Network as a surrogate
+model. The later illustrates how to extend the implementation,
+replacing part of ROMC with a user-defined component. Finally, we
+measure the execution speed-up achieved by the parallelised version of
+ROMC.
+
 \subsubsection*{Model Definition}
@@
 
-As already stated, it is straightforward for the two basic parts of
-the training part to be parallelised. There is a batch of \(n_1\)
-independent objectives functions to be optimised and \(n_1\) bounding
-boxes to be built. Both tasks can be applied in a parallel utilising
-all the available CPU cores. Our implementation supports
-parallelisation using the built-in \proglang{Python} package
-\pkg{multiprocess}. In Figure~\ref{fig:exec_parallel} we observe the
-execution times for performing the inference. The parallel version
-performs all tasks between 2.5 and 6 times faster compared to the
-sequential; the optimisation problems are solved almost 6 times
-faster, sampling 3.5 times faster and evaluating the posterior and
-constructing the bounding boxes almost 2.5 times faster.
+As stated above, ROMC is a ridiculously parallelisable
+method. Therefore, it is straightforward to parallelise the fitting
+part, i.e. (i) solving the optimisation problems and (ii) estimating
+the proposal regions. Our implementation supports exploiting all the
+available CPU cores through the built-in \proglang{Python} package
+\pkg{multiprocess}\footnote{https://docs.python.org/3/library/multiprocessing.html}. In
+Figure~\ref{fig:exec_parallel} we observe the execution times for
+performing the inference on the ma2 model; the parallel version
+performs both tasks almost six times faster than the sequential. The
+result is reasonable given that the experiments have run in a single
+machine with the Intel® Core™ i7-8750H Processor, which has six
+separate cores.
 
@@
   \begin{center}
-    \resizebox{.47\columnwidth}{!}{%
+    \resizebox{.49\columnwidth}{!}{%
       \input{./latex_files/images/chapter4/exec_time_solve.tex}
@@
     \end{center}
-    \caption[Execution time exploiting parallelisation]{In the first
-      line, we compare the parallel and sequential execution of the
-      training part and at the second line the inference part. At the
-      left figure, we measure the execution time for sampling \(n_2=50\)
-      points per region. At the right figure we measure the execution
-      time for evaluating the posterior at a batch of \(50\) points.}
+    \caption[Execution time exploiting parallelisation]{Comparison
+      between parallel and sequential execution of the fitting part of
+      ROMC. We observe that the parallel version runs between 2.5 an 6
+      times faster.}
       \label{fig:exec_parallel}
 \end{figure}
-
-% \begin{figure}[ht]
-%   \centering
-%     \begin{minipage}[t]{.4\textwidth}
-%       \begin{tikzpicture}
-%         \begin{axis}[
-%           title={Solve optimisation problems},
-%           xlabel={\(n_1\)},
-%           ylabel={seconds},
-%           xmin=1, xmax=500,
-%           ymin=0, ymax=7,
-%           xtick={1,100,200,300,400,500},
-%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
-%           legend pos=north west,
-%           ymajorgrids=true,
-%           grid style=dashed,
-%           ]
-%           \addplot[color=blue, mark=o] coordinates {
-%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-%           };
-%           \addlegendentry{sequential}
-%           \addplot[color=red, mark=o] coordinates {
-%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-%           };
-%           \addlegendentry{parallel}          
-%         \end{axis}
-%       \end{tikzpicture}
-%     \end{minipage}\hfill
-%     \begin{minipage}[t]{.4\textwidth}
-%       \begin{tikzpicture}
-%         \begin{axis}[
-%           title={Proposal regions},
-%           xlabel={\(n_1\)},
-%           ylabel={seconds},
-%           xmin=1, xmax=500,
-%           ymin=0, ymax=7,
-%           xtick={1,100,200,300,400,500},
-%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
-%           legend pos=north west,
-%           ymajorgrids=true,
-%           grid style=dashed,
-%           ]
-%           \addplot[color=blue, mark=o] coordinates {
-%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-%           };
-%           \addlegendentry{sequential}
-%           \addplot[color=red, mark=o] coordinates {
-%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-%           };
-%           \addlegendentry{parallel}          
-%         \end{axis}
-%       \end{tikzpicture}
-%     \end{minipage}\hfill
-%     \begin{minipage}[t]{.4\textwidth}
-%       \begin{tikzpicture}
-%         \begin{axis}[
-%           title={Sample},
-%           xlabel={\(n_1\)},
-%           ylabel={seconds},
-%           xmin=1, xmax=500,
-%           ymin=0, ymax=7,
-%           xtick={1,100,200,300,400,500},
-%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
-%           legend pos=north west,
-%           ymajorgrids=true,
-%           grid style=dashed,
-%           ]
-%           \addplot[color=blue, mark=o] coordinates {
-%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-%           };
-%           \addlegendentry{sequential}
-%           \addplot[color=red, mark=o] coordinates {
-%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-%           };
-%           \addlegendentry{parallel}          
-%         \end{axis}
-%       \end{tikzpicture}
-%     \end{minipage}\hfill
-%     \begin{minipage}[t]{.4\textwidth}
-%       \begin{tikzpicture}
-%         \begin{axis}[
-%           title={Evaluate posterior},
-%           xlabel={\(n_1\)},
-%           ylabel={seconds},
-%           xmin=1, xmax=500,
-%           ymin=0, ymax=7,
-%           xtick={1,100,200,300,400,500},
-%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
-%           legend pos=north west,
-%           ymajorgrids=true,
-%           grid style=dashed,
-%           ]
-%           \addplot[color=blue, mark=o] coordinates {
-%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
-%           };
-%           \addlegendentry{sequential}
-%           \addplot[color=red, mark=o] coordinates {
-%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
-%           };
-%           \addlegendentry{parallel}          
-%         \end{axis}
-%       \end{tikzpicture}
-%     \end{minipage}\hfill
-%     \caption[Execution time exploiting parallelisation]{In the first
-%       line, we compare the parallel and sequential execution of the
-%       training part and at the second line the inference part. At the
-%       left figure, we measure the execution time for sampling \(n_2=50\)
-%       points per region. At the right figure we measure the execution
-%       time for evaluating the posterior at a batch of \(50\) points.}
-%       \label{fig:exec_parallel}
-% \end{figure}

### 2022-02-15 article/latex_files/ch6/6-summary.tex
final version

--- 2022-02-14 latex/article.tex
@@
-\documentclass[article]{jss}
-%% -- LaTeX packages and custom commands ---------------------------------------
-
-\usepackage[usenames,dvipsnames]{xcolor}
-
-%% recommended packages
-\usepackage{thumbpdf,lmodern}
-\usepackage{orcidlink}
-\usepackage{amsmath,amssymb,bbold, xfrac}
-
-\usepackage{multirow}
-\usepackage{caption}
-\usepackage{subcaption} % for subfigure
-\usepackage{float}
-
-%% TikZ packages
-\usepackage{tikz}
-\usetikzlibrary{shapes.geometric, arrows, backgrounds, scopes}
-\usepackage{pgfplots}
-\pgfplotsset{width=6.75cm, compat=newest}
-\usepackage[utf8]{inputenc}
-\DeclareUnicodeCharacter{2212}{−}
-\usepgfplotslibrary{groupplots,dateplot}
-\usetikzlibrary{patterns,shapes.arrows}
-
-%% for algorithms
-\usepackage{algorithm}
-\usepackage{algpseudocode} % algorithmicx package
-
-%% another package (only for this demo article)
-\usepackage{framed}
-\usepackage{marginnote}
-
-%% new custom commands
-\newcommand{\class}[1]{`\code{#1}'}
-\newcommand{\fct}[1]{\code{#1()}}
-
-%% Vasilis' custom commands
-\newcommand{\bigO}{\mathcal{O}}
-\newcommand{\X}{\mathbf{x}}
-\newcommand{\Z}{\mathbf{z}}
-\newcommand{\V}{\mathbf{V}}
-\newcommand{\Y}{\mathbf{Y}}
-\newcommand{\hess}{\mathbf{H}}
-\newcommand{\Thetab}{\mathbf{\Theta}}
-
-\newcommand{\vb}{\mathbf{v}}
-\newcommand{\yb}{\mathbf{y}}
-\newcommand{\xb}{\mathbf{x}}
-
-\newcommand{\jac}{\mathbf{J}}
-\newcommand{\hessian}{\mathbf{H}}
-\newcommand{\thetab}{\boldsymbol{\theta}}
-\newcommand{\thetabij}{\thetab_{ij}}
-\newcommand{\thetabi}{\thetab_i}
-
-\newcommand{\simulator}{g}
-\newcommand{\region}{B_{d,\epsilon}}
-\newcommand{\indicator}[1]{\mathbb{1}_{#1}}
-\newcommand{\regioni}{B_{d,\epsilon}^i}
-\newcommand{\data}{\mathbf{y_0}}
-\newcommand{\accregioni}{C^i_{\epsilon}}
-\newcommand{\accregionihat}{\hat{C}^i_{\epsilon}}
-
-\newcommand{\danger}{\fontencoding{U}\fontfamily{futs}\selectfont\char 66\relax}
-\newcommand{\TODO}[1]{\textcolor{red}{TODO #1}\marginpar{\textcolor{red}{\Large \danger}}}
-
-\newcommand{\R}{\mathbb{R}}
-
-\newcommand{\Ex}{\mathbb{E}}
-
-\DeclareMathOperator*{\argmax}{argmax}
-\DeclareMathOperator*{\argmin}{argmin}
-
-%% -- Article metainformation (author, title, ...) -----------------------------
-
-\author{Vasilis Gkolemis~\orcidlink{0000-0002-2636-0245}\\ATHENA RC \And
-  Michael Gutmann~\orcidlink{0000-0002-5329-9910}\\University of Edinburgh \And
-  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo
-}
-\Plainauthor{Vasilis Gkolemis, Michael Gutmann, Henri Pesonen}
-
-\title{An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo}
-\Plaintitle{An Extendable Python Implementation of ROMC}
-\Shorttitle{An Extendable ROMC implementation}
-
-%% - \Abstract{} almost as usual
-\Abstract{ Performing the inference in statistical models with
-  intractable likelihood is challenging; most likelihood-free
-  inference (LFI) methods encounter problems in accuracy and
-  efficiency. In this paper, we present the implementation of Robust
-  Optimisation Monte Carlo (ROMC) in the \proglang{Python} package
-  \pkg{ELFI}. ROMC is a novel and efficient (highly-parallelizable)
-  LFI framework that provides accurate weighted samples from the
-  posterior. Our implementation can be used in two ways. Firstly, a
-  scientist may use it as an out-of-the-box LFI algorithm; we provide
-  an easy-to-use API harmonised with the rest of the \pkg{ELFI}
-  framework.  Therefore, it is straightforward to compare it with the
-  rest of the methods included in \pkg{ELFI}. Afterwards, we have
-  carefully split ROMC into isolated algorithmic steps. Hence, we
-  allow the researcher to experiment by replacing part(s) of the
-  algorithm with novel subcomponent(s). In both scenarios, all steps
-  of ROMC run in a fully-parallelizable manner, exploiting all CPU
-  cores. We also provide helpful functionalities for (i) inspecting
-  the inference process and (ii) evaluating the obtained
-  samples. Finally, we test the robustness of our implementation on
-  some typical LFI examples.}
-
-%% - \Keywords{}
-\Keywords{Bayesian inference, implicit models, likelihood-free, \proglang{Python}, \pkg{ELFI}}
-\Plainkeywords{Bayesian inference, implicit models, likelihood-free, Python, ELFI}
-
-%% - \Address{} of at least one author
-\Address{
-  Vasilis Gkolemis\\
-  Information Management Systems Institute (IMSI)\\
-  ATHENA Research and Innovation Center\\
-  Athens, Greece\\
-  E-mail: \email{vgkolemis@athenarc.gr}\\
-  URL: \url{https://givasile.github.io}
-}
-
-\begin{document}
-
-%% -- Introduction -------------------------------------------------------------
-
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
-% \section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
-\section{Introduction} \label{sec:intro}
-
-Simulator-based models are particularly captivating due to the
-modelling freedom they provide. In essence, a simulator-based model can
-describe any data generating mechanism that can be written as a finite
-set of algorithmic steps. This modelling freedom comes at a cost;
-performing the inference, i.e., sample or evaluate the posterior
-distribution, is challenging.
-
-Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} was a
-novel LFI approach for approximating the posterior
-distribution. Afterwards, \citet{Forneron2016} provided a similar
-method naming it `reverse sampler'. The central idea is turning the
-stochastic data-generating mechanism into a set of
-fully-parallelizable deterministic optimisation processes. In their
-work, ~\citet{Ikonomov2019}, located some critical failure modes of
-OMC, and proposed Robust OMC (ROMC) an alternative version of OMC with
-the appropriate improvements.
-
-In this paper, we present an extendible and parallelisable
-implementation of ROMC at the \proglang{Python} package \pkg{Engine
-  for Likelihood-Free inference (ELFI)}. We choose carefully the
-designing principles for making the implementation extendible. As we
-describe analytically in the next chapter, ROMC can be understood not
-only as a specific algorithm, but mainly as a framework for obtaining
-weighted samples from the posterior; it defines a sequence of
-algorithmic steps, without enforcing a specific algorithm for solving
-them. Therefore, a researcher may adopt ROMC's backbone approach, but
-experiment with alternative methods to solve the tasks. We have
-designed ROMC for facilitating such alterations. Finally, we have
-tested the accuracy and the efficiency of our implementation on the
-LFI examples supported by the \pkg{ELFI} package.
-
-To the best of our knowledge, this is the first attempt to implement
-the ROMC inference method to a generic LFI framework. Therefore, there
-are not other implementation directly comparable to ours. For this
-reason, we test the implementation against (i) an artificial example
-with tractable likelihood and (ii) the second-order moving average
-(MA2) example from the \pkg{ELFI} package. In the latter, we generate
-data using known parameters and we compare our results with the
-typical Rejection ABC approach.
-
-% \begin{leftbar}
-% The introduction is in principle ``as usual''. However, it should usually embed
-% both the implemented \emph{methods} and the \emph{software} into the respective
-% relevant literature. For the latter both competing and complementary software
-% should be discussed (within the same software environment and beyond), bringing
-% out relative (dis)advantages. All software mentioned should be properly
-% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
-% \textsc{Bib}{\TeX}.)
-
-% For writing about software JSS requires authors to use the markup
-% \verb|\proglang{}| (programming languages and large programmable systems),
-% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
-% arguments, etc.). If there is such markup in (sub)section titles (as above), a
-% plain text version has to be provided in the {\LaTeX} command as well. Below we
-% also illustrate how abbrevations should be introduced and citation commands can
-% be employed. See the {\LaTeX} code for more details.
-% \end{leftbar}
-
-%% -- Manuscript ---------------------------------------------------------------
-
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
-
-% \begin{leftbar}
-% As the synopsis above is a code listing that is not meant to be executed,
-% one can use either the dedicated \verb|{Code}| environment or a simple
-% \verb|{verbatim}| environment for this. Again, spaces before and after should be
-% avoided.
-
-% Finally, there might be a reference to a \verb|{table}| such as
-% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
-% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
-% column headers and captions in sentence style, and if possible avoiding vertical
-% lines.
-% \end{leftbar}
-
-%% -- Illustrations ------------------------------------------------------------
-
-% - Virtually all JSS manuscripts list source code along with the generated
-%   output. The style files provide dedicated environments for this.
-% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
-%   or knitr using the render_sweave() hook - are used (without the need to
-%   load Sweave.sty).
-% - Equivalently, {CodeInput} and {CodeOutput} can be used.
-% - The code input should use "the usual" command prompt in the respective
-%   software system.
-% - For R code, the prompt "R> " should be used with "+  " as the
-%   continuation prompt.
-% - Comments within the code chunks should be avoided - these should be made
-%   within the regular LaTeX text.
-
-% \begin{leftbar}
-% For code input and output, the style files provide dedicated environments.
-% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
-% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
-% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
-% hook. Please make sure that all code is properly spaced, e.g., using
-% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
-% use ``the usual'' command prompt in the respective software system. For
-% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
-% the continuation prompt. Generally, comments within the code chunks should be
-% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
-% before and after code input/output should be avoided (see above).
-% \end{leftbar}
-
-% \clearpage
-\section{Background}
-We first give a short introduction to simulator-based models and
-LFI. We then focus on the methods forming the basis of this paper; OMC
-and its robust improvement, ROMC. Finally, we introduce ELFI, the
-software package used for our implementation.
-
-%\input{latex_files/ch2/2_1_implicit_models}
-
-\subsection{Simulator-based models and likelihood-free inference}
-\input{latex_files/ch2/2-1-implicit-models.tex}
-
-\subsection{Optimisation Monte Carlo (OMC)}
-\input{latex_files/ch2/2-2-OMC.tex}
-
-\subsection{Robust optimisation Monte Carlo (ROMC)}
-\input{latex_files/ch2/2-3-ROMC.tex}
-
-\subsection[Engine for likelihood-free inference (ELFI)]{\pkg{Engine for likelihood-free inference (ELFI)}}
-\label{subsec:ELFI}
-\input{latex_files/ch2/2-4-ELFI.tex}
-
-%% -- Design Principles -----------------------------------------------------
-\section{Design principles}  % better title needed
-
-\input{latex_files/ch3/3-0-design-principles.tex}
-
-\subsection{General implementation principles}
-\label{subsec:general_design}
-\input{latex_files/ch3/3-1-implementation-principles.tex}
-
-%% -- ROMC implementation ----------------------------------------------------
-\section{Implemented functionalities}
-
-In this section, we present the implementation. At~\ref{subsec:ELFI}
-we introduce the reader to the \proglang{Python} package \pkg{ELFI}. At~\ref{subsec:general_design} we explain the general
-design that we follow. Then, we divide the implementation into three
-parts; at~\ref{subsec:training} we present the training (fitting)
-part, at~\ref{subsec:inference} the inference part and
-at~\ref{subsec:evaluation} the evaluation part. Finally,
-at~\ref{subsec:extensibility} we describe how a user may extend ROMC
-with its custom modules.
-
-\subsection{Training part} 
-\label{subsec:training}
-\input{latex_files/ch4/4-1-training.tex}
-
-\subsection{Inference part} 
-\label{subsec:inference}
-\input{latex_files/ch4/4-2-inference.tex}
-
-\subsection{Evaluation part} 
-\label{subsec:evaluation}
-\input{latex_files/ch4/4-3-evaluation.tex}
-
-\subsection{Extensibility} % better title
-\label{subsec:extensibility}
-\input{latex_files/ch4/4-4-extensibility.tex}
-
-%% -- Use-case illustration ----------------------------------------------------
-\section{Use-case illustration}
-
-In this section, we test the implementation using the second-order
-moving average (MA2) example, which is one of the standard models of
-\pkg{ELFI}. We perform the inference using three different versions of
-ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
-Optimisation scheme and (iii) fitting a Neural Network as a surrogate
-model. The later illustrates how to extend the implementation,
-replacing part of ROMC with a user-defined component. Finally, we
-measure the execution speed-up achieved by the parallelised version of
-ROMC.
-
-% \subsection{Moving average example}
-% \label{subsec:ma2}
-\input{latex_files/ch5/5-1-ma2.tex}
-
-% \subsection{Execution time}
-% \label{subsec:exec_time}
-% \input{latex_files/ch5/5-2-execution-time.tex}
-
-%% -- Summary/conclusions/discussion -------------------------------------------
-\section{Summary and discussion} \label{sec:summary}
-
 In this paper, we presented the implementation details we followed for
@@
 high-dimensional parametric models.
-
-%% -- Optional special unnumbered sections -------------------------------------
-
-% \section*{Computational details}
-
-% \begin{leftbar}
-% If necessary or useful, information about certain computational details
-% such as version numbers, operating systems, or compilers could be included
-% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-% maps, tables, \dots) that are not cited in the main text can be credited here.
-% \end{leftbar}
-
-% The results in this paper were obtained using
-% \proglang{R}~3.4.1 with the
-% \pkg{MASS}~7.3.47 package. \proglang{R} itself
-% and all packages used are available from the Comprehensive
-% \proglang{R} Archive Network (CRAN) at
-% \url{https://CRAN.R-project.org/}.
-
-% \section*{Acknowledgments}
-
-% \begin{leftbar}
-% All acknowledgments (note the AE spelling) should be collected in this
-% unnumbered section before the references. It may contain the usual information
-% about funding and feedback from colleagues/reviewers/etc. Furthermore,
-% information such as relative contributions of the authors may be added here
-% (if any).
-% \end{leftbar}
-
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
-\bibliography{refs}
-
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-% \newpage
-
-% \begin{appendix}
-
-% \section{More technical details} \label{app:technical}
-
-% \begin{leftbar}
-% Appendices can be included after the bibliography (with a page break). Each
-% section within the appendix should have a proper section title (rather than
-% just \emph{Appendix}).
-
-% For more technical style details, please check out JSS's style FAQ at
-% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-% which includes the following topics:
-% \begin{itemize}
-%   \item Title vs.\ sentence case.
-%   \item Graphics formatting.
-%   \item Naming conventions.
-%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
-%   \item Trouble shooting.
-%   \item Many other potentially helpful details\dots
-% \end{itemize}
-% \end{leftbar}
-
-% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-% \begin{leftbar}
-% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-% formats of author-year citations and allow to include additional details (e.g.,
-% pages, chapters, \dots) in brackets. In case you are not familiar with these
-% commands see the JSS style FAQ for details.
-
-% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-% when acquiring the entries automatically from mixed online sources. However,
-% it is important that informations are complete and presented in a consistent
-% style to avoid confusions. JSS requires the following format.
-% \begin{itemize}
-%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-%     be used in the references.
-%   \item Titles should be in title case.
-%   \item Journal titles should not be abbreviated and in title case.
-%   \item DOIs should be included where available.
-%   \item Software should be properly cited as well. For \proglang{R} packages
-%     \code{citation("pkgname")} typically provides a good starting point.
-% \end{itemize}
-% \end{leftbar}
-
-% \end{appendix}
-
-%% -----------------------------------------------------------------------------
-
-\end{document}

### 2022-02-15 article/latex_files/python_scripts/example_MA2.py

@@ -1,174 +1,0 @@
-import logging
-import matplotlib.pyplot as plt
-from elfi.examples import ma2
-import elfi
-import numpy as np
-import scipy.stats
-import matplotlib
-import os
-import tikzplotlib as tplt
-matplotlib.rcParams['text.usetex'] = True
-
-logging.basicConfig(level=logging.INFO)
-prepath = './../images/chapter4/'
-savefig = False
-
-def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, range, ylim, savepath):
-    plt.figure()
-    plt.title(title)
-    plt.hist(x=samples,
-             weights=weights,
-             bins=bins, density=True, range=range)
-    plt.axvline(mean, 0, 1,
-                color="r", linestyle="--", label=r"$\mu = %.3f$" % (mean))
-    plt.axhline(1,
-                (mean-std-range[0])/(range[1] - range[0]),
-                (mean+std-range[0])/(range[1] - range[0]),
-                color="k",
-                linestyle="--", label=r"$\sigma = %.3f$" % (std))
-    plt.xlabel(xlabel)
-    plt.ylabel(ylabel)
-    plt.ylim(ylim)
-    plt.legend()
-    if savepath:
-        plt.savefig(savepath, bbox_inches='tight')
-        # tplt.clean_figure()
-        tplt.save(savepath[:-3] + "tex")
-    plt.show(block=False)
-
-# Set seed for reproducibility
-seed = 1
-np.random.seed(seed)
-model = ma2.get_model(seed_obs=seed)
-
-# plot prior samples
-
-x = model.generate(1000, outputs=["t1", "t2"])
-plt.figure()
-plt.title("Samples from the prior")
-plt.xlabel(r"$\theta_1$")
-plt.ylabel(r"$\theta_2$")
-plt.plot(x["t1"], x["t2"], "bo")
-if savefig:
-    plt.savefig(os.path.join(prepath, "mae2_prior_samples.png"),
-                bbox_inches="tight")
-    tplt.clean_figure()
-    tplt.save(os.path.join(prepath, "ma2_prior_samples.tex"))
-plt.show(block=False)
-
-####### ROMC with gradients ################
-n1 = 100
-n2 = 200
-bounds = [(-2, 2), (-1.25, 1.25)]
-eps = .01
-vis_ind_1 = 1
-vis_ind_2 = 3
-vis_ind_3 = 12
-
[…]

### 2022-02-15 replication_script/separate_scripts/execution_times_parallel.py

@@ -1,319 +1,0 @@
-"""Testing the execution times using parallelisation."""
-
-import timeit
-import time
-import matplotlib.pyplot as plt
-import numpy as np
-import scipy.integrate as integrate
-import scipy.ndimage as ndimage
-import scipy.stats as ss
-import os
-import elfi
-import matplotlib
-matplotlib.rcParams['text.usetex'] = True
-
-np.random.seed(21)
-prepath = './../images/chapter4' 
-savefig = True
-
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
-
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
-
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
-
-
-class Likelihood:
-    def rvs(self, th1, th2, seed=None):
-        assert isinstance(th1, np.ndarray)
-        assert isinstance(th2, np.ndarray)
-        assert th1.ndim == 2
-        assert th2.ndim == 2
-        assert np.allclose(th1.shape, th2.shape)
-
-        x = []
-        for i in range(th1.shape[0]):
-            cur_th = np.concatenate((th1[i], th2[i]))
-            x.append(ss.multivariate_normal(
-                mean=cur_th, cov=1).rvs(random_state=seed))
-        return np.array(x)
-
-    def pdf(self, x, th1, th2):
-        assert isinstance(th1, float)
-        assert isinstance(th2, float)
-        assert isinstance(x, np.ndarray)
-
-        th = np.stack((th1, th2))
-        rv = ss.multivariate_normal(mean=th, cov=1)
-        nof_points = x.shape[0]
-        prod = 1
-        for i in range(nof_points):
-            prod *= rv.pdf(x[i])
-        return prod
-
-
-def create_factor(x):
-    lik = Likelihood()
-    pr = Prior()
-
-    def tmp_func(th1, th2):
[…]

### 2022-02-20 article/article.tex
added requirements.txt file
final modifications

@@
 \newcommand{\vb}{\mathbf{v}}
+\newcommand{\ub}{\mathbf{u}}
 \newcommand{\yb}{\mathbf{y}}
@@
 %% - \Abstract{} almost as usual
-\Abstract{ Performing the inference in statistical models with
-  intractable likelihood is challenging; most likelihood-free
-  inference (LFI) methods encounter problems in accuracy and
-  efficiency. In this paper, we present the implementation of Robust
-  Optimisation Monte Carlo (ROMC) in the \proglang{Python} package
-  \pkg{ELFI}. ROMC is a novel and efficient (highly-parallelizable)
-  LFI framework that provides accurate weighted samples from the
-  posterior. Our implementation can be used in two ways. Firstly, a
-  scientist may use it as an out-of-the-box LFI algorithm; we provide
-  an easy-to-use API harmonised with the rest of the \pkg{ELFI}
-  framework.  Therefore, it is straightforward to compare it with the
-  rest of the methods included in \pkg{ELFI}. Afterwards, we have
-  carefully split ROMC into isolated algorithmic steps. Hence, we
-  allow the researcher to experiment by replacing part(s) of the
-  algorithm with novel subcomponent(s). In both scenarios, all steps
-  of ROMC run in a fully-parallelizable manner, exploiting all CPU
+\Abstract{Performing inference in statistical models with intractable
+  likelihood is challenging; most likelihood-free inference (LFI)
+  methods encounter accuracy and efficiency limitations. In this
+  paper, we present the implementation of Robust Optimisation Monte
+  Carlo (ROMC) in the \proglang{Python} package \pkg{ELFI}. ROMC is a
+  novel and efficient (highly-parallelisable) LFI framework that
+  provides accurate weighted samples from the posterior. Our
+  implementation can be used in two ways. First, a scientist may use
+  it as an out-of-the-box LFI algorithm; we provide an easy-to-use API
+  harmonised with the principles of \pkg{ELFI}, enabling effortless
+  comparisons with the rest of the methods included in the
+  package. Additionally, we have carefully split ROMC into isolated
+  components for supporting extensibility. A researcher may experiment
+  with novel method(s) for solving part(s) of ROMC without
+  reimplementing everything from scratch. In both scenarios, all steps
+  run in a fully-parallelisable manner, exploiting all CPU
   cores. We also provide helpful functionalities for (i) inspecting
@@
 %% -- Design Principles -----------------------------------------------------
-\section{Design principles}  % better title needed
+\section{Implementation}
+This section is split in two parts. We first express ROMC as an
+algorithm and then we present the general implementation principles we
+follow.
+
+\subsection{Algotirhmic view of ROMC}
 \input{latex_files/ch3/3-0-design-principles.tex}
@@
 \section*{Computational details}
-
-The results in this paper were obtained using \proglang{Python}~3.7.1
-with the \pkg{ELFI}~0.8.2 package. The experiments have been executed
+The results in this paper were obtained using \proglang{Python}~3.7.9
+with the \pkg{ELFI}~0.8.3 package. The experiments have been executed
 in a single machine with an Intel® Core™ i7-8750H Processor an with
@@
 
-% \section*{Acknowledgments}
+\section*{Acknowledgments}
+
+HP was funded by European Research Council grant 742158 (SCARABEE,
+Scalable inference algorithms for Bayesian evolutionary epidemiology).
 

### 2022-02-20 article/latex_files/ch2/2-2-OMC.tex
final modifications

--- 2022-02-07 latex_files/ch2/2-2-OMC.tex
@@
 \end{gather}
-which we can approximate as a sample average
+which can be approximated as a sample's average
 \begin{equation}
 L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
- \quad \text{ where } \yb_i = \simulator(\thetab, \vb_i),\; \vb_i \sim p(\vb). \label{eq:alt_view}
+ \quad \text{ where } \yb_i = \simulator(\thetab, \ub_i),\; \ub_i \sim p(\ub). \label{eq:alt_view}
 \end{equation}
-We thus approximate the likelihood of a specific \(\thetab\) by counting
-the fraction of samples that lie inside the volume around the
-observations. The fact that we isolated the randomness of the
-simulator via the \(\vb_i\) has two crucial consequences: First, we can
-sample the nuisance variables \(\vb_i \sim p(\vb)\) only once and use
-the same set of samples to approximate \(L_{d,\epsilon}(\thetab)\) for
-different values of \(\thetab\). Secondly, since
-\(\yb_i=\simulator(\thetab, \vb_i)\) is a function of \(\thetab\),
-checking whether \(\yb_i\) is contained in region \(\region (\data)\) for
-a particular value of \(\thetab\) is the same as checking whether
-\(\thetab\) is in the acceptance region \(\accregioni
-= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}\). This
-leads to the likelihood approximation
+We thus approximate the likelihood of a specific \(\thetab\) by
+counting the fraction of samples that lie inside a volume around the
+observations. Isolating the randomness of the simulator via the
+\(\ub_i\) has two crucial consequences: First, we can sample the
+nuisance variables \(\ub_i \sim p(\ub)\) only once and reuse them to
+approximate \(L_{d,\epsilon}(\thetab)\) for different
+\(\thetab\). Second, since \(\yb_i=\simulator(\thetab, \ub_i)\) is a
+function of \(\thetab\), checking whether \(\yb_i\) is contained in
+\(\region (\data)\) is the same as checking whether \(\thetab\) is in
+the acceptance region
+\(\accregioni = \{ \thetab: \simulator(\thetab, \ub_i) \in
+\region(\data) \}\). This leads to the likelihood approximation
 \begin{equation}
@@
 the geometry of the acceptance regions \(\accregioni\). Indeed, the
-task of approximating the likelihood and/or posterior becomes a task
+task of approximating the likelihood and the posterior becomes a task
 of characterising the sets \(\accregioni\). OMC by \citet{Meeds2015}
@@
 and generated data, and that the \(\accregioni\) can be well
-approximated by infinitesimally small ellipses. These assumptions
-leads to an approximation of the posterior in terms of weighted
-samples \(\thetab_i^*\) that achieve the smallest distance between
-observed and simulated data for each realisation
-\(\vb_i \sim p(\vb)\), i.e.\
+approximated by infinitesimally small ellipses. These assumptions lead
+to an approximation of the posterior in terms of weighted samples
+\(\thetab_i^*\) that achieve the smallest distance between observed
+and simulated data for each realisation \(\ub_i \sim p(\ub)\), i.e.\
 \begin{equation} \label{eq:omc_opt_prob}
-\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
+\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2  , \quad \ub_i \sim p(\ub).
 \end{equation}

### 2022-02-20 article/latex_files/ch2/2-3-ROMC.tex
reordered repo
final modifications

--- 2022-02-07 latex_files/ch2/2-3-ROMC.tex
@@
 only local information can be misleading. For example if the curvature
-of \(||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2\) at
+of \(||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2\) at
 \(\thetab_i^*\) is nearly flat, the curvature alone may wrongly
@@
 
-ROMC approximates the acceptance regions \(\accregioni\) and on them defines
-local proposal distributions \(q_i(\thetab)\) from which the
-posterior samples \(\thetab_{ij} \sim q_i\) are generated. The samples
-are assigned (importance) weights \(w_{ij}\) that compensate for using
-the proposal distributions \(q_i(\thetab)\) and not the prior
-\(p(\thetab)\),
+ROMC, firstly, approximates the acceptance regions \(\accregioni\),
+and defines proposal distributions \(q_i(\thetab)\) on them. The
+proposal distributions are used for generating posterior samples
+\(\thetab_{ij} \sim q_i\). The samples are assigned (importance)
+weights \(w_{ij}\) that compensate for using the proposal
+distributions \(q_i(\thetab)\) and not the prior \(p(\thetab)\),
 \begin{equation}
@@
 Given the weighted samples, any expectation
-\(\Ex_{p(\thetab|\data)}[h(\thetab)]\) of some function \(h(\thetab)\), can be approximated as
+\(\E_{p(\thetab|\data)}[h(\thetab)]\) of some function \(h(\thetab)\), can be approximated as
 \begin{equation} \label{eq:expectation}
-  \Ex_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
+  \E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
 \end{equation}
@@
 \subsubsection*{Solving the deterministic optimisation problems}
-For each set of nuisance variables \(\vb_i, i = \{1,2,\ldots,n_1 \}\), we
-search for a point \(\thetab^*_i\) such that
-\(d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon\). For notational
-convenience, we denote the distance \(d(\simulator(\thetab,\vb_i), \data)\) by
-\(d_i(\thetab)\).  Obtaining \(\theta_i^*\) involves solving the following
-optimisation problem:
+For each set of nuisance variables \(\ub_i, i = \{1,2,\ldots,n_1 \}\),
+we search for a point \(\thetab^*_i\) such that
+\(d(\simulator(\thetab^*_i,\ub_i), \data) \le \epsilon\). In general,
+\(d\) can be any valid distance, but for the rest of the paper we
+consider \(d\) as the squared Euclidean distance. For notational
+convenience, we denote \(d(\simulator(\thetab,\ub_i), \data)\) as
+\(d_i(\thetab)\).  Obtaining \(\theta_i^*\) involves solving the
+following optimisation problem:
 \begin{subequations}
@@
 \(\hat{d}_i\) can then substitute the actual distance function in
-downstream steps of the algorithms, which can lead to computational
-gains if running the simulator for evaluating the actual distance
-\(d_i(\thetab)\) is expensive.
+downstream steps of the algorithms, with possible computational gains
+especially if evaluating the actual distance \(d_i(\thetab)\) is
+expensive.
 
@@
 box \(\accregionihat\). Ideally, we want the bounding box to be as
-tight as possible to \(\mathcal{S}_i\), to ensure high acceptance rate
-in the importance sampling, but big enough for not discarding valid
-parts of \(\mathcal{S}_i\). The bounding boxes are built in two
-steps. First, we define their axes \(\mathbf{v}_d\) based on the
+tight as possible to \(\accregioni\) to ensure high acceptance rate in
+the importance sampling, but big enough for not discarding valid
+parts. The bounding boxes are built in two steps. First, we define
+their axes \(\mathbf{v}_m\), \(m = \{1, \ldots, D\}\) based on the
 (estimated) curvature of the distance at \(\thetab_i^*\). Second, we
 determine the size of the box via a one-dimensional line-search method
-along each axix, see Algorithm~\ref{alg:region_construction} for the
+along each axis, see Algorithm~\ref{alg:region_construction} for the
 details. After the bounding boxes construction, a uniform distribution

### 2022-02-20 article/latex_files/ch2/2-4-ELFI.tex
final modifications

@@
 likelihood-free inference methods, making it straightforward to
-perform comparisons between methods.
-
-% \subsubsection{Modelling at ELFI}
-% \label{sec:modelling}
-
+perform comparisons.
+%
 \pkg{ELFI} handles the probabilistic model as a Directed Acyclic Graph
@@
 parameters of interest, i.e. those we try to infer a posterior
-distribution, as the ones defined through the \code{elfi.Prior}
-class. Finally, the observations are passed to the appropriate node
-through the argument \code{observed}.
+distribution, the ones include in the \code{elfi.Prior}
+class.
 
@@
 
-% \subsubsection{Inference at ELFI}
-% \label{sec:inference-methods}
-
-All inference methods in \pkg{ELFI} share the following rules;
+All inference methods in \pkg{ELFI} share two rules;
 
@@
 \item they follow the signature \code{elfi.<Class name>(<output node>,
-    *arg)}; the initial argument is the output node model and the rest
+    *arg)}; the initial argument is the output node of the model and the rest
   of the arguments are hyper-parameters of the method.
-\item they provide a central sampling functionality, which in most
-  cases is named \\ \code{<method_name>.sample()}.
+\item they obtain posterior samples under the API \code{<method_name>.sample()}
 \end{itemize}

### 2022-02-20 article/latex_files/ch3/3-0-design-principles.tex
final modifications

@@
-For designing an extendable implementation of ROMC, we firstly define
-ROMC as a sequence of algorithmic steps. The algorithmic view of ROMC
-serves as the driver for the implementation at \pkg{ELFI}. At a high
-level, ROMC can be split into the training and the inference part. In
-general terms, the training part covers all steps for estimating the
-proposal regions and the inference part calculates the weighted
+For designing an extendable implementation, we firstly define ROMC as
+a sequence of algorithmic steps, which is the driver for the
+implementation that follows. At a high level, ROMC can be split into
+the training and the inference part. In broad view, the training part
+covers all steps for estimating the proposal regions and the inference
+part calculates the weighted
 samples. Algorithm~\ref{alg:romc_algorithm} defines ROMC formally;
-Steps 2-10 (before the horizontal line) form the training part and
-steps 12-17 the inference part, respectively.
+Steps 2-11 (before the horizontal line) form the training part and
+steps 13-18 the inference part.
 
@@
     \For{\(i \gets 1 \textrm{ to } n_1\)}
-    \State \(\vb_i \sim p(\vb)\) \Comment{Draw nuisance variables}
-    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \vb=\vb_i) \) \Comment{Define deterministic simulator}
-      \State \( d_i(\thetab) = ||g(\thetab, \vb=\vb_i) - \yb_0||_2 \) \Comment{Define distance function}
+    \State \(\ub_i \sim p(\ub)\) \Comment{Draw nuisance variables}
+    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \ub=\ub_i) \) \Comment{Define deterministic simulator}
+      \State \( d_i(\thetab) = d(g(\thetab, \ub=\ub_i), \yb_0) \) \Comment{Define distance function}
       \State \(\thetab_i^* = \text{argmin}_{\thetab} d_i\), \(d_i^*=d_i(\thetab_i^*)\) \Comment{Solve optimisation problem}
@@
       \EndIf
-      \State Estimate \(\mathcal{\hat{S}}_i\) and define \(q_i\) \Comment{Estimate proposal area}
-      \State (Optional) Fit \(\Tilde{d}_i\) on \(\mathcal{\hat{S}}_i\) \Comment{Fit surrogate model}
+      \State Estimate \(\hat{\accregioni}\) and define \(q_i\) \Comment{Estimate proposal area}
+      \State (Optional) Fit \(\Tilde{d}_i\) on \(\hat{\accregioni}\) \Comment{Fit surrogate model}
       \\\hrulefill
@@
 At the training (fitting) part, the goal is the estimation of the
-proposal regions \(\mathcal{\hat{S}}_i\). The tasks are; (a) sampling
-the nuisance variables \(\vb_i \sim p(\vb)\) for obtaining the
-deterministic functions \(d_i(\thetab)\), (b) defining the optimisation
-problems \(\min_{\thetab} \: d_i(\thetab)\), (c) obtaining
-\(\thetab_i^*\), \(d_i^*\) (d) checking whether \(d_i^* \leq \epsilon\) and
-(e) computing the proposal distribution \(q_i\).
+proposal regions \(\hat{\accregioni}\). The tasks are; (a) sampling
+the nuisance variables \(\ub_i \sim p(\ub)\) for obtaining the
+deterministic functions \(d_i(\thetab)\), (b) solving the optimisation
+problems \(\min_{\thetab} \: d_i(\thetab)\) for obtaining
+\(\thetab_i^*\), \(d_i^*\), and (c) estimate the proposal distribution
+\(q_i\).
 
@@
 evaluations of \(d_i\) for \emph{each} parameter
-\(\theta_d, d \in \{1, \ldots, D\}\)\footnote{\(D\) is the
+\(\theta_m, m \in \{1, \ldots, D\}\)\footnote{\(D\) is the
   dimensionality of \(\thetab\), i.e.  \(\thetab \in \mathbb{R}^D\)},
@@
 important step for the proposal region estimation is deciding the axes
-of the bounding box; the directions \(\vb_d, d = \{1, \ldots, D\}\) we
+of the bounding box; the directions \(\vb_m, m = \{1, \ldots, D\}\) we
 follow for reaching the boundaries. We approximate them as the
@@
 is the Jacobian matrix of the summary statistics
-\(\Phi(g(\thetab, \vb=\vb_i))\) at \(\thetab_i^*\). The approximation
+\(\Phi(g(\thetab, \ub=\ub_i))\) at \(\thetab_i^*\). The approximation
 through the Jacobian matrix has the computational advantage of using
@@
 \begin{algorithm}[!ht]
-	\caption{Approximation of \(\mathcal{\hat{S}}_i\) with a bounding box; 
+	\caption{Approximation \(\accregioni\) with a bounding box \(\hat{\accregioni}\); 
   Requires: a model of distance \(d_i(\thetab)\), 
@@
 	\begin{algorithmic}[1]
-	\State Compute eigenvectors \(\vb_{d}\) of \(\hess_i\) {\scriptsize (\(d = 1,\ldots,||\thetab ||)\)}
-	\For{\(d \gets 1 \textrm{ to } D\)}
+	\State Compute eigenvectors \(\vb_{m}\) of \(\hess_i\) {\scriptsize (\(d = 1,\ldots,D)\)}
+	\For{\(m \gets 1 \textrm{ to } D\)}
 		\State \(\Tilde{\thetab} \gets \thetab_i^*\) \label{algstep:box_constr_start}
@@
         	\Repeat
-            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{d}\) \Comment{Large step size \(\eta\).}
+            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{m}\) \Comment{Large step size \(\eta\).}
             \State \(j \gets j + 1\)
-        	\Until{\(d(f_i(\Tilde{\thetab}), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
-        	\State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{d}\)
+        	\Until{\(d(g(\Tilde{\thetab}, \ub=\ub_i), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
+        	\State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{m}\)
         	\State \(\eta \gets \eta/2\) \Comment{More accurate region boundary}
@@
       \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
-        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_d\) \Comment{Then, make the minimum step}
+        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \ub_m\) \Comment{Then, make the minimum step}
       \EndIf \label{algstep:box_constr_end}
-    	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\mathbf{v}_{d}\)
-    	\State Runc steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\mathbf{v}_{d} = - \mathbf{v}_{d}\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{d}\)
+    	\State Set \(\Tilde{\thetab}\) as the positive end point along \(\vb_m\)
+    	\State Runc steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\vb_m = - \vb_m\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{m}\)
 	\EndFor

### 2022-02-20 article/latex_files/ch3/3-1-implementation-principles.tex
final modifications

--- 2022-02-07 latex_files/ch3/3-1-implementation-principles.tex
@@
 represent internal (private) functions, whereas the rest (blue
-ellipses) are the methods exposed as the public API. In
+rectangles) are the methods exposed as the public API. In
 Figure~\ref{fig:romc_overview}, it can be easily observed that the
@@
 
-ROMC has the significant advantage of being fully parallelisable. We
-exploit this fact by implementing a parallel version of all the major
-components of the method; (a) solving the optimisation problems, (b)
-constructing bounding box regions, (c) sampling and (d) evaluating the
-posterior. We parallelise these processes using the built-in
-\proglang{Python} package \pkg{multiprocessing}. The specific package
-enables concurrency, using subprocesses instead of threads, for
-side-stepping the Global Interpreter (GIL). For activating the
-parallel version of the algorithm, the user simpy has to pass the
-argument \code{parallelize=True} when instantiating \code{ROMC}.
+As discussed, ROMC has the significant advantage of being fully
+parallelisable. We exploit this fact by implementing a parallel
+version of the major fitting components; (a) solving the optimisation
+problems, (b) constructing bounding box regions. We parallelise these
+processes using the built-in \proglang{Python} package
+\pkg{multiprocessing}. The specific package enables concurrency, using
+subprocesses instead of threads, for side-stepping the Global
+Interpreter (GIL). For activating the parallel version of the
+algorithm, the user simpy has to pass the argument
+\code{parallelize=True} when instantiating \code{ROMC}.
 

### 2022-02-20 article/latex_files/ch4/4-1-training.tex
final modifications

@@
   seed = 21 # seed for solving the optimisation problems
-  eps = .75 
   use_bo = False # set to True for switching to Bayesian optimisation
@@
 
-  eps = .75 # threshold for the bounding box based on histogram inspection
+  eps = .75 # threshold for the bounding box 
   romc.estimate_regions(eps=eps) # build the bounding boxes

### 2022-02-20 article/latex_files/ch4/4-4-extensibility.tex
final modifications

--- 2022-02-14 latex/latex_files/ch4/4-4-extensibility.tex
@@
-ROMC can be conceived as a meta-algorithm; it describes a sequence of
-steps for approximating the posterior distribution, without
-\emph{explicitly} enforcing the methods that solve these steps. Even
-though for each step a specific algorithm is proposed
-by~\cite{Ikonomov2019}, in general ROMC is functional if a practitioner
-thinks of an alternative way of approaching a specific
+ROMC describes a sequence of steps for approximating the posterior
+distribution, without explicitly enforcing the methods that solve
+these steps. Even though for each step a specific algorithm is
+proposed by~\cite{Ikonomov2019}, in general ROMC is functional if a
+practitioner thinks of an alternative way of approaching a specific
 step. Considering this particularity, we designed the implementation
-to support such \emph{extensibility} capability.
+to support extensibility.
 
@@
 ROMC procedure, whereas the rest of the code is the backbone of the
-algorithm; each of them can be swapped to a different
-algorithm with the same functionality than the one we propose, without the rest of the method to
-collapse.
+algorithm.
 
@@
 \linebreak (c) \code{build_region(**kwargs)}, (d)
-\code{fit_local_surrogate(**kwargs)}. Therefore, the user should
-create a custom class that inherits the basic
-\code{OptimisatioProblem} class and overwrites one or more of the
-above four functions with custom ones.
-
-We illustrate that in the
-following example; suppose we want to blindly set the proposal region
-as a predefined n-dimensional bounding box around the optimal
-point. The following snippet illustrates how:
+\code{fit_local_surrogate(**kwargs)}. The user can create a custom
+class that inherits the basic \code{OptimisatioProblem} and, then,
+overwrite any of the four functions with custom ones. Suppose a user
+wants to fit Neural Networks as local surrogate models
+\(\tilde{d}_i\). The user should overwrite the
+\code{fit_local_surrogate(**kwargs)} function with one that internally
+trains a neural network. In the following snippet we illustrate that,
+using the \code{neural_network.MLPRegressor} class of the
+\pkg{scikit-learn} package.
 
@@
 ---------------------------------- python ----------------------------------
-class customOptim(OptimisationProblem):
+class CustomOptim(OptimisationProblem):
     def __init__(self, **kwargs):
-        super(customOptim, self).__init__(**kwargs)
+        super(CustomOptim, self).__init__(**kwargs)
 
@@
         objective = self.objective
+
+        # helper function
+        def local_surrogate(theta, model_scikit):
+            assert theta.ndim == 1
+            theta = np.expand_dims(theta, 0)
+            return float(model_scikit.predict(theta))
+
+        # create local surrogate model as a function of theta
+        def create_local_surrogate(model):
+            return partial(local_surrogate, model_scikit=model)
 
@@
             model = model.fit(x, y)
-            local_surrogates.append(self.create_local_surrogate(model))
+            local_surrogates.append(create_local_surrogate(model))
 
@@
         self.state["local_surrogates"] = True
-
-    def create_local_surrogate(self, model):
-        def _local_surrogate(th):
-            th = np.expand_dims(th, 0)
-            return float(model.predict(th))
-        return _local_surrogate
 ----------------------------------------------------------------------------
@@
 class level. In the following four snippets we present the signature
-of each method, pointing out with a left arrow the side-effects that
-must be included in the custom routine that will replace them.
+of each method. We name the class-level variables that must be set
+under the comment \code{\# side-effects}.
 
@@
 def solve_gradients(self, **kwargs):
+  # custom solution procedure
+  x = ...
+  y = ...
+  jac = ...
+  hess_inv = ...
+
+  # side-effects
   self.state["attempted"] = True
-
-  # custom solution procedure
-
   if success:
-    self.result = RomcOpimisationResult(res.x, res.y, jac, hess_inv)
+    self.result = RomcOpimisationResult(x, y, jac, hess_inv)
     self.state["solved"] = True
@@
 def solve_bo(self, **kwargs):
+  # custom procedure
+  x = ...
+  y = ...
+  custom_surrogate = ...
+
+  # side-effects
   self.state["attempted"] = True
-
-  # custom procedure
-
   if success:
     self.result = RomcOpimisationResult(x, y)
-    self.surrogate <- Callable
+    self.surrogate = custom_surrogate
     self.state["solved"] = True
@@
 def build_region(self, **kwargs):
+  # custom build_region method
+  bounding_box: List[NDimBoudningBox] = ...
+  success = True/False # whether region built correctly
+
+  # side-effects
   self.eps_region = eps_region
-
   if success:
     # construct region
-        self.regions <- List[NDimBoudningBox]
+    self.regions = bounding_box
     self.state["region"] = True
@@
 def fit_local_surrogate(self, **kwargs):
+  # custom local surrogates
+  custom_surrogates = ...
+  success = True/False # whether local surrogates fit correctly
+
+  # side-effects
   if success:

### 2022-02-20 replication_script/replication_script.py

@@ -681,11 +681,13 @@ class CustomOptim(OptimisationProblem):
         nof_samples = 500
         objective = self.objective
 
+        # helper function
         def local_surrogate(theta, model_scikit):
             assert theta.ndim == 1
             theta = np.expand_dims(theta, 0)
             return float(model_scikit.predict(theta))
 
+        # create local surrogate model as a function of theta
         def create_local_surrogate(model):
             return partial(local_surrogate, model_scikit=model)
 

### 2022-02-20 article/article.tex
final changes

@@
 \newcommand{\vb}{\mathbf{v}}
-\newcommand{\ub}{\mathbf{u}}
 \newcommand{\yb}{\mathbf{y}}
@@
   provides accurate weighted samples from the posterior. Our
-  implementation can be used in two ways. First, a scientist may use
+  implementation can be used in two ways. Firstly, a scientist may use
   it as an out-of-the-box LFI algorithm; we provide an easy-to-use API
-  harmonised with the principles of \pkg{ELFI}, enabling effortless
-  comparisons with the rest of the methods included in the
-  package. Additionally, we have carefully split ROMC into isolated
-  components for supporting extensibility. A researcher may experiment
-  with novel method(s) for solving part(s) of ROMC without
-  reimplementing everything from scratch. In both scenarios, all steps
-  run in a fully-parallelisable manner, exploiting all CPU
-  cores. We also provide helpful functionalities for (i) inspecting
-  the inference process and (ii) evaluating the obtained
+  harmonised with the rest of the \pkg{ELFI} framework for enabling
+  effortless comparisons with the rest of the LFI methods implemented
+  there. Additionally, we support extensibility by splitting ROMC into
+  isolated components. A researcher may experiment with novel
+  method(s) for solving specific subpart(s) of ROMC without
+  reimplementing everything from scratch. In both scenarios, the
+  algorithm runs in a fully-parallelisable manner for exploiting all
+  CPU cores. We also provide helpful functionalities for (i)
+  inspecting the inference process and (ii) evaluating the obtained
   samples. Finally, we test the robustness of our implementation on
-  some typical LFI examples.}
+  some typical LFI examples.  }
 
@@
 %% -- Design Principles -----------------------------------------------------
-\section{Implementation}
-This section is split in two parts. We first express ROMC as an
-algorithm and then we present the general implementation principles we
-follow.
-
-\subsection{Algotirhmic view of ROMC}
+\section{Design principles}  % better title needed
 \input{latex_files/ch3/3-0-design-principles.tex}
@@
 \section*{Computational details}
+
 The results in this paper were obtained using \proglang{Python}~3.7.9
@@
 
-HP was funded by European Research Council grant 742158 (SCARABEE,
+Henri Pesonen was funded by European Research Council grant 742158 (SCARABEE,
 Scalable inference algorithms for Bayesian evolutionary epidemiology).

### 2022-02-20 article/article.tex

--- 2022-02-20 article/article.tex
@@
   samples. Finally, we test the robustness of our implementation on
-  some typical LFI examples.}
+  some typical LFI examples.  }
 

### 2022-02-20 article/latex_files/ch1/1-introduction.tex

@@
 Simulator-based models are particularly captivating due to the
-modelling freedom they provide. In essence, a simulator-based model can
-describe any data generating mechanism that can be written as a finite
-set of algorithmic steps. This modelling freedom comes at a cost;
-performing the inference, i.e., sample or evaluate the posterior
-distribution, is challenging.
+modelling freedom they provide. In essence, a simulator-based model
+can describe any data generating mechanism that can be written as a
+finite set of algorithmic steps. Such freedom comes at a cost;
+performing the inference, i.e., sampling or evaluating the posterior,
+is challenging.
 
-Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} was a
-novel LFI approach for approximating the posterior
-distribution. Afterwards, \citet{Forneron2016} provided a similar
-method naming it `reverse sampler'. The central idea is turning the
-stochastic data-generating mechanism into a set of
-fully-parallelizable deterministic optimisation processes. In their
-work, ~\citet{Ikonomov2019}, located some critical failure modes of
-OMC, and proposed Robust OMC (ROMC) an alternative version of OMC with
-the appropriate improvements.
+Optimisation Monte Carlo (OMC) proposed by~\citet{Meeds2015} is a
+novel LFI approach for approximating the posterior distribution. The
+central idea is turning the stochastic data-generating mechanism into
+a set of deterministic optimisation processes. Afterwards,
+\citet{Forneron2016} provided a similar method under the name `reverse
+sampler'. In their work, ~\citet{Ikonomov2019}, located some critical
+limitations of OMC. They proposed Robust OMC (ROMC), an alternative
+version of OMC with appropriate modifications.
 
-In this paper, we present an extendible and parallelisable
-implementation of ROMC at the \proglang{Python} package \pkg{Engine
-  for Likelihood-Free inference (ELFI)}. We choose carefully the
-designing principles for making the implementation extendible. As we
-describe analytically in the next chapter, ROMC can be understood not
-only as a specific algorithm, but mainly as a framework for obtaining
-weighted samples from the posterior; it defines a sequence of
-algorithmic steps, without enforcing a specific algorithm for solving
-them. Therefore, a researcher may adopt ROMC's backbone approach, but
-experiment with alternative methods to solve the tasks. We have
-designed ROMC for facilitating such alterations. Finally, we have
-tested the accuracy and the efficiency of our implementation on the
-LFI examples supported by the \pkg{ELFI} package.
+In this paper, we present an implementation of ROMC at the
+\proglang{Python} package \pkg{Engine for Likelihood-Free inference
+(ELFI)}. We follow appropriate design principles for ensuring
+extensibility. As we describe analytically in the next chapter, ROMC
+is a general framework for obtaining weighted samples from the
+posterior; it defines a sequence of algorithmic steps without
+enforcing a specific algorithm for solving each step\footnote{For
+being a ready-to-use algorithm,~\citet{Ikonomov2019} proposed a
+default method for each step, but this choice is by no means
+restrictive.}. A researcher may adopt ROMC and develop novel methods
+to solve each task. We have designed our software for facilitating
+such alterations. Finally, we have tested the accuracy and the
+efficiency of our implementation on some LFI examples supported by the
+\pkg{ELFI} package.
 
 To the best of our knowledge, this is the first attempt to implement
-the ROMC inference method to a generic LFI framework. Therefore, there
-are not other implementation directly comparable to ours. For this
-reason, we test the implementation against (i) an artificial example
-with tractable likelihood and (ii) the second-order moving average
-(MA2) example from the \pkg{ELFI} package. In the latter, we generate
-data using known parameters and we compare our results with the
-typical Rejection ABC approach.
+the ROMC inference method to a generic LFI framework; there are no
+other packages for making direct comparisons. Therefore, we test it
+against (i) an artificial example with tractable likelihood and (ii)
+the second-order moving average (MA2) example from the \pkg{ELFI}
+package. In the latter, we generate data using known parameters and
+compare our results with the typical Rejection ABC~\citet{Beaumont2002}.

### 2022-02-20 article/latex_files/ch2/2-1-implicit-models.tex

@@
 \(\{ p(\yb|\thetab) \}_{\thetab}\) whose closed-form is either unknown
-or computationally intractable. Practically, a simulator
-\( M_r(\thetab) \) can be understood as a random black-box machine
-(computer code) that generates samples \(\yb\) in a stochastic manner
-for any given a set of parameters \(\thetab\), i.e.
+or computationally intractable. Practically, in these scenarios, we
+can only a access a simulator \( M_r(\thetab) \), i.e. a random
+black-box machine (computer code) that generates samples \(\yb\) in a
+stochastic manner for any given a set of parameters \(\thetab\);
 \( M_r(\thetab) \rightarrow \yb \). We can isolate the randomness by
-introducing the stochastic nuisance variables \(\vb \sim p(\vb)\) so
+introducing the stochastic nuisance variables \(\ub \sim p(\ub)\) so
 that the simulator becomes a deterministic mapping \(\simulator\) that
-maps \((\thetab, \vb)\) to the data \(\yb\), i.e.\
-\(\yb=\simulator(\thetab,\vb)\).  In terms of computer code,
-\(p(\vb)\) is defined as a random number generating process within the
-computer code.  % Importantly, we can fix the value of \(\vb\) while
-% changing \(\thetab\) by fixing the random seed used.
+maps \((\thetab, \ub)\) to the data \(\yb\), i.e.\
+\(\yb=\simulator(\thetab,\ub)\). Within the computer code, the
+distribution \(p(\ub)\) is defined as the random number generating
+process.
 
@@
 The modelling freedom of simulator-based models, however, comes at the
-price of difficulties in inferring their parameters. Denoting observed
-data by \(\data\), the main difficulty is that the likelihood function
-\(L(\thetab) = p(\data|\thetab)\) is generally intractable. To better
-see the sources of the intractability, and to address them, we go back
-to the basic characterisation of the likelihood as the (rescaled)
-probability of a parameter of the model to generate data \(\yb\) that
-is similar to the observed data \(\data\). More formally, the
-likelihood \(L(\thetab)\) equals
+price of difficulties in inferring their parameters. Denoting the
+observed data as \(\data\), the main difficulty is that the likelihood
+function \(L(\thetab) = p(\data|\thetab)\) is generally
+intractable. To better see the sources of the intractability, and to
+address them, we go back to the basic characterisation of the
+likelihood as the (rescaled) probability of a parameter of the model
+to generate data \(\yb\) that is similar to the observed data
+\(\data\). More formally, the likelihood \(L(\thetab)\) equals
 \begin{equation} \label{eq:likelihood}
   L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
-  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data)  \mid \theta)
+  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
 \end{equation}
@@
 of the probability
-\(\Pr(\simulator(\thetab,\vb) \in \region(\data))\), the second is the
+\(\Pr(\simulator(\thetab,\ub) \in \region(\data))\), the second is the
 limit of \(\epsilon \to 0\).  Approximating the probability with
@@
 \(\epsilon > 0\), which leads to the approximate likelihood function
-\(L_{d, \epsilon}(\thetab)\),
+\(L_{d, \epsilon}(\thetab)\)
 \begin{equation} \label{eq:approx_likelihood}
@@
 
-and, in turn, to the approximate posterior,
+and, in turn, to the approximate posterior
 

### 2022-02-20 article/latex_files/ch4/4-2-inference.tex

@@
 remained after filtering\footnote{From the \(n_1\) optimisation
-  problems, only the ones with \(d_i(\thetab_*) < \epsilon\) are
+  problems, only the ones with \(d_i(\thetab^*) < \epsilon\) are
   maintained for building a bounding box}. The samples are drawn from
@@
 expression~\eqref{eq:expectation}. The argument \code{h} can be any
-python \code{Callable}. Method \code{eval_unnorm_posterior} computes
-the unnormalised posterior approximation using
-expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
+python \code{Callable}. The method \code{eval_unnorm_posterior} computes
+the unnormalised posterior approximation using the
+expression~\eqref{eq:approx_posterior}. The method \code{eval_posterior}
 evaluates the normalised posterior estimating the partition function
@@
 integral approximation. The approximation is computationally feasible
-only in low-dimensional parametric spaces.
+only in a low-dimensional parametric space.
 

### 2022-02-20 article/latex_files/ch5/5-ma2.tex
small change

@@
 
-% \begin{figure}[ht]
-%     \begin{center}
-%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter4/mae2_prior_samples.png}
-%     \end{center}
-%     \caption[MA2 example, prior distribution.]{Prior distribution proposed by \cite{Marin2012}. The
-%       samples follow a triangular shape.}
-%   \label{fig:ma2_1}
-% \end{figure}
-
 \noindent
@@
 
-\begin{gather} \label{eq:ma2_summary_1}
-  s_1(\yb) = \frac{1}{T-1} \sum_{t=2}^T y_ty_{t-1}\\ \label{eq:ma2_summary_2}
-  s_2(\yb) = \frac{1}{T-2} \sum_{t=3}^T y_ty_{t-2} \\
-  s(\yb) = (s_1(\yb), s_2(\yb))\\ \label{eq:ma2_summary_4}
-  d = ||s(\yb) - s(\yb_0)||_2^2 
+\begin{gather}
+  \label{eq:ma2_summary_1} s_1(\yb) = \frac{1}{T-1} \sum_{t=2}^T y_ty_{t-1}\\
+  \label{eq:ma2_summary_2} s_2(\yb) = \frac{1}{T-2} \sum_{t=3}^T y_ty_{t-2} \\
+  s(\yb) = (s_1(\yb), s_2(\yb))\\
+  \label{eq:ma2_summary_4} d = ||s(\yb) - s(\yb_0)||_2^2
 \end{gather}
@@
 \hline
-ROMC (gradient-based) & 0.503 & 0.143 & 0.032 & 0.17 \\
+ROMC (gradient-based) & 0.501 & 0.142 & 0.033 & 0.169 \\
 \hline
@@
 \hline
-ROMC (Neural Network) & 0.503 & 0.140 & 0.03 & 0.171 \\
+ROMC (Neural Network) & 0.491 & 0.138 & 0.04 & 0.172 \\
 \hline
@@
     \begin{center}
-        % \input{./latex_files/images/chapter4/ma2_region_3.tex}
-        % \input{./latex_files/images/chapter4/ma2_region_3_bo.tex}
         \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
@@
 Figure~\ref{fig:exec_parallel} we observe the execution times for
-performing the inference on the ma2 model; the parallel version
-performs both tasks almost six times faster than the sequential. The
+performing the inference on the MA2 model; the parallel version
+performs both tasks almost five times faster than the sequential. The
 result is reasonable given that the experiments have run in a single
@@
     \caption[Execution time exploiting parallelisation]{Comparison
-      between parallel and sequential execution of the fitting part of
-      ROMC. We observe that the parallel version runs between 2.5 an 6
-      times faster.}
+      between parallel and sequential execution of ROMC. We observe
+      that the parallel version runs almost 5 times faster.}
       \label{fig:exec_parallel}

### 2022-02-20 article/latex_files/ch4/4-4-extensibility.tex
final version

@@
 \code{fit_local_surrogate(**kwargs)} function with one that internally
-trains a neural network. In the following snippet we illustrate that,
-using the \code{neural_network.MLPRegressor} class of the
+trains a neural network. In the following snippet we illustrate how to
+do that, using the \code{neural_network.MLPRegressor} class of the
 \pkg{scikit-learn} package.
@@
 In the same way, the user may replace each of the other three
-functionalities. The only restriction that must be respected, concerns
+functionalities. The only restriction that must be respected concerns
 the side-effects each method has at the \code{OptimisatioProblem}
 class level. In the following four snippets we present the signature
-of each method. We name the class-level variables that must be set
+of each method and we name the class-level variables that must be set
 under the comment \code{\# side-effects}.
@@
 \end{Code}
-
-% Let's say we have observed that the local area around $\theta_i^*$ is
-% too complex to be represented by a simple quadratic model\footnote{as
-%   in the current implementation}. Hence, the user selects a neural
-% network as a good alternative. In the following snippet, we
-% demonstrate how they could implement this enhancement without much
-% effort; (a) they have to develop the neural network using the package
-% of their choice (b) they must create a custom optimisation class which
-% inherits the basic \code{OptimisationClass} and (c) they have to
-% overwrite the \code{fit_local_surrogate} routine, with one that
-% sets the neural network's prediction function as the
-% \code{local_surrogate} attribute. The argument \code{**kwargs}
-% may be used for passing all the important arguments, e.g.\ training
-% epochs, gradient step etc. If, for example, they would like to set the
-% size of the training set dynamically, we may replace \code{x =
-%   self.regions[0].sample(30)} with \code{x =
-%   self.regions[0].sample(kwargs["nof_examples"])}. Finally, they must
-% pass the custom optimisation class, when calling the \code{ROMC}
-% method.
-
-% \begin{Code}
-% ------------------------------ python snippet ------------------------------  
-%   class NeuralNetwork:
-%       def __init__(self, **kwargs):
-%           # set the input arguments
-
-%       def train(x, y):
-%           # training code
-
-%       def predict(x):
-%           # prediction code
-
-%   # Inherit the base optimisation class
-%   class customOptim(elfi.methods.parameter_inference.OptimisationProblem):
-%       def __init__(self, **kwargs):
-%           super(customOptim, self).__init__(**kwargs)
-
-%       # overwrite the function you want to replace
-%       def fit_local_surrogate(**kwargs):
-%           # init and train the NN
-%           x = self.regions[0].sample(30) # 30 training points
-%           y = [np.array([self.objective(ii) for ii in x])]
-%           nn = NeuralNet()
-%           nn.train(x,y)
-
-%           # set the appropriate attribute
-%           self.local_surrogate = nn.predict
-
-%           # update the state
-%           self.state["local_surrogate"] = True
-
-%   # pass the custom inference method as argument
-%   romc = elfi.ROMC(dist, bounds, custom_optim_class=customOptim)
-% ----------------------------------------------------------------------------
-% \end{Code}

### 2022-08-28 article/article.tex
added refs.bib

@@
   Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo
-}
+
 \Plainauthor{Vasilis Gkolemis, Michael Gutmann, Henri Pesonen}

### 2022-12-21 article/article.tex
fixed references

@@
   Michael Gutmann~\orcidlink{0000-0002-5329-9910}\\University of Edinburgh \And
-  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo
+  Henri Pesonen~\orcidlink{0000-0003-4500-2926}\\University of Oslo}
 

### 2022-12-21 article/latex_files/ch1/1-introduction.tex
fixed references

@@
 
-To the best of our knowledge, this is the first attempt to implement
-the ROMC inference method to a generic LFI framework; there are no
-other packages for making direct comparisons. Therefore, we test it
-against (i) an artificial example with tractable likelihood and (ii)
-the second-order moving average (MA2) example from the \pkg{ELFI}
+To the best of our knowledge, this is the first implementation of the
+ROMC inference method to a generic LFI framework, so there are no
+other packages for direct comparisons. Therefore, we test ROMC against
+(i) an artificial example with tractable likelihood and (ii) the
+second-order moving average (MA2) example from the \pkg{ELFI}
 package. In the latter, we generate data using known parameters and
-compare our results with the typical Rejection ABC~\citet{Beaumont2002}.
+compare our results with the typical Rejection
+ABC~\citet{lintusaari2017}.

### 2022-12-21 article/latex_files/ch2/2-1-implicit-models.tex
fixed references

@@
 that we can sample data points, but we cannot evaluate the
-likelihood. Formally, a simulator-based model is a parameterised
+likelihood. formally, a simulator-based model is a parameterised
 family of probability density functions
 \(\{ p(\yb|\thetab) \}_{\thetab}\) whose closed-form is either unknown
-or computationally intractable. Practically, in these scenarios, we
-can only a access a simulator \( M_r(\thetab) \), i.e. a random
+or computationally intractable. practically, in these scenarios, we
+can only a access a simulator \( m_r(\thetab) \), i.e. a random
 black-box machine (computer code) that generates samples \(\yb\) in a
 stochastic manner for any given a set of parameters \(\thetab\);
-\( M_r(\thetab) \rightarrow \yb \). We can isolate the randomness by
+\( m_r(\thetab) \rightarrow \yb \). we can isolate the randomness by
 introducing the stochastic nuisance variables \(\ub \sim p(\ub)\) so
@@
 maps \((\thetab, \ub)\) to the data \(\yb\), i.e.\
-\(\yb=\simulator(\thetab,\ub)\). Within the computer code, the
+\(\yb=\simulator(\thetab,\ub)\). within the computer code, the
 distribution \(p(\ub)\) is defined as the random number generating
@@
 
-Simulator-based models provide considerable modelling freedom; any
+simulator-based models provide considerable modelling freedom; any
 physical process that can be conceptualised as a computer program of
 finite steps can be modelled as a simulator-based model without any
-compromise. The modelling freedom allows for any amount of hidden
-(unobserved) internal variables or rule-based decisions. Hence,
+compromise. the modelling freedom allows for any amount of hidden
+(unobserved) internal variables or rule-based decisions. hence,
 implicit models are often used to model physical phenomena in the
 natural sciences such as e.g.~genetics, epidemiology, or neuroscience. 
-Further background on simulator-based models and
+further background on simulator-based models and
 example applications can be found in the articles
-by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.
+by \citet{gutmann2016, lintusaari2017, sisson2018, cranmer2020}.
 
-The modelling freedom of simulator-based models, however, comes at the
-price of difficulties in inferring their parameters. Denoting the
+the modelling freedom of simulator-based models, however, comes at the
+price of difficulties in inferring their parameters. denoting the
 observed data as \(\data\), the main difficulty is that the likelihood
-function \(L(\thetab) = p(\data|\thetab)\) is generally
-intractable. To better see the sources of the intractability, and to
+function \(l(\thetab) = p(\data|\thetab)\) is generally
+intractable. to better see the sources of the intractability, and to
 address them, we go back to the basic characterisation of the
@@
 to generate data \(\yb\) that is similar to the observed data
-\(\data\). More formally, the likelihood \(L(\thetab)\) equals
+\(\data\). more formally, the likelihood \(l(\thetab)\) equals
 \begin{equation} \label{eq:likelihood}
-  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
+  l(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in b_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
   \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
@@
 
-The basic characterisation of the likelihood in \eqref{eq:likelihood}
-highlights two sources of intractability: The first is the computation
+the basic characterisation of the likelihood in \eqref{eq:likelihood}
+highlights two sources of intractability: the first is the computation
 of the probability
 \(\Pr(\simulator(\thetab,\ub) \in \region(\data))\), the second is the
-limit of \(\epsilon \to 0\).  Approximating the probability with
+limit of \(\epsilon \to 0\).  approximating the probability with
 samples becomes computationally infeasible if \(\epsilon\) is too
-small. Hence, a large class of inference methods work with
+small. hence, a large class of inference methods work with
 \(\epsilon > 0\), which leads to the approximate likelihood function
-\(L_{d, \epsilon}(\thetab)\)
+\(l_{d, \epsilon}(\thetab)\)
 \begin{equation} \label{eq:approx_likelihood}
-  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
+  l_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
 \end{equation}
@@
 \begin{equation} \label{eq:approx_posterior}
-  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
+  p_{d,\epsilon}(\thetab|\data) \propto l_{d, \epsilon}(\thetab) p(\thetab).
 \end{equation}
 
-The approximation in \eqref{eq:approx_likelihood} is by no means the
+the approximation in \eqref{eq:approx_likelihood} is by no means the
 only strategy to deal with the intractabilities of the likelihood
-function in \eqref{eq:likelihood}. Other strategies include modelling
+function in \eqref{eq:likelihood}. other strategies include modelling
 the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
 reverse, or framing likelihood-free inference as a ratio estimation
-problem, see for example \citet{Blum2010, Wood2010, Papamakarios2016,
+problem, see for example \citet{blum2010, Wood2006, Papamakarios2016,
   Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both

### 2022-12-21 replication_script/replication_script.py

@@ -17,7 +17,7 @@ import tikzplotlib as tplt
 import matplotlib
 matplotlib.rcParams['text.usetex'] = True
 logging.basicConfig(level=logging.INFO)
-savefig = True
+savefig = False
 
 ######################################################################
 # Part 1. Generates the subfigures used in the running example in the#
@@ -50,7 +50,7 @@ class Likelihood:
 
         """
         assert isinstance(theta, np.ndarray)
-        theta = theta.astype(np.float)
+        theta = theta.astype(float)
         samples = np.empty_like(theta)
 
         c = 0.5 - 0.5 ** 4
@@ -83,7 +83,7 @@ class Likelihood:
 
         BS = theta.shape[0]
         N = x.shape[0]
-        theta = theta.astype(np.float)
+        theta = theta.astype(float)
 
         pdf_eval = np.zeros((BS))
         c = 0.5 - 0.5 ** 4

### 2022-12-21 article/latex_files/ch2/2-1-implicit-models.tex
fixed spaces and commas

--- 2022-02-20 article/latex_files/ch2/2-1-implicit-models.tex
@@
 \(\{ p(\yb|\thetab) \}_{\thetab}\) whose closed-form is either unknown
-or computationally intractable. Practically, in these scenarios, we
-can only a access a simulator \( M_r(\thetab) \), i.e. a random
+or computationally intractable. In these scenarios, we
+can only a access a simulator \( m_r(\thetab) \), i.e., a random
 black-box machine (computer code) that generates samples \(\yb\) in a
 stochastic manner for any given a set of parameters \(\thetab\);
-\( M_r(\thetab) \rightarrow \yb \). We can isolate the randomness by
+\( m_r(\thetab) \rightarrow \yb \). We can isolate the randomness by
 introducing the stochastic nuisance variables \(\ub \sim p(\ub)\) so
 that the simulator becomes a deterministic mapping \(\simulator\) that
-maps \((\thetab, \ub)\) to the data \(\yb\), i.e.\
+maps \((\thetab, \ub)\) to the data \(\yb\), i.e., \
 \(\yb=\simulator(\thetab,\ub)\). Within the computer code, the
@@
 implicit models are often used to model physical phenomena in the
-natural sciences such as e.g.~genetics, epidemiology, or neuroscience. 
-Further background on simulator-based models and
+natural sciences such as, e.g., genetics, epidemiology or
+neuroscience. Further background on simulator-based models and
 example applications can be found in the articles
-by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.
+by~\citet{gutmann2016, lintusaari2017, sisson2018, cranmer2020}.
 
@@
 observed data as \(\data\), the main difficulty is that the likelihood
-function \(L(\thetab) = p(\data|\thetab)\) is generally
+function \(l(\thetab) = p(\data|\thetab)\) is generally
 intractable. To better see the sources of the intractability, and to
@@
 to generate data \(\yb\) that is similar to the observed data
-\(\data\). More formally, the likelihood \(L(\thetab)\) equals
+\(\data\). More formally, the likelihood \(l(\thetab)\) equals
 \begin{equation} \label{eq:likelihood}
-  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
+  l(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in b_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
   \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
@@
 \(\epsilon\) and \(\region(\data)\) is an \(\epsilon\) region around \(\data\)
-that is defined via a distance function \(d\), i.e.\ \(\region(\data)
+that is defined via a distance function \(d\), i.e., \ \(\region(\data)
 := \{\yb: d(\yb, \data) \leq \epsilon \}\).
 
-The basic characterisation of the likelihood in \eqref{eq:likelihood}
-highlights two sources of intractability: The first is the computation
+The basic characterisation of the likelihood in~\eqref{eq:likelihood}
+highlights two sources of intractability; the first is the computation
 of the probability
@@
 \(\epsilon > 0\), which leads to the approximate likelihood function
-\(L_{d, \epsilon}(\thetab)\)
+\(l_{d, \epsilon}(\thetab)\)
 \begin{equation} \label{eq:approx_likelihood}
-  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
+  l_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
 \end{equation}
@@
 \begin{equation} \label{eq:approx_posterior}
-  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
+  p_{d,\epsilon}(\thetab|\data) \propto l_{d, \epsilon}(\thetab) p(\thetab).
 \end{equation}
 
-The approximation in \eqref{eq:approx_likelihood} is by no means the
+The approximation in~\eqref{eq:approx_likelihood} is by no means the
 only strategy to deal with the intractabilities of the likelihood
-function in \eqref{eq:likelihood}. Other strategies include modelling
+function in~\eqref{eq:likelihood}. Other strategies include modelling
 the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
 reverse, or framing likelihood-free inference as a ratio estimation
-problem, see for example \citet{Blum2010, Wood2010, Papamakarios2016,
+problem, see for example \citet{blum2010, Wood2006, Papamakarios2016,
   Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both
 OMC and robust OMC, which we introduce next, are based on the
-approximation in \eqref{eq:approx_likelihood}.
+approximation in~\eqref{eq:approx_likelihood}.

### 2022-12-21 article/latex_files/ch2/2-2-OMC.tex
fixed spaces and commas

@@
 who base their explanation of OMC on the approximate likelihood
-function in \eqref{eq:approx_likelihood}. With the indicator function
+function in~\eqref{eq:approx_likelihood}. With the indicator function
 %
@@
 \(\thetab_i^*\) that achieve the smallest distance between observed
-and simulated data for each realisation \(\ub_i \sim p(\ub)\), i.e.\
+and simulated data for each realisation \(\ub_i \sim p(\ub)\), i.e.,\
 \begin{equation} \label{eq:omc_opt_prob}

### 2022-12-21 article/latex_files/ch3/3-0-design-principles.tex
final version
fixed spaces and commas

@@
 \(\theta_m, m \in \{1, \ldots, D\}\)\footnote{\(D\) is the
-  dimensionality of \(\thetab\), i.e.  \(\thetab \in \mathbb{R}^D\)},
+  dimensionality of \(\thetab\), i.e.,  \(\thetab \in \mathbb{R}^D\)},
 which works in low-dimensional problems. If \(d_i(\thetab)\) is not
@@
       \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
-        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \ub_m\) \Comment{Then, make the minimum step}
+        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_m\) \Comment{Then, make the minimum step}
       \EndIf \label{algstep:box_constr_end}

### 2022-12-21 article/latex_files/ch3/3-1-implementation-principles.tex
fixed spaces and commas

@@
 training part includes all the steps until the computation of the
-proposal regions i.e.~sampling the nuisance variables, defining the
+proposal regions, i.e., ~sampling the nuisance variables, defining the
 optimisation problems, solving them, constructing the regions and
@@
   available; However, this functionality is useful in cases where the
-  posterior can be computed numerically or with an alternative method
-  (i.e.\ ABC Rejection Sampling), and we would like to measure the
+  posterior can be computed numerically or with an alternative method,
+  i.e., ABC Rejection Sampling), and we would like to measure the
   discrepancy between the two approximations.}
@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> elfi.ROMC(<output_node>, parallelize=True)
-----------------------------------------------------------------------------
+elfi.ROMC(<output_node>, parallelize = True)
 \end{Code}
@@
 
-In the following code snippet, we code the model at \pkg{ELFI}.
+In the following code snippet, we code the model at \pkg{ELFI}:
 
 \begin{Code}
------------------------------- python snippet ------------------------------
 import elfi
@@
 
-  def simulator(t1, batch_size=1,random_state=None):
+def simulator(t1, batch_size = 1, random_state = None):
     c = 0.5 - 0.5**4
     if t1 < -0.5:
-        y = ss.norm(loc=-t1-c, scale=1).rvs(random_state=random_state)
+        y = ss.norm(loc = -t1-c, scale = 1).rvs(random_state = random_state)
     elif t1 <= 0.5:
-        y = ss.norm(loc=t1**4, scale=1).rvs(random_state=random_state)
+        y = ss.norm(loc = t1**4, scale = 1).rvs(random_state = random_state)
     else:
-        y = ss.norm(loc=t1-c, scale=1).rvs(random_state=random_state)
+        y = ss.norm(loc = t1-c, scale = 1).rvs(random_state = random_state)
     return y
@@
 t1 = elfi.Prior('uniform', -2.5, 5)
-  sim = elfi.Simulator(simulator, t1, observed=y)
+sim = elfi.Simulator(simulator, t1, observed = y)
 d = elfi.Distance('euclidean', sim)
@@
 # Initialise the ROMC inference method
-  bounds = [(-2.5, 2.5)] # limits of the prior
+bounds = [(-2.5, 2.5)] # bounds of the prior
 parallelize = True # activate parallel execution
-  romc = elfi.ROMC(d, bounds=bounds, parallelize=parallelize)
-----------------------------------------------------------------------------
+romc = elfi.ROMC(d, bounds = bounds, parallelize = parallelize)
 \end{Code}

### 2022-12-21 article/latex_files/ch4/4-1-training.tex
fixed spaces and commas

@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> romc.solve_problems(n1, use_bo=False, optimizer_args=None)
-----------------------------------------------------------------------------
+romc.solve_problems(n1, use_bo = False, optimizer_args = None)
 \end{Code}
@@
 random simulator. The user may select the Bayesian Optimisation scheme
-by setting the argument \code{use_bo=True} and pass custom arguments
+by setting the argument \code{use_bo = True} and pass custom arguments
 to the optimizer through \code{optimize_args}.
@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> romc.distance_hist(**kwargs)
-----------------------------------------------------------------------------
+romc.distance_hist(**kwargs)
 \end{Code}
@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> romc.estimate_regions(eps_filter, use_surrogate=None, fit_models=False)
-----------------------------------------------------------------------------
+romc.estimate_regions(eps_filter, use_surrogate=None, fit_models=False)
 \end{Code}
@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> romc.fit_posterior(args*)  # training in a single call
->>> romc.visualize_region(i)   # acceptance region
->>> romc.compute_eps(quantile) # estimates eps
+romc.fit_posterior(args*)  # training in a single call
+romc.visualize_region(i)   # acceptance region
+romc.compute_eps(quantile) # estimates eps
 ----------------------------------------------------------------------------
@@
 of accepted solutions. It can be used to automate the selection of the
-threshold \(\epsilon\), e.g.\
+threshold \(\epsilon\), e.g.,
 \code{eps=romc.compute_eps(quantile=0.9)}.
@@
 \begin{Code}
------------------------------- python snippet ------------------------------
 # Training (fitting) part
@@
 # Training step-by-step
-  romc.solve_problems(n1=n1, seed=seed, use_bo=use_bo)
-  romc.theta_hist(bins=100) # plot hist to decide which eps to use
+romc.solve_problems(n1 = n1, seed = seed, use_bo = use_bo)
+romc.theta_hist(bins = 100) # plot hist to decide which eps to use
 
 eps = .75 # threshold for the bounding box 
-  romc.estimate_regions(eps=eps) # build the bounding boxes
+romc.estimate_regions(eps = eps) # build the bounding boxes
 
-  romc.visualize_region(i=1) # for inspecting visually the bounding box
+romc.visualize_region(i = 1) # for inspecting visually the bounding box
 
 # Equivalent one-line command
-  # romc.fit_posterior(n1=n1, eps=eps, use_bo=use_bo, seed=seed)
-----------------------------------------------------------------------------  
+# romc.fit_posterior(n1 = n1, eps = eps, use_bo = use_bo, seed = seed)
 \end{Code}

### 2022-12-21 article/latex_files/ch4/4-2-inference.tex
fixed spaces and commas

@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> romc.sample(n2)
-----------------------------------------------------------------------------
+romc.sample(n2)
 \end{Code}
@@
 \begin{Code}
----------------------------------- python ----------------------------------
->>> romc.compute_expectation(h) # E[h(x)|theta]
->>> romc.eval_unorm_posterior(theta, eps_cutoff=False) # unnorm p(theta)
->>> romc.eval_posterior(theta, eps_cutoff=False) # normalised p(theta)
-----------------------------------------------------------------------------
+romc.compute_expectation(h) # E[h(x)|theta]
+romc.eval_unorm_posterior(theta, eps_cutoff = False) # unnorm p(theta)
+romc.eval_posterior(theta, eps_cutoff = False) # normalised p(theta)
 \end{Code}
@@
 \begin{Code}
------------------------------- python snippet ------------------------------  
 # Inference part
@@
 n2 = 50
-  romc.sample(n2=n2, seed=seed)
+romc.sample(n2 = n2, seed = seed)
 
 # visualize region, adding the samples now
-  romc.visualize_region(i=1)
+romc.visualize_region(i = 1)
 
 # Visualise marginal (built-in ELFI tool)
-  romc.result.plot_marginals(weights=romc.result.weights,
-                             bins=100, range=(-3,3))
+weights = romc.result.weights
+romc.result.plot_marginals(weights = weights, bins = 100, range = (-3,3))
 
@@
 romc.result.summary()
-  # Number of samples: 19300
+# Number of samples  : 19300
 # Sample means: theta: -0.0116
@@
 # compute expectation
-  exp_val = romc.compute_expectation(h=lambda x: np.squeeze(x))
+exp_val = romc.compute_expectation(h = lambda x: np.squeeze(x))
 print("Expected value   : %.3f" % exp_val)
-  # Expected value   : -0.012
+# Expected value: -0.012
 
-  exp_var = romc.compute_expectation(h=lambda x: np.squeeze(x)**2)
+exp_var = romc.compute_expectation(h = lambda x: np.squeeze(x)**2)
 print("Expected variance: %.3f" % exp_var)
@@
 # eval unnorm posterior
-  romc.eval_unnorm_posterior(theta=0)
+romc.eval_unnorm_posterior(theta = 0)
 
 # check eval posterior
-  romc.eval_posterior(theta=0)
-----------------------------------------------------------------------------
+romc.eval_posterior(theta = 0)
 \end{Code}
-
-% \begin{figure}[h]
-%     \begin{center}
-%       \includegraphics[width=0.48\textwidth]{./latex_files/images/chapter3/example_marginal.png}
-%     \end{center}
-%   \caption[Histogram of the obtained samples at the 1D example.]{(a) Left: Histogram of the obtained samples.
-(b) Right: Acceptance region around \(\theta_1^*\) with the obtained samples plotted inside.}
-%   \label{fig:example_sampling}
-% \end{figure}
-
-% \begin{figure}[h]
-%     \begin{center}
-%       \includegraphics[width=0.6\textwidth]{./latex_files/images/chapter3/example_posterior.png}
-%     \end{center}
-%   \caption[Approximate posterior evaluation, at the 1D example.]{Approximate posterior evaluation.}
-%   \label{fig:approx_posterior}
-% \end{figure}

### 2022-12-21 article/latex_files/ch4/4-3-evaluation.tex
fixed references
fixed spaces and commas

@@
 \begin{Code}
----------------------------------- python ----------------------------------  
->>> romc.compute_divergence(gt_posterior, bounds, step, distance)
->>> romc.compute_ess()
-----------------------------------------------------------------------------
+romc.compute_divergence(gt_posterior, bounds, step, distance)
+romc.compute_ess()
 \end{Code}
@@
 
-The method \code{compute_ess} computes the Effective Sample Size (ESS)
-using the following expression~\cite{Sudman1967},
+The method \code{compute_ess} computes the Effective Sample Size
+(ESS) using the following expression,
 
@@
 \begin{Code}
------------------------------- python snippet ------------------------------  
 # Evaluation part
-  res = romc.compute_divergence(wrapper, distance="Jensen-Shannon")                                 
+res = romc.compute_divergence(wrapper, distance = "Jensen-Shannon")
 print("Jensen-Shannon divergence: %.3f" % res)
@@
 # Nof Samples: 19300, ESS: 16196.214
-----------------------------------------------------------------------------  
 \end{Code}

### 2022-12-21 article/latex_files/ch4/4-4-extensibility.tex
fixed spaces and commas

@@
 \begin{Code}
----------------------------------- python ----------------------------------
 class CustomOptim(OptimisationProblem):
@@
         def create_local_surrogate(model):
-            return partial(local_surrogate, model_scikit=model)
+            return partial(local_surrogate, model_scikit = model)
 
@@
             # train Neural Network
-            mlp = MLPRegressor(hidden_layer_sizes=(10,10), solver='adam')
+            mlp = MLPRegressor(hidden_layer_sizes = (10,10), solver = 'adam')
             model = Pipeline([('linear', mlp)])
@@
         self.state["local_surrogates"] = True
-----------------------------------------------------------------------------
 \end{Code}
@@
 \begin{Code}
----------------------------------- python ----------------------------------
+----------------------------------------------------------------------------
 def solve_gradients(self, **kwargs):
@@
 \begin{Code}
----------------------------------- python ----------------------------------
+----------------------------------------------------------------------------
 class RomcOpimisationResult:

### 2022-12-21 article/latex_files/ch5/5-ma2.tex
fixed spaces and commas

@@
 \(T\) is a constant and set to \(T=100\). For securing
-that the inference problem is identifiable, i.e.\ the likelihood has
+that the inference problem is identifiable, i.e., the likelihood has
 only one mode, we use the prior proposed by~\cite{Marin2012},
@@
 distance function \(d_i\) inside the proposal region
-\(q_i~ \forall i\) at the inference phase i.e.~sampling,
+\(q_i~ \forall i\) at the inference phase, i.e., sampling,
 computing an expectation and evaluating the posterior. In our example
@@
 method. Therefore, it is straightforward to parallelise the fitting
-part, i.e. (i) solving the optimisation problems and (ii) estimating
+part, i.e., (i) solving the optimisation problems and (ii) estimating
 the proposal regions. Our implementation supports exploiting all the

### 2022-12-21 replication_script/replication_script.py

@@ -823,8 +823,7 @@ for i, n in enumerate(n1):
     romc = elfi.ROMC(model, bounds, discrepancy_name="d", parallelize=False)
 
     tic = timeit.default_timer()
-    romc.solve_problems(n1=int(n), seed=seed,
-                        use_bo=False)
+    romc.solve_problems(n1=int(n), seed=seed, use_bo=False)
     toc = timeit.default_timer()
     solve_grad.append(toc-tic)
 

### 2022-12-21 replication_script/test.py

@@ -1,0 +1,28 @@
+import numpy as np
+import elfi
+
+# Set seed for reproducibility
+seed = 1
+np.random.seed(seed)
+model = ma2.get_model(seed_obs=seed)
+
+# Hyperparameters
+n1 = 5
+n2 = 200
+bounds = [(-2, 2), (-1.25, 1.25)]
+eps = .01
+
+# define model
+romc = elfi.ROMC(model, bounds=bounds, parallelize=False, discrepancy_name="d")
+
+# fitting part
+romc.solve_problems(n1=n1, use_bo=True, seed=seed)
+romc.estimate_regions(eps_filter=eps, fit_models=False)
+
+# sampling part
+tmp = romc.sample(n2=n2, seed=seed)
+romc.visualize_region(vis_ind_1)
+
+# print summary statistics
+print(romc.result.summary())
+print(romc.result.samples_cov())

### 2022-12-22 article/article.tex
finished revision until 2.1

@@
 %% - \Abstract{} almost as usual
-\Abstract{Performing inference in statistical models with intractable
-  likelihood is challenging; most likelihood-free inference (LFI)
-  methods encounter accuracy and efficiency limitations. In this
-  paper, we present the implementation of Robust Optimisation Monte
-  Carlo (ROMC) in the \proglang{Python} package \pkg{ELFI}. ROMC is a
-  novel and efficient (highly-parallelisable) LFI framework that
-  provides accurate weighted samples from the posterior. Our
-  implementation can be used in two ways. First, a scientist may use
-  it as an out-of-the-box LFI algorithm; we provide an easy-to-use API
-  harmonised with the principles of \pkg{ELFI}, enabling effortless
-  comparisons with the rest of the methods included in the
-  package. Additionally, we have carefully split ROMC into isolated
-  components for supporting extensibility. A researcher may experiment
-  with novel method(s) for solving part(s) of ROMC without
-  reimplementing everything from scratch. In both scenarios, all steps
-  run in a fully-parallelisable manner, exploiting all CPU
-  cores. We also provide helpful functionalities for (i) inspecting
-  the inference process and (ii) evaluating the obtained
-  samples. Finally, we test the robustness of our implementation on
-  some typical LFI examples.  }
+\Abstract{Performing inference in statistical models with an
+  intractable likelihood is challenging, therefore, most
+  likelihood-free inference (LFI) methods encounter accuracy and
+  efficiency limitations. In this paper, we present the implementation
+  of the LFI method Robust Optimisation Monte Carlo (ROMC) in the
+  \proglang{Python} package \pkg{ELFI}. ROMC is a novel and efficient
+  (highly-parallelisable) LFI framework that provides accurate
+  weighted samples from the posterior. Our implementation can be used
+  in two ways. First, a scientist may use it as an out-of-the-box LFI
+  algorithm; we provide an easy-to-use API harmonised with the
+  principles of \pkg{ELFI}, enabling effortless comparisons with the
+  rest of the methods included in the package. Additionally, we have
+  carefully split ROMC into isolated components for supporting
+  extensibility. A researcher may experiment with novel method(s) for
+  solving part(s) of ROMC without reimplementing everything from
+  scratch. In both scenarios, the ROMC parts can run in a
+  fully-parallelised manner, exploiting all CPU cores. We also provide
+  helpful functionalities for (i) inspecting the inference process and
+  (ii) evaluating the obtained samples. Finally, we test the
+  robustness of our implementation on some typical LFI examples.}
 
@@
 \section{Introduction} \label{sec:intro}
-\input{latex_files/ch1/1-introduction.tex}
+
+Simulator-based models are particularly captivating due to the
+provided modeling freedom. In essence, any data generating mechanism
+that can be written as a finite set of algorithmic steps can be
+programmed as a simulator-based model. In these cases, it is feasible
+to generate samples using the simulator but it is infeasible to
+evaluate the likelihood function. The intractability of the likelihood
+makes the likelihood-free inference (LFI), i.e., the approximation of
+the posterior distribution without using the likelihood function,
+particularly challenging.
+
+Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} is a
+novel LFI approach for approximating the posterior distribution. The
+central idea is turning the stochastic data-generating mechanism into
+a set of deterministic optimization processes. Afterwards,
+\citet{Forneron2016} provided a similar method under the name `reverse
+sampler'. In their work,~\citet{Ikonomov2019}, located some critical
+limitations of OMC, so they proposed Robust OMC (ROMC), an alternative
+version of OMC with the appropriate modifications.
+
+In this paper, we present the implementation of ROMC at the
+\proglang{Python} package \pkg{ELFI} (Engine for likelihood-free
+inference)~\citet{1708.00707}. As we illustrate at
+Section~\ref{sec:implementation}, we have carefully designed the
+implementation to ensure extensibility. ROMC is a general framework
+for obtaining weighted samples from the posterior, i.e., it defines a
+sequence of algorithmic steps without enforcing a specific algorithm
+for solving each step. Therefore, a researcher may use ROMC as the
+backbone algorithm and develop novel methods to solve each separate
+step.\footnote{For being a ready-to-use
+algorithm,~\citet{Ikonomov2019} proposed a default method for each
+step, but this choice is by no means restrictive.} We have designed
+our software for facilitating such experimentation.
+
+To the best of our knowledge, this is the first implementation of the
+ROMC inference method to a generic LFI framework. We organize the
+evaluation in three steps. First, for securing that our implementation
+is accurate, we test it against an artificial example with a tractable
+likelihood. The artificial example also serves as a step-by-step guide
+for showcasing how to use the various functionalities of our
+implementation. Second, we use the second-order moving average (MA2)
+example from the \pkg{ELFI} package, using as ground truth the samples
+obtained with Rejection ABC~\citet{lintusaari2017}, with a very large
+number of trials. Finally, we present the execution times of ROMC,
+measuring the speed-up obtained by using the parallel version of the
+implementation.
 
@@
 \section{Background}
-\input{latex_files/ch2/2-0-intro.tex}
+
+We first give a short introduction to simulator-based models, we then
+focus on OMC and its robust improvement, ROMC, and we, finally,
+introduce \pkg{ELFI}, the underlying package that is used for the
+implementation of ROMC.
 
 \subsection{Simulator-based models and likelihood-free inference}
-\input{latex_files/ch2/2-1-implicit-models.tex}
 
-\subsection{Optimisation Monte Carlo (OMC)}
+An implicit or simulator-based model is a parameterized stochastic
+data generating mechanism. The key characteristic of these models is
+that we can sample data points, but we cannot evaluate the
+likelihood. Formally, a simulator-based model is a parameterized
+family of probability density functions
+\(\{ p(\yb|\thetab)\}_{\thetab}\) whose closed-form is either unknown
+or computationally intractable. In these scenarios, we can only a
+access the simulator \( m_r(\thetab) \), i.e., a black-box mechanism
+(computer code) that generates samples \(\yb\) in a stochastic manner
+from a set of parameters \(\thetab\). We denote the process of
+obtaining samples from the simulator with
+\( m_r(\thetab) \rightarrow \yb \). It is feasible to isolate the
+randomness of the simulator by introducing the nuisance random
+variables denoted by \(\ub \sim p(\ub)\). Therefore, for a tuple
+\((\thetab, \ub)\), the simulator turns to a deterministic mapping
+\(g\), such that \(\yb=\simulator(\thetab,\ub)\). In terms of computer
+code, the randomness of a random process is governed by the global
+seed. Even though each software may handle the random generation
+process in a different way\footnote{For example, at
+  \pkg{Numpy}~\citet{harris2020array}, the pseudo-random number
+  generation is based on a global state, whereas, in
+  \pkg{JAX}~\citet{jax2018github} random functions consume a key that
+  is passed as parameter. }, in all cases, by sampling an integer value
+and setting it as the initial seed converts the simulation to a
+deterministic sequence of steps.
+
+Simulator-based models provide considerable modeling freedom; any
+physical process that can be conceptualized as a computer program of
+finite steps can be modeled as a simulator-based model without any
+compromise. The modeling freedom allows for any amount of hidden
+(unobserved) internal variables or rule-based decisions. Hence,
+implicit models are often used to model physical phenomena in the
+natural sciences such as, e.g., genetics, epidemiology or
+neuroscience. Further background on simulator-based models and
+example applications can be found in the articles
+by~\citet{gutmann2016, lintusaari2017, sisson2018, cranmer2020}.
+
+The modeling freedom of simulator-based models, however, comes at the
+price of difficulties in inferring their parameters. Denoting the
+observed data as \(\data\), the main difficulty is that the likelihood
+function \(l(\thetab) = p(\data|\thetab)\) is generally
+intractable. To better see the sources of the intractability, and to
+address them, we go back to the basic characterization of the
+likelihood as the (rescaled) probability of a parameter of the model
+to generate data \(\yb\) that is similar to the observed data
+\(\data\). More formally, the likelihood \(l(\thetab)\) equals
+\begin{equation} \label{eq:likelihood}
+  l(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in b_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
+  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
+\end{equation}
+where \(c_\epsilon\) is a proportionality factor that depends on
+\(\epsilon\) and \(\region(\data)\) is an \(\epsilon\) region around \(\data\)
+that is defined via a distance function \(d\), i.e., \ \(\region(\data)
+:= \{\yb: d(\yb, \data) \leq \epsilon \}\).
+
+The basic characterization of the likelihood in~\eqref{eq:likelihood}
+highlights two sources of intractability; the first is the computation
+of the probability
+\(\Pr(\simulator(\thetab,\ub) \in \region(\data))\), the second is the
+limit of \(\epsilon \to 0\). Approximating the probability with
+samples becomes computationally infeasible if \(\epsilon\) is too
+small. Hence, a large class of inference methods work with
+\(\epsilon > 0\), which leads to the approximate likelihood function
+\(l_{d, \epsilon}(\thetab)\)
+\begin{equation} \label{eq:approx_likelihood}
+  l_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
+\end{equation}
+
+and, in turn, to the approximate posterior
+
+\begin{equation} \label{eq:approx_posterior}
+  p_{d,\epsilon}(\thetab|\data) \propto l_{d, \epsilon}(\thetab) p(\thetab).
+\end{equation}
+
+The approximation in~\eqref{eq:approx_likelihood} is by no means the
+only strategy to deal with the intractability of the likelihood
+function in~\eqref{eq:likelihood}. Other strategies include modeling
+the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
+reverse, or framing likelihood-free inference as a ratio estimation
+problem, see for example \citet{blum2010, Wood2006, Papamakarios2016,
+  Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both
+OMC and robust OMC, which we introduce next, are based on the
+approximation in~\eqref{eq:approx_likelihood}.
+
+\subsection{Optimization Monte Carlo (OMC)}
 \input{latex_files/ch2/2-2-OMC.tex}
 
-\subsection{Robust optimisation Monte Carlo (ROMC)}
+\subsection{Robust optimization Monte Carlo (ROMC)}
 \input{latex_files/ch2/2-3-ROMC.tex}
@@
 \section{Implementation}
+\label{sec:implementation}
 This section is split in two parts. We first express ROMC as an
@@
 
-\subsection{Algotirhmic view of ROMC}
+\subsection{Algorithmic view of ROMC}
 \input{latex_files/ch3/3-0-design-principles.tex}

### 2022-12-22 article/latex_files/ch1/1-introduction.tex
fixed spaces and commas
finished revision until 2.1

@@
 Simulator-based models are particularly captivating due to the
-modelling freedom they provide. In essence, a simulator-based model
-can describe any data generating mechanism that can be written as a
-finite set of algorithmic steps. Such freedom comes at a cost;
-performing the inference, i.e., sampling or evaluating the posterior,
-is challenging.
+provided modeling freedom. In essence, any data generating mechanism
+that can be written as a finite set of algorithmic steps can be
+programmed as a simulator-based model. In these cases, it is feasible
+to generate samples using the simulator but it is infeasible to
+evaluate the likelihood function. The intractability of the likelihood
+makes the likelihood-free inference (LFI), i.e., the approximation of
+the posterior distribution without using the likelihood function,
+particularly challenging.
 
-Optimisation Monte Carlo (OMC) proposed by~\citet{Meeds2015} is a
+Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} is a
 novel LFI approach for approximating the posterior distribution. The
 central idea is turning the stochastic data-generating mechanism into
-a set of deterministic optimisation processes. Afterwards,
+a set of deterministic optimization processes. Afterwards,
 \citet{Forneron2016} provided a similar method under the name `reverse
-sampler'. In their work, ~\citet{Ikonomov2019}, located some critical
-limitations of OMC. They proposed Robust OMC (ROMC), an alternative
-version of OMC with appropriate modifications.
+sampler'. In their work,~\citet{Ikonomov2019}, located some critical
+limitations of OMC, so they proposed Robust OMC (ROMC), an alternative
+version of OMC with the appropriate modifications.
 
-In this paper, we present an implementation of ROMC at the
-\proglang{Python} package \pkg{Engine for Likelihood-Free inference
-(ELFI)}. We follow appropriate design principles for ensuring
-extensibility. As we describe analytically in the next chapter, ROMC
-is a general framework for obtaining weighted samples from the
-posterior; it defines a sequence of algorithmic steps without
-enforcing a specific algorithm for solving each step\footnote{For
-being a ready-to-use algorithm,~\citet{Ikonomov2019} proposed a
-default method for each step, but this choice is by no means
-restrictive.}. A researcher may adopt ROMC and develop novel methods
-to solve each task. We have designed our software for facilitating
-such alterations. Finally, we have tested the accuracy and the
-efficiency of our implementation on some LFI examples supported by the
-\pkg{ELFI} package.
+In this paper, we present the implementation of ROMC at the
+\proglang{Python} package \pkg{ELFI} (Engine for likelihood-free
+inference)~\citet{1708.00707}. As we illustrate at
+Section~\ref{sec:implementation}, we have carefully designed the
+implementation to ensure extensibility. ROMC is a general framework
+for obtaining weighted samples from the posterior, i.e., it defines a
+sequence of algorithmic steps without enforcing a specific algorithm
+for solving each step. Therefore, a researcher may use ROMC as the
+backbone algorithm and develop novel methods to solve each separate
+step.\footnote{For being a ready-to-use
+algorithm,~\citet{Ikonomov2019} proposed a default method for each
+step, but this choice is by no means restrictive.} We have designed
+our software for facilitating such experimentation.
 
 To the best of our knowledge, this is the first implementation of the
-ROMC inference method to a generic LFI framework, so there are no
-other packages for direct comparisons. Therefore, we test ROMC against
-(i) an artificial example with tractable likelihood and (ii) the
-second-order moving average (MA2) example from the \pkg{ELFI}
-package. In the latter, we generate data using known parameters and
-compare our results with the typical Rejection
-ABC~\citet{lintusaari2017}.
+ROMC inference method to a generic LFI framework. We organize the
+evaluation in three steps. First, for securing that our implementation
+is accurate, we test it against an artificial example with a tractable
+likelihood. The artificial example also serves as a step-by-step guide
+for showcasing how to use the various functionalities of our
+implementation. Second, we use the second-order moving average (MA2)
+example from the \pkg{ELFI} package, using as ground truth the samples
+obtained with Rejection ABC~\citet{lintusaari2017}, with a very large
+number of trials. Finally, we present the execution times of ROMC,
+measuring the speed-up obtained by using the parallel version of the
+implementation.

### 2022-12-22 replication_script/test.py

@@ -12,7 +12,7 @@ n2 = 200
 bounds = [(-2, 2), (-1.25, 1.25)]
 eps = .01
 
-# define model
+# ROMC
 romc = elfi.ROMC(model, bounds=bounds, parallelize=False, discrepancy_name="d")
 
 # fitting part
@@ -26,3 +26,13 @@ romc.visualize_region(vis_ind_1)
 # print summary statistics
 print(romc.result.summary())
 print(romc.result.samples_cov())
+
+
+# BOLFI
+bolfi = elfi.BOLFI(model, target_name = "d",
+                   batch_size=1,
+                   initial_evidence=20,
+                   update_interval=10,
+                   bounds={'t1':(-2, 2), 't2':(-1.25, 1.25)},
+                   acq_noise_var=0.1,
+                   seed=seed)

### 2022-12-27 article/article.tex
revised until 2.2

@@
 An implicit or simulator-based model is a parameterized stochastic
-data generating mechanism. The key characteristic of these models is
-that we can sample data points, but we cannot evaluate the
+data generating mechanism. The key characteristic of simulator-based
+models is that we can sample data points, but we cannot evaluate the
 likelihood. Formally, a simulator-based model is a parameterized
@@
 or computationally intractable. In these scenarios, we can only a
-access the simulator \( m_r(\thetab) \), i.e., a black-box mechanism
+access the simulator \( m_r(\thetab) \), i.e., the black-box mechanism
 (computer code) that generates samples \(\yb\) in a stochastic manner
@@
 obtaining samples from the simulator with
-\( m_r(\thetab) \rightarrow \yb \). It is feasible to isolate the
-randomness of the simulator by introducing the nuisance random
-variables denoted by \(\ub \sim p(\ub)\). Therefore, for a tuple
-\((\thetab, \ub)\), the simulator turns to a deterministic mapping
-\(g\), such that \(\yb=\simulator(\thetab,\ub)\). In terms of computer
-code, the randomness of a random process is governed by the global
-seed. Even though each software may handle the random generation
-process in a different way\footnote{For example, at
+\( m_r(\thetab) \rightarrow \yb \). As shown by~\citet{Meeds2015}, it
+is feasible to isolate the randomness of the simulator by introducing
+a set of nuisance random variables denoted by \(\ub \sim
+p(\ub)\). Therefore, for a specific tuple \((\thetab, \ub)\) the
+simulator becomes a deterministic mapping \(g\), such that
+\(\yb=\simulator(\thetab,\ub)\). In terms of computer code, the
+stochasticity of a random process is governed by the global
+seed. Although each software may handle the randomness in slightly
+different ways\footnote{For example, at
   \pkg{Numpy}~\citet{harris2020array}, the pseudo-random number
@@
   \pkg{JAX}~\citet{jax2018github} random functions consume a key that
-  is passed as parameter. }, in all cases, by sampling an integer value
-and setting it as the initial seed converts the simulation to a
-deterministic sequence of steps.
+  is passed as parameter. }, in all cases, freezing the initial seed
+to a specific integer converts the simulation to a deterministic piece
+of code.
 
-Simulator-based models provide considerable modeling freedom; any
+Simulator-based models provide modeling freedom, including hidden
+(unobserved) random variables or rule-based decisions. In fact, any
 physical process that can be conceptualized as a computer program of
-finite steps can be modeled as a simulator-based model without any
-compromise. The modeling freedom allows for any amount of hidden
-(unobserved) internal variables or rule-based decisions. Hence,
+finite steps can be modeled as a simulator-based model. Hence,
 implicit models are often used to model physical phenomena in the
 natural sciences such as, e.g., genetics, epidemiology or
-neuroscience. Further background on simulator-based models and
-example applications can be found in the articles
-by~\citet{gutmann2016, lintusaari2017, sisson2018, cranmer2020}.
+neuroscience. Further background on simulator-based models and example
+applications can be found in the articles by~\citet{gutmann2016,
+  lintusaari2017, sisson2018, cranmer2020}.
 
 The modeling freedom of simulator-based models, however, comes at the
-price of difficulties in inferring their parameters. Denoting the
-observed data as \(\data\), the main difficulty is that the likelihood
-function \(l(\thetab) = p(\data|\thetab)\) is generally
-intractable. To better see the sources of the intractability, and to
-address them, we go back to the basic characterization of the
-likelihood as the (rescaled) probability of a parameter of the model
-to generate data \(\yb\) that is similar to the observed data
-\(\data\). More formally, the likelihood \(l(\thetab)\) equals
-\begin{equation} \label{eq:likelihood}
-  l(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in b_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
+price of difficulties in inferring the parameters of
+interest. Denoting the observed data as \(\data\), the main difficulty
+lies at the intractability of the likelihood function
+\(L(\thetab) = p(\data|\thetab)\). To better see the sources of the
+intractability, and to address them, we go back to the basic
+characterization of the likelihood as the (rescaled) probability of
+generating data \(\yb\) that is similar to the observed data
+\(\data\), using parameters \(\thetab\). Formally, the likelihood
+\(L(\thetab)\) equals
+
+\begin{equation}
+  \label{eq:likelihood}
+  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
   \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
@@
 where \(c_\epsilon\) is a proportionality factor that depends on
-\(\epsilon\) and \(\region(\data)\) is an \(\epsilon\) region around \(\data\)
-that is defined via a distance function \(d\), i.e., \ \(\region(\data)
-:= \{\yb: d(\yb, \data) \leq \epsilon \}\).
+\(\epsilon\) and \(\region(\data)\) is an \(\epsilon\) region around
+\(\data\) that is defined through a distance function \(d\), i.e., \
+\(\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon \}\).
 
-The basic characterization of the likelihood in~\eqref{eq:likelihood}
-highlights two sources of intractability; the first is the computation
-of the probability
-\(\Pr(\simulator(\thetab,\ub) \in \region(\data))\), the second is the
-limit of \(\epsilon \to 0\). Approximating the probability with
-samples becomes computationally infeasible if \(\epsilon\) is too
-small. Hence, a large class of inference methods work with
-\(\epsilon > 0\), which leads to the approximate likelihood function
-\(l_{d, \epsilon}(\thetab)\)
-\begin{equation} \label{eq:approx_likelihood}
-  l_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
+Equation~\ref{eq:likelihood} highlights the main source of
+intractability; computing
+\(\Pr(\simulator(\thetab,\ub) \in \region(\data)) | \thetab\) as the
+fraction of samples that lie inside the \(\epsilon\) region around
+\(\data\) is computationally infeasible in the limit of
+\(\epsilon \to 0\). Hence, the constrainted is relaxed to
+\(\epsilon > 0\), which leads to the approximate likelihood:
+
+\begin{equation}
+  \label{eq:approx_likelihood}
+  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
 \end{equation}
 
-and, in turn, to the approximate posterior
+and, in turn, to the approximate posterior:
 
 \begin{equation} \label{eq:approx_posterior}
-  p_{d,\epsilon}(\thetab|\data) \propto l_{d, \epsilon}(\thetab) p(\thetab).
+  p_{d,\epsilon}(\thetab|\data)
+  \propto L_{d, \epsilon}(\thetab) p(\thetab)
+  \propto L_{d, \epsilon}(\thetab) p(\thetab)
 \end{equation}
 
-The approximation in~\eqref{eq:approx_likelihood} is by no means the
-only strategy to deal with the intractability of the likelihood
-function in~\eqref{eq:likelihood}. Other strategies include modeling
-the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
-reverse, or framing likelihood-free inference as a ratio estimation
-problem, see for example \citet{blum2010, Wood2006, Papamakarios2016,
-  Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both
-OMC and robust OMC, which we introduce next, are based on the
-approximation in~\eqref{eq:approx_likelihood}.
+The approximation in
+Equations~\ref{eq:approx_likelihood},~\ref{eq:approx_posterior} is by
+no means the only strategy to deal with the intractability of the
+likelihood function in Equation~\ref{eq:likelihood}. Other strategies include
+modeling the (stochastic) relationship between \(\thetab\) and
+\(\yb\), and its reverse, or framing likelihood-free inference as a
+ratio estimation problem, see for example \citet{blum2010, Wood2006,
+  Papamakarios2016, Papamakarios2019, Chen2019, Thomas2020,
+  Hermans2020}. However, both OMC and robust OMC, which we introduce
+next, are based on the approximation in Equation~\ref{eq:approx_likelihood}.
 
 \subsection{Optimization Monte Carlo (OMC)}
-\input{latex_files/ch2/2-2-OMC.tex}
+
+Our description of OMC~\citet{Meeds2015}
+follows~\citet{Ikonomov2019}. We define the indicator function (boxcar
+kernel) that equals one only if \(\xb\) lies in \(\region(\yb)\):
+%
+\begin{equation}
+  \label{eq:indicator}
+  \indicator{\region(\yb)}(\xb)=
+  \left\{
+    \begin{array}{ll}
+      1 & \mbox{if } \xb \in \region(\yb) \\
+      0 & \mbox{otherwise}
+    \end{array} \right. \end{equation}
+%
+
+We, then, rewrite the approximate likelihood function
+\(L_{d, \epsilon}(\thetab)\) of Equation~\ref{eq:approx_likelihood} as:
+
+\begin{gather}
+  \label{eq:approx_likelihood_omc}
+  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) | \thetab) =
+  \int_{\ub} \indicator{\region(\data)}(g(\thetab, \ub)) d \ub
+\end{gather}
+
+which can be approximated using samples from the simulator:
+
+\begin{equation}
+  \label{eq:samples_approx_likelihood_omc}
+  L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\simulator(\thetab, \ub_i))
+ \quad \text{ where } \ub_i \sim p(\ub).
+\end{equation}
+
+In Equation~\ref{eq:samples_approx_likelihood_omc}, for each
+\(\ub_i\), there is a region \(\accregioni\) in the parameter space
+\(\thetab\) where the indicator function returns one, i.e.,
+\(\accregioni = \{ \thetab: \simulator(\thetab, \ub_i) \in
+\region(\data) \}\). Therefore, we can approximate the likelihood as:
+
+\begin{equation}
+  \label{eq:alt_view_theta}
+  L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\accregioni}(\thetab)
+\end{equation}
+
+and the corresponding approximate posterior:
+
+\begin{equation}
+  \label{eq:approx_posterior_omc}
+  p_{d,\epsilon}(\thetab|\data) \propto
+  p(\thetab) \sum_i^N  \indicator{\accregioni}(\thetab).
+\end{equation}
+
+As argued by \citet{Ikonomov2019}, these derivations provide a unique
+perspective for likelihood-free inference by shifting the focus onto
+the geometry of the acceptance regions \(\accregioni\). Indeed, the
+task of approximating the likelihood and the posterior becomes a task
+of characterising the sets \(\accregioni\). OMC by \citet{Meeds2015}
+assumes that the distance \(d\) is the Euclidean distance
+\(||\cdot||_2\) between summary statistics \(\Phi\) of the observed
+and generated data, and that the \(\accregioni\) can be well
+approximated by infinitesimally small ellipses. These assumptions lead
+to an approximation of the posterior in terms of weighted samples
+\(\thetab_i^*\) that achieve the smallest distance between observed
+and simulated data for each realisation \(\ub_i \sim p(\ub)\), i.e.,\
+\begin{equation} \label{eq:omc_opt_prob}
+\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2  , \quad \ub_i \sim p(\ub).
+\end{equation}
+The weighting for each \(\thetab_i^*\) is proportional to the prior
+density at \(\thetab_i^*\) and inversely proportional to the determinant
+of the Jacobian matrix of the summary statistics at \(\thetab_i^*\). For
+further details on OMC we refer the reader to \citep{Meeds2015,
+  Ikonomov2019}.
 
 \subsection{Robust optimization Monte Carlo (ROMC)}
-\input{latex_files/ch2/2-3-ROMC.tex}
+
+\citet{Ikonomov2019} showed that considering infinitesimally small
+ellipses can lead to highly overconfident posteriors. We refer the
+reader to their paper for the technical details and conditions for
+this issue to occur. Intuitively, it happens because the weights in
+OMC are only computed from information at \(\thetab_i^*\), and using
+only local information can be misleading. For example if the curvature
+of \(||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2\) at
+\(\thetab_i^*\) is nearly flat, the curvature alone may wrongly
+indicate that \(\accregioni\) is much larger than it actually is. In
+our software package we implement the robust generalisation of OMC by
+\citet{Ikonomov2019} that resolves this issue.
+
+ROMC, firstly, approximates the acceptance regions \(\accregioni\),
+and defines proposal distributions \(q_i(\thetab)\) on them. The
+proposal distributions are used for generating posterior samples
+\(\thetab_{ij} \sim q_i\). The samples are assigned (importance)
+weights \(w_{ij}\) that compensate for using the proposal
+distributions \(q_i(\thetab)\) and not the prior \(p(\thetab)\),
+\begin{equation}
+  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
+  \label{eq:sampling}
+\end{equation}
+Given the weighted samples, any expectation
+\(\E_{p(\thetab|\data)}[h(\thetab)]\) of some function \(h(\thetab)\), can be approximated as
+\begin{equation} \label{eq:expectation}
+  \E_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
+\end{equation}
+\citet{Ikonomov2019} considered uniform distributions as proposal
+distributions so that the main task is to approximate the acceptance
+regions \(\accregioni\) and to represent them so that uniform sampling
+is easy. The approximation of the acceptance regions contains two
+compulsory and one optional step: (1) solving the optimisation
+problems as in OMC, (2) constructing bounding boxes around
+\(\accregioni\) and optionally, (3) refining the approximation via a
+surrogate model of the distance.
+
+\subsubsection*{Solving the deterministic optimisation problems}
+For each set of nuisance variables \(\ub_i, i = \{1,2,\ldots,n_1 \}\),
+we search for a point \(\thetab^*_i\) such that
+\(d(\simulator(\thetab^*_i,\ub_i), \data) \le \epsilon\). In general,
+\(d\) can be any valid distance, but for the rest of the paper we
+consider \(d\) as the squared Euclidean distance. For notational
+convenience, we denote \(d(\simulator(\thetab,\ub_i), \data)\) as
+\(d_i(\thetab)\).  Obtaining \(\theta_i^*\) involves solving the
+following optimisation problem:
+\begin{subequations}
+\begin{alignat}{2}
+  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
+  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
+\end{alignat}
+\end{subequations}
+%
+The optimisation problem can be treated as unconstrained, accepting
+the optimal point
+\(\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)\) only if
+\(d_i(\thetab_i^*) \le \epsilon\). If \(d_i(\thetab)\) is
+differentiable any gradient-based optimizer can be used
+for~\ref{eq:optProb}. The gradients \(\nabla_{\thetab} d_i(\thetab)\)
+can be either provided in closed form or approximated by finite
+differences. In case \(d_i\) is not differentiable, Bayesian
+Optimisation~\citep{Shahriari2016} provides an alternative
+approach. In this scenario, apart from obtaining an optimal
+\(\thetab_i^* \), a surrogate model \(\hat{d}_i(\thetab)\) of the
+distance function \(d_i(\thetab)\) is also automatically obtained;
+\(\hat{d}_i\) can then substitute the actual distance function in
+downstream steps of the algorithms, with possible computational gains
+especially if evaluating the actual distance \(d_i(\thetab)\) is
+expensive.
+
+\subsubsection*{Estimating the acceptance regions}
+The acceptance region \(\accregioni\) is approximated by a bounding
+box \(\accregionihat\). Ideally, we want the bounding box to be as
+tight as possible to \(\accregioni\) to ensure high acceptance rate in
+the importance sampling, but big enough for not discarding valid
+parts. The bounding boxes are built in two steps. First, we define
+their axes \(\mathbf{v}_m\), \(m = \{1, \ldots, D\}\) based on the
+(estimated) curvature of the distance at \(\thetab_i^*\). Second, we
+determine the size of the box via a one-dimensional line-search method
+along each axis, see Algorithm~\ref{alg:region_construction} for the
+details. After the bounding boxes construction, a uniform distribution
+\(q_i\) is defined on each bounding box, and is used as the proposal
+region for importance sampling.
+
+\subsubsection*{Refining the estimate via a local surrogate model (optional)}
+When computing the weight \(w_{ij}\) in \eqref{eq:sampling}, we need
+to check whether the samples \(\thetab_{ij} \sim q_i\) lie inside the
+acceptance region \(\accregioni\). This can be considered to be a
+safety-mechanism that corrects for any inaccuracies in the
+construction of \(\accregionihat\) above. However, this check involves
+evaluting the distance function \(d_i(\thetab_{ij})\), which can be
+expensive if the model is complex. \citet{Ikonomov2019} thus proposed
+to fit a surrogate model \(\tilde{d}_i(\thetab)\) of the distance
+function \(d_i(\thetab)\), on data points that lie inside
+\(\accregionihat\). They used a simple quadratic model whilst other
+regression models are, in principle, possible too. The advantage of
+using a quadratic model is that it has ellipsoidal isocontours, which
+thus naturally allowed \citet{Ikonomov2019} to replace the bounding
+box approximation of \(\accregioni\) with a tigher-fitting ellipsoidal
+approximation.\footnote{The difference to the infinitesimal
+  ellipsoidal model in OMC is the estimation procedure: OMC uses
+  information at \(\thetab_i^*\) whilst, here, information in
+  \(\accregionihat\) is used, which results in a more stable fit.}
+
+The training data for the quadratic model is obtained by sampling
+\(\thetab_{ij} \sim q_i\) and accessing the distances
+\(d_i(\thetab_{ij})\). The generation of the training data adds an
+extra computational cost, but leads to a significant speed-up when
+evaluating the weights \(w_{ij}\). Moreover, the extra cost is largely
+eliminated if Bayesian Optimisation with a Gaussian process (GP)
+surrogate model \(\hat{d}_i(\thetab)\) was used to obtain
+\(\thetab_i^*\) in the first step. In this case, we can use
+\(\hat{d}_i(\thetab)\) instead of \(d_i(\thetab)\) to generate the
+training data. This essentially replaces the global GP model with a
+simpler local quadratic model which is typically more robust.
 

### 2022-12-27 article/article.tex
finished chapter 2

@@
 \newcommand{\accregioni}{C^i_{\epsilon}}
+\newcommand{\accregionitheta}{C^i_{\epsilon, \theta_i^*}}
 \newcommand{\accregionihat}{\hat{C}^i_{\epsilon}}
@@
 \(\yb=\simulator(\thetab,\ub)\). In terms of computer code, the
-stochasticity of a random process is governed by the global
+randomness of a random process is governed by the global
 seed. Although each software may handle the randomness in slightly
@@
 \(\data\) is computationally infeasible in the limit of
-\(\epsilon \to 0\). Hence, the constrainted is relaxed to
+\(\epsilon \to 0\). Hence, the constraint is relaxed to
 \(\epsilon > 0\), which leads to the approximate likelihood:
@@
 task of approximating the likelihood and the posterior becomes a task
-of characterising the sets \(\accregioni\). OMC by \citet{Meeds2015}
+of characterizing the sets \(\accregioni\). OMC by \citet{Meeds2015}
 assumes that the distance \(d\) is the Euclidean distance
@@
 \(\thetab_i^*\) that achieve the smallest distance between observed
-and simulated data for each realisation \(\ub_i \sim p(\ub)\), i.e.,\
+and simulated data for each realization \(\ub_i \sim p(\ub)\), i.e.,\
 \begin{equation} \label{eq:omc_opt_prob}
@@
 OMC are only computed from information at \(\thetab_i^*\), and using
-only local information can be misleading. For example if the curvature
-of \(||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2\) at
-\(\thetab_i^*\) is nearly flat, the curvature alone may wrongly
-indicate that \(\accregioni\) is much larger than it actually is. In
-our software package we implement the robust generalisation of OMC by
+only local information can be misleading. For example, if the
+curvature of \(||\Phi(\data)-\Phi(\simulator(\thetab, \ub_i))||_2\) at
+\(\thetab_i^*\) is nearly flat, it may wrongly indicate that
+\(\accregioni\) is much larger than it actually is. In our software
+package we implement the robust generalization of OMC by
 \citet{Ikonomov2019} that resolves this issue.
 
-ROMC, firstly, approximates the acceptance regions \(\accregioni\),
-and defines proposal distributions \(q_i(\thetab)\) on them. The
-proposal distributions are used for generating posterior samples
-\(\thetab_{ij} \sim q_i\). The samples are assigned (importance)
-weights \(w_{ij}\) that compensate for using the proposal
-distributions \(q_i(\thetab)\) and not the prior \(p(\thetab)\),
+ROMC approximates the acceptance regions \(\accregioni\) with
+\(D\)-dimensional bounding boxes \(\accregionihat\). A uniform
+distribution, \(q_i(\thetab)\), is defined on top of each bounding
+box. Each \(q_i(\thetab)\) serves a proposal distribution for
+generating posterior samples \(\thetab_{ij} \sim q_i\). The samples
+get an (importance) weight \(w_{ij}\) that compensate for using the
+proposal distributions \(q_i(\thetab)\) instead of the prior
+\(p(\thetab)\):
+
 \begin{equation}
-  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
   \label{eq:sampling}
+  w_{ij} = \indicator{\accregioni}(\thetab_{ij}) \frac{ p(\thetab_{ij})}{q(\thetab_{ij})}.
 \end{equation}
+
 Given the weighted samples, any expectation
@@
 is easy. The approximation of the acceptance regions contains two
-compulsory and one optional step: (1) solving the optimisation
-problems as in OMC, (2) constructing bounding boxes around
-\(\accregioni\) and optionally, (3) refining the approximation via a
+compulsory and one optional step: (i) solving the optimization
+problems as in OMC, (ii) constructing bounding boxes around
+\(\accregioni\) and, optionally, (iii) refining the approximation via a
 surrogate model of the distance.
 
-\subsubsection*{Solving the deterministic optimisation problems}
+\subsubsection*{(i) Solving the deterministic optimization problems}
 For each set of nuisance variables \(\ub_i, i = \{1,2,\ldots,n_1 \}\),
 we search for a point \(\thetab^*_i\) such that
-\(d(\simulator(\thetab^*_i,\ub_i), \data) \le \epsilon\). In general,
-\(d\) can be any valid distance, but for the rest of the paper we
-consider \(d\) as the squared Euclidean distance. For notational
-convenience, we denote \(d(\simulator(\thetab,\ub_i), \data)\) as
-\(d_i(\thetab)\).  Obtaining \(\theta_i^*\) involves solving the
-following optimisation problem:
-\begin{subequations}
-\begin{alignat}{2}
-  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
-  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
-\end{alignat}
-\end{subequations}
+\(d(\simulator(\thetab^*_i,\ub_i), \data) \le \epsilon\). In
+principle, \(d(\cdot)\) can refer to any valid distance function. For
+the rest of the paper we consider \(d(\cdot)\) as the squared
+Euclidean distance, as in~\citet{Ikonomov2019}. For easier notation,
+we use \(d_i(\thetab)\) to refer to
+\(d(\simulator(\thetab,\ub_i), \data)\). We search for \(\theta_i^*\)
+solving the following optimization problem:
+
+\begin{equation}
+  \label{eq:optProb}
+  \thetab_i^* = \!\argmin_{\thetab} d_i(\thetab)
+\end{equation}
 %
-The optimisation problem can be treated as unconstrained, accepting
-the optimal point
-\(\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)\) only if
-\(d_i(\thetab_i^*) \le \epsilon\). If \(d_i(\thetab)\) is
-differentiable any gradient-based optimizer can be used
-for~\ref{eq:optProb}. The gradients \(\nabla_{\thetab} d_i(\thetab)\)
-can be either provided in closed form or approximated by finite
-differences. In case \(d_i\) is not differentiable, Bayesian
-Optimisation~\citep{Shahriari2016} provides an alternative
-approach. In this scenario, apart from obtaining an optimal
-\(\thetab_i^* \), a surrogate model \(\hat{d}_i(\thetab)\) of the
-distance function \(d_i(\thetab)\) is also automatically obtained;
-\(\hat{d}_i\) can then substitute the actual distance function in
-downstream steps of the algorithms, with possible computational gains
-especially if evaluating the actual distance \(d_i(\thetab)\) is
-expensive.
+The solution is accepted only if it satisfies the constraint
+\(d_i(\thetab_i^*) \le \epsilon\). If not, there is no point into
+building an acceptance region around \(\accregioni\), so
+\(\thetab_i^*\) is discarded. There are two proposed options for
+solving Equation~\ref{eq:optProb} depending on the differentiability
+of \(d_i(\thetab)\). If \(d_i(\thetab)\) is differentiable any
+gradient-based optimizer can be used for
+Equation~\ref{eq:optProb}. The gradients
+\(\nabla_{\thetab} d_i(\thetab)\) can be either provided in closed
+form or approximated by finite differences. If \(d_i\) is not
+differentiable, Bayesian Optimization~\citep{Shahriari2016} can be
+used instead. In this scenario, apart from obtaining an optimal
+\(\thetab_i^* \), we can also automatically build a surrogate model
+\(\hat{d}_i(\thetab)\) of the distance function \(d_i(\thetab)\). The
+surrogate model \(\hat{d}_i\) can then substitute the actual distance
+function in downstream steps of the algorithms, with possible
+computational gains especially in cases where evaluating the actual
+distance \(d_i(\thetab)\) is expensive.
 
-\subsubsection*{Estimating the acceptance regions}
-The acceptance region \(\accregioni\) is approximated by a bounding
-box \(\accregionihat\). Ideally, we want the bounding box to be as
-tight as possible to \(\accregioni\) to ensure high acceptance rate in
-the importance sampling, but big enough for not discarding valid
-parts. The bounding boxes are built in two steps. First, we define
-their axes \(\mathbf{v}_m\), \(m = \{1, \ldots, D\}\) based on the
-(estimated) curvature of the distance at \(\thetab_i^*\). Second, we
-determine the size of the box via a one-dimensional line-search method
-along each axis, see Algorithm~\ref{alg:region_construction} for the
-details. After the bounding boxes construction, a uniform distribution
-\(q_i\) is defined on each bounding box, and is used as the proposal
-region for importance sampling.
+\subsubsection*{(ii) Estimating the acceptance regions}
 
-\subsubsection*{Refining the estimate via a local surrogate model (optional)}
-When computing the weight \(w_{ij}\) in \eqref{eq:sampling}, we need
-to check whether the samples \(\thetab_{ij} \sim q_i\) lie inside the
-acceptance region \(\accregioni\). This can be considered to be a
-safety-mechanism that corrects for any inaccuracies in the
-construction of \(\accregionihat\) above. However, this check involves
-evaluting the distance function \(d_i(\thetab_{ij})\), which can be
-expensive if the model is complex. \citet{Ikonomov2019} thus proposed
-to fit a surrogate model \(\tilde{d}_i(\thetab)\) of the distance
-function \(d_i(\thetab)\), on data points that lie inside
-\(\accregionihat\). They used a simple quadratic model whilst other
-regression models are, in principle, possible too. The advantage of
-using a quadratic model is that it has ellipsoidal isocontours, which
-thus naturally allowed \citet{Ikonomov2019} to replace the bounding
-box approximation of \(\accregioni\) with a tigher-fitting ellipsoidal
-approximation.\footnote{The difference to the infinitesimal
-  ellipsoidal model in OMC is the estimation procedure: OMC uses
-  information at \(\thetab_i^*\) whilst, here, information in
-  \(\accregionihat\) is used, which results in a more stable fit.}
+Each acceptance region \(\accregioni\) is approximated by a bounding
+box \(\accregionihat\). The acceptance regions \(\accregioni\) can
+contain any number of disjoint subsets in the \(D\)-dimensional space
+and any of these subsets can take any arbitrary shape. First, since
+the bounding boxes are built around \(\thetab_i^*\), we focus only on
+the connected subset of \(\accregioni\) that contains \(\thetab_i^*\),
+which we denote as \(\accregionitheta\). Second, we want to ensure
+that the bounding box \(\accregionihat\) is big enough to contain on
+its interior all the volume of \(\accregionitheta\). Third, we want to
+be as tight as possible to \(\accregionihat\) to ensure high
+acceptance rate on the importance sampling step that will
+follow. Therefore, the bounding boxes are built as follows. First, we
+compute their axes \(\mathbf{v}_m\), for \(m = \{1, \ldots, D\}\)
+based on the (estimated) curvature of the distance at \(\thetab_i^*\),
+and, second, we apply a line-search method along each axis to determine
+the size of the box, see Algorithm~\ref{alg:region_construction} for
+the details. After the bounding boxes construction, a uniform
+distribution \(q_i\) is defined on each bounding box, and is used as
+the proposal region for importance sampling.
+
+\subsubsection*{(iii) Refining the estimate via a local surrogate model (optional)}
+
+For computing the weight \(w_{ij}\) in Equation~\ref{eq:sampling}, we
+need to check whether the samples \(\thetab_{ij}\) drawn from the
+bounding boxes are inside the acceptance region \(\accregioni\). This
+can be considered to be a safety-mechanism that corrects for any
+inaccuracies in the construction of \(\accregionihat\) above. However,
+this check involves evaluating the distance function
+\(d_i(\thetab_{ij})\), which can be expensive if the model is
+complex. \citet{Ikonomov2019} proposed fitting a surrogate model
+\(\tilde{d}_i(\thetab)\) of the distance function \(d_i(\thetab)\), on
+data points that lie inside \(\accregionihat\). In principle, any
+regression model can be used as surrogate model. \citet{Ikonomov2019}
+used a simple quadratic model because it has ellipsoidal isocontours,
+which facilitates replacing the bounding box approximation of
+\(\accregioni\) with a tighter-fitting ellipsoidal approximation.
 
@@
 evaluating the weights \(w_{ij}\). Moreover, the extra cost is largely
-eliminated if Bayesian Optimisation with a Gaussian process (GP)
+eliminated if Bayesian Optimization with a Gaussian process (GP)
 surrogate model \(\hat{d}_i(\thetab)\) was used to obtain
@@
 \label{subsec:ELFI}
-\input{latex_files/ch2/2-4-ELFI.tex}
+
+\pkg{Engine for Likelihood-Free Inference (ELFI)}\footnote{Extended
+  documentation can be found
+  \href{https://elfi.readthedocs.io/en/latest/}{https://elfi.readthedocs.io}}~\cite{1708.00707}
+is a \proglang{Python} package for LFI. We selected to implemented
+ROMC in \pkg{ELFI} since it provides convenient modules for all the
+fundamental components of a probabilistic model (e.g.\ prior,
+simulator, summaries etc.). Furthermore, \pkg{ELFI} already supports
+some recently proposed likelihood-free inference methods. \pkg{ELFI}
+handles the probabilistic model as a Directed Acyclic Graph
+(DAG). This functionality is based on the package
+\pkg{NetworkX}~\cite{hagberg2008exploring}, which supports
+general-purpose graphs. In most cases, the structure of a
+likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
+some edges connect the prior distributions to the simulator, the
+simulator is connected to the summary statistics that, in turn, lead
+to the output node. Samples can be obtained from all nodes through
+sequential (ancestral) sampling. \pkg{ELFI} automatically considers as
+parameters of interest, i.e., those we try to infer a posterior
+distribution, the ones included in the \code{elfi.Prior} class.
+
+\begin{figure}[ht]
+    \begin{center}
+      \includegraphics[width=0.8\textwidth]{./latex_files/images/chapter2/elfi.png}
+    \end{center}
+    \caption[Baseline example for creating an \pkg{ELFI} model]{Baseline example for creating an \pkg{ELFI} model. Image taken from \cite{1708.00707}}
+    \label{fig:elfi}
+\end{figure}
+
+All inference methods are implemented as separate \pkg{ELFI} classes
+with the two conventions:
+
+\begin{itemize}
+\item their constructor is \code{elfi.<Class name>(<output node>, *arg)};
+  the initial argument is the output node of the simulator-based model
+  and the rest of the arguments are parameters of the method.
+\item they obtain posterior samples using \code{elfi.<Class name>.sample(*args)}
+\end{itemize}
 

### 2022-12-27 article/article.tex
finished chapter 2

@@
 %% -- Introduction -------------------------------------------------------------
-\section{Introduction} \label{sec:intro}
+\section{Introduction}
+\label{sec:intro}
 
 Simulator-based models are particularly captivating due to the
-provided modeling freedom. In essence, any data generating mechanism
-that can be written as a finite set of algorithmic steps can be
-programmed as a simulator-based model. In these cases, it is feasible
-to generate samples using the simulator but it is infeasible to
-evaluate the likelihood function. The intractability of the likelihood
-makes the likelihood-free inference (LFI), i.e., the approximation of
-the posterior distribution without using the likelihood function,
-particularly challenging.
+modeling freedom they provide. In essence, any data generating
+mechanism that can be written as a finite set of algorithmic steps can
+be programmed as a simulator-based model. Hence, these models are
+often used to model physical phenomena in the natural sciences such
+as, e.g., genetics, epidemiology or neuroscience~\citet{gutmann2016,
+  lintusaari2017, sisson2018, cranmer2020}. In simulator-based models,
+it is feasible to generate samples using the simulator but is
+infeasible to evaluate the likelihood function. The intractability of
+the likelihood makes the so-called likelihood-free inference (LFI),
+i.e., the approximation of the posterior distribution without using
+the likelihood function, particularly challenging.
 
-Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} is a
-novel LFI approach for approximating the posterior distribution. The
-central idea is turning the stochastic data-generating mechanism into
-a set of deterministic optimization processes. Afterwards,
-\citet{Forneron2016} provided a similar method under the name `reverse
-sampler'. In their work,~\citet{Ikonomov2019}, located some critical
-limitations of OMC, so they proposed Robust OMC (ROMC), an alternative
-version of OMC with the appropriate modifications.
+Optimization Monte Carlo (OMC), proposed by~\citet{Meeds2015}, is a
+novel LFI approach. The central idea is to convert the stochastic
+data-generating mechanism into a set of deterministic optimization
+processes. Afterwards, \citet{Forneron2016} described a similar method
+under the name `reverse sampler'. In their work,~\citet{Ikonomov2019},
+identified some critical limitations of OMC and proposed Robust OMC
+(ROMC) as an improved version of OMC with appropriate
+modifications.
 
@@
 \proglang{Python} package \pkg{ELFI} (Engine for likelihood-free
-inference)~\citet{1708.00707}. As we illustrate at
-Section~\ref{sec:implementation}, we have carefully designed the
-implementation to ensure extensibility. ROMC is a general framework
-for obtaining weighted samples from the posterior, i.e., it defines a
-sequence of algorithmic steps without enforcing a specific algorithm
-for solving each step. Therefore, a researcher may use ROMC as the
-backbone algorithm and develop novel methods to solve each separate
-step.\footnote{For being a ready-to-use
+inference)~\citet{1708.00707}. The implementation has been designed to
+ensure extensibility. ROMC is an LFI framework; it defines a sequence
+of algorithmic steps for aproximating the posterior without enforcing
+a specific algorithm for solving each step. Therefore, a researcher
+may use ROMC as the backbone algorithm and develop novel methods to
+solve each separate step.\footnote{For being a ready-to-use
   algorithm,~\citet{Ikonomov2019} proposed a default method for each
@@
 example from the \pkg{ELFI} package, using as ground truth the samples
-obtained with Rejection ABC~\citet{lintusaari2017}, with a very large
-number of trials. Finally, we present the execution times of ROMC,
-measuring the speed-up obtained by using the parallel version of the
+obtained with Rejection ABC~\citet{lintusaari2017} using a very high
+number of samples. Finally, we present the execution times of ROMC,
+measuring the speed-up achieved by the parallel version of the
 implementation.
@@
 We first give a short introduction to simulator-based models, we then
-focus on OMC and its robust improvement, ROMC, and we, finally,
-introduce \pkg{ELFI}, the underlying package that is used for the
-implementation of ROMC.
+focus on OMC and its robust version, ROMC, and we, finally, introduce
+\pkg{ELFI}, the \proglang{Python} package used for the implementation.
 
@@
 An implicit or simulator-based model is a parameterized stochastic
-data generating mechanism. The key characteristic of simulator-based
-models is that we can sample data points, but we cannot evaluate the
-likelihood. Formally, a simulator-based model is a parameterized
-family of probability density functions
+data generating mechanism, where we can sample data points but we
+cannot evaluate the likelihood. Formally, a simulator-based model is a
+parameterized family of probability density functions
 \(\{ p(\yb|\thetab)\}_{\thetab}\) whose closed-form is either unknown
-or computationally intractable. In these scenarios, we can only a
-access the simulator \( m_r(\thetab) \), i.e., the black-box mechanism
-(computer code) that generates samples \(\yb\) in a stochastic manner
-from a set of parameters \(\thetab\). We denote the process of
-obtaining samples from the simulator with
-\( m_r(\thetab) \rightarrow \yb \). As shown by~\citet{Meeds2015}, it
-is feasible to isolate the randomness of the simulator by introducing
-a set of nuisance random variables denoted by \(\ub \sim
-p(\ub)\). Therefore, for a specific tuple \((\thetab, \ub)\) the
-simulator becomes a deterministic mapping \(g\), such that
-\(\yb=\simulator(\thetab,\ub)\). In terms of computer code, the
-randomness of a random process is governed by the global
-seed. Although each software may handle the randomness in slightly
+or computationally intractable. In these cases, we can only access the
+simulator \( m_r(\thetab) \), i.e., the black-box mechanism (computer
+code) that generates samples \(\yb\) in a stochastic manner from a set
+of parameters \(\thetab\). We denote the process of obtaining samples
+from the simulator with \( m_r(\thetab) \rightarrow \yb \). As shown
+by~\citet{Meeds2015}, it is feasible to isolate the randomness of the
+simulator by introducing a set of nuisance random variables denoted by
+\(\ub \sim p(\ub)\). Therefore, for a specific tuple
+\((\thetab, \ub)\) the simulator becomes a deterministic mapping
+\(g\), such that \(\yb=\simulator(\thetab,\ub)\). In terms of computer
+code, the randomness of a random process is governed by the global
+seed. Although each package may handle the randomness in slightly
 different ways\footnote{For example, at
@@
   \pkg{JAX}~\citet{jax2018github} random functions consume a key that
-  is passed as parameter. }, in all cases, freezing the initial seed
-to a specific integer converts the simulation to a deterministic piece
-of code.
+  is passed as parameter. }, in all cases, setting the initial seed to
+a specific integer converts the simulation to a deterministic piece of
+code.
 
-Simulator-based models provide modeling freedom, including hidden
-(unobserved) random variables or rule-based decisions. In fact, any
-physical process that can be conceptualized as a computer program of
-finite steps can be modeled as a simulator-based model. Hence,
-implicit models are often used to model physical phenomena in the
-natural sciences such as, e.g., genetics, epidemiology or
-neuroscience. Further background on simulator-based models and example
-applications can be found in the articles by~\citet{gutmann2016,
-  lintusaari2017, sisson2018, cranmer2020}.
-
-The modeling freedom of simulator-based models, however, comes at the
-price of difficulties in inferring the parameters of
-interest. Denoting the observed data as \(\data\), the main difficulty
-lies at the intractability of the likelihood function
+As stated at the Introduction, the modeling freedom of simulator-based
+models comes at the price of difficulties in inferring the parameters
+of interest. Denoting the observed data as \(\data\), the main
+difficulty lies at the intractability of the likelihood function
 \(L(\thetab) = p(\data|\thetab)\). To better see the sources of the
@@
 \(\data\), using parameters \(\thetab\). Formally, the likelihood
-\(L(\thetab)\) equals
+\(L(\thetab)\) is:
 
@@
 \(\data\) that is defined through a distance function \(d\), i.e., \
-\(\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon \}\).
-
-Equation~\ref{eq:likelihood} highlights the main source of
+\(\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon
+\}\). Equation~\ref{eq:likelihood} highlights the main source of
 intractability; computing
-\(\Pr(\simulator(\thetab,\ub) \in \region(\data)) | \thetab\) as the
+\(\Pr(\simulator(\thetab,\ub) \in \region(\data) | \thetab) \) as the
 fraction of samples that lie inside the \(\epsilon\) region around
-\(\data\) is computationally infeasible in the limit of
+\(\data\) is computationally infeasible in the limit where
 \(\epsilon \to 0\). Hence, the constraint is relaxed to
@@
   p_{d,\epsilon}(\thetab|\data)
-  \propto L_{d, \epsilon}(\thetab) p(\thetab)
   \propto L_{d, \epsilon}(\thetab) p(\thetab)
@@
 \(\accregioni = \{ \thetab: \simulator(\thetab, \ub_i) \in
-\region(\data) \}\). Therefore, we can approximate the likelihood as:
+\region(\data) \}\). Therefore, we can rewrite the approximate
+likelihood and posterior as:
 
@@
 \end{equation}
-
-and the corresponding approximate posterior:
 
@@
 \(D\)-dimensional bounding boxes \(\accregionihat\). A uniform
-distribution, \(q_i(\thetab)\), is defined on top of each bounding
-box. Each \(q_i(\thetab)\) serves a proposal distribution for
-generating posterior samples \(\thetab_{ij} \sim q_i\). The samples
-get an (importance) weight \(w_{ij}\) that compensate for using the
-proposal distributions \(q_i(\thetab)\) instead of the prior
-\(p(\thetab)\):
+distribution, \(q_i(\thetab)\), is defined on top of each bounding box
+and serves as a proposal distribution for generating posterior samples
+\(\thetab_{ij} \sim q_i\). The samples get an (importance) weight
+\(w_{ij}\) that compensate for using the proposal distributions
+\(q_i(\thetab)\) instead of the prior \(p(\thetab)\):
 
@@
 \end{equation}
-\citet{Ikonomov2019} considered uniform distributions as proposal
-distributions so that the main task is to approximate the acceptance
-regions \(\accregioni\) and to represent them so that uniform sampling
-is easy. The approximation of the acceptance regions contains two
+
+The approximation of the acceptance regions contains two
 compulsory and one optional step: (i) solving the optimization
@@
 the rest of the paper we consider \(d(\cdot)\) as the squared
-Euclidean distance, as in~\citet{Ikonomov2019}. For easier notation,
-we use \(d_i(\thetab)\) to refer to
-\(d(\simulator(\thetab,\ub_i), \data)\). We search for \(\theta_i^*\)
-solving the following optimization problem:
+Euclidean distance, as in~\citet{Ikonomov2019}. For simplicity, we use
+\(d_i(\thetab)\) to refer to \(d(\simulator(\thetab,\ub_i),
+\data)\). We search for \(\theta_i^*\) solving:
 
@@
 %
-The solution is accepted only if it satisfies the constraint
-\(d_i(\thetab_i^*) \le \epsilon\). If not, there is no point into
-building an acceptance region around \(\accregioni\), so
-\(\thetab_i^*\) is discarded. There are two proposed options for
-solving Equation~\ref{eq:optProb} depending on the differentiability
-of \(d_i(\thetab)\). If \(d_i(\thetab)\) is differentiable any
-gradient-based optimizer can be used for
-Equation~\ref{eq:optProb}. The gradients
+and we accept the solution only if it satisfies the constraint
+\(d_i(\thetab_i^*) \le \epsilon\). If \(d_i(\thetab)\) is
+differentiable, Equation~\ref{eq:optProb} can be solved using any
+gradient-based optimizer. The gradients
 \(\nabla_{\thetab} d_i(\thetab)\) can be either provided in closed
@@
 contain any number of disjoint subsets in the \(D\)-dimensional space
-and any of these subsets can take any arbitrary shape. First, since
-the bounding boxes are built around \(\thetab_i^*\), we focus only on
-the connected subset of \(\accregioni\) that contains \(\thetab_i^*\),
-which we denote as \(\accregionitheta\). Second, we want to ensure
-that the bounding box \(\accregionihat\) is big enough to contain on
-its interior all the volume of \(\accregionitheta\). Third, we want to
-be as tight as possible to \(\accregionihat\) to ensure high
-acceptance rate on the importance sampling step that will
-follow. Therefore, the bounding boxes are built as follows. First, we
-compute their axes \(\mathbf{v}_m\), for \(m = \{1, \ldots, D\}\)
+and any of these subsets can take any arbitrary shape. We should make
+three important remarks. First, since the bounding boxes are built
+around \(\thetab_i^*\), we focus only on the connected subset of
+\(\accregioni\) that contains \(\thetab_i^*\), which we denote as
+\(\accregionitheta\). Second, we want to ensure that the bounding box
+\(\accregionihat\) is big enough to contain on its interior all the
+volume of \(\accregionitheta\). Third, we want \(\accregionihat\) to
+be as tight as possible to \(\accregionitheta\) to ensure high
+acceptance rate on the importance sampling step that
+follows. Therefore, the bounding boxes are built in two steps. First,
+we compute their axes \(\mathbf{v}_m\), for \(m = \{1, \ldots, D\}\)
 based on the (estimated) curvature of the distance at \(\thetab_i^*\),
-and, second, we apply a line-search method along each axis to determine
-the size of the box, see Algorithm~\ref{alg:region_construction} for
-the details. After the bounding boxes construction, a uniform
-distribution \(q_i\) is defined on each bounding box, and is used as
-the proposal region for importance sampling.
+and, second, we apply a line-search method along each axis to
+determine the size of the box. We refer the reader to
+Algorithm~\ref{alg:region_construction} for the details. After the
+bounding boxes construction, a uniform distribution \(q_i\) is defined
+on each bounding box, and is used as the proposal region for
+importance sampling.
 
@@
 
-All inference methods are implemented as separate \pkg{ELFI} classes
-with the two conventions:
-
-\begin{itemize}
-\item their constructor is \code{elfi.<Class name>(<output node>, *arg)};
-  the initial argument is the output node of the simulator-based model
-  and the rest of the arguments are parameters of the method.
-\item they obtain posterior samples using \code{elfi.<Class name>.sample(*args)}
-\end{itemize}
+All inference methods of \pkg{ELFI} are implemented following two
+conventions. First, their constructor follows the signature
+\code{elfi.<Class name>(<output node>, *arg)}, where \code{<output
+  node>} is the output node of the simulator-based model and
+\code{*arg} are the parameters of the method. Second, they provide a
+method \code{elfi.<Class name>.sample(*args)} for drawing samples from
+the approximate posterior.
 

### 2022-12-28 article/article.tex
finished chapter 4

@@
 ensure extensibility. ROMC is an LFI framework; it defines a sequence
-of algorithmic steps for aproximating the posterior without enforcing
+of algorithmic steps for approximating the posterior without enforcing
 a specific algorithm for solving each step. Therefore, a researcher
@@
 %% -- Design Principles -----------------------------------------------------
-\section{Implementation}
+\section{Overview of the implementation}
 \label{sec:implementation}
-This section is split in two parts. We first express ROMC as an
-algorithm and then we present the general implementation principles we
-follow.
+In this section, we express ROMC as an algorithm and then we present
+the general implementation principles.
 
 \subsection{Algorithmic view of ROMC}
-\input{latex_files/ch3/3-0-design-principles.tex}
+
+For designing an extendable implementation, we firstly define ROMC as
+a sequence of algorithmic steps. At a high level, ROMC can be split
+into the training and the inference part; the training part covers the
+steps for estimating the proposal regions and the inference part
+calculates the weighted samples. In
+Algorithm~\ref{alg:romc_algorithm}, that defines ROMC as an algorithm,
+steps 2-11 (before the horizontal line) refer to the training part and
+steps 13-18 to the inference part.
+
+\begin{algorithm}[!ht]
+  \caption{ROMC. Requires the prior \( p(\thetab) \), the simulator
+    \(M_r(\thetab)\), number of optimization problems \(n_1\), number
+    of samples per region \(n_2\), acceptance limit
+    \(\epsilon\)}\label{alg:romc_algorithm}
+  \begin{algorithmic}[1]
+    \Procedure{ROMC}{}
+    \For{\(i \gets 1 \textrm{ to } n_1\)}
+    \State \(\ub_i \sim p(\ub)\) \Comment{Draw nuisance variables}
+    \State Convert \(M_r(\thetab) \) to \( g(\thetab, \ub=\ub_i) \) \Comment{Define deterministic simulator}
+      \State \( d_i(\thetab) = d(g(\thetab, \ub=\ub_i), \yb_0) \) \Comment{Define distance function}
+      \State \(\thetab_i^* = \text{argmin}_{\thetab} d_i\), \(d_i^*=d_i(\thetab_i^*)\) \Comment{Solve optimization problem}
+      \If{\(d^*_i > \epsilon\)}
+        \State Go to 2 \Comment{Filter solution}
+      \EndIf
+      \State Estimate \(\hat{\accregioni}\) and define \(q_i\) \Comment{Estimate proposal area}
+      \State (Optional) Fit \(\Tilde{d}_i\) on \(\hat{\accregioni}\) \Comment{Fit surrogate model}
+      \\\hrulefill
+      \For{\(j \gets 1 \textrm{ to } n_2\)}
+      \State \(\thetab_{ij} \sim q_i\), compute \(w_{ij}\) as in Algorithm \ref{alg:sampling_GB} \Comment{Sample}
+      \EndFor
+    \EndFor
+    \State \(\E_{p(\thetab|\data)}[h(\thetab)]\) as in eq.~\eqref{eq:expectation} \Comment{Estimate an expectation}
+    \State \(p_{d,\epsilon}(\thetab) \) as in eq.~\eqref{eq:approx_posterior_omc} \Comment{Evaluate the unnormalized posterior}
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\subsubsection*{Training part}
+\noindent
+At the training (fitting) part, the goal is the estimation of the
+proposal regions \(\hat{\accregioni}\), which expands to three
+mandatory tasks; (a) sample the nuisance variables
+\(\ub_i \sim p(\ub)\) for defining the deterministic distance
+functions \(d_i(\thetab)\) (Steps 3-5), (b) solve the optimization
+problems for obtaining \(\thetab_i^*, d_i^*\) and keep the solutions
+inside the threshold \(\epsilon\)(Steps 6-9), and (c) estimate the
+bounding boxes \(\accregionihat\) to define uniform distributions
+\(q_i\) on them (Step 10). Optionally, a surrogate model
+\(\tilde{d}_i\) can be fitted for a faster inference phase.
+
+If \(d_i(\thetab)\) is differentiable, using a gradient-based method
+is advised for obtaining \(\thetab_i^*\) faster. In this case, the
+gradients \(\nabla_{\thetab} d_i\) gradients are approximated
+automatically with finite-differences. This approximation requires two
+evaluations of \(d_i\) for each parameter
+\(\theta_m, m \in \{1, \ldots, D\}\), which scales well only in
+low-dimensional problems. If \(d_i(\thetab)\) is not differentiable,
+Bayesian Optimization can be used as an alternative. In this scenario,
+the training part becomes slower due to fitting of the surrogate model
+and the blind optimization steps.
+
+After obtaining the optimal points \(\thetab^*_i\), we estimate the
+proposal regions. Algorithm~\ref{alg:region_construction} describes
+the line search approach for finding the region boundaries. The axes
+of each bounding box \(\vb_m, m = \{1, \ldots, D\}\) are defined as
+the directions with the highest curvature at \(\thetab_i^*\) computed
+by the eigenvalues of the Hessian matrix \(\hessian_i\) of \(d_i\) at
+\(\thetab_i\) (Step 1). Depending on the algorithm used in the
+optimization step, we either use the real distance \(d_i\) or the
+Gaussian Process approximation \(\hat{d}_i\). When the distance
+function is the Euclidean distance (default choice), the Hessian
+matrix can be also computed as \(\hessian_i = \jac_i^T\jac_i\), where
+\(\jac_i\) is the Jacobian matrix of the summary statistics
+\(\Phi(g(\thetab, \ub_i))\) at \(\thetab_i^*\). This approximation has
+the computational advantage of using only first-order
+derivatives. After defining the axes, we search for the bounding box
+limits with a line step algorithm (Steps 2-21). The key idea is to
+take long steps \(\eta\_\text{start}\) until crossing the boundary
+where \(d > \epsilon\) and then take small steps backwards to find the
+exact boundary position.
+
+\begin{algorithm}[!ht]
+  \caption{Approximation \(\accregioni\) with a bounding box \(\hat{\accregioni}\);
+  Requires: a model of distance \(d_i(\thetab)\),
+  an optimal point \(\thetab_i^*\),
+  a number of refinements \(K\),
+  a step size \(\eta\_\text{start}\),
+  maximum iterations \(M\) and
+  a curvature matrix \(\hessian_i\) (\(\jac_i^T\jac_i \) or GP Hessian)}\label{alg:region_construction}
+  \begin{algorithmic}[1]
+  \State Compute eigenvectors \(\vb_{m}\) of \(\hess_i\) {\scriptsize (\(m = 1,\ldots,D)\)}
+  \For{\(m \gets 1 \textrm{ to } D\)}
+    \State \(\Tilde{\thetab} \gets \thetab_i^*\) \label{algstep:box_constr_start}
+    \State \(k \gets 0\)
+    \State \(\eta \gets \eta\_\text{start}\) \Comment{Initialize \(\eta\)}
+    \Repeat
+          \State \(j \gets 0\)
+          \Repeat
+            \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \eta \ \vb_{m}\) \Comment{Large step size \(\eta\).}
+            \State \(j \gets j + 1\)
+          \Until{\(d(g(\Tilde{\thetab}, \ub=\ub_i), \data) > \epsilon\) or \(j \geq M\)} \Comment{Check distance or maximum iterations}
+          \State \(\Tilde{\thetab} \gets \Tilde{\thetab} - \eta \ \vb_{m}\)
+          \State \(\eta \gets \eta/2\) \Comment{More accurate region boundary}
+          \State \(k \gets k + 1\)
+      \Until \(k = K\)
+      \If{\(\Tilde{\thetab} = \thetab_i^*\)} \Comment{Check if no step has been done}
+        \State \(\Tilde{\thetab} \gets \Tilde{\thetab} + \dfrac{\eta\_{\text{start}}}{2^K} \vb_m\) \Comment{Then, make the minimum step}
+      \EndIf \label{algstep:box_constr_end}
+      \State Set \(\Tilde{\thetab}\) as the positive end point along \(\vb_m\)
+      \State Run steps~\ref{algstep:box_constr_start}~-~\ref{algstep:box_constr_end} for \(\vb_m = - \vb_m\) and set \(\Tilde{\thetab}\) as the negative end point along \(\mathbf{v}_{m}\)
+  \EndFor
+  \State Fit a rectangular box around the region end points and define \(q_i\) as uniform distribution
+  \end{algorithmic}
+\end{algorithm}
+
+\subsubsection*{Inference Part}
+
+The inference part includes one or more of the following three tasks;
+(a) sample from the posterior distribution
+\( \thetab_i \sim p_{d, \epsilon}(\thetab|\data)\)
+(Equation~\ref{eq:sampling}), (b) compute an expectation
+\(\E_{\thetab|\data}[h(\thetab)]\) (Equation~\ref{eq:expectation})
+and/or (c) evaluate the unnormalized posterior
+\(p_{d, \epsilon}(\thetab|\data)\)
+(Equation~\ref{eq:approx_posterior_omc}). Sampling is performed by
+getting \(n_2\) samples from each proposal distribution \(q_i\). For
+each sample \(\thetab_{ij}\), the distance function\footnote{As
+  before, if a surrogate model \(\hat{d}\) is available, it can be
+  utilized as the distance function.} is evaluated for checking if it
+lies inside the acceptance region. Algorithm~\ref{alg:sampling_GB}
+defines the steps for computing a weighted sample. After we obtain
+weighted samples, computing the expectation is straightforward using
+Equation~\ref{eq:expectation}. Finally, evaluating the unnormalized
+posterior at a specific point \(\thetab\) requires access to the
+distance functions \(d_i\) and the prior distribution
+\(p(\thetab)\). Following Equation~\ref{eq:approx_posterior_omc}, the
+evaluation involves counting how many distance functions rating over
+all \(d_i(\thetab) < \epsilon\). It is worth noticing that for
+evaluating the unnormalized posterior, there is no need for solving
+the optimization problems and building the proposal regions.
+
+\begin{algorithm}[H]
+    \centering
+    \caption{Sampling. Requires a function of distance \(d_i\), the prior distribution \(p(\thetab)\), the proposal distribution \(q_i\)}\label{alg:sampling_GB}
+    \begin{algorithmic}[1]
+      \State \(\thetab_{ij} \sim q_i\)
+          \If {\(d_i(\thetab_{ij}) > \epsilon\)}
+            \State Go to 2 \Comment{Reject sample}
+          \Else {}
+            \State \(w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}\) \Comment{Compute weight}
+            \State Store \((w_{ij}, \thetab_{ij})\) \Comment{Store weighted sample}
+          \EndIf
+    \end{algorithmic}
+\end{algorithm}
 
@@
 \label{subsec:general_design}
-\input{latex_files/ch3/3-1-implementation-principles.tex}
+
+\input{latex_files/graphs/romc_graph}
+
+The overview of our implementation is illustrated in
+Figure~\ref{fig:romc_overview}. Following \proglang{Python}'s naming
+principles, the methods starting with an underscore (green rectangles)
+represent internal (private) functions, whereas the rest (blue
+rectangles) are the methods exposed at the API. In
+Figure~\ref{fig:romc_overview}, it can be easily observed that the
+implementation follows Algorithm~\ref{alg:romc_algorithm}. The
+training part includes all the steps until the computation of the
+proposal regions, i.e., sampling the nuisance variables, defining the
+optimization problems, solving them, constructing the regions and
+fitting local surrogate models. The inference part comprises of
+evaluating the unnormalized posterior (and the normalized when is
+possible), sampling and computing an expectation. We also provide some
+utilities for inspecting the training process, such as plotting the
+histogram of the final distances or visualizing the constructed
+bounding boxes. Finally, in the evaluation part, we provide two
+methods for evaluating the inference; (a) computing the Effective
+Sample Size (ESS) of the samples and (b) measuring the divergence
+between the approximate posterior the ground-truth, if the latter is
+available.\footnote{Normally, the ground-truth posterior is not
+  available; However, this functionality is useful in cases where the
+  posterior can be computed numerically or with an alternative method,
+  e.g., Rejection Sampling, and we want to measure the discrepancy
+  between the two approximations.}
+
+\subsubsection*{Parallelizing ROMC method}
+
+As discussed, ROMC has the significant advantage of being fully
+parallelisable. We exploit this fact by implementing a parallel
+version of the major fitting components; (a) solving the optimization
+problems, (b) constructing bounding box regions. We parallelize these
+processes using the built-in \proglang{Python} package
+\pkg{multiprocessing}. The specific package enables concurrency, using
+sub-processes instead of threads, for side-stepping the Global
+Interpreter (GIL). For activating the parallel version of the
+algorithm, the user simply has to set \code{elfi.ROMC(<output_node>,
+  parallelize = True)}.
+
+\subsubsection*{Simple one-dimensional example}
+
+For illustrating the functionalities we will use the following running
+example introduced by \citet{Ikonomov2019},
+
+\begin{gather} \label{eq:1D_example}
+  p(\theta) = \mathcal{U}(\theta;-2.5,2.5)\\ \label{eq:1D_example_eq_2}
+  p(y|\theta) = \left\{
+    \begin{array}{ll} \theta^4 + u & \mbox{if } \theta \in [-0.5, 0.5]
+\\ |\theta| - c + u & \mbox{otherwise}
+    \end{array} \right.\\
+  u \sim \mathcal{N}(0,1)
+\end{gather}
+
+\noindent
+
+The prior is a uniform distribution in the range \([-2.5, 2.5]\) and
+the likelihood is defined at Equation~\ref{eq:1D_example_eq_2}. The
+constant \(c\) is \(0.5 - 0.5^4\) ensures that the PDF is
+continuous. There is only one observation \(y_0 = 0\). The inference
+in this particular example can be performed quite easily without using
+a likelihood-free inference approach. We can exploit this fact for
+validating the accuracy of our implementation. At the following code
+snippet, we code the model at \pkg{ELFI}:
+
+\begin{Code}
+import elfi
+import scipy.stats as ss
+import numpy as np
+
+def simulator(t1, batch_size = 1, random_state = None):
+    c = 0.5 - 0.5**4
+    if t1 < -0.5:
+        y = ss.norm(loc = -t1-c, scale = 1).rvs(random_state = random_state)
+    elif t1 <= 0.5:
+        y = ss.norm(loc = t1**4, scale = 1).rvs(random_state = random_state)
+    else:
+        y = ss.norm(loc = t1-c, scale = 1).rvs(random_state = random_state)
+    return y
+
+# observation
+y = 0
+
+# Elfi graph
+t1 = elfi.Prior('uniform', -2.5, 5)
+sim = elfi.Simulator(simulator, t1, observed = y)
+d = elfi.Distance('euclidean', sim)
+
+# Initialize the ROMC inference method
+bounds = [(-2.5, 2.5)] # bounds of the prior
+parallelize = True # activate parallel execution
+romc = elfi.ROMC(d, bounds = bounds, parallelize = parallelize)
+\end{Code}
 
@@
 \section{Implemented functionalities}
-\input{latex_files/ch4/4-0-intro.tex}
+
+At this section, we analyze the functionalities of the training, the
+inference and the evaluation part. Extended documentation for each
+method can be found in \pkg{ELFI}'s
+\href{https://elfi.readthedocs.io/en/latest/}{official
+  documentation}. Finally, we describe how a user may extend ROMC with
+its custom modules.
 
@@
 \label{subsec:training}
-\input{latex_files/ch4/4-1-training.tex}
+
+In this section, we describe the six functions of the training part:
+
+\begin{Code}
+>>> romc.solve_problems(n1, use_bo = False, optimizer_args = None)
+\end{Code}
+
+\noindent
+This method (a) draws integers for setting the seed, (b) defines the
+optimization problems and (c) solves them using either a
+gradient-based optimizer (default choice) or Bayesian optimization
+(BO), if \code{use_bo = True}. The tasks are completed sequentially,
+as shown in Figure~\ref{fig:romc_overview}. The optimization problems
+are defined after drawing \code{n1} integer numbers from a discrete
+uniform distribution \(u_i \sim \mathcal{U}\{1, 2^{32}-1\}\), where
+each integer \(u_i\) is the seed passed to \pkg{ELFI}'s random
+simulator. The user can pass a \code{Dict} with custom parameters to
+the optimizer through \code{optimizer_args}. For example, in the
+gradient-based case, the user may pass \code{optimizer_args =
+  \{"method": "L-BFGS-B", "jac": jac\}}, to select the
+\code{"L-BFGS-B"} optimizer and use the callable \code{jac} to compute
+the gradients in closed-form.
+
+\begin{Code}
+>>> romc.distance_hist(**kwargs)
+\end{Code}
+
+\noindent
+This function helps the user decide which threshold \(\epsilon\) to
+use by plotting a histogram of the distances at the optimal point
+\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in
+case \code{use_bo = True}. The function forwards the keyword arguments
+to the underlying \code{pyplot.hist()} of the \pkg{matplotlib}
+package. In this way the user may customize some properties of the
+histogram, e.g., the number of bins.
+
+\begin{Code}
+>>> romc.estimate_regions(eps_filter, use_surrogate = None, fit_models = False)
+\end{Code}
+
+\noindent
+This method estimates the proposal regions around the optimal points,
+following Algorithm~\ref{alg:region_construction}. The choice about
+the distance function follows the previous optimization step; if a
+gradient-based optimizer has been used, then estimating the proposal
+region is based on the real distance \(d_i\). If BO is used, then the
+surrogate distance \(\hat{d}\) is chosen. Setting
+\code{use_surrogate=False} enforces the use of the real distance \(d\)
+even after BO. Finally, the parameter \code{fit_models} selects
+whether to fit local surrogate models \(\tilde{d}\) after estimating
+the proposal regions.
+
+\noindent
+The training part includes three more functions. The function
+\code{romc.fit_posterior(args*)} which is a syntactic sugar for
+applying \code{.solve_problems()} and \code{.estimate_regions()} in a
+single step. The function \code{romc.visualize_region(i)} plots the
+bounding box around the optimal point of the \(i\)-th optimization
+problem, when the parameter space is up to \(2D\). Finally,
+\code{romc.compute_eps(quantile)} returns the appropriate distance
+value \(d_{i=\kappa}^*\) where
+\(\kappa = \lfloor quantile \cdot n \rfloor\) from the collection
+\(\{ d_i^* \}, i = \{1, \ldots, n\}\) where \(n\) is the number
+of accepted solutions. It can be used to automate the selection of the
+threshold \(\epsilon\), e.g.,
+\code{eps=romc.compute_eps(quantile=0.9)}.
+
+\subsubsection*{Example - Training part}
+
+In the following snippet, we put together the routines to code the
+training part of the running example.
+
+\begin{Code}
+# Training (fitting) part
+n1 = 500 # number of optimization problems
+seed = 21 # seed for solving the optimization problems
+use_bo = False # set to True for switching to Bayesian optimization
+
+# Training step-by-step
+romc.solve_problems(n1 = n1, seed = seed, use_bo = use_bo)
+romc.theta_hist(bins = 100) # plot hist to decide which eps to use
+
+eps = .75 # threshold for the bounding box
+romc.estimate_regions(eps = eps) # build the bounding boxes
+
+romc.visualize_region(i = 1) # for inspecting visually the bounding box
+
+# Equivalent one-line command
+# romc.fit_posterior(n1 = n1, eps = eps, use_bo = use_bo, seed = seed)
+\end{Code}
+
+\begin{figure}[ht]
+  \begin{center}
+    \input{latex_files/images/chapter3/example_gt.tex}
+    \input{latex_files/images/chapter3/example_post.tex}
+  \end{center}
+  \caption[Histogram of distances at the one-dimensional example.]{Histogram of
+    distances and visualization of a specific region.}
+  \label{fig:running_example_romc_inference}
+\end{figure}
 
@@
 \label{subsec:inference}
-\input{latex_files/ch4/4-2-inference.tex}
+
+In this section, we describe the four functions of the inference part:
+
+\begin{Code}
+>>> romc.sample(n2)
+\end{Code}
+
+\noindent
+This is the basic functionality of the inference, drawing \(n_2\)
+samples for each bounding box region, giving a total of
+\(k \cdot n_2\) samples, where \(k < n_1\) is the number of the
+optimal points remained after filtering. The samples are drawn from a
+uniform distribution \(q_i\) defined over the corresponding bounding
+box and the weight \(w_i\) is computed as in
+Algorithm~\ref{alg:sampling_GB}.
+
+\noindent
+The inference part includes three more function. The function
+\code{romc.compute_expectation(h)} computes the expectation
+\(\E_{p(\thetab|\data)}[h(\thetab)]\) as in
+Equation~\ref{eq:expectation}. The argument \code{h} can be any python
+\code{Callable}. The method \code{romc.eval_unnorm_posterior(theta,
+  eps_cutoff = False)} computes the unnormalized posterior
+approximation as in Equation~\ref{eq:approx_posterior}. The method \\
+\code{romc.eval_posterior(theta, eps_cutoff = False)} evaluates the
+normalized posterior estimating the partition function
+\(Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab\) using Riemann's
+integral approximation. The approximation is computationally feasible
+only in a low-dimensional parametric space.
+
+\subsubsection*{Example - Inference part}
+
+In the following code snippet, we use the inference utilities to (a)
+get weighted samples from the approximate posterior, (b) compute an
+expectation and (c) evaluate the approximate posterior. We also use
+some of \pkg{ELFI}'s built-in tools to get a summary of the obtained
+samples. For \code{romc.compute_expectation()}, we demonstrate its use
+to compute the samples mean and the samples variance. Finally, we
+evaluate \code{romc.eval_posterior()} at multiple points to plot the
+approximate posterior of
+Figure~\ref{fig:running_example_romc_inference}. We observe that the
+approximation is quite close to the ground-truth.
+
+\begin{Code}
+# Inference part
+seed = 21
+n2 = 50
+romc.sample(n2 = n2, seed = seed)
+
+# visualize region, adding the samples now
+romc.visualize_region(i = 1)
+
+# Visualize marginal (built-in ELFI tool)
+weights = romc.result.weights
+romc.result.plot_marginals(weights = weights, bins = 100, range = (-3,3))
+
+# Summarize the samples (built-in ELFI tool)
+romc.result.summary()
+# Number of samples  : 19300
+# Sample means: theta: -0.0116
+
+# compute expectation
+exp_val = romc.compute_expectation(h = lambda x: np.squeeze(x))
+print("Expected value   : %.3f" % exp_val)
+# Expected value: -0.012
+
+exp_var = romc.compute_expectation(h = lambda x: np.squeeze(x)**2)
+print("Expected variance: %.3f" % exp_var)
+# Expected variance: 1.120
+
+# eval unnorm posterior
+romc.eval_unnorm_posterior(theta = 0)
+
+# check eval posterior
+romc.eval_posterior(theta = 0)
+\end{Code}
 
@@
 \label{subsec:evaluation}
-\input{latex_files/ch4/4-3-evaluation.tex}
 
-\subsection{Extensibility} % better title
+The method \code{romc.compute_ess()} computes the Effective Sample
+Size (ESS) as \(\frac{(\sum_i w_i)^2}{\sum_i w_i^2}\), which is a
+useful quantity to measure how many samples actually contribute to an
+expectation. For example, in an extreme case of a big population of
+samples where only one has big weight, the ESS is much smaller than
+the samples population.
+
+\noindent
+The method \code{romc.compute_divergence(gt_posterior, bounds, step,
+  distance)} estimate the divergence between the ROMC approximation
+and the ground truth posterior. Since the estimation is performed
+using Riemann's approximation, the method can only work in low
+dimensional spaces. The method can be used for evaluation in synthetic
+examples where the ground truth is accessible. In a real-case
+scenarios, where it is not expected to have access to the ground-truth
+posterior, the user may set the approximate posterior obtained with
+any other inference approach for comparing the two methods. The
+argument \code{step} defines the step used in the Riemann's
+approximation and the argument \code{distance} can be either
+\code{"Jensen-Shannon"} or \code{"KL-divergence"}.
+
+\begin{Code}
+# Evaluation part
+res = romc.compute_divergence(wrapper, distance = "Jensen-Shannon")
+print("Jensen-Shannon divergence: %.3f" % res)
+# Jensen-Shannon divergence: 0.035
+
+nof_samples = len(romc.result.weights)
+ess = romc.compute_ess()
+print("Nof Samples: %d, ESS: %.3f" % (nof_samples, ess))
+# Nof Samples: 19300, ESS: 16196.214
+\end{Code}
+
+\subsection{Extend the implementation with custom modules}
 \label{subsec:extensibility}
-\input{latex_files/ch4/4-4-extensibility.tex}
+
+ROMC is a generic LFI framework; it describes a sequence of steps for
+approximating the posterior distribution without explicitly enforcing
+the algorithms that solve each step. Although a specific algorithm for
+each step is proposed by~\cite{Ikonomov2019}, in general, a user can
+experiment with alternative algorithms. Considering that, we designed
+the implementation to support flexibility.
+
+We have specified four critical parts where a user may intervene using
+custom solutions; (a) the gradient-based optimization, (b) the
+Bayesian optimization, (c) the proposal region construction and (d)
+the surrogate model fitting. Each of these parts corresponds to a
+function inside the \code{romc.OptimisationProblem} class; (a)
+\code{solve_gradients()}, (b) \code{solve_bo()}, (c)
+\code{build_region()} and (d) \code{fit_local_surrogate()}
+respectively. To replace any of these parts, the user must create a
+custom class that inherits the basic \code{OptimisatioProblem} class
+and overwrite the appropriate function.
+
+To illustrate that in practice, suppose a user wants to fit Deep
+Neural Networks instead of, the default, quadratic models as local
+surrogates \(\tilde{d}_i\). Therefore, the user must create a new
+class that inherits \code{OptimisationProblem} and overwrite the
+\code{fit_local_surrogate(**kwargs)} function with one that fits
+neural networks as local surrogates. We illustrate that in the
+following snippet using the \code{neural_network.MLPRegressor} class
+of the \pkg{scikit-learn} package.
+
+\begin{Code}
+class CustomOptim(OptimisationProblem):
+    def __init__(self, **kwargs):
+        super(CustomOptim, self).__init__(**kwargs)
+
+    def fit_local_surrogate(self, **kwargs):
+        nof_samples = 500
+        objective = self.objective # the distance function
+
+        # helper function
+        def local_surrogate(theta, model_scikit):
+            assert theta.ndim == 1
+            theta = np.expand_dims(theta, 0)
+            return float(model_scikit.predict(theta))
+
+        # create local surrogate model as a function of theta
+        def create_local_surrogate(model):
+            return partial(local_surrogate, model_scikit = model)
+
+        local_surrogates = []
+        for i in range(len(self.regions)):
+            # prepare dataset
+            x = self.regions[i].sample(nof_samples)
+            y = np.array([objective(ii) for ii in x])
+
+            # train Neural Network
+            mlp = MLPRegressor(hidden_layer_sizes = (10,10), solver = 'adam')
+            model = Pipeline([('linear', mlp)])
+            model = model.fit(x, y)
+            local_surrogates.append(create_local_surrogate(model))
+
+        self.local_surrogates = local_surrogates
+        self.state["local_surrogates"] = True
+\end{Code}
+
+In a similar way, the user can replace any of the other three
+functions. In each case, the custom function must change the value of
+some variables that hold the state of the training phase. In the
+following sections, we present which are these variables in each case.
+When implementing the custom functions, the user may need two helping
+classes; (a) \code{RomcOpimisationResult} that stores the result of
+the optimization step and (b) \code{NDimBoundingBox} that stores the
+bounding box. We present the definitions of these classes below and we
+illustrate how to use them in each case in the next sections. Both
+classes can be imported from the module
+\code{elfi.methods.inference.romc}.
+
+\begin{Code}
+class RomcOpimisationResult:
+    def __init__(self, x_min, f_min, hess_appr):
+        Parameters
+        ----------
+        x_min: np.ndarray (D,) or float, the minimum point
+        f_min: float, distance at the minimum point
+        hess_appr: np.ndarray (DxD), Hessian approximation at x_min
+        """
+
+class NDimBoundingBox:
+    def __init__(self, rotation, center, limits, eps_region):
+        Parameters
+        ----------
+        rotation: np.array (D,D),  rotation matrix for the bounding box
+        center: np.array (D,) center of the bounding box
+        limits: np.ndarray, shape: (D,2), the limits of the bounding box
+        eps_region: float, distance threshold
+\end{Code}
+
+\subsubsection*{(a) Extending the gradient-based optimization}
+
+For using a custom gradient-based optimization method, the user must
+overwrite the method \code{solve_gradients()}. The method can take any
+number of keyword arguments. The objective of the optimization problem
+(distance \(d_i\)) can be accessed through the variable
+\code{self.objective}, which is a callable. In the following snippet,
+after the comment \code{state variables}, we present the state
+variables that must be set by the method. The class
+\code{RomcOpimisationResult(x, y, jac, hess_inv)} must be used to
+store the optimization result.
+
+\begin{Code}
+def solve_gradients(self, **kwargs):
+    # inputs
+    # self.objective: the distance function, Callable: np.ndarray (D,) -> float
+
+    # custom solution procedure
+    x, y, jac, hess_inv = ...
+    success = True/False # whether gradient-based optimization was succesful
+
+    # state variables
+    self.state["attempted"] = True
+    if success:
+        self.result = RomcOpimisationResult(x, y, jac, hess_inv)
+        self.state["solved"] = True
+        return True
+    else:
+        return False
+\end{Code}
+
+\subsubsection*{(b) Extending the Bayesian Optimization}
+
+For replacing the Bayesian Optimization function, the procedure is
+similar to the gradient-based case. As presented in the following
+snippet, the only added state variables are (a) setting
+\code{self.surrogate = custom_surrogate}, where
+\code{custom_surrogate} is a \code{Callable} with signature
+\code{np.ndarray (D,) -> float}, and setting
+\code{self.state["has_fit_surrogate"] = True} if the optimization is
+successful.
+
+\begin{Code}
+def solve_bo(self, **kwargs):
+    # inputs
+    # self.objective: the distance function, Callable: np.ndarray (D,) -> float
+
+    # custom procedure
+    x, y, jac, hess_inv = ...
+    custom_surrogate = ...
+    success = True/False # whether Bayesian optimization was succesful
+
+    # state variables
+    self.state["attempted"] = True
+    if success:
+        self.result = RomcOpimisationResult(x, y, jac, hess_inv)
+        self.surrogate = custom_surrogate
+        self.state["solved"] = True
+        self.state["has_fit_surrogate"] = True
+        return True
+    else:
+        return False
+\end{Code}
+
+\subsubsection*{(c) Extending the proposal region construction}
+
+For replacing the proposal region construction the user must overwrite
+the \code{build_region} method. Using the objective function \(d_i\)
+(\code{self.objective}) the method must estimate a list of bounding
+boxes as a \code{List} with \code{NDimBoundingBox} instances and set
+the state variables presented below:
+
+\begin{Code}
+def build_region(self, **kwargs):
+    # inputs
+    # self.objective: the distance function, Callable: np.ndarray (D,) -> float
+
+    # custom build_region method
+    bounding_box: List[NDimBoundingBox] = ...
+    success = True/False # whether region built correctly
+
+    # state variables
+    self.eps_region = eps_region
+    if success:
+        # construct region
+        self.regions = bounding_box
+        self.state["region"] = True
+        return True
+    else:
+        return False
+\end{Code}
+
+\subsubsection*{(d) Extending the surrogate model fitting}
+
+For replacing the surrogate model fitting the user must overwrite the
+\code{fit_local_surrogate} method. Using the objective function
+\(d_i\) (\code{self.objective}) and the estimated bounding boxes
+(\code{self.regions}), the method must estimate a list of local
+surrogates, one for each region, as a \code{List} with callables and
+set the state variables presented below:
+
+\begin{Code}
+def fit_local_surrogate(self, **kwargs):
+    # inputs
+    # self.objective: Callable np.ndarray (D,) -> float, the distance function
+    # self.regions: List[NDimBoundingBox], the bounding boxes
+
+    # custom local surrogates
+    local_surrogates: List[Callable] = ...
+    success = True/False # whether local surrogates fit correctly
+
+    # side-effects
+    if success:
+        self.local_surrogates = custom_surrogates
+        self.state["local_surrogates"] = True
+        return True
+    else:
+        return False
+\end{Code}
 

### 2022-12-29 article/article.tex
finished all paper

@@
   \proglang{Python} package \pkg{ELFI}. ROMC is a novel and efficient
-  (highly-parallelisable) LFI framework that provides accurate
+  (highly-parallelizable) LFI framework that provides accurate
   weighted samples from the posterior. Our implementation can be used
   in two ways. First, a scientist may use it as an out-of-the-box LFI
-  algorithm; we provide an easy-to-use API harmonised with the
+  algorithm; we provide an easy-to-use API harmonized with the
   principles of \pkg{ELFI}, enabling effortless comparisons with the
@@
   scratch. In both scenarios, the ROMC parts can run in a
-  fully-parallelised manner, exploiting all CPU cores. We also provide
+  fully-parallelized manner, exploiting all CPU cores. We also provide
   helpful functionalities for (i) inspecting the inference process and
@@
 processes. Afterwards, \citet{Forneron2016} described a similar method
-under the name `reverse sampler'. In their work,~\citet{Ikonomov2019},
-identified some critical limitations of OMC and proposed Robust OMC
-(ROMC) as an improved version of OMC with appropriate
-modifications.
+under the name `reverse sampler'. In their work,~\citet{Ikonomov2019}
+identified some critical limitations of OMC, so they proposed Robust
+OMC (ROMC) an improved version of OMC with appropriate modifications.
 
 In this paper, we present the implementation of ROMC at the
-\proglang{Python} package \pkg{ELFI} (Engine for likelihood-free
-inference)~\citet{1708.00707}. The implementation has been designed to
-ensure extensibility. ROMC is an LFI framework; it defines a sequence
-of algorithmic steps for approximating the posterior without enforcing
-a specific algorithm for solving each step. Therefore, a researcher
-may use ROMC as the backbone algorithm and develop novel methods to
-solve each separate step.\footnote{For being a ready-to-use
+\proglang{Python} package \pkg{ELFI} (\pkg{Engine for likelihood-free
+  inference})~\citet{1708.00707}. The implementation has been designed
+to ensure extensibility. ROMC is an LFI framework; it defines a
+sequence of algorithmic steps for approximating the posterior without
+enforcing a specific algorithm for solving each step. Therefore, a
+researcher may use ROMC as the backbone algorithm and develop novel
+methods to solve each separate step. For being a ready-to-use
 algorithm,~\citet{Ikonomov2019} proposed a default method for each
-  step, but this choice is by no means restrictive.} We have designed
-our software for facilitating such experimentation.
+step, but this choice is by no means restrictive. We have designed our
+software for facilitating such experimentation.
 
@@
 ROMC inference method to a generic LFI framework. We organize the
-evaluation in three steps. First, for securing that our implementation
-is accurate, we test it against an artificial example with a tractable
-likelihood. The artificial example also serves as a step-by-step guide
-for showcasing how to use the various functionalities of our
-implementation. Second, we use the second-order moving average (MA2)
-example from the \pkg{ELFI} package, using as ground truth the samples
-obtained with Rejection ABC~\citet{lintusaari2017} using a very high
-number of samples. Finally, we present the execution times of ROMC,
-measuring the speed-up achieved by the parallel version of the
+illustration and the evaluation of our implementation in three
+steps. First, for securing that our implementation is accurate, we
+test it against an artificial example with a tractable likelihood. The
+artificial example also serves as a step-by-step guide for showcasing
+how to use the various functionalities of our implementation. Second,
+we use the second-order moving average (MA2) example from the
+\pkg{ELFI} package, using as ground truth the samples obtained with
+Rejection ABC~\citet{lintusaari2017} using a very high number of
+samples. Finally, we present the execution times of ROMC, measuring
+the speed-up achieved by the parallel version of the
 implementation.
+
+The code of the implementation is available at the official \pkg{ELFI} \href{https://github.com/elfi-dev/elfi}{repository}. Apart
+from the examples presented in the paper, there are five \pkg{Google
+  Colab} notebooks available online, with end-to-end examples
+illustrating:
+(i) \href{https://colab.research.google.com/drive/1lGRp0XrNfZ64NN0ASB_tYEKowXwlveDC?usp=sharing}{ROMC on a synthetic \(1D\) example},
+(ii) \href{https://colab.research.google.com/drive/1Fof_WmCi1YizzSI_63aEsbLXsno5gSZ3?usp=sharing}{ROMC on a synthetic \(2D\) example},
+(iii) \href{https://colab.research.google.com/drive/1nkdACQ370SSc0KB1bHv4sBRaxMlMqoNH?usp=sharing}{ROMC on the Moving Average example},
+(iv) \href{https://colab.research.google.com/drive/1_jHVxPSH3XcNOORZJpLU0SPzs0PF8CQ5?usp=sharing}{how to extend ROMC with a Neural Network as a surrogate model},
+(v) \href{https://colab.research.google.com/drive/1RzB-V1QueP1y1nyzv_VOqR1nVz3DUH3v?usp=sharing}{how to extend ROMC with a custom proposal region module}.
 
@@
 code) that generates samples \(\yb\) in a stochastic manner from a set
-of parameters \(\thetab\). We denote the process of obtaining samples
+of parameters \(\thetab \in \mathbb{R}^D\). We denote the process of obtaining samples
 from the simulator with \( m_r(\thetab) \rightarrow \yb \). As shown
@@
 
-As stated at the Introduction, the modeling freedom of simulator-based
-models comes at the price of difficulties in inferring the parameters
-of interest. Denoting the observed data as \(\data\), the main
-difficulty lies at the intractability of the likelihood function
+The modeling freedom of simulator-based models comes at the price of
+difficulties in inferring the parameters of interest. Denoting the
+observed data as \(\data\), the main difficulty lies at the
+intractability of the likelihood function
 \(L(\thetab) = p(\data|\thetab)\). To better see the sources of the
@@
   L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
-  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \theta)
+  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \ub) \in \region(\data)  \mid \thetab)
 \end{equation}
@@
 \(\data\) that is defined through a distance function \(d\), i.e., \
-\(\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon
-\}\). Equation~\ref{eq:likelihood} highlights the main source of
-intractability; computing
+\(\region(\data) := \{\yb: d(\yb, \data) \leq \epsilon \}\). In cases
+where the output \(\yb\) belongs to a high dimensional space, it is
+common to extract summary statistics \(\Phi\) before applying the
+distance \(d\). In these cases, the \(\epsilon\)-region is defined as
+\(\region(\data) := \{\yb: d(\Phi(\yb), \Phi(\data)) \leq \epsilon
+\}\). In our notation, the summary statistics are sometimes omitted
+for simplicity. Equation~\ref{eq:likelihood} highlights the main
+source of intractability; computing
 \(\Pr(\simulator(\thetab,\ub) \in \region(\data) | \thetab) \) as the
@@
   \label{eq:approx_likelihood}
-  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
+  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \thetab), \quad \text{where  } \epsilon > 0.
 \end{equation}
@@
 
-The approximation in
-Equations~\ref{eq:approx_likelihood},~\ref{eq:approx_posterior} is by
+Equations~\ref{eq:approx_likelihood} and~\ref{eq:approx_posterior} is by
 no means the only strategy to deal with the intractability of the
@@
 of the Jacobian matrix of the summary statistics at \(\thetab_i^*\). For
-further details on OMC we refer the reader to \citep{Meeds2015,
+further details on OMC we refer the reader to \citet{Meeds2015,
   Ikonomov2019}.
@@
 and, second, we apply a line-search method along each axis to
-determine the size of the box. We refer the reader to
+determine the size of the bounding box. We refer the reader to
 Algorithm~\ref{alg:region_construction} for the details. After the
@@
 
-For computing the weight \(w_{ij}\) in Equation~\ref{eq:sampling}, we
-need to check whether the samples \(\thetab_{ij}\) drawn from the
-bounding boxes are inside the acceptance region \(\accregioni\). This
+For computing the weight \(w_{ij}\) at Equation~\ref{eq:sampling}, we
+need to check whether the samples \(\thetab_{ij}\), drawn from the
+bounding boxes, are inside the acceptance region \(\accregioni\). This
 can be considered to be a safety-mechanism that corrects for any
@@
 evaluating the weights \(w_{ij}\). Moreover, the extra cost is largely
-eliminated if Bayesian Optimization with a Gaussian process (GP)
+eliminated if Bayesian optimization with a Gaussian process (GP)
 surrogate model \(\hat{d}_i(\thetab)\) was used to obtain
@@
 problems for obtaining \(\thetab_i^*, d_i^*\) and keep the solutions
-inside the threshold \(\epsilon\)(Steps 6-9), and (c) estimate the
+inside the threshold \(\epsilon\) (Steps 6-9), and (c) estimate the
 bounding boxes \(\accregionihat\) to define uniform distributions
 \(q_i\) on them (Step 10). Optionally, a surrogate model
-\(\tilde{d}_i\) can be fitted for a faster inference phase.
+\(\tilde{d}_i\) can be fitted for a faster inference phase (Step 11).
 
@@
 gradients \(\nabla_{\thetab} d_i\) gradients are approximated
-automatically with finite-differences. This approximation requires two
+automatically with finite-differences, if they are not provided in
+closed-form by the user. Finite-differences approximation requires two
 evaluations of \(d_i\) for each parameter
@@
 limits with a line step algorithm (Steps 2-21). The key idea is to
-take long steps \(\eta\_\text{start}\) until crossing the boundary
-where \(d > \epsilon\) and then take small steps backwards to find the
-exact boundary position.
+take long steps \(\eta\_\text{start}\) until crossing the boundary and
+then take small steps backwards to find the exact boundary position.
 
@@
 distance functions \(d_i\) and the prior distribution
-\(p(\thetab)\). Following Equation~\ref{eq:approx_posterior_omc}, the
-evaluation involves counting how many distance functions rating over
-all \(d_i(\thetab) < \epsilon\). It is worth noticing that for
+\(p(\thetab)\). Following Equation~\ref{eq:approx_posterior_omc}, we
+simply count for how many deterministic distance functions it holds
+that \(d_i(\thetab) < \epsilon\). It is worth noticing that for
 evaluating the unnormalized posterior, there is no need for solving
@@
     \begin{algorithmic}[1]
-      \State \(\thetab_{ij} \sim q_i\)
-          \If {\(d_i(\thetab_{ij}) > \epsilon\)}
-            \State Go to 2 \Comment{Reject sample}
-          \Else {}
+      \State \(\thetab_{ij} \sim q_i \forall i\) \Comment{Sample parameters}
+      \For {\(i \gets 1 \textrm{ to } n_1\)}
+      \For {\(j \gets 1 \textrm{ to } n_2\)}
+          \If {\(d_i(\thetab_{ij}) \leq \epsilon\)} \Comment{Accept sample}
             \State \(w_{ij} = \frac{p(\thetab_{ij})}{q(\thetab_{ij})}\) \Comment{Compute weight}
@@
             \EndIf
+            \EndFor
+    \EndFor
     \end{algorithmic}
@@
 The overview of our implementation is illustrated in
-Figure~\ref{fig:romc_overview}. Following \proglang{Python}'s naming
+Figure~\ref{fig:romc_overview}. Following \proglang{Python} naming
 principles, the methods starting with an underscore (green rectangles)
@@
 rectangles) are the methods exposed at the API. In
-Figure~\ref{fig:romc_overview}, it can be easily observed that the
+Figure~\ref{fig:romc_overview}, it can be observed that the
 implementation follows Algorithm~\ref{alg:romc_algorithm}. The
@@
 
-\subsubsection*{Parallelizing ROMC method}
+\subsubsection*{Parallel version of ROMC}
 
@@
 Interpreter (GIL). For activating the parallel version of the
-algorithm, the user simply has to set \code{elfi.ROMC(<output_node>,
-  parallelize = True)}.
+algorithm, the user simply has to set \\
+\code{elfi.ROMC(<output_node>, parallelize = True)}.
 
@@
 
-ROMC is a generic LFI framework; it describes a sequence of steps for
-approximating the posterior distribution without explicitly enforcing
-the algorithms that solve each step. Although a specific algorithm for
-each step is proposed by~\cite{Ikonomov2019}, in general, a user can
-experiment with alternative algorithms. Considering that, we designed
-the implementation to support flexibility.
+ROMC is a generic LFI framework as it describes a sequence of steps
+for approximating the posterior distribution without explicitly
+enforcing a specific algorithm for each step. For completeness,
+\cite{Ikonomov2019} propose a method for each step but, in general, a
+user can experiment with alternative methods. Considering that, we
+designed the implementation to support flexibility.
 
 We have specified four critical parts where a user may intervene using
-custom solutions; (a) the gradient-based optimization, (b) the
-Bayesian optimization, (c) the proposal region construction and (d)
-the surrogate model fitting. Each of these parts corresponds to a
-function inside the \code{romc.OptimisationProblem} class; (a)
+custom methods; (a) gradient-based optimization, (b) Bayesian
+optimization, (c) proposal region construction and (d) surrogate model
+fitting. Each of these parts corresponds to an internal function
+inside the \code{romc.OptimisationProblem} class; (a)
 \code{solve_gradients()}, (b) \code{solve_bo()}, (c)
-\code{build_region()} and (d) \code{fit_local_surrogate()}
+\code{build_region()} and (d) \code{fit_local_surrogate()},
 respectively. To replace any of these parts, the user must create a
-custom class that inherits the basic \code{OptimisatioProblem} class
-and overwrite the appropriate function.
-
-To illustrate that in practice, suppose a user wants to fit Deep
+custom class that inherits \code{OptimisatioProblem} and overwrite the
+appropriate function(s).
+
+To illustrate this in practice, suppose a user wants to fit Deep
 Neural Networks instead of, the default, quadratic models as local
@@
 following snippet using the \code{neural_network.MLPRegressor} class
-of the \pkg{scikit-learn} package.
+of the \pkg{scikit-learn} package. The reader can find the end-to-end
+example
+\href{https://colab.research.google.com/drive/1_jHVxPSH3XcNOORZJpLU0SPzs0PF8CQ5?usp=sharing}{here}
+as an online Colab notebook.
 
@@
 In a similar way, the user can replace any of the other three
-functions. In each case, the custom function must change the value of
-some variables that hold the state of the training phase. In the
-following sections, we present which are these variables in each case.
-When implementing the custom functions, the user may need two helping
-classes; (a) \code{RomcOpimisationResult} that stores the result of
-the optimization step and (b) \code{NDimBoundingBox} that stores the
-bounding box. We present the definitions of these classes below and we
-illustrate how to use them in each case in the next sections. Both
-classes can be imported from the module
+functions. In each case, the custom function must update some
+class-level variables that hold the state of the training phase. In
+the following sections, we present which are these variables in each
+function. Furthermore, when implementing custom functions, the user
+may use two helping classes; (a) \code{RomcOpimisationResult}, that
+stores the result of the optimization and (b) \code{NDimBoundingBox},
+that stores the bounding box. We present their definitions in the
+following snippets and we illustrate how to use them in the next
+sections. Both classes can be imported from the module
 \code{elfi.methods.inference.romc}.
@@
         ----------
-        x_min: np.ndarray (D,) or float, the minimum point
-        f_min: float, distance at the minimum point
-        hess_appr: np.ndarray (DxD), Hessian approximation at x_min
-        """
+        x_min: np.ndarray (D,), minimum
+        f_min: float, distance at x_min
+        hess_appr: np.ndarray (D,D), Hessian approximation at x_min
 
 class NDimBoundingBox:
-    def __init__(self, rotation, center, limits, eps_region):
+    def __init__(self, rotation, center, limits):
         Parameters
         ----------
-        rotation: np.array (D,D),  rotation matrix for the bounding box
-        center: np.array (D,) center of the bounding box
-        limits: np.ndarray, shape: (D,2), the limits of the bounding box
-        eps_region: float, distance threshold
+        rotation: np.ndarray (D,D), rotation matrix for the bounding box
+        center: np.ndarray (D,) center of the bounding box
+        limits: np.ndarray (D,2), the limits of the bounding box
 \end{Code}
 
-\subsubsection*{(a) Extending the gradient-based optimization}
-
-For using a custom gradient-based optimization method, the user must
-overwrite the method \code{solve_gradients()}. The method can take any
-number of keyword arguments. The objective of the optimization problem
-(distance \(d_i\)) can be accessed through the variable
-\code{self.objective}, which is a callable. In the following snippet,
-after the comment \code{state variables}, we present the state
-variables that must be set by the method. The class
-\code{RomcOpimisationResult(x, y, jac, hess_inv)} must be used to
-store the optimization result.
+\subsubsection*{(a) Extending gradient-based optimization}
+
+For replacing the default gradient-based optimization method, the user
+must overwrite the function \code{solve_gradients()}. Using the
+objective function \(d_i\) (\code{self.objective}), the custom method
+must store the result of the optimization as a
+\code{RomcOptimisationResult} instance. In the following snippet,
+after the comment \code{\# state variables}, we present the
+class-level variables that must be set by the method.
 
@@
 def solve_gradients(self, **kwargs):
-    # inputs
-    # self.objective: the distance function, Callable: np.ndarray (D,) -> float
-
-    # custom solution procedure
-    x, y, jac, hess_inv = ...
-    success = True/False # whether gradient-based optimization was succesful
+    # useful variables
+    # self.objective: Callable, the distance function
+
+    # code custom solution here
+    result = RomcOptimisationResult(x = .., y = .., jac = .., hess_inv = ..)
+    success: bool = ... # whether optimization was successful
 
@@
     if success:
-        self.result = RomcOpimisationResult(x, y, jac, hess_inv)
+        self.result = result
         self.state["solved"] = True
@@
 
-\subsubsection*{(b) Extending the Bayesian Optimization}
-
-For replacing the Bayesian Optimization function, the procedure is
-similar to the gradient-based case. As presented in the following
-snippet, the only added state variables are (a) setting
-\code{self.surrogate = custom_surrogate}, where
-\code{custom_surrogate} is a \code{Callable} with signature
-\code{np.ndarray (D,) -> float}, and setting
+\subsubsection*{(b) Extending Bayesian Optimization}
+
+For replacing the default Bayesian optimization function, the
+procedure is similar to the gradient-based case. As presented in the
+following snippet, the additional class-level variables that must be
+set are; (a) \code{self.surrogate = custom_surrogate}, where
+\code{custom_surrogate} is a \code{Callable} and (b)
 \code{self.state["has_fit_surrogate"] = True} if the optimization is
@@
 def solve_bo(self, **kwargs):
-    # inputs
-    # self.objective: the distance function, Callable: np.ndarray (D,) -> float
-
-    # custom procedure
-    x, y, jac, hess_inv = ...
-    custom_surrogate = ...
-    success = True/False # whether Bayesian optimization was succesful
+    # useful variables
+    # self.objective: Callable, the distance function
+
+    # code custom solution here
+    result = RomcOptimisationResult(x = .., y = .., jac = .., hess_inv = ..)
+    custom_surrogate = ... # store a Callable here
+    success: bool = ... # whether optimization was successful
 
@@
     if success:
-        self.result = RomcOpimisationResult(x, y, jac, hess_inv)
+        self.result = result
         self.surrogate = custom_surrogate
@@
 
-For replacing the proposal region construction the user must overwrite
-the \code{build_region} method. Using the objective function \(d_i\)
-(\code{self.objective}) the method must estimate a list of bounding
-boxes as a \code{List} with \code{NDimBoundingBox} instances and set
-the state variables presented below:
+For replacing the construction of the proposal region the user must
+overwrite the \code{build_region} method. Using the objective function
+\(d_i\) (\code{self.objective}) the method must estimate a list of
+bounding boxes as a \code{List} with \code{NDimBoundingBox} instances
+and set the state variables presented below. An end-to-end example for
+using a custom region construction module can be found
+\href{https://colab.research.google.com/drive/1RzB-V1QueP1y1nyzv_VOqR1nVz3DUH3v?usp=sharing}{here}.
 
@@
 def build_region(self, **kwargs):
-    # inputs
-    # self.objective: the distance function, Callable: np.ndarray (D,) -> float
+    # useful variables
+    # self.objective: Callable, the distance function
 
     # custom build_region method
+    eps: float = ... # epsilon used in region estimation
     bounding_box: List[NDimBoundingBox] = ...
-    success = True/False # whether region built correctly
+    success: bool = ... # whether region built successfully
 
     # state variables
-    self.eps_region = eps_region
+    self.eps_region = eps
     if success:
@@
 \(d_i\) (\code{self.objective}) and the estimated bounding boxes
-(\code{self.regions}), the method must estimate a list of local
-surrogates, one for each region, as a \code{List} with callables and
-set the state variables presented below:
+(\code{self.regions}), the method must create a list of local
+surrogates, one for each region, as a \code{List} with
+\code{Callables} and set the state variables as presented in the
+following snippet.
 
@@
 def fit_local_surrogate(self, **kwargs):
-    # inputs
-    # self.objective: Callable np.ndarray (D,) -> float, the distance function
+    # useful variables
+    # self.objective: Callable, the distance function
     # self.regions: List[NDimBoundingBox], the bounding boxes
@@
     # custom local surrogates
-    local_surrogates: List[Callable] = ...
-    success = True/False # whether local surrogates fit correctly
-
-    # side-effects
+    local_surrogates: List[Callable] = ... # the surrogate models
+    success: bool = ... # whether surrogates fit successfully
+
+    # state variables
     if success:
@@
 \section{Use-case illustration}
-\input{latex_files/ch5/5-ma2.tex}
+
+In this section, we test the implementation using the second-order
+moving average (MA2) example, which is one of the standard models of
+\pkg{ELFI}. We perform the inference using three different versions of
+ROMC; (i) with a gradient-based optimizer, (ii) with Bayesian
+Optimization and (iii) fitting a Neural Network as a surrogate
+model. The later illustrates how to extend the implementation,
+replacing part of ROMC with a user-defined component. Finally, we
+measure the execution speed-up using the parallel version of ROMC.
+
+\subsubsection*{Model Definition}
+
+MA2 is a probabilistic model for time series analysis. The observation
+at time \(t\) is given by,
+
+\begin{gather} \label{eq:ma2}
+y_t = w_t + \theta_1 w_{t-1} + \theta_2 w_{t-2}, \quad t=1, \ldots, T\\
+\theta_1, \theta_2 \in \R, \quad  w_k \sim \mathcal{N}(0,1), k \in \mathbb{Z}
+\end{gather}
+
+\noindent
+The random variable \(w_{k} \sim \mathcal{N}(0,1) \) is white noise
+and the two parameters of interest, \(\theta_1, \theta_2\), model the
+dependence from the previous observations. The parameter \(T\) is the
+number of sequential observations which is set to \(T=100\). For
+securing that the inference problem is identifiable, i.e., the
+likelihood has only one mode, we use the prior proposed
+by~\cite{Marin2012},
+
+\begin{equation} \label{eq:ma2_prior}
+p(\thetab) = p(\theta_1)p(\theta_2|\theta_1)
+= \mathcal{U}(\theta_1;-2,2)\mathcal{U}(\theta_2;\theta_1-1, \theta_1+1)
+\end{equation}
+
+\noindent
+The observation vector \(\yb_0 = (y_1, \ldots, y_{100})\) is generated
+with \(\thetab^*=(0.6, 0.2)\). The dimensionality of the output
+\(\yb\) is high, therefore we use summary statistics. Considering that
+the output vector represents a time-series signal, we select the
+autocovariances with \(\mathrm{lag}=1\) and \(\mathrm{lag}=2\), as
+shown in Equations~\ref{eq:ma2_summary_1}
+and~\ref{eq:ma2_summary_2}. The distance between the observation and
+the simulator output is measured with the squared Euclidean distance,
+as shown in Equation~\ref{eq:ma2_summary_4}.
+
+\begin{gather}
+  \label{eq:ma2_summary_1} s_1(\yb) = \frac{1}{T-1} \sum_{t=2}^T y_ty_{t-1}\\
+  \label{eq:ma2_summary_2} s_2(\yb) = \frac{1}{T-2} \sum_{t=3}^T y_ty_{t-2} \\
+  s(\yb) = (s_1(\yb), s_2(\yb))\\
+  \label{eq:ma2_summary_4} d = ||s(\yb) - s(\yb_0)||_2^2
+\end{gather}
+
+\subsubsection*{Inference}
+
+To demonstrate the full capabilities of our ROMC implementation, we
+perform inference using three different methods: (i) a gradient-based
+optimizer, (ii) Bayesian optimization, and (iii) fitting a neural
+network (NN) as a surrogate model. The use of a NN as a surrogate
+model serves as an example of the extensibility of our implementation,
+as described in Chapter~\ref{subsec:extensibility}. For the NN, we
+employ the \pkg{MLPRegressor} class from the \pkg{scikit-learn}
+package. The NN (\(\tilde{d}_i\)) substitutes the actual distance
+function (\(d_i\)) inside the proposal regions. Therefore, all
+inference actions, namely, sampling, expectation computation, and
+posterior evaluation are based on \(\tilde{d}_i\). We use a NN with
+two hidden layers of 10 neurons each and train it using 500 examples
+from each proposal region. To compare the results of ROMC inference
+with a traditional ABC algorithm, we also include Rejection Sampling
+in our analysis.
+
+In Figure \ref{fig:ma2_5}, we illustrate the acceptance region of the
+same deterministic simulator, in the gradient-based and the Bayesian
+optimization case. The acceptance regions are quite similar even though the
+different optimization schemes lead to different optimal points.
+
+In Figure \ref{fig:ma2_3}, we demonstrate the histograms of the
+marginal posteriors, for each approach; (a) Rejection ABC, (b) ROMC
+with gradient-based optimization (c) ROMC with Bayesian optimization
+and (d) ROMC with the NN extension. We observe a significant agreement
+between the different approaches. At Table \ref{tab:ma2} we present
+the empirical mean \(\mu\) and standard deviation \(\sigma\) for each
+inference approach and finally, in Figure \ref{fig:ma2_4}, we
+illustrate the unnormalized posterior for the three different
+variations of the ROMC method. The results show that all ROMC
+variations provide consistent results between them which are in
+agreement with the Rejection ABC algorithm.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ c|c|c|c|c }
+\hline
+& \(\mu_{\theta_1}\) & \(\sigma_{\theta_1}\) & \(\mu_{\theta_2}\) & \(\sigma_{\theta_2}\) \\
+\hline \hline
+Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
+\hline
+ROMC (gradient-based) & 0.501 & 0.142 & 0.033 & 0.169 \\
+\hline
+ROMC (Bayesian optimization) & 0.494 & 0.16 & 0.086 & 0.167 \\
+\hline
+ROMC (Neural Network) & 0.491 & 0.138 & 0.04 & 0.172 \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Comparison of the samples obtained from the estimated
+  posterior with (a) Rejection sampling and (b) the different versions
+  of ROMC. We observe that the obtained samples share similar
+  statistics along all methods. \label{tab:ma2}}
+\end{table}
+
+\begin{figure}[ht]
+    \begin{center}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.png}
+    \end{center}
+  \caption[The acceptance region of a specific deterministic simulator.]{The acceptance region in a specific optimization problem.
+In the left figure the region obtained with gradient-based optimizer and in the right one with Bayesian Optimization.}
+  \label{fig:ma2_5}
+\end{figure}
+
+\begin{figure}[ht]
+  \begin{center}
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex}
+    }\\
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}
+    }
+    \resizebox{.24\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex}
+    }
+    \end{center}
+    \caption[MA2 example, evaluation of the marginal
+    distributions.]{Histogram of the marginal posterior distributions
+      using three different inference approaches; (a) in the first
+      row, the samples are obtained using Rejection ABC sampling (b)
+      in the second row, using ROMC with a gradient-based optimizer
+      and (c) in the third row, using ROMC with Bayesian optimization
+      approach. The vertical (red) line represents the samples mean
+      \(\mu\) and the horizontal (black) the standard deviation
+      \(\sigma\).}
+  \label{fig:ma2_3}
+\end{figure}
+
+\begin{figure}[ht]
+  \begin{center}
+    \resizebox{.32\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_romc_posterior.tex}
+    }
+    \resizebox{.32\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_romc_posterior_bo.tex}
+    }
+    \resizebox{.32\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/mae2_romc_posterior_nn.tex}
+    }
+    \end{center}
+    \caption[MA2 example, posterior distribution.]{The unnormalized posterior distribution using the ROMC method with (a) a gradient-based optimization (b) Bayesian Optimization (c) gradient-based with a Neural Network as a surrogate model.}
+  \label{fig:ma2_4}
+\end{figure}
+
+\subsubsection*{Parallelize the implementation}
+
+As stated above, ROMC is an approach that can be executed in a
+fully-parallelized manner, exploiting all CPU cores. In our
+implementation, we support a parallel version of the the training
+part, namely, for solving the optimization problems and for estimating
+the proposal regions. The parallel version of the algorithm is built
+on top of the built-in \proglang{Python} package \pkg{multiprocess}
+for using all the available CPU cores. In
+Figure~\ref{fig:exec_parallel} we observe the execution times for
+performing the inference on the MA2 model; the parallel version
+performs both tasks almost five times faster than the sequential. The
+result is reasonable given that the experiments have run in a single
+machine with the Intel® Core™ i7-8750H Processor, which has six
+separate cores.
+
+\begin{figure}[ht]
+  \begin{center}
+    \resizebox{.49\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/exec_time_solve.tex}
+    }
+    \resizebox{.49\columnwidth}{!}{%
+      \input{./latex_files/images/chapter4/exec_time_regions.tex}
+    }
+    \end{center}
+    \caption[Execution time exploiting Parallelize]{Comparison
+      between parallel and sequential execution of ROMC. We observe
+      that the parallel version runs almost 5 times faster.}
+      \label{fig:exec_parallel}
+\end{figure}
 
@@
 \section{Summary and discussion} \label{sec:summary}
-\input{latex_files/ch6/6-summary.tex}
+
+In this paper, we presented the implementation of the LFI method ROMC
+at the \proglang{Python} package \pkg{ELFI}. We highlighted two
+different use-cases. First, we illustrated how a user may exploit the
+provided API to solve an LFI problem. Second, we focus on the scenario
+where a researcher wants to intervene and alter parts of the method to
+experiment with new approaches. Since (Robust) Optimization Monte
+Carlo is a novel approach for statistical inference and, to the best
+of our knowledge, this is the first open-source implementation on a
+generic package, we believe that the later is the biggest
+contribution.
+
+There are still open challenges for enabling ROMC to solve
+high-dimensional LFI problems efficiently. The first is enabling the
+execution of ROMC execution into a distributed environment, i.e., a
+cluster of computers. ROMC can be characterized as an
+\textit{embarrassingly parallel} workload; each optimization problem
+is an entirely independent task. Therefore, supporting inference into
+a cluster of computers can radically speed up the inference. The
+second is the implementation of the method on a framework that supports
+automatic differentiation. Automatic differentiation is necessary for
+efficiently solving optimization problems, especially in
+high-dimensional parametric models.
 

### 2022-12-30 article/article.tex
rerun script with figures

@@
   \begin{center}
-    \input{latex_files/images/chapter3/example_gt.tex}
-    \input{latex_files/images/chapter3/example_post.tex}
+    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_gt.pdf}
+    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_posterior.pdf}
+    % \input{latex_files/images/chapter3/example_gt.tex}
+    % \input{latex_files/images/chapter3/example_post.tex}
   \end{center}
@@
 \hline
-ROMC (Bayesian optimization) & 0.494 & 0.16 & 0.086 & 0.167 \\
+ROMC (Bayesian optimization) & 0.501 & 0.17 & 0.087 & 0.175 \\
 \hline
@@
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.png}
-        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.png}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.pdf}
+        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.pdf}
     \end{center}
@@
   \begin{center}
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t1_rejection.tex}
-    }
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t1_romc.tex}
-    }
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.tex}
-    }
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.tex}
-    }\\
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t2_rejection.tex}
-    }
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t2_romc.tex}
-    }
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex}
-    }
-    \resizebox{.24\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.tex}
-    }
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.pdf}
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.pdf}
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.pdf}
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.pdf}
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.pdf}
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.pdf}
+    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.pdf}
     \end{center}
@@
   \begin{center}
-    \resizebox{.32\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_romc_posterior.tex}
-    }
-    \resizebox{.32\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_romc_posterior_bo.tex}
-    }
-    \resizebox{.32\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/mae2_romc_posterior_nn.tex}
-    }
+    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.pdf}
+    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.pdf}
+    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_nn.pdf}
     \end{center}
@@
   \begin{center}
-    \resizebox{.49\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/exec_time_solve.tex}
-    }
-    \resizebox{.49\columnwidth}{!}{%
-      \input{./latex_files/images/chapter4/exec_time_regions.tex}
-    }
+    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_solve.pdf}
+    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_regions.pdf}
     \end{center}

### 2022-12-30 article/latex_files/images/images/chapter4/mae2_hist_t2_romc_nn.tex
rerun script with figures

--- 2022-02-15 article/latex_files/images/chapter4/mae2_hist_t2_romc_bo.tex
@@
 tick pos=left,
-title={ROMC (BO) - \(\displaystyle \theta_2\)},
+title={ROMC (Neural Network) - \(\displaystyle \theta_2\)},
 x grid style={white!69.0196078431373!black},
@@
 \draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0);
-\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0);
-\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0);
-\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0);
-\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.0617545870506399);
-\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.494036696405118);
-\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.308772935253199);
-\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.864564218708957);
-\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.11158256691152);
-\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,0.802809631658317);
-\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.49395771225155);
-\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.71306067268783);
-\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.17926096548025);
-\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.06759941441516);
-\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.5396445345721);
-\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,1.91609586839403);
-\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,1.79258669429275);
-\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,1.59403798219366);
-\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,1.58667684276459);
-\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,1.52979906798699);
-\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.15183142210051);
-\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.06111444803853);
-\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.47292129320938);
-\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.85529643854941);
-\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.86505006309548);
-\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,2.67369166423286);
-\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,1.96552066170352);
-\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,1.29148734508544);
-\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,1.68849292724457);
-\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,1.00470598608042);
-\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.703294190256297);
-\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.703294190256294);
-\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,1.2056471832965);
-\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.200941197216084);
-\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.100470598608042);
-\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0);
-\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0);
-\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0);
-\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0);
-\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0);
+\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.0557676847180302);
+\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.134391393873674);
+\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.232670307676795);
+\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.421300598168856);
+\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.599565962284508);
+\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.819487590375745);
+\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,1.02995995908051);
+\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,1.25665578855782);
+\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.57888096147667);
+\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.58957701821668);
+\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.49600073368982);
+\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.52507821911024);
+\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.62625669151916);
+\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.92752590520997);
+\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.09127677525018);
+\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.20245981684059);
+\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.32191116112789);
+\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.50351769805879);
+\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.54297681090209);
+\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.14701105787528);
+\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,1.97050861768377);
+\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,1.65952774547258);
+\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.41333954655085);
+\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.1362319294878);
+\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,0.992545439733498);
+\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,0.95508724072649);
+\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.820509887123889);
+\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.693510954529521);
+\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.644349818384951);
+\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.49941685245109);
+\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.304905545384957);
+\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.137853649275808);
+\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.144181418842862);
+\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.0965680200621962);
+\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.0871794625561498);
+\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.087179462556149);
+\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.0871794625561498);
+\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.107297800069107);
+\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0603550125388724);
 \draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0);
@@
 table {%
-0.100012161271839 0
-0.100012161271839 3
+0.0396457758194439 0
+0.0396457758194439 3
 };
-\addlegendentry{$\mu = 0.100$}
+\addlegendentry{$\mu = 0.040$}
 \addplot [semithick, black, dashed]
 table {%
--0.108452084584719 1
-0.278478839382765 1
+-0.170712448191162 1
+0.207933154993939 1
 };
-\addlegendentry{$\sigma = 0.176$}
+\addlegendentry{$\sigma = 0.172$}
 \end{axis}

### 2022-12-30 replication_script/replication_script.py

@@ -184,8 +184,10 @@ plt.legend()
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_gt.png"), bbox_inches='tight')
-    tplt.clean_figure()
-    tplt.save(os.path.join(prepath, "chapter3/example_gt.tex"))
+    plt.savefig(os.path.join(
+        prepath, "chapter3/example_gt.pdf"), bbox_inches='tight')
+    # tplt.clean_figure()
+    # tplt.save(os.path.join(prepath, "chapter3/example_gt.tex"))
 plt.show(block=False)
 
 
@@ -298,8 +300,10 @@ plt.legend()
 if savefig:
     plt.savefig(os.path.join(
         prepath, "chapter3/example_posterior.png"), bbox_inches='tight')
+    plt.savefig(os.path.join(
+        prepath, "chapter3/example_posterior.pdf"), bbox_inches='tight')
     # tplt.clean_figure()
-    tplt.save(os.path.join(prepath, "chapter3/example_post.tex"))
+    # tplt.save(os.path.join(prepath, "chapter3/example_post.tex"))
 plt.show(block=False)
 
 
@@ -348,7 +352,7 @@ def plot_marginal(samples, weights, mean, std, title, xlabel, ylabel, bins, rang
     plt.legend()
     if savepath:
         plt.savefig(savepath, bbox_inches='tight')
-        tplt.save(savepath[:-3] + "tex")
+        # tplt.save(savepath[:-3] + "tex")
     plt.show(block=False)
 
 # helping functions for ploting the 2D posterior pdf
@@ -371,7 +375,7 @@ def plot_romc_posterior(title, posterior, nof_points, savefig):
     plt.colorbar()
     if savefig:
         plt.savefig(savefig, bbox_inches='tight')
-        tplt.save(savefig[:-3] + "tex")
+        # tplt.save(savefig[:-3] + "tex")
     plt.show(block=False)
 
 # Set seed for reproducibility
@@ -418,6 +422,8 @@ tmp = romc.sample(n2=n2, seed=seed)
 if savefig:
     romc.visualize_region(vis_ind_1, savefig=os.path.join(
         prepath, "ma2_region_1.png"))
+    romc.visualize_region(vis_ind_1, savefig=os.path.join(
+        prepath, "ma2_region_1.pdf"))
 else:
     romc.visualize_region(vis_ind_1)
 
@@ -448,6 +454,14 @@ if savefig:
[…]

### 2022-12-30 article/article.tex
finalized code

@@
-\documentclass[article]{jss}
+\documentclass[article, shortnames]{jss}
 %% -- LaTeX packages and custom commands ---------------------------------------
@@
 %% -- Introduction -------------------------------------------------------------
-\section{Introduction}
+\section[Introduction]{Introduction}
 \label{sec:intro}
@@
 % \clearpage
-\section{Background}
+\section[Background]{Background}
 
@@
 code) that generates samples \(\yb\) in a stochastic manner from a set
-of parameters \(\thetab \in \mathbb{R}^D\). We denote the process of obtaining samples
-from the simulator with \( m_r(\thetab) \rightarrow \yb \). As shown
-by~\citet{Meeds2015}, it is feasible to isolate the randomness of the
-simulator by introducing a set of nuisance random variables denoted by
-\(\ub \sim p(\ub)\). Therefore, for a specific tuple
-\((\thetab, \ub)\) the simulator becomes a deterministic mapping
-\(g\), such that \(\yb=\simulator(\thetab,\ub)\). In terms of computer
-code, the randomness of a random process is governed by the global
-seed. Although each package may handle the randomness in slightly
-different ways\footnote{For example, at
-  \pkg{Numpy}~\citet{harris2020array}, the pseudo-random number
-  generation is based on a global state, whereas, in
-  \pkg{JAX}~\citet{jax2018github} random functions consume a key that
-  is passed as parameter. }, in all cases, setting the initial seed to
-a specific integer converts the simulation to a deterministic piece of
-code.
+of parameters \(\thetab \in \mathbb{R}^D\). We denote the process of
+obtaining samples from the simulator with
+\( m_r(\thetab) \rightarrow \yb \). As shown by~\citet{Meeds2015}, it
+is feasible to isolate the randomness of the simulator by introducing
+a set of nuisance random variables denoted by \(\ub \sim
+p(\ub)\). Therefore, for a specific tuple \((\thetab, \ub)\) the
+simulator becomes a deterministic mapping \(g\), such that
+\(\yb=\simulator(\thetab,\ub)\). In terms of computer code, the
+randomness of a random process is governed by the global seed. There
+are some differences on how each scientific package handles the
+randomness; for example, at \pkg{Numpy}~\citet{harris2020array} the
+pseudo-random number generation is based on a global state, whereas,
+at \pkg{JAX}~\citet{jax2018github} random functions consume a key that
+is passed as a parameter. However, in all cases, setting the initial
+seed to a specific integer converts the simulation to a deterministic
+piece of code.
 
@@
 each sample \(\thetab_{ij}\), the distance function\footnote{As
-  before, if a surrogate model \(\hat{d}\) is available, it can be
-  utilized as the distance function.} is evaluated for checking if it
-lies inside the acceptance region. Algorithm~\ref{alg:sampling_GB}
+  before, a surrogate model \(\hat{d}\) can be utilized as the
+  distance function if it is available.} is evaluated for checking if
+it lies inside the acceptance region. Algorithm~\ref{alg:sampling_GB}
 defines the steps for computing a weighted sample. After we obtain
@@
 
-The results in this paper were obtained using \proglang{Python}~3.7.9
-with the \pkg{ELFI}~0.8.3 package. The experiments have been executed
-in a single machine with an Intel® Core™ i7-8750H Processor an with
-Ubuntu 20.04 lts operating system.
+The results in this paper were obtained using \proglang{Python~3.7.9}
+with the \pkg{ELFI~0.8.3} package. The experiments have been executed
+in a single machine with an \proglang{Intel® Core™ i7-8750H} Processor
+an with \proglang{Ubuntu 20.04} lts operating system.
 

### 2022-12-30 replication_script/replication_script.py

@@ -13,8 +13,8 @@ import matplotlib.pyplot as plt
 from elfi.examples import ma2
 import numpy as np
 import elfi
-import tikzplotlib as tplt
 import matplotlib
+
 matplotlib.rcParams['text.usetex'] = True
 logging.basicConfig(level=logging.INFO)
 savefig = False
@@ -183,11 +183,7 @@ plt.plot(theta, y, 'r-.', label=r'Posterior: $p(\theta|y_0)$')
 plt.legend()
 if savefig:
     plt.savefig(os.path.join(
-        prepath, "chapter3/example_gt.png"), bbox_inches='tight')
-    plt.savefig(os.path.join(
         prepath, "chapter3/example_gt.pdf"), bbox_inches='tight')
-    # tplt.clean_figure()
-    # tplt.save(os.path.join(prepath, "chapter3/example_gt.tex"))
 plt.show(block=False)
 
 
@@ -213,56 +209,18 @@ n1 = 500
 seed = 21
 eps = .75
 romc.solve_problems(n1=n1, seed=seed, use_bo=False, optimizer_args=None)
-# if savefig:
-#     romc.distance_hist(bins=100,
-#                        savefig=os.path.join(prepath, "chapter3/example_theta_dist.png"))
-# else:
-#     romc.distance_hist(bins=100)
 romc.estimate_regions(eps_filter=eps, use_surrogate=False,
                       region_args=None, fit_models=False)
-# if savefig:
-#     romc.visualize_region(0,
-#                          savefig=os.path.join(prepath, "chapter3/example_region.png"))
-# else:
-#     romc.visualize_region(0)
 
-# sampling
+# sampling part
 n2 = 50
 romc.sample(n2=n2)
-# if savefig:
-#     romc.visualize_region(i=0, savefig=os.path.join(
-#         prepath, "chapter3/example_region_samples.png"))
-# else:
-#     romc.visualize_region(i=0)
-
-# print summary statistics
-romc.result.summary()
 
 # compute expectation
 print("Expected value   : %.3f" %
       romc.compute_expectation(h=lambda x: np.squeeze(x)))
 print("Expected variance: %.3f" %
       romc.compute_expectation(h=lambda x: np.squeeze(x)**2))
-
-# ############# PLOT HISTOGRAM OF SAMPLES  ##################
-# plt.figure()
-# plt.title("Histogram of the samples drawn")
-# plt.hist(x=romc.result.samples_array,
-#          weights=romc.result.weights,
[…]

### 2022-12-30 replication_script/test.py

@@ -1,38 +1,0 @@
-import numpy as np
-import elfi
-
-# Set seed for reproducibility
-seed = 1
-np.random.seed(seed)
-model = ma2.get_model(seed_obs=seed)
-
-# Hyperparameters
-n1 = 5
-n2 = 200
-bounds = [(-2, 2), (-1.25, 1.25)]
-eps = .01
-
-# ROMC
-romc = elfi.ROMC(model, bounds=bounds, parallelize=False, discrepancy_name="d")
-
-# fitting part
-romc.solve_problems(n1=n1, use_bo=True, seed=seed)
-romc.estimate_regions(eps_filter=eps, fit_models=False)
-
-# sampling part
-tmp = romc.sample(n2=n2, seed=seed)
-romc.visualize_region(vis_ind_1)
-
-# print summary statistics
-print(romc.result.summary())
-print(romc.result.samples_cov())
-
-
-# BOLFI
-bolfi = elfi.BOLFI(model, target_name = "d",
-                   batch_size=1,
-                   initial_evidence=20,
-                   update_interval=10,
-                   bounds={'t1':(-2, 2), 't2':(-1.25, 1.25)},
-                   acq_noise_var=0.1,
-                   seed=seed)

### 2022-12-31 article/article.tex
finished paper

@@
   inference})~\citet{1708.00707}. The implementation has been designed
-to ensure extensibility. ROMC is an LFI framework; it defines a
+to facilitate extensibility. ROMC is an LFI framework; it defines a
 sequence of algorithmic steps for approximating the posterior without
-enforcing a specific algorithm for solving each step. Therefore, a
-researcher may use ROMC as the backbone algorithm and develop novel
-methods to solve each separate step. For being a ready-to-use
-algorithm,~\citet{Ikonomov2019} proposed a default method for each
-step, but this choice is by no means restrictive. We have designed our
-software for facilitating such experimentation.
+enforcing a specific algorithm for each step. Therefore, a researcher
+may use ROMC as the backbone method and apply novel algorithms to each
+separate step. For being a ready-to-use LFI
+method,~\citet{Ikonomov2019} propose a particular (default) algorithm
+for each step, but this choice is by no means restrictive. We have
+designed our software for facilitating such experimentation.
 
@@
 artificial example also serves as a step-by-step guide for showcasing
-how to use the various functionalities of our implementation. Second,
-we use the second-order moving average (MA2) example from the
+the various functionalities of our implementation. Second, we use the
+second-order moving average (MA2) example~\citet{Marin2012} from the
 \pkg{ELFI} package, using as ground truth the samples obtained with
@@
 samples. Finally, we present the execution times of ROMC, measuring
-the speed-up achieved by the parallel version of the
-implementation.
-
-The code of the implementation is available at the official \pkg{ELFI} \href{https://github.com/elfi-dev/elfi}{repository}. Apart
-from the examples presented in the paper, there are five \pkg{Google
-  Colab} notebooks available online, with end-to-end examples
-illustrating:
-(i) \href{https://colab.research.google.com/drive/1lGRp0XrNfZ64NN0ASB_tYEKowXwlveDC?usp=sharing}{ROMC on a synthetic \(1D\) example},
-(ii) \href{https://colab.research.google.com/drive/1Fof_WmCi1YizzSI_63aEsbLXsno5gSZ3?usp=sharing}{ROMC on a synthetic \(2D\) example},
-(iii) \href{https://colab.research.google.com/drive/1nkdACQ370SSc0KB1bHv4sBRaxMlMqoNH?usp=sharing}{ROMC on the Moving Average example},
-(iv) \href{https://colab.research.google.com/drive/1_jHVxPSH3XcNOORZJpLU0SPzs0PF8CQ5?usp=sharing}{how to extend ROMC with a Neural Network as a surrogate model},
-(v) \href{https://colab.research.google.com/drive/1RzB-V1QueP1y1nyzv_VOqR1nVz3DUH3v?usp=sharing}{how to extend ROMC with a custom proposal region module}.
+the speed-up achieved by the parallel version of the implementation.
+
+The code of the implementation is available at the official \pkg{ELFI}
+\href{https://github.com/elfi-dev/elfi}{repository}. Apart from the
+examples presented in the paper, there are five \pkg{Google
+  Colab}~\citet{Bisong2019} notebooks available online, with
+end-to-end examples illustrating how to: (i)
+\href{https://colab.research.google.com/drive/1lGRp0XrNfZ64NN0ASB_tYEKowXwlveDC?usp=sharing}{use ROMC
+  on a synthetic \(1D\) example}, (ii)
+\href{https://colab.research.google.com/drive/1Fof_WmCi1YizzSI_63aEsbLXsno5gSZ3?usp=sharing}{use ROMC
+  on a synthetic \(2D\) example}, (iii)
+\href{https://colab.research.google.com/drive/1nkdACQ370SSc0KB1bHv4sBRaxMlMqoNH?usp=sharing}{use ROMC
+  on the Moving Average example}, (iv)
+\href{https://colab.research.google.com/drive/1_jHVxPSH3XcNOORZJpLU0SPzs0PF8CQ5?usp=sharing}{extend ROMC with a Neural Network as a surrogate model}, (v)
+\href{https://colab.research.google.com/drive/1RzB-V1QueP1y1nyzv_VOqR1nVz3DUH3v?usp=sharing}{extend ROMC with a custom proposal region module}.
 
@@
 \(\ub_i \sim p(\ub)\) for defining the deterministic distance
-functions \(d_i(\thetab)\) (Steps 3-5), (b) solve the optimization
+functions \(d_i(\thetab)\) (steps 3-5), (b) solve the optimization
 problems for obtaining \(\thetab_i^*, d_i^*\) and keep the solutions
-inside the threshold \(\epsilon\) (Steps 6-9), and (c) estimate the
+inside the threshold \(\epsilon\) (steps 6-9), and (c) estimate the
 bounding boxes \(\accregionihat\) to define uniform distributions
-\(q_i\) on them (Step 10). Optionally, a surrogate model
-\(\tilde{d}_i\) can be fitted for a faster inference phase (Step 11).
+\(q_i\) on them (step 10). Optionally, a surrogate model
+\(\tilde{d}_i\) can be fitted for a faster inference phase (step 11).
 
@@
 by the eigenvalues of the Hessian matrix \(\hessian_i\) of \(d_i\) at
-\(\thetab_i\) (Step 1). Depending on the algorithm used in the
+\(\thetab_i\) (step 1). Depending on the algorithm used in the
 optimization step, we either use the real distance \(d_i\) or the
@@
 derivatives. After defining the axes, we search for the bounding box
-limits with a line step algorithm (Steps 2-21). The key idea is to
+limits with a line step algorithm (steps 2-21). The key idea is to
 take long steps \(\eta\_\text{start}\) until crossing the boundary and
@@
 
-The results in this paper were obtained using \proglang{Python~3.7.9}
-with the \pkg{ELFI~0.8.3} package. The experiments have been executed
+The results in this paper were obtained using \proglang{Python~3.9}
+with the \pkg{ELFI~0.8.4} package. The experiments have been executed
 in a single machine with an \proglang{Intel® Core™ i7-8750H} Processor

### 2022-12-31 article/article.tex
final
submitted version

@@
 \hline
-ROMC (Bayesian optimization) & 0.501 & 0.17 & 0.087 & 0.175 \\
+ROMC (Bayesian optimization) & 0.494 & 0.16 & 0.086 & 0.167 \\
 \hline
@@
 
-The results in this paper were obtained using \proglang{Python~3.9}
-with the \pkg{ELFI~0.8.4} package. The experiments have been executed
-in a single machine with an \proglang{Intel® Core™ i7-8750H} Processor
-an with \proglang{Ubuntu 20.04} lts operating system.
+The results in this paper were obtained using \proglang{Python~3.9},
+\pkg{ELFI~0.8.4} at \proglang{Ubuntu 20.04 lts} operating system, at a
+single machine with \proglang{Intel® Core™ i7-8750H} processor.
 

### 2023-09-05 article/article.tex
after revision 3

@@
 bounds = [(-2.5, 2.5)] # bounds of the prior
-parallelize = True # activate parallel execution
+parallelize = False # True activates parallel execution
 romc = elfi.ROMC(d, bounds = bounds, parallelize = parallelize)
@@
 
-romc.visualize_region(i = 1) # for inspecting visually the bounding box
+romc.visualize_region(i = 2) # for inspecting visually the bounding box
 
@@
 romc.result.summary()
-# Number of samples  : 19300
-# Sample means: theta: -0.0116
+# Method: ROMC
+# Number of samples: 19300
+# Parameter                Mean               2.5%              97.5%
+# theta:                 -0.012             -1.985              1.987
 
@@
 # eval unnorm posterior
-romc.eval_unnorm_posterior(theta = 0)
+print("%.3f" % romc.eval_unnorm_posterior(theta = np.array([[0]])))
+# 58.800
 
 # check eval posterior
-romc.eval_posterior(theta = 0)
+print("%.3f" % romc.eval_posterior(theta = np.array([[0]])))
+# 0.289
 \end{Code}
@@
 \hline \hline
-Rejection ABC & 0.516 & 0.142 & 0.07 & 0.172 \\
+Rejection ABC & 0.516 & 0.142 & 0.070 & 0.172 \\
 \hline
@@
 \hline
-ROMC (Bayesian optimization) & 0.494 & 0.16 & 0.086 & 0.167 \\
+ROMC (Bayesian optimization) & 0.513 & 0.169 & 0.090 & 0.174 \\
 \hline
-ROMC (Neural Network) & 0.491 & 0.138 & 0.04 & 0.172 \\
+ROMC (Neural Network) & 0.491 & 0.138 & 0.040 & 0.172 \\
 \hline

### 2023-09-05 replication_script/replication_script.py

@@ -204,23 +204,57 @@ dim = data.shape[-1]
 # define romc inference
 romc = elfi.ROMC(dist, bounds)
 
-# fitting part
+# Example - Training part
+# Training (fitting) part
 n1 = 500
 seed = 21
+use_bo = False
+
+# Training step-by-step
+romc.solve_problems(n1=n1, seed=seed, use_bo=use_bo, optimizer_args=None)
+
 eps = .75
-romc.solve_problems(n1=n1, seed=seed, use_bo=False, optimizer_args=None)
 romc.estimate_regions(eps_filter=eps, use_surrogate=False,
                       region_args=None, fit_models=False)
 
-# sampling part
+romc.visualize_region(i=2)
+
+# Inference part
+seed = 21
 n2 = 50
-romc.sample(n2=n2)
+romc.sample(n2=n2, seed=seed)
+
+# visualize_region, adding the samples now
+romc.visualize_region(i=2)
+
+# visualize marginal (built-in ELFI tool)
+weights = romc.result.weights
+romc.result.plot_marginals(weights=weights, bins=100, range=(-3, 3))
+plt.show(block=False)
+
+# summarize the samples (built-in ElFI tool)
+romc.result.summary()
+# Method: ROMC
+# Number of samples: 19300
+# Parameter                Mean               2.5%              97.5%
+# theta:                 -0.012             -1.985              1.987
 
 # compute expectation
-print("Expected value   : %.3f" %
-      romc.compute_expectation(h=lambda x: np.squeeze(x)))
-print("Expected variance: %.3f" %
-      romc.compute_expectation(h=lambda x: np.squeeze(x)**2))
+exp_eval = romc.compute_expectation(h=lambda x: np.squeeze(x))
+print("Expected value   : %.3f" % exp_eval)
+# Expected value: -0.012
+
+exp_var = romc.compute_expectation(h=lambda x: np.squeeze(x)**2)
+print("Expected variance: %.3f" % exp_var)
+# Expected variance: 1.120
+
+# eval unnorm posterior
+print("%.3f" % romc.eval_unnorm_posterior(theta = np.array([[0]])))
+# 58.800
+
+# check eval posterior
+print("%.3f" % romc.eval_posterior(theta = np.array([[0]])))
+# 0.289
 
 # plot histogram of samples along with the ground-truth prior, likelihood, posterior
 plt.figure()
@@ -262,7 +296,7 @@ plt.show(block=False)
 
 
[…]

### 2023-09-06 replication_script/replication_script.py
added Henri's comments

@@ -16,7 +16,7 @@ import elfi
 import matplotlib
 
 # matplotlib.rcParams['text.usetex'] = True
-logging.basicConfig(level=logging.INFO)
+logging.basicConfig(level=logging.CRITICAL)
 savefig = False
 
 ######################################################################
@@ -204,6 +204,9 @@ dim = data.shape[-1]
 # define romc inference
 romc = elfi.ROMC(dist, bounds)
 
+print("\n######################################################################")
+print("PART 1 - Generates all and outputs until page 19 (running example of the paper)\n")
+
 # Example - Training part
 # Training (fitting) part
 n1 = 500
@@ -316,10 +319,16 @@ ess = romc.compute_ess()
 print("Nof Samples: %d, ESS: %.3f" % (nof_samples, ess))
 # Nof Samples: 19300, ESS: 16196.214
 
+print("######################################################################\n")
+
 ######################################################################
 # Part 2. Generates the subfigures used in the gradient-based        #
 # optimisation version of the example at chapter 5, i.e. figures 5, 6#
 ######################################################################
+
+print("\n##############################################################")
+print("PARTS 2,3,4 - Generate all figures and outputs for Section 5\n")
+
 
 # set the path for storing images
 prepath = './images/chapter4/'
@@ -380,6 +389,9 @@ n2 = 200
 bounds = [(-2, 2), (-1.25, 1.25)]
 eps = .01
 vis_ind_1 = 1
+
+print("\n######################################################################")
+print("PART 2 - Fit the gradient-based ROMC for Section 5\n")
 
 # define model
 romc = elfi.ROMC(model, bounds=bounds, discrepancy_name="d")
@@ -455,11 +467,16 @@ romc_gb_mu_th1, romc_gb_mu_th2 = romc.result.sample_means_array
 romc_gb_sigma_th1 = np.sqrt(romc.result.samples_cov()[0, 0])
 romc_gb_sigma_th2 = np.sqrt(romc.result.samples_cov()[1, 1])
 
+print("######################################################################\n")
 
 ######################################################################
 # Part 3. Generates the subfigures used in the Bayesian optimization #
 # version of the example at chapter 5, i.e. figures 5, 6             #
 ######################################################################
+
[…]

### 2023-09-07 replication_script/replication_script.py
added docker execution

@@ -14,6 +14,9 @@ from elfi.examples import ma2
 import numpy as np
 import elfi
 import matplotlib
+import random
+
+random.seed(21)
 
 # matplotlib.rcParams['text.usetex'] = True
 logging.basicConfig(level=logging.CRITICAL)

### 2023-09-19 arXiv/article.tex
added submission to arXiv

--- 2023-09-05 article/article.tex
@@
-\documentclass[article, shortnames]{jss}
+\documentclass[nojss]{jss}
 %% -- LaTeX packages and custom commands ---------------------------------------
-
+\pdfoutput=1
 \usepackage[usenames,dvipsnames]{xcolor}
@@
     \begin{center}
-      \includegraphics[width=0.8\textwidth]{./latex_files/images/chapter2/elfi.png}
+      \includegraphics[width=0.8\textwidth]{./images/chapter2/elfi.png}
     \end{center}
@@
 
-\input{latex_files/graphs/romc_graph}
+\input{graphs/romc_graph}
 
@@
   \begin{center}
-    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_gt.pdf}
-    \includegraphics[width=0.49\textwidth]{latex_files/images/chapter3/example_posterior.pdf}
-    % \input{latex_files/images/chapter3/example_gt.tex}
-    % \input{latex_files/images/chapter3/example_post.tex}
+    \includegraphics[width=0.49\textwidth]{images/chapter3/example_gt.pdf}
+    \includegraphics[width=0.49\textwidth]{images/chapter3/example_posterior.pdf}
+    % \input{images/chapter3/example_gt.tex}
+    % \input{images/chapter3/example_post.tex}
   \end{center}
@@
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1.pdf}
-        \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/ma2_region_1_bo.pdf}
+        \includegraphics[width=0.49\textwidth]{./images/chapter4/ma2_region_1.pdf}
+        \includegraphics[width=0.49\textwidth]{./images/chapter4/ma2_region_1_bo.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_rejection.pdf}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc.pdf}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_bo.pdf}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_rejection.pdf}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc.pdf}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_bo.pdf}
-    \includegraphics[width=0.24\textwidth]{./latex_files/images/chapter4/mae2_hist_t2_romc_nn.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_rejection.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc_bo.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_rejection.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc_bo.pdf}
+    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc_nn.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior.pdf}
-    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_bo.pdf}
-    \includegraphics[width=0.32\textwidth]{./latex_files/images/chapter4/mae2_romc_posterior_nn.pdf}
+    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior.pdf}
+    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior_bo.pdf}
+    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior_nn.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_solve.pdf}
-    \includegraphics[width=0.49\textwidth]{./latex_files/images/chapter4/exec_time_regions.pdf}
+    \includegraphics[width=0.49\textwidth]{./images/chapter4/exec_time_solve.pdf}
+    \includegraphics[width=0.49\textwidth]{./images/chapter4/exec_time_regions.pdf}
     \end{center}
@@
 The results in this paper were obtained using \proglang{Python~3.9},
-\pkg{ELFI~0.8.4} at \proglang{Ubuntu 20.04 lts} operating system, at a
+\pkg{ELFI~0.8.6} at \proglang{Ubuntu 20.04 lts} operating system, at a
 single machine with \proglang{Intel® Core™ i7-8750H} processor.

### 2023-09-19 arXiv/graphs/romc_graph.tex
added submission to arXiv

--- 2021-10-05 latex_files/graphs/romc_graph.tex
@@
   \end{center}
-  \caption{Overview of the
-ROMC implementation. The training part follows a sequential pattern;
-the functions in the green ellipses must be called in a sequential
-fashion for completing the training part and define the posterior
-distribution. The functions in blue ellipses are the functionalities
-provided to the user.}
+  \caption{Overview of the ROMC implementation. On the left side, we
+    depict ROMC as a sequence of algotirhmic steps. On the right side,
+    we present the functions that form our implementation; the green
+    rectangles (starting with underscore) are the internal
+    functionalities and the blue rectangles the publicly exposed
+    API. This side-by-side illustration highlights that our
+    implementation follows strictly the algorithmic view of ROMC.}
 \label{fig:romc_overview}

### 2023-09-19 jss-submission/images/chapter4/mae2_hist_t2_romc_bo.tex
added submission to arXiv

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (BO) - \(\displaystyle \theta_2\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_2\)},
xmin=-0.575, xmax=1.075,
xtick style={color=black},
xtick={-0.6,-0.4,-0.2,0,0.2,0.4,0.6,0.8,1,1.2},
xticklabels={
  \(\displaystyle {\ensuremath{-}0.6}\),
  \(\displaystyle {\ensuremath{-}0.4}\),
  \(\displaystyle {\ensuremath{-}0.2}\),
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\)
}
]
\draw[draw=none,fill=color0] (axis cs:-0.5,0) rectangle (axis cs:-0.475,0);
\draw[draw=none,fill=color0] (axis cs:-0.475,0) rectangle (axis cs:-0.45,0);
\draw[draw=none,fill=color0] (axis cs:-0.45,0) rectangle (axis cs:-0.425,0);
\draw[draw=none,fill=color0] (axis cs:-0.425,0) rectangle (axis cs:-0.4,0);
\draw[draw=none,fill=color0] (axis cs:-0.4,0) rectangle (axis cs:-0.375,0);
\draw[draw=none,fill=color0] (axis cs:-0.375,0) rectangle (axis cs:-0.35,0.00876430825090703);
\draw[draw=none,fill=color0] (axis cs:-0.35,0) rectangle (axis cs:-0.325,0.0553849594476448);
\draw[draw=none,fill=color0] (axis cs:-0.325,0) rectangle (axis cs:-0.3,0.0995221501181132);
\draw[draw=none,fill=color0] (axis cs:-0.3,0) rectangle (axis cs:-0.275,0.200116932597969);
\draw[draw=none,fill=color0] (axis cs:-0.275,0) rectangle (axis cs:-0.25,0.280170997234605);
\draw[draw=none,fill=color0] (axis cs:-0.25,0) rectangle (axis cs:-0.225,0.308892055918909);
\draw[draw=none,fill=color0] (axis cs:-0.225,0) rectangle (axis cs:-0.2,0.480795303488478);
\draw[draw=none,fill=color0] (axis cs:-0.2,0) rectangle (axis cs:-0.175,0.663720536159796);
\draw[draw=none,fill=color0] (axis cs:-0.175,0) rectangle (axis cs:-0.15,1.00951495102481);
\draw[draw=none,fill=color0] (axis cs:-0.15,0) rectangle (axis cs:-0.125,1.06099778651247);
\draw[draw=none,fill=color0] (axis cs:-0.125,0) rectangle (axis cs:-0.1,1.34254475295888);
\draw[draw=none,fill=color0] (axis cs:-0.1,0) rectangle (axis cs:-0.075,1.33497188533518);
\draw[draw=none,fill=color0] (axis cs:-0.075,0) rectangle (axis cs:-0.05,1.58789149652067);
\draw[draw=none,fill=color0] (axis cs:-0.05,0) rectangle (axis cs:-0.025,1.74539319951242);
\draw[draw=none,fill=color0] (axis cs:-0.025,0) rectangle (axis cs:0,2.04991013705781);
\draw[draw=none,fill=color0] (axis cs:0,0) rectangle (axis cs:0.025,2.2241094383069);
\draw[draw=none,fill=color0] (axis cs:0.025,0) rectangle (axis cs:0.05,2.46101072110277);
\draw[draw=none,fill=color0] (axis cs:0.05,0) rectangle (axis cs:0.0750000000000001,2.34426632949851);
\draw[draw=none,fill=color0] (axis cs:0.0750000000000001,0) rectangle (axis cs:0.1,2.64989491584813);
\draw[draw=none,fill=color0] (axis cs:0.1,0) rectangle (axis cs:0.125,2.22087507116604);
\draw[draw=none,fill=color0] (axis cs:0.125,0) rectangle (axis cs:0.15,2.40968203254366);
\draw[draw=none,fill=color0] (axis cs:0.15,0) rectangle (axis cs:0.175,2.13585232653887);
\draw[draw=none,fill=color0] (axis cs:0.175,0) rectangle (axis cs:0.2,1.87390968331577);
\draw[draw=none,fill=color0] (axis cs:0.2,0) rectangle (axis cs:0.225,1.73667558306975);
\draw[draw=none,fill=color0] (axis cs:0.225,0) rectangle (axis cs:0.25,1.42275664230349);
\draw[draw=none,fill=color0] (axis cs:0.25,0) rectangle (axis cs:0.275,1.33852589947831);
\draw[draw=none,fill=color0] (axis cs:0.275,0) rectangle (axis cs:0.3,0.966336541457652);
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.325,0.853157677045122);
\draw[draw=none,fill=color0] (axis cs:0.325,0) rectangle (axis cs:0.35,0.646826576462357);
\draw[draw=none,fill=color0] (axis cs:0.35,0) rectangle (axis cs:0.375,0.490570041657969);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.4,0.417168760177305);
\draw[draw=none,fill=color0] (axis cs:0.4,0) rectangle (axis cs:0.425,0.417508714918072);
\draw[draw=none,fill=color0] (axis cs:0.425,0) rectangle (axis cs:0.45,0.264031941738089);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.475,0.176949480734724);
\draw[draw=none,fill=color0] (axis cs:0.475,0) rectangle (axis cs:0.5,0.180814986804008);
\draw[draw=none,fill=color0] (axis cs:0.5,0) rectangle (axis cs:0.525,0.189172180863948);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.55,0.11666427997438);
\draw[draw=none,fill=color0] (axis cs:0.55,0) rectangle (axis cs:0.575,0.0907995766684012);
\draw[draw=none,fill=color0] (axis cs:0.575,0) rectangle (axis cs:0.6,0.0626949457948479);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.625,0.0202885500980677);
\draw[draw=none,fill=color0] (axis cs:0.625,0) rectangle (axis cs:0.65,0.0304328251471012);
\draw[draw=none,fill=color0] (axis cs:0.65,0) rectangle (axis cs:0.675,0.0202885500980677);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.7,0.0101442750490337);
\draw[draw=none,fill=color0] (axis cs:0.7,0) rectangle (axis cs:0.725,0);
\draw[draw=none,fill=color0] (axis cs:0.725,0) rectangle (axis cs:0.75,0);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.775,0);
\draw[draw=none,fill=color0] (axis cs:0.775,0) rectangle (axis cs:0.8,0);
\draw[draw=none,fill=color0] (axis cs:0.8,0) rectangle (axis cs:0.825,0);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.85,0);
\draw[draw=none,fill=color0] (axis cs:0.85,0) rectangle (axis cs:0.875,0);
\draw[draw=none,fill=color0] (axis cs:0.875,0) rectangle (axis cs:0.9,0);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.925,0);
\draw[draw=none,fill=color0] (axis cs:0.925,0) rectangle (axis cs:0.95,0);
\draw[draw=none,fill=color0] (axis cs:0.95,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:1,0);
\addplot [semithick, red, dashed]
table {%
0.0862621123909427 0
0.0862621123909427 3
};
\addlegendentry{$\mu = 0.086$}
\addplot [semithick, black, dashed]
table {%
-0.113758196522132 1
0.253534843782206 1
};
\addlegendentry{$\sigma = 0.167$}
\end{axis}

\end{tikzpicture}

### 2023-09-19 replication_script/replication_script.py

@@ -18,833 +18,835 @@ import random
 
 random.seed(21)
 
-# matplotlib.rcParams['text.usetex'] = True
-logging.basicConfig(level=logging.CRITICAL)
-savefig = False
+if __name__ == '__main__':
 
-######################################################################
-# Part 1. Generates the subfigures used in the running example in the#
-# presentation of our implementation i.e. figure 3                   #
-######################################################################
+    # matplotlib.rcParams['text.usetex'] = True
+    logging.basicConfig(level=logging.CRITICAL)
+    savefig = False
 
-class Prior:
-    def rvs(self, size=None, random_state=None):
-        # size from (BS,) -> (BS,1)
-        if size is not None:
-            size = np.concatenate((size, [1]))
-        return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+    ######################################################################
+    # Part 1. Generates the subfigures used in the running example in the#
+    # presentation of our implementation i.e. figure 3                   #
+    ######################################################################
 
-    def pdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+    class Prior:
+        def rvs(self, size=None, random_state=None):
+            # size from (BS,) -> (BS,1)
+            if size is not None:
+                size = np.concatenate((size, [1]))
+            return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
 
-    def logpdf(self, theta):
-        return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+        def pdf(self, theta):
+            return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+        def logpdf(self, theta):
+            return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
 
 
-class Likelihood:
-    r"""Implements the distribution
-    P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
-                 N(x; theta + 0.5 + 0.5^4 if theta > 0.5
-                 N(x; theta - 0.5 + 0.5^4 if theta < 0.5
-    """
-
[…]

### 2023-09-27 jss-camera-ready/Figures/chapter4/mae2_hist_t1_romc.tex
added camera ready

% This file was created with tikzplotlib v0.9.12.
\begin{tikzpicture}

\definecolor{color0}{rgb}{0.12156862745098,0.466666666666667,0.705882352941177}

\begin{axis}[
legend cell align={left},
legend style={fill opacity=0.8, draw opacity=1, text opacity=1, draw=white!80!black},
tick align=outside,
tick pos=left,
title={ROMC (gradient-based) - \(\displaystyle \theta_1\)},
x grid style={white!69.0196078431373!black},
xlabel={\(\displaystyle \theta_1\)},
xmin=0.255, xmax=1.245,
xtick style={color=black},
xtick={0.2,0.4,0.6,0.8,1,1.2,1.4},
xticklabels={
  \(\displaystyle {0.2}\),
  \(\displaystyle {0.4}\),
  \(\displaystyle {0.6}\),
  \(\displaystyle {0.8}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.2}\),
  \(\displaystyle {1.4}\)
},
y grid style={white!69.0196078431373!black},
ylabel={density},
ymin=0, ymax=3.5,
ytick style={color=black},
ytick={0,0.5,1,1.5,2,2.5,3,3.5},
yticklabels={
  \(\displaystyle {0.0}\),
  \(\displaystyle {0.5}\),
  \(\displaystyle {1.0}\),
  \(\displaystyle {1.5}\),
  \(\displaystyle {2.0}\),
  \(\displaystyle {2.5}\),
  \(\displaystyle {3.0}\),
  \(\displaystyle {3.5}\)
}
]
\draw[draw=none,fill=color0] (axis cs:0.3,0) rectangle (axis cs:0.315,1.31803081159554);
\draw[draw=none,fill=color0] (axis cs:0.315,0) rectangle (axis cs:0.33,1.43612870647347);
\draw[draw=none,fill=color0] (axis cs:0.33,0) rectangle (axis cs:0.345,1.5890027481777);
\draw[draw=none,fill=color0] (axis cs:0.345,0) rectangle (axis cs:0.36,1.70239598718898);
\draw[draw=none,fill=color0] (axis cs:0.36,0) rectangle (axis cs:0.375,1.99625549422168);
\draw[draw=none,fill=color0] (axis cs:0.375,0) rectangle (axis cs:0.39,2.34342226707839);
\draw[draw=none,fill=color0] (axis cs:0.39,0) rectangle (axis cs:0.405,2.7730601684882);
\draw[draw=none,fill=color0] (axis cs:0.405,0) rectangle (axis cs:0.42,2.88578570810527);
\draw[draw=none,fill=color0] (axis cs:0.42,0) rectangle (axis cs:0.435,3.1094248119564);
\draw[draw=none,fill=color0] (axis cs:0.435,0) rectangle (axis cs:0.45,3.41920821129824);
\draw[draw=none,fill=color0] (axis cs:0.45,0) rectangle (axis cs:0.465,3.54938990669914);
\draw[draw=none,fill=color0] (axis cs:0.465,0) rectangle (axis cs:0.48,3.44932850829471);
\draw[draw=none,fill=color0] (axis cs:0.48,0) rectangle (axis cs:0.495,3.56489218047191);
\draw[draw=none,fill=color0] (axis cs:0.495,0) rectangle (axis cs:0.51,3.19918772646129);
\draw[draw=none,fill=color0] (axis cs:0.51,0) rectangle (axis cs:0.525,2.84463114544444);
\draw[draw=none,fill=color0] (axis cs:0.525,0) rectangle (axis cs:0.54,2.48396273618897);
\draw[draw=none,fill=color0] (axis cs:0.54,0) rectangle (axis cs:0.555,2.36620853424246);
\draw[draw=none,fill=color0] (axis cs:0.555,0) rectangle (axis cs:0.57,2.15648120155311);
\draw[draw=none,fill=color0] (axis cs:0.57,0) rectangle (axis cs:0.585,2.03237770334023);
\draw[draw=none,fill=color0] (axis cs:0.585,0) rectangle (axis cs:0.6,1.7316295728278);
\draw[draw=none,fill=color0] (axis cs:0.6,0) rectangle (axis cs:0.615,1.76467905230175);
\draw[draw=none,fill=color0] (axis cs:0.615,0) rectangle (axis cs:0.63,1.63196846595208);
\draw[draw=none,fill=color0] (axis cs:0.63,0) rectangle (axis cs:0.645,1.59520767700748);
\draw[draw=none,fill=color0] (axis cs:0.645,0) rectangle (axis cs:0.66,1.51973180697253);
\draw[draw=none,fill=color0] (axis cs:0.66,0) rectangle (axis cs:0.675,1.31194646238609);
\draw[draw=none,fill=color0] (axis cs:0.675,0) rectangle (axis cs:0.69,1.2211118657454);
\draw[draw=none,fill=color0] (axis cs:0.69,0) rectangle (axis cs:0.705,1.12757407088164);
\draw[draw=none,fill=color0] (axis cs:0.705,0) rectangle (axis cs:0.72,0.960594674407477);
\draw[draw=none,fill=color0] (axis cs:0.72,0) rectangle (axis cs:0.735,0.940482485898332);
\draw[draw=none,fill=color0] (axis cs:0.735,0) rectangle (axis cs:0.75,0.688125745940285);
\draw[draw=none,fill=color0] (axis cs:0.75,0) rectangle (axis cs:0.765,0.663702220802231);
\draw[draw=none,fill=color0] (axis cs:0.765,0) rectangle (axis cs:0.78,0.579520830287313);
\draw[draw=none,fill=color0] (axis cs:0.78,0) rectangle (axis cs:0.795,0.556626697680565);
\draw[draw=none,fill=color0] (axis cs:0.795,0) rectangle (axis cs:0.81,0.48821785974726);
\draw[draw=none,fill=color0] (axis cs:0.81,0) rectangle (axis cs:0.825,0.344895036404317);
\draw[draw=none,fill=color0] (axis cs:0.825,0) rectangle (axis cs:0.84,0.304533264240393);
\draw[draw=none,fill=color0] (axis cs:0.84,0) rectangle (axis cs:0.855,0.196650595905097);
\draw[draw=none,fill=color0] (axis cs:0.855,0) rectangle (axis cs:0.87,0.138447961549637);
\draw[draw=none,fill=color0] (axis cs:0.87,0) rectangle (axis cs:0.885,0.157880556756463);
\draw[draw=none,fill=color0] (axis cs:0.885,0) rectangle (axis cs:0.9,0.198075404071845);
\draw[draw=none,fill=color0] (axis cs:0.9,0) rectangle (axis cs:0.915,0.13671760298444);
\draw[draw=none,fill=color0] (axis cs:0.915,0) rectangle (axis cs:0.93,0.100312400879399);
\draw[draw=none,fill=color0] (axis cs:0.93,0) rectangle (axis cs:0.945,0.0415876649706079);
\draw[draw=none,fill=color0] (axis cs:0.945,0) rectangle (axis cs:0.96,0.0472721327861164);
\draw[draw=none,fill=color0] (axis cs:0.96,0) rectangle (axis cs:0.975,0);
\draw[draw=none,fill=color0] (axis cs:0.975,0) rectangle (axis cs:0.99,0);
\draw[draw=none,fill=color0] (axis cs:0.99,0) rectangle (axis cs:1.005,0);
\draw[draw=none,fill=color0] (axis cs:1.005,0) rectangle (axis cs:1.02,0);
\draw[draw=none,fill=color0] (axis cs:1.02,0) rectangle (axis cs:1.035,0);
\draw[draw=none,fill=color0] (axis cs:1.035,0) rectangle (axis cs:1.05,0);
\draw[draw=none,fill=color0] (axis cs:1.05,0) rectangle (axis cs:1.065,0);
\draw[draw=none,fill=color0] (axis cs:1.065,0) rectangle (axis cs:1.08,0);
\draw[draw=none,fill=color0] (axis cs:1.08,0) rectangle (axis cs:1.095,0);
\draw[draw=none,fill=color0] (axis cs:1.095,0) rectangle (axis cs:1.11,0);
\draw[draw=none,fill=color0] (axis cs:1.11,0) rectangle (axis cs:1.125,0);
\draw[draw=none,fill=color0] (axis cs:1.125,0) rectangle (axis cs:1.14,0);
\draw[draw=none,fill=color0] (axis cs:1.14,0) rectangle (axis cs:1.155,0);
\draw[draw=none,fill=color0] (axis cs:1.155,0) rectangle (axis cs:1.17,0);
\draw[draw=none,fill=color0] (axis cs:1.17,0) rectangle (axis cs:1.185,0);
\draw[draw=none,fill=color0] (axis cs:1.185,0) rectangle (axis cs:1.2,0);
\addplot [semithick, red, dashed]
table {%
0.501096405407453 0
0.501096405407453 3.5
};
\addlegendentry{$\mu = 0.501$}
\addplot [semithick, black, dashed]
table {%
0.320058196552687 1
0.632353895343709 1
};
\addlegendentry{$\sigma = 0.142$}
\end{axis}

\end{tikzpicture}

### 2023-09-27 jss-camera-ready/jss4678.tex
added camera ready

--- 2023-09-19 arXiv/article.tex
@@
-\documentclass[nojss]{jss}
+\documentclass[article, shortnames]{jss}
 %% -- LaTeX packages and custom commands ---------------------------------------
-\pdfoutput=1
+
 \usepackage[usenames,dvipsnames]{xcolor}
@@
     \begin{center}
-      \includegraphics[width=0.8\textwidth]{./images/chapter2/elfi.png}
+      \includegraphics[width=0.8\textwidth]{./Figures/chapter2/elfi.png}
     \end{center}
@@
 
-\input{graphs/romc_graph}
+\input{Figures/romc_graph}
 
@@
   \begin{center}
-    \includegraphics[width=0.49\textwidth]{images/chapter3/example_gt.pdf}
-    \includegraphics[width=0.49\textwidth]{images/chapter3/example_posterior.pdf}
+    \includegraphics[width=0.49\textwidth]{Figures/chapter3/example_gt.pdf}
+    \includegraphics[width=0.49\textwidth]{Figures/chapter3/example_posterior.pdf}
     % \input{images/chapter3/example_gt.tex}
@@
     \begin{center}
-        \includegraphics[width=0.49\textwidth]{./images/chapter4/ma2_region_1.pdf}
-        \includegraphics[width=0.49\textwidth]{./images/chapter4/ma2_region_1_bo.pdf}
+        \includegraphics[width=0.49\textwidth]{./Figures/chapter4/ma2_region_1.pdf}
+        \includegraphics[width=0.49\textwidth]{./Figures/chapter4/ma2_region_1_bo.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_rejection.pdf}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc.pdf}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc_bo.pdf}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t1_romc_nn.pdf}\\
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_rejection.pdf}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc.pdf}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc_bo.pdf}
-    \includegraphics[width=0.24\textwidth]{./images/chapter4/mae2_hist_t2_romc_nn.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t1_rejection.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t1_romc.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t1_romc_bo.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t1_romc_nn.pdf}\\
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t2_rejection.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t2_romc.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t2_romc_bo.pdf}
+    \includegraphics[width=0.24\textwidth]{./Figures/chapter4/mae2_hist_t2_romc_nn.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior.pdf}
-    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior_bo.pdf}
-    \includegraphics[width=0.32\textwidth]{./images/chapter4/mae2_romc_posterior_nn.pdf}
+    \includegraphics[width=0.32\textwidth]{./Figures/chapter4/mae2_romc_posterior.pdf}
+    \includegraphics[width=0.32\textwidth]{./Figures/chapter4/mae2_romc_posterior_bo.pdf}
+    \includegraphics[width=0.32\textwidth]{./Figures/chapter4/mae2_romc_posterior_nn.pdf}
     \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.49\textwidth]{./images/chapter4/exec_time_solve.pdf}
-    \includegraphics[width=0.49\textwidth]{./images/chapter4/exec_time_regions.pdf}
+    \includegraphics[width=0.49\textwidth]{./Figures/chapter4/exec_time_solve.pdf}
+    \includegraphics[width=0.49\textwidth]{./Figures/chapter4/exec_time_regions.pdf}
     \end{center}
@@
 \clearpage
-\bibliography{refs}
+\bibliography{jss4678}
 

### 2023-09-27 jss-camera-ready/Code/elfi-dev/docs/conf.py

@@ -1,0 +1,307 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+#
+# ELFI documentation build configuration file, created by
+# sphinx-quickstart on Thu Oct 27 09:49:40 2016.
+#
+# This file is execfile()d with the current directory set to its
+# containing dir.
+#
+# Note that not all possible configuration values are present in this
+# autogenerated file.
+#
+# All configuration values have a default; values that are commented out
+# serve to show the default.
+
+import os
+import sys
+# http://docs.readthedocs.io/en/latest/faq.html#i-get-import-errors-on-libraries-that-depend-on-c-modules
+from unittest.mock import MagicMock
+
+class Mock(MagicMock):
+    @classmethod
+    def __getattr__(cls, name):
+        return MagicMock()
+
+on_RTD = os.environ.get('READTHEDOCS', None) == 'True'
+if on_RTD:
+    MOCK_MODULES = [
+        'pygtk', 'gtk', 'gobject', 'argparse', 'numpy', 'pandas', 'scipy', 'unqlite', 'dask',
+        'distributed', 'distributed.client', 'graphviz', 'matplotlib', 'sobol_seq', 'GPy',
+        'GPy.kern', 'GPy.models', 'toolz', 'toolz.functoolz',
+        'dask.delayed', 'scipy.linalg', 'scipy.optimize', 'scipy.stats', 'scipy.spatial',
+        'scipy.sparse', 'scipy.special', 'matplotlib.pyplot', 'numpy.random', 'networkx',
+        'ipyparallel', 'numpy.lib', 'numpy.lib.format', 'sklearn.linear_model',
+        'sklearn.pipeline', 'sklearn.preprocessing', 'numdifftools', 'GPy.kern', 'GPy.models',
+        'sklearn.covariance', 'sklearn.exceptions'
+    ]
+    sys.modules.update((mod_name, Mock()) for mod_name in MOCK_MODULES)
+
+    html_theme = 'sphinx_rtd_theme'
+
+else:
+    html_theme = 'sphinx_rtd_theme'
+
+# https://github.com/rtfd/readthedocs.org/issues/1139
+"""
+def run_apidoc(_):
+    from sphinx.apidoc import main
+    import os
+    import sys
+    sys.path.append(os.path.join(os.path.dirname(__file__), '..'))
+    cur_dir = os.path.abspath(os.path.dirname(__file__))
+    module = '../elfi'
+    output_path = os.path.join(cur_dir, 'source')
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/scripts/MA2_run.py

@@ -1,0 +1,12 @@
+import elfi
+from elfi.examples import ma2
+
+# load the model from elfi.examples
+model = ma2.get_model()
+
+# setup and run rejection sampling
+rej = elfi.Rejection(model['d'], batch_size=10000)
+result = rej.sample(1000, quantile=0.01)
+
+# show summary of results on stdout
+result.summary()

### 2023-09-27 jss-camera-ready/Code/elfi-dev/setup.py

@@ -1,0 +1,41 @@
+import os
+from io import open
+
+from setuptools import find_packages, setup
+
+packages = ['elfi'] + ['elfi.' + p for p in find_packages('elfi')]
+
+# include C++ examples
+package_data = {'elfi.examples': ['cpp/Makefile', 'cpp/*.txt', 'cpp/*.cpp']}
+
+with open('requirements.txt', 'r') as f:
+    requirements = f.read().splitlines()
+
+optionals = {'doc': ['Sphinx'], 'graphviz': ['graphviz>=0.7.1']}
+
+# read version number
+__version__ = open('elfi/__init__.py').readlines()[-1].split(' ')[-1].strip().strip("'\"")
+
+setup(
+    name='elfi',
+    keywords='abc likelihood-free statistics',
+    packages=packages,
+    package_data=package_data,
+    version=__version__,
+    author='ELFI authors',
+    author_email='elfi-support@hiit.fi',
+    url='http://elfi.readthedocs.io',
+    install_requires=requirements,
+    extras_require=optionals,
+    description='ELFI - Engine for Likelihood-free Inference',
+    long_description=(open('docs/description.rst').read()),
+    license='BSD',
+    classifiers=[
+        'Programming Language :: Python :: 3.7', 'Topic :: Scientific/Engineering',
+        'Topic :: Scientific/Engineering :: Artificial Intelligence',
+        'Topic :: Scientific/Engineering :: Bio-Informatics',
+        'Topic :: Scientific/Engineering :: Mathematics', 'Operating System :: OS Independent',
+        'Development Status :: 4 - Beta', 'Intended Audience :: Science/Research',
+        'License :: OSI Approved :: BSD License'
+    ],
+    zip_safe=False)

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/conftest.py

@@ -1,0 +1,308 @@
+import logging
+import os
+import time
+
+import numpy as np
+import pytest
+import scipy.stats as ss
+
+import elfi
+import elfi.clients.dask as dask
+import elfi.clients.ipyparallel as eipp
+import elfi.clients.multiprocessing as mp
+import elfi.clients.native as native
+import elfi.examples.gauss
+import elfi.examples.ma2
+from elfi.methods.bo.gpy_regression import GPyRegression
+from elfi.methods.bo.acquisition import ExpIntVar, MaxVar, RandMaxVar
+from elfi.model.extensions import ModelPrior
+
+elfi.clients.native.set_as_default()
+
+
+# Add command line options
+def pytest_addoption(parser):
+    parser.addoption(
+        "--client",
+        action="store",
+        default="all",
+        help="perform the tests for the specified client (default all)")
+
+
+"""Functional fixtures"""
+
+
+@pytest.fixture(scope="session", params=[eipp, dask, mp, native])
+def client(request):
+    """Provides a fixture for all the different supported clients
+    """
+
+    client_module = request.param
+    client_name = client_module.__name__.split('.')[-1]
+    use_client = request.config.getoption('--client')
+
+    if use_client != 'all' and use_client != client_name:
+        pytest.skip("Skipping client {}".format(client_name))
+
+    try:
+        client = client_module.Client()
+    except BaseException:
+        pytest.skip("Client {} not available".format(client_name))
+
+    yield client
+
+    # Run cleanup code here if needed
+
+
+@pytest.fixture()
+def with_all_clients(client):
+    pre = elfi.get_client()
+    elfi.client.set_client(client)
+
+    yield
+
+    elfi.client.set_client(pre)
+
+
+@pytest.fixture()
+def use_logging():
+    logging.basicConfig(level=logging.DEBUG)
+    logging.getLogger('elfi.executor').setLevel(logging.WARNING)
+    logging.getLogger('elfi.compiler').setLevel(logging.WARNING)
+
+
+@pytest.fixture()
+def skip_travis():
+    if "TRAVIS" in os.environ and os.environ['TRAVIS'] == "true":
+        pytest.skip("Skipping this test in Travis CI due to very slow run-time. Tested "
+                    "locally!")
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_bolfire.py

@@ -1,0 +1,100 @@
+import pytest
+
+import numpy as np
+
+import elfi
+from elfi.methods.bo.acquisition import LCBSC
+from elfi.methods.classifier import LogisticRegression
+from elfi.model.extensions import ModelPrior
+
+
+def simple_gaussian_model(true_param, seed, n_summaries=10):
+    """The simple gaussian model that has been used as a toy example in the LFIRE paper."""
+
+    def power(x, y):
+        return x**y
+
+    m = elfi.ElfiModel()
+    mu = elfi.Prior('uniform', -5, 10, model=m, name='mu')
+    y = elfi.Simulator(gauss, *[mu], observed=gauss(true_param, seed=seed), name='y')
+    for i in range(n_summaries):
+        elfi.Summary(power, y, i, model=m, name=f'power_{i}')
+    return m
+
+
+def gauss(mu, sigma=3, n_obs=1, batch_size=1, seed=None, *args, **kwargs):
+    if isinstance(seed, int):
+        np.random.seed(seed)
+    mu = np.asanyarray(mu).reshape((-1, 1))
+    sigma = np.asanyarray(sigma).reshape((-1, 1))
+    return np.random.normal(mu, sigma, size=(batch_size, n_obs))
+
+
+@pytest.fixture
+def true_param():
+    return 2.6
+
+
+@pytest.fixture
+def seed():
+    return 4
+
+
+def test_bolfire_init(true_param, seed):
+    # define the simple gaussian elfi model
+    m = simple_gaussian_model(true_param, seed)
+
+    # define the bolfire method
+    bolfire_method = elfi.BOLFIRE(model=m, n_training_data=10)
+
+    # check the size of mariginal data (should be the size of training data x number of summaries)
+    assert bolfire_method.marginal.shape == (10, 10)
+    # check the feature names
+    assert bolfire_method.feature_names == [f'power_{i}' for i in range(10)]
+    # check the type of a default classifier
+    assert isinstance(bolfire_method.classifier, LogisticRegression)
+    # check the length of observed feature values
+    assert len(bolfire_method.observed[0]) == 10
+    # check the type of the prior
+    assert isinstance(bolfire_method.prior, ModelPrior)
+    # check the acquisition function and GP regression related parameters
+    assert bolfire_method.bounds is None
+    assert bolfire_method.acq_noise_var == 0
+    assert bolfire_method.exploration_rate == 10
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_compilation.py

@@ -1,0 +1,46 @@
+import ipyparallel
+import pytest
+
+import elfi
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_meta_param(ma2):
+    sim = ma2.get_reference('MA2')
+
+    # Test that it is passed
+    try:
+        # Add to state
+        sim.uses_meta = True
+        sim.generate()
+        assert False, "Should raise an error"
+    except TypeError:
+        assert True
+    except ipyparallel.error.RemoteError:
+        assert True
+
+
+# TODO: add ipyparallel, maybe use dill or cloudpickle
+# client.ipp_client[:].use_dill() or .use_cloudpickle()
+def test_batch_index_value(ma2):
+    def bi(meta):
+        return meta['batch_index']
+
+    # Test the correct batch_index value
+    op = elfi.Operation(bi, model=ma2, name='op')
+    op.uses_meta = True
+    client = elfi.get_client()
+    c = elfi.ComputationContext()
+    compiled_net = client.compile(ma2.source_net, ma2.nodes)
+    loaded_net = client.load_data(compiled_net, c, batch_index=3)
+    res = client.compute(loaded_net)
+
+    assert res['op'] == 3
+
+
+def test_reduce_compiler(ma2, client):
+    compiled_net = client.compile(ma2.source_net)
+    assert compiled_net.has_node('S1')
+
+    compiled_net2 = client.compile(ma2.source_net, ['MA2'])
+    assert not compiled_net2.has_node('S1')

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_consistency.py

@@ -1,0 +1,193 @@
+import logging
+
+import numpy as np
+import pytest
+
+import elfi
+
+"""This module tests the consistency of results when using the same seed."""
+
+
+def check_consistent_sample(sample, sample_diff, sample_same):
+    assert not np.array_equal(sample.outputs['t1'], sample_diff.outputs['t1'])
+
+    assert np.allclose(sample.outputs['t1'], sample_same.outputs['t1'])
+    assert np.allclose(sample.outputs['t2'], sample_same.outputs['t2'])
+
+    # BOLFI does not have d in its outputs
+    if 'd' in sample.outputs:
+        assert np.allclose(sample.outputs['d'], sample_same.outputs['d'])
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_rejection(ma2):
+    bs = 3
+    n_samples = 3
+    n_sim = 9
+
+    rej = elfi.Rejection(ma2, 'd', batch_size=bs)
+    sample = rej.sample(n_samples, n_sim=n_sim)
+    seed = rej.seed
+
+    rej = elfi.Rejection(ma2, 'd', batch_size=bs)
+    sample_diff = rej.sample(n_samples, n_sim=n_sim)
+
+    rej = elfi.Rejection(ma2, 'd', batch_size=bs, seed=seed)
+    sample_same = rej.sample(n_samples, n_sim=n_sim)
+
+    check_consistent_sample(sample, sample_diff, sample_same)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_smc(ma2):
+    bs = 3
+    n_samples = 10
+    thresholds = [1, .9, .8]
+
+    smc = elfi.SMC(ma2, 'd', batch_size=bs)
+    sample = smc.sample(n_samples, thresholds=thresholds)
+    seed = smc.seed
+
+    smc = elfi.SMC(ma2, 'd', batch_size=bs, seed=seed)
+    sample_same = smc.sample(n_samples, thresholds=thresholds)
+
+    smc = elfi.SMC(ma2, 'd', batch_size=bs)
+    sample_diff = smc.sample(n_samples, thresholds=thresholds)
+
+    check_consistent_sample(sample, sample_diff, sample_same)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_adaptive_distance_smc(ma2):
+    bs = 3
+    n_samples = 10
+    rounds = 3
+    quantile = 0.9
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_custom_outputs.py

@@ -1,0 +1,87 @@
+import numpy as np
+import pytest
+
+import elfi
+from elfi.utils import is_array
+
+
+def simulator(p, random_state=None):
+    n = 30
+    rs = random_state or np.random.RandomState()
+    data = rs.multinomial(n, p)
+
+    # Make it a dict for testing purposes
+    return dict(zip(range(n), data))
+
+
+def summary(dict_data):
+    n = len(dict_data)
+    data = np.array([dict_data[i] for i in range(n)])
+    return data / n
+
+
+def lsimulator(p, random_state=None):
+    n = 30
+    rs = random_state or np.random.RandomState()
+    data = rs.multinomial(n, p)
+
+    # Make it a list for testing purposes
+    return list(data) + ['test']
+
+
+def lsummary(list_data):
+    n = len(list_data)
+    data = np.array(list_data[:-1])
+    return data / (n - 1)
+
+
+def test_dict_output():
+    vsim = elfi.tools.vectorize(simulator)
+    vsum = elfi.tools.vectorize(summary)
+
+    obs = simulator([.2, .8])
+
+    elfi.new_model()
+    p = elfi.Prior('dirichlet', [2, 2])
+    sim = elfi.Simulator(vsim, p, observed=obs)
+    S = elfi.Summary(vsum, sim)
+    d = elfi.Distance('euclidean', S)
+
+    pool = elfi.OutputPool(['sim'])
+    rej = elfi.Rejection(d, batch_size=100, pool=pool, output_names=['sim'])
+    sample = rej.sample(100, n_sim=1000)
+    mean = np.mean(sample.samples['p'], axis=0)
+
+    # Crude test
+    assert mean[1] > mean[0]
+
+
+def test_list_output():
+    vsim = elfi.tools.vectorize(lsimulator)
+    vsum = elfi.tools.vectorize(lsummary)
+
+    v = vsim(np.array([[.2, .8], [.3, .7]]))
+    assert is_array(v)
+    assert not isinstance(v[0], list)
+
+    vsim = elfi.tools.vectorize(lsimulator, dtype=False)
+
+    v = vsim(np.array([[.2, .8], [.3, .7]]))
+    assert is_array(v)
+    assert isinstance(v[0], list)
+
+    obs = lsimulator([.2, .8])
+
+    elfi.new_model()
+    p = elfi.Prior('dirichlet', [2, 2])
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_inference.py

@@ -1,0 +1,337 @@
+"""Test inference methods point estimates with informative data from the MA2 process."""
+
+from collections import OrderedDict
+
+import numpy as np
+import pytest
+
+import elfi
+from elfi.examples import ma2
+from elfi.methods.bo.utils import minimize, stochastic_optimization
+from elfi.methods.bsl.pre_sample_methods import estimate_whitening_matrix, select_penalty
+from elfi.methods.bsl.pdf_methods import standard_likelihood, unbiased_likelihood, semiparametric_likelihood, robust_likelihood
+from elfi.model.elfi_model import NodeReference
+
+
+def setup_ma2_with_informative_data(n_obs=100):
+    true_params = OrderedDict([('t1', .6), ('t2', .2)])
+
+    # In our implementation, seed 4 gives informative (enough) synthetic observed
+    # data of length 100 for quite accurate inference of the true parameters using
+    # posterior mean as the point estimate
+    m = ma2.get_model(n_obs=n_obs, true_params=true_params.values(), seed_obs=4)
+    return m, true_params
+
+
+def check_inference_with_informative_data(outputs, N, true_params, error_bound=0.05):
+    t1 = outputs['t1']
+    t2 = outputs['t2']
+
+    if N > 1:
+        assert len(t1) == N
+
+    assert np.abs(np.mean(t1) - true_params['t1']) < error_bound, \
+        "\n\nNot |{} - {}| < {}\n".format(np.mean(t1), true_params['t1'], error_bound)
+    assert np.abs(np.mean(t2) - true_params['t2']) < error_bound, \
+        "\n\nNot |{} - {}| < {}\n".format(np.mean(t2), true_params['t2'], error_bound)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_rejection_with_quantile():
+    m, true_params = setup_ma2_with_informative_data()
+
+    quantile = 0.01
+    N = 1000
+    batch_size = 20000
+    rej = elfi.Rejection(m['d'], batch_size=batch_size)
+    res = rej.sample(N, quantile=quantile)
+
+    check_inference_with_informative_data(res.samples, N, true_params)
+
+    # Check that there are no repeating values indicating a seeding problem
+    assert len(np.unique(res.discrepancies)) == N
+
+    assert res.accept_rate == quantile
+    assert res.n_sim == int(N / quantile)
+
+
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_post_processing.py

@@ -1,0 +1,126 @@
+from functools import partial
+
+import numpy as np
+import pytest
+
+import elfi
+import elfi.methods.post_processing as pp
+from elfi.examples import gauss, ma2
+from elfi.methods.post_processing import LinearAdjustment, adjust_posterior
+
+
+def _statistics(arr):
+    return arr.mean(), arr.var()
+
+
+def test_get_adjustment():
+    with pytest.raises(ValueError):
+        pp._get_adjustment('doesnotexist')
+
+
+def test_single_parameter_linear_adjustment():
+    """A regression test against values obtained in the notebook."""
+    seed = 20170616
+    n_obs = 50
+    batch_size = 1000
+    mu, sigma = (5, 1)
+
+    # Hyperparameters
+    mu0, sigma0 = (10, 100)
+
+    y_obs = gauss.gauss(
+        mu, sigma, n_obs=n_obs, batch_size=1, random_state=np.random.RandomState(seed))
+    sim_fn = partial(gauss.gauss, sigma=sigma, n_obs=n_obs)
+
+    # Posterior
+    n = y_obs.shape[1]
+    mu1 = (mu0 / sigma0**2 + y_obs.sum() / sigma**2) / (1 / sigma0**2 + n / sigma**2)
+    sigma1 = (1 / sigma0**2 + n / sigma**2)**(-0.5)
+
+    # Model
+    m = elfi.ElfiModel()
+    elfi.Prior('norm', mu0, sigma0, model=m, name='mu')
+    elfi.Simulator(sim_fn, m['mu'], observed=y_obs, name='gauss')
+    elfi.Summary(lambda x: x.mean(axis=1), m['gauss'], name='ss_mean')
+    elfi.Distance('euclidean', m['ss_mean'], name='d')
+
+    res = elfi.Rejection(m['d'], output_names=['ss_mean'], batch_size=batch_size,
+                         seed=seed).sample(1000, threshold=1)
+    adj = elfi.adjust_posterior(model=m, sample=res, parameter_names=['mu'], summary_names=['ss_mean'])
+
+    assert np.allclose(_statistics(adj.outputs['mu']), (4.9772879640569778, 0.02058680115402544))
+
+
+# TODO: Use a fixture for the model
+def test_nonfinite_values():
+    """A regression test against values obtained in the notebook."""
+    seed = 20170616
+    n_obs = 50
+    batch_size = 1000
+    mu, sigma = (5, 1)
+
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_randomness.py

@@ -1,0 +1,76 @@
+import numpy as np
+import pytest
+import scipy.stats as ss
+
+import elfi
+from elfi.utils import get_sub_seed
+
+
+def test_randomness(simple_model):
+    k1 = simple_model['k1']
+
+    gen1 = k1.generate(10)
+    gen2 = k1.generate(10)
+
+    assert not np.array_equal(gen1, gen2)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_randomness2(simple_model):
+    k1 = simple_model['k1']
+
+    n = 30
+    samples1 = elfi.Rejection(simple_model['k1'], batch_size=3).sample(n).samples['k1']
+    assert len(np.unique(samples1)) == n
+
+    samples2 = elfi.Rejection(simple_model['k1'], batch_size=3).sample(n).samples['k1']
+    assert not np.array_equal(samples1, samples2)
+
+
+# If we want to test this with all clients, we need to to set the worker's random state
+def test_global_random_state_usage(simple_model):
+    n_gen = 10
+
+    np.random.seed(0)
+    k2 = simple_model['k2']
+    k2.generate(n_gen)
+    st1 = np.random.get_state()
+
+    np.random.seed(0)
+    mu = ss.uniform.rvs(0, 10, size=(n_gen, 1))
+    ss.norm.rvs(mu, size=(n_gen, 3))
+    st2 = np.random.get_state()
+
+    assert random_state_equal(st1, st2)
+
+
+def test_get_sub_seed():
+    n = 100
+    seed = np.random.randint(2**31)
+    sub_seeds = []
+    for i in range(n):
+        sub_seeds.append(get_sub_seed(seed, i, n))
+
+    assert len(np.unique(sub_seeds)) == n
+
+    # Test the cached version
+    cache = {}
+    sub_seeds_cached = []
+    for i in range(n):
+        sub_seed = get_sub_seed(seed, i, n, cache=cache)
+        sub_seeds_cached.append(sub_seed)
+
+    assert np.array_equal(sub_seeds, sub_seeds_cached)
+
+
+# Helpers
+
+
+def random_state_equal(st1, st2):
+    # 1. the string 'MT19937'.
+    tf = st1[0] == st2[0]
+    # 2. a 1-D array of 624 unsigned integer keys.
+    tf = tf and np.array_equal(st1[1], st2[1])
+    # 3. an integer ``pos``.
+    tf = tf and st1[2] == st2[2]
+    return tf

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_serialization.py

@@ -1,0 +1,43 @@
+import pickle
+
+import numpy as np
+
+from elfi.client import ClientBase
+from elfi.examples import ma2
+from elfi.executor import Executor
+from elfi.model.elfi_model import ComputationContext
+
+
+def test_pickle_ma2():
+    m = ma2.get_model()
+    d = m.get_reference('d')
+
+    np.random.seed(0)
+    res1 = d.generate(10)
+
+    serialized = pickle.dumps(m)
+    m = pickle.loads(serialized)
+    d = m.get_reference('d')
+
+    np.random.seed(0)
+    res2 = d.generate(10)
+
+    assert np.array_equal(res1, res2)
+
+
+def test_pickle_ma2_compiled_and_loaded(ma2):
+    compiled = ClientBase.compile(ma2.source_net, ['d'])
+    loaded = ClientBase.load_data(compiled, ComputationContext(), 0)
+
+    np.random.seed(0)
+    result = Executor.execute(loaded)
+    res1 = result['d']
+
+    serialized = pickle.dumps(loaded)
+    loaded = pickle.loads(serialized)
+
+    np.random.seed(0)
+    result = Executor.execute(loaded)
+    res2 = result['d']
+
+    assert np.array_equal(res1, res2)

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/functional/test_simulation_reuse.py

@@ -1,0 +1,94 @@
+import os
+import time
+
+import numpy as np
+import pytest
+
+import elfi
+
+
+@pytest.mark.parametrize('sleep_model', [.2], indirect=['sleep_model'])
+def test_pool_usage(sleep_model):
+    # Add nodes to the pool
+    pool = elfi.OutputPool(outputs=sleep_model.parameter_names + ['slept', 'summary', 'd'])
+
+    rej = elfi.Rejection(sleep_model['d'], batch_size=5, pool=pool)
+    quantile = .25
+    ts = time.time()
+    res = rej.sample(5, quantile=quantile)
+    td = time.time() - ts
+    # Will make 5/.25 = 20 evaluations with mean time of .1 secs, so 2 secs total on
+    # average. Allow some slack although still on rare occasions this may fail.
+    assert td > 1.2
+
+    # Instantiating new inference with the same pool should be faster because we
+    # use the prepopulated pool
+    rej = elfi.Rejection(sleep_model['d'], batch_size=5, pool=pool)
+    ts = time.time()
+    res = rej.sample(5, quantile=quantile)
+    td = time.time() - ts
+    assert td < 1.2
+
+    # It should work if we remove the simulation, since the Rejection sampling
+    # only requires the parameters and the discrepancy
+    pool.remove_store('slept')
+    rej = elfi.Rejection(sleep_model['d'], batch_size=5, pool=pool)
+    ts = time.time()
+    res = rej.sample(5, quantile=quantile)
+    td = time.time() - ts
+    assert td < 1.2
+
+    # It should work even if we remove the discrepancy, since the discrepancy can be recomputed
+    # from the stored summary
+    pool.remove_store('d')
+    rej = elfi.Rejection(sleep_model['d'], batch_size=5, pool=pool)
+    ts = time.time()
+    res = rej.sample(5, quantile=quantile)
+    td = time.time() - ts
+    assert td < 1.2
+
+
+def test_pool_restarts(ma2):
+    pool = elfi.ArrayPool(['t1', 'd'], name='test')
+    rej = elfi.Rejection(ma2, 'd', batch_size=10, pool=pool, seed=123)
+
+    rej.sample(1, n_sim=30)
+    pool.save()
+
+    # Do not save the pool...
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_bo.py

@@ -1,0 +1,237 @@
+import numpy as np
+import pytest
+
+import elfi
+import elfi.methods.bo.acquisition as acquisition
+from elfi.methods.bo.gpy_regression import GPyRegression
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_BO(ma2):
+    # Log transform of the distance usually smooths the distance surface
+    log_d = elfi.Operation(np.log, ma2['d'], name='log_d')
+
+    n_init = 20
+    res_init = elfi.Rejection(log_d, batch_size=5).sample(n_init, quantile=1)
+
+    bounds = {n: (-2, 2) for n in ma2.parameter_names}
+    bo = elfi.BayesianOptimization(
+        log_d, initial_evidence=res_init.outputs, update_interval=10, batch_size=5, bounds=bounds)
+    assert bo.target_model.n_evidence == n_init
+    assert bo.n_evidence == n_init
+    assert bo.n_precomputed_evidence == n_init
+    assert bo.n_initial_evidence == n_init
+
+    n1 = 5
+    bo.infer(n_init + n1)
+
+    assert bo.target_model.n_evidence == n_init + n1
+    assert bo.n_evidence == n_init + n1
+    assert bo.n_precomputed_evidence == n_init
+    assert bo.n_initial_evidence == n_init
+
+    n2 = 5
+    bo.infer(n_init + n1 + n2)
+
+    assert bo.target_model.n_evidence == n_init + n1 + n2
+    assert bo.n_evidence == n_init + n1 + n2
+    assert bo.n_precomputed_evidence == n_init
+    assert bo.n_initial_evidence == n_init
+
+    assert np.array_equal(bo.target_model._gp.X[:n_init, 0], res_init.samples_array[:, 0])
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_async(ma2):
+    bounds = {n: (-2, 2) for n in ma2.parameter_names}
+    bo = elfi.BayesianOptimization(
+        ma2, 'd', initial_evidence=0, update_interval=2, batch_size=2, bounds=bounds, async_acq=True)
+    n_samples = 5
+    bo.infer(n_samples)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_BO_works_with_zero_init_samples(ma2):
+    log_d = elfi.Operation(np.log, ma2['d'], name='log_d')
+    bounds = {n: (-2, 2) for n in ma2.parameter_names}
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_client.py

@@ -1,0 +1,47 @@
+import numpy as np
+import pytest
+import scipy.stats as ss
+
+import elfi
+import elfi.client
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_batch_handler(simple_model):
+    m = simple_model
+    computation_context = elfi.ComputationContext(seed=123, batch_size=10)
+    batches = elfi.client.BatchHandler(m, computation_context, 'k2')
+
+    batches.submit()
+    out0, i0 = batches.wait_next()
+
+    batches.submit()
+    out1, i1 = batches.wait_next()
+
+    batches.reset()
+    batches.submit()
+    out0_, i0_ = batches.wait_next()
+
+    assert i0 == 0
+    assert i1 == 1
+    assert i0_ == 0
+    assert np.array_equal(out0['k2'], out0_['k2'])
+    assert not np.array_equal(out0['k2'], out1['k2'])
+
+
+def test_multiprocessing_kwargs(simple_model):
+    pre = elfi.get_client()
+
+    m = simple_model
+    num_proc = 2
+    elfi.set_client('multiprocessing', num_processes=num_proc)
+    rej = elfi.Rejection(m['k1'])
+    assert rej.client.num_cores == num_proc
+
+    elfi.set_client('multiprocessing', processes=num_proc)
+    rej = elfi.Rejection(m['k1'])
+    assert rej.client.num_cores == num_proc
+
+    elfi.set_client(pre)
+
+# TODO: add testing that client is cleared from tasks after they are retrieved

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_document_examples.py

@@ -1,0 +1,77 @@
+"""
+This file contains tests for some of the examples used in the documentation that
+are not automatically produced from the notebooks.
+
+Note that if you change anything in this file (e.g. imports), you should change
+the documentation accordingly. For this reason the imports and class definitions
+are inside their respective functions.
+"""
+
+
+def test_implementing_new_algorithm():
+    import numpy as np
+
+    from elfi.methods.inference.parameter_inference import ParameterInference
+    from elfi.methods.results import Sample
+
+    import elfi.examples.ma2 as ma2
+
+    class CustomMethod(ParameterInference):
+        def __init__(self, model, discrepancy_name, threshold, **kwargs):
+            # Create a name list of nodes whose outputs we wish to receive
+            output_names = [discrepancy_name] + model.parameter_names
+            super(CustomMethod, self).__init__(model, output_names, **kwargs)
+
+            self.threshold = threshold
+            self.discrepancy_name = discrepancy_name
+
+            # Prepare lists to push the filtered outputs into
+            self.state['filtered_outputs'] = {name: [] for name in output_names}
+
+        def set_objective(self, n_sim):
+            self.objective['n_sim'] = n_sim
+
+        def update(self, batch, batch_index):
+            super(CustomMethod, self).update(batch, batch_index)
+
+            # Make a filter mask (logical numpy array) from the distance array
+            filter_mask = batch[self.discrepancy_name] <= self.threshold
+
+            # Append the filtered parameters to their lists
+            for name in self.output_names:
+                values = batch[name]
+                self.state['filtered_outputs'][name].append(values[filter_mask])
+
+        def extract_result(self):
+            filtered_outputs = self.state['filtered_outputs']
+            outputs = {name: np.concatenate(filtered_outputs[name]) for name in self.output_names}
+
+            return Sample(
+                method_name='CustomMethod',
+                outputs=outputs,
+                parameter_names=self.parameter_names,
+                discrepancy_name=self.discrepancy_name,
+                n_sim=self.state['n_sim'],
+                threshold=self.threshold)
+
+    # Below is from the part where we demonstrate iterative advancing
+
+    # Run it
+    m = ma2.get_model()
+    custom_method = CustomMethod(m, 'd', threshold=.1, batch_size=1000)
+
+    # Continue inference from the previous state (with n_sim=2000)
+    custom_method.infer(n_sim=4000)
+
+    # Or use it iteratively
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_elfi_model.py

@@ -1,0 +1,253 @@
+import os
+
+import numpy as np
+import pytest
+
+import elfi
+import elfi.model.elfi_model as em
+from elfi.examples import ma2 as ema2
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_generate(ma2):
+    n_gen = 10
+
+    d = ma2.get_reference('d')
+    res = d.generate(n_gen)
+
+    assert res.shape[0] == n_gen
+    assert res.ndim == 1
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_generate_outputs(ma2):
+    n_gen = 10
+
+    res = ma2.generate(n_gen)
+
+    assert 'd' in res
+    assert res['d'].shape[0] == n_gen
+    assert res['d'].ndim == 1
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_observed():
+    true_params = [.6, .2]
+    m = ema2.get_model(100, true_params=true_params)
+    y = m.observed['MA2']
+    S1 = m.get_reference('S1')
+    S2 = m.get_reference('S2')
+
+    S1_observed = ema2.autocov(y)
+    S2_observed = ema2.autocov(y, 2)
+
+    assert np.array_equal(S1.observed, S1_observed)
+    assert np.array_equal(S2.observed, S2_observed)
+
+
+def euclidean_discrepancy(*simulated, observed):
+    """Euclidean discrepancy between data.
+
+    Parameters
+    ----------
+    *simulated
+        simulated summaries
+    observed : tuple of 1d or 2d np.arrays of length n
+
+    Returns
+    -------
+    d : np.array of size (n,)
+    """
+    d = np.linalg.norm(np.column_stack(simulated) - np.column_stack(observed), ord=2, axis=1)
+    return d
+
+
+class TestElfiModel:
+    def test_remove_node(self, ma2):
+        ma2.remove_node('MA2')
+
+        assert not ma2.has_node('MA2')
+        assert ma2.has_node('t2')
+
+        parents = ma2.get_parents('t2')
+        # This needs to have at least 2 parents so that the test below makes sense
+        assert len(parents) > 1
+
+        ma2.remove_node('t2')
+        for p in parents:
+            if p[0] == '_':
+                assert not ma2.has_node(p)
+            else:
+                assert ma2.has_node(p)
+
+        assert 'MA2' not in ma2.observed
+
+    def test_save_load(self, ma2):
+        name = ma2.name
+        ma2.save()
+        ma2 = elfi.load_model(name)
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_examples.py

@@ -1,0 +1,142 @@
+"""Simple running tests for examples."""
+
+import os
+
+import pytest
+
+import elfi
+from elfi.examples import arch, bdm, bignk, gauss, gnk, lotka_volterra, \
+                          ricker, daycare, lorenz, ar1, mg1, toad, \
+                          scratch_assay, stochastic_volatility_model
+
+
+def test_bdm():
+    """Currently only works in unix-like systems and with a cloned repository."""
+    cpp_path = bdm.get_sources_path()
+
+    do_cleanup = False
+    if not os.path.isfile(cpp_path + '/bdm'):
+        os.system('make -C {}'.format(cpp_path))
+        do_cleanup = True
+
+    assert os.path.isfile(cpp_path + '/bdm')
+
+    # Remove the executable if it already exists
+    if os.path.isfile('bdm'):
+        os.system('rm bdm')
+
+    with pytest.warns(RuntimeWarning):
+        m = bdm.get_model()
+
+    # Copy the file here to run the test
+    os.system('cp {}/bdm .'.format(cpp_path))
+
+    # Should no longer warn
+    m = bdm.get_model()
+
+    # Test that you can run the inference
+
+    rej = elfi.Rejection(m, 'd', batch_size=100)
+    rej.sample(20)
+
+    # TODO: test the correctness of the result
+
+    os.system('rm ./bdm')
+    if do_cleanup:
+        os.system('rm {}/bdm'.format(cpp_path))
+
+
+def test_gauss():
+    m = gauss.get_model()
+    rej = elfi.Rejection(m, m['d'], batch_size=10)
+    rej.sample(20)
+
+
+def test_gauss_1d_mean():
+    params_true = [4]
+    cov_matrix = [1]
+
+    m = gauss.get_model(true_params=params_true, nd_mean=True, cov_matrix=cov_matrix)
+    rej = elfi.Rejection(m, m['d'], batch_size=10)
+    rej.sample(20)
+
+
+def test_gauss_2d_mean():
+    params_true = [4, 4]
+    cov_matrix = [[1, .5], [.5, 1]]
+
+    m = gauss.get_model(true_params=params_true, nd_mean=True, cov_matrix=cov_matrix)
+    rej = elfi.Rejection(m, m['d'], batch_size=10)
+    rej.sample(20)
+
+
+def test_Ricker():
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_mcmc.py

@@ -1,0 +1,68 @@
+import numpy as np
+import pytest
+
+from elfi.methods import mcmc
+
+# construct a covariance matrix and calculate the precision matrix
+n = 5
+true_cov = np.random.rand(n, n) * 0.5
+true_cov += true_cov.T
+true_cov += n * np.eye(n)
+prec = np.linalg.inv(true_cov)
+
+
+# log multivariate Gaussian pdf
+def log_pdf(x):
+    return -0.5 * x.dot(prec).dot(x)
+
+
+# gradient of log multivariate Gaussian pdf
+def grad_log_pdf(x):
+    return -x.dot(prec)
+
+
+class TestMetropolis():
+    def test_metropolis(self):
+        n_samples = 200000
+        x_init = np.random.rand(n)
+        sigma = np.ones(n)
+        samples = mcmc.metropolis(n_samples, x_init, log_pdf, sigma)
+        assert samples.shape == (n_samples, n)
+        cov = np.cov(samples[100000:, :].T)
+        assert np.allclose(cov, true_cov, atol=0.3, rtol=0.1)
+
+
+@pytest.mark.slowtest
+class TestNUTS():
+    def test_nuts(self):
+        n_samples = 100000
+        n_adapt = 10000
+        x_init = np.random.rand(n)
+        samples = mcmc.nuts(n_samples, x_init, log_pdf, grad_log_pdf, n_adapt=n_adapt)
+        assert samples.shape == (n_samples, n)
+        cov = np.cov(samples[n_adapt:, :].T)
+        assert np.allclose(cov, true_cov, atol=0.1, rtol=0.1)
+
+
+# some data generated in PyStan
+chains_Stan = np.array([[0.2955857, 1.27937191, 1.05884099, 0.91236858], [
+    0.38128885, 1.34242613, 0.49102573, 0.76061715
+], [0.38128885, 1.18404563, 0.49102573,
+    0.78910512], [0.38128885, 0.72150199, 0.49102573,
+                  1.13845618], [0.38128885, 0.72150199, 0.38102685,
+                                0.81298041], [0.26917982, 0.72150199, 0.38102685, 0.81298041],
+                        [0.26917982, 0.68149163, 0.45830605,
+                         0.86364605], [0.51213898, 0.68149163, 0.29170172, 0.80734373],
+                        [0.51213898, 0.85560228, 0.29170172,
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_methods.py

@@ -1,0 +1,122 @@
+import numpy as np
+import pytest
+
+import elfi
+import elfi.examples.ma2 as exma2
+from elfi.methods.inference.parameter_inference import ParameterInference
+
+
+def test_no_model_parameters(simple_model):
+    simple_model.parameter_names = []
+
+    with pytest.raises(Exception):
+        ParameterInference(simple_model, [])
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_smc(ma2):
+    thresholds = [.5, .2]
+    N = 1000
+    smc = elfi.SMC(ma2['d'], batch_size=20000)
+    res = smc.sample(N, thresholds=thresholds)
+    dens = smc._prior.logpdf(res.samples_array)
+    # Test that the density is uniform
+    assert np.allclose(dens, dens[0])
+
+    # Test that weighted mean is computed
+    w = res.weights
+    assert w is not None
+
+    samples = res.samples_array
+    means = res.sample_means_array
+    assert np.allclose(means, np.average(samples, 0, w))
+    assert not np.allclose(means, np.mean(samples, 0))
+
+    # Test result summaries
+    res.summary()
+    res.summary(all=True)
+
+    res.sample_means_summary()
+    res.sample_means_summary(all=True)
+
+    # Ensure prior pdf > 0 for samples
+    assert np.all(exma2.CustomPrior1.pdf(samples[:, 0], 2) > 0)
+    assert np.all(exma2.CustomPrior2.pdf(samples[:, 1], samples[:, 0], 1) > 0)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_threshold_evolution_in_smc(ma2):
+    threshold_selection_quantiles = [0.5] * 5
+    N = 100
+    smc = elfi.SMC(ma2['d'], batch_size=100)
+    res = smc.sample(N, quantiles=threshold_selection_quantiles)
+
+    # Check that tolerance threshold decreased between iterations
+    thresholds = smc.objective['thresholds'][1:]
+    assert np.all(np.diff(thresholds)<0)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_threshold_evolution_in_adaptive_threshold_smc(ma2):
+    N = 100
+    adathsmc = elfi.AdaptiveThresholdSMC(ma2['d'], batch_size=100)
+    res = adathsmc.sample(N, max_iter=5)
+
+    # Check that tolerance threshold decreased between iterations
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_results.py

@@ -1,0 +1,87 @@
+import numpy as np
+import pytest
+
+import elfi
+
+
+def test_sample():
+    n_samples = 10
+    parameter_names = ['a', 'b']
+    distance_name = 'dist'
+    samples = [
+        np.random.random(n_samples),
+        np.random.random(n_samples),
+        np.random.random(n_samples)
+    ]
+    outputs = dict(zip(parameter_names + [distance_name], samples))
+    sample = elfi.methods.results.Sample(
+        method_name="TestRes",
+        outputs=outputs,
+        parameter_names=parameter_names,
+        discrepancy_name=distance_name,
+        something='x',
+        something_else='y',
+        n_sim=0, )
+
+    assert sample.method_name == "TestRes"
+    assert hasattr(sample, 'samples')
+    assert sample.n_samples == n_samples
+    assert sample.dim == len(parameter_names)
+    assert not sample.is_multivariate
+
+    assert np.allclose(samples[0], sample.samples_array[:, 0])
+    assert np.allclose(samples[1], sample.samples_array[:, 1])
+    assert np.allclose(samples[-1], sample.discrepancies)
+
+    assert hasattr(sample, 'something')
+    assert sample.something_else == 'y'
+
+    with pytest.raises(AttributeError):
+        sample.not_here
+
+    # Test summary
+    sample.summary()
+
+
+def test_bolfi_sample():
+    n_chains = 3
+    n_iters = 10
+    warmup = 5
+    parameter_names = ['a', 'b']
+    chains = np.random.random((n_chains, n_iters, len(parameter_names)))
+
+    result = elfi.methods.results.BolfiSample(
+        method_name="TestRes",
+        chains=chains,
+        parameter_names=parameter_names,
+        warmup=warmup,
+        something='x',
+        something_else='y',
+        n_sim=0, )
+
+    assert result.method_name == "TestRes"
+    assert hasattr(result, 'samples')
+    assert hasattr(result, 'chains')
+    assert hasattr(result, 'outputs')
+    assert result.n_samples == n_chains * (n_iters - warmup)
+    assert result.dim == len(parameter_names)
+    assert not result.is_multivariate
+
+    # verify that chains are merged correctly
+    s0 = np.concatenate([chains[i, warmup:, 0] for i in range(n_chains)])
+    s1 = np.concatenate([chains[i, warmup:, 1] for i in range(n_chains)])
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_store.py

@@ -1,0 +1,231 @@
+import os
+import pickle
+
+import numpy as np
+import pytest
+
+import elfi
+from elfi.store import ArrayPool, ArrayStore, NpyArray, NpyStore, OutputPool
+
+
+def test_npy_array():
+    filename = 'test.npy'
+
+    original = np.random.rand(3, 2)
+    append = np.random.rand(10, 2)
+    ones = np.ones((10, 2))
+    append2 = np.random.rand(23, 2)
+
+    arr = NpyArray(filename, truncate=True)
+    arr.append(original)
+    assert np.array_equal(original, arr[:])
+    arr.close()
+    loaded = np.load(filename)
+    assert np.array_equal(original, loaded)
+
+    # Test appending and reading
+    arr = NpyArray(filename)
+    arr.append(append)
+    arr.flush()
+    loaded = np.load(filename)
+    assert np.array_equal(np.r_[original, append], loaded)
+
+    # Test further appending
+    arr.append(append2)
+    assert np.array_equal(np.r_[original, append, append2], arr[:])
+    arr.flush()
+    loaded = np.load(filename)
+    assert np.array_equal(np.r_[original, append, append2], loaded)
+
+    # Test that writing over the array fails
+    with pytest.raises(Exception):
+        arr[len(loaded):len(loaded) + 10, :] = ones
+
+    # Test rewriting
+    arr[3:13, :] = ones
+    arr.close()
+    loaded = np.load(filename)
+    assert np.array_equal(np.r_[original, ones, append2], loaded)
+
+    # Test pickling
+    arr = NpyArray(filename)
+    serialized = pickle.dumps(arr)
+    arr = pickle.loads(serialized)
+    assert np.array_equal(np.r_[original, ones, append2], arr[:])
+
+    # Test truncate method
+    arr = NpyArray(filename)
+    arr.truncate(len(original))
+    assert np.array_equal(original, arr[:])
+    arr.close()
+    loaded = np.load(filename)
+    assert np.array_equal(original, loaded)
+
+    # Try that truncation in initialization works
+    arr = NpyArray(filename, truncate=True)
+    arr.append(append)
+    arr.close()
+    loaded = np.load(filename)
+    assert np.array_equal(append, loaded)
+
+    os.remove(filename)
+
+
+def test_npy_array_multiple_instances():
+    original = np.random.rand(3, 2)
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_tools.py

@@ -1,0 +1,113 @@
+import pickle
+
+import numpy as np
+import pytest
+
+import elfi
+
+from elfi.examples import ma2
+
+def test_vectorize_decorator():
+    batch_size = 3
+    a = np.array([1, 2, 3])
+    b = np.array([3, 2, 1])
+
+    def simulator(a, b, random_state=None):
+        return a * b
+
+    vsim = elfi.tools.vectorize(simulator)
+    assert np.array_equal(a * b, vsim(a, b, batch_size=batch_size))
+
+    def simulator(a, constant, random_state=None):
+        return a * constant
+
+    vsim = elfi.tools.vectorize(simulator, constants=[1])
+    assert np.array_equal(a * 5, vsim(a, 5, batch_size=batch_size))
+
+    vsim = elfi.tools.vectorize(simulator, [1])
+    assert np.array_equal(a * 5, vsim(a, 5, batch_size=batch_size))
+
+    def simulator(constant0, b, constant2, random_state=None):
+        return constant0 * b * constant2
+
+    vsim = elfi.tools.vectorize(simulator, constants=(0, 2))
+    assert np.array_equal(2 * b * 7, vsim(2, b, 7, batch_size=batch_size))
+
+    # Invalid batch size in b
+    with pytest.raises(ValueError):
+        vsim(2, b, 7, batch_size=2 * batch_size)
+
+
+def simulator():
+    pass
+
+
+def test_vectorized_pickling():
+    sim = elfi.tools.vectorize(simulator)
+    pickle.dumps(sim)
+
+
+def test_external_operation():
+    # Note that the test string has intentionally not uniform formatting with spaces
+    elfi.new_model()
+    op = elfi.tools.external_operation('echo 1 {0} 4  5    6 {seed}')
+    constant = elfi.Constant(123)
+    simulator = elfi.Simulator(op, constant)
+    v = simulator.generate(1)
+    assert np.array_equal(v[:5], [1, 123, 4, 5, 6])
+
+    # Can be pickled
+    pickle.dumps(op)
+
+
+@pytest.mark.usefixtures('with_all_clients')
+def test_vectorized_and_external_combined():
+    constant = elfi.Constant(123)
+    kwargs_sim = elfi.tools.external_operation(
+        'echo {seed} {batch_index} {index_in_batch} {submission_index}', process_result='int32')
[…]

### 2023-09-27 jss-camera-ready/Code/elfi-dev/tests/unit/test_utils.py

@@ -1,0 +1,302 @@
+import json
+from collections import OrderedDict
+
+import numpy as np
+import scipy.stats as ss
+
+import elfi
+from elfi.examples.ma2 import get_model
+from elfi.methods.bo.utils import CostFunction, minimize, stochastic_optimization
+from elfi.methods.density_ratio_estimation import DensityRatioEstimation
+from elfi.methods.utils import (GMDistribution, normalize_weights, numgrad, numpy_to_python_type,
+                                sample_object_to_dict, weighted_sample_quantile, weighted_var)
+from elfi.model.extensions import ModelPrior
+
+
+def test_stochastic_optimization():
+    def fun(x):
+        return x**2
+
+    bounds = ((-1, 1), )
+    its = int(1e3)
+    polish = True
+    loc, val = stochastic_optimization(fun, bounds, its, polish)
+    assert abs(loc - 0.0) < 1e-5
+    assert abs(val - 0.0) < 1e-5
+
+
+def test_minimize_with_known_gradient():
+    def fun(x):
+        return x[0]**2 + (x[1] - 1)**4
+
+    def grad(x):
+        return np.array([2 * x[0], 4 * (x[1] - 1)**3])
+
+    bounds = ((-2, 2), (-2, 3))
+    loc, val = minimize(fun, bounds, grad=grad)
+    assert np.isclose(val, 0, atol=0.01)
+    assert np.allclose(loc, np.array([0, 1]), atol=0.02)
+
+
+def test_minimize_with_approx_gradient():
+    def fun(x):
+        return x[0]**2 + (x[1] - 1)**4
+
+    bounds = ((-2, 2), (-2, 3))
+    loc, val = minimize(fun, bounds)
+    assert np.isclose(val, 0, atol=0.01)
+    assert np.allclose(loc, np.array([0, 1]), atol=0.02)
+
+
+def test_minimize_with_constraints():
+    def fun(x):
+        return x[0]**2 + (x[1] - 1)**4
+
+    bounds = ((-2, 2), (-2, 3))
+    # Test constraint y >= x
+    constraints = ({'type': 'ineq',
+                    'fun': lambda x : x[1] - x[0]})
+    loc, val = minimize(fun, bounds, constraints=constraints, method='SLSQP')
+    assert np.isclose(val, 0, atol=0.01)
+    assert np.allclose(loc, np.array([0, 1]), atol=0.02)
+
+
+def test_weighted_sample_quantile():
[…]

### 2023-09-27 jss-camera-ready/Code/replication_script.py

@@ -1,0 +1,852 @@
+from elfi.methods.inference.romc import OptimisationProblem
+from sklearn.pipeline import Pipeline
+from sklearn.neural_network import MLPRegressor
+from functools import partial
+
+import os
+import timeit
+import time
+import scipy.stats as ss
+import scipy.integrate as integrate
+import logging
+import matplotlib.pyplot as plt
+from elfi.examples import ma2
+import numpy as np
+import elfi
+import matplotlib
+import random
+
+random.seed(21)
+
+if __name__ == '__main__':
+
+    # matplotlib.rcParams['text.usetex'] = True
+    logging.basicConfig(level=logging.CRITICAL)
+    savefig = False
+
+    ######################################################################
+    # Part 1. Generates the subfigures used in the running example in the#
+    # presentation of our implementation i.e. figure 3                   #
+    ######################################################################
+
+    class Prior:
+        def rvs(self, size=None, random_state=None):
+            # size from (BS,) -> (BS,1)
+            if size is not None:
+                size = np.concatenate((size, [1]))
+            return ss.uniform(loc=-2.5, scale=5).rvs(size=size, random_state=random_state)
+
+        def pdf(self, theta):
+            return ss.uniform(loc=-2.5, scale=5).pdf(theta)
+
+        def logpdf(self, theta):
+            return ss.uniform(loc=-2.5, scale=5).logpdf(theta)
+
+
+    class Likelihood:
+        r"""Implements the distribution
+        P(x|theta) = N(x; theta^4, 1)         if theta in [-0.5, 0.5]
+                     N(x; theta + 0.5 + 0.5^4 if theta > 0.5
+                     N(x; theta - 0.5 + 0.5^4 if theta < 0.5
+        """
+
+        def rvs(self, theta, seed=None):
+            """Vectorised sampling from likelihood.
+
+            """
+            assert isinstance(theta, np.ndarray)
+            theta = theta.astype(float)
+            samples = np.empty_like(theta)
+
+            c = 0.5 - 0.5 ** 4
+
+            tmp_theta = theta[theta <= -0.5]
+            samples[theta <= -
+                    0.5] = ss.norm(loc=-tmp_theta - c, scale=1).rvs(random_state=seed)
+            theta[theta <= -0.5] = np.inf
+
[…]
